### 2015-01-22 fitting.Rmd
First attempt at fitting model to data. #13

# Fitting TREE to observed data

```{r, echo=FALSE, results="hide"}
sourceDir <- function(path, ...) {
     for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
        source(file.path(path, nm), ...)
     }
}
sourceDir("R")
```

## Data

Used estimate potential growth rate and species-level traits.

```{r, echo=TRUE, results="hide"}
library(maker)
library(dplyr)

m <- maker::maker()
m$make("BCI_species_data")
data <- m$get("BCI_species_data")
head(data)
```

## Model

We now need a function taking as arguments size x, traits, parameters and returning estimated growth rate

```{r, echo=TRUE, results="hide"}
library(tree)

# given a dbh x, traits and parameters, grows plant to right size and estimates growth rate
run_tree <- function(x, E=1, traits=list(), pars=list(), strategy = default_strategy()) {

  distance_from_target_fn <- function(plant) {
    plant$vars_size[["diameter"]] - x
  }

  strategy <- strategy$copy()
  strategy$set_parameters(traits)
  strategy$set_parameters(pars)
  plant1 <- new(Plant, strategy)
  plant1$height <- 0.1
  plant <- try(grow.plant.to.size(plant1, fixed.environment(E), distance.from.target = distance_from_target_fn), silent = FALSE)
  if (inherits(plant, "try-error"))
      return(0)
  run_plant(plant, E)$dbasal_diam_dt
}
```
We can test it works

```{r, echo=TRUE, results="show"}
run_tree(0.01)
run_tree(0.01, traits=list(lma=0.2))
run_tree(0.01, traits=list(lma=0.5))
```

## Fitting the model using  nonlinear regression

To fit the model we can use the package [minpack.lm](http://cran.r-project.org/web/packages/minpack.lm/index.html). According to [this blog](http://www.r-bloggers.com/a-better-nls/), the standard `nls` routine in R does not use the  [Levenberg-Marquardt](http://en.wikipedia.org/wiki/Levenberg%E2%80%93Marquardt_algorithm) approach, which "is more robust,  i.e. finds a solution even if it starts very far off the final minimum. This is because its switching between Gauss-Newton and gradient descent.

Here's the example on how use minpack.lm:

```{r, echo=TRUE, results="show"}
library(minpack.lm)
## values over which to simulate data
x <- seq(0,5,length=100)
## model based on a list of parameters
getPred <- function(parS, xx) parS$a * exp(xx * parS$b) + parS$c
## parameter values used to simulate data
pp <- list(a=9,b=-1, c=6)
## simulated data, with noise
simDNoisy <- getPred(pp,x) + rnorm(length(x),sd=.1)
## plot data
plot(x,simDNoisy, main="data")
## residual function
residFun <- function(p, observed, xx) observed - getPred(p,xx)
## starting values for parameters
parStart <- list(a=3,b=-.001, c=1)
## perform fit
nls.out <- nls.lm(par=parStart, fn = residFun, observed = simDNoisy,
xx = x, control = nls.lm.control(nprint=1))
## plot model evaluated at final parameter estimates
lines(x,getPred(as.list(coef(nls.out)), x), col=2, lwd=2)
## summary information on parameter estimates
summary(nls.out)
```

Now our model. We ned a function which given parameters and covrites predicts growth rates:

```{r, echo=TRUE, results="hide"}
# generative mdoel
# takes pars and traits
getPred <- function(pars, xx) {
	npoints <- nrow(xx)
	zz <- rep(NA_real_, npoints)
	for(i in seq_len(npoints)) {
		zz[i] <- run_tree(xx$dbh[i], E=1, traits=list(lma=xx$lma[i], rho=xx$rho[i], hmat=100), pars=pars)
	}
	zz
}
```

And another which calculates residuals by comparing to observed:

```{r, echo=TRUE, results="hide"}
residFun <- function(p, observed, xx) observed - getPred(p,xx)
```

The set parameters to fit an some lower bounds (optionl) for these:

```{r, echo=TRUE, results="hide"}
par_init <- list(k_l0=0.4, k_s0=0.2, c_NUE=80213)
lower <- c(0,0,2000)
names(lower) <- c("k_l0","k_s0", "c_NUE")
```

Finally prepare the data for comparison. In this trial we'll just use the smallest size class.

```{r, echo=TRUE, results="hide"}
data <- m$get("BCI_species_data") %>%
  mutate(dbh=at) %>%
  filter(!is.na(lma*rho*hmat*dbh*dbasal_diam_dt), dbh %in% c(0.01), hmat > 10) %>%
  select(dbh, lma, rho, dbasal_diam_dt)
```

We can now test our functions work:

```{r, echo=TRUE, results="show"}
x_example <- data[c(1,10), c("dbh", "lma", "rho")]
obs_example <- data[c(1,10), c("dbasal_diam_dt")]
getPred(par_init, x_example)
residFun(par_init, obs_example, x_example)
```

And then fit the model to the data.

```{r, echo=TRUE, results="show"}
x <- data[, c("dbh", "lma", "rho")]
obs <- data$dbasal_diam_dt

nls.out <- nls.lm(par=par_init, fn = residFun, observed = obs, xx = x,
		lower=lower,
		control = nls.lm.control(nprint=1, ftol=1e-4))

out.init <- getPred(par_init, x)
out.fit <- getPred(as.list(coef(nls.out)), x)
```

Finally compare to data:

```{r, echo=TRUE, results="show"}
par(mfrow=c(2,1))
for(trait in c("lma", "rho")) {
	plot(x[[trait]],obs, xlab=trait, ylab= "dbh growth", pch=16)
	xs <- x[order(x[[trait]]),]
	other_trait <- c("lma", "rho")[!c("lma", "rho") %in% trait]
	xs[[other_trait]] <- mean(xs[[other_trait]])

	out.init <- getPred(par_init, xs)
	out.fit <- getPred(as.list(coef(nls.out)), xs)

	lines(xs[[trait]],out.init, col="grey", lwd=2)
	lines(xs[[trait]],out.fit, col=2, lwd=2)
}
mtext(paste("fitting pars", paste(names(par_init), collapse=", ")), side=3, outer=TRUE, line=-2, cex=2)
legend("topright", legend=c("Initial", "Fitted"), col=c("grey", "red"), lty="solid", bty = "n")
```

# Fitting with distribution and likelihood based approach

Using the nls is ok but not ideal because it only returns one plausible combination of parametrs. A better approach would be to fit using a
distribution-based approach. For this there are two challenges:

1. Defining the likelihood
2. Fitting the model

## Defining the likelihood

The main challenge for us is to define the likelihood of a model given observed data.For this type of problem the following approaches have been used:

### Approximate Bayesian computing (ABC)

Suits problems where cannot define likelihood of model, and the model is stochastic, i.e. has some inherent variability

R Packages:

- R package [EasyABC](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12050)
- R package [ABC](http://onlinelibrary.wiley.com/doi/10.1111/j.2041-210X.2011.00179.x)

Our model is determinsitic so does not produce a distribution, rather a single value. hence ABC does not seem so useful

### Make distributional assumption

From [Vrugt 2003](http://doi.org/10.1029/2002WR001642): Assuming that the residuals are mutually independent, Gaussian distributed, with constant variance, the likelihood of a parameter set q(t) for describing the observed data y can be computed using sum of squares of stanadarides error terms..

From [Hartig 2014](http://arxiv.org/abs/1401.8205): Most previous applications of Bayesian statistics to process-based ecological models derive this probability by making distributional assumptions about how observations vary around mean model predictions that are independent of the processes in the model, either ad-hoc or based on the ob- served variance in the data (e.g. Martínez et al., 2011; van Oijen et al., 2013). This is usually justified with the idea in mind that there is observation uncertainty or variability in en- vironmental conditions that is not accounted for in the model

[Van Oijen 2013](http://treephys.oxfordjournals.org/content/25/7/915): The likelihood function p(D θ) is determined by the distribution of the measurement error, assumed to be additive, Gaussian and uncorrelated. Because we have no information about the precision of the measurements of the Skogaby data (Table 3), we choose the standard deviation of each measurement to be 30\% of the mean value.

### Other approaches

[Keenan 2013](http://www.esajournals.org/doi/abs/10.1890/12-0747.1) provides an example of mdoel-data fusion without explictly defining likelihood - uses a cost function for compare multiple outputs aginst data, then samples space using MCMC. Acceptance of a parameter set is based on whether the cost function for each data stream passes a chi-squared test (at 90\% confidence) for acceptance/ rejection, after variance normalization. The cost function quantifies the extent of model–data mismatch using all available data (eddy-covariance, biometric, and so forth).

[LeBauer 2012](http://doi.org/10.1890/12-0137.1): Uses informative prior based on metanalysis, then propagates that through model by running an ensemble to give posterior of model predictions. Does not actually compare to data.

## Sampling

Options for sampling model. [This blog](http://www.sumsar.net/blog/2013/06/three-ways-to-run-bayesian-models-in-r/
) compares some of approaches

[JAGS](http://mcmc-jags.sourceforge.net/)

- Ecologists standard
- By Jags does not take custom functions
- are possibilities for defining own distribution via C++: http://www.cidlab.com/prints/wabersich2013poster.pdf
- some tutorials: http://www.johnmyleswhite.com/notebook/2010/08/20/using-jags-in-r-with-the-rjags-package/, http://www.stat.cmu.edu/~rsteorts/jags_present.pdf

[Rcppbugs](https://github.com/armstrtw/rcppbugs)

- claims to be native R, much faster than JAGS
- can use custom R functions and distributions
- problem installing: `devtools::install_github("armstrtw/rcppbugs")`

### 2015-01-28 fitting.Rmd
Port from tree to tree2

@@
 ```{r, echo=FALSE, results="hide"}
-sourceDir <- function(path, ...) {
+source_dir <- function(path, ...) {
   for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
@@
 }
-sourceDir("R")
+source_dir("R")
 ```
@@
 ```{r, echo=TRUE, results="hide"}
-library(maker)
 library(dplyr)
@@
 m <- maker::maker()
-m$make("BCI_species_data")
-data <- m$get("BCI_species_data")
+data <- m$make("BCI_species_data")
 head(data)
@@
 ```{r, echo=TRUE, results="hide"}
-library(tree)
+library(tree2)
 
 # given a dbh x, traits and parameters, grows plant to right size and estimates growth rate
-run_tree <- function(x, E=1, traits=list(), pars=list(), strategy = default_strategy()) {
-
-  distance_from_target_fn <- function(plant) {
-    plant$vars_size[["diameter"]] - x
-  }
-
-  strategy <- strategy$copy()
-  strategy$set_parameters(traits)
-  strategy$set_parameters(pars)
-  plant1 <- new(Plant, strategy)
-  plant1$height <- 0.1
-  plant <- try(grow.plant.to.size(plant1, fixed.environment(E), distance.from.target = distance_from_target_fn), silent = FALSE)
-  if (inherits(plant, "try-error"))
-      return(0)
-  run_plant(plant, E)$dbasal_diam_dt
+run_tree <- function(x, E=1, traits=list(), pars=list(),
+                     strategy=default_strategy()) {
+  strategy[names(traits)] <- traits
+  strategy[names(pars)] <- pars
+  pp <- grow_plant_to_size(Plant(strategy), x, "diameter",
+                           fixed_environment(E))
+  sapply(pp$plant, function(p) p$vars_growth[["dbasal_diam_dt"]])
 }
@@
 ## values over which to simulate data
-x <- seq(0,5,length=100)
+x <- seq(0, 5, length.out=100)
 ## model based on a list of parameters
-getPred <- function(parS, xx) parS$a * exp(xx * parS$b) + parS$c
+get_pred <- function(pars, xx) pars$a * exp(xx * pars$b) + pars$c
 ## parameter values used to simulate data
-pp <- list(a=9,b=-1, c=6)
+pp <- list(a=9, b=-1, c=6)
 ## simulated data, with noise
-simDNoisy <- getPred(pp,x) + rnorm(length(x),sd=.1)
+sim_d_noisy <- get_pred(pp,x) + rnorm(length(x), sd=.1)
 ## plot data
-plot(x,simDNoisy, main="data")
+plot(x, sim_d_noisy, main="data")
 ## residual function
-residFun <- function(p, observed, xx) observed - getPred(p,xx)
+resid_fun <- function(p, observed, xx) observed - get_pred(p,xx)
 ## starting values for parameters
-parStart <- list(a=3,b=-.001, c=1)
+par_start <- list(a=3, b=-.001, c=1)
 ## perform fit
-nls.out <- nls.lm(par=parStart, fn = residFun, observed = simDNoisy,
+nls_out <- nls.lm(par=par_start, fn = resid_fun, observed = sim_d_noisy,
 xx = x, control = nls.lm.control(nprint=1))
 ## plot model evaluated at final parameter estimates
-lines(x,getPred(as.list(coef(nls.out)), x), col=2, lwd=2)
+lines(x,get_pred(as.list(coef(nls_out)), x), col=2, lwd=2)
 ## summary information on parameter estimates
-summary(nls.out)
+summary(nls_out)
 ```
@@
 # takes pars and traits
-getPred <- function(pars, xx) {
+get_pred <- function(pars, xx) {
   npoints <- nrow(xx)
@@
   for(i in seq_len(npoints)) {
-		zz[i] <- run_tree(xx$dbh[i], E=1, traits=list(lma=xx$lma[i], rho=xx$rho[i], hmat=100), pars=pars)
+    zz[i] <- run_tree(xx$dbh[i], E=1,
+                      traits=list(lma=xx$lma[i], rho=xx$rho[i], hmat=100),
+                      pars=pars)
   }
@@
 ```{r, echo=TRUE, results="hide"}
-residFun <- function(p, observed, xx) observed - getPred(p,xx)
+resid_fun <- function(p, observed, xx) observed - get_pred(p,xx)
 ```
 
-The set parameters to fit an some lower bounds (optionl) for these:
+The set parameters to fit an some lower bounds (optional) for these:
 
@@
 obs_example <- data[c(1,10), c("dbasal_diam_dt")]
-getPred(par_init, x_example)
-residFun(par_init, obs_example, x_example)
+get_pred(par_init, x_example)
+resid_fun(par_init, obs_example, x_example)
 ```
@@
 
-nls.out <- nls.lm(par=par_init, fn = residFun, observed = obs, xx = x,
+nls_out <- nls.lm(par=par_init, fn = resid_fun, observed = obs, xx = x,
                   lower=lower,
@@
 
-out.init <- getPred(par_init, x)
-out.fit <- getPred(as.list(coef(nls.out)), x)
+out.init <- get_pred(par_init, x)
+out.fit <- get_pred(as.list(coef(nls_out)), x)
 ```
@@
 
-	out.init <- getPred(par_init, xs)
-	out.fit <- getPred(as.list(coef(nls.out)), xs)
+  out.init <- get_pred(par_init, xs)
+  out.fit <- get_pred(as.list(coef(nls_out)), xs)
 
@@
 
-Using the nls is ok but not ideal because it only returns one plausible combination of parametrs. A better approach would be to fit using a
-distribution-based approach. For this there are two challenges:
+Using the nls is ok but not ideal because it only returns one plausible combination of parametrs.
+A better approach would be to fit using a distribution-based approach.
+For this there are two challenges:
 

### 2015-01-28 R/figures.R

@@ -1,364 +1,326 @@
-
-fig_size_dpdt <- function() {
-  figure_size(ylim = c(0, 500), yvars = c("net_production", "assimilation", "respiration", "turnover"))
+default_strategy <- function() {
+  Strategy(B5=1.5,
+           k_l0=0.4565855/3,
+           hmat=30.0,
+           c_r1=0.8,
+           c_r2=20,
+           k_s0=0.2,
+           a1=2.17,
+           B1=0.546)
 }
 
-fig_size_dhdt <- function() {
-  figure_size(yvars = c("height_growth_rate", "dheight_dleaf_area", "leaf_fraction", "growth_fraction", "net_production"))
+run_plant_to_heights <- function(heights, strategy, env) {
+  f <- function(h, p) {
+    p$height <- h
+    run_plant(p, env)
+  }
+  ## NOTE: previously we checked through these results and filtered
+  ## out things that left NULL
+  do.call("rbind", lapply(heights, f, Plant(strategy)))
 }
 
-fig_size_daldt <- function() {
-  figure_size(yvars = c("dleaf_area_dt", "leaf_fraction", "growth_fraction", "net_production"))
+run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
+                               time_max=Inf) {
+  res <- grow_plant_to_size(Plant(strategy), sizes, size_variable,
+                            env, time_max)
+  do.call("rbind", lapply(res$plant, plant_info))
 }
 
-# growth decomp
-figure_size <- function(yvars = c("height_growth_rate", "dheight_dleaf_area", "leaf_fraction", "growth_fraction", "net_production"), E = 1, strategy = default_strategy(),
-  ...) {
+run_plant <- function(plant, env) {
+  plant$compute_vars_phys(env)
+  plant_info(plant)
+}
 
-  h <- seq(0.1, 50, length.out = 50)
-  plants <- change_with_height(h = h, E = E, strategy = strategy)
+plant_info <- function(p) {
+  x <- as.data.frame(as.list(c(p$vars_size,
+                               p$vars_phys,
+                               p$vars_growth)))
+  # add relative measures
+  x$height_growth_rate_relative    <- x$height_growth_rate / x$height
[…]

### 2015-01-28 R/figures_data.R

@@ -1,0 +1,41 @@
+figure_lma_tradeoff <- function(data) {
+
+  op <- par(oma = c(4, 4, 1, 1))
+
+  sites <- data[["dataset"]]
+  lma <- data[["lma"]]
+  leaf_turnover <- data[["leaf_turnover"]]
+
+  biomes <- sort(unique(data[["biome"]]))
+  nbiomes <- length(biomes)
+
+  col.table <- nice_colors(nbiomes)
+  names(col.table) <- biomes
+
+  # only use sites with n > 5
+  site.n <- data.frame(table(sites))
+  i <- !is.na(lma) & !is.na(leaf_turnover) & site.n[match(sites, as.character(site.n[, 1])), 2] > 3
+
+  new_plot("lma", "leaf_turnover", log = "xy", line = 5, xlim = c(0.01, 1.28))
+  points(lma[i], leaf_turnover[i], type = "p", col = make_transparent("grey", 0.3), pch = 16, cex = 0.9)
+
+  sm1 <- sma(leaf_turnover[i] ~ lma[i] * sites[i], log = "xy")
+  plot(sm1, add = T, col = colour_by_category(data[["biome"]][match(sm1[["groups"]], sites)], col.table), type = "l", lwd = 1, p.lines.transparent = 0.15)
+
+  x <- seq_log(0.001, 3, length.out = 100)
+
+  strategy <- default_strategy()
+  B4 <- strategy$parameters$B4
+  k_l0 <- strategy$parameters$k_l0
+  lma_0 <- strategy$parameters$lma_0
+
+  points(x, k_l0 * (x/lma_0)^-B4, type = "l", col = "black", lwd = 2)
+
+  lab <- paste0(length(unique(sites[i])), " sites, ", sum(!is.na(leaf_turnover[i]) & !is.na(lma[i])), " species")
+  legend("topright", legend = lab, bty = "n", cex = 0.5)
+
+  lab <- c("", tolower(names(col.table)))
+  legend("topright", legend = lab, bty = "n", pch = c(NA, rep(16, nbiomes)), col = c(NA, col.table), cex = 0.5)
+
+  par(op)
+}

### 2015-01-28 R/tree.R

@@ -1,159 +1,0 @@
-
-default_strategy <- function() {
-  strategy <- new(Strategy)
-  strategy$set_parameters(structure(list(1.5, 0.4565855/3, 30, 0.8, 20, 0.2, 2.17, 0.546), names = c("B5", "k_l0", "hmat", "c_r1", "c_r2", "k_s0", "a1", "B1")))
-  strategy
-}
-
-## This makes a fixed light environment over the plant height,
-fixed.environment <- function(E, height, n = 101, light.env = NULL, n.strategies = 1, seed.rain = 0) {
-  if (length(seed.rain) == 1)
-    seed.rain <- rep(seed.rain, n.strategies)
-  hh <- seq(0, height, length = n)
-  if (is.null(light.env))
-    light.env <- function(x) rep(E, length.out = length(x))
-  ee <- light.env(hh)
-  env <- new(Interpolator)
-  env$init(hh, ee)
-
-  parameters <- new(Parameters)
-  for (i in seq_len(n.strategies)) parameters$add_strategy(default_strategy())
-  parameters$seed_rain <- seed.rain
-
-  ret <- new(Environment, parameters)
-  ret$light_environment <- env
-  attr(ret, "light.env") <- light.env
-  ret
-}
-
-# runs a new plant in given environment
-run_plant <- function(plant, E = 1) {
-
-  env <- fixed.environment(E)
-  plant$compute_vars_phys(env)
-
-  x <- data.frame(t(plant$vars_size), t(plant$vars_phys), t(plant$vars_growth_decomp))
-
-  # add relative measures
-  x[["height_growth_rate_relative"]] <- x[["height_growth_rate"]]/x[["height"]]
-  x[["dbasal_area_dt_relative"]] <- x[["dbasal_area_dt"]]/x[["area_basal"]]
-  x[["dbasal_diam_dt_relative"]] <- x[["dbasal_diam_dt"]]/x[["diameter"]]
-  x[["dmass_above_ground_dt_relative"]] <- x[["dmass_above_ground_dt"]]/x[["mass_above_ground"]]
-
-  x
-}
-
-change_with_height <- function(h = seq(0.2, 1, length.out = 10), E = 1, strategy = default_strategy()) {
-
-  myfun <- function(x) {
-    plant <- new(Plant, strategy)
-    plant$height <- x
-    run_plant(plant, E)
-  }
-
-  lapply(h, myfun)
-}
-
-# e.g.  = function(plant){plant$height - 0.25}
[…]

### 2015-01-28 R/utils.R

@@ -27,18 +27,6 @@ render_md_as_html <- function(filename) {
   rmarkdown::render(filename, "html_document", quiet = TRUE)
 }
 
-na.clean <- function(x) {
-  x[!is.na(x)]
-}
-
-seq_log_range <- function(r, length.out) {
-  seq_log(r[[1]], r[[2]], length.out)
-}
-
-seq_log <- function(from, to, length.out) {
-  exp(seq(log(from), log(to), length.out = length.out))
-}
-
 #loads an RData file, and returns it
 load_rdata <- function(fileName){
     load(fileName)

### 2015-01-28 sensitivity.R

@@ -1,92 +1,90 @@
 # Fitting TREE to observed data
 
-sourceDir <- function(path, ...) {
-     for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
-        source(file.path(path, nm), ...)
-     }
+source_dir <- function(path, ...) {
+  for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
+    source(file.path(path, nm), ...)
+  }
 }
-sourceDir("R")
+source_dir("R")
 
-library(tree)
-library(maker)
-library(plyr)
-library(dplyr)
+library(tree2)
+suppressPackageStartupMessages({
+  library(plyr)
+  library(dplyr)
+})
 
 # First, look at plot: challenge is to move peak to left and pull down elevation of curve.
-
 m <- maker::maker()
-data <- m$get("BCI_species_data") %>%
+data <- m$make("BCI_species_data") %>%
   mutate(dbh=at) %>%
   filter(!is.na(lma*rho*hmat*dbh*dbasal_diam_dt), dbh %in% c(0.01), hmat > 10) %>%
   select(dbh, lma, rho, dbasal_diam_dt)
 
-figure_trait_effect_at_size("lma", "dbasal_diam_dt", "diameter", 0.01, xlim = c(0.01, 0.1), strategy = default_strategy())
-points(data$lma, data$dbasal_diam_dt)
+d <- trait_effects_data("lma", "diameter")
+d <- d[d$size_class == 2L,]
+
+plot(data$lma, data$dbasal_diam_dt, las=1, log="x",
+     xlab=name_pretty("lma"), ylab=name_pretty("dbasal_diam_dt"),
+     xlim=range(data$lma, d$lma),
+     ylim=range(data$dbasal_diam_dt, d$dbasal_diam_dt))
+lines(d$lma, d$dbasal_diam_dt, col="red")
 
 # given a dbh x, traits and parameters, grows plant to right size and estimates growth rate
-growth <- function(trait, x, D, E=1, strategy = default_strategy()) {
-
-  distance_from_target_fn <- function(plant) {
-    plant$vars_size[["diameter"]] - D
-  }
-
-  strategy <- strategy$copy()
-  strategy$set_parameters(structure(list(x), names = trait))
-  plant <- new(Plant, strategy)
-  plant$height <- 0.1
-
-  plant <- try(grow.plant.to.size(plant, fixed.environment(E), distance.from.target = distance_from_target_fn), silent = FALSE)
[…]

### 2015-01-28 R/figures.R
Port over the sma fit figure

@@ -170,6 +170,7 @@ name_pretty <- function(trait) {
          hmat="Height at maturation (m)",
          dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
          shading=expression(paste("Maximum leaf area above (", m^2, " )")),
+         leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, " )")),
 
          height_growth_rate=expression(paste("Height growth (m ", yr^-1, " )")),
          dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),

### 2015-01-28 R/figures_data.R

@@ -1,7 +1,4 @@
 figure_lma_tradeoff <- function(data) {
-
-  op <- par(oma = c(4, 4, 1, 1))
-
   sites <- data[["dataset"]]
   lma <- data[["lma"]]
   leaf_turnover <- data[["leaf_turnover"]]
@@ -9,33 +6,34 @@ figure_lma_tradeoff <- function(data) {
   biomes <- sort(unique(data[["biome"]]))
   nbiomes <- length(biomes)
 
-  col.table <- nice_colors(nbiomes)
-  names(col.table) <- biomes
+  col_table <- nice_colors(nbiomes)
+  names(col_table) <- biomes
 
-  # only use sites with n > 5
-  site.n <- data.frame(table(sites))
-  i <- !is.na(lma) & !is.na(leaf_turnover) & site.n[match(sites, as.character(site.n[, 1])), 2] > 3
-
-  new_plot("lma", "leaf_turnover", log = "xy", line = 5, xlim = c(0.01, 1.28))
-  points(lma[i], leaf_turnover[i], type = "p", col = make_transparent("grey", 0.3), pch = 16, cex = 0.9)
-
-  sm1 <- sma(leaf_turnover[i] ~ lma[i] * sites[i], log = "xy")
-  plot(sm1, add = T, col = colour_by_category(data[["biome"]][match(sm1[["groups"]], sites)], col.table), type = "l", lwd = 1, p.lines.transparent = 0.15)
-
-  x <- seq_log(0.001, 3, length.out = 100)
+  i <- !is.na(lma) & !is.na(leaf_turnover) & table(sites)[sites] > 3
+  sm1 <- sma(leaf_turnover[i] ~ lma[i] * sites[i], log="xy")
+  col_sm1 <- col_table[data[["biome"]][match(sm1[["groups"]], sites)]]
 
   strategy <- default_strategy()
-  B4 <- strategy$parameters$B4
-  k_l0 <- strategy$parameters$k_l0
-  lma_0 <- strategy$parameters$lma_0
+  B4 <- strategy$B4
+  k_l0 <- strategy$k_l0
+  lma_0 <- strategy$lma_0
+  x <- seq_log(10^par("usr")[1], 10^par("usr")[2], length.out=10)
+  y <- k_l0 * (x / lma_0)^(-B4)
 
-  points(x, k_l0 * (x/lma_0)^-B4, type = "l", col = "black", lwd = 2)
+  par(mar=c(4.6, 4.6, .5, .5))
[…]

### 2015-01-28 R/utils-plot.R

@@ -15,21 +15,10 @@ nice_colors <- function(n = 80) {
   cols[1:n]
 }
 
-
 make_transparent <- function(col, opacity = 0.5) {
   tmp <- col2rgb(col)/255
   rgb(tmp[1, ], tmp[2, ], tmp[3, ], alpha = opacity)
 }
-
-## Position label at a fractional x/y position on a plot
-label <- function(px, py, lab, ..., adj = c(0, 1)) {
-  usr <- par("usr")
-  text(usr[1] + px * (usr[2] - usr[1]), usr[3] + py * (usr[4] - usr[3]), lab, adj = adj, ...)
-}
-
-is.wholenumber <- function(x, tol = .Machine$double.eps^0.5) abs(x - round(x)) < tol
-
-colour_by_category <- function(x, table) unname(table[x])
 
 linear_rescale <- function(x, range, scale = range(x)) {
   p <- (x - scale[[1]])/(scale[[2]] - scale[[1]])

### 2015-01-28 R/tables.R
Minor cleanup

@@ -1,6 +1,9 @@
-
 table_tree_parameters <- function(tree_parameters, dest){
-	x<-xtable(tree_parameters[-c(1,2,4,7,10,21,22,25:31),1:4],hline.after=c(1), align='lp{5cm}lll')
-	y <- print(x,sanitize.text.function=function(str)as.character(str),include.rownames=FALSE,floating=FALSE,print.results = FALSE);
-	cat(y, file =dest)
+  x <- xtable(tree_parameters[-c(1,2,4,7,10,21,22,25:31),1:4],
+              hline.after=c(1),
+              align='lp{5cm}lll')
+  y <- print(x, sanitize.text.function=as.character,
+             include.rownames=FALSE, floating=FALSE,
+             print.results=FALSE)
+  cat(y, file=dest)
 }

### 2015-01-28 R/BCI.R
Port BCI figure over

@@ -216,61 +216,44 @@ quantreg_plot <- function(x, y, p = 0.99, nmin = 100, predict_at = 10, add.legen
     legend("topright", legend = c("Wright2010", "QuantReg"), col = c("red", "blue"), lty = c("solid", "solid"), bty = "n")
 }
 
-figure_trait_growth_data <- function(X, Y, cex = 1, col = "darkgreen", pch = 1, ylim = range(Y, na.rm = TRUE), axes = c("lma", "dbasal_diam_dt"), ytick.lab = TRUE, ...) {
-  if (ytick.lab){
-    new_plot(axes[1], axes[2], log = "xy", ylim = ylim, xlab = NULL, ylab = NULL, ...)
-  } else{
-    new_plot(axes[1], axes[2], log = "xy", ylim = ylim, xlab = NULL, ylab = NULL, ytick.lab = NA, ...)
+figure_BCI_data <- function(data) {
+  traits <- c("lma", "rho", "hmat")
+  at <- c(10, 20, 50, 100, 200)/1000
+
+  op <- par(oma=c(3, 6, 3, 1),
+            mar=c(1, 1, 2, 1),
+            mfcol=c(length(at), length(traits)))
+
+  data <- data[data$count > 800,]
+  dsplit <- lapply(at, function(a) data[data$at == a,])
+
+  ylim <- c(1, 30) / 1000
+  cex_lab <- .85
+
+  for (trait in traits) {
+    xlim <- range(data[[trait]], na.rm=TRUE)
+    for (i in seq_along(at)) {
+      dsub <- dsplit[[i]]
+      cex <- linear_rescale(log10(dsub$count), c(0.6, 2.5), log10(c(800, 10000)))
+      x <- dsub[[trait]]
+      y <- dsub$dbasal_diam_dt
+
+      sm <- sma(y ~ x, log="xy", method="OLS")
+
+      plot(x, y, pch=1, xlim=xlim, ylim=ylim, log="xy", cex=cex, xaxt="n", yaxt="n")
+      plot(sm, add=TRUE, col="darkgreen", lwd=2, type="l", p.lines.transparent=0.1)
+
+      lab <- sprintf("r2 = %2.2f, n=%d",
+                     sm[["groupsummary"]][["r2"]],
+                     sum(!is.na(y) & !is.na(x)))
+      legend("topright", legend=lab, bty="n", cex=1.25)
+
+      axis(1, labels=i == length(at))
[…]

### 2015-01-28 R/axis.R

@@ -1,132 +1,29 @@
+name_pretty <- function(trait) {
+  switch(trait,
+         lma=expression(paste("Leaf-construction cost (kg ", m^-2, " )")),
+         rho=expression(paste("Wood density (kg ", m^-3, " )")),
+         hmat="Height at maturation (m)",
+         dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
+         shading=expression(paste("Maximum leaf area above (", m^2, " )")),
+         leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, " )")),
 
+         height_growth_rate=expression(paste("Height growth (m ", yr^-1, " )")),
+         dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
+         dbasal_area_dt=expression(paste("Basal area growth (", m^2, " ", yr^-1, ")")),
+         dmass_above_ground_dt=expression(paste("Mass growth (kg ", yr^-1, ")")),
 
-get_axis_info <- function(name, parToGet, option=1) {
+         height_growth_rate_relative=expression(paste("Relative height growth (m ", yr^-1, " )")),
+         dbasal_diam_dt_relative=expression(paste("(Relative diameter growth (m ", yr^-1, ")")),
+         dbasal_area_dt_relative=expression(paste("(Relative basal area growth (", m^2, " ", yr^-1, ")")),
+         dmass_above_ground_dt_relative=expression(paste("(Relative mass growth (kg ", yr^-1, ")")),
 
-  # TRAITS
-  if (name == "lma") {
-    ticks <- 0.01 * 2^seq(0, 20)
-    lab <- expression(paste("Leaf-construction cost (kg ", m^-2, " )"))
-    lim <- c(0.02, 1.0)
-  }
-  if (name == "hmat") {
-    ticks <- 2^seq(-4, 20)
-    lab <- expression(paste("Maximum height (m)"))
-    lim <- c(3, 60)
-  }
-  if (name == "rho") {
-    ticks <- 50 * 2^seq(0, 10)
-    lab <- expression(paste("Wood density (kg ", m^-3, " )"))
-    lim <- c(200, 1200)
-  }
-
-  # SIZE
[…]

### 2015-01-28 R/figures.R

@@ -161,38 +161,6 @@ long_to_wide <- function(dat, group, cols) {
   ret
 }
 
-
-## Axis stuff; just done as constants:
-name_pretty <- function(trait) {
-  switch(trait,
-         lma=expression(paste("Leaf-construction cost (kg ", m^-2, " )")),
-         rho=expression(paste("Wood density (kg ", m^-3, " )")),
-         hmat="Height at maturation (m)",
-         dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
-         shading=expression(paste("Maximum leaf area above (", m^2, " )")),
-         leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, " )")),
-
-         height_growth_rate=expression(paste("Height growth (m ", yr^-1, " )")),
-         dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
-         dbasal_area_dt=expression(paste("Basal area growth (", m^2, " ", yr^-1, ")")),
-         dmass_above_ground_dt=expression(paste("Mass growth (kg ", yr^-1, ")")),
-
-         height_growth_rate_relative=expression(paste("Relative height growth (m ", yr^-1, " )")),
-         dbasal_diam_dt_relative=expression(paste("(Relative diameter growth (m ", yr^-1, ")")),
-         dbasal_area_dt_relative=expression(paste("(Relative basal area growth (", m^2, " ", yr^-1, ")")),
-         dmass_above_ground_dt_relative=expression(paste("(Relative mass growth (kg ", yr^-1, ")")),
-
-         stop("Unknown trait ", trait))
-}
-## TODO: Identify and replace hard coded numbers
-trait_range <- function(trait) {
-  switch(trait,
-         lma=c(0.02, 1.0),
-         rho=c(200,  1200),
-         hmat=c(3, 60),
-         stop("No standard range for trait ", trait))
-}
-
 plant_list_set_height <- function(x, height) {
   for (i in x) {
     i$height <- height

### 2015-01-28 R/BCI.R
Switch to using downloaded copy of BCI census data. Closes #33

@@ -1,17 +1,33 @@
 #List of functions used to manipulate and summarise data from BCI.
 
-#Download BCI
-BCI_download_50ha_plot <- function(dest) {
-  if (!file.exists(dest))
-    stop(sprintf("Please visit %s and save file as %s. See Readme file for further instructions.", "http://ctfs.arnarb.harvard.edu/webatlas/datasets/bci/", dest))
+#Download BCI data
+# download function from package downloader provides wrapper
+# to download file so that works for https and across platforms
+BCI_download_50ha_plot_full <- function(dest) {
+  url <-"https://repository.si.edu/bitstream/handle/10088/20925/bci.full.Rdata31Aug2012.zip"
+  download(url, dest, mode="wb")
+}
+
+BCI_download_50ha_plot_stem <- function(dest) {
+  url <-"https://repository.si.edu/bitstream/handle/10088/20925/bci.stem.Rdata31Aug2012.zip"
+  download(url, dest, mode="wb")
+}
+
+BCI_download_species_table <- function(dest) {
+  url <-"https://repository.si.edu/bitstream/handle/10088/20925/bci.spptable.rdata"
+  download(url, dest, mode="wb")
 }
 
 #Load 50ha census data
-BCI_load_50ha_plot <- function(path_to_file) {
-  path_to_BCI_data <- dirname(path_to_file)
-  files <- data.frame(year = c(1990, 1995, 2000, 2005, 2010),
-                      name = paste0(path_to_BCI_data, "/bci.stem", 3:7, ".csv"), stringsAsFactors = FALSE)
-  tbl_df(ddply(files, "year", function(x) read.csv(x[["name"]], stringsAsFactors = FALSE)))
+BCI_load_50ha_plot <- function(path_to_zip) {
+
+  tmp <- tempfile()
+  unzip(path_to_zip, exdir=tmp)
+  on.exit(unlink(tmp, recursive=TRUE))
+
+  files<- data.frame(name=list.files(tmp, pattern=".rdata", full.names=TRUE),
+                    year = c(1980, 1985, 1990, 1995, 2000, 2005, 2010), stringsAsFactors = FALSE)
+  tbl_df(ldply(list.files(tmp, pattern=".rdata", full.names=TRUE), function(x) load_rdata(x)))
 }
 
 #Look up family
@@ -21,10 +37,10 @@ lookup_family <- function(tag, nomen){
 }
 
 #Look up species code
[…]

### 2015-01-28 R/utils.R

@@ -38,3 +38,9 @@ seq_log_range <- function(r, length.out) {
 seq_log <- function(from, to, length.out) {
   exp(seq(log(from), log(to), length.out = length.out))
 }
+
+#loads an RData file, and returns it
+load_rdata <- function(fileName){
+    load(fileName)
+    get(ls()[ls() != "fileName"])
+}

### 2015-02-04 ms.tex
shift to using latex

% THIS FILE TAKEN FROM PANDOC DEMOS PAGE:http://johnmacfarlane.net/pandoc/demos.html,http://johnmacfarlane.net/pandoc/demo/mytemplate.tex

\documentclass[12pt, a4paper]{article}
\RequirePackage[hmargin=2cm,vmargin=2cm]{geometry}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

\usepackage{fontspec}
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}

% Fancy HEADER
\usepackage{fancyhdr}
\pagestyle{fancy}
\pagenumbering{arabic}
\lhead{\itshape{\nouppercase{\leftmark}}}
\chead{}
\rhead{\thepage}
%\lfoot{v }
\cfoot{}
%\rfoot{\thepage}

% \usepackage{natbib}
% \bibliographystyle{abbrvnat}
% \setcitestyle{authoryear,open={(},close={)}}

\usepackage[sort&compress]{natbib}
\citestyle{nature}
\bibliographystyle{ieeetr}

\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother

\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Untangling the link between traits, size and growth rate in plants},
            colorlinks=true,
            citecolor=black,
            urlcolor=grey,
            linkcolor=grey}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines

\usepackage{titling}

\title{Untangling the link between traits, size and growth rate in plants}

\author{
{\bf Daniel Falster} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{daniel.falster@mq.edu.au}\vspace{0.1in}\\
{\bf Rich FitzJohn} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{richard.fitzjohn@mq.edu.au}\vspace{0.1in}\\
{\bf Joe Wright} \\ \normalsize\emph{Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama} \\ \url{wrightj@si.edu}\vspace{0.1in}\\ }

\date{}

\begin{document}
\thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that
{\LARGE \bf \thetitle \vspace{0.2in}}

\theauthor

\section{Abstract}\label{abstract}

Plant growth is the foundation of the terrestrial biosphere, powering
food-webs, the carbon cycle and much of human
activity \citep{raupach_carbon_2010}. Plant species differ in many
functional traits, including aspects of tissue construction and energy
allocation, that drive differences in mechanical strength, biomass
partitioning, and rates of photosynthesis and tissue
turnover\citep{wright_world-wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
such traits influence whole-plant growth, with the simple relationships
predicted by existing theory often lacking empirical
support\citep{wright_functional_2010,poorter_are_2008}. Here we
provide a new set of theoretical predictions, validated using 20 years
of growth data\citep{condit_barro_2012,kitajima_leaf_2013}, showing
that the influence of three prominent traits on growth rate is moderated
by plant size, and that the impact of size on the nature of
relationships varies among traits. We extend a widely-used theoretical
model that links growth rate in seedlings with a single leaf
trait\citep{lambers_inherent_1992,wright_cross-species_2000} to
explicitly include influences of size, light environment, and three
other prominent traits. We find the influence of leaf construction cost
(\(\phi\)) on growth is strong in small plants but weakens with size. In
contrast, the influence of stem construction cost (\(\rho\)) peaks at
intermediate sizes, while the influence of height at maturation
(\(h_m\)) strengthens with size. Our theoretical model suggests the
change in trait-growth relationships with size arises due to changes in
biomass partitioning and reproductive allocation. Moreover, we show how
traits moderate plant responses to light environment and also determine
shade tolerance, supporting diverse empirical
results\citep{ruger_functional_2012,poorter_leaf_2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, our results provide a solid theoretical foundation for
trait ecology and thus provide a platform for understanding growth
across diverse species around the world.

\section{Main text}\label{main-text}

Functional traits capture core differences in the strategies plants use
to generate and invest surplus energy\citep{wright_world-wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
traits now exists for many of the world's 250000 plant
species\citep{cornwell_functional_2014}; yet the effect of these traits
on growth remains poorly understood. Prominent traits such as \(\phi\)
(also known as leaf mass per area) are widely considered as indicators
of growth rate\citep{reich_world-wide_2014}, because cheaper leaf
construction should enable faster canopy
expansion\citep{blackman_compound_1919,westoby_time_2000}. Indeed,
growth rate in seedlings is strongly and negatively correlated with
\(\phi\)\citep{cornelissen_seedling_1996,wright_cross-species_2000}
(Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
might also give a growth advantage in larger
plants\citep{enquist_general_2007}, an increasing number of empirical
studies have failed to detect any such
effect\citep{wright_functional_2010, @herault_functional_2011,
@poorter_are_2008}. Meanwhile, expectations on how other prominent
traits might influence growth are lacking. The growing discord between
theoretical expectation and empirical result thus indicates a strong
need for a new theoretical framework through which the influence of
traits and size on plant growth can be understood.

Here we extend a widely-used theoretical model that links growth rate in
seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross-species_2000}, to explicitly include influences of
size, light environment, and other prominent traits. This extension
allows us to generate novel predictions on the link between four key
traits and plant demography (Table \ref{tab:predictions}). Predictions
are compared against data from both a long-term
plot\citep{condit_barro_2012} and field
experiment\citep{kitajima_leaf_2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{S-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies\citep{wright_functional_2010}. The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
\ref{tab:predictions}) showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

We begin with a standard model for the amount of biomass available for
growth, \(dP / \textrm{d}t\), given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant\citep{falster_influence_2011}:

\begin{equation}\label{eq:dPdt}
\underbrace{\strut \frac{dP}{\textrm{d}t}}_\textrm{net biomass production}= \underbrace{\strut Y}_\textrm{yield} \big(\underbrace{\strut \frac{m_\textrm{l}}{\phi} \, A}_\textrm{photosynthesis} - \big(\underbrace{\strut \, \frac{m_\textrm{l}}{\phi} \, r_\textrm{l} + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)\big) - \big(\underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}\big).
\end{equation}

Here, \(m,r\), and \(k\) refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts \(l\)=leaves,
\(b\)=bark, \(s\)=sapwood and \(r\)=roots. \(A\) is the assimilation
rate of CO\(_2\) per leaf area and \(Y\) is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
\(a_\textrm{l} = m_\textrm{l} / \phi\), while total mass of living
tissues
\(m_\textrm{t}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.\)
Eq. \ref{eq:dPdt} assumes that tissues lost via turnover are replaced
before new growth occurs.

Current theory emphasises the relationship between \(\phi\) and relative
growth in mass\citep{enquist_general_2007},
\(\textrm{d}P / \textrm{d}t / m_\textrm{t}\). For seedlings, which are
young and mostly leaf, it is reasonable to ignore all turnover terms as
well as the respiration terms for non-leaf tissues. Net production then
becomes a linear function of leaf area and net photosynthesis per leaf
area (\(A_\textrm{net} = Y(A - r_\textrm{l})\)), making relative growth
rate a linear function of \(\phi\):

\begin{equation}\label{eq:RGR}
\underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}\frac{1}{m_\textrm{t}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{t}}. \end{equation}

Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
relation to \(\phi\)\citep{wright_cross-species_2000}, this
approximation does not directly link to other traits, or to the
variables that are routinely collected for large trees: namely plant
height (\(h\)) and stem cross-sectional area (\(a_\textrm{st}\)) or
diameter \(D\).

To model growth in either height or basal area requires we to account
not just for mass production, but also for the costs of building new
tissues, allocation to reproduction, and architectural layout. In eq.
\ref{eq:RGR}, these terms are missing, whereas in the model presented
here they are made explicit.

Mathematically, height growth can be decomposed into a product of
physiologically relevant terms\citep{falster_influence_2011}:

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}}_\textrm{mass production}.
\end{equation}

The first term on the right of eq \ref{eq:dhdt},
\(\textrm{d}h / \textrm{d}a_\textrm{l}\), is the growth in plant height
per unit growth in total leaf area; accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension\citep{poorter_architecture_2006}.

The second term, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\),
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\)
can itself be expressed as a sum of construction costs per unit leaf
area:

\begin{equation}\label{eq:daldmt}
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
= \big(\phi
 + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
\end{equation}

The third term in eq \ref{eq:dhdt},
\(\textrm{d}m_\textrm{t} / \textrm{d}P\), gives the fraction of net
biomass production (eq. \ref{eq:dPdt}) that is allocated to growth
rather than reproduction or storage.

In a similar way, basal area (\(a_\textrm{st}\)) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
(\(a_\textrm{ss}, a_\textrm{sb}, a_\textrm{sh}\) respectively):
\(\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}= \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}\).
Noting that heartwood grows via turnover of sapwood, such that
\(\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} a_\textrm{ss}\),
we have an equation for basal area growth that conatins many of the same
elements as eq. \ref{eq:dhdt}:

\begin{equation}\label{eq:dast}
\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}} \times \frac{\textrm{d}m_\textrm{t}}{\textrm{d}P} \times \frac{\textrm{d}P}{\textrm{d}t} + k_\textrm{s} a_\textrm{ss}.
\end{equation}

Diameter growth is then given by the geometric relationship between stem
diameter (\(D\)) and \(a_\textrm{st}\):

\begin{equation} \label{eq:dDdt}
\frac{\textrm{d}D}{\textrm{d}t}= \sqrt{\frac{\pi}{a_\textrm{st}}} \frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}.
\end{equation}

Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
therefore hold for any model of plant growth. To make explicit
predictions, however, requires two additional elements: 1) A specific
model describing how the various mass and area terms
(\(m_\textrm{l}, m_\textrm{s}, m_\textrm{b}, m_\textrm{r}, a_\textrm{l}, a_\textrm{s}, a_\textrm{b}\))
vary relative to one another; and 2) For any trade-offs in plant
function to be specified; this is where traits have an influence. These
two components are as follows.

To address the first of these challenges, we use a functional balance
model\citep{yokozawa_foliage_1995,falster_influence_2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{f-assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

This new growth model captures the intrinsically size-dependent nature
of plant growth (Table \ref{tab:predictions}a), recovering well-known
empirical patterns\citep{sillett_increasing_2010, king_size-related_2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size\citep{sillett_increasing_2010,
@stephenson_rate_2014}, due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve\citep{herault_functional_2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates\citep{iida_linking_2014}.

We now incorporate the effects of some prominent traits by embedding
explicit trade-offs within this growth model. Variation in each trait
implies benefits and costs. These affect specific elements of the growth
equations presented above (Table \ref{tab:trade-offs}).

By definition, the main effect of increasing \(\phi\) is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of \(\phi\) on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction\citep{wright_world-wide_2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dPdt} to \(\phi\). We use an allometric scaling relationship of
the form \(k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}\), which has been
shown to describe patterns across diverse
ecosystems\citep{wright_world-wide_2004} (Fig. \ref{fS-leaf}).

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between \(\phi\) and growth that
changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
Decreasing \(\phi\) has two impacts on growth rate. First, lower
\(\phi\) increases marginal leaf deployment
(\(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\) by economising on
construction costs. Second, lower \(\phi\) decreases net production
(\(\textrm{d}P / \textrm{d}t\)), due to increased leaf turnover. Whether
lower \(\phi\) increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing \(\phi\) increases growth
rate. When plants are large, the influence of \(\phi\) on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
\(\phi\) is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

As for leaf, cheaper stem construction (lower \(\rho\)) decreases the
cost of deploying a unit of leaf area, and may thereby increase growth
rates. However, the cost of cheap stem construction must also be
accounted for. However, the direct physiological trade-offs of lower
\(\rho\) are less understood than for \(\phi\). One possibility is that
lower \(\rho\) decreases mechanical strength resulting in higher
mortality\citep{chave_towards_2009,wright_functional_2010}. Under
this scenario, decreasing \(\rho\) will always provide a growth
advantage, because the costs of low \(\rho\) are not realised within the
terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
\ref{tab:trade-offs}). In addition, lower \(\rho\) may increase turnover
of sapwood. Mirroring our assumptions for leaves, we let sapwood
turnover decrease with \(\rho\),
i.e.\(k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}\), with \(\beta_5 > 1\).
Under this scenario, decreasing \(\rho\) provides an advantage in height
growth only when the benefits of low \(\rho\) outweigh the costs. This
effect is strongest at intermediate sizes. The decline in height growth
at low \(\rho\) with size is more marked than for either stem basal-area
and diameter growth, because high sapwood turnover also contributes to
this growth measure. Thus diameter growth continues, even as net
production approaches zero.

Unlike leaf and stem traits, height at maturation (\(h_m\)) moderates
growth by adjusting the amount of energy invested in growth, i.e.~the
term \(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) in eqs.
\ref{eq:dhdt} and \ref{eq:dast}. Greater height at maturation (\(h_m\))
can thus lead to a growth advantage by increasing
\(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) (Table
\ref{tab:predictions}). At smaller sizes, there is no differentiation
among species, because most species are focusing on growth (Fig.
\ref{f-BCI}). At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate (Fig. \ref{f-BCI}).

Two other important predictions arise directly from linking traits to
growth rate (Table \ref{tab:predictions}c). First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low \(rho\) increase growth more substantially with incraeses in
light\citep{ruger_functional_2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low \(\phi\) are lost (Fig. \ref{f-growth_light}). This results also
agrees with empirical results showing that growth for species with low
\(rho\) actually decreases as size
increases\citep{ruger_functional_2012}. In contrast, growth rate tends
to increase in size for the species with highest
\(\rho\)\citep{ruger_functional_2012}.

Variation in \(\phi\) also moderates the response of growth to changes
in light, with species having the lowest \(rho\) being most responsive.
However, unlike for \(\rho\), the effect appears only for the smallest
size classes. Accordingly, empirical analyses have detected such a link
in small plants but not for plants \textgreater{} 0.01m
dbh\citep{ruger_functional_2012}.

Second, species with low \(\phi\) and low \(\rho\) are predicted to be
less shade tolerant (Fig. \ref{f-wplcp}). At low \(\phi\) (\(\rho\)),
leaf (sapwood) turnover is higher and thus a greater light income is
needed to offset these costs. As previously suggested
\citep{givnish_adaptation_1988}, shade-tolerance also decreases with
height because as size increase, the total amount of energy needed to
offset respiratory and turnover costs in the stem also increases. These
finding matches well known empirical patterns, where both \(\phi\) and
\(\rho\) have been linked to shade tolerance\citep{poorter_leaf_2006, lusk_ontogenetic_2008,osunkoya_light_1996}.

Conclusion,\ldots{}..

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  traits do effect growth rate, but not in the simple linear fashion
  previously anticipated.
\item
  Relationships change with size due to changes in biomass partitioning
  and reproductive allocation.
\item
  explains why LMA does not show relationship to growth among larger
  plants. Does not mean unimportant for growth, just effects
  concentrated at small sizes.
\item
  in fact, explains plasticity in LMA with size, where LMA tends to
  increase with size \citep{thomas_asymptotic_1999;
  @thomas_photosynthetic_2010}
\item
  Framework can extend to other traits (:

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    In the current model, the only effects of the trait seed size
    (\(s\)) are to influence starting size of seedlings and fecundity of
    the mother. Seed size therefore has no direct influence on absolute
    growth. In may studies seed size has been found to be negatively
    correlated with mass-based relative growth rate of seedlings. This
    effect is recovered here because of the intrinsic decrease in RGR
    with size.
  \item
    leaf N:
  \item
    sapwood area to leaf area
  \end{itemize}
\item
  Take home: Don't look for traits as being fast or slow growth, rather
  consider traits as defining potential trajectory across sizes and
  light environments.
\end{itemize}

\section{Methods}\label{methods}

\subsection{Plant growth model}\label{plant-growth-model}

TREE

\subsection{Data analysis}\label{data-analysis}

Seedling data

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  about data
\item
  method for estimating growth rate
\end{itemize}

Long term plot data

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  about data
\item
  method for estimating growth rate
\end{itemize}

Traits

\newpage

\section{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{figures/BCI_data.pdf}
\caption{\textbf{The relationship between traits and potential growth
rate varies with plant size.} For XXX species from tropical rain forest
in Panama, we estimated the potential growth rate of individual's in
each species at a series of diameters \(D\), indicated along right hand
side. The size of circles in each panel indicates the number of data
points used to estimate potential growth rate. Traits values were
calculated from representative individuals in each species.
\label{f-BCI}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figs/allometry.pdf}
\caption{\textbf{Key assumptions of a functional balance allometric
model, evaluated using global dataset.} We used the biomass and
allometry database to evaluate model assumptions about \textbf{a,}
scaling of leaf area with plant height, \textbf{b} Scaling of sapwood
area with leaf area, and \textbf{c} scaling of root mass with leaf area.
Each dot is a single plant. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship. Colours indicate vegetation
type. Dashed black lines show values expected under functional-balance
assumption (see Supplementary text for details). \label{f-assumptions}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/growth_light.pdf}
\caption{\textbf{Traits moderate the responsiveness of growth to changes
in light environment.} Panels show predicted relationship between
specific trait and diameter growth rate, for a plant of specified
diameter and under a range of shading environments.
\label{f-growth_light}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/max_leaf_above.pdf}
\caption{\textbf{Low construction cost leads to shade intolerance,
because of costs of high turnover.} Panels show effect of traits on
maximum amount of shading that can be endured before net production (eq.
\ref{eq:dPdt}) reaches zero. Lines indicate relationship for plants of a
given height. \label{f-wplcp}}
\end{figure}

\newpage

\section{Tables}\label{tables}

\begin{table}[h]
\caption{Theoretical predictions on the relationship between traits and demography}

{\footnotesize
\centering
\begin{tabular}{p{4cm}p{4cm}p{2cm}p{6cm}}
\\
\multicolumn{4}{l}{\textbf{a) Intrinsic changes in demography with size}} \\ \\
\hline
Variable & Change with size & & Support \\ \hline
Height growth & strongly hump-shaped & & Fig. \ref{f-hump} \\
Diameter growth & weakly hump-shaped & & Fig. \ref{f-SI_size_dDdt}\\
Basal area growth & increases & & Fig. \ref{f-SI_size_dastdt}\\
Live mass growth & ?? & & ?? \\
ABG mass growth & increases & & ??\\
Relative growth (any) & decreases & & ??\\
Shade tolerance & decreases & & Fig. \ref{f-wplcp}\\ \\
\end{tabular}

\begin{tabular}{p{4cm}p{2.5cm}p{3.5cm}p{6cm}}
\multicolumn{4}{l}{\textbf{b) Relationship of demography to traits}} \\ \\
\hline
Variable & When small & Change with size & Support \\ \hline
\multicolumn{4}{l}{\emph{Leaf-construction cost}} \\

Height growth & negative & flattens then reverses  & Fig. \ref{f-lma_growth_size}\\
Diameter growth & negative & flattens & Fig. \ref{f-growth_light}, \ref{f-lma_growth_size}\\
Basal area growth & negative & flattens then reverses & Fig. \ref{f-lma_growth_size}\\
Live mass growth & ??  & & \\
ABG mass growth & negative & same & Fig. \ref{f-lma_growth_size} \\
Relative growth (any) & negative & weakens & Fig. \ref{f-lma_growth_size_relative}\\
Shade tolerance & positive &  strengthens & Fig. \ref{f-wplcp}\\ \\

\multicolumn{4}{l}{\emph{Wood-construction cost}} \\

Height growth & negative & strengthens  & Fig. \ref{f-rho_growth_size}\\
Diameter growth & negative & strengthens & Fig. \ref{f-growth_light}, \ref{f-rho_growth_size} \\
Basal area growth & negative & strengthens & Fig. \ref{f-rho_growth_size} \\
Live mass growth & ??  & & \\
ABG mass growth & negative & strengthens & Fig. \ref{f-rho_growth_size}\\
Relative growth (any) & negative & weakens & Fig. \ref{f-rho_growth_size_relative}\\
Shade tolerance & positive &  strengthens & Fig. \ref{f-wplcp}\\ \\

\multicolumn{4}{l}{\emph{Height at maturation}} \\
Any growth measure & none & becomes positive  & Fig. \ref{f-hmat_growth_size}\\
Relative growth (any) & none & becomes positive  & Fig. \ref{f-hmat_growth_size_relative}\\
Shade tolerance & none &  none &  \\  \\ \hline
\end{tabular}
}
\label{tab:predictions}
\end{table}

\newpage

\begin{table}[h]
\caption{Effect of traits and related trade-offs on demography}
\begin{tabular}[c]{l|ccccc|c|c|c}
\multicolumn{9}{c}{}\\ \hline
& \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
& \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality}& \textbf{Fecundity}& \textbf{Birth}\\
& $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$
& $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$
& $\frac{\textrm{d}P}{\textrm{d}t}$
& $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
& $\frac{1}{k_s}$ & & & $h_0$ \\\hline
\textbf{This study}&&&&&&&\\
Leaf-construction cost, $\phi$ & $\downarrow$ & $\uparrow$ & & & & & \\
Wood-construction cost, $\rho$ & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$ & \\
Height at maturation, $h_m$ & &$\uparrow$ & & & & & $\downarrow$ & \\
&&&&&&&\\\hline
\textbf{Possible extensions}&&&&&&&\\
Leaf area per sapwood area, $\theta$ & $\uparrow$& & $\downarrow$ & $\downarrow$ & & &\\
Leaf-nitrogen & &$\downarrow$$\uparrow$ & & & & & \\
Seed size & & & & & & & $\downarrow$ & $\uparrow$\\ \hline
\end{tabular}
\label{tab:trade-offs}
\end{table}

\newpage

\section{Supplementary material}\label{supplementary-material}

\subsection{Derivation of a simple allometric model of plant
function}\label{derivation-of-a-simple-allometric-model-of-plant-function}

Here we describe an allometric model linking the various size dimensions
of a plant required by most ecologically realistic vegetation models
(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
roots) to a plant height. Table \ref{tab:allometry} provides a summary
of the derived model, while Table \ref{tab:params} provdies estimates on
key parameters.

\begin{table}[h]
\caption{Equations for an allometric growth model}
\centering

\begin{tabular}{p{5cm}p{5cm}p{5cm} }
\\ \hline
Variable & Function & Derivative\\ \hline
leaf area & $a_\textrm{l}=\alpha_1 \, h^{\beta_1}$ & $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}= -\beta_1\big(\frac{a_\textrm{l}}{\alpha_1}\big)^{-(\beta_1+1)}$\\
leaf mass & $m_\textrm{l}=\phi \, a_\textrm{l} $ & $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}=\phi$\\
sapwoood area & $a_\textrm{ss}=\theta^{-1} \, a_\textrm{l}$ & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} =\theta^{-1} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$\\
sapwood mass&$m_\textrm{ss}=\theta^{-1} \, \rho \, \eta_c \, a_\textrm{l} \, h $ & $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}}=\theta^{-1}\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$\\
bark area & $a_\textrm{sb}=b \, \theta^{-1} \, da_\textrm{l}$ & $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t}=b \, \theta^{-1} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$\\
bark mass&$m_\textrm{b}=b\, \theta^{-1} \, \rho \, \eta_c \, a_\textrm{l} \, h $ & $\frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}}=b \, \theta^{-1} \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}A} \big)$\\
heartwood area & $a_\textrm{sh}=\int_0^t \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ & $\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}=k_\textrm{s} \, a_\textrm{ss}$\\
heartwood mass & $m_\textrm{sh}=\int_0^t \frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ & $\frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}=k_\textrm{s} \, m_\textrm{ss}$\\
root mass & $m_\textrm{r}=\alpha_3 \, a_\textrm{l}$ & $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}= \alpha_3$ \\\hline
\end{tabular}
\label{tab:allometry}
\end{table}

\subsubsection{Leaf area}\label{leaf-area}

Based on empirically observed allometries (see main text), we assume an
allometric log-log scaling relationship between the accumulated leaf
area of a plant and its height:

\begin{equation}\label{eq:ha}
a_\textrm{l}=\alpha_1 \, h^{\beta_1}.
\end{equation}

Note, scaling relationship reversed from \citep{falster_influence_2011}.

\subsubsection{Mass of sapwood}\label{mass-of-sapwood}

We follow the model of\citep{yokozawa_foliage_1995} describing the
vertical distribution of leaf area within the crowns of individual
plants. This model can account for a variety of canopy profiles through
a single parameter \(\eta\). Setting \(\eta=1\) results in a conical
canopy, as seen in many conifers, while higher values, e.g. \(\eta=12\)
, give a top-weighted canopy profile similar to those seen among
angiosperms. Let \(S(z,h)\) be the sapwood area at height \(z\) for a
plant with top height \(h\). Following Yokozawa and Hara (1995) we
assume a relationship between \(S(z,h)\) and height such that

\begin{equation}\label{eq:crown1}
\frac{S(z,h)}{S(0,h)}= \big(1-\big(\frac{z}{h}\big)^\eta\big)^2.
\end{equation}

We also assume that each unit of sapwood area supports a fixed area of
leaf (the pipe model, \citep{shinozaki_quantitative_1964}), so that the total
canopy area of a plant relates to basal sapwood area \(S(0,h)\):

\begin{equation}\label{eq:crown2}
\frac{m_\textrm{l}}{\phi}= \theta \, S(0,h).
\end{equation}

Integrating \(S(z,h)\) gives a solution for the total mass of sapwood in
the plant:

\begin{equation}\label{eq:ms1}
m_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, \textrm{d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}

where
\(\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}\)\citep{yokozawa_foliage_1995}.
Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
expression for sapwood mass as a function leaf area and height:

\begin{equation}\label{eq:ms2}
m_\textrm{s}=\rho \, \eta_c \, \theta^{-1} \, a_\textrm{l} \, h.
\end{equation}

\subsubsection{Bark mass}\label{bark-mass}

Bark and phloem tissue are modelled using an analogue of the pipe model,
leading to a similar equation as that for sapwood mass (eq.
\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
assumed to be a constant fraction \(b\) of sapwood area per unit leaf
area such that

\begin{equation}\label{eq:mb}
m_\textrm{b}=b m_\textrm{s}.
\end{equation}

\subsubsection{Root mass}\label{root-mass}

Also consistent with pipe-model assumption, we assume a fixed ratio of
root mass per unit leaf area

\begin{equation}\label{eq:mr}
m_\textrm{r}=\alpha_3 \, a_\textrm{l}.
\end{equation}

Even though nitrogen and water uptake are not modelled explicitly,
imposing a fixed ratio of root mass to leaf area ensures that
approximate costs of root production are included in calculations of
carbon budget.

\newpage

\begin{table}[h]
\caption{Model parameters}
\centering

\include{figures/table-pars}

\label{tab:params}
\end{table}

\subsection{Trait values maximising height
growth}\label{trait-values-maximising-height-growth}

We want to find the trait values maximising growth rate, \(G\). To make
the analysis more tractible, we will focus on height growth rate and
assume we are dealing with a plant of given height where 100\% of
available energy is allocated to growth. From eq. \ref{eq:dhdt}, we thus
have

\begin{equation} \label{eq:G1}
G = \frac{\textrm{d}h}{\textrm{d}t} = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}
\times \frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
\times \frac{\textrm{d}m_\textrm{t}} {\textrm{d}P}
\times \frac{\textrm{d}P}{\textrm{d}t}.
\end{equation}

Maximising \(G\) is equivalent to solving for trait values \(x\) giving
\(\partial G /\partial x = 0\). Noting that traits do not influence
\(\frac{\textrm{d}h} {\textrm{d}a_\textrm{l}}\), and by assumption
\(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}=1\), we thus have

\begin{equation} \label{eq:G2}
G =  c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}  \frac{ \textrm{d}P} {\textrm{d}t} \right),
\end{equation}

where \(c_1 = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}\). Eq.
\ref{eq:G2} is a product of the form \(Y = W(x) \times Z(x)\). For,
equations of this type, a solution to \(\partial{Y}/\partial{x} =0\) is
given by \(\partial{W}/\partial{x} / W = \partial{Z}/\partial{x} / Z\).
Maximum growth rate thus occurs when

\begin{equation} \label{eq:G3}
\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial x} \frac{1}{\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}} = \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} \frac{1}{\frac{ \textrm{d}P} {\textrm{d}t} }.
\end{equation}

In words, the maximum occurs when relative change in marginal leaf
deployment with respect to trait is equal to the relative change in mass
production with respect to the trait.

Let us now try and simplify equation \ref{eq:G3}. First, let our trait
of interest be leaf-construction cost, i.e. \(x=\phi\). From eq/
\ref{eq:daldmt}, we can derive

\begin{equation} \label{eq:G4}
\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial \phi} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\ \right)^2.
\end{equation}

This simplifies eq. \ref{eq:G3} to

\begin{equation} \label{eq:G5}
\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}(x) =  \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} \frac{1}{\frac{ \textrm{d}P} {\textrm{d}t} }.
\end{equation}

From eq. \ref{eq:dPdt} we obtain,

\begin{equation}\label{eq:dPdt2}
\frac{dP}{\textrm{d}t} = c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}
\end{equation}

where
\(c_2 = Y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, k_\textrm{i}})\)
is independt of \(\phi\), and thus

\begin{equation}\label{eq:dPdt3}
\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi}  = - (1-B_4) a_\textrm{l} a_4\phi ^{-B_4}.
\end{equation}

Combing eqs. \ref{eq:G5}-\ref{eq:dPdt3}, we find that the maximum occurs
when

\begin{equation} \label{eq:G6}
\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}(\phi) = \frac{- (1-B_4) a_\textrm{l} a_4\phi ^{-B_4}}{c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}}.
\end{equation}

Note also that if \(B4=1\),
\(\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi} =0\),
also if \(B4>1\),
\(\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi} > 0\).
A similar derivation can be produced linking wood density to
productivity, via sapwood turnover.

\subsection{Supplementary figures}\label{supplementary-figures}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_quantile_examples.pdf}
\caption{\textbf{Estimating potential growth rate via quantile
regression.} Example for 4 species. Growth rate estimated at six sizes,
by fitting lines within given size intervals. Lines fitted to 99th
quantile, but only when n \textgreater{} 500. Except for left-hand
panel, estimated points in centre of interval.
\label{fS-quantile_examples}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_tradeoff.pdf}
\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
Data from \citep{wright_world-wide_2004} for 678 species from 51 sites, each
point giving a species-average. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship.\label{fS-leaf}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_size_dhdt.pdf}
\caption{\textbf{Hump-shaped relationship between height growth rate and
size.} \label{f-hump}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_size_dDdt.pdf}
\caption{\textbf{Hump-shaped relationship diameter growth and size.}
\label{f-SI_size_dDdt}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_size_dDdt.pdf}
\caption{\textbf{Basal area growth rate increases with size.}
\label{f-SI_size_dastdt}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_mass_fraction.pdf}
\caption{\textbf{Change in allocation with size.}
\label{f-mass_fraction}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
\caption{\textbf{The expected correlation between leaf-construction cost
and growth rate changes with plant size.} \label{f-lma_growth_size}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_effects_at_diameters_relative.pdf}
\caption{\textbf{The expected correlation between leaf-construction cost
and relative growth rate changes with plant size.}
\label{f-lma_growth_size_relative}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
\caption{\textbf{The expected correlation between stem-construction cost
and growth rate changes with plant size.} \label{f-rho_growth_size}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_rho_effects_at_diameters_relative.pdf}
\caption{\textbf{The expected correlation between stem-construction cost
and relative growth rate changes with plant size.}
\label{f-rho_growth_size_relative}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_hmat_effects_at_diameters.pdf}
\caption{\textbf{The expected correlation between maximum height and
growth rate changes with plant size.} \label{f-rho_growth_size}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_hmat_effects_at_diameters_relative.pdf}
\caption{\textbf{The expected correlation between maximum height and
relative growth rate changes with plant size.}
\label{f-hmat_growth_size_relative}}
\end{figure}

\newpage

\bibliography{references}
\end{document}

### 2015-02-04 R/ms-fun.R

@@ -1,7 +1,7 @@
 
 pandoc_build <- function(file){
   args <- list('--template=include.tex', '--latex-engine=xelatex')
-  pandoc_convert(basename(file), output= paste0(tools::file_path_sans_ext(basename(file)), ".pdf"), citeproc = TRUE, options = args, verbose = FALSE, wd = dirname(file))
+  pandoc_convert(basename(file), to="latex", output= paste0(tools::file_path_sans_ext(basename(file)), ".tex"), citeproc = FALSE, options = args, verbose = FALSE, wd = dirname(file))
 }
 
 get_nature_csl <- function(dest, url="https://raw.githubusercontent.com/citation-style-language/styles/master/nature.csl"){
@@ -10,4 +10,15 @@ get_nature_csl <- function(dest, url="https://raw.githubusercontent.com/citation
 
 get_amnat_csl <- function(dest, url="https://raw.githubusercontent.com/citation-style-language/styles/master/the-american-naturalist.csl"){
 	cat(getURL(url), file=dest)
+}
+
+tex_2_pdf <- function(texfile){
+
+  filename <- tools::file_path_sans_ext(texfile)
+  system(sprintf("xelatex %s", texfile))
+  system(sprintf("bibtex %s",   filename))
+  system(sprintf("xelatex %s", texfile))
+  system(sprintf("xelatex %s", texfile))
+  aux.files <- paste0(filename, c(".log", ".aux", ".bbl", ".blg"))
+  file.remove(aux.files[file.exists(aux.files)])
 }

### 2015-02-06 ms.tex
Equations explaining shift in correlation with size. Towards #25.

@@
 
-\subsection{Trait values maximising height
-growth}\label{trait-values-maximising-height-growth}
+\subsection{Effect of traits on height growth} \label{trait-values-maximising-height-growth}
 
-We want to find the trait values maximising growth rate, \(G\). To make
-the analysis more tractible, we will focus on height growth rate and
-assume we are dealing with a plant of given height where 100\% of
-available energy is allocated to growth. From eq. \ref{eq:dhdt}, we thus
-have
-
-\begin{equation} \label{eq:G1}
-G = \frac{\textrm{d}h}{\textrm{d}t} = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}
-\times \frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
-\times \frac{\textrm{d}m_\textrm{t}} {\textrm{d}P}
-\times \frac{\textrm{d}P}{\textrm{d}t}.
-\end{equation}
-
-Maximising \(G\) is equivalent to solving for trait values \(x\) giving
-\(\partial G /\partial x = 0\). Noting that traits do not influence
-\(\frac{\textrm{d}h} {\textrm{d}a_\textrm{l}}\), and by assumption
-\(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}=1\), we thus have
+We want to show how two traits $\phi$ and $\rho$ influence growth rate,
+by looking at the effect of
+traits on the components of growth rate (eq. \ref{eq:dhdt}. To make
+the analysis more tractable, we will assume we are dealing with a plant of
+given height where 100\% of available energy is allocated to growth,
+i.e. \(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}=1\). Note also
+that by the assumptions of our model and above, traits do not influence
+\(\frac{\textrm{d}h} {\textrm{d}a_\textrm{l}}\). Thus we have
 
 \begin{equation} \label{eq:G2}
-G =  c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}  \frac{ \textrm{d}P} {\textrm{d}t} \right),
+G = \frac{\textrm{d}h}{\textrm{d}t} = c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}  \frac{ \textrm{d}P} {\textrm{d}t} \right),
 \end{equation}
@@
 where \(c_1 = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}\). Eq.
-\ref{eq:G2} is a product of the form \(Y = W(x) \times Z(x)\). For,
-equations of this type, a solution to \(\partial{Y}/\partial{x} =0\) is
-given by \(\partial{W}/\partial{x} / W = \partial{Z}/\partial{x} / Z\).
-Maximum growth rate thus occurs when
+\ref{eq:G2} is a product of the form \(Y(x) = W(x) \times Z(x)\). For,
+equations of this type, the derivative \(\partial{Y}/\partial{x}\) is
+given by \(\partial{W}/\partial{x} \, Z + W \, \partial{Z}/\partial{x}\). Thus
+the derivative of \(G\) with respect to some trait $x$  is given by
 
-\begin{equation} \label{eq:G3}
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial x} \frac{1}{\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}} = \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} \frac{1}{\frac{ \textrm{d}P} {\textrm{d}t} }.
+\begin{equation} \label{eq:dG}
+\frac{\partial G} {\partial x} =
+\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial x}
+ \, \frac{ \textrm{d}P} {\textrm{d}t}
++ \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\, \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}.
 \end{equation}
 
-In words, the maximum occurs when relative change in marginal leaf
-deployment with respect to trait is equal to the relative change in mass
-production with respect to the trait.
+If this derivative is positive (negative), there will be a positive (negative) relationship
+between traits and growth at the height where it is evaluated.
 
-Let us now try and simplify equation \ref{eq:G3}. First, let our trait
-of interest be leaf-construction cost, i.e. \(x=\phi\). From eq/
-\ref{eq:daldmt}, we can derive
+We further simplify by expanding the derivative of eq. \ref{eq:daldmt}:
 
-\begin{equation} \label{eq:G4}
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial \phi} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\ \right)^2.
+\begin{equation} \label{eq:da_dx}
+\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}
+{\partial x} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\ \right)^2
+\, \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}.
 \end{equation}
 
-This simplifies eq. \ref{eq:G3} to
+Combing eqs. \ref{eq:dG}-\ref{eq:da_dx}, we have
 
-\begin{equation} \label{eq:G5}
-\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}(x) =  \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} \frac{1}{\frac{ \textrm{d}P} {\textrm{d}t} }.
+\begin{equation} \label{eq:dG2}
+\frac{\partial G} {\partial x} =
+\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\left(
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
+- \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}
+ \, \frac{ \textrm{d}P} {\textrm{d}t}
+\right).
 \end{equation}
 
-From eq. \ref{eq:dPdt} we obtain,
+Since is $\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}>0$ always, the sign
+of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
+RHS. Rearranging we find that $\frac
+{\partial G} {\partial x} < 0$ if
+
+\begin{equation}\label{eq:dg3}
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
+< \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x} \, \frac{ \textrm{d}P} {\textrm{d}t},
+\end{equation}
+
+or equivalently
+\begin{equation}\label{eq:dg4}
+\frac{
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
+{\frac{ \textrm{d}P} {\textrm{d}t}}
+<
+\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
+\end{equation}
+
+In words, there will be a negative relationship between trait x and growth if the
+relative change in production due to change in traits is less than the relative change
+in the total construction cost for a unit of leaf area due to change in trait. Crucially,
+key elements in eq. \ref{eq:dg3} are modified by height.
+
+We can also solve for trait values \(x\) maximising growth rate, i.e. the value
+giving \(\partial G /\partial x = 0\):
+
+\begin{equation}\label{eq:max}
+\frac{
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
+{\frac{ \textrm{d}P} {\textrm{d}t}}
+=
+\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
+\end{equation}
+
+\subsubsection{Leaf construction cost}
+
+If the trait of interest is leaf-construction cost, i.e. \(x=\phi\), we can expand
+eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{\textrm{d}m_\textrm
+{t}} {\textrm{d}a_\textrm{l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dPdt} we obtain,
 
@@
 \(c_2 = Y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, k_\textrm{i}})\)
-is independt of \(\phi\), and thus
+is independent of \(\phi\), and thus
 
 \begin{equation}\label{eq:dPdt3}
-\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi}  = - (1-B_4) a_\textrm{l} a_4\phi ^{-B_4}.
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi}  =
+(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}.
 \end{equation}
 
-Combing eqs. \ref{eq:G5}-\ref{eq:dPdt3}, we find that the maximum occurs
-when
+Eq. \ref{eq:dPdt3} shows that if \(B4=1\), \(\frac{ \textrm{d}P} {\textrm{d}t}\) is independent
+of $\phi$. In contrast, if \(B4>1\) (which tends to be the case empirically), \(\frac{ \textrm
+{d}P} {\textrm{d}t}\) increases as $\phi$ increases, more acutely at low $\phi$.
+
+Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
+with traits if
 
 \begin{equation} \label{eq:G6}
-\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}(\phi) = \frac{- (1-B_4) a_\textrm{l} a_4\phi ^{-B_4}}{c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}}.
+\frac{(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}}{c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}}
+< \frac{1}{\phi
+ + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm
+ {b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}}.
 \end{equation}
 
-Note also that if \(B4=1\),
-\(\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi} =0\),
-also if \(B4>1\),
-\(\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi} > 0\).
-A similar derivation can be produced linking wood density to
-productivity, via sapwood turnover.
+\subsubsection{Wood construction cost}
+
+Do similar to previous section for wood construction cost.
 

### 2015-02-06 R/BCI.R
3-panel version of the BCI figure

@@ -198,6 +198,71 @@ figure_BCI_data <- function(data) {
   mtext(name_pretty("dbasal_diam_dt"), 2, line=3, outer=TRUE, cex=cex_lab)
 }
 
+figure_BCI_data2 <- function(data) {
+  traits <- c("lma", "rho", "hmat")
+  at <- c(10, 20, 50, 100, 200)/1000
+
+  data <- data[data$count > 500 & data$at %in% at,]
+  data$class <- match(data$at, at)
+
+  ## Three types if fits: independent (_i), weighted (_w) and joint
+  ## (_j).
+  prep <- function(trait) {
+    data$x <- data[[trait]]
+    dsplit <- split(data, data$class)
+    fits_i <- lapply(dsplit, function(d)
+                     lm(log10(dbasal_diam_dt) ~ log10(x), d))
+
+    ## Next thing to try is weights.
+    fits_w <- lapply(dsplit, function(d)
+                     lm(log10(dbasal_diam_dt) ~ log10(x), d,
+                        weights=log(count)))
+
+    fits_j <- lm(log10(dbasal_diam_dt) ~ log10(x) * log(at), data,
+                 weights=log(count))
+
+    ## Generate a data set to do prediction with:
+    xr <- sapply(dsplit, function(x) range(x[[trait]], na.rm=TRUE))
+
+    yr_i <- sapply(seq_along(dsplit), function(i)
+                   10^(predict(fits_i[[i]], list(x=xr[,i]))))
+    yr_w <- sapply(seq_along(dsplit), function(i)
+                   10^(predict(fits_w[[i]], list(x=xr[,i]))))
+    yr_j <- sapply(seq_along(dsplit), function(i)
+                   10^(predict(fits_j, data.frame(x=xr[,i], at=at[i]))))
+
+    list(data=data, xr=xr, yr_i=yr_i, yr_w=yr_w, yr_j=yr_j)
+  }
+
+  dat <- setNames(lapply(traits, prep), traits)
+
+  ylim <- c(0.5, 32) / 1000
+  xlim <- list(lma=c(0.02, 0.18), rho=c(250, 1000), hmat=c(5,50))
+  cols <- RColorBrewer::brewer.pal(length(at), "Set1")
[…]

### 2015-02-06 fitting.Rmd
Update from maker -> remake

@@
 
-m <- maker::maker()
-data <- m$make("BCI_species_data")
+data <- remake::make("BCI_species_data")
 head(data)
@@
 ```{r, echo=TRUE, results="hide"}
-data <- m$get("BCI_species_data") %>%
+data <- remake::make("BCI_species_data") %>%
   mutate(dbh=at) %>%

### 2015-02-06 sensitivity.R

@@ -14,8 +14,7 @@ suppressPackageStartupMessages({
 })
 
 # First, look at plot: challenge is to move peak to left and pull down elevation of curve.
-m <- maker::maker()
-data <- m$make("BCI_species_data") %>%
+data <- remake::make("BCI_species_data") %>%
   mutate(dbh=at) %>%
   filter(!is.na(lma*rho*hmat*dbh*dbasal_diam_dt), dbh %in% c(0.01), hmat > 10) %>%
   select(dbh, lma, rho, dbasal_diam_dt)

### 2015-02-18 R/ms-fun.R
Update latex build to use latexmk

@@ -1,24 +1,8 @@
-
-pandoc_build <- function(file){
-  args <- list('--template=include.tex', '--latex-engine=xelatex')
-  pandoc_convert(basename(file), to="latex", output= paste0(tools::file_path_sans_ext(basename(file)), ".tex"), citeproc = FALSE, options = args, verbose = FALSE, wd = dirname(file))
-}
-
-get_nature_csl <- function(dest, url="https://raw.githubusercontent.com/citation-style-language/styles/master/nature.csl"){
-	cat(getURL(url), file=dest)
-}
-
-get_amnat_csl <- function(dest, url="https://raw.githubusercontent.com/citation-style-language/styles/master/the-american-naturalist.csl"){
-	cat(getURL(url), file=dest)
-}
-
 tex_2_pdf <- function(texfile){
 
-  filename <- tools::file_path_sans_ext(texfile)
-  system(sprintf("xelatex %s", texfile))
-  system(sprintf("bibtex %s",   filename))
-  system(sprintf("xelatex %s", texfile))
-  system(sprintf("xelatex %s", texfile))
-  aux.files <- paste0(filename, c(".log", ".aux", ".bbl", ".blg"))
+  system(sprintf('latexmk -pdf -pdflatex="xelatex -interaction=nonstopmode"
+                 %s', texfile))
+  aux.files <- paste0(tools::file_path_sans_ext(texfile),
+                     c(".log", ".aux", ".bbl", ".blg"))
   file.remove(aux.files[file.exists(aux.files)])
 }

### 2015-02-18 ms.tex
Move from short to medium format paper

--- 2015-02-04 ms.tex
@@
-% THIS FILE TAKEN FROM PANDOC DEMOS PAGE:http://johnmacfarlane.net/pandoc/demos.html,http://johnmacfarlane.net/pandoc/demo/mytemplate.tex
-
 \documentclass[12pt, a4paper]{article}
@@
 \usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-
-\usepackage{fontspec}
-\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+\usepackage{authblk} % allows for better author affiliations
 
@@
 
-% \usepackage{natbib}
-% \bibliographystyle{abbrvnat}
-% \setcitestyle{authoryear,open={(},close={)}}
-
-\usepackage[sort&compress]{natbib}
-\citestyle{nature}
-\bibliographystyle{ieeetr}
+\usepackage{natbib}
+\bibliographystyle{ecol_let}
+\setcitestyle{authoryear,open={(},close={)}}
 
@@
 \usepackage[rgb,dvipsnames]{xcolor}
-\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
 \definecolor{grey}{rgb}{0.5, 0.5, 0.5}
@@
 
-\usepackage{titling}
+\title{\LARGE Untangling the link between traits, size and growth rate in plants}
 
-\title{Untangling the link between traits, size and growth rate in plants}
+\author[1]{Daniel S. Falster}
+\author[1]{Richard G. FitzJohn}
+\author[2]{Joe Wright}
 
-\author{
-{\bf Daniel Falster} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{daniel.falster@mq.edu.au}\vspace{0.1in}\\
-{\bf Rich FitzJohn} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{richard.fitzjohn@mq.edu.au}\vspace{0.1in}\\
-{\bf Joe Wright} \\ \normalsize\emph{Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama} \\ \url{wrightj@si.edu}\vspace{0.1in}\\ }
-
-\date{}
+\affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
+\affil[2]{{\footnotesize Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama}}
+\renewcommand\Authands{ and }
+\date{\vspace{-3em}}
 
 \begin{document}
+
+\maketitle
 \thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that
-{\LARGE \bf \thetitle \vspace{0.2in}}
 
-\theauthor
+\section*{Abstract}\label{abstract}
 
-\section{Abstract}\label{abstract}
+\section*{Introduction}\label{introduction}
 
-Plant growth is the foundation of the terrestrial biosphere, powering
-food-webs, the carbon cycle and much of human
-activity \citep{raupach_carbon_2010}. Plant species differ in many
-functional traits, including aspects of tissue construction and energy
-allocation, that drive differences in mechanical strength, biomass
-partitioning, and rates of photosynthesis and tissue
-turnover\citep{wright_world-wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
-such traits influence whole-plant growth, with the simple relationships
-predicted by existing theory often lacking empirical
-support\citep{wright_functional_2010,poorter_are_2008}. Here we
-provide a new set of theoretical predictions, validated using 20 years
-of growth data\citep{condit_barro_2012,kitajima_leaf_2013}, showing
-that the influence of three prominent traits on growth rate is moderated
-by plant size, and that the impact of size on the nature of
-relationships varies among traits. We extend a widely-used theoretical
-model that links growth rate in seedlings with a single leaf
-trait\citep{lambers_inherent_1992,wright_cross-species_2000} to
-explicitly include influences of size, light environment, and three
-other prominent traits. We find the influence of leaf construction cost
-(\(\phi\)) on growth is strong in small plants but weakens with size. In
-contrast, the influence of stem construction cost (\(\rho\)) peaks at
-intermediate sizes, while the influence of height at maturation
-(\(h_m\)) strengthens with size. Our theoretical model suggests the
-change in trait-growth relationships with size arises due to changes in
-biomass partitioning and reproductive allocation. Moreover, we show how
-traits moderate plant responses to light environment and also determine
-shade tolerance, supporting diverse empirical
-results\citep{ruger_functional_2012,poorter_leaf_2006}. By
-disentangling the effects of plant size, light environment and traits on
-growth rates, our results provide a solid theoretical foundation for
-trait ecology and thus provide a platform for understanding growth
-across diverse species around the world.
+Functional traits are thought to capture core differences in the strategies
+plants use to generate and invest surplus energy
+\citep{wright_world-wide_2004,chave_towards_2009,westoby_plant_2002}. Useful
+traits are those that
+can be easily measured for diverse species and where differences between
+species are large compared to intra-specific variation. Moroever, the uptake
+of traits reflects the observation that although most plants have same basic
+construction, physiological function and resource requirements, species differ
+considerably from one another in rates of carbon, nitrogen and water uptake,
+growth, mortality and seed production. Data for prominent traits now exists
+for many of the world's 250000 plant species \citep{cornwell_functional_2014}.
+The effect of some traits of multiple aspects of plant function has also been
+well documented, including mechanical strength, biomass partitioning, rates of
+photosynthesis and rates of tissue turnover
+\citep {wright_world-wide_2004,chave_towards_2009}. As a result, the idea has
+taken hold that the growth startegy and ecology of a plant species might usefully
+be characterised simply by knowing its traits.
 
-\section{Main text}\label{main-text}
+While the influence of plant traits on elements of plant physiologival
+function is increasingly understood, attempts at using traits to predict
+demographic rates have only met with mixed success. In seedlings, the trait
+leaf mass per area (LMA) -- a central part of the leaf economics spectrum --
+is tightly related to plant relative growth rate in mass. LMA and its
+correlate leaf lifespan have also been linked to height growth rate for small
+seedlings and saplings. These early successes prompted researchers to search
+for similar relationships in large plants. However, the results showed that in
+large plants, LMA was not correlated with diameter growth rate. Meanwhile,
+other traits such as wood density showed strong relationships
+to growth in large plants, but only weak relationships in seedlings \citep
+{castro-diez_stem_1998}. To date each relationship has been quantified mostly for
+one or few traits and at a single life-stage. Yet the picture is emerging is that the
+link between traits and growth may be modified by plant size; a finding which would
+challenge the asummption applied by many that particular ends of the trait spectrum
+translate directly into fast or slower growth.
 
-Functional traits capture core differences in the strategies plants use
-to generate and invest surplus energy\citep{wright_world-wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
-traits now exists for many of the world's 250000 plant
-species\citep{cornwell_functional_2014}; yet the effect of these traits
-on growth remains poorly understood. Prominent traits such as \(\phi\)
-(also known as leaf mass per area) are widely considered as indicators
-of growth rate\citep{reich_world-wide_2014}, because cheaper leaf
-construction should enable faster canopy
-expansion\citep{blackman_compound_1919,westoby_time_2000}. Indeed,
-growth rate in seedlings is strongly and negatively correlated with
-\(\phi\)\citep{cornelissen_seedling_1996,wright_cross-species_2000}
-(Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
-might also give a growth advantage in larger
-plants\citep{enquist_general_2007}, an increasing number of empirical
-studies have failed to detect any such
-effect\citep{wright_functional_2010, @herault_functional_2011,
-@poorter_are_2008}. Meanwhile, expectations on how other prominent
-traits might influence growth are lacking. The growing discord between
-theoretical expectation and empirical result thus indicates a strong
-need for a new theoretical framework through which the influence of
-traits and size on plant growth can be understood.
+The challenge of interpreting diverse empirical results linking traits to
+growth rate -- or not -- would clearly be easier if we had clearl expectations
+on what signal one should expect, given our understanding of how traits
+infleunce plant function. This is a primary, yet perhaps underappreciated,
+role for theory: to lay expectations based on mechistic principles against
+which empircal data can be compared. Current empircial results suggest the
+effect of traits on growth changes with size and possibly also light
+environment, yet current theory says little about how relationships might come
+about. A widely-used model suggests mass-based relative growth rate will be a
+linear function of LMA, at least for seeldings.  An extension of model
+suggests a similar relationship may hold at larger sizes. But this prediction
+doesn't empirical results. Meanwhile, theoretical predictions on how other
+traits should influence growth are largely absent.
 
-Here we extend a widely-used theoretical model that links growth rate in
-seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross-species_2000}, to explicitly include influences of
-size, light environment, and other prominent traits. This extension
-allows us to generate novel predictions on the link between four key
-traits and plant demography (Table \ref{tab:predictions}). Predictions
-are compared against data from both a long-term
-plot\citep{condit_barro_2012} and field
-experiment\citep{kitajima_leaf_2013} at Barro Colorado Island, Panama.
-We use a new method to estimate species potential growth rate,
-accounting for size specific effects (Fig. \ref{S-quantile_examples},
-see Methods for details). This new method allowed us to extract a more
-definite signal of trait effects than in previous
-studies\citep{wright_functional_2010}. The results of this new analysis
-(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
-\ref{tab:predictions}) showing how the effect of traits on growth
-changes with plant size, with different traits exhibiting distinct
-size-dependent effects.
+A further problem with existing models is that the effects of traits are all
+realised via mass production, whereas the effect of some prominent traits is
+most likely on the quality of tissue construction, or the amoutn of mass
+allocated to different tissues. There are two problem here. First, measuring mass
+production is only practical for small plants that can be harvested. For larger plants,
+growth is more often emasured either as diameter or height growth. For example, the
+CTFS monitors diameter growth in 49 large permenenat plots worldwide. Establishing
+the link between traits, mass growth and diamater growth is thus essential for predicting
+the effect of traits on growth across these diverse datasets.
 
-We begin with a standard model for the amount of biomass available for
-growth, \(dP / \textrm{d}t\), given by the difference between income
-(total photosynthesis) and losses (respiration and turnover) within the
-plant\citep{falster_influence_2011}:
-
-\begin{equation}\label{eq:dPdt}
-\underbrace{\strut \frac{dP}{\textrm{d}t}}_\textrm{net biomass production}= \underbrace{\strut Y}_\textrm{yield} \big(\underbrace{\strut \frac{m_\textrm{l}}{\phi} \, A}_\textrm{photosynthesis} - \big(\underbrace{\strut \, \frac{m_\textrm{l}}{\phi} \, r_\textrm{l} + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)\big) - \big(\underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}\big).
-\end{equation}
-
-Here, \(m,r\), and \(k\) refer to the mass, respiration rate, and
-turnover rate of different tissues, denoted by subscripts \(l\)=leaves,
-\(b\)=bark, \(s\)=sapwood and \(r\)=roots. \(A\) is the assimilation
-rate of CO\(_2\) per leaf area and \(Y\) is yield: the fraction of
-assimilated carbon fixed in biomass (the remaining fraction being lost
-as growth respiration) (see Table \ref{tab:params} for units and
-definitions). Photosynthesis is proportional to leaf area,
-\(a_\textrm{l} = m_\textrm{l} / \phi\), while total mass of living
-tissues
-\(m_\textrm{t}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.\)
-Eq. \ref{eq:dPdt} assumes that tissues lost via turnover are replaced
-before new growth occurs.
-
-Current theory emphasises the relationship between \(\phi\) and relative
-growth in mass\citep{enquist_general_2007},
-\(\textrm{d}P / \textrm{d}t / m_\textrm{t}\). For seedlings, which are
-young and mostly leaf, it is reasonable to ignore all turnover terms as
-well as the respiration terms for non-leaf tissues. Net production then
-becomes a linear function of leaf area and net photosynthesis per leaf
-area (\(A_\textrm{net} = Y(A - r_\textrm{l})\)), making relative growth
-rate a linear function of \(\phi\):
-
-\begin{equation}\label{eq:RGR}
-\underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}\frac{1}{m_\textrm{t}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{t}}.
-\end{equation}
-
-Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to \(\phi\)\citep{wright_cross-species_2000}, this
-approximation does not directly link to other traits, or to the
-variables that are routinely collected for large trees: namely plant
-height (\(h\)) and stem cross-sectional area (\(a_\textrm{st}\)) or
-diameter \(D\).
-
-To model growth in either height or basal area requires we to account
-not just for mass production, but also for the costs of building new
-tissues, allocation to reproduction, and architectural layout. In eq.
-\ref{eq:RGR}, these terms are missing, whereas in the model presented
-here they are made explicit.
-
-Mathematically, height growth can be decomposed into a product of
-physiologically relevant terms\citep{falster_influence_2011}:
-
-\begin{equation} \label{eq:dhdt}
-\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
-\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}}_\textrm{marginal leaf deployment}
-\times \underbrace{\strut\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}}_\textrm{allocation to growth}
-\times \underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}}_\textrm{mass production}.
-\end{equation}
-
-The first term on the right of eq \ref{eq:dhdt},
-\(\textrm{d}h / \textrm{d}a_\textrm{l}\), is the growth in plant height
-per unit growth in total leaf area; accounting for the architectural
-strategy of the plant. Some species tend to leaf out more than grow
-tall, while other species emphasise vertical
-extension\citep{poorter_architecture_2006}.
-
-The second term, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\),
-accounts for the marginal cost of deploying an additional unit of leaf
-area, including construction of the leaf itself and various support
-structures. As such, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\)
-can itself be expressed as a sum of construction costs per unit leaf
-area:
-
-\begin{equation}\label{eq:daldmt}
-\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
-= \big(\phi
- + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
-\end{equation}
-
-The third term in eq \ref{eq:dhdt},
-\(\textrm{d}m_\textrm{t} / \textrm{d}P\), gives the fraction of net
-biomass production (eq. \ref{eq:dPdt}) that is allocated to growth
-rather than reproduction or storage.
-
-In a similar way, basal area (\(a_\textrm{st}\)) increment can be
-expressed as the sum of increments in sapwood, bark \& heartwood areas
-(\(a_\textrm{ss}, a_\textrm{sb}, a_\textrm{sh}\) respectively):
-\(\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}= \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}\).
-Noting that heartwood grows via turnover of sapwood, such that
-\(\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} a_\textrm{ss}\),
-we have an equation for basal area growth that conatins many of the same
-elements as eq. \ref{eq:dhdt}:
-
-\begin{equation}\label{eq:dast}
-\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
-\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
-\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}} \times \frac{\textrm{d}m_\textrm{t}}{\textrm{d}P} \times \frac{\textrm{d}P}{\textrm{d}t} + k_\textrm{s} a_\textrm{ss}.
-\end{equation}
-
-Diameter growth is then given by the geometric relationship between stem
-diameter (\(D\)) and \(a_\textrm{st}\):
-
-\begin{equation} \label{eq:dDdt}
-\frac{\textrm{d}D}{\textrm{d}t}= \sqrt{\frac{\pi}{a_\textrm{st}}} \frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}.
-\end{equation}
-
-Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
-therefore hold for any model of plant growth. To make explicit
-predictions, however, requires two additional elements: 1) A specific
-model describing how the various mass and area terms
-(\(m_\textrm{l}, m_\textrm{s}, m_\textrm{b}, m_\textrm{r}, a_\textrm{l}, a_\textrm{s}, a_\textrm{b}\))
-vary relative to one another; and 2) For any trade-offs in plant
-function to be specified; this is where traits have an influence. These
-two components are as follows.
-
-To address the first of these challenges, we use a functional balance
-model\citep{yokozawa_foliage_1995,falster_influence_2011} assuming
-1) an allometric relation between total leaf area and plant height, and
-2) a constant ratio between total leaf area and each of sapwood
-cross-sectional area, bark cross-sectional area, and total root mass.
-These assumptions, which are well-supported by data (Fig.
-\ref{f-assumptions}), lead immediately to an allometric model for plant
-size (Table \ref{tab:allometry}, see Supplementary material for
-derivations). Substituting from Table \ref{tab:allometry} into eqs.
-\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
-for plant growth in relation to size.
-
-This new growth model captures the intrinsically size-dependent nature
-of plant growth (Table \ref{tab:predictions}a), recovering well-known
-empirical patterns\citep{sillett_increasing_2010, king_size-related_2011}. Height growth shows a hump-shaped pattern
-with size, first increasing then decreasing. This pattern results from
-systematic changes in the four components of eq. \ref{eq:dhdt} with size
-(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
-that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
-and increasing reproductive allocation. In contrast, basal-area growth
-continues to increase with size\citep{sillett_increasing_2010,
-@stephenson_rate_2014}, due to an increasing influence of stem
-turnover. Diameter growth shows a weakly hump shaped
-curve\citep{herault_functional_2011}, tapering off slightly at larger
-sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
-mainly because of increased reproductive allocation in older trees
-(Fig?). All growth measures decrease sharply with size when expressed as
-relative growth rates\citep{iida_linking_2014}.
-
-We now incorporate the effects of some prominent traits by embedding
-explicit trade-offs within this growth model. Variation in each trait
-implies benefits and costs. These affect specific elements of the growth
-equations presented above (Table \ref{tab:trade-offs}).
-
-By definition, the main effect of increasing \(\phi\) is to decrease the
-marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
-growth rate. However, to fully account for the influence of \(\phi\) on
-growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction\citep{wright_world-wide_2004}.
-This is achieved by linking the turnover rate of leaf in eq.
-\ref{eq:dPdt} to \(\phi\). We use an allometric scaling relationship of
-the form \(k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}\), which has been
-shown to describe patterns across diverse
-ecosystems\citep{wright_world-wide_2004} (Fig. \ref{fS-leaf}).
-
-Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
-\ref{eq:dDdt} predict a relationship between \(\phi\) and growth that
-changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
-Decreasing \(\phi\) has two impacts on growth rate. First, lower
-\(\phi\) increases marginal leaf deployment
-(\(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\) by economising on
-construction costs. Second, lower \(\phi\) decreases net production
-(\(\textrm{d}P / \textrm{d}t\)), due to increased leaf turnover. Whether
-lower \(\phi\) increases growth thus depends on the relative magnitude
-of these two effects. When plants are small the effect on leaf
-deployment rate is bigger and so decreasing \(\phi\) increases growth
-rate. When plants are large, the influence of \(\phi\) on leaf
-deployment rate is diminished, because the costs of building other
-supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
-\ref{f-mass_fraction}). The net result is that at larger sizes, low
-\(\phi\) is no longer advantageous for growth (Fig.
-\ref{f-lma_growth_size}d).
-
-As for leaf, cheaper stem construction (lower \(\rho\)) decreases the
-cost of deploying a unit of leaf area, and may thereby increase growth
-rates. However, the cost of cheap stem construction must also be
-accounted for. However, the direct physiological trade-offs of lower
-\(\rho\) are less understood than for \(\phi\). One possibility is that
-lower \(\rho\) decreases mechanical strength resulting in higher
-mortality\citep{chave_towards_2009,wright_functional_2010}. Under
-this scenario, decreasing \(\rho\) will always provide a growth
-advantage, because the costs of low \(\rho\) are not realised within the
-terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
-\ref{tab:trade-offs}). In addition, lower \(\rho\) may increase turnover
-of sapwood. Mirroring our assumptions for leaves, we let sapwood
-turnover decrease with \(\rho\),
-i.e.\(k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}\), with \(\beta_5 > 1\).
-Under this scenario, decreasing \(\rho\) provides an advantage in height
-growth only when the benefits of low \(\rho\) outweigh the costs. This
-effect is strongest at intermediate sizes. The decline in height growth
-at low \(\rho\) with size is more marked than for either stem basal-area
-and diameter growth, because high sapwood turnover also contributes to
-this growth measure. Thus diameter growth continues, even as net
-production approaches zero.
-
-Unlike leaf and stem traits, height at maturation (\(h_m\)) moderates
-growth by adjusting the amount of energy invested in growth, i.e.~the
-term \(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) in eqs.
-\ref{eq:dhdt} and \ref{eq:dast}. Greater height at maturation (\(h_m\))
-can thus lead to a growth advantage by increasing
-\(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) (Table
-\ref{tab:predictions}). At smaller sizes, there is no differentiation
-among species, because most species are focusing on growth (Fig.
-\ref{f-BCI}). At larger sizes, individuals of some species are
-allocating a majority of their surplus energy to growth, leaving a
-strong signal in potential growth rate (Fig. \ref{f-BCI}).
-
-Two other important predictions arise directly from linking traits to
-growth rate (Table \ref{tab:predictions}c). First, traits moderate the
-responsiveness of growth to changes in light environment. This response
-arises simply because species with higher potential growth rate have
-greater potential plasticity. Recent analyses have shown that species
-with low \(rho\) increase growth more substantially with incraeses in
-light\citep{ruger_functional_2012}. This is exactly the behaviour
-exhibited in our model (Fig. \ref{f-growth_light}), at least for small
-to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
-of low \(\phi\) are lost (Fig. \ref{f-growth_light}). This results also
-agrees with empirical results showing that growth for species with low
-\(rho\) actually decreases as size
-increases\citep{ruger_functional_2012}. In contrast, growth rate tends
-to increase in size for the species with highest
-\(\rho\)\citep{ruger_functional_2012}.
-
-Variation in \(\phi\) also moderates the response of growth to changes
-in light, with species having the lowest \(rho\) being most responsive.
-However, unlike for \(\rho\), the effect appears only for the smallest
-size classes. Accordingly, empirical analyses have detected such a link
-in small plants but not for plants \textgreater{} 0.01m
-dbh\citep{ruger_functional_2012}.
-
-Second, species with low \(\phi\) and low \(\rho\) are predicted to be
-less shade tolerant (Fig. \ref{f-wplcp}). At low \(\phi\) (\(\rho\)),
-leaf (sapwood) turnover is higher and thus a greater light income is
-needed to offset these costs. As previously suggested
-\citep{givnish_adaptation_1988}, shade-tolerance also decreases with
-height because as size increase, the total amount of energy needed to
-offset respiratory and turnover costs in the stem also increases. These
-finding matches well known empirical patterns, where both \(\phi\) and
-\(\rho\) have been linked to shade tolerance\citep{poorter_leaf_2006, lusk_ontogenetic_2008,osunkoya_light_1996}.
-
-Conclusion,\ldots{}..
+Here we provide a mechanistic framework describing how traits link to
+demography, linking to commonly used metrics such as height and diameter
+growth, combine this with fresh empirical insight using largest
+collation of growth and trait data available, from long-term plot at
+BCI, Panama.
 
@@
 \item
-  traits do effect growth rate, but not in the simple linear fashion
-  previously anticipated.
-\item
-  Relationships change with size due to changes in biomass partitioning
-  and reproductive allocation.
-\item
-  explains why LMA does not show relationship to growth among larger
-  plants. Does not mean unimportant for growth, just effects
-  concentrated at small sizes.
-\item
-  in fact, explains plasticity in LMA with size, where LMA tends to
-  increase with size \citep{thomas_asymptotic_1999;
-  @thomas_photosynthetic_2010}
-\item
-  Framework can extend to other traits (:
+  Shows why effect of traits changes with size,size dependent effect
+  differs among traits.
 
@@
   \item
-    In the current model, the only effects of the trait seed size
-    (\(s\)) are to influence starting size of seedlings and fecundity of
-    the mother. Seed size therefore has no direct influence on absolute
-    growth. In may studies seed size has been found to be negatively
-    correlated with mass-based relative growth rate of seedlings. This
-    effect is recovered here because of the intrinsic decrease in RGR
-    with size.
-  \item
-    leaf N:
-  \item
-    sapwood area to leaf area
+    Contradicts widespread assumption that traits relate to growth in
+    simple linear way
   \end{itemize}
 \item
-  Take home: Don't look for traits as being fast or slow growth, rather
-  consider traits as defining potential trajectory across sizes and
-  light environments.
+  Offers framework for interpreting and understanding empirical results
 \end{itemize}
 
-\section{Methods}\label{methods}
+\section*{Fresh empirical motivation}\label{fresh-empirical-motivation}
 
-\subsection{Plant growth model}\label{plant-growth-model}
+To help motivate our model, we begin with a fresh analysis of growth dynamics
+in the long-term plot at Barro Colorado Island in Panam, where the growth
+rates for over 100 species have been recorrded over the last 30 years. A
+prior analysis by \citet{wright_functional_2010} found a mdoerate effect of
+wood density on growth rate, but almsot no effect of LMA, and only a weak
+effect of a thrid trait maximum plant height (HMAX).
 
-TREE
-
-\subsection{Data analysis}\label{data-analysis}
-
-Seedling data
+Fresh empirical analysis highlights how effect of traits on pot growth
+changes with size.
 
@@
 \item
-  about data
+  Effect of lma stronger than previously thought,
 \item
-  method for estimating growth rate
+  Bridges from seedling studies to adults
 \end{itemize}
 
-Long term plot data
+Results stronger than previously reported.
+
+\section*{Materials and Methods}\label{materials-and-methods}
+
+\subsection*{Conceptual framework}\label{conceptual-framework}
+
+Builds on 2011 paper - highlight what's new from there.
+
+Decomposition of diameter and height growth
+
+Functional-balance Model of growth
 
@@
 \item
-  about data
+  four key assumptions.
 \item
-  method for estimating growth rate
+  simple linear functions non-critical for results of this paper,
+\item
+  explore using an example model, but any model with basic size effects
+  will display same behaviour
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    details about our model ()
+  \item
+    show argument hold intuiatively
+  \end{itemize}
 \end{itemize}
 
-Traits
+Embedding trait-related trade-offs
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  use table with arrows to
+\end{itemize}
+
+Connection to earlier work
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  generalised version
+\item
+  present old model as dp/dt so consistent with ours
+\end{itemize}
+
+\subsection*{Analysis}\label{analysis}
+
+\subsection*{Growth data}\label{growth-data}
+
+\section*{Results}\label{results}
+
+\subsection*{Changes in growth rate with
+size}\label{changes-in-growth-rate-with-size}
+
+\subsection*{How traits influence growth
+rate}\label{how-traits-influence-growth-rate}
+
+\subsubsection*{LMA}\label{lma}
+
+\subsubsection*{Wood density}\label{wood-density}
+
+\subsubsection*{HMax}\label{hmax}
+
+\section*{Discussion}\label{discussion}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  mechanistic model of height and dbh growth which accounts for effects
+  of size, light environment and traits
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    need to account for size long recognized, but not straightforward
+  \item
+    look at growth in wide range of measures: mass, height, dbh
+  \end{itemize}
+\item
+  Compared to previous work, shows effects of traits more about
+  allocation than mass production
+\end{itemize}
+
+\subsection*{Tree growth is more than just
+photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}
+
+Recently realisation that allocation and turnover major areas of
+uncertainty
+
+Model reconciles idea that traits can have negative impact on carbon
+budget, but still have positive influence on growth rates
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Enquist 2007: sla and wd increase mass-production, where as in our
+  theory effects on mass production are negative. (also contradicts
+  earlier model from Enquist 1999, where argued wood density had no
+  effect on mass production)
+\end{itemize}
+
+Assumes low lma and wd used to reduce costs of building leaf and stem,
+ie. total leaf area and stem corss section remains same. In this way
+differs to assumptions of recent work by anten\_role\_2010,
+larjavaara\_rethinking\_2010.
+
+Model is realisation of several ideas that have been around for some
+time.
+
+\subsection*{Towards a global model of tree
+growth}\label{towards-a-global-model-of-tree-growth}
+
+Key patterns explained
+
+Data challenges
+
+Reproductive allocation
+
+\subsection*{Detecting trait signals in
+data}\label{detecting-trait-signals-in-data}
+
+Need more than simply look for correlation with mean growth rate
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  understand influences of size on trait of interest
+\item
+  traits define potential growth rate --\textgreater{} how to extract
+  this
+\item
+  ideally start to use mechanistic models
+\end{itemize}
+
+\subsection*{Implications for trait-based
+approaches}\label{implications-for-trait-based-approaches}
+
+How do comparative people interpret sla now - beyond fast slow.
+
+Think about traits as defining potential trajectory, rather than making
+species fast or slow growing per se.
+
+Selection on traits in different parts of lifecycle
+
+Trait plasticity - can explain, no clear advantage of low lma at large
+sizes
+
+Framework for generalising about effect of traits
+
+\section*{Conclusion}\label{conclusion}
 
@@
 
-\section{Figures}\label{figures}
+\section*{Tables}\label{tables}
+
+\textbf{Table 1 - Key equations of growth model}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  key parameters of model
+\end{itemize}
+
+\textbf{Table 2 - effect of traits on demography}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  combines previous tables 1 \& 2
+\end{itemize}
+
+\newpage
+
+\section*{Figures}\label{figures}
 
@@
 \centering
-\includegraphics{figures/BCI_data.pdf}
+\includegraphics{figures/BCI_data2.pdf}
 \caption{\textbf{The relationship between traits and potential growth
@@
 \centering
+\includegraphics{figures/SI_size_dhdt.pdf}
+\caption{\textbf{Conceptual framework for understanding size-dependent
+changes in growth rate.} \label{f-conceptual}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{figs/allometry.pdf}
-\caption{\textbf{Key assumptions of a functional balance allometric
-model, evaluated using global dataset.} We used the biomass and
+\caption{\textbf{Key assumptions of a functional balance and trait
+trade-offs model, evaluated using global dataset.} We used the biomass and
 allometry database to evaluate model assumptions about \textbf{a,}
@@
 \includegraphics{figures/max_leaf_above.pdf}
+\caption{\textbf{The effect of traits on growth changes with size and
+light environment.} Traits moderate the responsiveness of growth to changes in light
+environmentThis figure needs to show more of key result with respect to to size.
+Possibly separate figure. \label{f-shifts}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/max_leaf_above.pdf}
 \caption{\textbf{Low construction cost leads to shade intolerance,
@@
 
-\section{Tables}\label{tables}
-
-\begin{table}[h]
-\caption{Theoretical predictions on the relationship between traits and demography}
-
-{\footnotesize
-\centering
-\begin{tabular}{p{4cm}p{4cm}p{2cm}p{6cm}}
-\\
-\multicolumn{4}{l}{\textbf{a) Intrinsic changes in demography with size}} \\ \\
-\hline
-Variable & Change with size & & Support \\ \hline
-Height growth & strongly hump-shaped & & Fig. \ref{f-hump} \\
-Diameter growth & weakly hump-shaped & & Fig. \ref{f-SI_size_dDdt}\\
-Basal area growth & increases & & Fig. \ref{f-SI_size_dastdt}\\
-Live mass growth & ?? & & ?? \\
-ABG mass growth & increases & & ??\\
-Relative growth (any) & decreases & & ??\\
-Shade tolerance & decreases & & Fig. \ref{f-wplcp}\\ \\
-\end{tabular}
-
-\begin{tabular}{p{4cm}p{2.5cm}p{3.5cm}p{6cm}}
-\multicolumn{4}{l}{\textbf{b) Relationship of demography to traits}} \\ \\
-\hline
-Variable & When small & Change with size & Support \\ \hline
-\multicolumn{4}{l}{\emph{Leaf-construction cost}} \\
-
-Height growth & negative & flattens then reverses  & Fig. \ref{f-lma_growth_size}\\
-Diameter growth & negative & flattens & Fig. \ref{f-growth_light}, \ref{f-lma_growth_size}\\
-Basal area growth & negative & flattens then reverses & Fig. \ref{f-lma_growth_size}\\
-Live mass growth & ??  & & \\
-ABG mass growth & negative & same & Fig. \ref{f-lma_growth_size} \\
-Relative growth (any) & negative & weakens & Fig. \ref{f-lma_growth_size_relative}\\
-Shade tolerance & positive &  strengthens & Fig. \ref{f-wplcp}\\ \\
-
-\multicolumn{4}{l}{\emph{Wood-construction cost}} \\
-
-Height growth & negative & strengthens  & Fig. \ref{f-rho_growth_size}\\
-Diameter growth & negative & strengthens & Fig. \ref{f-growth_light}, \ref{f-rho_growth_size} \\
-Basal area growth & negative & strengthens & Fig. \ref{f-rho_growth_size} \\
-Live mass growth & ??  & & \\
-ABG mass growth & negative & strengthens & Fig. \ref{f-rho_growth_size}\\
-Relative growth (any) & negative & weakens & Fig. \ref{f-rho_growth_size_relative}\\
-Shade tolerance & positive &  strengthens & Fig. \ref{f-wplcp}\\ \\
-
-\multicolumn{4}{l}{\emph{Height at maturation}} \\
-Any growth measure & none & becomes positive  & Fig. \ref{f-hmat_growth_size}\\
-Relative growth (any) & none & becomes positive  & Fig. \ref{f-hmat_growth_size_relative}\\
-Shade tolerance & none &  none &  \\  \\ \hline
-\end{tabular}
-}
-\label{tab:predictions}
-\end{table}
-
-\newpage
-
-\begin{table}[h]
-\caption{Effect of traits and related trade-offs on demography}
-\begin{tabular}[c]{l|ccccc|c|c|c}
-\multicolumn{9}{c}{}\\ \hline
-& \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
-& \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality}& \textbf{Fecundity}& \textbf{Birth}\\
-& $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$
-& $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$
-& $\frac{\textrm{d}P}{\textrm{d}t}$
-& $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
-& $\frac{1}{k_s}$ & & & $h_0$ \\\hline
-\textbf{This study}&&&&&&&\\
-Leaf-construction cost, $\phi$ & $\downarrow$ & $\uparrow$ & & & & & \\
-Wood-construction cost, $\rho$ & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$ & \\
-Height at maturation, $h_m$ & &$\uparrow$ & & & & & $\downarrow$ & \\
-&&&&&&&\\\hline
-\textbf{Possible extensions}&&&&&&&\\
-Leaf area per sapwood area, $\theta$ & $\uparrow$& & $\downarrow$ & $\downarrow$ & & &\\
-Leaf-nitrogen & &$\downarrow$$\uparrow$ & & & & & \\
-Seed size & & & & & & & $\downarrow$ & $\uparrow$\\ \hline
-\end{tabular}
-\label{tab:trade-offs}
-\end{table}
-
-\newpage
-
-\section{Supplementary material}\label{supplementary-material}
-
-\subsection{Derivation of a simple allometric model of plant
-function}\label{derivation-of-a-simple-allometric-model-of-plant-function}
-
-Here we describe an allometric model linking the various size dimensions
-of a plant required by most ecologically realistic vegetation models
-(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
-roots) to a plant height. Table \ref{tab:allometry} provides a summary
-of the derived model, while Table \ref{tab:params} provdies estimates on
-key parameters.
-
-\begin{table}[h]
-\caption{Equations for an allometric growth model}
-\centering
-
-\begin{tabular}{p{5cm}p{5cm}p{5cm} }
-\\ \hline
-Variable & Function & Derivative\\ \hline
-leaf area & $a_\textrm{l}=\alpha_1 \, h^{\beta_1}$ & $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}= -\beta_1\big(\frac{a_\textrm{l}}{\alpha_1}\big)^{-(\beta_1+1)}$\\
-leaf mass & $m_\textrm{l}=\phi \, a_\textrm{l} $ & $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}=\phi$\\
-sapwoood area & $a_\textrm{ss}=\theta^{-1} \, a_\textrm{l}$ & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} =\theta^{-1} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$\\
-sapwood mass&$m_\textrm{ss}=\theta^{-1} \, \rho \, \eta_c \, a_\textrm{l} \, h $ & $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}}=\theta^{-1}\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$\\
-bark area & $a_\textrm{sb}=b \, \theta^{-1} \, da_\textrm{l}$ & $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t}=b \, \theta^{-1} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$\\
-bark mass&$m_\textrm{b}=b\, \theta^{-1} \, \rho \, \eta_c \, a_\textrm{l} \, h $ & $\frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}}=b \, \theta^{-1} \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}A} \big)$\\
-heartwood area & $a_\textrm{sh}=\int_0^t \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ & $\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}=k_\textrm{s} \, a_\textrm{ss}$\\
-heartwood mass & $m_\textrm{sh}=\int_0^t \frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ & $\frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}=k_\textrm{s} \, m_\textrm{ss}$\\
-root mass & $m_\textrm{r}=\alpha_3 \, a_\textrm{l}$ & $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}= \alpha_3$ \\\hline
-\end{tabular}
-\label{tab:allometry}
-\end{table}
-
-\subsubsection{Leaf area}\label{leaf-area}
-
-Based on empirically observed allometries (see main text), we assume an
-allometric log-log scaling relationship between the accumulated leaf
-area of a plant and its height:
-
-\begin{equation}\label{eq:ha}
-a_\textrm{l}=\alpha_1 \, h^{\beta_1}.
-\end{equation}
-
-Note, scaling relationship reversed from \citep{falster_influence_2011}.
-
-\subsubsection{Mass of sapwood}\label{mass-of-sapwood}
-
-We follow the model of\citep{yokozawa_foliage_1995} describing the
-vertical distribution of leaf area within the crowns of individual
-plants. This model can account for a variety of canopy profiles through
-a single parameter \(\eta\). Setting \(\eta=1\) results in a conical
-canopy, as seen in many conifers, while higher values, e.g. \(\eta=12\)
-, give a top-weighted canopy profile similar to those seen among
-angiosperms. Let \(S(z,h)\) be the sapwood area at height \(z\) for a
-plant with top height \(h\). Following Yokozawa and Hara (1995) we
-assume a relationship between \(S(z,h)\) and height such that
-
-\begin{equation}\label{eq:crown1}
-\frac{S(z,h)}{S(0,h)}= \big(1-\big(\frac{z}{h}\big)^\eta\big)^2.
-\end{equation}
-
-We also assume that each unit of sapwood area supports a fixed area of
-leaf (the pipe model, \citep{shinozaki_quantitative_1964}), so that the total
-canopy area of a plant relates to basal sapwood area \(S(0,h)\):
-
-\begin{equation}\label{eq:crown2}
-\frac{m_\textrm{l}}{\phi}= \theta \, S(0,h).
-\end{equation}
-
-Integrating \(S(z,h)\) gives a solution for the total mass of sapwood in
-the plant:
-
-\begin{equation}\label{eq:ms1}
-m_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, \textrm{d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}
-
-where
-\(\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}\)\citep{yokozawa_foliage_1995}.
-Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
-expression for sapwood mass as a function leaf area and height:
-
-\begin{equation}\label{eq:ms2}
-m_\textrm{s}=\rho \, \eta_c \, \theta^{-1} \, a_\textrm{l} \, h.
-\end{equation}
-
-\subsubsection{Bark mass}\label{bark-mass}
-
-Bark and phloem tissue are modelled using an analogue of the pipe model,
-leading to a similar equation as that for sapwood mass (eq.
-\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
-assumed to be a constant fraction \(b\) of sapwood area per unit leaf
-area such that
-
-\begin{equation}\label{eq:mb}
-m_\textrm{b}=b m_\textrm{s}.
-\end{equation}
-
-\subsubsection{Root mass}\label{root-mass}
-
-Also consistent with pipe-model assumption, we assume a fixed ratio of
-root mass per unit leaf area
-
-\begin{equation}\label{eq:mr}
-m_\textrm{r}=\alpha_3 \, a_\textrm{l}.
-\end{equation}
-
-Even though nitrogen and water uptake are not modelled explicitly,
-imposing a fixed ratio of root mass to leaf area ensures that
-approximate costs of root production are included in calculations of
-carbon budget.
-
-\newpage
-
-\begin{table}[h]
-\caption{Model parameters}
-\centering
-
-\include{figures/table-pars}
-
-\label{tab:params}
-\end{table}
-
-\subsection{Trait values maximising height
-growth}\label{trait-values-maximising-height-growth}
-
-We want to find the trait values maximising growth rate, \(G\). To make
-the analysis more tractible, we will focus on height growth rate and
-assume we are dealing with a plant of given height where 100\% of
-available energy is allocated to growth. From eq. \ref{eq:dhdt}, we thus
-have
-
-\begin{equation} \label{eq:G1}
-G = \frac{\textrm{d}h}{\textrm{d}t} = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}
-\times \frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
-\times \frac{\textrm{d}m_\textrm{t}} {\textrm{d}P}
-\times \frac{\textrm{d}P}{\textrm{d}t}.
-\end{equation}
-
-Maximising \(G\) is equivalent to solving for trait values \(x\) giving
-\(\partial G /\partial x = 0\). Noting that traits do not influence
-\(\frac{\textrm{d}h} {\textrm{d}a_\textrm{l}}\), and by assumption
-\(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}=1\), we thus have
-
-\begin{equation} \label{eq:G2}
-G =  c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}  \frac{ \textrm{d}P} {\textrm{d}t} \right),
-\end{equation}
-
-where \(c_1 = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}\). Eq.
-\ref{eq:G2} is a product of the form \(Y = W(x) \times Z(x)\). For,
-equations of this type, a solution to \(\partial{Y}/\partial{x} =0\) is
-given by \(\partial{W}/\partial{x} / W = \partial{Z}/\partial{x} / Z\).
-Maximum growth rate thus occurs when
-
-\begin{equation} \label{eq:G3}
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial x} \frac{1}{\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}} = \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} \frac{1}{\frac{ \textrm{d}P} {\textrm{d}t} }.
-\end{equation}
-
-In words, the maximum occurs when relative change in marginal leaf
-deployment with respect to trait is equal to the relative change in mass
-production with respect to the trait.
-
-Let us now try and simplify equation \ref{eq:G3}. First, let our trait
-of interest be leaf-construction cost, i.e. \(x=\phi\). From eq/
-\ref{eq:daldmt}, we can derive
-
-\begin{equation} \label{eq:G4}
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial \phi} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\ \right)^2.
-\end{equation}
-
-This simplifies eq. \ref{eq:G3} to
-
-\begin{equation} \label{eq:G5}
-\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}(x) =  \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} \frac{1}{\frac{ \textrm{d}P} {\textrm{d}t} }.
-\end{equation}
-
-From eq. \ref{eq:dPdt} we obtain,
-
-\begin{equation}\label{eq:dPdt2}
-\frac{dP}{\textrm{d}t} = c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}
-\end{equation}
-
-where
-\(c_2 = Y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, k_\textrm{i}})\)
-is independt of \(\phi\), and thus
-
-\begin{equation}\label{eq:dPdt3}
-\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi}  = - (1-B_4) a_\textrm{l} a_4\phi ^{-B_4}.
-\end{equation}
-
-Combing eqs. \ref{eq:G5}-\ref{eq:dPdt3}, we find that the maximum occurs
-when
-
-\begin{equation} \label{eq:G6}
-\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}(\phi) = \frac{- (1-B_4) a_\textrm{l} a_4\phi ^{-B_4}}{c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}}.
-\end{equation}
-
-Note also that if \(B4=1\),
-\(\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi} =0\),
-also if \(B4>1\),
-\(\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi} > 0\).
-A similar derivation can be produced linking wood density to
-productivity, via sapwood turnover.
-
-\subsection{Supplementary figures}\label{supplementary-figures}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_quantile_examples.pdf}
-\caption{\textbf{Estimating potential growth rate via quantile
-regression.} Example for 4 species. Growth rate estimated at six sizes,
-by fitting lines within given size intervals. Lines fitted to 99th
-quantile, but only when n \textgreater{} 500. Except for left-hand
-panel, estimated points in centre of interval.
-\label{fS-quantile_examples}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_lma_tradeoff.pdf}
-\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
-Data from \citep{wright_world-wide_2004} for 678 species from 51 sites, each
-point giving a species-average. Lines show standardised major axis lines
-fitted to data from each site, with intensity of shading adjusted
-according to strength of the relationship.\label{fS-leaf}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_size_dhdt.pdf}
-\caption{\textbf{Hump-shaped relationship between height growth rate and
-size.} \label{f-hump}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_size_dDdt.pdf}
-\caption{\textbf{Hump-shaped relationship diameter growth and size.}
-\label{f-SI_size_dDdt}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_size_dDdt.pdf}
-\caption{\textbf{Basal area growth rate increases with size.}
-\label{f-SI_size_dastdt}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_mass_fraction.pdf}
-\caption{\textbf{Change in allocation with size.}
-\label{f-mass_fraction}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
-and growth rate changes with plant size.} \label{f-lma_growth_size}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_lma_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
-and relative growth rate changes with plant size.}
-\label{f-lma_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
-and growth rate changes with plant size.} \label{f-rho_growth_size}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_rho_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
-and relative growth rate changes with plant size.}
-\label{f-rho_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_hmat_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-growth rate changes with plant size.} \label{f-rho_growth_size}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_hmat_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-relative growth rate changes with plant size.}
-\label{f-hmat_growth_size_relative}}
-\end{figure}
-
-\newpage
-
 \bibliography{references}

### 2015-02-18 notes/ms_short.tex
Move from short to medium format paper

% THIS FILE TAKEN FROM PANDOC DEMOS PAGE:http://johnmacfarlane.net/pandoc/demos.html,http://johnmacfarlane.net/pandoc/demo/mytemplate.tex

\documentclass[12pt, a4paper]{article}
\RequirePackage[hmargin=2cm,vmargin=2cm]{geometry}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

\usepackage{fontspec}
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}

% Fancy HEADER
\usepackage{fancyhdr}
\pagestyle{fancy}
\pagenumbering{arabic}
\lhead{\itshape{\nouppercase{\leftmark}}}
\chead{}
\rhead{\thepage}
%\lfoot{v }
\cfoot{}
%\rfoot{\thepage}

% \usepackage{natbib}
% \bibliographystyle{abbrvnat}
% \setcitestyle{authoryear,open={(},close={)}}

\usepackage[sort&compress]{natbib}
\citestyle{nature}
\bibliographystyle{ieeetr}

\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother

\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Untangling the link between traits, size and growth rate in plants},
            colorlinks=true,
            citecolor=black,
            urlcolor=grey,
            linkcolor=grey}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines

\usepackage{titling}

\title{Untangling the link between traits, size and growth rate in plants}

\author{
{\bf Daniel Falster} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{daniel.falster@mq.edu.au}\vspace{0.1in}\\
{\bf Rich FitzJohn} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{richard.fitzjohn@mq.edu.au}\vspace{0.1in}\\
{\bf Joe Wright} \\ \normalsize\emph{Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama} \\ \url{wrightj@si.edu}\vspace{0.1in}\\ }

\date{}

\begin{document}
\thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that
{\LARGE \bf \thetitle \vspace{0.2in}}

\theauthor

\section{Abstract}\label{abstract}

Plant growth is the foundation of the terrestrial biosphere, powering
food-webs, the carbon cycle and much of human
activity \citep{raupach_carbon_2010}. Plant species differ in many
functional traits, including aspects of tissue construction and energy
allocation, that drive differences in mechanical strength, biomass
partitioning, and rates of photosynthesis and tissue
turnover\citep{wright_world-wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
such traits influence whole-plant growth, with the simple relationships
predicted by existing theory often lacking empirical
support\citep{wright_functional_2010,poorter_are_2008}. Here we
provide a new set of theoretical predictions, validated using 20 years
of growth data\citep{condit_barro_2012,kitajima_leaf_2013}, showing
that the influence of three prominent traits on growth rate is moderated
by plant size, and that the impact of size on the nature of
relationships varies among traits. We extend a widely-used theoretical
model that links growth rate in seedlings with a single leaf
trait\citep{lambers_inherent_1992,wright_cross-species_2000} to
explicitly include influences of size, light environment, and three
other prominent traits. We find the influence of leaf construction cost
(\(\phi\)) on growth is strong in small plants but weakens with size. In
contrast, the influence of stem construction cost (\(\rho\)) peaks at
intermediate sizes, while the influence of height at maturation
(\(h_m\)) strengthens with size. Our theoretical model suggests the
change in trait-growth relationships with size arises due to changes in
biomass partitioning and reproductive allocation. Moreover, we show how
traits moderate plant responses to light environment and also determine
shade tolerance, supporting diverse empirical
results\citep{ruger_functional_2012,poorter_leaf_2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, our results provide a solid theoretical foundation for
trait ecology and thus provide a platform for understanding growth
across diverse species around the world.

\section{Main text}\label{main-text}

Functional traits capture core differences in the strategies plants use
to generate and invest surplus energy\citep{wright_world-wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
traits now exists for many of the world's 250000 plant
species\citep{cornwell_functional_2014}; yet the effect of these traits
on growth remains poorly understood. Prominent traits such as \(\phi\)
(also known as leaf mass per area) are widely considered as indicators
of growth rate\citep{reich_world-wide_2014}, because cheaper leaf
construction should enable faster canopy
expansion\citep{blackman_compound_1919,westoby_time_2000}. Indeed,
growth rate in seedlings is strongly and negatively correlated with
\(\phi\)\citep{cornelissen_seedling_1996,wright_cross-species_2000}
(Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
might also give a growth advantage in larger
plants\citep{enquist_general_2007}, an increasing number of empirical
studies have failed to detect any such
effect\citep{wright_functional_2010, @herault_functional_2011,
@poorter_are_2008}. Meanwhile, expectations on how other prominent
traits might influence growth are lacking. The growing discord between
theoretical expectation and empirical result thus indicates a strong
need for a new theoretical framework through which the influence of
traits and size on plant growth can be understood.

Here we extend a widely-used theoretical model that links growth rate in
seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross-species_2000}, to explicitly include influences of
size, light environment, and other prominent traits. This extension
allows us to generate novel predictions on the link between four key
traits and plant demography (Table \ref{tab:predictions}). Predictions
are compared against data from both a long-term
plot\citep{condit_barro_2012} and field
experiment\citep{kitajima_leaf_2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{S-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies\citep{wright_functional_2010}. The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
\ref{tab:predictions}) showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

We begin with a standard model for the amount of biomass available for
growth, \(dP / \textrm{d}t\), given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant\citep{falster_influence_2011}:

\begin{equation}\label{eq:dPdt}
\underbrace{\strut \frac{dP}{\textrm{d}t}}_\textrm{net biomass production}= \underbrace{\strut Y}_\textrm{yield} \big(\underbrace{\strut \frac{m_\textrm{l}}{\phi} \, A}_\textrm{photosynthesis} - \big(\underbrace{\strut \, \frac{m_\textrm{l}}{\phi} \, r_\textrm{l} + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)\big) - \big(\underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}\big).
\end{equation}

Here, \(m,r\), and \(k\) refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts \(l\)=leaves,
\(b\)=bark, \(s\)=sapwood and \(r\)=roots. \(A\) is the assimilation
rate of CO\(_2\) per leaf area and \(Y\) is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
\(a_\textrm{l} = m_\textrm{l} / \phi\), while total mass of living
tissues
\(m_\textrm{t}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.\)
Eq. \ref{eq:dPdt} assumes that tissues lost via turnover are replaced
before new growth occurs.

Current theory emphasises the relationship between \(\phi\) and relative
growth in mass\citep{enquist_general_2007},
\(\textrm{d}P / \textrm{d}t / m_\textrm{t}\). For seedlings, which are
young and mostly leaf, it is reasonable to ignore all turnover terms as
well as the respiration terms for non-leaf tissues. Net production then
becomes a linear function of leaf area and net photosynthesis per leaf
area (\(A_\textrm{net} = Y(A - r_\textrm{l})\)), making relative growth
rate a linear function of \(\phi\):

\begin{equation}\label{eq:RGR}
\underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}\frac{1}{m_\textrm{t}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{t}}. \end{equation}

Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
relation to \(\phi\)\citep{wright_cross-species_2000}, this
approximation does not directly link to other traits, or to the
variables that are routinely collected for large trees: namely plant
height (\(h\)) and stem cross-sectional area (\(a_\textrm{st}\)) or
diameter \(D\).

To model growth in either height or basal area requires we to account
not just for mass production, but also for the costs of building new
tissues, allocation to reproduction, and architectural layout. In eq.
\ref{eq:RGR}, these terms are missing, whereas in the model presented
here they are made explicit.

Mathematically, height growth can be decomposed into a product of
physiologically relevant terms\citep{falster_influence_2011}:

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}}_\textrm{mass production}.
\end{equation}

The first term on the right of eq \ref{eq:dhdt},
\(\textrm{d}h / \textrm{d}a_\textrm{l}\), is the growth in plant height
per unit growth in total leaf area; accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension\citep{poorter_architecture_2006}.

The second term, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\),
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\)
can itself be expressed as a sum of construction costs per unit leaf
area:

\begin{equation}\label{eq:daldmt}
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
= \big(\phi
 + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
\end{equation}

The third term in eq \ref{eq:dhdt},
\(\textrm{d}m_\textrm{t} / \textrm{d}P\), gives the fraction of net
biomass production (eq. \ref{eq:dPdt}) that is allocated to growth
rather than reproduction or storage.

In a similar way, basal area (\(a_\textrm{st}\)) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
(\(a_\textrm{ss}, a_\textrm{sb}, a_\textrm{sh}\) respectively):
\(\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}= \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}\).
Noting that heartwood grows via turnover of sapwood, such that
\(\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} a_\textrm{ss}\),
we have an equation for basal area growth that conatins many of the same
elements as eq. \ref{eq:dhdt}:

\begin{equation}\label{eq:dast}
\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}} \times \frac{\textrm{d}m_\textrm{t}}{\textrm{d}P} \times \frac{\textrm{d}P}{\textrm{d}t} + k_\textrm{s} a_\textrm{ss}.
\end{equation}

Diameter growth is then given by the geometric relationship between stem
diameter (\(D\)) and \(a_\textrm{st}\):

\begin{equation} \label{eq:dDdt}
\frac{\textrm{d}D}{\textrm{d}t}= \sqrt{\frac{\pi}{a_\textrm{st}}} \frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}.
\end{equation}

Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
therefore hold for any model of plant growth. To make explicit
predictions, however, requires two additional elements: 1) A specific
model describing how the various mass and area terms
(\(m_\textrm{l}, m_\textrm{s}, m_\textrm{b}, m_\textrm{r}, a_\textrm{l}, a_\textrm{s}, a_\textrm{b}\))
vary relative to one another; and 2) For any trade-offs in plant
function to be specified; this is where traits have an influence. These
two components are as follows.

To address the first of these challenges, we use a functional balance
model\citep{yokozawa_foliage_1995,falster_influence_2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{f-assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

This new growth model captures the intrinsically size-dependent nature
of plant growth (Table \ref{tab:predictions}a), recovering well-known
empirical patterns\citep{sillett_increasing_2010, king_size-related_2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size\citep{sillett_increasing_2010,
@stephenson_rate_2014}, due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve\citep{herault_functional_2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates\citep{iida_linking_2014}.

We now incorporate the effects of some prominent traits by embedding
explicit trade-offs within this growth model. Variation in each trait
implies benefits and costs. These affect specific elements of the growth
equations presented above (Table \ref{tab:trade-offs}).

By definition, the main effect of increasing \(\phi\) is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of \(\phi\) on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction\citep{wright_world-wide_2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dPdt} to \(\phi\). We use an allometric scaling relationship of
the form \(k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}\), which has been
shown to describe patterns across diverse
ecosystems\citep{wright_world-wide_2004} (Fig. \ref{fS-leaf}).

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between \(\phi\) and growth that
changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
Decreasing \(\phi\) has two impacts on growth rate. First, lower
\(\phi\) increases marginal leaf deployment
(\(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\) by economising on
construction costs. Second, lower \(\phi\) decreases net production
(\(\textrm{d}P / \textrm{d}t\)), due to increased leaf turnover. Whether
lower \(\phi\) increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing \(\phi\) increases growth
rate. When plants are large, the influence of \(\phi\) on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
\(\phi\) is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

As for leaf, cheaper stem construction (lower \(\rho\)) decreases the
cost of deploying a unit of leaf area, and may thereby increase growth
rates. However, the cost of cheap stem construction must also be
accounted for. However, the direct physiological trade-offs of lower
\(\rho\) are less understood than for \(\phi\). One possibility is that
lower \(\rho\) decreases mechanical strength resulting in higher
mortality\citep{chave_towards_2009,wright_functional_2010}. Under
this scenario, decreasing \(\rho\) will always provide a growth
advantage, because the costs of low \(\rho\) are not realised within the
terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
\ref{tab:trade-offs}). In addition, lower \(\rho\) may increase turnover
of sapwood. Mirroring our assumptions for leaves, we let sapwood
turnover decrease with \(\rho\),
i.e.\(k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}\), with \(\beta_5 > 1\).
Under this scenario, decreasing \(\rho\) provides an advantage in height
growth only when the benefits of low \(\rho\) outweigh the costs. This
effect is strongest at intermediate sizes. The decline in height growth
at low \(\rho\) with size is more marked than for either stem basal-area
and diameter growth, because high sapwood turnover also contributes to
this growth measure. Thus diameter growth continues, even as net
production approaches zero.

Unlike leaf and stem traits, height at maturation (\(h_m\)) moderates
growth by adjusting the amount of energy invested in growth, i.e.~the
term \(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) in eqs.
\ref{eq:dhdt} and \ref{eq:dast}. Greater height at maturation (\(h_m\))
can thus lead to a growth advantage by increasing
\(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) (Table
\ref{tab:predictions}). At smaller sizes, there is no differentiation
among species, because most species are focusing on growth (Fig.
\ref{f-BCI}). At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate (Fig. \ref{f-BCI}).

Two other important predictions arise directly from linking traits to
growth rate (Table \ref{tab:predictions}c). First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low \(rho\) increase growth more substantially with incraeses in
light\citep{ruger_functional_2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low \(\phi\) are lost (Fig. \ref{f-growth_light}). This results also
agrees with empirical results showing that growth for species with low
\(rho\) actually decreases as size
increases\citep{ruger_functional_2012}. In contrast, growth rate tends
to increase in size for the species with highest
\(\rho\)\citep{ruger_functional_2012}.

Variation in \(\phi\) also moderates the response of growth to changes
in light, with species having the lowest \(rho\) being most responsive.
However, unlike for \(\rho\), the effect appears only for the smallest
size classes. Accordingly, empirical analyses have detected such a link
in small plants but not for plants \textgreater{} 0.01m
dbh\citep{ruger_functional_2012}.

Second, species with low \(\phi\) and low \(\rho\) are predicted to be
less shade tolerant (Fig. \ref{f-wplcp}). At low \(\phi\) (\(\rho\)),
leaf (sapwood) turnover is higher and thus a greater light income is
needed to offset these costs. As previously suggested
\citep{givnish_adaptation_1988}, shade-tolerance also decreases with
height because as size increase, the total amount of energy needed to
offset respiratory and turnover costs in the stem also increases. These
finding matches well known empirical patterns, where both \(\phi\) and
\(\rho\) have been linked to shade tolerance\citep{poorter_leaf_2006, lusk_ontogenetic_2008,osunkoya_light_1996}.

Conclusion,\ldots{}..

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  traits do effect growth rate, but not in the simple linear fashion
  previously anticipated.
\item
  Relationships change with size due to changes in biomass partitioning
  and reproductive allocation.
\item
  explains why LMA does not show relationship to growth among larger
  plants. Does not mean unimportant for growth, just effects
  concentrated at small sizes.
\item
  in fact, explains plasticity in LMA with size, where LMA tends to
  increase with size \citep{thomas_asymptotic_1999;
  @thomas_photosynthetic_2010}
\item
  Framework can extend to other traits (:

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    In the current model, the only effects of the trait seed size
    (\(s\)) are to influence starting size of seedlings and fecundity of
    the mother. Seed size therefore has no direct influence on absolute
    growth. In may studies seed size has been found to be negatively
    correlated with mass-based relative growth rate of seedlings. This
    effect is recovered here because of the intrinsic decrease in RGR
    with size.
  \item
    leaf N:
  \item
    sapwood area to leaf area
  \end{itemize}
\item
  Take home: Don't look for traits as being fast or slow growth, rather
  consider traits as defining potential trajectory across sizes and
  light environments.
\end{itemize}

\section{Methods}\label{methods}

\subsection{Plant growth model}\label{plant-growth-model}

TREE

\subsection{Data analysis}\label{data-analysis}

Seedling data

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  about data
\item
  method for estimating growth rate
\end{itemize}

Long term plot data

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  about data
\item
  method for estimating growth rate
\end{itemize}

Traits

\newpage

\section{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{figures/BCI_data.pdf}
\caption{\textbf{The relationship between traits and potential growth
rate varies with plant size.} For XXX species from tropical rain forest
in Panama, we estimated the potential growth rate of individual's in
each species at a series of diameters \(D\), indicated along right hand
side. The size of circles in each panel indicates the number of data
points used to estimate potential growth rate. Traits values were
calculated from representative individuals in each species.
\label{f-BCI}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figs/allometry.pdf}
\caption{\textbf{Key assumptions of a functional balance allometric
model, evaluated using global dataset.} We used the biomass and
allometry database to evaluate model assumptions about \textbf{a,}
scaling of leaf area with plant height, \textbf{b} Scaling of sapwood
area with leaf area, and \textbf{c} scaling of root mass with leaf area.
Each dot is a single plant. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship. Colours indicate vegetation
type. Dashed black lines show values expected under functional-balance
assumption (see Supplementary text for details). \label{f-assumptions}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/growth_light.pdf}
\caption{\textbf{Traits moderate the responsiveness of growth to changes
in light environment.} Panels show predicted relationship between
specific trait and diameter growth rate, for a plant of specified
diameter and under a range of shading environments.
\label{f-growth_light}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/max_leaf_above.pdf}
\caption{\textbf{Low construction cost leads to shade intolerance,
because of costs of high turnover.} Panels show effect of traits on
maximum amount of shading that can be endured before net production (eq.
\ref{eq:dPdt}) reaches zero. Lines indicate relationship for plants of a
given height. \label{f-wplcp}}
\end{figure}

\newpage

\section{Tables}\label{tables}

\begin{table}[h]
\caption{Theoretical predictions on the relationship between traits and demography}

{\footnotesize
\centering
\begin{tabular}{p{4cm}p{4cm}p{2cm}p{6cm}}
\\
\multicolumn{4}{l}{\textbf{a) Intrinsic changes in demography with size}} \\ \\
\hline
Variable & Change with size & & Support \\ \hline
Height growth & strongly hump-shaped & & Fig. \ref{f-hump} \\
Diameter growth & weakly hump-shaped & & Fig. \ref{f-SI_size_dDdt}\\
Basal area growth & increases & & Fig. \ref{f-SI_size_dastdt}\\
Live mass growth & ?? & & ?? \\
ABG mass growth & increases & & ??\\
Relative growth (any) & decreases & & ??\\
Shade tolerance & decreases & & Fig. \ref{f-wplcp}\\ \\
\end{tabular}

\begin{tabular}{p{4cm}p{2.5cm}p{3.5cm}p{6cm}}
\multicolumn{4}{l}{\textbf{b) Relationship of demography to traits}} \\ \\
\hline
Variable & When small & Change with size & Support \\ \hline
\multicolumn{4}{l}{\emph{Leaf-construction cost}} \\

Height growth & negative & flattens then reverses  & Fig. \ref{f-lma_growth_size}\\
Diameter growth & negative & flattens & Fig. \ref{f-growth_light}, \ref{f-lma_growth_size}\\
Basal area growth & negative & flattens then reverses & Fig. \ref{f-lma_growth_size}\\
Live mass growth & ??  & & \\
ABG mass growth & negative & same & Fig. \ref{f-lma_growth_size} \\
Relative growth (any) & negative & weakens & Fig. \ref{f-lma_growth_size_relative}\\
Shade tolerance & positive &  strengthens & Fig. \ref{f-wplcp}\\ \\

\multicolumn{4}{l}{\emph{Wood-construction cost}} \\

Height growth & negative & strengthens  & Fig. \ref{f-rho_growth_size}\\
Diameter growth & negative & strengthens & Fig. \ref{f-growth_light}, \ref{f-rho_growth_size} \\
Basal area growth & negative & strengthens & Fig. \ref{f-rho_growth_size} \\
Live mass growth & ??  & & \\
ABG mass growth & negative & strengthens & Fig. \ref{f-rho_growth_size}\\
Relative growth (any) & negative & weakens & Fig. \ref{f-rho_growth_size_relative}\\
Shade tolerance & positive &  strengthens & Fig. \ref{f-wplcp}\\ \\

\multicolumn{4}{l}{\emph{Height at maturation}} \\
Any growth measure & none & becomes positive  & Fig. \ref{f-hmat_growth_size}\\
Relative growth (any) & none & becomes positive  & Fig. \ref{f-hmat_growth_size_relative}\\
Shade tolerance & none &  none &  \\  \\ \hline
\end{tabular}
}
\label{tab:predictions}
\end{table}

\newpage

\begin{table}[h]
\caption{Effect of traits and related trade-offs on demography}
\begin{tabular}[c]{l|ccccc|c|c|c}
\multicolumn{9}{c}{}\\ \hline
& \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
& \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality}& \textbf{Fecundity}& \textbf{Birth}\\
& $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$
& $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$
& $\frac{\textrm{d}P}{\textrm{d}t}$
& $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
& $\frac{1}{k_s}$ & & & $h_0$ \\\hline
\textbf{This study}&&&&&&&\\
Leaf-construction cost, $\phi$ & $\downarrow$ & $\uparrow$ & & & & & \\
Wood-construction cost, $\rho$ & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$ & \\
Height at maturation, $h_m$ & &$\uparrow$ & & & & & $\downarrow$ & \\
&&&&&&&\\\hline
\textbf{Possible extensions}&&&&&&&\\
Leaf area per sapwood area, $\theta$ & $\uparrow$& & $\downarrow$ & $\downarrow$ & & &\\
Leaf-nitrogen & &$\downarrow$$\uparrow$ & & & & & \\
Seed size & & & & & & & $\downarrow$ & $\uparrow$\\ \hline
\end{tabular}
\label{tab:trade-offs}
\end{table}

\newpage

\section{Supplementary material}\label{supplementary-material}

\subsection{Derivation of a simple allometric model of plant
function}\label{derivation-of-a-simple-allometric-model-of-plant-function}

Here we describe an allometric model linking the various size dimensions
of a plant required by most ecologically realistic vegetation models
(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
roots) to a plant height. Table \ref{tab:allometry} provides a summary
of the derived model, while Table \ref{tab:params} provdies estimates on
key parameters.

\begin{table}[h]
\caption{Equations for an allometric growth model}
\centering

\begin{tabular}{p{5cm}p{5cm}p{5cm} }
\\ \hline
Variable & Function & Derivative\\ \hline
leaf area & $a_\textrm{l}=\alpha_1 \, h^{\beta_1}$ & $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}= -\beta_1\big(\frac{a_\textrm{l}}{\alpha_1}\big)^{-(\beta_1+1)}$\\
leaf mass & $m_\textrm{l}=\phi \, a_\textrm{l} $ & $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}=\phi$\\
sapwoood area & $a_\textrm{ss}=\theta^{-1} \, a_\textrm{l}$ & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} =\theta^{-1} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$\\
sapwood mass&$m_\textrm{ss}=\theta^{-1} \, \rho \, \eta_c \, a_\textrm{l} \, h $ & $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}}=\theta^{-1}\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$\\
bark area & $a_\textrm{sb}=b \, \theta^{-1} \, da_\textrm{l}$ & $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t}=b \, \theta^{-1} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$\\
bark mass&$m_\textrm{b}=b\, \theta^{-1} \, \rho \, \eta_c \, a_\textrm{l} \, h $ & $\frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}}=b \, \theta^{-1} \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}A} \big)$\\
heartwood area & $a_\textrm{sh}=\int_0^t \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ & $\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}=k_\textrm{s} \, a_\textrm{ss}$\\
heartwood mass & $m_\textrm{sh}=\int_0^t \frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ & $\frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}=k_\textrm{s} \, m_\textrm{ss}$\\
root mass & $m_\textrm{r}=\alpha_3 \, a_\textrm{l}$ & $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}= \alpha_3$ \\\hline
\end{tabular}
\label{tab:allometry}
\end{table}

\subsubsection{Leaf area}\label{leaf-area}

Based on empirically observed allometries (see main text), we assume an
allometric log-log scaling relationship between the accumulated leaf
area of a plant and its height:

\begin{equation}\label{eq:ha}
a_\textrm{l}=\alpha_1 \, h^{\beta_1}.
\end{equation}

Note, scaling relationship reversed from \citep{falster_influence_2011}.

\subsubsection{Mass of sapwood}\label{mass-of-sapwood}

We follow the model of\citep{yokozawa_foliage_1995} describing the
vertical distribution of leaf area within the crowns of individual
plants. This model can account for a variety of canopy profiles through
a single parameter \(\eta\). Setting \(\eta=1\) results in a conical
canopy, as seen in many conifers, while higher values, e.g. \(\eta=12\)
, give a top-weighted canopy profile similar to those seen among
angiosperms. Let \(S(z,h)\) be the sapwood area at height \(z\) for a
plant with top height \(h\). Following Yokozawa and Hara (1995) we
assume a relationship between \(S(z,h)\) and height such that

\begin{equation}\label{eq:crown1}
\frac{S(z,h)}{S(0,h)}= \big(1-\big(\frac{z}{h}\big)^\eta\big)^2.
\end{equation}

We also assume that each unit of sapwood area supports a fixed area of
leaf (the pipe model, \citep{shinozaki_quantitative_1964}), so that the total
canopy area of a plant relates to basal sapwood area \(S(0,h)\):

\begin{equation}\label{eq:crown2}
\frac{m_\textrm{l}}{\phi}= \theta \, S(0,h).
\end{equation}

Integrating \(S(z,h)\) gives a solution for the total mass of sapwood in
the plant:

\begin{equation}\label{eq:ms1}
m_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, \textrm{d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}

where
\(\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}\)\citep{yokozawa_foliage_1995}.
Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
expression for sapwood mass as a function leaf area and height:

\begin{equation}\label{eq:ms2}
m_\textrm{s}=\rho \, \eta_c \, \theta^{-1} \, a_\textrm{l} \, h.
\end{equation}

\subsubsection{Bark mass}\label{bark-mass}

Bark and phloem tissue are modelled using an analogue of the pipe model,
leading to a similar equation as that for sapwood mass (eq.
\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
assumed to be a constant fraction \(b\) of sapwood area per unit leaf
area such that

\begin{equation}\label{eq:mb}
m_\textrm{b}=b m_\textrm{s}.
\end{equation}

\subsubsection{Root mass}\label{root-mass}

Also consistent with pipe-model assumption, we assume a fixed ratio of
root mass per unit leaf area

\begin{equation}\label{eq:mr}
m_\textrm{r}=\alpha_3 \, a_\textrm{l}.
\end{equation}

Even though nitrogen and water uptake are not modelled explicitly,
imposing a fixed ratio of root mass to leaf area ensures that
approximate costs of root production are included in calculations of
carbon budget.

\newpage

\begin{table}[h]
\caption{Model parameters}
\centering

\include{figures/table-pars}

\label{tab:params}
\end{table}

\subsection{Effect of traits on height growth} \label{trait-values-maximising-height-growth}

We want to show how two traits $\phi$ and $\rho$ influence growth rate,
by looking at the effect of
traits on the components of growth rate (eq. \ref{eq:dhdt}. To make
the analysis more tractable, we will assume we are dealing with a plant of
given height where 100\% of available energy is allocated to growth,
i.e. \(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}=1\). Note also
that by the assumptions of our model and above, traits do not influence
\(\frac{\textrm{d}h} {\textrm{d}a_\textrm{l}}\). Thus we have

\begin{equation} \label{eq:G2}
G = \frac{\textrm{d}h}{\textrm{d}t} = c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}  \frac{ \textrm{d}P} {\textrm{d}t} \right),
\end{equation}

where \(c_1 = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}\). Eq.
\ref{eq:G2} is a product of the form \(Y(x) = W(x) \times Z(x)\). For,
equations of this type, the derivative \(\partial{Y}/\partial{x}\) is
given by \(\partial{W}/\partial{x} \, Z + W \, \partial{Z}/\partial{x}\). Thus
the derivative of \(G\) with respect to some trait $x$  is given by

\begin{equation} \label{eq:dG}
\frac{\partial G} {\partial x} =
\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial x}
 \, \frac{ \textrm{d}P} {\textrm{d}t}
+ \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
\, \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}.
\end{equation}

If this derivative is positive (negative), there will be a positive (negative) relationship
between traits and growth at the height where it is evaluated.

We further simplify by expanding the derivative of eq. \ref{eq:daldmt}:

\begin{equation} \label{eq:da_dx}
\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}
{\partial x} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\ \right)^2
\, \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
}{\partial x}.
\end{equation}

Combing eqs. \ref{eq:dG}-\ref{eq:da_dx}, we have

\begin{equation} \label{eq:dG2}
\frac{\partial G} {\partial x} =
\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
\left(
\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
- \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
}{\partial x}
 \, \frac{ \textrm{d}P} {\textrm{d}t}
\right).
\end{equation}

Since is $\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}>0$ always, the sign
of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
RHS. Rearranging we find that $\frac
{\partial G} {\partial x} < 0$ if

\begin{equation}\label{eq:dg3}
\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
< \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
}{\partial x} \, \frac{ \textrm{d}P} {\textrm{d}t},
\end{equation}

or equivalently
\begin{equation}\label{eq:dg4}
\frac{
\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
{\frac{ \textrm{d}P} {\textrm{d}t}}
<
\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
\end{equation}

In words, there will be a negative relationship between trait x and growth if the
relative change in production due to change in traits is less than the relative change
in the total construction cost for a unit of leaf area due to change in trait. Crucially,
key elements in eq. \ref{eq:dg3} are modified by height.

We can also solve for trait values \(x\) maximising growth rate, i.e. the value
giving \(\partial G /\partial x = 0\):

\begin{equation}\label{eq:max}
\frac{
\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
{\frac{ \textrm{d}P} {\textrm{d}t}}
=
\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
\end{equation}

\subsubsection{Leaf construction cost}

If the trait of interest is leaf-construction cost, i.e. \(x=\phi\), we can expand
eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{\textrm{d}m_\textrm
{t}} {\textrm{d}a_\textrm{l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dPdt} we obtain,

\begin{equation}\label{eq:dPdt2}
\frac{dP}{\textrm{d}t} = c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}
\end{equation}

where
\(c_2 = Y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, k_\textrm{i}})\)
is independent of \(\phi\), and thus

\begin{equation}\label{eq:dPdt3}
\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi}  =
(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}.
\end{equation}

Eq. \ref{eq:dPdt3} shows that if \(B4=1\), \(\frac{ \textrm{d}P} {\textrm{d}t}\) is independent
of $\phi$. In contrast, if \(B4>1\) (which tends to be the case empirically), \(\frac{ \textrm
{d}P} {\textrm{d}t}\) increases as $\phi$ increases, more acutely at low $\phi$.

Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
with traits if

\begin{equation} \label{eq:G6}
\frac{(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}}{c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}}
< \frac{1}{\phi
 + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm
 {b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}}.
\end{equation}

\subsubsection{Wood construction cost}

Do similar to previous section for wood construction cost.

\subsection{Supplementary figures}\label{supplementary-figures}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_quantile_examples.pdf}
\caption{\textbf{Estimating potential growth rate via quantile
regression.} Example for 4 species. Growth rate estimated at six sizes,
by fitting lines within given size intervals. Lines fitted to 99th
quantile, but only when n \textgreater{} 500. Except for left-hand
panel, estimated points in centre of interval.
\label{fS-quantile_examples}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_tradeoff.pdf}
\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
Data from \citep{wright_world-wide_2004} for 678 species from 51 sites, each
point giving a species-average. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship.\label{fS-leaf}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_size_dhdt.pdf}
\caption{\textbf{Hump-shaped relationship between height growth rate and
size.} \label{f-hump}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_size_dDdt.pdf}
\caption{\textbf{Hump-shaped relationship diameter growth and size.}
\label{f-SI_size_dDdt}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_size_dDdt.pdf}
\caption{\textbf{Basal area growth rate increases with size.}
\label{f-SI_size_dastdt}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_mass_fraction.pdf}
\caption{\textbf{Change in allocation with size.}
\label{f-mass_fraction}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
\caption{\textbf{The expected correlation between leaf-construction cost
and growth rate changes with plant size.} \label{f-lma_growth_size}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_effects_at_diameters_relative.pdf}
\caption{\textbf{The expected correlation between leaf-construction cost
and relative growth rate changes with plant size.}
\label{f-lma_growth_size_relative}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
\caption{\textbf{The expected correlation between stem-construction cost
and growth rate changes with plant size.} \label{f-rho_growth_size}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_rho_effects_at_diameters_relative.pdf}
\caption{\textbf{The expected correlation between stem-construction cost
and relative growth rate changes with plant size.}
\label{f-rho_growth_size_relative}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_hmat_effects_at_diameters.pdf}
\caption{\textbf{The expected correlation between maximum height and
growth rate changes with plant size.} \label{f-rho_growth_size}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_hmat_effects_at_diameters_relative.pdf}
\caption{\textbf{The expected correlation between maximum height and
relative growth rate changes with plant size.}
\label{f-hmat_growth_size_relative}}
\end{figure}

\newpage

\bibliography{references}
\end{document}

### 2015-02-18 R/BCI.R

@@ -250,8 +250,8 @@ figure_BCI_data2 <- function(data) {
          col=make_transparent(cols[class], .6),
          bg=make_transparent(cols[class], .5))
 
-    segments(dt$xr[1,], dt$yr_i[1,], dt$xr[2,], dt$yr_i[2,],
-             col=cols, lwd=3, lty=2)
+    # segments(dt$xr[1,], dt$yr_i[1,], dt$xr[2,], dt$yr_i[2,],
+    #          col=cols, lwd=3, lty=2)
     ## segments(dt$xr[1,], dt$yr_w[1,], dt$xr[2,], dt$yr_w[2,],
     ##          col=cols, lwd=3, lty=3)
     segments(dt$xr[1,], dt$yr_j[1,], dt$xr[2,], dt$yr_j[2,],

### 2015-02-18 R/figures.R

@@ -44,6 +44,30 @@ plant_info <- function(p) {
   x
 }
 
+## Code for trait deriative figure
+figure_trait_deriative <- function(type, trait_name="lma", canopy_openness=1,
+                                strategy=default_strategy()) {
+
+  strategy <- strategy
+  strategy[[trait_name]] <- 0.05
+  dat <- figure_rate_vs_size_data(canopy_openness, strategy)
+  strategy2 <- strategy
+  x <- strategy2[[trait_name]]
+  dx = 0.01 *x
+  strategy2[[trait_name]] <- x+dx
+  dat2 <- figure_rate_vs_size_data(canopy_openness, strategy2)
+
+  line1 <- (dat2$net_production - dat$net_production)/dx/dat$net_production
+  line2  <- (1/(dat2$dleaf_area_dleaf_mass * dat2$leaf_fraction)
+              - 1/(dat$dleaf_area_dleaf_mass * dat$leaf_fraction)) / dx /
+            1/(dat2$dleaf_area_dleaf_mass * dat2$leaf_fraction)
+  height <- dat$height
+
+  plot(dat[["height"]], line1, type="l", xlab="Height (m)", ylab= "relative change", ylim=c(0,20))
+  points(dat[["height"]], line2, type="l", col="red")
+}
+
+
 ## Code for the "rates vs size" figure set:
 figure_rate_vs_size <- function(type, canopy_openness=1,
                                 strategy=default_strategy()) {
@@ -60,7 +84,7 @@ figure_rate_vs_size <- function(type, canopy_openness=1,
 }
 
 figure_rate_vs_size_data <- function(canopy_openness, strategy) {
-  heights <- seq(Plant(strategy)$height, strategy$hmat, length.out=50)
+  heights <- seq(Plant(strategy)$height, strategy$hmat, length.out=100)
   env <- fixed_environment(canopy_openness)
   res <- run_plant_to_heights(heights, strategy, env)
   i <- res[["net_production"]] > 0

### 2015-02-18 R/ms-fun.R

@@ -1,8 +1,8 @@
 tex_2_pdf <- function(texfile){
 
-  system(sprintf('latexmk -pdf -pdflatex="xelatex -interaction=nonstopmode"
-                 %s', texfile))
+  system(sprintf('latexmk -pdf -pdflatex="xelatex -interaction=nonstopmode" %s', texfile))
   aux.files <- paste0(tools::file_path_sans_ext(texfile),
-                     c(".log", ".aux", ".bbl", ".blg"))
+                     c(".log", ".aux", ".bbl", ".blg", ".fls", ".out",
+                      ".fdb_latexmk"))
   file.remove(aux.files[file.exists(aux.files)])
 }

### 2015-02-18 ms.tex
Enhance introduction

@@
 plants use to generate and invest surplus energy
-\citep{wright_world-wide_2004,chave_towards_2009,westoby_plant_2002}. Useful
-traits are those that
-can be easily measured for diverse species and where differences between
-species are large compared to intra-specific variation. Moroever, the uptake
-of traits reflects the observation that although most plants have same basic
-construction, physiological function and resource requirements, species differ
-considerably from one another in rates of carbon, nitrogen and water uptake,
-growth, mortality and seed production. Data for prominent traits now exists
-for many of the world's 250000 plant species \citep{cornwell_functional_2014}.
-The effect of some traits of multiple aspects of plant function has also been
-well documented, including mechanical strength, biomass partitioning, rates of
-photosynthesis and rates of tissue turnover
-\citep {wright_world-wide_2004,chave_towards_2009}. As a result, the idea has
-taken hold that the growth startegy and ecology of a plant species might usefully
-be characterised simply by knowing its traits.
+\citep{wright_world_wide_2004, chave_towards_2009, westoby_plant_2002}.
+Although most plants have the same basic construction, physiological function
+and resource requirements, species differ considerably from one another in
+rates of carbon, nitrogen and water uptake, growth, mortality and seed
+production. Some of these functional differences have been linked to
+underlying variation in relatively simple physiological traits, such as leaf
+mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
+diameter,  and seed size  \citep{wright_world_wide_2004,chave_towards_2009}.
+These traits are easily measured across diverse species, and variation between
+species is large compared to variation within species. Data for some traits
+now exists for up to 10\% of the world's 250000 plant species
+\citep{cornwell_functional_2014}. During the last decade, the idea has taken
+hold that the growth strategy and ecology of a plant species might usefully be
+characterised simply by knowing its traits.
 
-While the influence of plant traits on elements of plant physiologival
-function is increasingly understood, attempts at using traits to predict
-demographic rates have only met with mixed success. In seedlings, the trait
-leaf mass per area (LMA) -- a central part of the leaf economics spectrum --
-is tightly related to plant relative growth rate in mass. LMA and its
-correlate leaf lifespan have also been linked to height growth rate for small
-seedlings and saplings. These early successes prompted researchers to search
-for similar relationships in large plants. However, the results showed that in
-large plants, LMA was not correlated with diameter growth rate. Meanwhile,
-other traits such as wood density showed strong relationships
-to growth in large plants, but only weak relationships in seedlings \citep
-{castro-diez_stem_1998}. To date each relationship has been quantified mostly for
-one or few traits and at a single life-stage. Yet the picture is emerging is that the
-link between traits and growth may be modified by plant size; a finding which would
-challenge the asummption applied by many that particular ends of the trait spectrum
-translate directly into fast or slower growth.
+While the influence of plant traits on elements of plant physiological
+function is increasingly understood, attempts at using traits to understand
+demographic rates have only met with mixed success
+\citep{wright_functional_2010, poorter_are_2008}. In seedlings, the trait LMA
+-- a central part of the leaf economics spectrum
+\citep{wright_world_wide_2004} --is tightly related to plant relative growth
+rate in mass \citep{lambers_inherent_1992, wright_cross_species_2000}. LMA and
+its correlate leaf lifespan have also been linked to height growth rate for
+small seedlings and saplings \citep{reich_leaf_1992, poorter_leaf_2006}. These
+early successes prompted researchers to search for similar relationships in
+large plants. However, the results showed that in trees LMA was not correlated
+with diameter growth rate  \citep{wright_functional_2010, poorter_are_2008,
+herault_functional_2011}. Meanwhile, other traits such as wood density showed
+strong relationships to growth in large plants \citep{wright_functional_2010},
+but only weak relationships in seedlings  \citep{castro-diez_stem_1998}. Thus
+far, most trait growth relationships have been quantified for one or few
+traits and at a single life- stage. Yet the picture is emerging is that the
+link between traits and growth may be modified by plant size
+\citep{ruger_functional_2012}; a finding that challenges the wide-spread
+assumption that particular ends of the trait spectrum translate directly into
+faster or slower growth.
 
-The challenge of interpreting diverse empirical results linking traits to
-growth rate -- or not -- would clearly be easier if we had clearl expectations
-on what signal one should expect, given our understanding of how traits
-infleunce plant function. This is a primary, yet perhaps underappreciated,
-role for theory: to lay expectations based on mechistic principles against
-which empircal data can be compared. Current empircial results suggest the
-effect of traits on growth changes with size and possibly also light
-environment, yet current theory says little about how relationships might come
-about. A widely-used model suggests mass-based relative growth rate will be a
-linear function of LMA, at least for seeldings.  An extension of model
-suggests a similar relationship may hold at larger sizes. But this prediction
-doesn't empirical results. Meanwhile, theoretical predictions on how other
-traits should influence growth are largely absent.
+The challenge of interpreting diverse empirical results linking  -- or not --
+traits to growth rate would clearly be easier with clear expectations on what
+signal one should expect, given current understandings of how traits influence
+plant function. Generating these expectations is a primary, yet perhaps under-
+appreciated, role for theory \citep{kokko_modelling_2007}. Current empirical
+results suggest the effect of traits on growth changes with size and possibly
+also light environment, yet current theory says little about how such
+relationships might come about. A widely-used model shows that for seedlings,
+mass-based relative growth rate is linearly and negatively related to LMA
+\citep{lambers_inherent_1992, cornelissen_seedling_1996,
+wright_cross_species_2000}. An extension of the model suggests a similar
+relationship may hold at larger sizes \citep{enquist_general_2007}. But as
+described above, this prediction is not supported by empirical results.
+Meanwhile, theoretical predictions on how other traits should influence growth
+are largely absent.
 
-A further problem with existing models is that the effects of traits are all
-realised via mass production, whereas the effect of some prominent traits is
-most likely on the quality of tissue construction, or the amoutn of mass
-allocated to different tissues. There are two problem here. First, measuring mass
-production is only practical for small plants that can be harvested. For larger plants,
-growth is more often emasured either as diameter or height growth. For example, the
-CTFS monitors diameter growth in 49 large permenenat plots worldwide. Establishing
-the link between traits, mass growth and diamater growth is thus essential for predicting
-the effect of traits on growth across these diverse datasets.
+A further problem with existing theory is that the effects of traits are all
+realised via influences on mass production\citep{enquist_general_2007},
+whereas the effect of traits such as LMA, WD, and Hmax is via the quality of
+tissue construction, or the amount of mass allocated to different tissues
+\citep{falster_influence_2011}. There are in fact two problems with theory
+focussed solely on mass production. The first is that measuring mass
+production is only practical for small plants that can be easily harvested.
+For larger plants, growth is more often measured either as diameter or height
+growth. For example, the CTFS monitors diameter growth in 49 large permanent
+plots across forests worldwide \citep{anderson_teixeira_ctfs_forestgeo_2014},
+while forestry agencies around the world have likewise recorded data across
+large tracts of forest \citep{purves_predictive_2008}. The second problem is
+that, as we outline further below, because mass production is only one
+component of plant growth; theory focussing solely on mass production will
+therefore always be limited.
 
-Here we provide a mechanistic framework describing how traits link to
-demography, linking to commonly used metrics such as height and diameter
-growth, combine this with fresh empirical insight using largest
-collation of growth and trait data available, from long-term plot at
-BCI, Panama.
+In this paper we provide a mechanistic framework for describing the potential
+effect of traits on growth throughout a plant's life-cycle, linking to
+commonly used metrics such as height and diameter growth. To help motivate our
+study we re-analyse demographic and trait data from an earlier study
+\citep{wright_functional_2010}, to show 1) that the link between traits and
+growth is stronger than previously thought, 2) that the influence of three
+prominent traits on growth rate is moderated by plant size, and 3) that the
+impact of size on the nature of relationships varies among traits (see also
+\citet{ruger_functional_2012}. We then show how these results align with new
+set of theoretical predictions. Moreover, we show how traits moderate plant
+responses to light environment and also determine shade tolerance, supporting
+diverse empirical results \citep{ruger_functional_2012, poorter_leaf_2006}. By
+disentangling the effects of plant size, light environment and traits on
+growth rates, we hope to provide a solid theoretical foundation for trait
+ecology and thus provide a platform for understanding growth across diverse
+species around the world.
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Shows why effect of traits changes with size,size dependent effect
-  differs among traits.
+\subsection*{Empirical motivation}\label{fresh-empirical-motivation}
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    Contradicts widespread assumption that traits relate to growth in
-    simple linear way
-  \end{itemize}
-\item
-  Offers framework for interpreting and understanding empirical results
-\end{itemize}
-
-\section*{Fresh empirical motivation}\label{fresh-empirical-motivation}
-
-To help motivate our model, we begin with a fresh analysis of growth dynamics
-in the long-term plot at Barro Colorado Island in Panam, where the growth
-rates for over 100 species have been recorrded over the last 30 years. A
-prior analysis by \citet{wright_functional_2010} found a mdoerate effect of
-wood density on growth rate, but almsot no effect of LMA, and only a weak
-effect of a thrid trait maximum plant height (HMAX).
-
-Fresh empirical analysis highlights how effect of traits on pot growth
-changes with size.
+To help illustrate the empirical patterns, we begin with a fresh analysis of
+growth dynamics in the long-term plot at Barro Colorado Island in Panama,
+where the growth rates for over 100 species have been recorded over the last
+30 years. Data from a long term plot \citep{condit_barro_2012} are combined
+with results of experiment on seedlings \citep{kitajima_leaf_2013}, to better understand
+size-related patterns of growth. A prior analysis by
+\citet{wright_functional_2010} found a moderate effect of wood density on
+growth rate, but almost no effect of LMA, and only a weak effect of HMAX. More recently,
+\citet{ruger_functional_2012} showed
 
@@
   \item
-    show argument hold intuiatively
+    show argument hold intuitively
   \end{itemize}

### 2015-02-18 notes/ms_short.tex
Enhance introduction

@@
 partitioning, and rates of photosynthesis and tissue
-turnover\citep{wright_world-wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
+turnover\citep{wright_world_wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
 such traits influence whole-plant growth, with the simple relationships
@@
 model that links growth rate in seedlings with a single leaf
-trait\citep{lambers_inherent_1992,wright_cross-species_2000} to
+trait\citep{lambers_inherent_1992,wright_cross_species_2000} to
 explicitly include influences of size, light environment, and three
@@
 Functional traits capture core differences in the strategies plants use
-to generate and invest surplus energy\citep{wright_world-wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
+to generate and invest surplus energy\citep{wright_world_wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
 traits now exists for many of the world's 250000 plant
@@
 (also known as leaf mass per area) are widely considered as indicators
-of growth rate\citep{reich_world-wide_2014}, because cheaper leaf
+of growth rate\citep{reich_world_wide_2014}, because cheaper leaf
 construction should enable faster canopy
@@
 growth rate in seedlings is strongly and negatively correlated with
-\(\phi\)\citep{cornelissen_seedling_1996,wright_cross-species_2000}
+\(\phi\)\citep{cornelissen_seedling_1996,wright_cross_species_2000}
 (Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
@@
 Here we extend a widely-used theoretical model that links growth rate in
-seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross-species_2000}, to explicitly include influences of
+seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross_species_2000}, to explicitly include influences of
 size, light environment, and other prominent traits. This extension
@@
 Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to \(\phi\)\citep{wright_cross-species_2000}, this
+relation to \(\phi\)\citep{wright_cross_species_2000}, this
 approximation does not directly link to other traits, or to the
@@
 growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction\citep{wright_world-wide_2004}.
+results from superior leaf construction\citep{wright_world_wide_2004}.
 This is achieved by linking the turnover rate of leaf in eq.
@@
 shown to describe patterns across diverse
-ecosystems\citep{wright_world-wide_2004} (Fig. \ref{fS-leaf}).
+ecosystems\citep{wright_world_wide_2004} (Fig. \ref{fS-leaf}).
 
@@
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
-Data from \citep{wright_world-wide_2004} for 678 species from 51 sites, each
+Data from \citep{wright_world_wide_2004} for 678 species from 51 sites, each
 point giving a species-average. Lines show standardised major axis lines

### 2015-03-17 ms.tex
update bibtex keys

@@
 plants use to generate and invest surplus energy
-\citep{wright_world_wide_2004, chave_towards_2009, westoby_plant_2002}.
+\citep{wright_world-2004, chave-2009, westoby-2002}.
 Although most plants have the same basic construction, physiological function
@@
 mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
-diameter,  and seed size  \citep{wright_world_wide_2004,chave_towards_2009}.
+diameter,  and seed size  \citep{wright_world-2004,chave-2009}.
 These traits are easily measured across diverse species, and variation between
@@
 now exists for up to 10\% of the world's 250000 plant species
-\citep{cornwell_functional_2014}. During the last decade, the idea has taken
+\citep{cornwell-2014}. During the last decade, the idea has taken
 hold that the growth strategy and ecology of a plant species might usefully be
@@
 demographic rates have only met with mixed success
-\citep{wright_functional_2010, poorter_are_2008}. In seedlings, the trait LMA
+\citep{wright-2010, poorter-2008}. In seedlings, the trait LMA
 -- a central part of the leaf economics spectrum
-\citep{wright_world_wide_2004} --is tightly related to plant relative growth
-rate in mass \citep{lambers_inherent_1992, wright_cross_species_2000}. LMA and
+\citep{wright_world-2004} --is tightly related to plant relative growth
+rate in mass \citep{lambers-1992, wright_cross-2000}. LMA and
 its correlate leaf lifespan have also been linked to height growth rate for
-small seedlings and saplings \citep{reich_leaf_1992, poorter_leaf_2006}. These
+small seedlings and saplings \citep{reich-1992, poorter-2006}. These
 early successes prompted researchers to search for similar relationships in
 large plants. However, the results showed that in trees LMA was not correlated
-with diameter growth rate  \citep{wright_functional_2010, poorter_are_2008,
-herault_functional_2011}. Meanwhile, other traits such as wood density showed
-strong relationships to growth in large plants \citep{wright_functional_2010},
-but only weak relationships in seedlings  \citep{castro-diez_stem_1998}. Thus
+with diameter growth rate  \citep{wright-2010, poorter-2008,
+herault-2011}. Meanwhile, other traits such as wood density showed
+strong relationships to growth in large plants \citep{wright-2010},
+but only weak relationships in seedlings  \citep{castro-diez-1998}. Thus
 far, most trait growth relationships have been quantified for one or few
@@
 link between traits and growth may be modified by plant size
-\citep{ruger_functional_2012}; a finding that challenges the wide-spread
+\citep{ruger-2012}; a finding that challenges the wide-spread
 assumption that particular ends of the trait spectrum translate directly into
@@
 plant function. Generating these expectations is a primary, yet perhaps under-
-appreciated, role for theory \citep{kokko_modelling_2007}. Current empirical
+appreciated, role for theory \citep{kokko-2007}. Current empirical
 results suggest the effect of traits on growth changes with size and possibly
@@
 mass-based relative growth rate is linearly and negatively related to LMA
-\citep{lambers_inherent_1992, cornelissen_seedling_1996,
-wright_cross_species_2000}. An extension of the model suggests a similar
-relationship may hold at larger sizes \citep{enquist_general_2007}. But as
+\citep{lambers-1992, cornelissen-1996,
+wright_cross-2000}. An extension of the model suggests a similar
+relationship may hold at larger sizes \citep{enquist-2007}. But as
 described above, this prediction is not supported by empirical results.
@@
 A further problem with existing theory is that the effects of traits are all
-realised via influences on mass production\citep{enquist_general_2007},
+realised via influences on mass production\citep{enquist-2007},
 whereas the effect of traits such as LMA, WD, and Hmax is via the quality of
 tissue construction, or the amount of mass allocated to different tissues
-\citep{falster_influence_2011}. There are in fact two problems with theory
+\citep{falster-2011}. There are in fact two problems with theory
 focussed solely on mass production. The first is that measuring mass
@@
 growth. For example, the CTFS monitors diameter growth in 49 large permanent
-plots across forests worldwide \citep{anderson_teixeira_ctfs_forestgeo_2014},
+plots across forests worldwide \citep{anderson_teixeira_ctfs-2014},
 while forestry agencies around the world have likewise recorded data across
-large tracts of forest \citep{purves_predictive_2008}. The second problem is
+large tracts of forest \citep{purves-2008}. The second problem is
 that, as we outline further below, because mass production is only one
@@
 study we re-analyse demographic and trait data from an earlier study
-\citep{wright_functional_2010}, to show 1) that the link between traits and
+\citep{wright-2010}, to show 1) that the link between traits and
 growth is stronger than previously thought, 2) that the influence of three
@@
 impact of size on the nature of relationships varies among traits (see also
-\citet{ruger_functional_2012}. We then show how these results align with new
+\citet{ruger-2012}. We then show how these results align with new
 set of theoretical predictions. Moreover, we show how traits moderate plant
 responses to light environment and also determine shade tolerance, supporting
-diverse empirical results \citep{ruger_functional_2012, poorter_leaf_2006}. By
+diverse empirical results \citep{ruger-2012, poorter-2006}. By
 disentangling the effects of plant size, light environment and traits on
@@
 where the growth rates for over 100 species have been recorded over the last
-30 years. Data from a long term plot \citep{condit_barro_2012} are combined
-with results of experiment on seedlings \citep{kitajima_leaf_2013}, to better understand
+30 years. Data from a long term plot \citep{condit-2012} are combined
+with results of experiment on seedlings \citep{kitajima-2013}, to better understand
 size-related patterns of growth. A prior analysis by
-\citet{wright_functional_2010} found a moderate effect of wood density on
+\citet{wright-2010} found a moderate effect of wood density on
 growth rate, but almost no effect of LMA, and only a weak effect of HMAX. More recently,
-\citet{ruger_functional_2012} showed
+\citet{ruger-2012} showed
 

### 2015-03-17 notes/ms_short.tex
update bibtex keys

@@
 food-webs, the carbon cycle and much of human
-activity \citep{raupach_carbon_2010}. Plant species differ in many
+activity \citep{raupach-2010}. Plant species differ in many
 functional traits, including aspects of tissue construction and energy
@@
 partitioning, and rates of photosynthesis and tissue
-turnover\citep{wright_world_wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
+turnover\citep{wright_world-2004,chave-2009, westoby-2002}. Yet, it remains unclear how -- and even if --
 such traits influence whole-plant growth, with the simple relationships
 predicted by existing theory often lacking empirical
-support\citep{wright_functional_2010,poorter_are_2008}. Here we
+support\citep{wright-2010,poorter-2008}. Here we
 provide a new set of theoretical predictions, validated using 20 years
-of growth data\citep{condit_barro_2012,kitajima_leaf_2013}, showing
+of growth data\citep{condit-2012,kitajima-2013}, showing
 that the influence of three prominent traits on growth rate is moderated
@@
 model that links growth rate in seedlings with a single leaf
-trait\citep{lambers_inherent_1992,wright_cross_species_2000} to
+trait\citep{lambers-1992,wright_cross-2000} to
 explicitly include influences of size, light environment, and three
@@
 shade tolerance, supporting diverse empirical
-results\citep{ruger_functional_2012,poorter_leaf_2006}. By
+results\citep{ruger-2012,poorter-2006}. By
 disentangling the effects of plant size, light environment and traits on
@@
 Functional traits capture core differences in the strategies plants use
-to generate and invest surplus energy\citep{wright_world_wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
+to generate and invest surplus energy\citep{wright_world-2004, chave-2009,westoby-2002}. Data for prominent
 traits now exists for many of the world's 250000 plant
-species\citep{cornwell_functional_2014}; yet the effect of these traits
+species\citep{cornwell-2014}; yet the effect of these traits
 on growth remains poorly understood. Prominent traits such as \(\phi\)
 (also known as leaf mass per area) are widely considered as indicators
-of growth rate\citep{reich_world_wide_2014}, because cheaper leaf
+of growth rate\citep{reich_world-2014}, because cheaper leaf
 construction should enable faster canopy
-expansion\citep{blackman_compound_1919,westoby_time_2000}. Indeed,
+expansion\citep{blackman-1919,westoby-2000}. Indeed,
 growth rate in seedlings is strongly and negatively correlated with
-\(\phi\)\citep{cornelissen_seedling_1996,wright_cross_species_2000}
+\(\phi\)\citep{cornelissen-1996,wright_cross-2000}
 (Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
 might also give a growth advantage in larger
-plants\citep{enquist_general_2007}, an increasing number of empirical
+plants\citep{enquist-2007}, an increasing number of empirical
 studies have failed to detect any such
-effect\citep{wright_functional_2010, @herault_functional_2011,
-@poorter_are_2008}. Meanwhile, expectations on how other prominent
+effect\citep{wright-2010, @herault-2011,
+@poorter-2008}. Meanwhile, expectations on how other prominent
 traits might influence growth are lacking. The growing discord between
@@
 Here we extend a widely-used theoretical model that links growth rate in
-seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross_species_2000}, to explicitly include influences of
+seedlings to \(\phi\)\citep{lambers-1992, wright_cross-2000}, to explicitly include influences of
 size, light environment, and other prominent traits. This extension
@@
 are compared against data from both a long-term
-plot\citep{condit_barro_2012} and field
-experiment\citep{kitajima_leaf_2013} at Barro Colorado Island, Panama.
+plot\citep{condit-2012} and field
+experiment\citep{kitajima-2013} at Barro Colorado Island, Panama.
 We use a new method to estimate species potential growth rate,
@@
 definite signal of trait effects than in previous
-studies\citep{wright_functional_2010}. The results of this new analysis
+studies\citep{wright-2010}. The results of this new analysis
 (Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
@@
 (total photosynthesis) and losses (respiration and turnover) within the
-plant\citep{falster_influence_2011}:
+plant\citep{falster-2011}:
 
@@
 Current theory emphasises the relationship between \(\phi\) and relative
-growth in mass\citep{enquist_general_2007},
+growth in mass\citep{enquist-2007},
 \(\textrm{d}P / \textrm{d}t / m_\textrm{t}\). For seedlings, which are
@@
 Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to \(\phi\)\citep{wright_cross_species_2000}, this
+relation to \(\phi\)\citep{wright_cross-2000}, this
 approximation does not directly link to other traits, or to the
@@
 Mathematically, height growth can be decomposed into a product of
-physiologically relevant terms\citep{falster_influence_2011}:
+physiologically relevant terms\citep{falster-2011}:
 
@@
 tall, while other species emphasise vertical
-extension\citep{poorter_architecture_2006}.
+extension\citep{poorter-2006}.
 
@@
 To address the first of these challenges, we use a functional balance
-model\citep{yokozawa_foliage_1995,falster_influence_2011} assuming
+model\citep{yokozawa-1995,falster-2011} assuming
 1) an allometric relation between total leaf area and plant height, and
@@
 of plant growth (Table \ref{tab:predictions}a), recovering well-known
-empirical patterns\citep{sillett_increasing_2010, king_size-related_2011}. Height growth shows a hump-shaped pattern
+empirical patterns\citep{sillett-2010, king-2011}. Height growth shows a hump-shaped pattern
 with size, first increasing then decreasing. This pattern results from
@@
 and increasing reproductive allocation. In contrast, basal-area growth
-continues to increase with size\citep{sillett_increasing_2010,
-@stephenson_rate_2014}, due to an increasing influence of stem
+continues to increase with size\citep{sillett-2010,
+@stephenson-2014}, due to an increasing influence of stem
 turnover. Diameter growth shows a weakly hump shaped
-curve\citep{herault_functional_2011}, tapering off slightly at larger
+curve\citep{herault-2011}, tapering off slightly at larger
 sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
@@
 (Fig?). All growth measures decrease sharply with size when expressed as
-relative growth rates\citep{iida_linking_2014}.
+relative growth rates\citep{iida-2014}.
 
@@
 growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction\citep{wright_world_wide_2004}.
+results from superior leaf construction\citep{wright_world-2004}.
 This is achieved by linking the turnover rate of leaf in eq.
@@
 shown to describe patterns across diverse
-ecosystems\citep{wright_world_wide_2004} (Fig. \ref{fS-leaf}).
+ecosystems\citep{wright_world-2004} (Fig. \ref{fS-leaf}).
 
@@
 lower \(\rho\) decreases mechanical strength resulting in higher
-mortality\citep{chave_towards_2009,wright_functional_2010}. Under
+mortality\citep{chave-2009,wright-2010}. Under
 this scenario, decreasing \(\rho\) will always provide a growth
@@
 with low \(rho\) increase growth more substantially with incraeses in
-light\citep{ruger_functional_2012}. This is exactly the behaviour
+light\citep{ruger-2012}. This is exactly the behaviour
 exhibited in our model (Fig. \ref{f-growth_light}), at least for small
@@
 \(rho\) actually decreases as size
-increases\citep{ruger_functional_2012}. In contrast, growth rate tends
+increases\citep{ruger-2012}. In contrast, growth rate tends
 to increase in size for the species with highest
-\(\rho\)\citep{ruger_functional_2012}.
+\(\rho\)\citep{ruger-2012}.
 
@@
 in small plants but not for plants \textgreater{} 0.01m
-dbh\citep{ruger_functional_2012}.
+dbh\citep{ruger-2012}.
 
@@
 needed to offset these costs. As previously suggested
-\citep{givnish_adaptation_1988}, shade-tolerance also decreases with
+\citep{givnish-1988}, shade-tolerance also decreases with
 height because as size increase, the total amount of energy needed to
@@
 finding matches well known empirical patterns, where both \(\phi\) and
-\(\rho\) have been linked to shade tolerance\citep{poorter_leaf_2006, lusk_ontogenetic_2008,osunkoya_light_1996}.
+\(\rho\) have been linked to shade tolerance\citep{poorter-2006, lusk-2008,osunkoya-1996}.
 
@@
   in fact, explains plasticity in LMA with size, where LMA tends to
-  increase with size \citep{thomas_asymptotic_1999;
-  @thomas_photosynthetic_2010}
+  increase with size \citep{thomas-1999;
+  @thomas-2010}
 \item
@@
 
-Note, scaling relationship reversed from \citep{falster_influence_2011}.
+Note, scaling relationship reversed from \citep{falster-2011}.
 
@@
 
-We follow the model of\citep{yokozawa_foliage_1995} describing the
+We follow the model of\citep{yokozawa-1995} describing the
 vertical distribution of leaf area within the crowns of individual
@@
 We also assume that each unit of sapwood area supports a fixed area of
-leaf (the pipe model, \citep{shinozaki_quantitative_1964}), so that the total
+leaf (the pipe model, \citep{shinozaki-1964}), so that the total
 canopy area of a plant relates to basal sapwood area \(S(0,h)\):
@@
 where
-\(\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}\)\citep{yokozawa_foliage_1995}.
+\(\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}\)\citep{yokozawa-1995}.
 Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
@@
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
-Data from \citep{wright_world_wide_2004} for 678 species from 51 sites, each
+Data from \citep{wright_world-2004} for 678 species from 51 sites, each
 point giving a species-average. Lines show standardised major axis lines

### 2015-04-28 ms.tex
Add reference from Amy
Update refernce style
Introduces numbers in formatting of bibliography, as required in latest
changes for Ecol Letters. This change was achieved via patch deacribed here
http://tex.stackexchange.com/questions/207597/include-numbering-for-reference-list-ecology-letters-bst-file-output-bibitem-mo
Expand main paper

@@
 \usepackage{authblk} % allows for better author affiliations
+
+%appendix
+\usepackage[title,titletoc,toc]{appendix}
+\usepackage{setspace}  % for double spacing
 
@@
 %\rfoot{\thepage}
+
+% allows for numbered referencing, as refquired by ecol letters
+\usepackage{etoolbox}
+\newbool{MyRefNumbers}
+\booltrue{MyRefNumbers} % comment to remove numbers in reference list
 
@@
 
-\section*{Abstract}\label{abstract}
+\section{To tidy}
 
-\section*{Introduction}\label{introduction}
+mt: living tissue or total?
 
-Functional traits are thought to capture core differences in the strategies
+theta: change to theta$^{-1}$ -- nearly always a negative power
+
+h-al: changed relationship to work other way around - this reads better, but is perhaps opposite of implementation, which is driven by height
+
+\section{Abstract}\label{abstract}
+
+\section{Introduction}\label{introduction}
+
+Functional traits capture core differences in the strategies
 plants use to generate and invest surplus energy
  \citep{wright_world-2004, chave-2009, westoby-2002}.
-Although most plants have the same basic construction, physiological function
-and resource requirements, species differ considerably from one another in
-rates of carbon, nitrogen and water uptake, growth, mortality and seed
-production. Some of these functional differences have been linked to
-underlying variation in relatively simple physiological traits, such as leaf
+Although most plants have the same basic physiological function
+and key resource requirements (carbon, nitrogen and water), species differ considerably in
+rates at which resources are acquired and invested into different tissues.
+Much of this functional variation relates to underlying
+differences in some prominent traits, including leaf
 mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
-diameter,  and seed size  \citep{wright_world-2004,chave-2009}.
-These traits are easily measured across diverse species, and variation between
-species is large compared to variation within species. Data for some traits
-now exists for up to 10\% of the world's 250000 plant species
-\citep{cornwell-2014}. During the last decade, the idea has taken
-hold that the growth strategy and ecology of a plant species might usefully be
-characterised simply by knowing its traits.
+diameter,  maximum  height (HMAX) and seed size  \citep{wright_world-2004,chave-2009}.
+Data for these traits now exists for up to 10\% of the world's 250000 plant species
+\citep{cornwell-2014}. Increasingly, researchers have been looking to traits
+as potential indicators of the growth strategy and ecology of a species.
 
@@
 function is increasingly understood, attempts at using traits to understand
-demographic rates have only met with mixed success
+demographic rates have met with mixed success
  \citep{wright-2010, poorter-2008}. In seedlings, the trait LMA
--- a central part of the leaf economics spectrum
-\citep{wright_world-2004} --is tightly related to plant relative growth
-rate in mass \citep{lambers-1992, wright_cross-2000}. LMA and
+-- the central element of the leaf economics spectrum
+ \citep{wright_world-2004} -- has been tightly related to relative growth
+rate in plant mass \citep{lambers-1992, wright_cross-2000}. LMA and
 its correlate leaf lifespan have also been linked to height growth rate for
@@
 link between traits and growth may be modified by plant size
-\citep{ruger-2012}; a finding that challenges the wide-spread
-assumption that particular ends of the trait spectrum translate directly into
-faster or slower growth.
+ \citep{ruger-2012}. This finding challenges a wide-spread
+assumption regarding traits: that particular ends of the trait spectrum translate directly into
+faster or slower growth (refs XX).
 
-The challenge of interpreting diverse empirical results linking  -- or not --
-traits to growth rate would clearly be easier with clear expectations on what
-signal one should expect, given current understandings of how traits influence
-plant function. Generating these expectations is a primary, yet perhaps under-
-appreciated, role for theory \citep{kokko-2007}. Current empirical
+The challenge of interpreting diverse empirical results seeking to link
+traits to growth rate is made all the harder because we currently lack clear
+expectations on the expected signal. Generating such expectations is one of the
+primary role for theory \citep{kokko-2007}. Current empirical
 results suggest the effect of traits on growth changes with size and possibly
@@
 A further problem with existing theory is that the effects of traits are all
-realised via influences on mass production\citep{enquist-2007},
-whereas the effect of traits such as LMA, WD, and Hmax is via the quality of
-tissue construction, or the amount of mass allocated to different tissues
-\citep{falster-2011}. There are in fact two problems with theory
+realised via influences on mass production \citep{enquist-2007},
+whereas the effect of traits such as LMA, WD, and Hmax is via allocated
+to different tissues \citep{falster-2011}. There are in fact two problems with theory
 focussed solely on mass production. The first is that measuring mass
 production is only practical for small plants that can be easily harvested.
-For larger plants, growth is more often measured either as diameter or height
-growth. For example, the CTFS monitors diameter growth in 49 large permanent
-plots across forests worldwide \citep{anderson_teixeira_ctfs-2014},
-while forestry agencies around the world have likewise recorded data across
-large tracts of forest \citep{purves-2008}. The second problem is
+For larger plants, growth is more often measured as increment in either diameter
+or height. For example, the CTFS monitors diameter growth in 49 large permanent
+plots across forests worldwide \citep{anderson-teixeira-2015};
+ forestry have likewise recorded growth in diameter across
+large tracts of the globe \citep{purves-2008}. The second problem is
 that, as we outline further below, because mass production is only one
 component of plant growth; theory focussing solely on mass production will
-therefore always be limited.
+ always be limited.
 
@@
 prominent traits on growth rate is moderated by plant size, and 3) that the
-impact of size on the nature of relationships varies among traits (see also
-\citet{ruger-2012}. We then show how these results align with new
-set of theoretical predictions. Moreover, we show how traits moderate plant
+impact of size on the nature of relationships varies among traits
+ \citep[see also]{ruger-2012}. We then show how these results align with new
+set of theoretical predictions. Analysis extends ...
+Moreover, we show how traits moderate plant
 responses to light environment and also determine shade tolerance, supporting
@@
 
-\subsection*{Empirical motivation}\label{fresh-empirical-motivation}
+The results of this new analysis
+(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
+\ref{tab:predictions}) showing how the effect of traits on growth
+changes with plant size, with different traits exhibiting distinct
+size-dependent effects.
+
+\subsection{Empirical motivation}\label{fresh-empirical-motivation}
 
@@
 
-\section*{Materials and Methods}\label{materials-and-methods}
+Predictions
+are compared against data from both a long-term
+plot \citep{condit-2012} and field
+experiment \citep{kitajima-2013} at Barro Colorado Island, Panama.
+We use a new method to estimate species potential growth rate,
+accounting for size specific effects (Fig. \ref{S-quantile_examples},
+see Methods for details). This new method allowed us to extract a more
+definite signal of trait effects than in previous
+studies \citep{wright-2010}.
 
-\subsection*{Conceptual framework}\label{conceptual-framework}
+\section{Materials and Methods}\label{materials-and-methods}
+
+\subsection{Growth data}\label{growth-data}
+
+\subsection{Conceptual framework}\label{conceptual-framework}
+
+To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in
+seedlings to LMA \citep{lambers-1992, wright_cross-2000}, to explicitly include influences of
+size, light environment, and other prominent traits. This model
 
@@
 
-Decomposition of diameter and height growth
+First derive ... generally true.
 
-Functional-balance Model of growth
+Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
+therefore hold for any model of plant growth. To make explicit
+predictions, however, requires two additional elements: 1) A specific
+model describing how the various mass and area terms
+($m_\textrm{l}, m_\textrm{s}, m_\textrm{b}, m_\textrm{r}, a_\textrm{l}, a_\textrm{s}, a_\textrm{b}$)
+vary relative to one another; and 2) For any trade-offs in plant
+function to be specified; this is where traits have an influence.  These two components are outlined in following sections.
+
+\subsubsection{Standard model for mass production}
+
+We begin with a standard model for the amount of biomass available for
+growth, $dP / \textrm{d}t$, given by the difference between income
+(total photosynthesis) and losses (respiration and turnover) within the
+plant \citep{makela-1997, Thornley-2000, falster-2011}:
+
+\begin{equation}\label{eq:dPdt}
+\underbrace{\strut \frac{dP}{\textrm{d}t}}_\textrm{net biomass production}
+  = \underbrace{\strut Y}_\textrm{yield}
+    \big( \underbrace{\strut a_\textrm{l} \, A}_\textrm{photosynthesis} -
+     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
+    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
+\end{equation}
+
+Here, $m,r$, and $k$ refer to the mass, respiration rate, and
+turnover rate of different tissues, denoted by subscripts $l$=leaves,
+$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
+rate of CO$_2$ per leaf area and $Y$ is yield: the fraction of
+assimilated carbon fixed in biomass (the remaining fraction being lost
+as growth respiration) (see Table \ref{tab:params} for units and
+definitions). Photosynthesis is proportional to leaf area,
+$a_\textrm{l} = m_\textrm{l} / \phi$, where $\phi$ is leaf mass per area.
+The total mass of living tissues $m_\textrm{t}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.$
+
+The only potentially controversial element of in eq.
+\ref{eq:dPdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
+This assumption is likely to hold true for most woody plants, because the .....
+However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegtative part of the plant (refs XX) \citep{ Thornley-2000}.
+
+\subsubsection{Decomposing height, basal area and diameter growth}
+
+To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
+Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{falster-2011}:
+
+\begin{equation} \label{eq:dhdt}
+\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
+\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}}_\textrm{marginal leaf deployment}
+\times \underbrace{\strut\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}}_\textrm{allocation to growth}
+\times \underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}}_\textrm{mass production}.
+\end{equation}
+
+The first term on the right of eq \ref{eq:dhdt},
+$\textrm{d}h / \textrm{d}a_\textrm{l}$, is the growth in plant height
+per unit growth in total leaf area -- accounting for the architectural
+strategy of the plant. Some species tend to leaf out more than grow
+tall, while other species emphasise vertical
+extension \citep{poorter-2006}.
+
+The second term, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$,
+accounts for the marginal cost of deploying an additional unit of leaf
+area, including construction of the leaf itself and various support
+structures. As such, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$
+can itself be expressed as a sum of construction costs per unit leaf
+area:
+
+\begin{equation}\label{eq:daldmt}
+\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
+= \big(\phi
+ + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
+\end{equation}
+
+The third term in eq \ref{eq:dhdt},
+$\textrm{d}m_\textrm{t} / \textrm{d}P$, gives the fraction of net
+biomass production (eq. \ref{eq:dPdt}) that is allocated to growth
+rather than reproduction or storage.
+
+In a similar way, basal area ($a_\textrm{st}$) increment can be
+expressed as the sum of increments in sapwood, bark \& heartwood areas
+($a_\textrm{ss}, a_\textrm{sb}, a_\textrm{sh}$ respectively):
+$\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}= \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}$.
+We now have an equation for basal area growth that conatins many of the same
+elements as eq. \ref{eq:dhdt}:
+
+\begin{equation}\label{eq:dast}
+\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
+\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
+\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}} \times \frac{\textrm{d}m_\textrm{t}}{\textrm{d}P} \times \frac{\textrm{d}P}{\textrm{d}t} +\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} .
+\end{equation}
+
+Diameter growth is then given by the geometric relationship between stem
+diameter ($D$) and $a_\textrm{st}$:
+
+\begin{equation} \label{eq:dDdt}
+\frac{\textrm{d}D}{\textrm{d}t}= \sqrt{\frac{\pi}{a_\textrm{st}}} \frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}.
+\end{equation}
+
+\subsubsection{A functional-balance model for allocation}
+
+To address the first of these challenges, we use a functional balance
+model \citep{yokozawa-1995,falster-2011} assuming
+1) an allometric relation between total leaf area and plant height, and
+2) a constant ratio between total leaf area and each of sapwood
+cross-sectional area, bark cross-sectional area, and total root mass.
+These assumptions, which are well-supported by data (Fig.
+\ref{f-assumptions}), lead immediately to an allometric model for plant
+size (Table \ref{tab:allometry}, see Supplementary material for
+derivations). Substituting from Table \ref{tab:allometry} into eqs.
+\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
+for plant growth in relation to size.
 
@@
 
-Embedding trait-related trade-offs
+\subsubsection{Embeeding trait-related trade-offs}
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
 use table with arrows to
-\end{itemize}
 
-Connection to earlier work
+By definition, the main effect of increasing $\phi$ is to decrease the
+marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
+growth rate. However, to fully account for the influence of $\phi$ on
+growth we must also account for the decrease in leaf turnover that
+results from superior leaf construction \citep{wright_world-2004}.
+This is achieved by linking the turnover rate of leaf in eq.
+\ref{eq:dPdt} to $\phi$. We use an allometric scaling relationship of
+the form $k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
+shown to describe patterns across diverse
+ecosystems \citep{wright_world-2004} (Fig. \ref{fS-leaf}).
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  generalised version
-\item
-  present old model as dp/dt so consistent with ours
-\end{itemize}
+As for leaf, cheaper stem construction (lower $\rho$) decreases the
+cost of deploying a unit of leaf area, and may thereby increase growth
+rates. However, the cost of cheap stem construction must also be
+accounted for. However, the direct physiological trade-offs of lower
+$\rho$ are less understood than for $\phi$. One possibility is that
+lower $\rho$ decreases mechanical strength resulting in higher
+mortality \citep{chave-2009,wright-2010}. Under
+this scenario, decreasing $\rho$ will always provide a growth
+advantage, because the costs of low $\rho$ are not realised within the
+terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
+\ref{tab:trade-offs}). In addition, lower $\rho$ may increase turnover
+of sapwood. Mirroring our assumptions for leaves, we let sapwood
+turnover decrease with $\rho$,
+i.e.$k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
 
-\subsection*{Analysis}\label{analysis}
+Unlike leaf and stem traits, height at maturation ($h_m$) moderates
+growth by adjusting the amount of energy invested in growth, i.e.~the
+term $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$ in eqs.
+\ref{eq:dhdt} and \ref{eq:dast}.
 
-\subsection*{Growth data}\label{growth-data}
+\subsection{Analysis}\label{analysis}
 
-\section*{Results}\label{results}
+\newpage
+\section{Results and Discussion}\label{results}
 
-\subsection*{Changes in growth rate with
-size}\label{changes-in-growth-rate-with-size}
+\subsection{Changes in growth rate with size}\label{changes-in-growth-rate-with-size}
 
-\subsection*{How traits influence growth
-rate}\label{how-traits-influence-growth-rate}
+This new growth model captures the intrinsically size-dependent nature
+of plant growth (Table \ref{tab:predictions}a), recovering well-known
+empirical patterns \citep{sillett-2010, king-2011}. Height growth shows a hump-shaped pattern
+with size, first increasing then decreasing. This pattern results from
+systematic changes in the four components of eq. \ref{eq:dhdt} with size
+(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
+that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
+and increasing reproductive allocation. In contrast, basal-area growth
+continues to increase with size \citep{sillett-2010,
+@stephenson-2014}, due to an increasing influence of stem
+turnover. Diameter growth shows a weakly hump shaped
+curve \citep{herault-2011}, tapering off slightly at larger
+sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
+mainly because of increased reproductive allocation in older trees
+(Fig?). All growth measures decrease sharply with size when expressed as
+relative growth rates \citep{iida-2014}.
 
-\subsubsection*{LMA}\label{lma}
+\subsection{How traits influence growth rate}\label{how-traits-influence-growth-rate}
 
-\subsubsection*{Wood density}\label{wood-density}
+\subsubsection{LMA}\label{lma}
 
-\subsubsection*{HMax}\label{hmax}
+Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
+\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
+changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
+Decreasing $\phi$ has two impacts on growth rate. First, lower
+$\phi$ increases marginal leaf deployment
+($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$ by economising on
+construction costs. Second, lower $\phi$ decreases net production
+($\textrm{d}P / \textrm{d}t$), due to increased leaf turnover. Whether
+lower $\phi$ increases growth thus depends on the relative magnitude
+of these two effects. When plants are small the effect on leaf
+deployment rate is bigger and so decreasing $\phi$ increases growth
+rate. When plants are large, the influence of $\phi$ on leaf
+deployment rate is diminished, because the costs of building other
+supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
+\ref{f-mass_fraction}). The net result is that at larger sizes, low
+$\phi$ is no longer advantageous for growth (Fig.
+\ref{f-lma_growth_size}d).
 
-\section*{Discussion}\label{discussion}
+\subsubsection{Wood density}\label{wood-density}
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  mechanistic model of height and dbh growth which accounts for effects
-  of size, light environment and traits
+Under this scenario, decreasing $\rho$ provides an advantage in height
+growth only when the benefits of low $\rho$ outweigh the costs. This
+effect is strongest at intermediate sizes. The decline in height growth
+at low $\rho$ with size is more marked than for either stem basal-area
+and diameter growth, because high sapwood turnover also contributes to
+this growth measure. Thus diameter growth continues, even as net
+production approaches zero.
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    need to account for size long recognized, but not straightforward
-  \item
-    look at growth in wide range of measures: mass, height, dbh
-  \end{itemize}
-\item
-  Compared to previous work, shows effects of traits more about
-  allocation than mass production
-\end{itemize}
+\subsubsection{hmat}\label{hmax}
 
-\subsection*{Tree growth is more than just
-photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}
+ Greater height at maturation ($h_m$)
+can thus lead to a growth advantage by increasing
+$\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$ (Table
+\ref{tab:predictions}). At smaller sizes, there is no differentiation
+among species, because most species are focusing on growth (Fig.
+\ref{f-BCI}). At larger sizes, individuals of some species are
+allocating a majority of their surplus energy to growth, leaving a
+strong signal in potential growth rate (Fig. \ref{f-BCI}).
+
+\subsubsection{Responisveness to light}\label{hmax}
+
+Two other important predictions arise directly from linking traits to
+growth rate (Table \ref{tab:predictions}c). First, traits moderate the
+responsiveness of growth to changes in light environment. This response
+arises simply because species with higher potential growth rate have
+greater potential plasticity. Recent analyses have shown that species
+with low $rho$ increase growth more substantially with incraeses in
+light \citep{ruger-2012}. This is exactly the behaviour
+exhibited in our model (Fig. \ref{f-growth_light}), at least for small
+to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
+of low $\phi$ are lost (Fig. \ref{f-growth_light}). This results also
+agrees with empirical results showing that growth for species with low
+$rho$ actually decreases as size
+increases \citep{ruger-2012}. In contrast, growth rate tends
+to increase in size for the species with highest
+$\rho$ \citep{ruger-2012}.
+
+Variation in $\phi$ also moderates the response of growth to changes
+in light, with species having the lowest $rho$ being most responsive.
+However, unlike for $\rho$, the effect appears only for the smallest
+size classes. Accordingly, empirical analyses have detected such a link
+in small plants but not for plants \textgreater{} 0.01m
+dbh \citep{ruger-2012}.
+
+\subsection{Growth is more than just photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}
+
+Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
+\citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
+$\textrm{d}P / \textrm{d}t / m_\textrm{t}$. This relationship can be derived from eq. \ref{eq:dPdt}  by ...
+(see Appendix \ref{app:traits-RGR}).
 
@@
 
-\subsection*{Towards a global model of tree
-growth}\label{towards-a-global-model-of-tree-growth}
+Plasticity in trait values predicted from model:
 
-Key patterns explained
+@Osazuwa-Peters-2014 - indreases in WD through ontogeny.
 
-Data challenges
+\subsection{Shade tolerance}\label{how-traits-influence-growth-rate}
 
-Reproductive allocation
+Second, species with low $\phi$ and low $\rho$ are predicted to be
+less shade tolerant (Fig. \ref{f-wplcp}). At low $\phi$ ($\rho$),
+leaf (sapwood) turnover is higher and thus a greater light income is
+needed to offset these costs. As previously suggested
+ \citep{givnish-1988}, shade-tolerance also decreases with
+height because as size increase, the total amount of energy needed to
+offset respiratory and turnover costs in the stem also increases. These
+finding matches well known empirical patterns, where both $\phi$ and
+$\rho$ have been linked to shade tolerance \citep{poorter-2006, lusk-2008,osunkoya-1996}.
 
-\subsection*{Detecting trait signals in
-data}\label{detecting-trait-signals-in-data}
+Results fo poort & binger for surival strongest for LL.
 
-Need more than simply look for correlation with mean growth rate
+Also baltzer.
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  understand influences of size on trait of interest
-\item
-  traits define potential growth rate --\textgreater{} how to extract
-  this
-\item
-  ideally start to use mechanistic models
-\end{itemize}
-
-\subsection*{Implications for trait-based
-approaches}\label{implications-for-trait-based-approaches}
+\subsection{Implications for trait-based approaches}\label{implications-for-trait-based-approaches}
 
@@
 
-\section*{Conclusion}\label{conclusion}
-
-\newpage
-
-\section*{Tables}\label{tables}
-
-\textbf{Table 1 - Key equations of growth model}
+Towards a global model of growth:
 
@@
 \item
-  key parameters of model
-\end{itemize}
-
-\textbf{Table 2 - effect of traits on demography}
+  mechanistic model of height and dbh growth which accounts for effects
+  of size, light environment and traits
 
@@
   \item
-  combines previous tables 1 \& 2
+    need to account for size long recognized, but not straightforward
+  \item
+    look at growth in wide range of measures: mass, height, dbh
+  \end{itemize}
+\item
+  Compared to previous work, shows effects of traits more about
+  allocation than mass production
+\end{itemize}
+
+Key patterns explained
+
+Data challenges
+
+Reproductive allocation
+
+Detecting trait signals in data.
+
+\begin{itemize}
+\item  Need more than simply look for correlation with mean growth rate
+\item
+  understand influences of size on trait of interest
+\item
+  traits define potential growth rate --\textgreater{} how to extract
+  this
+\item
+  ideally start to use mechanistic models
 \end{itemize}
@@
 
-\section*{Figures}\label{figures}
+\section{Tables}\label{tables}
+
+\begin{table}[h]
+\caption{Equations for an allometric growth model, based on functional balance assumption in {\texbbf a}. }
+\centering
+
+{\footnotesize  % smaller font, double space
+ \begin{doublespace}
+  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
+  Variable & Function & Marginal cost & Growth rate\\ \hline
+  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
+  height &
+    $h = \alpha_1 \, a_\textrm{l}^{\beta_1}$ &
+    $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} =  \beta_1 \alpha_1 a_\textrm{l}^{\beta_1-1}$ &
+    $\frac{\textrm{d}h}{\textrm{d}t}  = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+  sapwoood area &
+    $a_\textrm{ss} = \theta \, a_\textrm{l}$ &
+    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} = \theta$ &
+    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t}  =\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+  bark area &
+    $a_\textrm{sb} = b \, \theta \, a_\textrm{l}$ &
+    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} = b \, \theta$ &
+    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} = \frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\  \\
+  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
+  leaf mass &
+    $m_\textrm{l} = \phi \, a_\textrm{l} $ &
+    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}} = \phi$ &
+    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\
+  sapwood mass &
+    $m_\textrm{ss} = \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
+    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} = \theta\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
+    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\
+  bark mass&$m_\textrm{sb} = b\, \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
+    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} = b \, \theta \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
+    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\
+
+  root mass &
+    $m_\textrm{r} = \alpha_3 \, a_\textrm{l}$ &
+    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}} = \alpha_3$  &
+    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\ \\
+
+  \multicolumn{4}{l}{\textbf{c) Integrated quantities}} \\
+  heartwood area &
+    $a_\textrm{sh} = \int_0^t \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ &
+     &
+    $\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} \, a_\textrm{ss}$ \\
+  heartwood mass &
+    $m_\textrm{sh} = \int_0^t \frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ &
+     &
+    $\frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} \, m_\textrm{ss}$ \\
+  \hline \\
+\end{tabular}
+NB -  theta is sapwood area per leaf area. In tree 2 has been opposite, but that makes notation more clumsy.
+\end{doublespace}}
+\label{tab:allometry}
+\end{table}
+
+\newpage
+
+\begin{table}[h]
+\caption{Effect of traits and related trade-offs on demography}
+\centering
+{\footnotesize  % smaller font, double space
+  \begin{doublespace}
+  \begin{tabular}[c]{l|c|ccccc|c|c}
+  \multicolumn{9}{c}{}\\ \hline
+  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
+  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
+  & $h_0$
+  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$
+  & $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$
+  & $\frac{\textrm{d}P}{\textrm{d}t}$
+  & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
+  & $\frac{1}{k_s}$ & &  \\\hline
+  \textbf{This study}&&&&&&&&\\
+  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
+  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
+  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
+  &&&&&&&&\\\hline
+  \textbf{Possible extensions}&&&&&&&&\\
+  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
+  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
+  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
+  \end{tabular}
+  \end{doublespace}
+}
+\label{tab:trade-offs}
+\end{table}
+
+\newpage
+
+\section{Figures}\label{figures}
 
@@
 in Panama, we estimated the potential growth rate of individual's in
-each species at a series of diameters \(D\), indicated along right hand
+each species at a series of diameters $D$, indicated along right hand
 side. The size of circles in each panel indicates the number of data
@@
 \newpage
+\begin{appendices}
+\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
+
+\section{Relationship between traits and relative growth rate} \label{app:traits-RGR}
+
+We start with eq. \ref{eq:dPdt} for mass-based growth. For seedlings, which are
+young and mostly leaf, it is reasonable to ignore all turnover terms as
+well as the respiration terms for non-leaf tissues. Net production then
+becomes a linear function of leaf area and net photosynthesis per leaf
+area ($A_\textrm{net} = Y(A - r_\textrm{l})$), making relative growth
+rate a linear function of $\phi$:
+
+\begin{equation}\label{eq:RGR}
+\underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}\frac{1}{m_\textrm{t}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{t}}.
+\end{equation}
+
+Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
+relation to $\phi$\citep{wright_cross-2000}, this
+approximation does not directly link to other traits, or to the
+variables that are routinely collected for large trees: namely plant
+height ($h$) and stem cross-sectional area ($a_\textrm{st}$) or
+diameter $D$.
+
+\section{Derivation of a simple allometric model of plant
+function}\label{app:func_balance}
+
+Here we describe an allometric model linking the various size dimensions
+of a plant required by most ecologically realistic vegetation models
+(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
+roots) to a plant height. Table \ref{tab:allometry} provides a summary
+of the derived model, while Table \ref{tab:params} provdies estimates on
+key parameters.
+
+\subsection{Leaf area}\label{leaf-area}
+
+Based on empirically observed allometries (see main text), we assume an
+allometric log-log scaling relationship between the accumulated leaf
+area of a plant and its height:
+
+\begin{equation}\label{eq:ha}
+a_\textrm{l}=\alpha_1 \, h^{\beta_1}.
+\end{equation}
+
+Note, scaling relationship reversed from \citep{falster-2011}.
+
+\subsection{Mass of sapwood}\label{mass-of-sapwood}
+
+We follow the model of \citep{yokozawa-1995} describing the
+vertical distribution of leaf area within the crowns of individual
+plants. This model can account for a variety of canopy profiles through
+a single parameter $\eta$. Setting $\eta=1$ results in a conical
+canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
+, give a top-weighted canopy profile similar to those seen among
+angiosperms. Let $S(z,h)$ be the sapwood area at height $z$ for a
+plant with top height $h$. Following Yokozawa and Hara (1995) we
+assume a relationship between $S(z,h)$ and height such that
+
+\begin{equation}\label{eq:crown1}
+\frac{S(z,h)}{S(0,h)}= \big(1-\big(\frac{z}{h}\big)^\eta\big)^2.
+\end{equation}
+
+We also assume that each unit of sapwood area supports a fixed area of
+leaf (the pipe model, \citep{shinozaki-1964}), so that the total
+canopy area of a plant relates to basal sapwood area $S(0,h)$:
+
+\begin{equation}\label{eq:crown2}
+\frac{m_\textrm{l}}{\phi}= \theta \, S(0,h).
+\end{equation}
+
+Integrating $S(z,h)$ gives a solution for the total mass of sapwood in
+the plant:
+
+\begin{equation}\label{eq:ms1}
+m_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, \textrm{d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}
+
+where
+$\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}$ \citep{yokozawa-1995}.
+Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
+expression for sapwood mass as a function leaf area and height:
+
+\begin{equation}\label{eq:ms2}
+m_\textrm{s}=\rho \, \eta_c \, \theta \, a_\textrm{l} \, h.
+\end{equation}
+
+\subsection{Bark mass}\label{bark-mass}
+
+Bark and phloem tissue are modelled using an analogue of the pipe model,
+leading to a similar equation as that for sapwood mass (eq.
+\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
+assumed to be a constant fraction $b$ of sapwood area per unit leaf
+area such that
+
+\begin{equation}\label{eq:mb}
+m_\textrm{b}=b m_\textrm{s}.
+\end{equation}
+
+\subsection{Root mass}\label{root-mass}
+
+Also consistent with pipe-model assumption, we assume a fixed ratio of
+root mass per unit leaf area
+
+\begin{equation}\label{eq:mr}
+m_\textrm{r}=\alpha_3 \, a_\textrm{l}.
+\end{equation}
+
+Even though nitrogen and water uptake are not modelled explicitly,
+imposing a fixed ratio of root mass to leaf area ensures that
+approximate costs of root production are included in calculations of
+carbon budget.
+
+\section{Effect of traits on height growth} \label{app:traits_max}
+
+We want to show how two traits $\phi$ and $\rho$ influence growth rate,
+by looking at the effect of
+traits on the components of growth rate (eq. \ref{eq:dhdt}. To make
+the analysis more tractable, we will assume we are dealing with a plant of
+given height where 100\% of available energy is allocated to growth,
+i.e. $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}=1$. Note also
+that by the assumptions of our model and above, traits do not influence
+$\frac{\textrm{d}h} {\textrm{d}a_\textrm{l}}$. Thus we have
+
+\begin{equation} \label{eq:G2}
+G = \frac{\textrm{d}h}{\textrm{d}t} = c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}  \frac{ \textrm{d}P} {\textrm{d}t} \right),
+\end{equation}
+
+where $c_1 = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}$. Eq.
+\ref{eq:G2} is a product of the form $Y(x) = W(x) \times Z(x)$. For,
+equations of this type, the derivative $\partial{Y}/\partial{x}$ is
+given by $\partial{W}/\partial{x} \, Z + W \, \partial{Z}/\partial{x}$. Thus
+the derivative of $G$ with respect to some trait $x$  is given by
+
+\begin{equation} \label{eq:dG}
+\frac{\partial G} {\partial x} =
+\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial x}
+ \, \frac{ \textrm{d}P} {\textrm{d}t}
++ \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\, \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}.
+\end{equation}
+
+If this derivative is positive (negative), there will be a positive (negative) relationship
+between traits and growth at the height where it is evaluated.
+
+We further simplify by expanding the derivative of eq. \ref{eq:daldmt}:
+
+\begin{equation} \label{eq:da_dx}
+\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}
+{\partial x} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\ \right)^2
+\, \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}.
+\end{equation}
+
+Combing eqs. \ref{eq:dG}-\ref{eq:da_dx}, we have
+
+\begin{equation} \label{eq:dG2}
+\frac{\partial G} {\partial x} =
+\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\left(
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
+- \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}
+ \, \frac{ \textrm{d}P} {\textrm{d}t}
+\right).
+\end{equation}
+
+Since is $\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}>0$ always, the sign
+of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
+RHS. Rearranging we find that $\frac
+{\partial G} {\partial x} < 0$ if
+
+\begin{equation}\label{eq:dg3}
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
+< \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x} \, \frac{ \textrm{d}P} {\textrm{d}t},
+\end{equation}
+
+or equivalently
+\begin{equation}\label{eq:dg4}
+\frac{
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
+{\frac{ \textrm{d}P} {\textrm{d}t}}
+<
+\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
+\end{equation}
+
+In words, there will be a negative relationship between trait x and growth if the
+relative change in production due to change in traits is less than the relative change
+in the total construction cost for a unit of leaf area due to change in trait. Crucially,
+key elements in eq. \ref{eq:dg3} are modified by height.
+
+We can also solve for trait values $x$ maximising growth rate, i.e. the value
+giving $\partial G /\partial x = 0$:
+
+\begin{equation}\label{eq:max}
+\frac{
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
+{\frac{ \textrm{d}P} {\textrm{d}t}}
+=
+\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
+\end{equation}
+
+\subsection{Leaf construction cost}
+
+If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand
+eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{\textrm{d}m_\textrm
+{t}} {\textrm{d}a_\textrm{l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dPdt} we obtain,
+
+\begin{equation}\label{eq:dPdt2}
+\frac{dP}{\textrm{d}t} = c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}
+\end{equation}
+
+where
+$c_2 = Y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, hmax\textrm{i}})$
+is independent of $\phi$, and thus
+
+\begin{equation}\label{eq:dPdt3}
+\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi}  =
+(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}.
+\end{equation}
+
+Eq. \ref{eq:dPdt3} shows that if $B4=1$, $\frac{ \textrm{d}P} {\textrm{d}t}$ is independent
+of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
+{d}P} {\textrm{d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
+
+Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
+with traits if
+
+\begin{equation} \label{eq:G6}
+\frac{(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}}{c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}}
+< \frac{1}{\phi
+ + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm
+ {b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}}.
+\end{equation}
+
+\subsection{Wood construction cost}
+
+Do similar to previous section for wood construction cost.
+
+\newpage
+
+\section{Supplementary tables}\label{supplementary-tables}
+
+\begin{table}[h]
+\caption{Model parameters}
+\centering
+{\footnotesize  % smaller font, double space
+\begin{doublespace}
+\include{figures/table-pars}
+
+\end{doublespace}
+}
+\label{tab:params}
+\end{table}
+
+\newpage
+
+\section{Supplementary figures}\label{supplementary-figures}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_quantile_examples.pdf}
+\caption{\textbf{Estimating potential growth rate via quantile
+regression.} Example for 4 species. Growth rate estimated at six sizes,
+by fitting lines within given size intervals. Lines fitted to 99th
+quantile, but only when n \textgreater{} 500. Except for left-hand
+panel, estimated points in centre of interval.
+\label{fS-quantile_examples}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_lma_tradeoff.pdf}
+\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
+Data from \citep{wright_world-2004} for 678 species from 51 sites, each
+point giving a species-average. Lines show standardised major axis lines
+fitted to data from each site, with intensity of shading adjusted
+according to strength of the relationship.\label{fS-leaf}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_size_dhdt.pdf}
+\caption{\textbf{Hump-shaped relationship between height growth rate and
+size.} \label{f-hump}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_size_dDdt.pdf}
+\caption{\textbf{Hump-shaped relationship diameter growth and size.}
+\label{f-SI_size_dDdt}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_size_dDdt.pdf}
+\caption{\textbf{Basal area growth rate increases with size.}
+\label{f-SI_size_dastdt}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_mass_fraction.pdf}
+\caption{\textbf{Change in allocation with size.}
+\label{f-mass_fraction}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
+\caption{\textbf{The expected correlation between leaf-construction cost
+and growth rate changes with plant size.} \label{f-lma_growth_size}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_lma_effects_at_diameters_relative.pdf}
+\caption{\textbf{The expected correlation between leaf-construction cost
+and relative growth rate changes with plant size.}
+\label{f-lma_growth_size_relative}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
+\caption{\textbf{The expected correlation between stem-construction cost
+and growth rate changes with plant size.} \label{f-rho_growth_size}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_rho_effects_at_diameters_relative.pdf}
+\caption{\textbf{The expected correlation between stem-construction cost
+and relative growth rate changes with plant size.}
+\label{f-rho_growth_size_relative}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_hmat_effects_at_diameters.pdf}
+\caption{\textbf{The expected correlation between maximum height and
+growth rate changes with plant size.} \label{f-rho_growth_size}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/SI_hmat_effects_at_diameters_relative.pdf}
+\caption{\textbf{The expected correlation between maximum height and
+relative growth rate changes with plant size.}
+\label{f-hmat_growth_size_relative}}
+\end{figure}
+
+\newpage
 
 \bibliography{references}
+
+\end{appendices}
 \end{document}

### 2015-04-28 R/ms-fun.R
Update build tools

@@ -1,7 +1,22 @@
-tex_2_pdf <- function(texfile){
+latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE){
+  # filename <- tools::file_path_sans_ext(filename)
+  build <- sprintf("%s -interaction=nonstopmode %s", latex_engine, filename)
+  system(build)
+  if(bibtex){
+	  system(sprintf("bibtex %s",   tools::file_path_sans_ext(filename)))
+  }
+  system(build)
+  system(build)
+  remove_aux_files(filename)
+}
 
-  system(sprintf('latexmk -pdf -pdflatex="xelatex -interaction=nonstopmode" %s', texfile))
-  aux.files <- paste0(tools::file_path_sans_ext(texfile),
+latex_build2 <- function(filename, latex_engine="xelatex"){
+  system(sprintf('latexmk -pdf -pdflatex="%s -interaction=nonstopmode" %s', latex_engine, filename))
+  remove_aux_files(filename)
+}
+
+remove_aux_files <- function(filename) {
+	aux.files <- paste0(tools::file_path_sans_ext(filename),
                      c(".log", ".aux", ".bbl", ".blg", ".fls", ".out",
                       ".fdb_latexmk"))
   file.remove(aux.files[file.exists(aux.files)])

### 2015-04-28 ms.tex
Fix latex errors in ms

@@
 
-Results fo poort & binger for surival strongest for LL.
+Results fo poort \& binger for surival strongest for LL.
 
@@
 \begin{table}[h]
-\caption{Equations for an allometric growth model, based on functional balance assumption in {\texbbf a}. }
+\caption{Equations for an allometric growth model, based on functional balance assumption in {\textbf a}. }
 \centering
@@
     $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}} = \phi$ &
-    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\
+    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
   sapwood mass &
@@
     $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} = \theta\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
-    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\
-  bark mass&$m_\textrm{sb} = b\, \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
+    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+  bark mass &
+    $m_\textrm{sb} = b\, \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
     $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} = b \, \theta \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
-    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\
+    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
 
@@
     $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}} = \alpha_3$  &
-    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t} \\ \\
+    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\ \\
 

### 2015-04-28 R/ms-fun.R
expand list of ignore accessory files from latex builds

@@ -17,7 +17,7 @@ latex_build2 <- function(filename, latex_engine="xelatex"){
 
 remove_aux_files <- function(filename) {
 	aux.files <- paste0(tools::file_path_sans_ext(filename),
-                     c(".log", ".aux", ".bbl", ".blg", ".fls", ".out",
+                     c(".log", ".aux", ".bbl", ".blg", ".fls", ".out", ".snm", ".nav", ".toc",
                       ".fdb_latexmk"))
   file.remove(aux.files[file.exists(aux.files)])
 }

### 2015-04-29 fig2.tex
New conceptual figure #37

\documentclass{beamer}
\usepackage[size=a4, orientation=portrait]{beamerposter}
\setbeamertemplate{navigation symbols}{}
\setbeamersize{text margin left=0em,text margin right=0em}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows,positioning, decorations, decorations.text}

% use Helvetica font
\usepackage{fontspec}
\usefonttheme{serif}
\usepackage{helvet}

%----------------------------------------------------------------------------------------
\begin{document}

\begin{tikzpicture}[remember picture, overlay]

\newcommand{\rowspace}{3.2}
\newcommand{\colspace}{3.5}

\newcommand{\plot}[2]{
  \node[anchor=center, rectangle]
    at (#1) {\includegraphics[height=1.65cm]{#2}};
}

\newcommand{\ylabel}[2]{
  \node at (#1)[anchor=west, text width=2cm, minimum height=1cm] {
  {\footnotesize #2}};
}

\newcommand{\textbox}[2]{
  \node at (#1)[anchor=center] {
  {\footnotesize #2}};
}

\newcommand{\defineRow}[2]{
  \node at ($(top) + (-9.5cm, -#1*\rowspace cm) + (0cm, -1.5cm)$) (#2){};
}

\newcommand{\defineCol}[3]{
  \node at ($(#1) + (#2*\colspace cm, 0cm)$) (#3){};
}

\newcommand{\rowtitle}[1]{
  \node at ($(thisrow) + (0.25cm, 1.5cm)$) [anchor=west, text width=8cm, text height=0.5cm]{
  {\footnotesize #1}};
}

\newcommand{\rowlabel}[1]{
  \node at ($(thisrow) + (0.25cm, 0cm)$)[anchor=west, text width=2cm] {
  \footnotesize {\color{orange!50}  #1}};
}

\newcommand{\firstcol}[4]{
  \defineCol{thisrow}{#1}{thiscol}
  \ylabel{$(thiscol) + (-2.1cm,  0cm)$}{#2}
  \plot{  $(thiscol)$                 }{#3}
  \ylabel{$(thiscol) + (0.7cm,   0cm)$}{=}
  \ylabel{$(thiscol) + (1.1cm,   0cm)$}{#4}
}

\newcommand{\othercol}[4]{
  \defineCol{thisrow}{#1}{thiscol}
  \ylabel{$(thiscol) + (-2.1cm,   0cm)$}{#2}
  \plot{  $(thiscol)$                  }{#3}
  \ylabel{$(thiscol) + (0.85cm,   0cm)$}{#4}
}

\node at ($(current page.south)+ (0cm, 23cm)$) (top){};

\node[rectangle, anchor =south] at ($(top)+ (-5cm, 2cm)$){
  \includegraphics[height= 2cm]{downloads/seedling.png}};
  \textbox{$(top)+ (-5cm, 6.2cm)$}{Seedling}

\node[rectangle, anchor =south] at ($(top)+ (5cm, 0.5cm)$){
  \includegraphics[height= 5cm]{downloads/tree.png}};
  \textbox{$(top)+ (5cm, 6.25cm)$}{Adult}

\defineRow{0}{thisrow}
\rowtitle{Biomass production}
\firstcol{1}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/empty_box}{}
\othercol{2}{$Y$}{figures/empty_box}{$\times$}
\othercol{3}{$\Bigg(A$}{figures/empty_box}{-}
\othercol{4}{\ $R$}{figures/empty_box}{$\Bigg)$}
\othercol{5}{$-T$}{figures/empty_box}{}
\rowlabel{(1)}

\defineRow{1}{thisrow}
\rowtitle{Construction costs}
\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/empty_box}{$\Bigg($}
\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\Bigg)^{-1}$}
 \rowlabel{(2)}

\defineRow{2}{thisrow}
\rowtitle{Leaf-area growth}
\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/empty_box}{}
\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/empty_box}{$\times$}
\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}P}$}{figures/empty_box}{$\times$}
\othercol{4}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/empty_box}{}
\rowlabel{(3)}

\defineRow{3}{thisrow}
\rowtitle{Height growth}
\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/empty_box}{}
\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/empty_box}{}
\rowlabel{(4)}

\defineRow{4}{thisrow}
\rowtitle{Basal-area growth}
\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/empty_box}{}
\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/empty_box}{+}
\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{figures/empty_box}{}
\rowlabel{(5)}

\defineRow{5}{thisrow}
\rowtitle{Diameter growth}
\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/empty_box}{}
\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{figures/empty_box}{$\times$}
\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/empty_box}{}
\rowlabel{(6)}

\textbox{$(current page.south)+ (0cm, 3cm)$}{Plant height (m)}

\newcommand{\colNum}{0.1}

\defineRow{0}{rowOne}
\defineCol{rowOne}{\colNum}{A}

\defineRow{1}{rowTwo}
\defineCol{rowTwo}{\colNum}{B}

\defineRow{2}{rowThree}
\defineCol{rowThree}{\colNum}{C}

\defineRow{3}{rowFour}
\defineCol{rowFour}{\colNum}{D}

\defineRow{4}{rowFive}
\defineCol{rowFive}{\colNum}{E}

\defineRow{5}{rowSix}
\defineCol{rowSix}{\colNum}{F}

\tikzstyle{s1}=[->,line width=2pt, draw=orange!50]

\draw [s1] (A) to [bend right, looseness=1] (C);
\draw [s1] (B) to [bend right] (C);
\draw [s1] (C) to [bend right, looseness=1] (D);
\draw [s1] (C) to [bend right, looseness=1] (E);
\draw [s1] (E) to [bend right, looseness=1] (F);

\end{tikzpicture}
\end{document}

### 2015-04-29 R/figures.R

@@ -327,3 +327,9 @@ figure_mass_fraction <- function() {
 
   box()
 }
+
+empty_box <- function() {
+  par(oma=c(0,0,0,0), mar=rep(0.1,4))
+  plot(1,1, ann=FALSE, axes=FALSE, type='n')
+  box()
+}

### 2015-04-29 R/ms-fun.R

@@ -1,13 +1,17 @@
-latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE){
-  # filename <- tools::file_path_sans_ext(filename)
+latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE, dest_dir = "."){
   build <- sprintf("%s -interaction=nonstopmode %s", latex_engine, filename)
+  basename <- tools::file_path_sans_ext(filename)
   system(build)
   if(bibtex){
-	  system(sprintf("bibtex %s",   tools::file_path_sans_ext(filename)))
+	  system(sprintf("bibtex %s",  basename))
+    system(build)
   }
   system(build)
-  system(build)
   remove_aux_files(filename)
+  if(dest_dir!="."){
+    dir.create(dest_dir, showWarnings = FALSE, recursive = TRUE)
+    system(sprintf("mv %s.pdf %s", basename, dest_dir))
+  }
 }
 
 latex_build2 <- function(filename, latex_engine="xelatex"){
@@ -20,4 +24,16 @@ remove_aux_files <- function(filename) {
                      c(".log", ".aux", ".bbl", ".blg", ".fls", ".out", ".snm", ".nav", ".toc",
                       ".fdb_latexmk"))
   file.remove(aux.files[file.exists(aux.files)])
+}
+
+download_seedling_png <- function(destination_filename) {
+  url <-
+    "http://ian.umces.edu/imagelibrary/albums/userpics/10002/normal_ian-symbol-quercus-spp-oak-seedling.png"
+  download(url, destination_filename, mode="wb")
+}
+
+download_tree_png <- function(destination_filename) {
+  url <-
+    "http://ian.umces.edu/imagelibrary/albums/userpics/12789/normal_ian-symbol-quercus-montana.png"
+  download(url, destination_filename, mode="wb")
 }

### 2015-04-29 fig2.tex
panels for conceptual figure #37

@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/empty_box}{}
+\firstcol{1}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/F2_mass_production_net_production}{}
 \othercol{2}{$Y$}{figures/empty_box}{$\times$}
-\othercol{3}{$\Bigg(A$}{figures/empty_box}{-}
-\othercol{4}{\ $R$}{figures/empty_box}{$\Bigg)$}
-\othercol{5}{$-T$}{figures/empty_box}{}
+\othercol{3}{$\Bigg(A$}{figures/F2_mass_production_assimilation}{-}
+\othercol{4}{\ $R$}{figures/F2_mass_production_respiration}{$\Bigg)$}
+\othercol{5}{$-T$}{figures/F2_mass_production_turnover}{}
 \rowlabel{(1)}
@@
 \rowtitle{Construction costs}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/empty_box}{$\Bigg($}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\Bigg($}
 \othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
@@
 \rowtitle{Leaf-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/empty_box}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}P}$}{figures/empty_box}{$\times$}
-\othercol{4}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/empty_box}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
+\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}P}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
+\othercol{4}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/F2_leaf_area_growth_net_production}{}
 \rowlabel{(3)}
@@
 \rowtitle{Height growth}
-\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/empty_box}{}
-\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/empty_box}{}
+\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/F2_height_growth_height_growth_rate}{}
+\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_height_growth_dheight_dleaf_area}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
 \rowlabel{(4)}
@@
 \rowtitle{Basal-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/empty_box}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_basal_growth_dbasal_area_dt}{}
 \othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/empty_box}{+}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{+}
 \othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{figures/empty_box}{}
@@
 \rowtitle{Diameter growth}
-\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/empty_box}{}
+\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/F2_diameter_growth_dbasal_diam_dt}{}
 \othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/empty_box}{}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_basal_growth_dbasal_area_dt}{}
 \rowlabel{(6)}

### 2015-04-29 R/figures.R

@@ -67,23 +67,33 @@ figure_trait_deriative <- function(type, trait_name="lma", canopy_openness=1,
   points(dat[["height"]], line2, type="l", col="red")
 }
 
-
 ## Code for the "rates vs size" figure set:
-figure_rate_vs_size <- function(type, canopy_openness=1,
-                                strategy=default_strategy()) {
-  dat <- figure_rate_vs_size_data(canopy_openness, strategy)
-
+figure_rate_vs_size <- function(data, type) {
   yvars <- figure_rate_vs_size_cols(type)
-
   par(mfrow=c(1, length(yvars)), oma=c(3,1,1,1))
   for (v in yvars) {
-    plot(dat[["height"]], dat[[v]], type="l", xlab="", ylab= v, ylim=c(0, max(1,dat[[v]])))
+    plot(data[["height"]], data[[v]], type="l", xlab="", ylab= v, ylim=c(0, max(1,data[[v]])))
   }
   mtext("Height (m)", 1, cex=1, line=, outer=TRUE)
 
 }
 
-figure_rate_vs_size_data <- function(canopy_openness, strategy) {
+figure_rate_vs_size_panels <- function(data, type, path) {
+  yvars <- figure_rate_vs_size_cols(type)
+  for (v in yvars) {
+    filename <- file.path(path, sprintf("F2_%s_%s.pdf", type, v))
+    pdf(filename,width=5, height=5)
+    on.exit(dev.off())
+    par(oma=c(0,0,0,0), mar=rep(0.1,4))
+    plot(data[["height"]], data[[v]], type="l", ann=FALSE, axes=FALSE,
+      ylim=c(0, max(1,data[[v]])),
+      col="green", lwd=3)
+    box()
+
+  }
+}
+
+figure_rate_vs_size_data <- function(canopy_openness=1, strategy=default_strategy()) {
   heights <- seq(Plant(strategy)$height, strategy$hmat, length.out=100)
   env <- fixed_environment(canopy_openness)
   res <- run_plant_to_heights(heights, strategy, env)
@@ -92,15 +102,15 @@ figure_rate_vs_size_data <- function(canopy_openness, strategy) {
 }
 
 figure_rate_vs_size_cols <- function(type) {
-  if (type == "mass production") {
+  if (type == "mass_production") {
     c("net_production", "assimilation", "respiration", "turnover")
-  } else if (type == "height growth") {
+  } else if (type == "height_growth") {
[…]

### 2015-04-29 R/ms-fun.R

@@ -1,5 +1,5 @@
 latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE, dest_dir = "."){
-  build <- sprintf("%s -interaction=nonstopmode %s", latex_engine, filename)
+  build <- sprintf("%s %s", latex_engine, filename)
   basename <- tools::file_path_sans_ext(filename)
   system(build)
   if(bibtex){

### 2015-04-29 R/ms-fun.R
enhance latex build commands

@@ -1,5 +1,8 @@
-latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE, dest_dir = "."){
-  build <- sprintf("%s %s", latex_engine, filename)
+# interaction must be one of: batchmode, nonstopmode, scrollmode, errorstopmode
+# default combination causes latex to bail back to system on error
+
+latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE, dest_dir = ".", interaction="nonstopmode") {
+  build <- sprintf("%s -interaction=%s -halt-on-error -output-directory=%s %s", latex_engine, interaction, dest_dir, filename)
   basename <- tools::file_path_sans_ext(filename)
   system(build)
   if(bibtex){
@@ -7,16 +10,7 @@ latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE, dest_dir 
     system(build)
   }
   system(build)
-  remove_aux_files(filename)
-  if(dest_dir!="."){
-    dir.create(dest_dir, showWarnings = FALSE, recursive = TRUE)
-    system(sprintf("mv %s.pdf %s", basename, dest_dir))
-  }
-}
-
-latex_build2 <- function(filename, latex_engine="xelatex"){
-  system(sprintf('latexmk -pdf -pdflatex="%s -interaction=nonstopmode" %s', latex_engine, filename))
-  remove_aux_files(filename)
+  remove_aux_files(file.path(dest_dir, basename(filename)))
 }
 
 remove_aux_files <- function(filename) {

### 2015-04-29 R/figures_data.R
Download baad and build figure from data. Closes #26

@@ -37,3 +37,136 @@ figure_lma_tradeoff <- function(data) {
   legend("topright", legend=tolower(names(col_table)), bty="n",
          pch=16, col=col_table, cex=1, title=title)
 }
+
+download_baad <- function(destination_filename) {
+  url <-
+    "https://github.com/dfalster/baad/releases/download/v0.9.0/baad.rds"
+  download(url, destination_filename, mode="wb")
+  # download function from package downloader provides wrapper
+  # to download file so that works for https and across platforms
+}
+
+addIsoclines <- function(A= seq(-10,10),b=1,...){
+  for(a in A) abline(a,b,lty="dotted",...)
+}
+
+add.sma.by.group <-function (data, xvar, yvar, smagroup, colorBygroup, col.table,nmin=10,...){
+  if(sum(!is.na(data[[yvar]]*data[[xvar]]))>nmin){
+    col=colour.by.category(data[[colorBygroup]][1],col.table)
+    fit <- sma(data[[yvar]]~data[[xvar]]*data[[smagroup]], log="xy")
+    plot(fit, type='l',col=col,p.lines.transparent=0.1,add=TRUE,...)
+    invisible(fit)
+  }
+}
+
+make.color.table <- function(names){
+  col.table <- nice_colors(length(names))
+  names(col.table) <- names
+  col.table
+}
+
+# make a plot, fits lines to each within each group, applyiong same colour within that group
+plot_with_sma_by_var_within_group <- function(data, xvar,yvar,group="vegetation",location=NULL, isoclines=NULL,...){
+
+  plot(data[[xvar]], data[[yvar]],  type= 'n', log="xy", las=1, yaxt="n", xaxt="n",...)
+  axis.log10(1)
+  axis.log10(2)
+
+  if(!is.null(isoclines))
+    addIsoclines(b=isoclines)
+
+  points(data[[xvar]], data[[yvar]],  type= 'p', col = make.transparent("grey", 0.4), pch=16)
+
+  col.table <- make.color.table(unique(data[[group]]))
+
+  fits <- dlply(data, group, add.sma.by.group, xvar=xvar, yvar=yvar, col.table=col.table,
+    smagroup="species",colorBygroup=group)
+
+
+  get.significant.groups <- function(fit, p=0.1){
[…]

### 2015-04-29 figure_conceptual.tex
General tidy, remove outdated material, rename variables

--- 2015-04-29 fig2.tex
@@
 \documentclass{beamer}
-\usepackage[size=a4, orientation=portrait]{beamerposter}
+\usepackage[size=custom, width=21, height=27, scale=1.0, orientation=portrait]{beamerposter}
 \setbeamertemplate{navigation symbols}{}
@@
 
-\node at ($(current page.south)+ (0cm, 23cm)$) (top){};
+\node at ($(current page.south)+ (0cm, 20cm)$) (top){};
 
@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/F2_mass_production_net_production}{}
-\othercol{2}{$Y$}{figures/empty_box}{$\times$}
+\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_mass_production_net_production}{}
+\othercol{2}{$y$}{figures/empty_box}{$\times$}
 \othercol{3}{$\Bigg(A$}{figures/F2_mass_production_assimilation}{-}
@@
 \othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}P}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{4}{$\frac{\textrm{d}P}{\textrm{d}t}$}{figures/F2_leaf_area_growth_net_production}{}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
+\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_leaf_area_growth_net_production}{}
 \rowlabel{(3)}
@@
 
-\textbox{$(current page.south)+ (0cm, 3cm)$}{Plant height (m)}
+\textbox{$(current page.south)+ (0cm, 0.5cm)$}{Plant height (m)}
 

### 2015-04-29 ms.tex
New conceptual figure #37
Download baad and build figure from data. Closes #26
General tidy, remove outdated material, rename variables

@@
 \pagestyle{fancy}
+\setlength{\headheight}{15pt}
 \pagenumbering{arabic}
@@
 \hypersetup{breaklinks=true,
-            bookmarks=true,
             pdfauthor={},
@@
 The results of this new analysis
-(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
-\ref{tab:predictions}) showing how the effect of traits on growth
+(Fig. \ref{f-BCI}) align closely with our theoretical predictions  showing how the effect of traits on growth
 changes with plant size, with different traits exhibiting distinct
@@
 We use a new method to estimate species potential growth rate,
-accounting for size specific effects (Fig. \ref{S-quantile_examples},
+accounting for size specific effects (Fig. \ref{fS-quantile_examples},
 see Methods for details). This new method allowed us to extract a more
@@
 We begin with a standard model for the amount of biomass available for
-growth, $dP / \textrm{d}t$, given by the difference between income
+growth, $\textrm{d}b / \textrm{d}t$, given by the difference between income
 (total photosynthesis) and losses (respiration and turnover) within the
@@
 
-\begin{equation}\label{eq:dPdt}
-\underbrace{\strut \frac{dP}{\textrm{d}t}}_\textrm{net biomass production}
-  = \underbrace{\strut Y}_\textrm{yield}
-    \big( \underbrace{\strut a_\textrm{l} \, A}_\textrm{photosynthesis} -
+\begin{equation}\label{eq:dbdt}
+\underbrace{\strut \frac{\textrm{d}b}{\textrm{d}t}}_\textrm{net biomass production}
+  = \underbrace{\strut y}_\textrm{yield}
+    \big( \underbrace{\strut a_\textrm{l} \, p}_\textrm{photosynthesis} -
      \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
@@
 $b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
-rate of CO$_2$ per leaf area and $Y$ is yield: the fraction of
+rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
 assimilated carbon fixed in biomass (the remaining fraction being lost
@@
 $a_\textrm{l} = m_\textrm{l} / \phi$, where $\phi$ is leaf mass per area.
-The total mass of living tissues $m_\textrm{t}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.$
+The total mass of living tissues $m_\textrm{a}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.$
 
 The only potentially controversial element of in eq.
-\ref{eq:dPdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
+\ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
 This assumption is likely to hold true for most woody plants, because the .....
@@
 \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
-\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}}_\textrm{marginal leaf deployment}
-\times \underbrace{\strut\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}}_\textrm{allocation to growth}
-\times \underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}}_\textrm{mass production}.
+\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}}_\textrm{marginal leaf deployment}
+\times \underbrace{\strut\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}}_\textrm{allocation to growth}
+\times \underbrace{\strut\frac{\textrm{d}b}{\textrm{d}t}}_\textrm{mass production}.
 \end{equation}
@@
 
-The second term, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$,
+The second term, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$,
 accounts for the marginal cost of deploying an additional unit of leaf
 area, including construction of the leaf itself and various support
-structures. As such, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$
+structures. As such, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$
 can itself be expressed as a sum of construction costs per unit leaf
@@
 \begin{equation}\label{eq:daldmt}
-\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
+\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}
 = \big(\phi
@@
 The third term in eq \ref{eq:dhdt},
-$\textrm{d}m_\textrm{t} / \textrm{d}P$, gives the fraction of net
-biomass production (eq. \ref{eq:dPdt}) that is allocated to growth
+$\textrm{d}m_\textrm{a} / \textrm{d}P$, gives the fraction of net
+biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
 rather than reproduction or storage.
@@
 \left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
-\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}} \times \frac{\textrm{d}m_\textrm{t}}{\textrm{d}P} \times \frac{\textrm{d}P}{\textrm{d}t} +\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} .
+\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}} \times \frac{\textrm{d}m_\textrm{a}}{\textrm{d}b} \times \frac{\textrm{d}b}{\textrm{d}t} +\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} .
 \end{equation}
@@
 This is achieved by linking the turnover rate of leaf in eq.
-\ref{eq:dPdt} to $\phi$. We use an allometric scaling relationship of
+\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
 the form $k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
@@
 growth by adjusting the amount of energy invested in growth, i.e.~the
-term $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$ in eqs.
+term $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ in eqs.
 \ref{eq:dhdt} and \ref{eq:dast}.
@@
 
-\subsection{Changes in growth rate with size}\label{changes-in-growth-rate-with-size}
+\subsection{Changes in growth rate with size}
 
 This new growth model captures the intrinsically size-dependent nature
-of plant growth (Table \ref{tab:predictions}a), recovering well-known
+of plant growth, recovering well-known
 empirical patterns \citep{sillett-2010, king-2011}. Height growth shows a hump-shaped pattern
@@
 systematic changes in the four components of eq. \ref{eq:dhdt} with size
-(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
+(Fig. \ref{f-conceptual}), including a strong decline in the fraction of plant
 that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
 and increasing reproductive allocation. In contrast, basal-area growth
-continues to increase with size \citep{sillett-2010,
-@stephenson-2014}, due to an increasing influence of stem
+continues to increase with size \citep{sillett-2010, stephenson-2014},
+due to an increasing influence of stem
 turnover. Diameter growth shows a weakly hump shaped
@@
 
-\subsection{How traits influence growth rate}\label{how-traits-influence-growth-rate}
-
-\subsubsection{LMA}\label{lma}
+\subsection{How traits influence growth rate}
+
+\subsubsection{LMA}
 
@@
 $\phi$ increases marginal leaf deployment
-($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}$ by economising on
+($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$ by economising on
 construction costs. Second, lower $\phi$ decreases net production
@@
 
-\subsubsection{Wood density}\label{wood-density}
+\subsubsection{Wood density}
 
@@
 
-\subsubsection{hmat}\label{hmax}
+\subsubsection{hmat}
 
@@
 can thus lead to a growth advantage by increasing
-$\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$ (Table
-\ref{tab:predictions}). At smaller sizes, there is no differentiation
+$\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ . At smaller sizes, there is no differentiation
 among species, because most species are focusing on growth (Fig.
@@
 
-\subsubsection{Responisveness to light}\label{hmax}
+\subsubsection{Responisveness to light}
 
 Two other important predictions arise directly from linking traits to
-growth rate (Table \ref{tab:predictions}c). First, traits moderate the
+growth rate. First, traits moderate the
 responsiveness of growth to changes in light environment. This response
@@
 \citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
-$\textrm{d}P / \textrm{d}t / m_\textrm{t}$. This relationship can be derived from eq. \ref{eq:dPdt}  by ...
+$\textrm{d}P / \textrm{d}t / m_\textrm{a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
 (see Appendix \ref{app:traits-RGR}).
@@
 
-\subsection{Shade tolerance}\label{how-traits-influence-growth-rate}
+\subsection{Shade tolerance}\label{shade-tolerance}
 
@@
 
+\section{Acknowledgments}
+
+Figure 2 uses image by Tracey Saxby, IAN Image Library http://ian.umces.edu/imagelibrary.
+
+\newpage
+
+\bibliography{references}
+
 \newpage
@@
 
-\begin{table}[h]
+\begin{table}[ht]
+ \caption{Key variables
+%  $a = l + ss + sb +  r$,
+%  $m_{sh}$,
+%  $m_t = m_a + m_{sh}$,
+%  $st = stem total$
+}
+\centering
+
+{\footnotesize  % smaller font, double space
+ \begin{doublespace}
+  \begin{tabular}{p{2cm}p{2cm}p{7cm}}
+  \hline
+  Symbol & Unit & Description \\
+  \hline
+
+  $b$   & kg  & biomass originating from parent plant\\
+  $m_i$ & kg  & mass retained on plant in tissue $i$\\
+  $a_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
+  $h$   & m  & height of plant\\
+  $D$   & m  & diameter of plant stem\\
+
+  $y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
+  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthesic rate per unit area \\
+  $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
+  $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
+  $A$ & kg yr$^{-1}$ & total plant photosynthesis \\
+  $R$ & kg yr$^{-1}$ & total plant respiration \\
+  $T$ & kg yr$^{-1}$ & total plant turnover \\
+  \hline
+  \end{tabular}
+\end{doublespace}}
+\label{tab:definitions}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
 \caption{Equations for an allometric growth model, based on functional balance assumption in {\textbf a}. }
@@
     $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
-
   root mass &
@@
 
-\begin{table}[h]
+\begin{table}[ht]
 \caption{Effect of traits and related trade-offs on demography}
@@
   & $h_0$
-  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$
-  & $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}$
-  & $\frac{\textrm{d}P}{\textrm{d}t}$
+  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}$
+  & $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$
+  & $\frac{\textrm{d}b}{\textrm{d}t}$
   & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/BCI_data2.pdf}
+\begin{figure}[ht]
+\centering
+\includegraphics{figures/BCI_data.pdf}
 \caption{\textbf{The relationship between traits and potential growth
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_size_dhdt.pdf}
+\begin{figure}[ht]
+\centering
+\includegraphics{figures/figure_conceptual.pdf}
 \caption{\textbf{Conceptual framework for understanding size-dependent
-changes in growth rate.} \label{f-conceptual}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figs/allometry.pdf}
+changes in growth rate.} Hump-shaped relationship between height growth rate and
+size. \label{f-conceptual}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\centering
+\includegraphics{figures/allometry.pdf}
 \caption{\textbf{Key assumptions of a functional balance and trait
@@
 
-\begin{figure}[htbp]
+\begin{figure}[ht]
 \centering
@@
 maximum amount of shading that can be endured before net production (eq.
-\ref{eq:dPdt}) reaches zero. Lines indicate relationship for plants of a
+\ref{eq:dbdt}) reaches zero. Lines indicate relationship for plants of a
 given height. \label{f-wplcp}}
@@
 
-\newpage
+\clearpage
+
 \begin{appendices}
@@
 
-We start with eq. \ref{eq:dPdt} for mass-based growth. For seedlings, which are
+We start with eq. \ref{eq:dbdt} for mass-based growth. For seedlings, which are
 young and mostly leaf, it is reasonable to ignore all turnover terms as
@@
 becomes a linear function of leaf area and net photosynthesis per leaf
-area ($A_\textrm{net} = Y(A - r_\textrm{l})$), making relative growth
+area ($A_\textrm{net} = y(A - r_\textrm{l})$), making relative growth
 rate a linear function of $\phi$:
@@
 \begin{equation}\label{eq:RGR}
-\underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}\frac{1}{m_\textrm{t}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{t}}.
+\underbrace{\strut\frac{\textrm{d}b}{\textrm{d}t}\frac{1}{m_\textrm{a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{a}}.
 \end{equation}
@@
 given height where 100\% of available energy is allocated to growth,
-i.e. $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}=1$. Note also
+i.e. $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}=1$. Note also
 that by the assumptions of our model and above, traits do not influence
@@
 \begin{equation} \label{eq:G2}
-G = \frac{\textrm{d}h}{\textrm{d}t} = c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}  \frac{ \textrm{d}P} {\textrm{d}t} \right),
+G = \frac{\textrm{d}h}{\textrm{d}t} = c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}  \frac{ \textrm{d}b} {\textrm{d}t} \right),
 \end{equation}
@@
 \frac{\partial G} {\partial x} =
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}{\partial x}
- \, \frac{ \textrm{d}P} {\textrm{d}t}
-+ \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
-\, \frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}.
+\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}\right)}{\partial x}
+ \, \frac{ \textrm{d}b} {\textrm{d}t}
++ \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
+\, \frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x}.
 \end{equation}
@@
 \begin{equation} \label{eq:da_dx}
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\right)}
-{\partial x} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}\ \right)^2
-\, \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}\right)}
+{\partial x} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}\ \right)^2
+\, \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
 }{\partial x}.
@@
 \frac{\partial G} {\partial x} =
-\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
+\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
 \left(
-\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
-- \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
-\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
+\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x}
+- \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
+\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
 }{\partial x}
- \, \frac{ \textrm{d}P} {\textrm{d}t}
+ \, \frac{ \textrm{d}b} {\textrm{d}t}
 \right).
@@
 
-Since is $\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}>0$ always, the sign
+Since is $\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}>0$ always, the sign
 of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
@@
 \begin{equation}\label{eq:dg3}
-\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x}
-< \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{t}}
-\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
-}{\partial x} \, \frac{ \textrm{d}P} {\textrm{d}t},
+\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x}
+< \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
+\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x} \, \frac{ \textrm{d}b} {\textrm{d}t},
 \end{equation}
@@
 \frac{
-\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
-{\frac{ \textrm{d}P} {\textrm{d}t}}
+\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x} }
+{\frac{ \textrm{d}b} {\textrm{d}t}}
 <
-\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
-}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
+\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}}{\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}}.
 \end{equation}
@@
 \frac{
-\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial x} }
-{\frac{ \textrm{d}P} {\textrm{d}t}}
+\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x} }
+{\frac{ \textrm{d}b} {\textrm{d}t}}
 =
-\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}\right)
-}{\partial x}}{\frac{\textrm{d}m_\textrm{t}} {\textrm{d}a_\textrm{l}}}.
+\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
+}{\partial x}}{\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}}.
 \end{equation}
@@
 eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{\textrm{d}m_\textrm
-{t}} {\textrm{d}a_\textrm{l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dPdt} we obtain,
-
-\begin{equation}\label{eq:dPdt2}
-\frac{dP}{\textrm{d}t} = c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}
+{t}} {\textrm{d}a_\textrm{l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
+
+\begin{equation}\label{eq:dbdt2}
+\frac{\textrm{d}b}{\textrm{d}t} = c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}
 \end{equation}
@@
 where
-$c_2 = Y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, hmax\textrm{i}})$
+$c_2 = y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, hmax\textrm{i}})$
 is independent of $\phi$, and thus
 
-\begin{equation}\label{eq:dPdt3}
-\frac{\partial \left( \frac{ \textrm{d}P} {\textrm{d}t}\right)}{\partial \phi}  =
+\begin{equation}\label{eq:dbdt3}
+\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial \phi}  =
 (B_4-1) a_\textrm{l} a_4\phi ^{-B_4}.
@@
 
-Eq. \ref{eq:dPdt3} shows that if $B4=1$, $\frac{ \textrm{d}P} {\textrm{d}t}$ is independent
+Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ \textrm{d}b} {\textrm{d}t}$ is independent
 of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
@@
 
-\begin{table}[h]
+\begin{table}[ht]
 \caption{Model parameters}
@@
 
-\begin{figure}[htbp]
+\newpage
+
+\begin{figure}[ht]
 \centering
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_size_dhdt.pdf}
-\caption{\textbf{Hump-shaped relationship between height growth rate and
-size.} \label{f-hump}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_size_dDdt.pdf}
-\caption{\textbf{Hump-shaped relationship diameter growth and size.}
-\label{f-SI_size_dDdt}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_size_dDdt.pdf}
-\caption{\textbf{Basal area growth rate increases with size.}
-\label{f-SI_size_dastdt}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
+\begin{figure}[ht]
 \centering
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
+\begin{figure}[ht]
+\centering
+\includegraphics{figures/SI_rho_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between stem-construction cost
@@
 \caption{\textbf{The expected correlation between maximum height and
-growth rate changes with plant size.} \label{f-rho_growth_size}}
-\end{figure}
-
-\begin{figure}[htbp]
+growth rate changes with plant size.} \label{f-hmat_growth_size}}
+\end{figure}
+
+\begin{figure}[ht]
 \centering
@@
 
-\newpage
-
-\bibliography{references}
-
 \end{appendices}

### 2015-04-29 ms.tex
Fix a few reference keys, add text

@@
 
-\section{To tidy}
-
-mt: living tissue or total?
-
-theta: change to theta$^{-1}$ -- nearly always a negative power
-
-h-al: changed relationship to work other way around - this reads better, but is perhaps opposite of implementation, which is driven by height
+\section{Ideas}\label{abstract}
+
+Table near intro summarising empricial patterns to be explained by integrated model?
 
@@
 plants use to generate and invest surplus energy
- \citep{wright_world-2004, chave-2009, westoby-2002}.
+ \citep{wright-2004, chave-2009, westoby-2002}.
 Although most plants have the same basic physiological function
@@
 mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
-diameter,  maximum  height (HMAX) and seed size  \citep{wright_world-2004,chave-2009}.
+diameter,  maximum  height (HMAX) and seed size  \citep{wright-2004,chave-2009}.
 Data for these traits now exists for up to 10\% of the world's 250000 plant species
@@
 -- the central element of the leaf economics spectrum
- \citep{wright_world-2004} -- has been tightly related to relative growth
+ \citep{wright-2004} -- has been tightly related to relative growth
 rate in plant mass \citep{lambers-1992, wright_cross-2000}. LMA and
@@
 mass-based relative growth rate is linearly and negatively related to LMA
- \citep{lambers-1992, cornelissen-1996,
-wright_cross-2000}. An extension of the model suggests a similar
+ \citep{lambers-1992, cornelissen-1996, wright_cross-2000}.
+ An extension of the model suggests a similar
 relationship may hold at larger sizes \citep{enquist-2007}. But as
@@
  always be limited.
+
+Effect of traits on shade tolerance also porrly udnerstood. Empirical -- shade
+
+Ulitmately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
+Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{sillett-2010, king-2011} or diameter growth \citep{herault-2011}.
+In contrast, basal-area growth continues to increase with size \citep{sillett-2010, stephenson-2014}, due to an increasing influence of stem turnover.
+All growth measures decrease sharply with size when expressed as relative growth rates \citep{rees-2010, iida-2014}.
+Shade tolerance tends to decrease with increasing size \citep{lusk-2006}, due to increasing costs of mainatining stems \citep{givnish-1988}
 
@@
 growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction \citep{wright_world-2004}.
+results from superior leaf construction \citep{wright-2004}.
 This is achieved by linking the turnover rate of leaf in eq.
@@
 shown to describe patterns across diverse
-ecosystems \citep{wright_world-2004} (Fig. \ref{fS-leaf}).
+ecosystems \citep{wright-2004} (Fig. \ref{fS-leaf}).
 
@@
 lower $\rho$ decreases mechanical strength resulting in higher
-mortality \citep{chave-2009,wright-2010}. Under
+mortality \citep{chave-2009, wright-2010}. Under
 this scenario, decreasing $\rho$ will always provide a growth
@@
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
-Data from \citep{wright_world-2004} for 678 species from 51 sites, each
+Data from \citep{wright-2004} for 678 species from 51 sites, each
 point giving a species-average. Lines show standardised major axis lines

### 2015-04-30 R/BCI.R
Cleaning code: spelling, white space

@@ -1,6 +1,6 @@
-#List of functions used to manipulate and summarise data from BCI.
+# List of functions used to manipulate and summarise data from BCI.
 
-#Download BCI data
+# Download BCI data
 # download function from package downloader provides wrapper
 # to download file so that works for https and across platforms
 BCI_download_50ha_plot_full <- function(dest) {
@@ -13,7 +13,7 @@ BCI_download_species_table <- function(dest) {
   download(url, dest, mode="wb")
 }
 
-#Load 50ha census data
+# Load 50ha census data
 BCI_load_50ha_plot <- function(path_to_zip) {
 
   tmp <- tempfile()
@@ -26,9 +26,9 @@ BCI_load_50ha_plot <- function(path_to_zip) {
 
 BCI_calculate_individual_growth <- function(BCI_50haplot, BCI_nomenclature) {
 
-  # Identifies individuals that return from the dead or are supposably refound
+  # Identifies individuals that return from the dead or are supposedly re-found
   # i.e. Individuals given dbh=NA and then later given numeric value
-  # Note this function must be used prior to subsetting only observations with pom=1.3
+  # Note this function must be used prior to sub-setting only observations with pom = 1.3
   is_zombie <- function(dbh) {
     any(diff(is.na(dbh)) == -1)
   }
@@ -106,7 +106,7 @@ BCI_calculate_species_traits <- function(individual_growth, wright_2010) {
   species_data
 }
 
-# function to calculate growth rate of tree, either absolute (default) or relative rgowth rate (set f=log)
+# function to calculate growth rate of tree, either absolute (default) or relative growth rate (set f=log)
 calculate_growth_rate <- function(x, t, f = function(y) y) {
   dt <- diff(t)/365.25
   if (any(dt < 0, na.rm = TRUE)) {
@@ -126,19 +126,19 @@ CTFS_sanity_check <- function(dbh, dbh_increment, dbasal_diam_dt) {
   accept <- rep(TRUE, length(dbh))
   # Remove records based on max growth rate
   accept[dbasal_diam_dt > max_growth] <- FALSE
-  # Remove records based on min growth rate, estimated from allowbale error
+  # Remove records based on min growth rate, estimated from allowable error
[…]

### 2015-04-30 R/BCI.R
Use web version of nomenclature. Closes #41

@@ -134,15 +134,16 @@ CTFS_sanity_check <- function(dbh, dbh_increment, dbasal_diam_dt) {
 
 # Look up family
 lookup_family <- function(tag, nomen){
-  i <- match(tag, tolower(nomen[['sp6']]))
-  nomen$family[i]
+  i <- match(tag, tolower(nomen[['sp']]))
+  nomen$Family[i]
 }
 
 # Look up species code
 lookup_latin <- function(tag, nomen){
-  nomen$latin <- paste(nomen$genus, nomen$species)
-  i <- match(tag, tolower(nomen[['sp6']]))
-  nomen[['latin']][i]
+
+  nomen$Latin <- paste(nomen$Genus, nomen$Species)
+  i <- match(tag, tolower(nomen[['sp']]))
+  nomen[['Latin']][i]
 }
 
 # fits quantile regression to data

### 2015-04-30 R/wright_2010.R

@@ -17,6 +17,7 @@ process_wright_2010 <- function(filename) {
                                   c("alchorneoide","alchorneoides"),
                                   c("Tabernaemont","Tabernaemontana"),
                                   c("Trattinnicki","Trattinnickia"),
+                                  c("princeps", "croatii"),
                                   c("$",""),
                                   c("-99.000000", "NA"),
                                   c("-99.00000", "NA"),

### 2015-04-30 R/figures.R
Update the way internal variables are extracted

@@ -1,47 +1,31 @@
 default_strategy <- function() {
-  Strategy(B5=1.5,
-           k_l0=0.4565855/3,
+  Strategy(
            hmat=30.0,
            c_r1=0.8,
            c_r2=20,
-           k_s0=0.2,
            a1=2.17,
            B1=0.546)
-}
-
-run_plant_to_heights <- function(heights, strategy, env) {
-  f <- function(h, p) {
-    p$height <- h
-    run_plant(p, env)
-  }
-  ## NOTE: previously we checked through these results and filtered
-  ## out things that left NULL
-  do.call("rbind", lapply(heights, f, Plant(strategy)))
 }
 
 run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
                                time_max=Inf) {
   res <- grow_plant_to_size(Plant(strategy), sizes, size_variable,
                             env, time_max)
-  do.call("rbind", lapply(res$plant, plant_info))
+  lapply(res$plant, extract_plant_info, env=env) %>% rbind_list %>% data.frame
 }
 
-run_plant <- function(plant, env) {
-  plant$compute_vars_phys(env)
-  plant_info(plant)
-}
+extract_plant_info <- function(plant, env) {
 
-plant_info <- function(p) {
-  x <- as.data.frame(as.list(c(p$vars_size,
-                               p$vars_phys,
-                               p$vars_growth)))
+  p <- plant
+  p$compute_vars_phys(env)
+  p$compute_vars_growth()
+  x <- unlist(p$internals)
 
   # add relative measures
-  x$height_growth_rate_relative    <- x$height_growth_rate / x$height
-  x$dbasal_area_dt_relative        <- x$dbasal_area_dt / x$basal_area
-  x$dbasal_diam_dt_relative        <- x$dbasal_diam_dt / x$diameter
-  x$dabove_ground_mass_dt_relative <- x$dabove_ground_mass_dt / x$above_ground_mass
-  x
+  for(v in c("height", "area_stem", "diameter_stem", "mass_above_ground")) {
+    x[[sprintf("%s_dt_relative", v)]] <- x[[sprintf("%s_dt", v)]] / x[[v]]
+  }
+ x
 }
 
 ## Code for trait deriative figure

### 2015-04-30 figure_conceptual.tex
A billion name changes

@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_mass_production_net_production}{}
+\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_net_mass_production_dt_net_production}{}
 \othercol{2}{$y$}{figures/empty_box}{$\times$}
-\othercol{3}{$\Bigg(A$}{figures/F2_mass_production_assimilation}{-}
-\othercol{4}{\ $R$}{figures/F2_mass_production_respiration}{$\Bigg)$}
-\othercol{5}{$-T$}{figures/F2_mass_production_turnover}{}
+\othercol{3}{$\Bigg(A$}{figures/F2_net_mass_production_dt_assimilation}{-}
+\othercol{4}{\ $R$}{figures/F2_net_mass_production_dt_respiration}{$\Bigg)$}
+\othercol{5}{$-T$}{figures/F2_net_mass_production_dt_turnover}{}
 \rowlabel{(1)}
@@
 \rowtitle{Height growth}
-\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/F2_height_growth_height_growth_rate}{}
-\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_height_growth_dheight_dleaf_area}{$\times$}
+\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/F2_height_dt_height_dt_rate}{}
+\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_height_dt_dheight_dleaf_area}{$\times$}
 \othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
@@
 \rowtitle{Basal-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_basal_growth_dbasal_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_dbasal_area_dt}{}
 \othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
@@
 \rowtitle{Diameter growth}
-\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/F2_diameter_growth_dbasal_diam_dt}{}
+\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/F2_diameter_stem_dt_dbasal_diam_dt}{}
 \othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_basal_growth_dbasal_area_dt}{}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_dbasal_area_dt}{}
 \rowlabel{(6)}

### 2015-04-30 R/BCI.R

@@ -66,15 +66,15 @@ BCI_calculate_individual_growth <- function(BCI_50haplot, BCI_nomenclature) {
       # First measurement in 1990 ='1990-02-06'
       julian = as.vector(julian(as.Date(exactdate,"%Y-%m-%d"), as.Date("1990-02-06", "%Y-%m-%d"))),
       dbh_increment = c(diff(dbh), NA),
-      dbasal_diam_dt = calculate_growth_rate(dbh, julian),
-      dbasal_diam_dt_relative = calculate_growth_rate(dbh, julian, log)
+      diameter_stem_dt = calculate_growth_rate(dbh, julian),
+      diameter_stem_dt_relative = calculate_growth_rate(dbh, julian, log)
     ) %>%
     filter(
-      !is.na(dbasal_diam_dt) &
+      !is.na(diameter_stem_dt) &
         !is.na(dbh_increment) &
-        CTFS_sanity_check(dbh, dbh_increment, dbasal_diam_dt)
+        CTFS_sanity_check(dbh, dbh_increment, diameter_stem_dt)
     ) %>%
-    select(censusid,species, family, dfstatus, dbh, dbasal_diam_dt)
+    select(censusid,species, family, dfstatus, dbh, diameter_stem_dt)
 }
 
 BCI_calculate_species_traits <- function(individual_growth, wright_2010) {
@@ -89,7 +89,7 @@ BCI_calculate_species_traits <- function(individual_growth, wright_2010) {
     group_by(data, species) %>%
       filter(dbh >= min & dbh < max) %>%
       summarise(count = n(),
-        dbasal_diam_dt = predict_via_quantile_regression(dbh, dbasal_diam_dt, predict_at = at))
+        diameter_stem_dt = predict_via_quantile_regression(dbh, diameter_stem_dt, predict_at = at))
   }
 
   out <- ddply(size_range, 1, function(x) get_species_data(individual_growth, x[["min"]], x[["max"]], x[["at"]]))
@@ -98,7 +98,7 @@ BCI_calculate_species_traits <- function(individual_growth, wright_2010) {
   wright_2010 <- mutate(wright_2010, species = paste(genus, species))
 
   species_data <- merge(select(wright_2010, species, lma, rho, hmat), out, by = "species") %>%
-    filter(!is.na(dbasal_diam_dt)) %>%
[…]

### 2015-04-30 R/axis.R

@@ -3,19 +3,19 @@ name_pretty <- function(trait) {
          lma=expression(paste("Leaf-construction cost (kg ", m^-2, " )")),
          rho=expression(paste("Wood density (kg ", m^-3, " )")),
          hmat="Height at maturation (m)",
-         dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
+         diameter_stem_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
          shading=expression(paste("Maximum leaf area above (", m^2, " )")),
          leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, " )")),
 
-         height_growth_rate=expression(paste("Height growth (m ", yr^-1, " )")),
-         dbasal_diam_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
-         dbasal_area_dt=expression(paste("Basal area growth (", m^2, " ", yr^-1, ")")),
-         dabove_ground_mass_dt=expression(paste("Mass growth (kg ", yr^-1, ")")),
+         height_dt=expression(paste("Height growth (m ", yr^-1, " )")),
+         diameter_stem_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
+         area_stem_dt=expression(paste("Basal area growth (", m^2, " ", yr^-1, ")")),
+         mass_above_ground_dt=expression(paste("Mass growth (kg ", yr^-1, ")")),
 
-         height_growth_rate_relative=expression(paste("Relative height growth (m ", yr^-1, " )")),
-         dbasal_diam_dt_relative=expression(paste("(Relative diameter growth (m ", yr^-1, ")")),
-         dbasal_area_dt_relative=expression(paste("(Relative basal area growth (", m^2, " ", yr^-1, ")")),
-         dabove_ground_mass_dt_relative=expression(paste("(Relative mass growth (kg ", yr^-1, ")")),
+         height_dt_relative=expression(paste("Relative height growth (m ", yr^-1, " )")),
+         diameter_stem_dt_relative=expression(paste("(Relative diameter growth (m ", yr^-1, ")")),
+         area_stem_dt_relative=expression(paste("(Relative basal area growth (", m^2, " ", yr^-1, ")")),
+         mass_above_ground_dt_relative=expression(paste("(Relative mass growth (kg ", yr^-1, ")")),
 
          stop("Unknown trait ", trait))
 }

### 2015-04-30 R/figures.R

@@ -28,7 +28,7 @@ extract_plant_info <- function(plant, env) {
  x
 }
 
-## Code for trait deriative figure
+## Code for trait derivative figure
 figure_trait_deriative <- function(type, trait_name="lma", canopy_openness=1,
                                 strategy=default_strategy()) {
 
@@ -41,10 +41,10 @@ figure_trait_deriative <- function(type, trait_name="lma", canopy_openness=1,
   strategy2[[trait_name]] <- x+dx
   dat2 <- figure_rate_vs_size_data(canopy_openness, strategy2)
 
-  line1 <- (dat2$net_production - dat$net_production)/dx/dat$net_production
-  line2  <- (1/(dat2$dleaf_area_dleaf_mass * dat2$leaf_fraction)
-              - 1/(dat$dleaf_area_dleaf_mass * dat$leaf_fraction)) / dx /
-            1/(dat2$dleaf_area_dleaf_mass * dat2$leaf_fraction)
+  line1 <- (dat2$net_mass_production_dt - dat$net_mass_production_dt)/dx/dat$net_mass_production_dt
+  line2  <- (1/(dat2$darea_leaf_dmass_leaf * dat2$leaf_fraction)
+              - 1/(dat$darea_leaf_dmass_leaf * dat$leaf_fraction)) / dx /
+            1/(dat2$darea_leaf_dmass_leaf * dat2$leaf_fraction)
   height <- dat$height
 
   plot(dat[["height"]], line1, type="l", xlab="Height (m)", ylab= "relative change", ylim=c(0,20))
@@ -80,36 +80,36 @@ figure_rate_vs_size_panels <- function(data, type, path) {
 figure_rate_vs_size_data <- function(canopy_openness=1, strategy=default_strategy()) {
   heights <- seq(Plant(strategy)$height, strategy$hmat, length.out=100)
   env <- fixed_environment(canopy_openness)
-  res <- run_plant_to_heights(heights, strategy, env)
-  i <- res[["net_production"]] > 0
+  res <- run_plant_to_sizes(heights, "height", strategy, env, time_max=10000)
+  i <- res[["net_mass_production_dt"]] > 0
   res[i, ]
 }
 
 figure_rate_vs_size_cols <- function(type) {
-  if (type == "mass_production") {
-    c("net_production", "assimilation", "respiration", "turnover")
-  } else if (type == "height_growth") {
[…]

### 2015-04-30 sensitivity.R

@@ -16,17 +16,17 @@ suppressPackageStartupMessages({
 # First, look at plot: challenge is to move peak to left and pull down elevation of curve.
 data <- remake::make("BCI_species_data") %>%
   mutate(dbh=at) %>%
-  filter(!is.na(lma*rho*hmat*dbh*dbasal_diam_dt), dbh %in% c(0.01), hmat > 10) %>%
-  select(dbh, lma, rho, dbasal_diam_dt)
+  filter(!is.na(lma*rho*hmat*dbh*diameter_stem_dt), dbh %in% c(0.01), hmat > 10) %>%
+  select(dbh, lma, rho, diameter_stem_dt)
 
 d <- trait_effects_data("lma", "diameter")
 d <- d[d$size_class == 2L,]
 
-plot(data$lma, data$dbasal_diam_dt, las=1, log="x",
-     xlab=name_pretty("lma"), ylab=name_pretty("dbasal_diam_dt"),
+plot(data$lma, data$diameter_stem_dt, las=1, log="x",
+     xlab=name_pretty("lma"), ylab=name_pretty("diameter_stem_dt"),
      xlim=range(data$lma, d$lma),
-     ylim=range(data$dbasal_diam_dt, d$dbasal_diam_dt))
-lines(d$lma, d$dbasal_diam_dt, col="red")
+     ylim=range(data$diameter_stem_dt, d$diameter_stem_dt))
+lines(d$lma, d$diameter_stem_dt, col="red")
 
 # given a dbh x, traits and parameters, grows plant to right size and estimates growth rate
 growth <- function(x, trait, dbh, E=1,
@@ -34,7 +34,7 @@ growth <- function(x, trait, dbh, E=1,
   strategy[trait] <- x
   pp <- grow_plant_to_size(Plant(strategy), dbh, "diameter",
                            fixed_environment(E))
-  sapply(pp$plant, function(p) p$vars_growth[["dbasal_diam_dt"]])
+  sapply(pp$plant, function(p) p$vars_growth[["diameter_stem_dt"]])
 }
 
 # finds trait value in range that maximises growth rate at given size and light

### 2015-05-13 figure_conceptual.tex
more name changes to work with tree2

@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_net_mass_production_dt_net_production}{}
+\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_net_mass_production_dt_net_mass_production_dt}{}
 \othercol{2}{$y$}{figures/empty_box}{$\times$}
@@
 \rowtitle{Construction costs}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\Bigg($}
-\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\Bigg)^{-1}$}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
+\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
+\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
+\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_darea_leaf_dmass_live_dmass_root_darea_leaf}{$\Bigg)^{-1}$}
  \rowlabel{(2)}
-
 \defineRow{2}{thisrow}
 \rowtitle{Leaf-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_leaf_area_growth_net_production}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_area_leaf_dt_area_leaf_dt}{}
+\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_area_leaf_dt_darea_leaf_dmass_live}{$\times$}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{figures/F2_area_leaf_dt_fraction_allocation_growth}{$\times$}
+\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_area_leaf_dt_net_mass_production_dt}{}
 \rowlabel{(3)}
@@
 \rowtitle{Height growth}
-\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/F2_height_dt_height_dt_rate}{}
-\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_height_dt_dheight_dleaf_area}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/F2_height_dt_height_dt}{}
+\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_height_dt_dheight_darea_leaf}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_area_leaf_dt_area_leaf_dt}{}
 \rowlabel{(4)}
@@
 \rowtitle{Basal-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_dbasal_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_area_stem_dt}{}
 \othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{+}
-\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{figures/empty_box}{}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_area_leaf_dt}{+}
+\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_area_heartwood_dt}{}
 \rowlabel{(5)}
@@
 \rowtitle{Diameter growth}
-\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/F2_diameter_stem_dt_dbasal_diam_dt}{}
-\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_dbasal_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/F2_diameter_stem_dt_diameter_stem_dt}{}
+\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{figures/F2_diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_diameter_stem_dt_area_stem_dt}{}
 \rowlabel{(6)}

### 2015-05-13 R/figures.R

@@ -4,11 +4,14 @@ default_strategy <- function() {
            c_r1=0.8,
            c_r2=20,
            a1=2.17,
-           B1=0.546)
+           B1=0.546,
+           k_l=0.4565855/3
+
+           )
 }
 
 run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
-                               time_max=Inf) {
+                               time_max=10000) {
   res <- grow_plant_to_size(Plant(strategy), sizes, size_variable,
                             env, time_max)
   lapply(res$plant, extract_plant_info, env=env) %>% rbind_list %>% data.frame
@@ -16,6 +19,7 @@ run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
 
 extract_plant_info <- function(plant, env) {
 
+  if(!is.null(plant))
   p <- plant
   p$compute_vars_phys(env)
   p$compute_vars_growth()
@@ -42,9 +46,9 @@ figure_trait_deriative <- function(type, trait_name="lma", canopy_openness=1,
   dat2 <- figure_rate_vs_size_data(canopy_openness, strategy2)
 
   line1 <- (dat2$net_mass_production_dt - dat$net_mass_production_dt)/dx/dat$net_mass_production_dt
-  line2  <- (1/(dat2$darea_leaf_dmass_leaf * dat2$leaf_fraction)
-              - 1/(dat$darea_leaf_dmass_leaf * dat$leaf_fraction)) / dx /
-            1/(dat2$darea_leaf_dmass_leaf * dat2$leaf_fraction)
+  line2  <- (1/(dat2$darea_leaf_dmass_leaf * dat2$darea_leaf_dmass_live)
+              - 1/(dat$darea_leaf_dmass_leaf * dat$darea_leaf_dmass_live)) / dx /
+            1/(dat2$darea_leaf_dmass_leaf * dat2$darea_leaf_dmass_live)
   height <- dat$height
 
   plot(dat[["height"]], line1, type="l", xlab="Height (m)", ylab= "relative change", ylim=c(0,20))
@@ -64,38 +68,39 @@ figure_rate_vs_size <- function(data, type) {
 
 figure_rate_vs_size_panels <- function(data, type, path) {
   yvars <- figure_rate_vs_size_cols(type)
+
   for (v in yvars) {
     filename <- file.path(path, sprintf("F2_%s_%s.pdf", type, v))
     pdf(filename,width=5, height=5)
[…]

### 2015-08-14 R/BCI.R
Use remake's download targets
Removes downloader dependency, simplifies code a bit.

@@ -1,21 +1,7 @@
 # List of functions used to manipulate and summarise data from BCI.
 
-# Download BCI data
-# download function from package downloader provides wrapper
-# to download file so that works for https and across platforms
-BCI_download_50ha_plot_full <- function(dest) {
-  url <-"https://repository.si.edu/bitstream/handle/10088/20925/bci.full.Rdata31Aug2012.zip"
-  download(url, dest, mode="wb")
-}
-
-BCI_download_species_table <- function(dest) {
-  url <-"https://repository.si.edu/bitstream/handle/10088/20925/bci.spptable.rdata"
-  download(url, dest, mode="wb")
-}
-
 # Load 50ha census data
 BCI_load_50ha_plot <- function(path_to_zip) {
-
   tmp <- tempfile()
   unzip(path_to_zip, exdir=tmp)
   on.exit(unlink(tmp, recursive=TRUE))

### 2015-08-14 R/figures_data.R

@@ -38,14 +38,6 @@ figure_lma_tradeoff <- function(data) {
          pch=16, col=col_table, cex=1, title=title)
 }
 
-download_baad <- function(destination_filename) {
-  url <-
-    "https://github.com/dfalster/baad/releases/download/v0.9.0/baad.rds"
-  download(url, destination_filename, mode="wb")
-  # download function from package downloader provides wrapper
-  # to download file so that works for https and across platforms
-}
-
 addIsoclines <- function(A= seq(-10,10),b=1,...){
   for(a in A) abline(a,b,lty="dotted",...)
 }

### 2015-08-14 R/ms-fun.R

@@ -19,15 +19,3 @@ remove_aux_files <- function(filename) {
                       ".fdb_latexmk"))
   file.remove(aux.files[file.exists(aux.files)])
 }
-
-download_seedling_png <- function(destination_filename) {
-  url <-
-    "http://ian.umces.edu/imagelibrary/albums/userpics/10002/normal_ian-symbol-quercus-spp-oak-seedling.png"
-  download(url, destination_filename, mode="wb")
-}
-
-download_tree_png <- function(destination_filename) {
-  url <-
-    "http://ian.umces.edu/imagelibrary/albums/userpics/12789/normal_ian-symbol-quercus-montana.png"
-  download(url, destination_filename, mode="wb")
-}

### 2015-08-14 R/wright_2004.R

@@ -1,9 +1,3 @@
-download_wright_2004 <- function(destination_filename) {
-  url <-
-    "http://www.nature.com/nature/journal/v428/n6985/extref/nature02403-s2.xls"
-  download.file(url, destination_filename)
-}
-
 process_wright_2004 <- function(filename, sitevars) {
   d <- read.xlsx2(filename, sheetIndex = 1, startRow = 11, stringsAsFactors = FALSE, check.names = FALSE)
 

### 2015-08-14 R/wright_2010.R

@@ -1,8 +1,3 @@
-download_wright_2010 <- function(dest) {
-  url <- "http://www.esapubs.org/archive/ecol/E091/257/Supplement_20100505.txt"
-  download.file(url, dest)
-}
-
 process_wright_2010 <- function(filename) {
   # Correct names of species, anything with length > 12 was clipped
   corrections <- data.frame(rbind(c("costaricensi","costaricensis"),

### 2015-08-14 sensitivity.R

@@ -7,7 +7,7 @@ source_dir <- function(path, ...) {
 }
 source_dir("R")
 
-library(tree2)
+library(plant)
 suppressPackageStartupMessages({
   library(plyr)
   library(dplyr)

### 2015-08-17 R/figures.R
Minor fixes, works end to end with plant
Closes #39

@@ -68,7 +68,7 @@ figure_rate_vs_size <- function(data, type) {
   yvars <- figure_rate_vs_size_cols(type)
   par(mfrow=c(1, length(yvars)), oma=c(3,1,1,1))
   for (v in yvars) {
-    plot(data[["height"]], data[[v]], type="l", xlab="", ylab= v, ylim=c(0, max(1,data[[v]])))
+    plot(data[["height"]], data[[v]], type="l", xlab="", ylab= v, ylim=c(0, max(1,data[[v]], na.rm=TRUE)))
   }
   mtext("Height (m)", 1, cex=1, line=, outer=TRUE)
 }
@@ -81,7 +81,7 @@ figure_rate_vs_size_panels <- function(data, type, path) {
     pdf(filename,width=5, height=5)
     par(oma=c(0,0,0,0), mar=rep(0.1,4))
     plot(data[["height"]], data[[v]], type="l", ann=FALSE, axes=FALSE,
-      ylim=c(0, max(1,data[[type]])),
+      ylim=c(0, max(1,data[[type]], na.rm=TRUE)),
       col="green", lwd=3)
     box()
     dev.off()
@@ -122,7 +122,7 @@ figure_diameter_stem_dt <- function(dat=NULL) {
   lai <- dat$lai
   traits <- dat$traits
 
-  ymax <- max(sapply(dat[traits], function(x) max(x$diameter_stem_dt)))
+  ymax <- max(sapply(dat[traits], function(x) max(x$diameter_stem_dt, na.rm=TRUE)))
   ylim <- c(0, ymax * 1.1)
   cols <- rev(RColorBrewer::brewer.pal(length(lai) + 3, "Blues")[-(1:3)])
   par(mfcol=c(length(diameters), length(traits)),
@@ -291,7 +291,7 @@ trait_effects_plot <- function(d) {
   growth_measures <- info$growth_measures
   dsplit <- split(d, d$size_class)
   trait <- info$trait_name
-  ylim <- lapply(d[growth_measures], range)
+  ylim <- lapply(d[growth_measures], range, na.rm=TRUE)
 
   par(oma=c(6, 6, 2, 1), mar=c(1, 1, 2, 1),
       mfrow=c(length(growth_measures), length(info$size_values)))

### 2015-08-17 R/figures_data.R

@@ -14,9 +14,15 @@ figure_lma_tradeoff <- function(data) {
   col_sm1 <- col_table[data[["biome"]][match(sm1[["groups"]], sites)]]
 
   strategy <- default_strategy()
-  B4 <- strategy$B4
-  k_l0 <- strategy$k_l0
-  lma_0 <- strategy$lma_0
+
+  B4 <- 1.71
+  lma_0 <- 0.1978791
+  k_l0 <- 0.4565855
+
+  # TODO: These values no longer accessible via strategy. See traitecoevo/plant#166
+  # B4 <- strategy$B4
+  # k_l0 <- strategy$k_l0
+  # lma_0 <- strategy$lma_0
   x <- seq_log(10^par("usr")[1], 10^par("usr")[2], length.out=10)
   y <- k_l0 * (x / lma_0)^(-B4)
 

### 2015-08-17 R/tables.R

@@ -1,5 +1,5 @@
-table_tree_parameters <- function(tree_parameters, dest){
-  x <- xtable(tree_parameters[-c(1,2,4,7,10,21,22,25:31),1:4],
+table_plant_parameters <- function(plant_parameters, dest){
+  x <- xtable(plant_parameters[-c(1,2,4,7,10,21,22,25:31),1:4],
               hline.after=c(1),
               align='lp{5cm}lll')
   y <- print(x, sanitize.text.function=as.character,

### 2015-11-19 fitting.Rmd
Use remake's download targets
Removes downloader dependency, simplifies code a bit.
Update to reflect upstream breaking changes in plant
- renaming plant physiology from FFW16 -> FF16
- renaming solver from EBT to SCM (capturing correct method used)
- Introduced at traitecoevo/plant/commit/bf4208066e8f83c31ced90435cc3744b5ea97df6

@@
-# Fitting TREE to observed data
+# Fitting plant to observed data
 
@@
 ```{r, echo=TRUE, results="hide"}
-library(tree2)
+library(plant)
 
@@
                      strategy=default_strategy()) {
+
+  x <- c(0.005, 0.01)
+
+  env <- fixed_environment(1)
+  pl <- run_plant_to_sizes(x, "diameter", strategy, env, time_max=3000)
+
+  strategy=FF16_Strategy()
+  env <- fixed_environment(1)  
+  pl <- FF16_PlantPlus(strategy)
+  res <- grow_plant_to_size(pl, x, "diameter", env, time_max=3000)
+
+  x <- seq(FF16_Plant(strategy)$height, strategy$hmat, length.out=100)
+  env <- fixed_environment(canopy_openness)
+  run_plant_to_sizes(heights, "height", strategy, env, time_max=300)
+
+  pl <- FF16_PlantPlus(strategy)
+  res <- grow_plant_to_size(pl, sizes, size_variable, env, time_max,
+                            warn=FALSE, filter=filter)
+
+run_plant_to_sizes
+
   strategy[names(traits)] <- traits
   strategy[names(pars)] <- pars
-  pp <- grow_plant_to_size(Plant(strategy), x, "diameter",
+  pp <- grow_plant_to_size(pl, x, "diameter",
                            fixed_environment(E))

### 2015-11-19 R/figures.R

@@ -1,16 +1,17 @@
 default_strategy <- function() {
-  FFW16_Strategy(
+  FF16_Strategy(
     hmat = 30.0,
-    c_r1 = 0.8,
-    c_r2 = 20,
-    a1   = 2.17,
-    B1   = 0.546,
-    k_l  = 0.4565855 / 3)
+    a_f1 = 0.8,
+    a_f2 = 20,
+    a_l1 = 2.17,
+    a_l2 = 0.546,
+    k_l  = 0.4565855 / 3
+      )
 }
 
 run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
                                time_max=300, filter=FALSE) {
-  pl <- FFW16_PlantPlus(strategy)
+  pl <- FF16_PlantPlus(strategy)
   res <- grow_plant_to_size(pl, sizes, size_variable, env, time_max,
                             warn=FALSE, filter=filter)
   data.frame(rbind_list(lapply(res$plant, extract_plant_info, env=env)))
@@ -18,7 +19,7 @@ run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
 
 extract_plant_info <- function(plant, env) {
   if (is.null(plant)) {
-    x <- unlist(FFW16_PlantPlus(FFW16_Strategy())$internals)
+    x <- unlist(FF16_PlantPlus(FF16_Strategy())$internals)
     x[] <- NA_real_
   } else {
     plant$compute_vars_phys(env)
@@ -35,7 +36,7 @@ extract_plant_info <- function(plant, env) {
 ## TODO: Can grader help here?
 figure_trait_derivative <- function(type, trait_name="lma", canopy_openness=1,
                                     strategy=default_strategy()) {
-  p0 <- FFW16_Parameters(strategy_default=strategy)
+  p0 <- FF16_Parameters(strategy_default=strategy)
   ## strategy(trait_matrix( traits, p)
 
   ## TODO: This seems tragically broken because it apparently computes
@@ -90,7 +91,7 @@ figure_rate_vs_size_panels <- function(data, type, path) {
 
 figure_rate_vs_size_data <- function(canopy_openness=1,
                                      strategy=default_strategy()) {
-  heights <- seq(FFW16_Plant(strategy)$height, strategy$hmat, length.out=100)
+  heights <- seq(FF16_Plant(strategy)$height, strategy$hmat, length.out=100)
   env <- fixed_environment(canopy_openness)
   run_plant_to_sizes(heights, "height", strategy, env, time_max=300)
 }
[…]

### 2016-02-19 ms.tex
Minor fixes, works end to end with plant
Closes #39
Generalise growth_light plot to use any size metric

@@
 light \citep{ruger-2012}. This is exactly the behaviour
-exhibited in our model (Fig. \ref{f-growth_light}), at least for small
+exhibited in our model (Fig. \ref{f-growth_light_dia}), at least for small
 to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
-of low $\phi$ are lost (Fig. \ref{f-growth_light}). This results also
+of low $\phi$ are lost (Fig. \ref{f-growth_light_dia}). This results also
 agrees with empirical results showing that growth for species with low
@@
 \centering
-\includegraphics{figures/growth_light.pdf}
+\includegraphics{figures/growth_light_dia.pdf}
 \caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment.} Panels show predicted relationship between
+in light environment (diameter growth).} Panels show predicted relationship between
 specific trait and diameter growth rate, for a plant of specified
 diameter and under a range of shading environments.
-\label{f-growth_light}}
+\label{f-growth_light_dia}}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics{figures/growth_light_mass.pdf}
+\caption{\textbf{Traits moderate the responsiveness of growth to changes
+in light environment (mass growth).} Panels show predicted relationship between
+specific trait and diameter growth rate, for a plant of specified
+diameter and under a range of shading environments.
+\label{f-growth_light_mass}}
 \end{figure}
@@
 \begin{doublespace}
-\include{figures/table-pars}
+% \include{figures/table-pars}
 

### 2016-02-19 R/figures.R

@@ -115,56 +115,75 @@ figure_rate_vs_size_cols <- function(type) {
   }
 }
 
-figure_diameter_stem_dt <- function(dat=NULL) {
-  if (is.null(dat)) {
-    dat <- figure_diameter_stem_dt_data()
-  }
-  diameters <- dat$diameters
-  lai <- dat$lai
-  traits <- dat$traits
+figure_dY_dt <- function(dat) {
 
-  ymax <- max(sapply(dat[traits], function(x) max(x$diameter_stem_dt, na.rm=TRUE)))
+  vars <- dat[["vars"]]
+  sizes <- dat[["sizes"]]
+  lai <- dat[["lai"]]
+  traits <- dat[["traits"]]
+
+  ymax <- max(sapply(dat[traits], function(x) max(x[[vars[2]]], na.rm=TRUE)))
   ylim <- c(0, ymax * 1.1)
   cols <- rev(RColorBrewer::brewer.pal(length(lai) + 3, "Blues")[-(1:3)])
-  par(mfcol=c(length(diameters), length(traits)),
+  par(mfcol=c(length(sizes), length(traits)),
       oma=c(4, 5, 0, 1.5), mar=c(1, 1, 1, 1))
   for (v in traits) {
-    dat_v <- unname(split(dat[[v]], dat[[v]]$diameter_class))
-    for (i in seq_along(diameters)) {
+    dat_v <- unname(split(dat[[v]], dat[[v]]$class))
+    for (i in seq_along(sizes)) {
       dsub <- long_to_wide(dat_v[[i]], "canopy_openness",
-                           c(v, "diameter_stem_dt"))
+                           c(v, vars[2]))
       matplot(dsub[[1]], dsub[[2]], type="l", col=cols, lty=1, log="x",
-              ylim=ylim, xaxt="n", yaxt="n", xlab="", ylab="")
-      axis(1, labels=i == length(diameters), las=1)
+              #ylim = ylim,
+              xaxt="n", yaxt="n", xlab="", ylab="")
+      axis(1, labels=i == length(sizes), las=1)
       axis(2, labels=v == "lma", las=1)
       if (v == last(traits)) {
-        mtext(sprintf("D=%sm", diameters[[i]]), 4, cex=1, line=1)
+        mtext(dat[["label"]](sizes[[i]]), 4, cex=1, line=1)
       }
     }
[…]

### 2016-07-28 ms.tex
Update build process, remove docker
- import build functions used in other papers

@@
 Table near intro summarising empricial patterns to be explained by integrated model?
+
+Section 3.3 of King (2005) shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
+
+King 1999 also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
+Basically, the much greater path length from roots to leaves in trees vs.
+seedlings results in a much greater increment in support cost per increment in crown area for trees.
+This explanation is treated by the growth model of King (1999), which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
 
@@
 
+(reich 1992, Niinements 2010, Gower-1993 - both show higher LAI in high LMA / long LL stands)
+
 Ulitmately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
@@
 
-\subsubsection{Responisveness to light}
+\subsubsection{Responsiveness to light}
 

### 2016-07-28 R/build.R

@@ -1,0 +1,76 @@
+#!/usr/bin/env Rscript
+
+latex_build <- function(filename, bibliography=NULL, output=NULL,
+                        chdir=TRUE, interaction="nonstopmode",
+                        max_attempts=5L, clean=FALSE, engine="pdflatex") {
+  if (chdir && dirname(filename) != "") {
+    owd <- setwd(dirname(filename))
+    on.exit(setwd(owd))
+    filename <- basename(filename)
+  }
+
+  res <- run_latex(filename, interaction, engine)
+  if(engine=="xelatex") {
+      res <- run_latex(filename, interaction, engine)
+  }
+  if (!is.null(bibliography)) {
+    run_bibtex(filename)
+    res <- run_latex(filename, interaction, engine)
+  }
+
+  Encoding(res) <- "UTF-8"
+
+  pat <- c("Rerun to get cross-references right", # labels
+           "Rerun to get citations correct",      # bibtex
+           "Rerun to get outlines right")         # tikz
+  isin <- function(p, x) {
+    any(grepl(p, x))
+  }
+  for (i in seq_len(max_attempts)) {
+    if (any(vapply(pat, isin, logical(1), res))) {
+      res <- run_latex(filename, interaction, engine, output_directory)
+    } else {
+      break
+    }
+  }
+
+  if (clean) {
+    latex_clean(filename)
+  }
+
+  if(!is.null(output)) {
+    fileout <- sub(".tex$", ".pdf", filename)
+    file.copy(fileout, output)
+    file.remove(fileout)
+  }
+
+  invisible(NULL)
+}
+
+latex_clean <- function(filename) {
+  filebase <- sub(".tex$", "", filename)
+  exts <- c(".log", ".aux", ".bbl", ".blg", ".fls", ".out", ".snm",
+            ".nav", ".tdo", ".toc")
+  aux <- paste0(filebase, exts)
+  file.remove(aux[file.exists(aux)])
+}
+
+run_latex <- function(filename, interaction="nonstopmode", engine="pdflatex") {
+  args <- c(paste0("-interaction=", interaction),
+            "-halt-on-error",
+            filename)
+  callr::call_system(Sys_which(engine), args)
+}
+
+run_bibtex <- function(filename) {
+  callr::call_system(Sys_which("bibtex"), sub(".tex$", "", filename))
+}
+
+Sys_which <- function(x) {
+  ret <- Sys.which(x)
[…]

### 2016-07-28 R/ms-fun.R

@@ -1,21 +1,0 @@
-# interaction must be one of: batchmode, nonstopmode, scrollmode, errorstopmode
-# default combination causes latex to bail back to system on error
-
-latex_build <- function(filename, latex_engine="xelatex", bibtex=TRUE, dest_dir = ".", interaction="nonstopmode") {
-  build <- sprintf("%s -interaction=%s -halt-on-error -output-directory=%s %s", latex_engine, interaction, dest_dir, filename)
-  basename <- tools::file_path_sans_ext(filename)
-  system(build)
-  if(bibtex){
-	  system(sprintf("bibtex %s",  basename))
-    system(build)
-  }
-  system(build)
-  remove_aux_files(file.path(dest_dir, basename(filename)))
-}
-
-remove_aux_files <- function(filename) {
-	aux.files <- paste0(tools::file_path_sans_ext(filename),
-                     c(".log", ".aux", ".bbl", ".blg", ".fls", ".out", ".snm", ".nav", ".toc",
-                      ".fdb_latexmk"))
-  file.remove(aux.files[file.exists(aux.files)])
-}

### 2016-07-28 docker/growth_trajectories/launch.sh

@@ -1,4 +1,0 @@
-#!/bin/bash
-set -e
-$(boot2docker shellinit 2> /dev/null)
-docker run -v ${PWD}/self:/root/growth_trajectories -it traitecoevo/growth_trajectories $*

### 2016-07-28 figure_conceptual.tex
Manuscript in subdirectory, all outputs in "output"

--- 2015-04-29 figure_conceptual.tex
@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_mass_production_net_production}{}
-\othercol{2}{$y$}{figures/empty_box}{$\times$}
-\othercol{3}{$\Bigg(A$}{figures/F2_mass_production_assimilation}{-}
-\othercol{4}{\ $R$}{figures/F2_mass_production_respiration}{$\Bigg)$}
-\othercol{5}{$-T$}{figures/F2_mass_production_turnover}{}
+\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{output/F2_net_mass_production_dt_net_mass_production_dt}{}
+\othercol{2}{$y$}{output/empty_box}{$\times$}
+\othercol{3}{$\Bigg(A$}{output/F2_net_mass_production_dt_assimilation}{-}
+\othercol{4}{\ $R$}{output/F2_net_mass_production_dt_respiration}{$\Bigg)$}
+\othercol{5}{$-T$}{output/F2_net_mass_production_dt_turnover}{}
 \rowlabel{(1)}
@@
 \rowtitle{Construction costs}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\Bigg($}
-\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\Bigg)^{-1}$}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{output/F2_darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
+\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
+\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
+\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_root_darea_leaf}{$\Bigg)^{-1}$}
  \rowlabel{(2)}
-
 \defineRow{2}{thisrow}
 \rowtitle{Leaf-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_leaf_area_growth_net_production}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/F2_area_leaf_dt_area_leaf_dt}{}
+\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{output/F2_area_leaf_dt_darea_leaf_dmass_live}{$\times$}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{output/F2_area_leaf_dt_fraction_allocation_growth}{$\times$}
+\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{output/F2_area_leaf_dt_net_mass_production_dt}{}
 \rowlabel{(3)}
@@
 \rowtitle{Height growth}
-\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/F2_height_growth_height_growth_rate}{}
-\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_height_growth_dheight_dleaf_area}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{output/F2_height_dt_height_dt}{}
+\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{output/F2_height_dt_dheight_darea_leaf}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/F2_area_leaf_dt_area_leaf_dt}{}
 \rowlabel{(4)}
@@
 \rowtitle{Basal-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_basal_growth_dbasal_area_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{+}
-\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{figures/empty_box}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/F2_area_stem_dt_area_stem_dt}{}
+\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/empty_box}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/F2_area_stem_dt_area_leaf_dt}{+}
+\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{output/F2_area_stem_dt_area_heartwood_dt}{}
 \rowlabel{(5)}
@@
 \rowtitle{Diameter growth}
-\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/F2_diameter_growth_dbasal_diam_dt}{}
-\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_basal_growth_dbasal_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{output/F2_diameter_stem_dt_diameter_stem_dt}{}
+\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{output/F2_diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/F2_diameter_stem_dt_area_stem_dt}{}
 \rowlabel{(6)}

### 2016-07-28 ms/ms.tex
Manuscript in subdirectory, all outputs in "output"

--- 2016-07-28 ms.tex
@@
 \usepackage{graphicx}
+\graphicspath{{../output/}}
+
 % We will generate all images so they have a width \maxwidth. This means
@@
 \centering
-\includegraphics{figures/BCI_data.pdf}
+\includegraphics{BCI_data.pdf}
 \caption{\textbf{The relationship between traits and potential growth
@@
 \centering
-\includegraphics{figures/figure_conceptual.pdf}
+\includegraphics{figure_conceptual.pdf}
 \caption{\textbf{Conceptual framework for understanding size-dependent
@@
 \centering
-\includegraphics{figures/allometry.pdf}
+\includegraphics{allometry.pdf}
 \caption{\textbf{Key assumptions of a functional balance and trait
@@
 \centering
-\includegraphics{figures/growth_light_dia.pdf}
+\includegraphics{growth_light_dia.pdf}
 \caption{\textbf{Traits moderate the responsiveness of growth to changes
@@
 \centering
-\includegraphics{figures/growth_light_mass.pdf}
+\includegraphics{growth_light_mass.pdf}
 \caption{\textbf{Traits moderate the responsiveness of growth to changes
@@
 \centering
-\includegraphics{figures/max_leaf_above.pdf}
+\includegraphics{max_leaf_above.pdf}
 \caption{\textbf{The effect of traits on growth changes with size and
@@
 \centering
-\includegraphics{figures/max_leaf_above.pdf}
+\includegraphics{max_leaf_above.pdf}
 \caption{\textbf{Low construction cost leads to shade intolerance,
@@
 \begin{doublespace}
-% \include{figures/table-pars}
+% \include{table-pars}
 
@@
 \centering
-\includegraphics{figures/SI_quantile_examples.pdf}
+\includegraphics{SI_quantile_examples.pdf}
 \caption{\textbf{Estimating potential growth rate via quantile
@@
 \centering
-\includegraphics{figures/SI_lma_tradeoff.pdf}
+\includegraphics{SI_lma_tradeoff.pdf}
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
@@
 \centering
-\includegraphics{figures/SI_mass_fraction.pdf}
+\includegraphics{SI_mass_fraction.pdf}
 \caption{\textbf{Change in allocation with size.}
@@
 \centering
-\includegraphics{figures/SI_lma_effects_at_diameters.pdf}
+\includegraphics{SI_lma_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
@@
 \centering
-\includegraphics{figures/SI_lma_effects_at_diameters_relative.pdf}
+\includegraphics{SI_lma_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
@@
 \centering
-\includegraphics{figures/SI_rho_effects_at_diameters.pdf}
+\includegraphics{SI_rho_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between stem-construction cost
@@
 \centering
-\includegraphics{figures/SI_rho_effects_at_diameters_relative.pdf}
+\includegraphics{SI_rho_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between stem-construction cost
@@
 \centering
-\includegraphics{figures/SI_hmat_effects_at_diameters.pdf}
+\includegraphics{SI_hmat_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between maximum height and
@@
 \centering
-\includegraphics{figures/SI_hmat_effects_at_diameters_relative.pdf}
+\includegraphics{SI_hmat_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between maximum height and

### 2016-07-28 R/build.R

@@ -3,7 +3,10 @@
 latex_build <- function(filename, bibliography=NULL, output=NULL,
                         chdir=TRUE, interaction="nonstopmode",
                         max_attempts=5L, clean=FALSE, engine="pdflatex") {
-  if (chdir && dirname(filename) != "") {
+
+  fileout <- sub(".tex$", ".pdf", filename)
+
+  if (chdir) {
     owd <- setwd(dirname(filename))
     on.exit(setwd(owd))
     filename <- basename(filename)
@@ -28,7 +31,7 @@ latex_build <- function(filename, bibliography=NULL, output=NULL,
   }
   for (i in seq_len(max_attempts)) {
     if (any(vapply(pat, isin, logical(1), res))) {
-      res <- run_latex(filename, interaction, engine, output_directory)
+      res <- run_latex(filename, interaction, engine)
     } else {
       break
     }
@@ -39,8 +42,10 @@ latex_build <- function(filename, bibliography=NULL, output=NULL,
   }
 
   if(!is.null(output)) {
-    fileout <- sub(".tex$", ".pdf", filename)
-    file.copy(fileout, output)
+    if (chdir) {
+      setwd(owd)
+    }
+    file.copy(fileout, output, TRUE)
     file.remove(fileout)
   }
 

### 2016-07-28 ms/ms.tex
Remove data component

@@
 
-% allows for numbered referencing, as refquired by ecol letters
+% allows for numbered referencing, as required by ecol letters
 \usepackage{etoolbox}
@@
 \author[1]{Richard G. FitzJohn}
-\author[2]{Joe Wright}
 
 \affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
-\affil[2]{{\footnotesize Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama}}
 \renewcommand\Authands{ and }
@@
 
-Ulitmately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
+Ultimately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
 Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{sillett-2010, king-2011} or diameter growth \citep{herault-2011}.
@@
 
-The results of this new analysis
-(Fig. \ref{f-BCI}) align closely with our theoretical predictions  showing how the effect of traits on growth
-changes with plant size, with different traits exhibiting distinct
-size-dependent effects.
-
-\subsection{Empirical motivation}\label{fresh-empirical-motivation}
-
-To help illustrate the empirical patterns, we begin with a fresh analysis of
-growth dynamics in the long-term plot at Barro Colorado Island in Panama,
-where the growth rates for over 100 species have been recorded over the last
-30 years. Data from a long term plot \citep{condit-2012} are combined
-with results of experiment on seedlings \citep{kitajima-2013}, to better understand
-size-related patterns of growth. A prior analysis by
- \citet{wright-2010} found a moderate effect of wood density on
-growth rate, but almost no effect of LMA, and only a weak effect of HMAX. More recently,
- \citet{ruger-2012} showed
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Effect of lma stronger than previously thought,
-\item
-  Bridges from seedling studies to adults
-\end{itemize}
-
-Results stronger than previously reported.
-
-Predictions
-are compared against data from both a long-term
-plot \citep{condit-2012} and field
-experiment \citep{kitajima-2013} at Barro Colorado Island, Panama.
-We use a new method to estimate species potential growth rate,
-accounting for size specific effects (Fig. \ref{fS-quantile_examples},
-see Methods for details). This new method allowed us to extract a more
-definite signal of trait effects than in previous
-studies \citep{wright-2010}.
-
 \section{Materials and Methods}\label{materials-and-methods}
@@
 \ref{eq:dDdt} predict a relationship between $\phi$ and growth that
-changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
+changes with plant size ( \ref{f-lma_growth_size}).
 Decreasing $\phi$ has two impacts on growth rate. First, lower
@@
 $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ . At smaller sizes, there is no differentiation
-among species, because most species are focusing on growth (Fig.
-\ref{f-BCI}). At larger sizes, individuals of some species are
+among species, because most species are focusing on growth. At larger sizes, individuals of some species are
 allocating a majority of their surplus energy to growth, leaving a
-strong signal in potential growth rate (Fig. \ref{f-BCI}).
+strong signal in potential growth rate.
 
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics{BCI_data.pdf}
-\caption{\textbf{The relationship between traits and potential growth
-rate varies with plant size.} For XXX species from tropical rain forest
-in Panama, we estimated the potential growth rate of individual's in
-each species at a series of diameters $D$, indicated along right hand
-side. The size of circles in each panel indicates the number of data
-points used to estimate potential growth rate. Traits values were
-calculated from representative individuals in each species.
-\label{f-BCI}}
-\end{figure}
-
 \newpage
@@
 \centering
-\includegraphics{SI_quantile_examples.pdf}
-\caption{\textbf{Estimating potential growth rate via quantile
-regression.} Example for 4 species. Growth rate estimated at six sizes,
-by fitting lines within given size intervals. Lines fitted to 99th
-quantile, but only when n \textgreater{} 500. Except for left-hand
-panel, estimated points in centre of interval.
-\label{fS-quantile_examples}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
 \includegraphics{SI_lma_tradeoff.pdf}

### 2016-07-28 R/BCI.R

@@ -1,305 +1,0 @@
-# List of functions used to manipulate and summarise data from BCI.
-
-# Load 50ha census data
-BCI_load_50ha_plot <- function(path_to_zip) {
-  tmp <- tempfile()
-  unzip(path_to_zip, exdir=tmp)
-  on.exit(unlink(tmp, recursive=TRUE))
-
-  files <- list.files(tmp, pattern=".rdata", full.names=TRUE)
-  tbl_df(ldply(list.files(tmp, pattern=".rdata", full.names=TRUE), function(x) load_rdata(x)))
-}
-
-BCI_calculate_individual_growth <- function(BCI_50haplot, BCI_nomenclature) {
-
-  # Identifies individuals that return from the dead or are supposedly re-found
-  # i.e. Individuals given dbh=NA and then later given numeric value
-  # Note this function must be used prior to sub-setting only observations with pom = 1.3
-  is_zombie <- function(dbh) {
-    any(diff(is.na(dbh)) == -1)
-  }
-
-  names(BCI_50haplot) <- tolower(names(BCI_50haplot)) # lower case for all column names
-
-  data <- BCI_50haplot %>%
-    arrange(sp, treeid, date) %>%
-    select(sp, treeid, nostems, censusid,exactdate, dfstatus, pom, dbh) %>%
-    mutate(
-      #census id for period 7 was entered incorrectly
-      censusid = ifelse(censusid==171, 7,censusid),
-      species = lookup_latin(sp, BCI_nomenclature),
-      family = lookup_family(sp, BCI_nomenclature),
-      dbh=dbh/1000) %>%
-    # Remove stems from earlier census, measured with course resolution
-    filter(censusid >= 3) %>%
-    # Remove families that don't exhibit dbh growth e.g. palms
-    filter(!family %in% c('Arecaceae', 'Cyatheaceae', 'Dicksoniaceae', 'Metaxyaceae',
-                          'Cibotiaceae', 'Loxomataceae', 'Culcitaceae', 'Plagiogyriaceae',
-                          'Thyrsopteridaceae')) %>%
-    # Remove observations without a species code
-    filter(!is.na(sp)) %>%
-    # For each individual..
-    group_by(treeid) %>%
-    # Filter plants with multiple stems
-    filter(max(nostems)==1) %>%
[…]

### 2016-07-28 R/wright_2010.R

@@ -1,47 +1,0 @@
-process_wright_2010 <- function(filename) {
-  # Correct names of species, anything with length > 12 was clipped
-  corrections <- data.frame(rbind(c("costaricensi","costaricensis"),
-                                  c("Beilschmiedi","Beilschmiedia"),
-                                  c("megalophyllu","megalophyllum"),
-                                  c("Chrysochlamy","Chrysochlamys"),
-                                  c("Chrysophyllu","Chrysophyllum"),
-                                  c("manzinellens","manzinellensis"),
-                                  c("billbergianu","billbergianus"),
-                                  c("costaricensi","costaricensis"),
-                                  c("coloradoensi","coloradoensis"),
-                                  c("alchorneoide","alchorneoides"),
-                                  c("Tabernaemont","Tabernaemontana"),
-                                  c("Trattinnicki","Trattinnickia"),
-                                  c("princeps", "croatii"),
-                                  c("$",""),
-                                  c("-99.000000", "NA"),
-                                  c("-99.00000", "NA"),
-                                  c("-99.0000", "NA"),
-                                  c("-99.000", "NA"),
-                                  c("-99.00", "NA"),
-                                  c("-99.0", "NA"),
-                                  c("-99", "NA"),
-                                  c("\t", ",")), stringsAsFactors = FALSE)
-  names(corrections) <- c("OLD", "NEW")
-
-  x <- readLines(filename)[26:160]
-
-  for (i in 1:nrow(corrections)) {
-    x <- gsub(corrections[["OLD"]][i], corrections[["NEW"]][i], x, fixed = TRUE)
-  }
-
-  ## This is a hack and should be looked at:
-  f <- tempfile()
-  on.exit(file.remove(f))
-  cat(x, file = f, sep = "\n")
-  data <- read.csv(f, stringsAsFactors = FALSE)
-
-  names(data) <- tolower(names(data))
-  names(data)[names(data) == "wsg"] <- "rho"
-  names(data)[names(data) == "height"] <- "hmat"
-
-  data[["lma"]] <- data[["lma"]] / 1000  #(convert from g/2 to kg/m2)
-  data[["rho"]] <- data[["rho"]] * 1000
-
-  data
-}

### 2016-07-28 figure_conceptual.tex
put figures from main & SI in subfolders

--- 2015-04-30 figure_conceptual.tex
@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_net_mass_production_dt_net_production}{}
-\othercol{2}{$y$}{figures/empty_box}{$\times$}
-\othercol{3}{$\Bigg(A$}{figures/F2_net_mass_production_dt_assimilation}{-}
-\othercol{4}{\ $R$}{figures/F2_net_mass_production_dt_respiration}{$\Bigg)$}
-\othercol{5}{$-T$}{figures/F2_net_mass_production_dt_turnover}{}
+\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{output/con/net_mass_production_dt_net_mass_production_dt}{}
+\othercol{2}{$y$}{output/con/empty_box}{$\times$}
+\othercol{3}{$\Bigg(A$}{output/con/net_mass_production_dt_assimilation}{-}
+\othercol{4}{\ $R$}{output/con/net_mass_production_dt_respiration}{$\Bigg)$}
+\othercol{5}{$-T$}{output/con/net_mass_production_dt_turnover}{}
 \rowlabel{(1)}
@@
 \rowtitle{Construction costs}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\Bigg($}
-\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{+}
-\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\Bigg)^{-1}$}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{output/con/darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
+\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
+\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
+\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/darea_leaf_dmass_live_dmass_root_darea_leaf}{$\Bigg)^{-1}$}
  \rowlabel{(2)}
-
 \defineRow{2}{thisrow}
 \rowtitle{Leaf-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{figures/F2_leaf_area_growth_leaf_fraction}{$\times$}
-\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{figures/F2_leaf_area_growth_net_production}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
+\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{output/con/area_leaf_dt_darea_leaf_dmass_live}{$\times$}
+\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{output/con/area_leaf_dt_fraction_allocation_growth}{$\times$}
+\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{output/con/area_leaf_dt_net_mass_production_dt}{}
 \rowlabel{(3)}
@@
 \rowtitle{Height growth}
-\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{figures/F2_height_dt_height_dt_rate}{}
-\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{figures/F2_height_dt_dheight_dleaf_area}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{output/con/height_dt_height_dt}{}
+\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{output/con/height_dt_dheight_darea_leaf}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
 \rowlabel{(4)}
@@
 \rowtitle{Basal-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_dbasal_area_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{figures/F2_leaf_area_growth_dleaf_area_dt}{+}
-\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{figures/empty_box}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_stem_dt}{}
+\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/empty_box}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_leaf_dt}{+}
+\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_heartwood_dt}{}
 \rowlabel{(5)}
@@
 \rowtitle{Diameter growth}
-\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{figures/F2_diameter_stem_dt_dbasal_diam_dt}{}
-\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{figures/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{figures/F2_area_stem_dt_dbasal_area_dt}{}
+\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{output/con/diameter_stem_dt_diameter_stem_dt}{}
+\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{output/con/diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/con/diameter_stem_dt_area_stem_dt}{}
 \rowlabel{(6)}

### 2016-07-28 ms/ms.tex
put figures from main & SI in subfolders

@@
 \centering
-\includegraphics{figure_conceptual.pdf}
+\includegraphics{main/figure_conceptual.pdf}
 \caption{\textbf{Conceptual framework for understanding size-dependent
@@
 \centering
-\includegraphics{allometry.pdf}
+\includegraphics{main/allometry.pdf}
 \caption{\textbf{Key assumptions of a functional balance and trait
@@
 \centering
-\includegraphics{growth_light_dia.pdf}
+\includegraphics{main/growth_light_dia.pdf}
 \caption{\textbf{Traits moderate the responsiveness of growth to changes
@@
 \centering
-\includegraphics{growth_light_mass.pdf}
+\includegraphics{main/growth_light_mass.pdf}
 \caption{\textbf{Traits moderate the responsiveness of growth to changes
@@
 \centering
-\includegraphics{max_leaf_above.pdf}
+\includegraphics{main/max_leaf_above.pdf}
 \caption{\textbf{The effect of traits on growth changes with size and
@@
 \centering
-\includegraphics{max_leaf_above.pdf}
+\includegraphics{main/max_leaf_above.pdf}
 \caption{\textbf{Low construction cost leads to shade intolerance,
@@
 \centering
-\includegraphics{SI_lma_tradeoff.pdf}
+\includegraphics{SI/lma_tradeoff.pdf}
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
@@
 \centering
-\includegraphics{SI_mass_fraction.pdf}
+\includegraphics{SI/mass_fraction.pdf}
 \caption{\textbf{Change in allocation with size.}
@@
 \centering
-\includegraphics{SI_lma_effects_at_diameters.pdf}
+\includegraphics{SI/lma_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
@@
 \centering
-\includegraphics{SI_lma_effects_at_diameters_relative.pdf}
+\includegraphics{SI/lma_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
@@
 \centering
-\includegraphics{SI_rho_effects_at_diameters.pdf}
+\includegraphics{SI/rho_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between stem-construction cost
@@
 \centering
-\includegraphics{SI_rho_effects_at_diameters_relative.pdf}
+\includegraphics{SI/rho_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between stem-construction cost
@@
 \centering
-\includegraphics{SI_hmat_effects_at_diameters.pdf}
+\includegraphics{SI/hmat_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between maximum height and
@@
 \centering
-\includegraphics{SI_hmat_effects_at_diameters_relative.pdf}
+\includegraphics{SI/hmat_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between maximum height and

### 2016-07-28 R/figures.R

@@ -78,11 +78,11 @@ figure_rate_vs_size_panels <- function(data, type, path) {
   yvars <- figure_rate_vs_size_cols(type)
 
   for (v in yvars) {
-    filename <- file.path(path, sprintf("F2_%s_%s.pdf", type, v))
+    filename <- file.path(path, sprintf("%s_%s.pdf", type, v))
     pdf(filename,width=5, height=5)
     par(oma=c(0,0,0,0), mar=rep(0.1,4))
     plot(data[["height"]], data[[v]], type="l", ann=FALSE, axes=FALSE,
-      ylim=c(0, max(1,data[[type]], na.rm=TRUE)),
+  #    ylim=c(0, max(1,data[[type]], na.rm=TRUE)),
       col="green", lwd=3)
     box()
     dev.off()

### 2016-07-28 R/utils.R

@@ -1,3 +1,4 @@
+last <- plant:::last
 
 read_csv <- function(filename, ...){
   read.csv(filename, stringsAsFactors=FALSE, ...)

### 2016-07-28 figure_conceptual.tex
Add missing panels from conceptual figure

@@
 \firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{output/con/net_mass_production_dt_net_mass_production_dt}{}
-\othercol{2}{$y$}{output/con/empty_box}{$\times$}
-\othercol{3}{$\Bigg(A$}{output/con/net_mass_production_dt_assimilation}{-}
-\othercol{4}{\ $R$}{output/con/net_mass_production_dt_respiration}{$\Bigg)$}
-\othercol{5}{$-T$}{output/con/net_mass_production_dt_turnover}{}
+\othercol{2}{$\alpha_y$}{output/con/net_mass_production_dt_yield}{$\times$}
+\othercol{3}{$\Bigg(\textrm{P}$}{output/con/net_mass_production_dt_assimilation}{-}
+\othercol{4}{\ $\textrm{R}$}{output/con/net_mass_production_dt_respiration}{$\Bigg)$}
+\othercol{5}{$-\textrm{T}$}{output/con/net_mass_production_dt_turnover}{}
 \rowlabel{(1)}
@@
 \rowtitle{Basal-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_stem_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_leaf_dt}{+}
-\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_heartwood_dt}{}
+\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_stem_dt}{$\Bigg($}
+\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/area_stem_dt_darea_sapwood_darea_leaf}{$+$}
+\othercol{3}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/con/area_stem_dt_darea_bark_darea_leaf}{$\Bigg)$}
+\othercol{4}{\hspace{-0.2em}$\times\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_leaf_dt}{$+$}
+\othercol{5}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{output/con/area_stem_dt_area_heartwood_dt}{}
 \rowlabel{(5)}
@@
 \firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{output/con/diameter_stem_dt_diameter_stem_dt}{}
-\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{output/con/diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
+\othercol{2}{$\frac{D}{\textrm{d}a_{\textrm{st}}}$}{output/con/diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
 \othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/con/diameter_stem_dt_area_stem_dt}{}

### 2016-07-28 R/figures.R

@@ -75,13 +75,22 @@ figure_rate_vs_size <- function(data, type) {
 }
 
 figure_rate_vs_size_panels <- function(data, type, path) {
+
+  strategy=default_strategy()
+
   yvars <- figure_rate_vs_size_cols(type)
+
+  data[["yield"]] <- strategy[["a_y"]]
+  data[["dmass_leaf_darea_leaf"]] <- data[["mass_leaf"]] / data[["area_leaf"]]
+  data[["darea_sapwood_darea_leaf"]] <- data[["area_sapwood"]] / data[["area_leaf"]]
+  data[["darea_bark_darea_leaf"]] <- data[["area_bark"]] / data[["area_leaf"]]
 
   for (v in yvars) {
     filename <- file.path(path, sprintf("%s_%s.pdf", type, v))
     pdf(filename,width=5, height=5)
     par(oma=c(0,0,0,0), mar=rep(0.1,4))
     plot(data[["height"]], data[[v]], type="l", ann=FALSE, axes=FALSE,
+       xlim=c(0, 25),
   #    ylim=c(0, max(1,data[[type]], na.rm=TRUE)),
       col="green", lwd=3)
     box()
@@ -98,7 +107,7 @@ figure_rate_vs_size_data <- function(canopy_openness=1,
 
 figure_rate_vs_size_cols <- function(type) {
   if (type == "net_mass_production_dt") {
-    c("net_mass_production_dt", "assimilation", "respiration", "turnover")
+    c("net_mass_production_dt", "yield", "assimilation", "respiration", "turnover")
   } else if (type == "darea_leaf_dmass_live") {
     c("darea_leaf_dmass_live", "dmass_leaf_darea_leaf", "dmass_sapwood_darea_leaf", "dmass_bark_darea_leaf", "dmass_root_darea_leaf")
   } else if (type == "area_leaf_dt") {
@@ -106,7 +115,7 @@ figure_rate_vs_size_cols <- function(type) {
   } else if (type == "height_dt") {
     c("height_dt", "dheight_darea_leaf", "area_leaf_dt")
   } else if (type == "area_stem_dt") {
-    c("area_stem_dt", "darea_leaf_dmass_live", "area_leaf_dt", "area_heartwood_dt")
+    c("area_stem_dt", "darea_sapwood_darea_leaf", "darea_bark_darea_leaf", "area_leaf_dt", "area_heartwood_dt")
[…]

### 2016-07-28 figure_conceptual.tex
Standardise maths with Falster et al 2016, Gibert et al 2016

--- 2016-07-28 figure_conceptual.tex
@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{\textrm{d}b}{\textrm{d}t}$}{output/F2_net_mass_production_dt_net_mass_production_dt}{}
-\othercol{2}{$y$}{output/empty_box}{$\times$}
-\othercol{3}{$\Bigg(A$}{output/F2_net_mass_production_dt_assimilation}{-}
-\othercol{4}{\ $R$}{output/F2_net_mass_production_dt_respiration}{$\Bigg)$}
-\othercol{5}{$-T$}{output/F2_net_mass_production_dt_turnover}{}
+\firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/net_mass_production_dt_net_mass_production_dt}{}
+\othercol{2}{$\alpha_y$}{output/con/net_mass_production_dt_yield}{$\times$}
+\othercol{3}{$\Bigg({\rm P}$}{output/con/net_mass_production_dt_assimilation}{$-$}
+\othercol{4}{${\rm R}$}{output/con/net_mass_production_dt_respiration}{$\Bigg)$}
+\othercol{5}{$-{\rm T}$}{output/con/net_mass_production_dt_turnover}{}
 \rowlabel{(1)}
@@
 \rowtitle{Construction costs}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{output/F2_darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
-\othercol{2}{$\frac{\textrm{d}m_{\textrm{l}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
-\othercol{4}{$\frac{\textrm{d}m_{\textrm{sb}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
-\othercol{5}{$\frac{\textrm{d}m_{\textrm{r}}}{\textrm{d}a_{\textrm{l}}}$}{output/F2_darea_leaf_dmass_live_dmass_root_darea_leaf}{$\Bigg)^{-1}$}
+\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm t}}}$}{output/con/darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
+\othercol{2}{$\frac{{\rm d}M_{{\rm l}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
+\othercol{3}{$\frac{{\rm d}M_{\textrm{ss}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
+\othercol{4}{$\frac{{\rm d}M_{{\rm sb}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
+\othercol{5}{$\frac{{\rm d}M_{{\rm r}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_root_darea_leaf}{$\Bigg)^{-1}$}
  \rowlabel{(2)}
@@
 \rowtitle{Leaf-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/F2_area_leaf_dt_area_leaf_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}m_{\textrm{t}}}$}{output/F2_area_leaf_dt_darea_leaf_dmass_live}{$\times$}
-\othercol{3}{$\frac{\textrm{d}m_{\textrm{t}}}{\textrm{d}b}$}{output/F2_area_leaf_dt_fraction_allocation_growth}{$\times$}
-\othercol{4}{$\frac{\textrm{d}b}{\textrm{d}t}$}{output/F2_area_leaf_dt_net_mass_production_dt}{}
+\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
+\othercol{2}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm t}}}$}{output/con/area_leaf_dt_darea_leaf_dmass_live}{$\times$}
+\othercol{3}{$\frac{{\rm d}M_{{\rm t}}}{{\rm d}B}$}{output/con/area_leaf_dt_fraction_allocation_growth}{$\times$}
+\othercol{4}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/area_leaf_dt_net_mass_production_dt}{}
 \rowlabel{(3)}
@@
 \rowtitle{Height growth}
-\firstcol{1}{$\frac{\textrm{d}h}{\textrm{d}t}$}{output/F2_height_dt_height_dt}{}
-\othercol{2}{$\frac{\textrm{d}h}{\textrm{d}a_{\textrm{l}}}$}{output/F2_height_dt_dheight_darea_leaf}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/F2_area_leaf_dt_area_leaf_dt}{}
+\firstcol{1}{$\frac{{\rm d}H}{{\rm d}t}$}{output/con/height_dt_height_dt}{}
+\othercol{2}{$\frac{{\rm d}H}{{\rm d}A_{{\rm l}}}$}{output/con/height_dt_dheight_darea_leaf}{$\times$}
+\othercol{3}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
 \rowlabel{(4)}
@@
 \rowtitle{Basal-area growth}
-\firstcol{1}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/F2_area_stem_dt_area_stem_dt}{}
-\othercol{2}{$\frac{\textrm{d}a_{\textrm{ss}}}{\textrm{d}a_{\textrm{l}}}$}{output/empty_box}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{l}}}{\textrm{d}t}$}{output/F2_area_stem_dt_area_leaf_dt}{+}
-\othercol{4}{$\frac{\textrm{d}a_{\textrm{sh}}}{\textrm{d}t}$}{output/F2_area_stem_dt_area_heartwood_dt}{}
+\firstcol{1}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/con/area_stem_dt_area_stem_dt}{$\Bigg($}
+\othercol{2}{$\frac{{\rm d}A_{\textrm{ss}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_sapwood_darea_leaf}{$+$}
+\othercol{3}{$\frac{{\rm d}A_{\textrm{sb}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_bark_darea_leaf}{\hspace{-0.3em}$\Bigg)$}
+\othercol{4}{\hspace{-0.4em}$\times\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_stem_dt_area_leaf_dt}{$+$}
+\othercol{5}{$\frac{{\rm d}A_{{\rm sh}}}{{\rm d}t}$}{output/con/area_stem_dt_area_heartwood_dt}{}
 \rowlabel{(5)}
@@
 \rowtitle{Diameter growth}
-\firstcol{1}{$\frac{\textrm{d}D}{\textrm{d}t}$}{output/F2_diameter_stem_dt_diameter_stem_dt}{}
-\othercol{2}{$\frac{\sqrt{\pi}}{\sqrt{a_{\textrm{st}}}}$}{output/F2_diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
-\othercol{3}{$\frac{\textrm{d}a_{\textrm{st}}}{\textrm{d}t}$}{output/F2_diameter_stem_dt_area_stem_dt}{}
+\firstcol{1}{$\frac{{\rm d}D}{{\rm d}t}$}{output/con/diameter_stem_dt_diameter_stem_dt}{}
+\othercol{2}{$\frac{{\rm d}D}{{\rm d}A_{{\rm st}}}$}{output/con/diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
+\othercol{3}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/con/diameter_stem_dt_area_stem_dt}{}
 \rowlabel{(6)}

### 2016-07-28 ms/ms.tex
Standardise maths with Falster et al 2016, Gibert et al 2016

@@
 model describing how the various mass and area terms
-($m_\textrm{l}, m_\textrm{s}, m_\textrm{b}, m_\textrm{r}, a_\textrm{l}, a_\textrm{s}, a_\textrm{b}$)
+($M_{\rm l}, M_\textrm{s}, M_{\rm b}, M_{\rm r}, A_{\rm l}, A_\textrm{s}, A_{\rm b}$)
 vary relative to one another; and 2) For any trade-offs in plant
@@
 We begin with a standard model for the amount of biomass available for
-growth, $\textrm{d}b / \textrm{d}t$, given by the difference between income
+growth, ${\rm d}B / {\rm d}t$, given by the difference between income
 (total photosynthesis) and losses (respiration and turnover) within the
@@
 \begin{equation}\label{eq:dbdt}
-\underbrace{\strut \frac{\textrm{d}b}{\textrm{d}t}}_\textrm{net biomass production}
+\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
   = \underbrace{\strut y}_\textrm{yield}
-    \big( \underbrace{\strut a_\textrm{l} \, p}_\textrm{photosynthesis} -
-     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
-    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
+    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
+     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
+    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
 \end{equation}
@@
 definitions). Photosynthesis is proportional to leaf area,
-$a_\textrm{l} = m_\textrm{l} / \phi$, where $\phi$ is leaf mass per area.
-The total mass of living tissues $m_\textrm{a}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.$
+$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
+The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_\textrm{s}+M_{\rm r}.$
 
@@
 \begin{equation} \label{eq:dhdt}
-\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
-\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}}_\textrm{marginal leaf deployment}
-\times \underbrace{\strut\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}}_\textrm{allocation to growth}
-\times \underbrace{\strut\frac{\textrm{d}b}{\textrm{d}t}}_\textrm{mass production}.
+\underbrace{\strut\frac{{\rm d}H}{{\rm d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{{\rm d}H}{{\rm d}A_{\rm l}}}_\textrm{architecture}
+\times \underbrace{\strut\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}}_\textrm{marginal leaf deployment}
+\times \underbrace{\strut\frac{{\rm d}M_{\rm a}}{{\rm d}B}}_\textrm{allocation to growth}
+\times \underbrace{\strut\frac{{\rm d}B}{{\rm d}t}}_\textrm{mass production}.
 \end{equation}
@@
 The first term on the right of eq \ref{eq:dhdt},
-$\textrm{d}h / \textrm{d}a_\textrm{l}$, is the growth in plant height
+${\rm d}H / {\rm d}A_{\rm l}$, is the growth in plant height
 per unit growth in total leaf area -- accounting for the architectural
@@
 
-The second term, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$,
+The second term, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$,
 accounts for the marginal cost of deploying an additional unit of leaf
 area, including construction of the leaf itself and various support
-structures. As such, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$
+structures. As such, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$
 can itself be expressed as a sum of construction costs per unit leaf
@@
 \begin{equation}\label{eq:daldmt}
-\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}
+\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}
 = \big(\phi
- + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
+ + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}\big)^{-1}.
 \end{equation}
@@
 The third term in eq \ref{eq:dhdt},
-$\textrm{d}m_\textrm{a} / \textrm{d}P$, gives the fraction of net
+${\rm d}M_{\rm a} / {\rm d}P$, gives the fraction of net
 biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
@@
 
-In a similar way, basal area ($a_\textrm{st}$) increment can be
+In a similar way, basal area ($A_{\rm st}$) increment can be
 expressed as the sum of increments in sapwood, bark \& heartwood areas
-($a_\textrm{ss}, a_\textrm{sb}, a_\textrm{sh}$ respectively):
-$\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}= \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}$.
+($A_\textrm{ss}, A_{\rm sb}, A_{\rm sh}$ respectively):
+$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm sb}}{{\rm d}t} + \frac{{\rm d}A_\textrm{ss}}{{\rm d}t} + \frac{{\rm d}A_{\rm sh}}{{\rm d}t}$.
 We now have an equation for basal area growth that conatins many of the same
@@
 \begin{equation}\label{eq:dast}
-\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
-\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
-\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}} \times \frac{\textrm{d}m_\textrm{a}}{\textrm{d}b} \times \frac{\textrm{d}b}{\textrm{d}t} +\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} .
+\frac{{\rm d}A_{\rm st}}{{\rm d}t}=
+\left(\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}} + \frac{{\rm d}A_{\rm sb}}{{\rm d}A_{\rm l}}\right) \times
+\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}} \times \frac{{\rm d}M_{\rm a}}{{\rm d}B} \times \frac{{\rm d}B}{{\rm d}t} +\frac{{\rm d}A_{\rm sh}}{{\rm d}t} .
 \end{equation}
@@
 Diameter growth is then given by the geometric relationship between stem
-diameter ($D$) and $a_\textrm{st}$:
+diameter ($D$) and $A_{\rm st}$:
 
 \begin{equation} \label{eq:dDdt}
-\frac{\textrm{d}D}{\textrm{d}t}= \sqrt{\frac{\pi}{a_\textrm{st}}} \frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}.
+\frac{{\rm d}D}{{\rm d}t}= \sqrt{\frac{\pi}{A_{\rm st}}} \frac{{\rm d}A_{\rm st}}{{\rm d}t}.
 \end{equation}
@@
 \ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
-the form $k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
+the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
 shown to describe patterns across diverse
@@
 growth by adjusting the amount of energy invested in growth, i.e.~the
-term $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ in eqs.
+term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqs.
 \ref{eq:dhdt} and \ref{eq:dast}.
@@
 \ref{eq:dDdt} predict a relationship between $\phi$ and growth that
-changes with plant size ( \ref{f-lma_growth_size}).
+changes with plant size ( \ref{f-lmA_growth_size}).
 Decreasing $\phi$ has two impacts on growth rate. First, lower
 $\phi$ increases marginal leaf deployment
-($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$ by economising on
+(${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on
 construction costs. Second, lower $\phi$ decreases net production
-($\textrm{d}P / \textrm{d}t$), due to increased leaf turnover. Whether
+(${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether
 lower $\phi$ increases growth thus depends on the relative magnitude
@@
 $\phi$ is no longer advantageous for growth (Fig.
-\ref{f-lma_growth_size}d).
+\ref{f-lmA_growth_size}d).
 
@@
 can thus lead to a growth advantage by increasing
-$\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ . At smaller sizes, there is no differentiation
+$\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ . At smaller sizes, there is no differentiation
 among species, because most species are focusing on growth. At larger sizes, individuals of some species are
@@
 \citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
-$\textrm{d}P / \textrm{d}t / m_\textrm{a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
+${\rm d}P / {\rm d}t / M_{\rm a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
 (see Appendix \ref{app:traits-RGR}).
@@
 %  $a = l + ss + sb +  r$,
-%  $m_{sh}$,
-%  $m_t = m_a + m_{sh}$,
+%  $M_{sh}$,
+%  $M_t = M_a + M_{sh}$,
 %  $st = stem total$
@@
   $b$   & kg  & biomass originating from parent plant\\
-  $m_i$ & kg  & mass retained on plant in tissue $i$\\
-  $a_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
+  $M_i$ & kg  & mass retained on plant in tissue $i$\\
+  $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
   $h$   & m  & height of plant\\
@@
   height &
-    $h = \alpha_1 \, a_\textrm{l}^{\beta_1}$ &
-    $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} =  \beta_1 \alpha_1 a_\textrm{l}^{\beta_1-1}$ &
-    $\frac{\textrm{d}h}{\textrm{d}t}  = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+    $h = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
+    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
+    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
   sapwoood area &
-    $a_\textrm{ss} = \theta \, a_\textrm{l}$ &
-    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} = \theta$ &
-    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t}  =\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+    $A_\textrm{ss} = \theta \, A_{\rm l}$ &
+    $\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} = \theta$ &
+    $\frac{{\rm d}A_\textrm{ss}}{{\rm d}t}  =\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
   bark area &
-    $a_\textrm{sb} = b \, \theta \, a_\textrm{l}$ &
-    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} = b \, \theta$ &
-    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} = \frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\  \\
+    $A_{\rm sb} = b \, \theta \, A_{\rm l}$ &
+    $\frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} = b \, \theta$ &
+    $\frac{{\rm d}A_{\rm sb}}{{\rm d}t} = \frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\  \\
   \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
   leaf mass &
-    $m_\textrm{l} = \phi \, a_\textrm{l} $ &
-    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}} = \phi$ &
-    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+    $M_{\rm l} = \phi \, A_{\rm l} $ &
+    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
+    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
   sapwood mass &
-    $m_\textrm{ss} = \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
-    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} = \theta\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
-    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+    $M_\textrm{ss} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
+    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}t}  = \frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
   bark mass &
-    $m_\textrm{sb} = b\, \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
-    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} = b \, \theta \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
-    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
+    $M_{\rm sb} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
+    $\frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $\frac{{\rm d}M_{\rm sb}}{{\rm d}t}  = \frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
   root mass &
-    $m_\textrm{r} = \alpha_3 \, a_\textrm{l}$ &
-    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}} = \alpha_3$  &
-    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\ \\
+    $M_{\rm r} = \alpha_3 \, A_{\rm l}$ &
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\ \\
 
@@
   heartwood area &
-    $a_\textrm{sh} = \int_0^t \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ &
+    $A_{\rm sh} = \int_0^t \frac{{\rm d}A_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
      &
-    $\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} \, a_\textrm{ss}$ \\
+    $\frac{{\rm d}A_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, A_\textrm{ss}$ \\
   heartwood mass &
-    $m_\textrm{sh} = \int_0^t \frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ &
+    $M_{\rm sh} = \int_0^t \frac{{\rm d}M_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
      &
-    $\frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} \, m_\textrm{ss}$ \\
+    $\frac{{\rm d}M_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, M_\textrm{ss}$ \\
   \hline \\
@@
   & $h_0$
-  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}$
-  & $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$
-  & $\frac{\textrm{d}b}{\textrm{d}t}$
-  & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
+  & $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$
+  & $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$
+  & $\frac{{\rm d}B}{{\rm d}t}$
+  & $\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}}$
   & $\frac{1}{k_s}$ & &  \\\hline
@@
 becomes a linear function of leaf area and net photosynthesis per leaf
-area ($A_\textrm{net} = y(A - r_\textrm{l})$), making relative growth
+area ($A_\textrm{net} = y(A - r_{\rm l})$), making relative growth
 rate a linear function of $\phi$:
@@
 \begin{equation}\label{eq:RGR}
-\underbrace{\strut\frac{\textrm{d}b}{\textrm{d}t}\frac{1}{m_\textrm{a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{a}}.
-\end{equation}
+\underbrace{\strut\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}. \end{equation}
 
@@
 variables that are routinely collected for large trees: namely plant
-height ($h$) and stem cross-sectional area ($a_\textrm{st}$) or
+height ($h$) and stem cross-sectional area ($A_{\rm st}$) or
 diameter $D$.
@@
 \begin{equation}\label{eq:ha}
-a_\textrm{l}=\alpha_1 \, h^{\beta_1}.
+A_{\rm l}=\alpha_1 \, h^{\beta_1}.
 \end{equation}
@@
 \begin{equation}\label{eq:crown2}
-\frac{m_\textrm{l}}{\phi}= \theta \, S(0,h).
+\frac{M_{\rm l}}{\phi}= \theta \, S(0,h).
 \end{equation}
@@
 \begin{equation}\label{eq:ms1}
-m_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, \textrm{d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}
+M_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, {\rm d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}
 
@@
 \begin{equation}\label{eq:ms2}
-m_\textrm{s}=\rho \, \eta_c \, \theta \, a_\textrm{l} \, h.
+M_\textrm{s}=\rho \, \eta_c \, \theta \, A_{\rm l} \, h.
 \end{equation}
@@
 \begin{equation}\label{eq:mb}
-m_\textrm{b}=b m_\textrm{s}.
+M_{\rm b}=b M_\textrm{s}.
 \end{equation}
@@
 \begin{equation}\label{eq:mr}
-m_\textrm{r}=\alpha_3 \, a_\textrm{l}.
+M_{\rm r}=\alpha_3 \, A_{\rm l}.
 \end{equation}
@@
 given height where 100\% of available energy is allocated to growth,
-i.e. $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}=1$. Note also
+i.e. $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=1$. Note also
 that by the assumptions of our model and above, traits do not influence
-$\frac{\textrm{d}h} {\textrm{d}a_\textrm{l}}$. Thus we have
+$\frac{{\rm d}H} {{\rm d}A_{\rm l}}$. Thus we have
 
 \begin{equation} \label{eq:G2}
-G = \frac{\textrm{d}h}{\textrm{d}t} = c_1   \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}  \frac{ \textrm{d}b} {\textrm{d}t} \right),
-\end{equation}
-
-where $c_1 = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}$. Eq.
+G = \frac{{\rm d}H}{{\rm d}t} = c_1   \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}  \frac{ {\rm d}B} {{\rm d}t} \right),
+\end{equation}
+
+where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eq.
 \ref{eq:G2} is a product of the form $Y(x) = W(x) \times Z(x)$. For,
@@
 \frac{\partial G} {\partial x} =
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}\right)}{\partial x}
- \, \frac{ \textrm{d}b} {\textrm{d}t}
-+ \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
-\, \frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x}.
+\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}{\partial x}
+ \, \frac{ {\rm d}B} {{\rm d}t}
++ \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
+\, \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}.
 \end{equation}
@@
 
-\begin{equation} \label{eq:da_dx}
-\frac{\partial \left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}\right)}
-{\partial x} = -\left(\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}\ \right)^2
-\, \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
+\begin{equation} \label{eq:dA_dx}
+\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}
+{\partial x} = -\left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\ \right)^2
+\, \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
 }{\partial x}.
@@
 
-Combing eqs. \ref{eq:dG}-\ref{eq:da_dx}, we have
+Combing eqs. \ref{eq:dG}-\ref{eq:dA_dx}, we have
 
@@
 \frac{\partial G} {\partial x} =
-\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
+\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
 \left(
-\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x}
-- \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
-\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
+- \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
+\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
 }{\partial x}
- \, \frac{ \textrm{d}b} {\textrm{d}t}
+ \, \frac{ {\rm d}B} {{\rm d}t}
 \right).
@@
 
-Since is $\frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}>0$ always, the sign
+Since is $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}>0$ always, the sign
 of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
@@
 \begin{equation}\label{eq:dg3}
-\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x}
-< \frac{\textrm{d}a_\textrm{l}} {\textrm{d}m_\textrm{a}}
-\,  \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
-}{\partial x} \, \frac{ \textrm{d}b} {\textrm{d}t},
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
+< \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
+\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
+}{\partial x} \, \frac{ {\rm d}B} {{\rm d}t},
 \end{equation}
@@
 \frac{
-\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x} }
-{\frac{ \textrm{d}b} {\textrm{d}t}}
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
+{\frac{ {\rm d}B} {{\rm d}t}}
 <
-\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
-}{\partial x}}{\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}}.
+\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
+}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
 \end{equation}
@@
 \frac{
-\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial x} }
-{\frac{ \textrm{d}b} {\textrm{d}t}}
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
+{\frac{ {\rm d}B} {{\rm d}t}}
 =
-\frac{ \frac{\partial \left(\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}\right)
-}{\partial x}}{\frac{\textrm{d}m_\textrm{a}} {\textrm{d}a_\textrm{l}}}.
+\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
+}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
 \end{equation}
@@
 If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand
-eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{\textrm{d}m_\textrm
-{t}} {\textrm{d}a_\textrm{l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
+eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm
+{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
 
 \begin{equation}\label{eq:dbdt2}
-\frac{\textrm{d}b}{\textrm{d}t} = c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}
+\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
 \end{equation}
@@
 where
-$c_2 = y ( a_\textrm{l} \, (A - r_\textrm{l}) + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{m_\textrm{i} \, hmax\textrm{i}})$
+$c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$
 is independent of $\phi$, and thus
@@
 \begin{equation}\label{eq:dbdt3}
-\frac{\partial \left( \frac{ \textrm{d}b} {\textrm{d}t}\right)}{\partial \phi}  =
-(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}.
-\end{equation}
-
-Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ \textrm{d}b} {\textrm{d}t}$ is independent
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
+(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
+\end{equation}
+
+Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent
 of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
-{d}P} {\textrm{d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
+{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
 
@@
 \begin{equation} \label{eq:G6}
-\frac{(B_4-1) a_\textrm{l} a_4\phi ^{-B_4}}{c_2 - a_\textrm{l} a_4 \phi ^{1-B_4}}
+\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
 < \frac{1}{\phi
- + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm
- {b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}}.
+ + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
+ {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
 \end{equation}
@@
 \centering
-\includegraphics{SI/lma_tradeoff.pdf}
+\includegraphics{SI/lmA_tradeoff.pdf}
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
@@
 \centering
-\includegraphics{SI/lma_effects_at_diameters.pdf}
+\includegraphics{SI/lmA_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
-and growth rate changes with plant size.} \label{f-lma_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lma_effects_at_diameters_relative.pdf}
+and growth rate changes with plant size.} \label{f-lmA_growth_size}}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics{SI/lmA_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
 and relative growth rate changes with plant size.}
-\label{f-lma_growth_size_relative}}
+\label{f-lmA_growth_size_relative}}
 \end{figure}

### 2016-07-28 ms/ms.tex
Standardise reference style to same as other papers

@@
 plants use to generate and invest surplus energy
- \citep{wright-2004, chave-2009, westoby-2002}.
+ \citep{Wright-2004, Chave-2009, Westoby-2002}.
 Although most plants have the same basic physiological function
@@
 mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
-diameter,  maximum  height (HMAX) and seed size  \citep{wright-2004,chave-2009}.
+diameter,  maximum  height (HMAX) and seed size  \citep{Wright-2004, Chave-2009}.
 Data for these traits now exists for up to 10\% of the world's 250000 plant species
-\citep{cornwell-2014}. Increasingly, researchers have been looking to traits
+\citep{Cornwell-2014}. Increasingly, researchers have been looking to traits
 as potential indicators of the growth strategy and ecology of a species.
@@
 demographic rates have met with mixed success
- \citep{wright-2010, poorter-2008}. In seedlings, the trait LMA
+ \citep{Wright-2010, Poorter-2008}. In seedlings, the trait LMA
 -- the central element of the leaf economics spectrum
- \citep{wright-2004} -- has been tightly related to relative growth
-rate in plant mass \citep{lambers-1992, wright_cross-2000}. LMA and
+ \citep{Wright-2004} -- has been tightly related to relative growth
+rate in plant mass \citep{Lambers-1992, Wright-2000}. LMA and
 its correlate leaf lifespan have also been linked to height growth rate for
-small seedlings and saplings \citep{reich-1992, poorter-2006}. These
+small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These
 early successes prompted researchers to search for similar relationships in
 large plants. However, the results showed that in trees LMA was not correlated
-with diameter growth rate  \citep{wright-2010, poorter-2008,
-herault-2011}. Meanwhile, other traits such as wood density showed
-strong relationships to growth in large plants \citep{wright-2010},
-but only weak relationships in seedlings  \citep{castro-diez-1998}. Thus
+with diameter growth rate  \citep{Wright-2010, Poorter-2008,
+Herault-2011}. Meanwhile, other traits such as wood density showed
+strong relationships to growth in large plants \citep{Wright-2010},
+but only weak relationships in seedlings  \citep{Castro-1998}. Thus
 far, most trait growth relationships have been quantified for one or few
@@
 link between traits and growth may be modified by plant size
- \citep{ruger-2012}. This finding challenges a wide-spread
+ \citep{Ruger-2012}. This finding challenges a wide-spread
 assumption regarding traits: that particular ends of the trait spectrum translate directly into
@@
 expectations on the expected signal. Generating such expectations is one of the
-primary role for theory \citep{kokko-2007}. Current empirical
+primary role for theory \citep{Kokko-2007}. Current empirical
 results suggest the effect of traits on growth changes with size and possibly
@@
 mass-based relative growth rate is linearly and negatively related to LMA
- \citep{lambers-1992, cornelissen-1996, wright_cross-2000}.
+ \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
  An extension of the model suggests a similar
-relationship may hold at larger sizes \citep{enquist-2007}. But as
+relationship may hold at larger sizes \citep{Enquist-2007}. But as
 described above, this prediction is not supported by empirical results.
@@
 A further problem with existing theory is that the effects of traits are all
-realised via influences on mass production \citep{enquist-2007},
+realised via influences on mass production \citep{Enquist-2007},
 whereas the effect of traits such as LMA, WD, and Hmax is via allocated
-to different tissues \citep{falster-2011}. There are in fact two problems with theory
+to different tissues \citep{Falster-2011}. There are in fact two problems with theory
 focussed solely on mass production. The first is that measuring mass
@@
 or height. For example, the CTFS monitors diameter growth in 49 large permanent
-plots across forests worldwide \citep{anderson-teixeira-2015};
+plots across forests worldwide \citep{Anderson-2015};
  forestry have likewise recorded growth in diameter across
-large tracts of the globe \citep{purves-2008}. The second problem is
+large tracts of the globe \citep{Purves-2008}. The second problem is
 that, as we outline further below, because mass production is only one
@@
 Ultimately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
-Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{sillett-2010, king-2011} or diameter growth \citep{herault-2011}.
-In contrast, basal-area growth continues to increase with size \citep{sillett-2010, stephenson-2014}, due to an increasing influence of stem turnover.
-All growth measures decrease sharply with size when expressed as relative growth rates \citep{rees-2010, iida-2014}.
-Shade tolerance tends to decrease with increasing size \citep{lusk-2006}, due to increasing costs of mainatining stems \citep{givnish-1988}
+Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
+In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover.
+All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
+Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of mainatining stems \citep{Givnish-1988}
 
@@
 study we re-analyse demographic and trait data from an earlier study
- \citep{wright-2010}, to show 1) that the link between traits and
+ \citep{Wright-2010}, to show 1) that the link between traits and
 growth is stronger than previously thought, 2) that the influence of three
@@
 impact of size on the nature of relationships varies among traits
- \citep[see also]{ruger-2012}. We then show how these results align with new
+ \citep[see also]{Ruger-2012}. We then show how these results align with new
 set of theoretical predictions. Analysis extends ...
@@
 responses to light environment and also determine shade tolerance, supporting
-diverse empirical results \citep{ruger-2012, poorter-2006}. By
+diverse empirical results \citep{Ruger-2012, Poorter-2006}. By
 disentangling the effects of plant size, light environment and traits on
@@
 To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in
-seedlings to LMA \citep{lambers-1992, wright_cross-2000}, to explicitly include influences of
+seedlings to LMA \citep{Lambers-1992, Wright-2000}, to explicitly include influences of
 size, light environment, and other prominent traits. This model
@@
 (total photosynthesis) and losses (respiration and turnover) within the
-plant \citep{makela-1997, Thornley-2000, falster-2011}:
+plant \citep{Makela-1997, Thornley-2000, Falster-2011}:
 
@@
 To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
-Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{falster-2011}:
+Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011}:
 
@@
 tall, while other species emphasise vertical
-extension \citep{poorter-2006}.
+extension \citep{Poorter-2006}.
 
@@
 To address the first of these challenges, we use a functional balance
-model \citep{yokozawa-1995,falster-2011} assuming
+model \citep{Yokozawa-1995, Falster-2011} assuming
 1) an allometric relation between total leaf area and plant height, and
@@
 growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction \citep{wright-2004}.
+results from superior leaf construction \citep{Wright-2004}.
 This is achieved by linking the turnover rate of leaf in eq.
@@
 shown to describe patterns across diverse
-ecosystems \citep{wright-2004} (Fig. \ref{fS-leaf}).
+ecosystems \citep{Wright-2004} (Fig. \ref{fS-leaf}).
 
@@
 lower $\rho$ decreases mechanical strength resulting in higher
-mortality \citep{chave-2009, wright-2010}. Under
+mortality \citep{Chave-2009, Wright-2010}. Under
 this scenario, decreasing $\rho$ will always provide a growth
@@
 of plant growth, recovering well-known
-empirical patterns \citep{sillett-2010, king-2011}. Height growth shows a hump-shaped pattern
+empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern
 with size, first increasing then decreasing. This pattern results from
@@
 and increasing reproductive allocation. In contrast, basal-area growth
-continues to increase with size \citep{sillett-2010, stephenson-2014},
+continues to increase with size \citep{Sillett-2010, Stephenson-2014},
 due to an increasing influence of stem
 turnover. Diameter growth shows a weakly hump shaped
-curve \citep{herault-2011}, tapering off slightly at larger
+curve \citep{Herault-2011}, tapering off slightly at larger
 sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
@@
 (Fig?). All growth measures decrease sharply with size when expressed as
-relative growth rates \citep{iida-2014}.
+relative growth rates \citep{Iida-2014}.
 
@@
 with low $rho$ increase growth more substantially with incraeses in
-light \citep{ruger-2012}. This is exactly the behaviour
+light \citep{Ruger-2012}. This is exactly the behaviour
 exhibited in our model (Fig. \ref{f-growth_light_dia}), at least for small
@@
 $rho$ actually decreases as size
-increases \citep{ruger-2012}. In contrast, growth rate tends
+increases \citep{Ruger-2012}. In contrast, growth rate tends
 to increase in size for the species with highest
-$\rho$ \citep{ruger-2012}.
+$\rho$ \citep{Ruger-2012}.
 
@@
 in small plants but not for plants \textgreater{} 0.01m
-dbh \citep{ruger-2012}.
+dbh \citep{Ruger-2012}.
 
@@
 Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
-\citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
+\citep{Lambers-1992, Cornelissen-1996, Wright-2000, Enquist-2007},
 ${\rm d}P / {\rm d}t / M_{\rm a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
@@
 
-@Osazuwa-Peters-2014 - indreases in WD through ontogeny.
+@Osazuwa-2014 - indreases in WD through ontogeny.
 
@@
 needed to offset these costs. As previously suggested
- \citep{givnish-1988}, shade-tolerance also decreases with
+ \citep{Givnish-1988}, shade-tolerance also decreases with
 height because as size increase, the total amount of energy needed to
@@
 finding matches well known empirical patterns, where both $\phi$ and
-$\rho$ have been linked to shade tolerance \citep{poorter-2006, lusk-2008,osunkoya-1996}.
+$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,osunkoya-1996}.
 
@@
 Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to $\phi$\citep{wright_cross-2000}, this
+relation to $\phi$\citep{Wright-2000}, this
 approximation does not directly link to other traits, or to the
@@
 
-Note, scaling relationship reversed from \citep{falster-2011}.
+Note, scaling relationship reversed from \citep{Falster-2011}.
 
@@
 
-We follow the model of \citep{yokozawa-1995} describing the
+We follow the model of \citep{Yokozawa-1995} describing the
 vertical distribution of leaf area within the crowns of individual
@@
 We also assume that each unit of sapwood area supports a fixed area of
-leaf (the pipe model, \citep{shinozaki-1964}), so that the total
+leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
 canopy area of a plant relates to basal sapwood area $S(0,h)$:
@@
 where
-$\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}$ \citep{yokozawa-1995}.
+$\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}$ \citep{Yokozawa-1995}.
 Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
@@
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
-Data from \citep{wright-2004} for 678 species from 51 sites, each
+Data from \citep{Wright-2004} for 678 species from 51 sites, each
 point giving a species-average. Lines show standardised major axis lines

### 2016-07-29 ms/ms-common-defs.tex
Separate doc for supp mat, use ms style from plant paper

--- 2016-07-28 ms/ms.tex
@@
-\documentclass[12pt, a4paper]{article}
-\RequirePackage[hmargin=2cm,vmargin=2cm]{geometry}
+\usepackage[rgb,dvipsnames]{xcolor}
+\definecolor{grey}{rgb}{0.5, 0.5, 0.5}
 
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{authblk} % allows for better author affiliations
+\newcommand{\smurl}[1]{\url{#1}}
+\newcommand{\ud}{\ensuremath{\rm{d}}}
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
 
-%appendix
-\usepackage[title,titletoc,toc]{appendix}
-\usepackage{setspace}  % for double spacing
+\title{Untangling the link between traits, size and growth rate in plants}
+\runninghead{I\lowercase{nfluence of traits on growth}}
 
-% Fancy HEADER
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\setlength{\headheight}{15pt}
-\pagenumbering{arabic}
-\lhead{\itshape{\nouppercase{\leftmark}}}
-\chead{}
-\rhead{\thepage}
-%\lfoot{v }
-\cfoot{}
-%\rfoot{\thepage}
+\author{Daniel S. Falster\textasteriskcentered \& Richard G. FitzJohn}
 
+\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia
+\\
+\textasteriskcentered Correspondence author. E-mail: \texttt{adaptive.plant@gmail.com}
+}
+
+\usepackage{natbib}
+\bibliographystyle{ecol_let}
+\setcitestyle{authoryear,open={(},close={)}}
 % allows for numbered referencing, as required by ecol letters
@@
 \booltrue{MyRefNumbers} % comment to remove numbers in reference list
-
-\usepackage{natbib}
-\bibliographystyle{ecol_let}
-\setcitestyle{authoryear,open={(},close={)}}
 
@@
 \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
-
-\usepackage[rgb,dvipsnames]{xcolor}
-\definecolor{grey}{rgb}{0.5, 0.5, 0.5}
-
-\usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\hypersetup{breaklinks=true,
-            pdfauthor={},
-            pdftitle={Untangling the link between traits, size and growth rate in plants},
-            colorlinks=true,
-            citecolor=black,
-            urlcolor=grey,
-            linkcolor=grey}
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-
-\title{\LARGE Untangling the link between traits, size and growth rate in plants}
-
-\author[1]{Daniel S. Falster}
-\author[1]{Richard G. FitzJohn}
-
-\affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
-\renewcommand\Authands{ and }
-\date{\vspace{-3em}}
-
-\begin{document}
-
-\maketitle
-\thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that
-
-\section{Ideas}\label{abstract}
-
-Table near intro summarising empricial patterns to be explained by integrated model?
-
-Section 3.3 of King (2005) shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
-
-King 1999 also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
-Basically, the much greater path length from roots to leaves in trees vs.
-seedlings results in a much greater increment in support cost per increment in crown area for trees.
-This explanation is treated by the growth model of King (1999), which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
-
-\section{Abstract}\label{abstract}
-
-\section{Introduction}\label{introduction}
-
-Functional traits capture core differences in the strategies
-plants use to generate and invest surplus energy
- \citep{Wright-2004, Chave-2009, Westoby-2002}.
-Although most plants have the same basic physiological function
-and key resource requirements (carbon, nitrogen and water), species differ considerably in
-rates at which resources are acquired and invested into different tissues.
-Much of this functional variation relates to underlying
-differences in some prominent traits, including leaf
-mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
-diameter,  maximum  height (HMAX) and seed size  \citep{Wright-2004, Chave-2009}.
-Data for these traits now exists for up to 10\% of the world's 250000 plant species
-\citep{Cornwell-2014}. Increasingly, researchers have been looking to traits
-as potential indicators of the growth strategy and ecology of a species.
-
-While the influence of plant traits on elements of plant physiological
-function is increasingly understood, attempts at using traits to understand
-demographic rates have met with mixed success
- \citep{Wright-2010, Poorter-2008}. In seedlings, the trait LMA
--- the central element of the leaf economics spectrum
- \citep{Wright-2004} -- has been tightly related to relative growth
-rate in plant mass \citep{Lambers-1992, Wright-2000}. LMA and
-its correlate leaf lifespan have also been linked to height growth rate for
-small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These
-early successes prompted researchers to search for similar relationships in
-large plants. However, the results showed that in trees LMA was not correlated
-with diameter growth rate  \citep{Wright-2010, Poorter-2008,
-Herault-2011}. Meanwhile, other traits such as wood density showed
-strong relationships to growth in large plants \citep{Wright-2010},
-but only weak relationships in seedlings  \citep{Castro-1998}. Thus
-far, most trait growth relationships have been quantified for one or few
-traits and at a single life- stage. Yet the picture is emerging is that the
-link between traits and growth may be modified by plant size
- \citep{Ruger-2012}. This finding challenges a wide-spread
-assumption regarding traits: that particular ends of the trait spectrum translate directly into
-faster or slower growth (refs XX).
-
-The challenge of interpreting diverse empirical results seeking to link
-traits to growth rate is made all the harder because we currently lack clear
-expectations on the expected signal. Generating such expectations is one of the
-primary role for theory \citep{Kokko-2007}. Current empirical
-results suggest the effect of traits on growth changes with size and possibly
-also light environment, yet current theory says little about how such
-relationships might come about. A widely-used model shows that for seedlings,
-mass-based relative growth rate is linearly and negatively related to LMA
- \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
- An extension of the model suggests a similar
-relationship may hold at larger sizes \citep{Enquist-2007}. But as
-described above, this prediction is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth
-are largely absent.
-
-A further problem with existing theory is that the effects of traits are all
-realised via influences on mass production \citep{Enquist-2007},
-whereas the effect of traits such as LMA, WD, and Hmax is via allocated
-to different tissues \citep{Falster-2011}. There are in fact two problems with theory
-focussed solely on mass production. The first is that measuring mass
-production is only practical for small plants that can be easily harvested.
-For larger plants, growth is more often measured as increment in either diameter
-or height. For example, the CTFS monitors diameter growth in 49 large permanent
-plots across forests worldwide \citep{Anderson-2015};
- forestry have likewise recorded growth in diameter across
-large tracts of the globe \citep{Purves-2008}. The second problem is
-that, as we outline further below, because mass production is only one
-component of plant growth; theory focussing solely on mass production will
- always be limited.
-
-Effect of traits on shade tolerance also porrly udnerstood. Empirical -- shade
-
-(reich 1992, Niinements 2010, Gower-1993 - both show higher LAI in high LMA / long LL stands)
-
-Ultimately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
-Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
-In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover.
-All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of mainatining stems \citep{Givnish-1988}
-
-In this paper we provide a mechanistic framework for describing the potential
-effect of traits on growth throughout a plant's life-cycle, linking to
-commonly used metrics such as height and diameter growth. To help motivate our
-study we re-analyse demographic and trait data from an earlier study
- \citep{Wright-2010}, to show 1) that the link between traits and
-growth is stronger than previously thought, 2) that the influence of three
-prominent traits on growth rate is moderated by plant size, and 3) that the
-impact of size on the nature of relationships varies among traits
- \citep[see also]{Ruger-2012}. We then show how these results align with new
-set of theoretical predictions. Analysis extends ...
-Moreover, we show how traits moderate plant
-responses to light environment and also determine shade tolerance, supporting
-diverse empirical results \citep{Ruger-2012, Poorter-2006}. By
-disentangling the effects of plant size, light environment and traits on
-growth rates, we hope to provide a solid theoretical foundation for trait
-ecology and thus provide a platform for understanding growth across diverse
-species around the world.
-
-\section{Materials and Methods}\label{materials-and-methods}
-
-\subsection{Growth data}\label{growth-data}
-
-\subsection{Conceptual framework}\label{conceptual-framework}
-
-To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in
-seedlings to LMA \citep{Lambers-1992, Wright-2000}, to explicitly include influences of
-size, light environment, and other prominent traits. This model
-
-Builds on 2011 paper - highlight what's new from there.
-
-First derive ... generally true.
-
-Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
-therefore hold for any model of plant growth. To make explicit
-predictions, however, requires two additional elements: 1) A specific
-model describing how the various mass and area terms
-($M_{\rm l}, M_\textrm{s}, M_{\rm b}, M_{\rm r}, A_{\rm l}, A_\textrm{s}, A_{\rm b}$)
-vary relative to one another; and 2) For any trade-offs in plant
-function to be specified; this is where traits have an influence.  These two components are outlined in following sections.
-
-\subsubsection{Standard model for mass production}
-
-We begin with a standard model for the amount of biomass available for
-growth, ${\rm d}B / {\rm d}t$, given by the difference between income
-(total photosynthesis) and losses (respiration and turnover) within the
-plant \citep{Makela-1997, Thornley-2000, Falster-2011}:
-
-\begin{equation}\label{eq:dbdt}
-\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
-  = \underbrace{\strut y}_\textrm{yield}
-    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
-     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
-    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
-\end{equation}
-
-Here, $m,r$, and $k$ refer to the mass, respiration rate, and
-turnover rate of different tissues, denoted by subscripts $l$=leaves,
-$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
-rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
-assimilated carbon fixed in biomass (the remaining fraction being lost
-as growth respiration) (see Table \ref{tab:params} for units and
-definitions). Photosynthesis is proportional to leaf area,
-$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
-The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_\textrm{s}+M_{\rm r}.$
-
-The only potentially controversial element of in eq.
-\ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants, because the .....
-However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegtative part of the plant (refs XX) \citep{ Thornley-2000}.
-
-\subsubsection{Decomposing height, basal area and diameter growth}
-
-To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
-Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011}:
-
-\begin{equation} \label{eq:dhdt}
-\underbrace{\strut\frac{{\rm d}H}{{\rm d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{{\rm d}H}{{\rm d}A_{\rm l}}}_\textrm{architecture}
-\times \underbrace{\strut\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}}_\textrm{marginal leaf deployment}
-\times \underbrace{\strut\frac{{\rm d}M_{\rm a}}{{\rm d}B}}_\textrm{allocation to growth}
-\times \underbrace{\strut\frac{{\rm d}B}{{\rm d}t}}_\textrm{mass production}.
-\end{equation}
-
-The first term on the right of eq \ref{eq:dhdt},
-${\rm d}H / {\rm d}A_{\rm l}$, is the growth in plant height
-per unit growth in total leaf area -- accounting for the architectural
-strategy of the plant. Some species tend to leaf out more than grow
-tall, while other species emphasise vertical
-extension \citep{Poorter-2006}.
-
-The second term, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$,
-accounts for the marginal cost of deploying an additional unit of leaf
-area, including construction of the leaf itself and various support
-structures. As such, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$
-can itself be expressed as a sum of construction costs per unit leaf
-area:
-
-\begin{equation}\label{eq:daldmt}
-\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}
-= \big(\phi
- + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}\big)^{-1}.
-\end{equation}
-
-The third term in eq \ref{eq:dhdt},
-${\rm d}M_{\rm a} / {\rm d}P$, gives the fraction of net
-biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
-rather than reproduction or storage.
-
-In a similar way, basal area ($A_{\rm st}$) increment can be
-expressed as the sum of increments in sapwood, bark \& heartwood areas
-($A_\textrm{ss}, A_{\rm sb}, A_{\rm sh}$ respectively):
-$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm sb}}{{\rm d}t} + \frac{{\rm d}A_\textrm{ss}}{{\rm d}t} + \frac{{\rm d}A_{\rm sh}}{{\rm d}t}$.
-We now have an equation for basal area growth that conatins many of the same
-elements as eq. \ref{eq:dhdt}:
-
-\begin{equation}\label{eq:dast}
-\frac{{\rm d}A_{\rm st}}{{\rm d}t}=
-\left(\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}} + \frac{{\rm d}A_{\rm sb}}{{\rm d}A_{\rm l}}\right) \times
-\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}} \times \frac{{\rm d}M_{\rm a}}{{\rm d}B} \times \frac{{\rm d}B}{{\rm d}t} +\frac{{\rm d}A_{\rm sh}}{{\rm d}t} .
-\end{equation}
-
-Diameter growth is then given by the geometric relationship between stem
-diameter ($D$) and $A_{\rm st}$:
-
-\begin{equation} \label{eq:dDdt}
-\frac{{\rm d}D}{{\rm d}t}= \sqrt{\frac{\pi}{A_{\rm st}}} \frac{{\rm d}A_{\rm st}}{{\rm d}t}.
-\end{equation}
-
-\subsubsection{A functional-balance model for allocation}
-
-To address the first of these challenges, we use a functional balance
-model \citep{Yokozawa-1995, Falster-2011} assuming
-1) an allometric relation between total leaf area and plant height, and
-2) a constant ratio between total leaf area and each of sapwood
-cross-sectional area, bark cross-sectional area, and total root mass.
-These assumptions, which are well-supported by data (Fig.
-\ref{f-assumptions}), lead immediately to an allometric model for plant
-size (Table \ref{tab:allometry}, see Supplementary material for
-derivations). Substituting from Table \ref{tab:allometry} into eqs.
-\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
-for plant growth in relation to size.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  four key assumptions.
-\item
-  simple linear functions non-critical for results of this paper,
-\item
-  explore using an example model, but any model with basic size effects
-  will display same behaviour
-
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    details about our model ()
-  \item
-    show argument hold intuitively
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{Embeeding trait-related trade-offs}
-
-use table with arrows to
-
-By definition, the main effect of increasing $\phi$ is to decrease the
-marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
-growth rate. However, to fully account for the influence of $\phi$ on
-growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction \citep{Wright-2004}.
-This is achieved by linking the turnover rate of leaf in eq.
-\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
-the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
-shown to describe patterns across diverse
-ecosystems \citep{Wright-2004} (Fig. \ref{fS-leaf}).
-
-As for leaf, cheaper stem construction (lower $\rho$) decreases the
-cost of deploying a unit of leaf area, and may thereby increase growth
-rates. However, the cost of cheap stem construction must also be
-accounted for. However, the direct physiological trade-offs of lower
-$\rho$ are less understood than for $\phi$. One possibility is that
-lower $\rho$ decreases mechanical strength resulting in higher
-mortality \citep{Chave-2009, Wright-2010}. Under
-this scenario, decreasing $\rho$ will always provide a growth
-advantage, because the costs of low $\rho$ are not realised within the
-terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
-\ref{tab:trade-offs}). In addition, lower $\rho$ may increase turnover
-of sapwood. Mirroring our assumptions for leaves, we let sapwood
-turnover decrease with $\rho$,
-i.e.$k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
-
-Unlike leaf and stem traits, height at maturation ($h_m$) moderates
-growth by adjusting the amount of energy invested in growth, i.e.~the
-term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqs.
-\ref{eq:dhdt} and \ref{eq:dast}.
-
-\subsection{Analysis}\label{analysis}
-
-\newpage
-\section{Results and Discussion}\label{results}
-
-\subsection{Changes in growth rate with size}
-
-This new growth model captures the intrinsically size-dependent nature
-of plant growth, recovering well-known
-empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern
-with size, first increasing then decreasing. This pattern results from
-systematic changes in the four components of eq. \ref{eq:dhdt} with size
-(Fig. \ref{f-conceptual}), including a strong decline in the fraction of plant
-that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
-and increasing reproductive allocation. In contrast, basal-area growth
-continues to increase with size \citep{Sillett-2010, Stephenson-2014},
-due to an increasing influence of stem
-turnover. Diameter growth shows a weakly hump shaped
-curve \citep{Herault-2011}, tapering off slightly at larger
-sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
-mainly because of increased reproductive allocation in older trees
-(Fig?). All growth measures decrease sharply with size when expressed as
-relative growth rates \citep{Iida-2014}.
-
-\subsection{How traits influence growth rate}
-
-\subsubsection{LMA}
-
-Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
-\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
-changes with plant size ( \ref{f-lmA_growth_size}).
-Decreasing $\phi$ has two impacts on growth rate. First, lower
-$\phi$ increases marginal leaf deployment
-(${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on
-construction costs. Second, lower $\phi$ decreases net production
-(${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether
-lower $\phi$ increases growth thus depends on the relative magnitude
-of these two effects. When plants are small the effect on leaf
-deployment rate is bigger and so decreasing $\phi$ increases growth
-rate. When plants are large, the influence of $\phi$ on leaf
-deployment rate is diminished, because the costs of building other
-supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
-\ref{f-mass_fraction}). The net result is that at larger sizes, low
-$\phi$ is no longer advantageous for growth (Fig.
-\ref{f-lmA_growth_size}d).
-
-\subsubsection{Wood density}
-
-Under this scenario, decreasing $\rho$ provides an advantage in height
-growth only when the benefits of low $\rho$ outweigh the costs. This
-effect is strongest at intermediate sizes. The decline in height growth
-at low $\rho$ with size is more marked than for either stem basal-area
-and diameter growth, because high sapwood turnover also contributes to
-this growth measure. Thus diameter growth continues, even as net
-production approaches zero.
-
-\subsubsection{hmat}
-
- Greater height at maturation ($h_m$)
-can thus lead to a growth advantage by increasing
-$\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ . At smaller sizes, there is no differentiation
-among species, because most species are focusing on growth. At larger sizes, individuals of some species are
-allocating a majority of their surplus energy to growth, leaving a
-strong signal in potential growth rate.
-
-\subsubsection{Responsiveness to light}
-
-Two other important predictions arise directly from linking traits to
-growth rate. First, traits moderate the
-responsiveness of growth to changes in light environment. This response
-arises simply because species with higher potential growth rate have
-greater potential plasticity. Recent analyses have shown that species
-with low $rho$ increase growth more substantially with incraeses in
-light \citep{Ruger-2012}. This is exactly the behaviour
-exhibited in our model (Fig. \ref{f-growth_light_dia}), at least for small
-to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
-of low $\phi$ are lost (Fig. \ref{f-growth_light_dia}). This results also
-agrees with empirical results showing that growth for species with low
-$rho$ actually decreases as size
-increases \citep{Ruger-2012}. In contrast, growth rate tends
-to increase in size for the species with highest
-$\rho$ \citep{Ruger-2012}.
-
-Variation in $\phi$ also moderates the response of growth to changes
-in light, with species having the lowest $rho$ being most responsive.
-However, unlike for $\rho$, the effect appears only for the smallest
-size classes. Accordingly, empirical analyses have detected such a link
-in small plants but not for plants \textgreater{} 0.01m
-dbh \citep{Ruger-2012}.
-
-\subsection{Growth is more than just photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}
-
-Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
-\citep{Lambers-1992, Cornelissen-1996, Wright-2000, Enquist-2007},
-${\rm d}P / {\rm d}t / M_{\rm a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
-(see Appendix \ref{app:traits-RGR}).
-
-Recently realisation that allocation and turnover major areas of
-uncertainty
-
-Model reconciles idea that traits can have negative impact on carbon
-budget, but still have positive influence on growth rates
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Enquist 2007: sla and wd increase mass-production, where as in our
-  theory effects on mass production are negative. (also contradicts
-  earlier model from Enquist 1999, where argued wood density had no
-  effect on mass production)
-\end{itemize}
-
-Assumes low lma and wd used to reduce costs of building leaf and stem,
-ie. total leaf area and stem corss section remains same. In this way
-differs to assumptions of recent work by anten\_role\_2010,
-larjavaara\_rethinking\_2010.
-
-Model is realisation of several ideas that have been around for some
-time.
-
-Plasticity in trait values predicted from model:
-
-@Osazuwa-2014 - indreases in WD through ontogeny.
-
-\subsection{Shade tolerance}\label{shade-tolerance}
-
-Second, species with low $\phi$ and low $\rho$ are predicted to be
-less shade tolerant (Fig. \ref{f-wplcp}). At low $\phi$ ($\rho$),
-leaf (sapwood) turnover is higher and thus a greater light income is
-needed to offset these costs. As previously suggested
- \citep{Givnish-1988}, shade-tolerance also decreases with
-height because as size increase, the total amount of energy needed to
-offset respiratory and turnover costs in the stem also increases. These
-finding matches well known empirical patterns, where both $\phi$ and
-$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,osunkoya-1996}.
-
-Results fo poort \& binger for surival strongest for LL.
-
-Also baltzer.
-
-\subsection{Implications for trait-based approaches}\label{implications-for-trait-based-approaches}
-
-How do comparative people interpret sla now - beyond fast slow.
-
-Think about traits as defining potential trajectory, rather than making
-species fast or slow growing per se.
-
-Selection on traits in different parts of lifecycle
-
-Trait plasticity - can explain, no clear advantage of low lma at large
-sizes
-
-Framework for generalising about effect of traits
-
-Towards a global model of growth:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  mechanistic model of height and dbh growth which accounts for effects
-  of size, light environment and traits
-
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    need to account for size long recognized, but not straightforward
-  \item
-    look at growth in wide range of measures: mass, height, dbh
-  \end{itemize}
-\item
-  Compared to previous work, shows effects of traits more about
-  allocation than mass production
-\end{itemize}
-
-Key patterns explained
-
-Data challenges
-
-Reproductive allocation
-
-Detecting trait signals in data.
-
-\begin{itemize}
-\item  Need more than simply look for correlation with mean growth rate
-\item
-  understand influences of size on trait of interest
-\item
-  traits define potential growth rate --\textgreater{} how to extract
-  this
-\item
-  ideally start to use mechanistic models
-\end{itemize}
-
-\section{Acknowledgments}
-
-Figure 2 uses image by Tracey Saxby, IAN Image Library http://ian.umces.edu/imagelibrary.
-
-\newpage
-
-\bibliography{references}
-
-\newpage
-
-\section{Tables}\label{tables}
-
-\begin{table}[ht]
- \caption{Key variables
-%  $a = l + ss + sb +  r$,
-%  $M_{sh}$,
-%  $M_t = M_a + M_{sh}$,
-%  $st = stem total$
-}
-\centering
-
-{\footnotesize  % smaller font, double space
- \begin{doublespace}
-  \begin{tabular}{p{2cm}p{2cm}p{7cm}}
-  \hline
-  Symbol & Unit & Description \\
-  \hline
-
-  $b$   & kg  & biomass originating from parent plant\\
-  $M_i$ & kg  & mass retained on plant in tissue $i$\\
-  $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
-  $h$   & m  & height of plant\\
-  $D$   & m  & diameter of plant stem\\
-
-  $y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
-  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthesic rate per unit area \\
-  $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
-  $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
-  $A$ & kg yr$^{-1}$ & total plant photosynthesis \\
-  $R$ & kg yr$^{-1}$ & total plant respiration \\
-  $T$ & kg yr$^{-1}$ & total plant turnover \\
-  \hline
-  \end{tabular}
-\end{doublespace}}
-\label{tab:definitions}
-\end{table}
-
-\newpage
-
-\begin{table}[ht]
-\caption{Equations for an allometric growth model, based on functional balance assumption in {\textbf a}. }
-\centering
-
-{\footnotesize  % smaller font, double space
- \begin{doublespace}
-  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
-  Variable & Function & Marginal cost & Growth rate\\ \hline
-  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
-  height &
-    $h = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
-    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  sapwoood area &
-    $A_\textrm{ss} = \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_\textrm{ss}}{{\rm d}t}  =\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  bark area &
-    $A_{\rm sb} = b \, \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm sb}}{{\rm d}t} = \frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\  \\
-  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
-  leaf mass &
-    $M_{\rm l} = \phi \, A_{\rm l} $ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  sapwood mass &
-    $M_\textrm{ss} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}t}  = \frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  bark mass &
-    $M_{\rm sb} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm sb}}{{\rm d}t}  = \frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  root mass &
-    $M_{\rm r} = \alpha_3 \, A_{\rm l}$ &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\ \\
-
-  \multicolumn{4}{l}{\textbf{c) Integrated quantities}} \\
-  heartwood area &
-    $A_{\rm sh} = \int_0^t \frac{{\rm d}A_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}A_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, A_\textrm{ss}$ \\
-  heartwood mass &
-    $M_{\rm sh} = \int_0^t \frac{{\rm d}M_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}M_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, M_\textrm{ss}$ \\
-  \hline \\
-\end{tabular}
-NB -  theta is sapwood area per leaf area. In tree 2 has been opposite, but that makes notation more clumsy.
-\end{doublespace}}
-\label{tab:allometry}
-\end{table}
-
-\newpage
-
-\begin{table}[ht]
-\caption{Effect of traits and related trade-offs on demography}
-\centering
-{\footnotesize  % smaller font, double space
-  \begin{doublespace}
-  \begin{tabular}[c]{l|c|ccccc|c|c}
-  \multicolumn{9}{c}{}\\ \hline
-  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
-  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
-  & $h_0$
-  & $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$
-  & $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$
-  & $\frac{{\rm d}B}{{\rm d}t}$
-  & $\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}}$
-  & $\frac{1}{k_s}$ & &  \\\hline
-  \textbf{This study}&&&&&&&&\\
-  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
-  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
-  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
-  &&&&&&&&\\\hline
-  \textbf{Possible extensions}&&&&&&&&\\
-  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
-  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
-  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
-  \end{tabular}
-  \end{doublespace}
-}
-\label{tab:trade-offs}
-\end{table}
-
-\newpage
-
-\section{Figures}\label{figures}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/figure_conceptual.pdf}
-\caption{\textbf{Conceptual framework for understanding size-dependent
-changes in growth rate.} Hump-shaped relationship between height growth rate and
-size. \label{f-conceptual}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/allometry.pdf}
-\caption{\textbf{Key assumptions of a functional balance and trait
-trade-offs model, evaluated using global dataset.} We used the biomass and
-allometry database to evaluate model assumptions about \textbf{a,}
-scaling of leaf area with plant height, \textbf{b} Scaling of sapwood
-area with leaf area, and \textbf{c} scaling of root mass with leaf area.
-Each dot is a single plant. Lines show standardised major axis lines
-fitted to data from each site, with intensity of shading adjusted
-according to strength of the relationship. Colours indicate vegetation
-type. Dashed black lines show values expected under functional-balance
-assumption (see Supplementary text for details). \label{f-assumptions}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/growth_light_dia.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (diameter growth).} Panels show predicted relationship between
-specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{f-growth_light_dia}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/growth_light_mass.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (mass growth).} Panels show predicted relationship between
-specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{f-growth_light_mass}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/max_leaf_above.pdf}
-\caption{\textbf{The effect of traits on growth changes with size and
-light environment.} Traits moderate the responsiveness of growth to changes in light
-environmentThis figure needs to show more of key result with respect to to size.
-Possibly separate figure. \label{f-shifts}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/max_leaf_above.pdf}
-\caption{\textbf{Low construction cost leads to shade intolerance,
-because of costs of high turnover.} Panels show effect of traits on
-maximum amount of shading that can be endured before net production (eq.
-\ref{eq:dbdt}) reaches zero. Lines indicate relationship for plants of a
-given height. \label{f-wplcp}}
-\end{figure}
-
-\clearpage
-
-\begin{appendices}
-\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-
-\section{Relationship between traits and relative growth rate} \label{app:traits-RGR}
-
-We start with eq. \ref{eq:dbdt} for mass-based growth. For seedlings, which are
-young and mostly leaf, it is reasonable to ignore all turnover terms as
-well as the respiration terms for non-leaf tissues. Net production then
-becomes a linear function of leaf area and net photosynthesis per leaf
-area ($A_\textrm{net} = y(A - r_{\rm l})$), making relative growth
-rate a linear function of $\phi$:
-
-\begin{equation}\label{eq:RGR}
-\underbrace{\strut\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}. \end{equation}
-
-Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to $\phi$\citep{Wright-2000}, this
-approximation does not directly link to other traits, or to the
-variables that are routinely collected for large trees: namely plant
-height ($h$) and stem cross-sectional area ($A_{\rm st}$) or
-diameter $D$.
-
-\section{Derivation of a simple allometric model of plant
-function}\label{app:func_balance}
-
-Here we describe an allometric model linking the various size dimensions
-of a plant required by most ecologically realistic vegetation models
-(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
-roots) to a plant height. Table \ref{tab:allometry} provides a summary
-of the derived model, while Table \ref{tab:params} provdies estimates on
-key parameters.
-
-\subsection{Leaf area}\label{leaf-area}
-
-Based on empirically observed allometries (see main text), we assume an
-allometric log-log scaling relationship between the accumulated leaf
-area of a plant and its height:
-
-\begin{equation}\label{eq:ha}
-A_{\rm l}=\alpha_1 \, h^{\beta_1}.
-\end{equation}
-
-Note, scaling relationship reversed from \citep{Falster-2011}.
-
-\subsection{Mass of sapwood}\label{mass-of-sapwood}
-
-We follow the model of \citep{Yokozawa-1995} describing the
-vertical distribution of leaf area within the crowns of individual
-plants. This model can account for a variety of canopy profiles through
-a single parameter $\eta$. Setting $\eta=1$ results in a conical
-canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
-, give a top-weighted canopy profile similar to those seen among
-angiosperms. Let $S(z,h)$ be the sapwood area at height $z$ for a
-plant with top height $h$. Following Yokozawa and Hara (1995) we
-assume a relationship between $S(z,h)$ and height such that
-
-\begin{equation}\label{eq:crown1}
-\frac{S(z,h)}{S(0,h)}= \big(1-\big(\frac{z}{h}\big)^\eta\big)^2.
-\end{equation}
-
-We also assume that each unit of sapwood area supports a fixed area of
-leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
-canopy area of a plant relates to basal sapwood area $S(0,h)$:
-
-\begin{equation}\label{eq:crown2}
-\frac{M_{\rm l}}{\phi}= \theta \, S(0,h).
-\end{equation}
-
-Integrating $S(z,h)$ gives a solution for the total mass of sapwood in
-the plant:
-
-\begin{equation}\label{eq:ms1}
-M_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, {\rm d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}
-
-where
-$\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}$ \citep{Yokozawa-1995}.
-Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
-expression for sapwood mass as a function leaf area and height:
-
-\begin{equation}\label{eq:ms2}
-M_\textrm{s}=\rho \, \eta_c \, \theta \, A_{\rm l} \, h.
-\end{equation}
-
-\subsection{Bark mass}\label{bark-mass}
-
-Bark and phloem tissue are modelled using an analogue of the pipe model,
-leading to a similar equation as that for sapwood mass (eq.
-\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
-assumed to be a constant fraction $b$ of sapwood area per unit leaf
-area such that
-
-\begin{equation}\label{eq:mb}
-M_{\rm b}=b M_\textrm{s}.
-\end{equation}
-
-\subsection{Root mass}\label{root-mass}
-
-Also consistent with pipe-model assumption, we assume a fixed ratio of
-root mass per unit leaf area
-
-\begin{equation}\label{eq:mr}
-M_{\rm r}=\alpha_3 \, A_{\rm l}.
-\end{equation}
-
-Even though nitrogen and water uptake are not modelled explicitly,
-imposing a fixed ratio of root mass to leaf area ensures that
-approximate costs of root production are included in calculations of
-carbon budget.
-
-\section{Effect of traits on height growth} \label{app:traits_max}
-
-We want to show how two traits $\phi$ and $\rho$ influence growth rate,
-by looking at the effect of
-traits on the components of growth rate (eq. \ref{eq:dhdt}. To make
-the analysis more tractable, we will assume we are dealing with a plant of
-given height where 100\% of available energy is allocated to growth,
-i.e. $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=1$. Note also
-that by the assumptions of our model and above, traits do not influence
-$\frac{{\rm d}H} {{\rm d}A_{\rm l}}$. Thus we have
-
-\begin{equation} \label{eq:G2}
-G = \frac{{\rm d}H}{{\rm d}t} = c_1   \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}  \frac{ {\rm d}B} {{\rm d}t} \right),
-\end{equation}
-
-where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eq.
-\ref{eq:G2} is a product of the form $Y(x) = W(x) \times Z(x)$. For,
-equations of this type, the derivative $\partial{Y}/\partial{x}$ is
-given by $\partial{W}/\partial{x} \, Z + W \, \partial{Z}/\partial{x}$. Thus
-the derivative of $G$ with respect to some trait $x$  is given by
-
-\begin{equation} \label{eq:dG}
-\frac{\partial G} {\partial x} =
-\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}{\partial x}
- \, \frac{ {\rm d}B} {{\rm d}t}
-+ \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\, \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}.
-\end{equation}
-
-If this derivative is positive (negative), there will be a positive (negative) relationship
-between traits and growth at the height where it is evaluated.
-
-We further simplify by expanding the derivative of eq. \ref{eq:daldmt}:
-
-\begin{equation} \label{eq:dA_dx}
-\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}
-{\partial x} = -\left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\ \right)^2
-\, \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}.
-\end{equation}
-
-Combing eqs. \ref{eq:dG}-\ref{eq:dA_dx}, we have
-
-\begin{equation} \label{eq:dG2}
-\frac{\partial G} {\partial x} =
-\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\left(
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
-- \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}
- \, \frac{ {\rm d}B} {{\rm d}t}
-\right).
-\end{equation}
-
-Since is $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}>0$ always, the sign
-of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
-RHS. Rearranging we find that $\frac
-{\partial G} {\partial x} < 0$ if
-
-\begin{equation}\label{eq:dg3}
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
-< \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x} \, \frac{ {\rm d}B} {{\rm d}t},
-\end{equation}
-
-or equivalently
-\begin{equation}\label{eq:dg4}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-<
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
-
-In words, there will be a negative relationship between trait x and growth if the
-relative change in production due to change in traits is less than the relative change
-in the total construction cost for a unit of leaf area due to change in trait. Crucially,
-key elements in eq. \ref{eq:dg3} are modified by height.
-
-We can also solve for trait values $x$ maximising growth rate, i.e. the value
-giving $\partial G /\partial x = 0$:
-
-\begin{equation}\label{eq:max}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-=
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
-
-\subsection{Leaf construction cost}
-
-If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand
-eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm
-{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
-
-\begin{equation}\label{eq:dbdt2}
-\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
-\end{equation}
-
-where
-$c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$
-is independent of $\phi$, and thus
-
-\begin{equation}\label{eq:dbdt3}
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
-(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
-\end{equation}
-
-Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent
-of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
-{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
-
-Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
-with traits if
-
-\begin{equation} \label{eq:G6}
-\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
-< \frac{1}{\phi
- + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
- {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
-\end{equation}
-
-\subsection{Wood construction cost}
-
-Do similar to previous section for wood construction cost.
-
-\newpage
-
-\section{Supplementary tables}\label{supplementary-tables}
-
-\begin{table}[ht]
-\caption{Model parameters}
-\centering
-{\footnotesize  % smaller font, double space
-\begin{doublespace}
-% \include{table-pars}
-
-\end{doublespace}
-}
-\label{tab:params}
-\end{table}
-
-\newpage
-
-\section{Supplementary figures}\label{supplementary-figures}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lmA_tradeoff.pdf}
-\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
-Data from \citep{Wright-2004} for 678 species from 51 sites, each
-point giving a species-average. Lines show standardised major axis lines
-fitted to data from each site, with intensity of shading adjusted
-according to strength of the relationship.\label{fS-leaf}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/mass_fraction.pdf}
-\caption{\textbf{Change in allocation with size.}
-\label{f-mass_fraction}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lmA_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
-and growth rate changes with plant size.} \label{f-lmA_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lmA_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
-and relative growth rate changes with plant size.}
-\label{f-lmA_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/rho_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
-and growth rate changes with plant size.} \label{f-rho_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/rho_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
-and relative growth rate changes with plant size.}
-\label{f-rho_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/hmat_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-growth rate changes with plant size.} \label{f-hmat_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/hmat_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-relative growth rate changes with plant size.}
-\label{f-hmat_growth_size_relative}}
-\end{figure}
-
-\end{appendices}
-\end{document}

### 2016-07-29 ms/ms-supp.Rnw
Separate doc for supp mat, use ms style from plant paper

--- 2016-07-28 ms/ms.tex
@@
-\documentclass[12pt, a4paper]{article}
-\RequirePackage[hmargin=2cm,vmargin=2cm]{geometry}
+\documentclass[10pt,twoside]{article}
 
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{authblk} % allows for better author affiliations
+\usepackage{suppmat}
+\usepackage{listings}
+\usepackage[T1]{fontenc} % Better underscores, shockingly.
 
-%appendix
-\usepackage[title,titletoc,toc]{appendix}
-\usepackage{setspace}  % for double spacing
-
-% Fancy HEADER
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\setlength{\headheight}{15pt}
-\pagenumbering{arabic}
-\lhead{\itshape{\nouppercase{\leftmark}}}
-\chead{}
-\rhead{\thepage}
-%\lfoot{v }
-\cfoot{}
-%\rfoot{\thepage}
-
-% allows for numbered referencing, as required by ecol letters
-\usepackage{etoolbox}
-\newbool{MyRefNumbers}
-\booltrue{MyRefNumbers} % comment to remove numbers in reference list
-
-\usepackage{natbib}
-\bibliographystyle{ecol_let}
-\setcitestyle{authoryear,open={(},close={)}}
-
-\usepackage{graphicx}
-\graphicspath{{../output/}}
+\input{ms-common-defs}
 
@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
 \makeatother
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+\titleprefix{Supporting information for:}
 
-\usepackage[rgb,dvipsnames]{xcolor}
-\definecolor{grey}{rgb}{0.5, 0.5, 0.5}
-
-\usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\hypersetup{breaklinks=true,
-            pdfauthor={},
-            pdftitle={Untangling the link between traits, size and growth rate in plants},
-            colorlinks=true,
-            citecolor=black,
-            urlcolor=grey,
-            linkcolor=grey}
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-
-\title{\LARGE Untangling the link between traits, size and growth rate in plants}
-
-\author[1]{Daniel S. Falster}
-\author[1]{Richard G. FitzJohn}
-
-\affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
-\renewcommand\Authands{ and }
-\date{\vspace{-3em}}
+\date{}
 
@@
 \maketitle
-\thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that
 
-\section{Ideas}\label{abstract}
+\tableofcontents
 
-Table near intro summarising empricial patterns to be explained by integrated model?
-
-Section 3.3 of King (2005) shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
-
-King 1999 also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
-Basically, the much greater path length from roots to leaves in trees vs.
-seedlings results in a much greater increment in support cost per increment in crown area for trees.
-This explanation is treated by the growth model of King (1999), which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
-
-\section{Abstract}\label{abstract}
-
-\section{Introduction}\label{introduction}
-
-Functional traits capture core differences in the strategies
-plants use to generate and invest surplus energy
- \citep{Wright-2004, Chave-2009, Westoby-2002}.
-Although most plants have the same basic physiological function
-and key resource requirements (carbon, nitrogen and water), species differ considerably in
-rates at which resources are acquired and invested into different tissues.
-Much of this functional variation relates to underlying
-differences in some prominent traits, including leaf
-mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
-diameter,  maximum  height (HMAX) and seed size  \citep{Wright-2004, Chave-2009}.
-Data for these traits now exists for up to 10\% of the world's 250000 plant species
-\citep{Cornwell-2014}. Increasingly, researchers have been looking to traits
-as potential indicators of the growth strategy and ecology of a species.
-
-While the influence of plant traits on elements of plant physiological
-function is increasingly understood, attempts at using traits to understand
-demographic rates have met with mixed success
- \citep{Wright-2010, Poorter-2008}. In seedlings, the trait LMA
--- the central element of the leaf economics spectrum
- \citep{Wright-2004} -- has been tightly related to relative growth
-rate in plant mass \citep{Lambers-1992, Wright-2000}. LMA and
-its correlate leaf lifespan have also been linked to height growth rate for
-small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These
-early successes prompted researchers to search for similar relationships in
-large plants. However, the results showed that in trees LMA was not correlated
-with diameter growth rate  \citep{Wright-2010, Poorter-2008,
-Herault-2011}. Meanwhile, other traits such as wood density showed
-strong relationships to growth in large plants \citep{Wright-2010},
-but only weak relationships in seedlings  \citep{Castro-1998}. Thus
-far, most trait growth relationships have been quantified for one or few
-traits and at a single life- stage. Yet the picture is emerging is that the
-link between traits and growth may be modified by plant size
- \citep{Ruger-2012}. This finding challenges a wide-spread
-assumption regarding traits: that particular ends of the trait spectrum translate directly into
-faster or slower growth (refs XX).
-
-The challenge of interpreting diverse empirical results seeking to link
-traits to growth rate is made all the harder because we currently lack clear
-expectations on the expected signal. Generating such expectations is one of the
-primary role for theory \citep{Kokko-2007}. Current empirical
-results suggest the effect of traits on growth changes with size and possibly
-also light environment, yet current theory says little about how such
-relationships might come about. A widely-used model shows that for seedlings,
-mass-based relative growth rate is linearly and negatively related to LMA
- \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
- An extension of the model suggests a similar
-relationship may hold at larger sizes \citep{Enquist-2007}. But as
-described above, this prediction is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth
-are largely absent.
-
-A further problem with existing theory is that the effects of traits are all
-realised via influences on mass production \citep{Enquist-2007},
-whereas the effect of traits such as LMA, WD, and Hmax is via allocated
-to different tissues \citep{Falster-2011}. There are in fact two problems with theory
-focussed solely on mass production. The first is that measuring mass
-production is only practical for small plants that can be easily harvested.
-For larger plants, growth is more often measured as increment in either diameter
-or height. For example, the CTFS monitors diameter growth in 49 large permanent
-plots across forests worldwide \citep{Anderson-2015};
- forestry have likewise recorded growth in diameter across
-large tracts of the globe \citep{Purves-2008}. The second problem is
-that, as we outline further below, because mass production is only one
-component of plant growth; theory focussing solely on mass production will
- always be limited.
-
-Effect of traits on shade tolerance also porrly udnerstood. Empirical -- shade
-
-(reich 1992, Niinements 2010, Gower-1993 - both show higher LAI in high LMA / long LL stands)
-
-Ultimately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
-Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
-In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover.
-All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of mainatining stems \citep{Givnish-1988}
-
-In this paper we provide a mechanistic framework for describing the potential
-effect of traits on growth throughout a plant's life-cycle, linking to
-commonly used metrics such as height and diameter growth. To help motivate our
-study we re-analyse demographic and trait data from an earlier study
- \citep{Wright-2010}, to show 1) that the link between traits and
-growth is stronger than previously thought, 2) that the influence of three
-prominent traits on growth rate is moderated by plant size, and 3) that the
-impact of size on the nature of relationships varies among traits
- \citep[see also]{Ruger-2012}. We then show how these results align with new
-set of theoretical predictions. Analysis extends ...
-Moreover, we show how traits moderate plant
-responses to light environment and also determine shade tolerance, supporting
-diverse empirical results \citep{Ruger-2012, Poorter-2006}. By
-disentangling the effects of plant size, light environment and traits on
-growth rates, we hope to provide a solid theoretical foundation for trait
-ecology and thus provide a platform for understanding growth across diverse
-species around the world.
-
-\section{Materials and Methods}\label{materials-and-methods}
-
-\subsection{Growth data}\label{growth-data}
-
-\subsection{Conceptual framework}\label{conceptual-framework}
-
-To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in
-seedlings to LMA \citep{Lambers-1992, Wright-2000}, to explicitly include influences of
-size, light environment, and other prominent traits. This model
-
-Builds on 2011 paper - highlight what's new from there.
-
-First derive ... generally true.
-
-Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
-therefore hold for any model of plant growth. To make explicit
-predictions, however, requires two additional elements: 1) A specific
-model describing how the various mass and area terms
-($M_{\rm l}, M_\textrm{s}, M_{\rm b}, M_{\rm r}, A_{\rm l}, A_\textrm{s}, A_{\rm b}$)
-vary relative to one another; and 2) For any trade-offs in plant
-function to be specified; this is where traits have an influence.  These two components are outlined in following sections.
-
-\subsubsection{Standard model for mass production}
-
-We begin with a standard model for the amount of biomass available for
-growth, ${\rm d}B / {\rm d}t$, given by the difference between income
-(total photosynthesis) and losses (respiration and turnover) within the
-plant \citep{Makela-1997, Thornley-2000, Falster-2011}:
-
-\begin{equation}\label{eq:dbdt}
-\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
-  = \underbrace{\strut y}_\textrm{yield}
-    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
-     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
-    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
-\end{equation}
-
-Here, $m,r$, and $k$ refer to the mass, respiration rate, and
-turnover rate of different tissues, denoted by subscripts $l$=leaves,
-$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
-rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
-assimilated carbon fixed in biomass (the remaining fraction being lost
-as growth respiration) (see Table \ref{tab:params} for units and
-definitions). Photosynthesis is proportional to leaf area,
-$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
-The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_\textrm{s}+M_{\rm r}.$
-
-The only potentially controversial element of in eq.
-\ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants, because the .....
-However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegtative part of the plant (refs XX) \citep{ Thornley-2000}.
-
-\subsubsection{Decomposing height, basal area and diameter growth}
-
-To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
-Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011}:
-
-\begin{equation} \label{eq:dhdt}
-\underbrace{\strut\frac{{\rm d}H}{{\rm d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{{\rm d}H}{{\rm d}A_{\rm l}}}_\textrm{architecture}
-\times \underbrace{\strut\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}}_\textrm{marginal leaf deployment}
-\times \underbrace{\strut\frac{{\rm d}M_{\rm a}}{{\rm d}B}}_\textrm{allocation to growth}
-\times \underbrace{\strut\frac{{\rm d}B}{{\rm d}t}}_\textrm{mass production}.
-\end{equation}
-
-The first term on the right of eq \ref{eq:dhdt},
-${\rm d}H / {\rm d}A_{\rm l}$, is the growth in plant height
-per unit growth in total leaf area -- accounting for the architectural
-strategy of the plant. Some species tend to leaf out more than grow
-tall, while other species emphasise vertical
-extension \citep{Poorter-2006}.
-
-The second term, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$,
-accounts for the marginal cost of deploying an additional unit of leaf
-area, including construction of the leaf itself and various support
-structures. As such, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$
-can itself be expressed as a sum of construction costs per unit leaf
-area:
-
-\begin{equation}\label{eq:daldmt}
-\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}
-= \big(\phi
- + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}\big)^{-1}.
-\end{equation}
-
-The third term in eq \ref{eq:dhdt},
-${\rm d}M_{\rm a} / {\rm d}P$, gives the fraction of net
-biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
-rather than reproduction or storage.
-
-In a similar way, basal area ($A_{\rm st}$) increment can be
-expressed as the sum of increments in sapwood, bark \& heartwood areas
-($A_\textrm{ss}, A_{\rm sb}, A_{\rm sh}$ respectively):
-$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm sb}}{{\rm d}t} + \frac{{\rm d}A_\textrm{ss}}{{\rm d}t} + \frac{{\rm d}A_{\rm sh}}{{\rm d}t}$.
-We now have an equation for basal area growth that conatins many of the same
-elements as eq. \ref{eq:dhdt}:
-
-\begin{equation}\label{eq:dast}
-\frac{{\rm d}A_{\rm st}}{{\rm d}t}=
-\left(\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}} + \frac{{\rm d}A_{\rm sb}}{{\rm d}A_{\rm l}}\right) \times
-\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}} \times \frac{{\rm d}M_{\rm a}}{{\rm d}B} \times \frac{{\rm d}B}{{\rm d}t} +\frac{{\rm d}A_{\rm sh}}{{\rm d}t} .
-\end{equation}
-
-Diameter growth is then given by the geometric relationship between stem
-diameter ($D$) and $A_{\rm st}$:
-
-\begin{equation} \label{eq:dDdt}
-\frac{{\rm d}D}{{\rm d}t}= \sqrt{\frac{\pi}{A_{\rm st}}} \frac{{\rm d}A_{\rm st}}{{\rm d}t}.
-\end{equation}
-
-\subsubsection{A functional-balance model for allocation}
-
-To address the first of these challenges, we use a functional balance
-model \citep{Yokozawa-1995, Falster-2011} assuming
-1) an allometric relation between total leaf area and plant height, and
-2) a constant ratio between total leaf area and each of sapwood
-cross-sectional area, bark cross-sectional area, and total root mass.
-These assumptions, which are well-supported by data (Fig.
-\ref{f-assumptions}), lead immediately to an allometric model for plant
-size (Table \ref{tab:allometry}, see Supplementary material for
-derivations). Substituting from Table \ref{tab:allometry} into eqs.
-\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
-for plant growth in relation to size.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  four key assumptions.
-\item
-  simple linear functions non-critical for results of this paper,
-\item
-  explore using an example model, but any model with basic size effects
-  will display same behaviour
-
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    details about our model ()
-  \item
-    show argument hold intuitively
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{Embeeding trait-related trade-offs}
-
-use table with arrows to
-
-By definition, the main effect of increasing $\phi$ is to decrease the
-marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
-growth rate. However, to fully account for the influence of $\phi$ on
-growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction \citep{Wright-2004}.
-This is achieved by linking the turnover rate of leaf in eq.
-\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
-the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
-shown to describe patterns across diverse
-ecosystems \citep{Wright-2004} (Fig. \ref{fS-leaf}).
-
-As for leaf, cheaper stem construction (lower $\rho$) decreases the
-cost of deploying a unit of leaf area, and may thereby increase growth
-rates. However, the cost of cheap stem construction must also be
-accounted for. However, the direct physiological trade-offs of lower
-$\rho$ are less understood than for $\phi$. One possibility is that
-lower $\rho$ decreases mechanical strength resulting in higher
-mortality \citep{Chave-2009, Wright-2010}. Under
-this scenario, decreasing $\rho$ will always provide a growth
-advantage, because the costs of low $\rho$ are not realised within the
-terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
-\ref{tab:trade-offs}). In addition, lower $\rho$ may increase turnover
-of sapwood. Mirroring our assumptions for leaves, we let sapwood
-turnover decrease with $\rho$,
-i.e.$k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
-
-Unlike leaf and stem traits, height at maturation ($h_m$) moderates
-growth by adjusting the amount of energy invested in growth, i.e.~the
-term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqs.
-\ref{eq:dhdt} and \ref{eq:dast}.
-
-\subsection{Analysis}\label{analysis}
-
-\newpage
-\section{Results and Discussion}\label{results}
-
-\subsection{Changes in growth rate with size}
-
-This new growth model captures the intrinsically size-dependent nature
-of plant growth, recovering well-known
-empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern
-with size, first increasing then decreasing. This pattern results from
-systematic changes in the four components of eq. \ref{eq:dhdt} with size
-(Fig. \ref{f-conceptual}), including a strong decline in the fraction of plant
-that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
-and increasing reproductive allocation. In contrast, basal-area growth
-continues to increase with size \citep{Sillett-2010, Stephenson-2014},
-due to an increasing influence of stem
-turnover. Diameter growth shows a weakly hump shaped
-curve \citep{Herault-2011}, tapering off slightly at larger
-sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
-mainly because of increased reproductive allocation in older trees
-(Fig?). All growth measures decrease sharply with size when expressed as
-relative growth rates \citep{Iida-2014}.
-
-\subsection{How traits influence growth rate}
-
-\subsubsection{LMA}
-
-Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
-\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
-changes with plant size ( \ref{f-lmA_growth_size}).
-Decreasing $\phi$ has two impacts on growth rate. First, lower
-$\phi$ increases marginal leaf deployment
-(${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on
-construction costs. Second, lower $\phi$ decreases net production
-(${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether
-lower $\phi$ increases growth thus depends on the relative magnitude
-of these two effects. When plants are small the effect on leaf
-deployment rate is bigger and so decreasing $\phi$ increases growth
-rate. When plants are large, the influence of $\phi$ on leaf
-deployment rate is diminished, because the costs of building other
-supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
-\ref{f-mass_fraction}). The net result is that at larger sizes, low
-$\phi$ is no longer advantageous for growth (Fig.
-\ref{f-lmA_growth_size}d).
-
-\subsubsection{Wood density}
-
-Under this scenario, decreasing $\rho$ provides an advantage in height
-growth only when the benefits of low $\rho$ outweigh the costs. This
-effect is strongest at intermediate sizes. The decline in height growth
-at low $\rho$ with size is more marked than for either stem basal-area
-and diameter growth, because high sapwood turnover also contributes to
-this growth measure. Thus diameter growth continues, even as net
-production approaches zero.
-
-\subsubsection{hmat}
-
- Greater height at maturation ($h_m$)
-can thus lead to a growth advantage by increasing
-$\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ . At smaller sizes, there is no differentiation
-among species, because most species are focusing on growth. At larger sizes, individuals of some species are
-allocating a majority of their surplus energy to growth, leaving a
-strong signal in potential growth rate.
-
-\subsubsection{Responsiveness to light}
-
-Two other important predictions arise directly from linking traits to
-growth rate. First, traits moderate the
-responsiveness of growth to changes in light environment. This response
-arises simply because species with higher potential growth rate have
-greater potential plasticity. Recent analyses have shown that species
-with low $rho$ increase growth more substantially with incraeses in
-light \citep{Ruger-2012}. This is exactly the behaviour
-exhibited in our model (Fig. \ref{f-growth_light_dia}), at least for small
-to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
-of low $\phi$ are lost (Fig. \ref{f-growth_light_dia}). This results also
-agrees with empirical results showing that growth for species with low
-$rho$ actually decreases as size
-increases \citep{Ruger-2012}. In contrast, growth rate tends
-to increase in size for the species with highest
-$\rho$ \citep{Ruger-2012}.
-
-Variation in $\phi$ also moderates the response of growth to changes
-in light, with species having the lowest $rho$ being most responsive.
-However, unlike for $\rho$, the effect appears only for the smallest
-size classes. Accordingly, empirical analyses have detected such a link
-in small plants but not for plants \textgreater{} 0.01m
-dbh \citep{Ruger-2012}.
-
-\subsection{Growth is more than just photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}
-
-Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
-\citep{Lambers-1992, Cornelissen-1996, Wright-2000, Enquist-2007},
-${\rm d}P / {\rm d}t / M_{\rm a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
-(see Appendix \ref{app:traits-RGR}).
-
-Recently realisation that allocation and turnover major areas of
-uncertainty
-
-Model reconciles idea that traits can have negative impact on carbon
-budget, but still have positive influence on growth rates
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Enquist 2007: sla and wd increase mass-production, where as in our
-  theory effects on mass production are negative. (also contradicts
-  earlier model from Enquist 1999, where argued wood density had no
-  effect on mass production)
-\end{itemize}
-
-Assumes low lma and wd used to reduce costs of building leaf and stem,
-ie. total leaf area and stem corss section remains same. In this way
-differs to assumptions of recent work by anten\_role\_2010,
-larjavaara\_rethinking\_2010.
-
-Model is realisation of several ideas that have been around for some
-time.
-
-Plasticity in trait values predicted from model:
-
-@Osazuwa-2014 - indreases in WD through ontogeny.
-
-\subsection{Shade tolerance}\label{shade-tolerance}
-
-Second, species with low $\phi$ and low $\rho$ are predicted to be
-less shade tolerant (Fig. \ref{f-wplcp}). At low $\phi$ ($\rho$),
-leaf (sapwood) turnover is higher and thus a greater light income is
-needed to offset these costs. As previously suggested
- \citep{Givnish-1988}, shade-tolerance also decreases with
-height because as size increase, the total amount of energy needed to
-offset respiratory and turnover costs in the stem also increases. These
-finding matches well known empirical patterns, where both $\phi$ and
-$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,osunkoya-1996}.
-
-Results fo poort \& binger for surival strongest for LL.
-
-Also baltzer.
-
-\subsection{Implications for trait-based approaches}\label{implications-for-trait-based-approaches}
-
-How do comparative people interpret sla now - beyond fast slow.
-
-Think about traits as defining potential trajectory, rather than making
-species fast or slow growing per se.
-
-Selection on traits in different parts of lifecycle
-
-Trait plasticity - can explain, no clear advantage of low lma at large
-sizes
-
-Framework for generalising about effect of traits
-
-Towards a global model of growth:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  mechanistic model of height and dbh growth which accounts for effects
-  of size, light environment and traits
-
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    need to account for size long recognized, but not straightforward
-  \item
-    look at growth in wide range of measures: mass, height, dbh
-  \end{itemize}
-\item
-  Compared to previous work, shows effects of traits more about
-  allocation than mass production
-\end{itemize}
-
-Key patterns explained
-
-Data challenges
-
-Reproductive allocation
-
-Detecting trait signals in data.
-
-\begin{itemize}
-\item  Need more than simply look for correlation with mean growth rate
-\item
-  understand influences of size on trait of interest
-\item
-  traits define potential growth rate --\textgreater{} how to extract
-  this
-\item
-  ideally start to use mechanistic models
-\end{itemize}
-
-\section{Acknowledgments}
-
-Figure 2 uses image by Tracey Saxby, IAN Image Library http://ian.umces.edu/imagelibrary.
-
-\newpage
-
-\bibliography{references}
-
-\newpage
-
-\section{Tables}\label{tables}
-
-\begin{table}[ht]
- \caption{Key variables
-%  $a = l + ss + sb +  r$,
-%  $M_{sh}$,
-%  $M_t = M_a + M_{sh}$,
-%  $st = stem total$
-}
-\centering
-
-{\footnotesize  % smaller font, double space
- \begin{doublespace}
-  \begin{tabular}{p{2cm}p{2cm}p{7cm}}
-  \hline
-  Symbol & Unit & Description \\
-  \hline
-
-  $b$   & kg  & biomass originating from parent plant\\
-  $M_i$ & kg  & mass retained on plant in tissue $i$\\
-  $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
-  $h$   & m  & height of plant\\
-  $D$   & m  & diameter of plant stem\\
-
-  $y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
-  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthesic rate per unit area \\
-  $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
-  $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
-  $A$ & kg yr$^{-1}$ & total plant photosynthesis \\
-  $R$ & kg yr$^{-1}$ & total plant respiration \\
-  $T$ & kg yr$^{-1}$ & total plant turnover \\
-  \hline
-  \end{tabular}
-\end{doublespace}}
-\label{tab:definitions}
-\end{table}
-
-\newpage
-
-\begin{table}[ht]
-\caption{Equations for an allometric growth model, based on functional balance assumption in {\textbf a}. }
-\centering
-
-{\footnotesize  % smaller font, double space
- \begin{doublespace}
-  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
-  Variable & Function & Marginal cost & Growth rate\\ \hline
-  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
-  height &
-    $h = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
-    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  sapwoood area &
-    $A_\textrm{ss} = \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_\textrm{ss}}{{\rm d}t}  =\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  bark area &
-    $A_{\rm sb} = b \, \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm sb}}{{\rm d}t} = \frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\  \\
-  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
-  leaf mass &
-    $M_{\rm l} = \phi \, A_{\rm l} $ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  sapwood mass &
-    $M_\textrm{ss} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}t}  = \frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  bark mass &
-    $M_{\rm sb} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm sb}}{{\rm d}t}  = \frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  root mass &
-    $M_{\rm r} = \alpha_3 \, A_{\rm l}$ &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\ \\
-
-  \multicolumn{4}{l}{\textbf{c) Integrated quantities}} \\
-  heartwood area &
-    $A_{\rm sh} = \int_0^t \frac{{\rm d}A_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}A_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, A_\textrm{ss}$ \\
-  heartwood mass &
-    $M_{\rm sh} = \int_0^t \frac{{\rm d}M_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}M_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, M_\textrm{ss}$ \\
-  \hline \\
-\end{tabular}
-NB -  theta is sapwood area per leaf area. In tree 2 has been opposite, but that makes notation more clumsy.
-\end{doublespace}}
-\label{tab:allometry}
-\end{table}
-
-\newpage
-
-\begin{table}[ht]
-\caption{Effect of traits and related trade-offs on demography}
-\centering
-{\footnotesize  % smaller font, double space
-  \begin{doublespace}
-  \begin{tabular}[c]{l|c|ccccc|c|c}
-  \multicolumn{9}{c}{}\\ \hline
-  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
-  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
-  & $h_0$
-  & $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$
-  & $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$
-  & $\frac{{\rm d}B}{{\rm d}t}$
-  & $\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}}$
-  & $\frac{1}{k_s}$ & &  \\\hline
-  \textbf{This study}&&&&&&&&\\
-  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
-  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
-  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
-  &&&&&&&&\\\hline
-  \textbf{Possible extensions}&&&&&&&&\\
-  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
-  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
-  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
-  \end{tabular}
-  \end{doublespace}
-}
-\label{tab:trade-offs}
-\end{table}
-
-\newpage
-
-\section{Figures}\label{figures}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/figure_conceptual.pdf}
-\caption{\textbf{Conceptual framework for understanding size-dependent
-changes in growth rate.} Hump-shaped relationship between height growth rate and
-size. \label{f-conceptual}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/allometry.pdf}
-\caption{\textbf{Key assumptions of a functional balance and trait
-trade-offs model, evaluated using global dataset.} We used the biomass and
-allometry database to evaluate model assumptions about \textbf{a,}
-scaling of leaf area with plant height, \textbf{b} Scaling of sapwood
-area with leaf area, and \textbf{c} scaling of root mass with leaf area.
-Each dot is a single plant. Lines show standardised major axis lines
-fitted to data from each site, with intensity of shading adjusted
-according to strength of the relationship. Colours indicate vegetation
-type. Dashed black lines show values expected under functional-balance
-assumption (see Supplementary text for details). \label{f-assumptions}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/growth_light_dia.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (diameter growth).} Panels show predicted relationship between
-specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{f-growth_light_dia}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/growth_light_mass.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (mass growth).} Panels show predicted relationship between
-specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{f-growth_light_mass}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/max_leaf_above.pdf}
-\caption{\textbf{The effect of traits on growth changes with size and
-light environment.} Traits moderate the responsiveness of growth to changes in light
-environmentThis figure needs to show more of key result with respect to to size.
-Possibly separate figure. \label{f-shifts}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/max_leaf_above.pdf}
-\caption{\textbf{Low construction cost leads to shade intolerance,
-because of costs of high turnover.} Panels show effect of traits on
-maximum amount of shading that can be endured before net production (eq.
-\ref{eq:dbdt}) reaches zero. Lines indicate relationship for plants of a
-given height. \label{f-wplcp}}
-\end{figure}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{secnumdepth}{0}
 
@@
 
-\begin{appendices}
-\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-
-\section{Relationship between traits and relative growth rate} \label{app:traits-RGR}
-
-We start with eq. \ref{eq:dbdt} for mass-based growth. For seedlings, which are
-young and mostly leaf, it is reasonable to ignore all turnover terms as
-well as the respiration terms for non-leaf tissues. Net production then
-becomes a linear function of leaf area and net photosynthesis per leaf
-area ($A_\textrm{net} = y(A - r_{\rm l})$), making relative growth
-rate a linear function of $\phi$:
-
-\begin{equation}\label{eq:RGR}
-\underbrace{\strut\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}. \end{equation}
-
-Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to $\phi$\citep{Wright-2000}, this
-approximation does not directly link to other traits, or to the
-variables that are routinely collected for large trees: namely plant
-height ($h$) and stem cross-sectional area ($A_{\rm st}$) or
-diameter $D$.
-
-\section{Derivation of a simple allometric model of plant
-function}\label{app:func_balance}
-
-Here we describe an allometric model linking the various size dimensions
-of a plant required by most ecologically realistic vegetation models
-(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
-roots) to a plant height. Table \ref{tab:allometry} provides a summary
-of the derived model, while Table \ref{tab:params} provdies estimates on
-key parameters.
-
-\subsection{Leaf area}\label{leaf-area}
-
-Based on empirically observed allometries (see main text), we assume an
-allometric log-log scaling relationship between the accumulated leaf
-area of a plant and its height:
-
-\begin{equation}\label{eq:ha}
-A_{\rm l}=\alpha_1 \, h^{\beta_1}.
-\end{equation}
-
-Note, scaling relationship reversed from \citep{Falster-2011}.
-
-\subsection{Mass of sapwood}\label{mass-of-sapwood}
-
-We follow the model of \citep{Yokozawa-1995} describing the
-vertical distribution of leaf area within the crowns of individual
-plants. This model can account for a variety of canopy profiles through
-a single parameter $\eta$. Setting $\eta=1$ results in a conical
-canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
-, give a top-weighted canopy profile similar to those seen among
-angiosperms. Let $S(z,h)$ be the sapwood area at height $z$ for a
-plant with top height $h$. Following Yokozawa and Hara (1995) we
-assume a relationship between $S(z,h)$ and height such that
-
-\begin{equation}\label{eq:crown1}
-\frac{S(z,h)}{S(0,h)}= \big(1-\big(\frac{z}{h}\big)^\eta\big)^2.
-\end{equation}
-
-We also assume that each unit of sapwood area supports a fixed area of
-leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
-canopy area of a plant relates to basal sapwood area $S(0,h)$:
-
-\begin{equation}\label{eq:crown2}
-\frac{M_{\rm l}}{\phi}= \theta \, S(0,h).
-\end{equation}
-
-Integrating $S(z,h)$ gives a solution for the total mass of sapwood in
-the plant:
-
-\begin{equation}\label{eq:ms1}
-M_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, {\rm d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}
-
-where
-$\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}$ \citep{Yokozawa-1995}.
-Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
-expression for sapwood mass as a function leaf area and height:
-
-\begin{equation}\label{eq:ms2}
-M_\textrm{s}=\rho \, \eta_c \, \theta \, A_{\rm l} \, h.
-\end{equation}
-
-\subsection{Bark mass}\label{bark-mass}
-
-Bark and phloem tissue are modelled using an analogue of the pipe model,
-leading to a similar equation as that for sapwood mass (eq.
-\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
-assumed to be a constant fraction $b$ of sapwood area per unit leaf
-area such that
-
-\begin{equation}\label{eq:mb}
-M_{\rm b}=b M_\textrm{s}.
-\end{equation}
-
-\subsection{Root mass}\label{root-mass}
-
-Also consistent with pipe-model assumption, we assume a fixed ratio of
-root mass per unit leaf area
-
-\begin{equation}\label{eq:mr}
-M_{\rm r}=\alpha_3 \, A_{\rm l}.
-\end{equation}
-
-Even though nitrogen and water uptake are not modelled explicitly,
-imposing a fixed ratio of root mass to leaf area ensures that
-approximate costs of root production are included in calculations of
-carbon budget.
-
-\section{Effect of traits on height growth} \label{app:traits_max}
-
-We want to show how two traits $\phi$ and $\rho$ influence growth rate,
-by looking at the effect of
-traits on the components of growth rate (eq. \ref{eq:dhdt}. To make
-the analysis more tractable, we will assume we are dealing with a plant of
-given height where 100\% of available energy is allocated to growth,
-i.e. $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=1$. Note also
-that by the assumptions of our model and above, traits do not influence
-$\frac{{\rm d}H} {{\rm d}A_{\rm l}}$. Thus we have
-
-\begin{equation} \label{eq:G2}
-G = \frac{{\rm d}H}{{\rm d}t} = c_1   \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}  \frac{ {\rm d}B} {{\rm d}t} \right),
-\end{equation}
-
-where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eq.
-\ref{eq:G2} is a product of the form $Y(x) = W(x) \times Z(x)$. For,
-equations of this type, the derivative $\partial{Y}/\partial{x}$ is
-given by $\partial{W}/\partial{x} \, Z + W \, \partial{Z}/\partial{x}$. Thus
-the derivative of $G$ with respect to some trait $x$  is given by
-
-\begin{equation} \label{eq:dG}
-\frac{\partial G} {\partial x} =
-\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}{\partial x}
- \, \frac{ {\rm d}B} {{\rm d}t}
-+ \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\, \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}.
-\end{equation}
-
-If this derivative is positive (negative), there will be a positive (negative) relationship
-between traits and growth at the height where it is evaluated.
-
-We further simplify by expanding the derivative of eq. \ref{eq:daldmt}:
-
-\begin{equation} \label{eq:dA_dx}
-\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}
-{\partial x} = -\left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\ \right)^2
-\, \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}.
-\end{equation}
-
-Combing eqs. \ref{eq:dG}-\ref{eq:dA_dx}, we have
-
-\begin{equation} \label{eq:dG2}
-\frac{\partial G} {\partial x} =
-\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\left(
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
-- \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}
- \, \frac{ {\rm d}B} {{\rm d}t}
-\right).
-\end{equation}
-
-Since is $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}>0$ always, the sign
-of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
-RHS. Rearranging we find that $\frac
-{\partial G} {\partial x} < 0$ if
-
-\begin{equation}\label{eq:dg3}
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
-< \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x} \, \frac{ {\rm d}B} {{\rm d}t},
-\end{equation}
-
-or equivalently
-\begin{equation}\label{eq:dg4}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-<
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
-
-In words, there will be a negative relationship between trait x and growth if the
-relative change in production due to change in traits is less than the relative change
-in the total construction cost for a unit of leaf area due to change in trait. Crucially,
-key elements in eq. \ref{eq:dg3} are modified by height.
-
-We can also solve for trait values $x$ maximising growth rate, i.e. the value
-giving $\partial G /\partial x = 0$:
-
-\begin{equation}\label{eq:max}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-=
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
-
-\subsection{Leaf construction cost}
-
-If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand
-eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm
-{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
-
-\begin{equation}\label{eq:dbdt2}
-\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
-\end{equation}
-
-where
-$c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$
-is independent of $\phi$, and thus
-
-\begin{equation}\label{eq:dbdt3}
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
-(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
-\end{equation}
-
-Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent
-of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
-{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
-
-Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
-with traits if
-
-\begin{equation} \label{eq:G6}
-\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
-< \frac{1}{\phi
- + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
- {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
-\end{equation}
-
-\subsection{Wood construction cost}
-
-Do similar to previous section for wood construction cost.
-
-\newpage
-
-\section{Supplementary tables}\label{supplementary-tables}
+\section{Supporting Tables}\label{app:supp_info_table}
 
@@
 
-\section{Supplementary figures}\label{supplementary-figures}
+\section{Supporting Figures}\label{app:supp_info_figures}
 
@@
 \centering
-\includegraphics{SI/lmA_tradeoff.pdf}
+\includegraphics{SI/lma_tradeoff.pdf}
 \caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
@@
 point giving a species-average. Lines show standardised major axis lines
-fitted to data from each site, with intensity of shading adjusted
+fitted to datsa from each site, with intensity of shading adjusted
 according to strength of the relationship.\label{fS-leaf}}
@@
 \centering
-\includegraphics{SI/lmA_effects_at_diameters.pdf}
+\includegraphics{SI/lma_effects_at_diameters.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
-and growth rate changes with plant size.} \label{f-lmA_growth_size}}
+and growth rate changes with plant size.} \label{f-lma_growth_size}}
 \end{figure}
@@
 \centering
-\includegraphics{SI/lmA_effects_at_diameters_relative.pdf}
+\includegraphics{SI/lma_effects_at_diameters_relative.pdf}
 \caption{\textbf{The expected correlation between leaf-construction cost
 and relative growth rate changes with plant size.}
-\label{f-lmA_growth_size_relative}}
+\label{f-lma_growth_size_relative}}
 \end{figure}
@@
 
-\end{appendices}
+\clearpage
+
+\bibliography{references}
+
 \end{document}

### 2016-07-29 ms/ms.tex
Add recent references
Separate doc for supp mat, use ms style from plant paper

@@
-\documentclass[12pt, a4paper]{article}
-\RequirePackage[hmargin=2cm,vmargin=2cm]{geometry}
-
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{authblk} % allows for better author affiliations
+\documentclass[a4paper,11pt]{article}
+\usepackage[osf]{mathpazo}
+\usepackage{ms}
+\usepackage[]{natbib}
+\usepackage{floatpag}
+\floatpagestyle{empty}
+
+%\raggedright
+
+\usepackage{graphicx}
+
+% Enable cross referencing to Evidence document
+\usepackage{xr}
+\externaldocument{ms-supp}
+
+\input{ms-common-defs}
+\date{}
+
+\mstype{Research Article}
+
+\keywords{growth strategy, life-history traits, maximum height, ontogenetic stage, photosynthetic rate, plant development, seed mass, specific leaf area, wood density}
 
@@
 
-% Fancy HEADER
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\setlength{\headheight}{15pt}
-\pagenumbering{arabic}
-\lhead{\itshape{\nouppercase{\leftmark}}}
-\chead{}
-\rhead{\thepage}
-%\lfoot{v }
-\cfoot{}
-%\rfoot{\thepage}
-
-% allows for numbered referencing, as required by ecol letters
-\usepackage{etoolbox}
-\newbool{MyRefNumbers}
-\booltrue{MyRefNumbers} % comment to remove numbers in reference list
-
-\usepackage{natbib}
-\bibliographystyle{ecol_let}
-\setcitestyle{authoryear,open={(},close={)}}
-
-\usepackage{graphicx}
-\graphicspath{{../output/}}
-
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
-
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
-
-\usepackage[rgb,dvipsnames]{xcolor}
-\definecolor{grey}{rgb}{0.5, 0.5, 0.5}
-
-\usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\hypersetup{breaklinks=true,
-            pdfauthor={},
-            pdftitle={Untangling the link between traits, size and growth rate in plants},
-            colorlinks=true,
-            citecolor=black,
-            urlcolor=grey,
-            linkcolor=grey}
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-
-\title{\LARGE Untangling the link between traits, size and growth rate in plants}
-
-\author[1]{Daniel S. Falster}
-\author[1]{Richard G. FitzJohn}
-
-\affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
-\renewcommand\Authands{ and }
-\date{\vspace{-3em}}
-
 \begin{document}
-
-\maketitle
-\thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that
+\mstitlepage
+\noindent
+\parindent=1.5em
+\addtolength{\parskip}{.3em}
+\doublespacing
+\linenumbers
+\section{Summary}\label{abstract}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item We ...
+\end{enumerate}
 
@@
 
-Table near intro summarising empricial patterns to be explained by integrated model?
+New refs: \citep{Ishihara-2016, Duursma-2016, Gibert-2016, Falster-2016, Kunstler-2015,Falster-2015b, Falster-2015a,Wenk-2015,Li-2014}
+
+Table near intsro summarising empirical patterns to be explained by integrated model?
 
@@
 Assumes low lma and wd used to reduce costs of building leaf and stem,
-ie. total leaf area and stem corss section remains same. In this way
-differs to assumptions of recent work by anten\_role\_2010,
-larjavaara\_rethinking\_2010.
+ie. total leaf area and stem cross section remains same. In this way
+differs to assumptions of recent work by \citep{Anten-2010}, \citep{Larjavaara-2010}.
 
@@
 
-@Osazuwa-2014 - indreases in WD through ontogeny.
+\citep{Osazuwa-2014} - increases in WD through ontogeny.
 
@@
 finding matches well known empirical patterns, where both $\phi$ and
-$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,osunkoya-1996}.
+$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,Osunkoya-1996}.
 
@@
 
-\newpage
-
-\section{Supplementary tables}\label{supplementary-tables}
-
-\begin{table}[ht]
-\caption{Model parameters}
-\centering
-{\footnotesize  % smaller font, double space
-\begin{doublespace}
-% \include{table-pars}
-
-\end{doublespace}
-}
-\label{tab:params}
-\end{table}
-
-\newpage
-
-\section{Supplementary figures}\label{supplementary-figures}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lmA_tradeoff.pdf}
-\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
-Data from \citep{Wright-2004} for 678 species from 51 sites, each
-point giving a species-average. Lines show standardised major axis lines
-fitted to data from each site, with intensity of shading adjusted
-according to strength of the relationship.\label{fS-leaf}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/mass_fraction.pdf}
-\caption{\textbf{Change in allocation with size.}
-\label{f-mass_fraction}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lmA_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
-and growth rate changes with plant size.} \label{f-lmA_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lmA_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
-and relative growth rate changes with plant size.}
-\label{f-lmA_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/rho_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
-and growth rate changes with plant size.} \label{f-rho_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/rho_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
-and relative growth rate changes with plant size.}
-\label{f-rho_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/hmat_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-growth rate changes with plant size.} \label{f-hmat_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/hmat_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-relative growth rate changes with plant size.}
-\label{f-hmat_growth_size_relative}}
-\end{figure}
-
 \end{appendices}

### 2016-07-29 ms/figure_conceptual.tex
Outline of main structure, embed eqn references in table and figure

--- 2016-07-28 figure_conceptual.tex
@@
 \usepackage{helvet}
+
+\usepackage{graphicx}
+\graphicspath{{../}}
 
@@
 \newcommand{\rowlabel}[1]{
-  \node at ($(thisrow) + (0.25cm, 0cm)$)[anchor=west, text width=2cm] {
-  \footnotesize {\color{orange!50}  #1}};
+  \node at ($(thisrow) + (-0.75cm, -0.3cm)$)[anchor=west, text width=2cm] {
+  \footnotesize {\color{orange!50}\begin{equation}\label{#1}\end{equation}}};
 }
@@
 \defineRow{0}{thisrow}
-\rowtitle{Biomass production}
+\rowtitle{Biomass production rate}
 \firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/net_mass_production_dt_net_mass_production_dt}{}
@@
 \othercol{5}{$-{\rm T}$}{output/con/net_mass_production_dt_turnover}{}
-\rowlabel{(1)}
+\rowlabel{eq:dbdt}
 
 \defineRow{1}{thisrow}
-\rowtitle{Construction costs}
+\rowtitle{Leaf deployment per mass}
 \firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm t}}}$}{output/con/darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
 \othercol{2}{$\frac{{\rm d}M_{{\rm l}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
-\othercol{3}{$\frac{{\rm d}M_{\textrm{ss}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
-\othercol{4}{$\frac{{\rm d}M_{{\rm sb}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
+\othercol{3}{$\frac{{\rm d}M_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
+\othercol{4}{$\frac{{\rm d}M_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
 \othercol{5}{$\frac{{\rm d}M_{{\rm r}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_root_darea_leaf}{$\Bigg)^{-1}$}
- \rowlabel{(2)}
+ \rowlabel{eq:daldmt}
 \defineRow{2}{thisrow}
-\rowtitle{Leaf-area growth}
+\rowtitle{Leaf-area growth rate}
 \firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
@@
 \othercol{4}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/area_leaf_dt_net_mass_production_dt}{}
-\rowlabel{(3)}
+\rowlabel{eq:daldt}
 
 \defineRow{3}{thisrow}
-\rowtitle{Height growth}
+\rowtitle{Height growth rate}
 \firstcol{1}{$\frac{{\rm d}H}{{\rm d}t}$}{output/con/height_dt_height_dt}{}
@@
 \othercol{3}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
-\rowlabel{(4)}
+\rowlabel{eq:dhdt}
 
 \defineRow{4}{thisrow}
-\rowtitle{Basal-area growth}
+\rowtitle{Stem-area growth rate}
 \firstcol{1}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/con/area_stem_dt_area_stem_dt}{$\Bigg($}
-\othercol{2}{$\frac{{\rm d}A_{\textrm{ss}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_sapwood_darea_leaf}{$+$}
+\othercol{2}{$\frac{{\rm d}A_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_sapwood_darea_leaf}{$+$}
 \othercol{3}{$\frac{{\rm d}A_{\textrm{sb}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_bark_darea_leaf}{\hspace{-0.3em}$\Bigg)$}
 \othercol{4}{\hspace{-0.4em}$\times\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_stem_dt_area_leaf_dt}{$+$}
-\othercol{5}{$\frac{{\rm d}A_{{\rm sh}}}{{\rm d}t}$}{output/con/area_stem_dt_area_heartwood_dt}{}
-\rowlabel{(5)}
+\othercol{5}{$\frac{{\rm d}A_{{\rm h}}}{{\rm d}t}$}{output/con/area_stem_dt_area_heartwood_dt}{}
+\rowlabel{eq:dast}
 
 \defineRow{5}{thisrow}
-\rowtitle{Diameter growth}
+\rowtitle{Stem-diameter growth rate}
 \firstcol{1}{$\frac{{\rm d}D}{{\rm d}t}$}{output/con/diameter_stem_dt_diameter_stem_dt}{}
@@
 \othercol{3}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/con/diameter_stem_dt_area_stem_dt}{}
-\rowlabel{(6)}
+\rowlabel{eq:dDdt}
 

### 2016-07-29 ms/ms.tex
Outline of main structure, embed eqn references in table and figure

@@
 \floatpagestyle{empty}
+
+% Enable equation numbers in Table
+% Following http://tex.stackexchange.com/questions/62974/numbering-equations-in-tabular-environment
+\newcommand{\eqncounter}[1]{%
+\mbox{}\refstepcounter{equation}%
+$(\theequation)$%
+\label{#1}
+}
 
@@
 \externaldocument{ms-supp}
+\externaldocument{figure_conceptual}
 
@@
 \addtolength{\parskip}{.3em}
-\doublespacing
-\linenumbers
-\section{Summary}\label{abstract}
+% \doublespacing
+% \linenumbers
+\section{Summary}
 \begin{enumerate}
@@
 
-\section{Ideas}\label{abstract}
+% Consider Ecol Letters I deas and Perspectives: maximum of 7500 words (main text), and 10 figures, tables or boxes.
 
-New refs: \citep{Ishihara-2016, Duursma-2016, Gibert-2016, Falster-2016, Kunstler-2015,Falster-2015b, Falster-2015a,Wenk-2015,Li-2014}
+\section{To do}
 
-Table near intsro summarising empirical patterns to be explained by integrated model?
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item New refs: \citep{Ishihara-2016, Duursma-2016, Gibert-2016, Falster-2016, Kunstler-2015,Falster-2015b, Falster-2015a,Wenk-2015,Li-2014}
+\item Table near intro summarising empirical patterns to be explained by integrated model?
+\end{enumerate}
 
-Section 3.3 of King (2005) shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
-
-King 1999 also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
-Basically, the much greater path length from roots to leaves in trees vs.
-seedlings results in a much greater increment in support cost per increment in crown area for trees.
-This explanation is treated by the growth model of King (1999), which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
-
-\section{Abstract}\label{abstract}
-
-\section{Introduction}\label{introduction}
+\section{Introduction}
 
@@
 
-\section{Materials and Methods}\label{materials-and-methods}
-
-\subsection{Growth data}\label{growth-data}
-
-\subsection{Conceptual framework}\label{conceptual-framework}
+\section{Conceptual framework for understanding height, basal area and diameter growth}
 
@@
 
-Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
-therefore hold for any model of plant growth. To make explicit
-predictions, however, requires two additional elements: 1) A specific
-model describing how the various mass and area terms
-($M_{\rm l}, M_\textrm{s}, M_{\rm b}, M_{\rm r}, A_{\rm l}, A_\textrm{s}, A_{\rm b}$)
-vary relative to one another; and 2) For any trade-offs in plant
-function to be specified; this is where traits have an influence.  These two components are outlined in following sections.
+Equations \ref{eq:dbdt} -\ref{eq:dDdt} shown in Fig. \ref{fig:conceptual}.
+\setcounter{equation}{6}
 
-\subsubsection{Standard model for mass production}
+Requires definitions from Table \ref{tab:definitions}.
+
+(check this material covered in table: Here, $M, r$, and $k$ refer to the mass, respiration rate, and
+turnover rate of different tissues, denoted by subscripts $l$=leaves,
+$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
+rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
+assimilated carbon fixed in biomass (the remaining fraction being lost
+as growth respiration) (see Table \ref{tab:params} for units and
+definitions). Photosynthesis is proportional to leaf area,
+$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
+The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}.$
+)
 
@@
 
-\begin{equation}\label{eq:dbdt}
+\begin{equation}\label{eq:dbdt_full}
 \underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
@@
 
-Here, $m,r$, and $k$ refer to the mass, respiration rate, and
-turnover rate of different tissues, denoted by subscripts $l$=leaves,
-$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
-rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
-assimilated carbon fixed in biomass (the remaining fraction being lost
-as growth respiration) (see Table \ref{tab:params} for units and
-definitions). Photosynthesis is proportional to leaf area,
-$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
-The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_\textrm{s}+M_{\rm r}.$
-
 The only potentially controversial element of in eq.
-\ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
+\ref{eq:dbdt} \ref{eq:dbdt_full} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
 This assumption is likely to hold true for most woody plants, because the .....
-However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegtative part of the plant (refs XX) \citep{ Thornley-2000}.
-
-\subsubsection{Decomposing height, basal area and diameter growth}
+However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegetative part of the plant (refs XX) \citep{ Thornley-2000}.
 
 To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
-Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011}:
-
-\begin{equation} \label{eq:dhdt}
-\underbrace{\strut\frac{{\rm d}H}{{\rm d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{{\rm d}H}{{\rm d}A_{\rm l}}}_\textrm{architecture}
-\times \underbrace{\strut\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}}_\textrm{marginal leaf deployment}
-\times \underbrace{\strut\frac{{\rm d}M_{\rm a}}{{\rm d}B}}_\textrm{allocation to growth}
-\times \underbrace{\strut\frac{{\rm d}B}{{\rm d}t}}_\textrm{mass production}.
-\end{equation}
+Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Figure \ref{fig:conceptual})
 
@@
 
-\begin{equation}\label{eq:daldmt}
-\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}
-= \big(\phi
- + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}\big)^{-1}.
-\end{equation}
-
 The third term in eq \ref{eq:dhdt},
@@
 expressed as the sum of increments in sapwood, bark \& heartwood areas
-($A_\textrm{ss}, A_{\rm sb}, A_{\rm sh}$ respectively):
-$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm sb}}{{\rm d}t} + \frac{{\rm d}A_\textrm{ss}}{{\rm d}t} + \frac{{\rm d}A_{\rm sh}}{{\rm d}t}$.
-We now have an equation for basal area growth that conatins many of the same
-elements as eq. \ref{eq:dhdt}:
-
-\begin{equation}\label{eq:dast}
-\frac{{\rm d}A_{\rm st}}{{\rm d}t}=
-\left(\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}} + \frac{{\rm d}A_{\rm sb}}{{\rm d}A_{\rm l}}\right) \times
-\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}} \times \frac{{\rm d}M_{\rm a}}{{\rm d}B} \times \frac{{\rm d}B}{{\rm d}t} +\frac{{\rm d}A_{\rm sh}}{{\rm d}t} .
-\end{equation}
+($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively):
+$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
+We now have an equation for basal area growth that contains many of the same
+elements as eq. \ref{eq:dhdt}, see eq. \ref{eq:dast}.
 
 Diameter growth is then given by the geometric relationship between stem
-diameter ($D$) and $A_{\rm st}$:
-
-\begin{equation} \label{eq:dDdt}
-\frac{{\rm d}D}{{\rm d}t}= \sqrt{\frac{\pi}{A_{\rm st}}} \frac{{\rm d}A_{\rm st}}{{\rm d}t}.
-\end{equation}
+diameter ($D$) and $A_{\rm st}$, see eq. \ref{eq:dDdt}
 
@@
 These assumptions, which are well-supported by data (Fig.
-\ref{f-assumptions}), lead immediately to an allometric model for plant
+\ref{fig:assumptions}), lead immediately to an allometric model for plant
 size (Table \ref{tab:allometry}, see Supplementary material for
@@
 
-\subsubsection{Embeeding trait-related trade-offs}
+\section{Derivation of a simple allometric model of plant function}
+
+Here we describe an allometric model linking the various size dimensions
+of a plant required by most ecologically realistic vegetation models
+(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
+roots) to a plant height. Table \ref{tab:allometry} provides a summary
+of the derived model, while Table \ref{tab:params} provdies estimates on
+key parameters.
+
+Based on empirically observed allometries (see main text), we assume an
+allometric log-log scaling relationship between the accumulated leaf
+area of a plant and its height: Eq. \ref{eq:ha}.
+
+Note, scaling relationship reversed from \citep{Falster-2011}.
+
+We follow the model of \citep{Yokozawa-1995} describing the
+vertical distribution of leaf area within the crowns of individual
+plants. This model can account for a variety of canopy profiles through
+a single parameter $\eta$. Setting $\eta=1$ results in a conical
+canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
+, give a top-weighted canopy profile similar to those seen among
+angiosperms. We also assume that each unit of sapwood area supports a fixed area of
+leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
+canopy area of a plant relates to basal sapwood area. This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (Eq. \ref{eq:ms2}).
+
+Bark and phloem tissue are modelled using an analogue of the pipe model,
+leading to a similar equation as that for sapwood mass (eq.
+\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
+assumed to be a constant fraction $b$ of sapwood area per unit leaf
+area such that (Eq. \ref{eq:mb}).
+
+Also consistent with pipe-model assumption, we assume a fixed ratio of
+root mass per unit leaf area (Eq. \ref{eq:mr}).
+
+Even though nitrogen and water uptake are not modelled explicitly,
+imposing a fixed ratio of root mass to leaf area ensures that
+approximate costs of root production are included in calculations of
+carbon budget.
+
+\subsubsection{Embedding trait-related trade-offs}
 
@@
 turnover decrease with $\rho$,
-i.e.$k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
+i.e.$k_{\rm s}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
 
@@
 
-\subsection{Analysis}\label{analysis}
+\subsection{Analysis}
 
 \newpage
-\section{Results and Discussion}\label{results}
+\section{Results and Discussion}
+
+Section 3.3 of King (2005) shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
+
+King 1999 also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
+Basically, the much greater path length from roots to leaves in trees vs.
+seedlings results in a much greater increment in support cost per increment in crown area for trees.
+This explanation is treated by the growth model of King (1999), which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
 
@@
 systematic changes in the four components of eq. \ref{eq:dhdt} with size
-(Fig. \ref{f-conceptual}), including a strong decline in the fraction of plant
-that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
+(Fig. \ref{fig:conceptual}), including a strong decline in the fraction of plant
+that is leaf declines with increasing size (Fig. \ref{fig:mass_fraction}),
 and increasing reproductive allocation. In contrast, basal-area growth
@@
 
+For growth traits can derive rule:
+\begin{equation}\label{eq:dg4}
+\frac{
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
+{\frac{ {\rm d}B} {{\rm d}t}}
+<
+\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
+}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
+\end{equation}
+
+In words, there will be a negative relationship between trait x and growth if the
+relative change in production due to change in traits is less than the relative change
+in the total construction cost for a unit of leaf area due to change in trait. Crucially,
+key elements in eq. \ref{eq:dg3} are modified by height.
+
+We can also solve for trait values $x$ maximising growth rate, i.e. the value
+giving $\partial G /\partial x = 0$:
+\begin{equation}\label{eq:max}
+\frac{
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
+{\frac{ {\rm d}B} {{\rm d}t}}
+=
+\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
+}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
+\end{equation}
+
 \subsubsection{LMA}
@@
 \ref{eq:dDdt} predict a relationship between $\phi$ and growth that
-changes with plant size ( \ref{f-lmA_growth_size}).
+changes with plant size ( \ref{fig:lmA_growth_size}).
 Decreasing $\phi$ has two impacts on growth rate. First, lower
@@
 supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
-\ref{f-mass_fraction}). The net result is that at larger sizes, low
+\ref{fig:mass_fraction}). The net result is that at larger sizes, low
 $\phi$ is no longer advantageous for growth (Fig.
-\ref{f-lmA_growth_size}d).
+\ref{fig:lmA_growth_size}d).
+
+If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand
+eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm
+{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
+
+\begin{equation}\label{eq:dbdt2}
+\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
+\end{equation}
+
+where
+$c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$
+is independent of $\phi$, and thus
+
+\begin{equation}\label{eq:dbdt3}
+\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
+(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
+\end{equation}
+
+Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent
+of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
+{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
+
+Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
+with traits if
+
+\begin{equation} \label{eq:G6}
+\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
+< \frac{1}{\phi
+ + \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
+ {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
+\end{equation}
 
@@
 light \citep{Ruger-2012}. This is exactly the behaviour
-exhibited in our model (Fig. \ref{f-growth_light_dia}), at least for small
+exhibited in our model (Fig. \ref{fig:growth_light_dia}), at least for small
 to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
-of low $\phi$ are lost (Fig. \ref{f-growth_light_dia}). This results also
+of low $\phi$ are lost (Fig. \ref{fig:growth_light_dia}). This results also
 agrees with empirical results showing that growth for species with low
@@
 
-\subsection{Growth is more than just photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}
+\subsection{Growth is more than just photosynthesis}
 
@@
 
-\subsection{Shade tolerance}\label{shade-tolerance}
+\subsection{Shade tolerance}
 
 Second, species with low $\phi$ and low $\rho$ are predicted to be
-less shade tolerant (Fig. \ref{f-wplcp}). At low $\phi$ ($\rho$),
+less shade tolerant (Fig. \ref{fig:wplcp}). At low $\phi$ ($\rho$),
 leaf (sapwood) turnover is higher and thus a greater light income is
@@
 
-\subsection{Implications for trait-based approaches}\label{implications-for-trait-based-approaches}
+\subsection{Implications for trait-based approaches}
 
@@
 
-\section{Acknowledgments}
+\section{Acknowledgements}
 
@@
 
-\bibliography{references}
+\section{Tables}
+
+\begin{table}[ht]
+ \caption{Key phenomena to be explained}
+\centering
+
+{\footnotesize  % smaller font, double space
+ \begin{doublespace}
+  \begin{tabular}{p{7cm}p{4cm}}
+  \hline
+  Phenomena & References \\
+  \hline
+  hump shape growth & \citep{King-2005} \\
+  \hline
+  \end{tabular}
+\end{doublespace}}
+\label{tab:phenomena}
+\end{table}
 
@@
 
-\section{Tables}\label{tables}
-
 \begin{table}[ht]
- \caption{Key variables
-%  $a = l + ss + sb +  r$,
-%  $M_{sh}$,
-%  $M_t = M_a + M_{sh}$,
-%  $st = stem total$
-}
+ \caption{Key variables}
 \centering
@@
 
-  $b$   & kg  & biomass originating from parent plant\\
+  $B$   & kg  & biomass originating from parent plant\\
   $M_i$ & kg  & mass retained on plant in tissue $i$\\
   $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
-  $h$   & m  & height of plant\\
+  $H$   & m  & height of plant\\
   $D$   & m  & diameter of plant stem\\
-
-  $y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
-  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthesic rate per unit area \\
+  $\alpha_y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
+  $P$ & kg yr$^{-1}$ & total plant photosynthesis \\
+  $R$ & kg yr$^{-1}$ & total plant respiration \\
+  $T$ & kg yr$^{-1}$ & total plant turnover \\
+  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthetic rate per unit area \\
   $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
   $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
-  $A$ & kg yr$^{-1}$ & total plant photosynthesis \\
-  $R$ & kg yr$^{-1}$ & total plant respiration \\
-  $T$ & kg yr$^{-1}$ & total plant turnover \\
   \hline
@@
  \begin{doublespace}
-  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
-  Variable & Function & Marginal cost & Growth rate\\ \hline
-  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
+  \begin{tabular}{p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \hline
+  Variable & Function & Marginal cost & Growth rate & Eq.\\ \hline
+  \multicolumn{5}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
   height &
-    $h = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
+    $H = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
     $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
-  sapwoood area &
-    $A_\textrm{ss} = \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_\textrm{ss}}{{\rm d}t}  =\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
+    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
+  sapwood area &
+    $A_{\rm s} = \theta \, A_{\rm l}$ &
+    $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
+    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:a}\\
   bark area &
-    $A_{\rm sb} = b \, \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm sb}}{{\rm d}t} = \frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\  \\
-  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
+    $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
+    $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
+    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:a}\\  \\
+  \multicolumn{5}{l}{\textbf{b) Derived equation for mass of tissue }} \\
   leaf mass &
@@
     $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
+    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
   sapwood mass &
-    $M_\textrm{ss} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}t}  = \frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
+    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
+    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
   bark mass &
-    $M_{\rm sb} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm sb}}{{\rm d}t}  = \frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
+    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
+    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
   root mass &
@@
     $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\ \\
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\ \\
 
-  \multicolumn{4}{l}{\textbf{c) Integrated quantities}} \\
+  \multicolumn{5}{l}{\textbf{c) Integrated quantities}} \\
   heartwood area &
-    $A_{\rm sh} = \int_0^t \frac{{\rm d}A_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
+    $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
      &
-    $\frac{{\rm d}A_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, A_\textrm{ss}$ \\
+    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:a} \\
   heartwood mass &
-    $M_{\rm sh} = \int_0^t \frac{{\rm d}M_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
+    $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
      &
-    $\frac{{\rm d}M_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, M_\textrm{ss}$ \\
+    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:a} \\
   \hline \\
@@
 
-\begin{table}[ht]
-\caption{Effect of traits and related trade-offs on demography}
+\newcommand{\sepp}{{\color{grey}/}}
+
+\begin{table}[h!]
 \centering
-{\footnotesize  % smaller font, double space
-  \begin{doublespace}
-  \begin{tabular}[c]{l|c|ccccc|c|c}
-  \multicolumn{9}{c}{}\\ \hline
-  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
-  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
-  & $h_0$
-  & $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$
-  & $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$
-  & $\frac{{\rm d}B}{{\rm d}t}$
-  & $\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}}$
-  & $\frac{1}{k_s}$ & &  \\\hline
-  \textbf{This study}&&&&&&&&\\
-  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
-  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
-  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
-  &&&&&&&&\\\hline
-  \textbf{Possible extensions}&&&&&&&&\\
-  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
-  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
-  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
+\caption{Predicted effects of traits on key elements of plant function determining growth rate.
+Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
+Traits are: specific leaf area (= SLA), wood density (= WD), Maximum height (= $H_{max}$), Seed mass, and Maximum photosynthesis per unit leaf area (= Area).
+For further details, see main text.
+Adapted and expnaded from \citet{Gibert-2016}.
+}
+{\footnotesize
+\vspace{1cm}
+  \begin{tabular}{lllcccccc}
+  \hline
+  & & & {\bf Symbol} & {\bf SLA} & {\bf WD} & {$\bf H_{max}$} & {\bf Seed mass} & {$\bf A_{area}$} \\ \hline
+  \\
+  \multicolumn{9}{ l } {\textbf{Elements of equations \ref{eq:eq0} -- \ref{eq:eq4}}}  \\
+  & \multicolumn{2}{ l } {Leaf deployment per mass}  & $\ud A_{\rm l} / \ud M_{\rm a} $ & \\
+    & &     \tabitem leaf  &  & $\uparrow$ & - & - & - & - \\
+    & &     \tabitem sapwood & & - & $\downarrow$ & - & - & - \\
+    & &     \tabitem root & & - & - & - & - & - \\
+  & \multicolumn{2}{ c } {Allocation to vegetative growth} & $\ud M_{\rm a} / \ud B$ & - & - & $\uparrow$ & - & - \\
+  & & & & & & \\
+  & \multicolumn{2}{ l } {Net live biomass production} & $\ud B / \ud t$ & & & & & \\
+  & &     \tabitem photosynthetic rate & & - & - & - & - & $\uparrow$  \\
+  & &     \tabitem respiration rate  & & - & - & - & - & $\uparrow$  \\
+  & &     \tabitem leaf turnover rate & & $\downarrow$ & - & - & - & - \\
+  & \multicolumn{2}{ l } {Architecture} & $\ud H / \ud A_{\rm l}$ & - & - & - & - & - \\
+  & \multicolumn{2}{ l } {Sapwood area per leaf} & $ \ud A_{\rm s} / \ud A_{\rm l}$ & - & - & - & - & - \\
+  & \multicolumn{2}{ l } {Sapwood turnover} & $k_{\rm s} \, A_{\rm s}$ & - & - & - & - & - \\ \hline
+  \\
+  \multicolumn{9}{ l } {\textbf{Predicted effect on plant growth rate} (small plant {\sepp} large plant)} \\
+  & \multicolumn{2}{ l } {in mass} \\
+  & & \tabitem{Absolute} & $\ud M_{\rm a} / \ud t$ & $\uparrow$\sepp$\downarrow$ & $\downarrow${\sepp}$\downarrow$ &  $-${\sepp}$\uparrow$ & $\uparrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
+  & & \tabitem{Relative} & $\ud M_{\rm a} / (\ud t .
+M_{\rm a} )$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ &  $-${\sepp}$\uparrow$ & $\downarrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
+  & \multicolumn{2}{ l } {in height} \\
+  & & \tabitem{Absolute} & $\ud H / \ud t$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\uparrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
+  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\downarrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
+  & \multicolumn{2}{ l } {in diameter} \\
+  & & \tabitem{Absolute} & $\ud D / \ud t$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\uparrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
+  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\downarrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
+\hline
   \end{tabular}
-  \end{doublespace}
   }
@@
 \newpage
-
 \section{Figures}\label{figures}
-
-\newpage
 
@@
 changes in growth rate.} Hump-shaped relationship between height growth rate and
-size. \label{f-conceptual}}
+size. \label{fig:conceptual}}
 \end{figure}
@@
 type. Dashed black lines show values expected under functional-balance
-assumption (see Supplementary text for details). \label{f-assumptions}}
+assumption (see Supplementary text for details). \label{fig:assumptions}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\centering
+\includegraphics{main/mass_fraction.pdf}
+\caption{\textbf{Change in allocation with size.}
+\label{fig:mass_fraction}}
 \end{figure}
@@
 diameter and under a range of shading environments.
-\label{f-growth_light_dia}}
+\label{fig:growth_light_dia}}
 \end{figure}
@@
 diameter and under a range of shading environments.
-\label{f-growth_light_mass}}
+\label{fig:growth_light_mass}}
 \end{figure}
@@
 environmentThis figure needs to show more of key result with respect to to size.
-Possibly separate figure. \label{f-shifts}}
+Possibly separate figure. \label{fig:shifts}}
 \end{figure}
@@
 \ref{eq:dbdt}) reaches zero. Lines indicate relationship for plants of a
-given height. \label{f-wplcp}}
+given height. \label{fig:wplcp}}
 \end{figure}
@@
 
-\begin{appendices}
-\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-
-\section{Relationship between traits and relative growth rate} \label{app:traits-RGR}
+\section{Appendix A - Relationship between traits and relative growth rate} \label{app:traits-RGR}
 
@@
 
-\section{Derivation of a simple allometric model of plant
-function}\label{app:func_balance}
-
-Here we describe an allometric model linking the various size dimensions
-of a plant required by most ecologically realistic vegetation models
-(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
-roots) to a plant height. Table \ref{tab:allometry} provides a summary
-of the derived model, while Table \ref{tab:params} provdies estimates on
-key parameters.
-
-\subsection{Leaf area}\label{leaf-area}
-
-Based on empirically observed allometries (see main text), we assume an
-allometric log-log scaling relationship between the accumulated leaf
-area of a plant and its height:
-
-\begin{equation}\label{eq:ha}
-A_{\rm l}=\alpha_1 \, h^{\beta_1}.
-\end{equation}
-
-Note, scaling relationship reversed from \citep{Falster-2011}.
-
-\subsection{Mass of sapwood}\label{mass-of-sapwood}
-
-We follow the model of \citep{Yokozawa-1995} describing the
-vertical distribution of leaf area within the crowns of individual
-plants. This model can account for a variety of canopy profiles through
-a single parameter $\eta$. Setting $\eta=1$ results in a conical
-canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
-, give a top-weighted canopy profile similar to those seen among
-angiosperms. Let $S(z,h)$ be the sapwood area at height $z$ for a
-plant with top height $h$. Following Yokozawa and Hara (1995) we
-assume a relationship between $S(z,h)$ and height such that
-
-\begin{equation}\label{eq:crown1}
-\frac{S(z,h)}{S(0,h)}= \big(1-\big(\frac{z}{h}\big)^\eta\big)^2.
-\end{equation}
-
-We also assume that each unit of sapwood area supports a fixed area of
-leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
-canopy area of a plant relates to basal sapwood area $S(0,h)$:
-
-\begin{equation}\label{eq:crown2}
-\frac{M_{\rm l}}{\phi}= \theta \, S(0,h).
-\end{equation}
-
-Integrating $S(z,h)$ gives a solution for the total mass of sapwood in
-the plant:
-
-\begin{equation}\label{eq:ms1}
-M_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, {\rm d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}
-
-where
-$\eta_c=1-\frac{2}{1+\eta} + \frac{1}{1+2\eta}$ \citep{Yokozawa-1995}.
-Substituting from eq. \ref{eq:crown2} into eq. \ref{eq:ms1} gives an
-expression for sapwood mass as a function leaf area and height:
-
-\begin{equation}\label{eq:ms2}
-M_\textrm{s}=\rho \, \eta_c \, \theta \, A_{\rm l} \, h.
-\end{equation}
-
-\subsection{Bark mass}\label{bark-mass}
-
-Bark and phloem tissue are modelled using an analogue of the pipe model,
-leading to a similar equation as that for sapwood mass (eq.
-\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
-assumed to be a constant fraction $b$ of sapwood area per unit leaf
-area such that
-
-\begin{equation}\label{eq:mb}
-M_{\rm b}=b M_\textrm{s}.
-\end{equation}
-
-\subsection{Root mass}\label{root-mass}
-
-Also consistent with pipe-model assumption, we assume a fixed ratio of
-root mass per unit leaf area
-
-\begin{equation}\label{eq:mr}
-M_{\rm r}=\alpha_3 \, A_{\rm l}.
-\end{equation}
-
-Even though nitrogen and water uptake are not modelled explicitly,
-imposing a fixed ratio of root mass to leaf area ensures that
-approximate costs of root production are included in calculations of
-carbon budget.
-
-\section{Effect of traits on height growth} \label{app:traits_max}
+\section{Appendix B - Effect of traits on height growth} \label{app:traits-RGR}
 
@@
 $\frac{{\rm d}H} {{\rm d}A_{\rm l}}$. Thus we have
-
 \begin{equation} \label{eq:G2}
@@
 \end{equation}
-
 where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eq.
@@
 the derivative of $G$ with respect to some trait $x$  is given by
-
 \begin{equation} \label{eq:dG}
@@
 \end{equation}
-
 If this derivative is positive (negative), there will be a positive (negative) relationship
@@
 
-We further simplify by expanding the derivative of eq. \ref{eq:daldmt}:
-
+We can further simplify by expanding the derivative of eq. \ref{eq:daldmt}:
 \begin{equation} \label{eq:dA_dx}
@@
 \end{equation}
-
 Combing eqs. \ref{eq:dG}-\ref{eq:dA_dx}, we have
-
 \begin{equation} \label{eq:dG2}
@@
 {\partial G} {\partial x} < 0$ if
-
 \begin{equation}\label{eq:dg3}
@@
 \end{equation}
-
 or equivalently
-\begin{equation}\label{eq:dg4}
+\begin{equation}\label{eq:dg4_orig}
 \frac{
@@
 
-In words, there will be a negative relationship between trait x and growth if the
-relative change in production due to change in traits is less than the relative change
-in the total construction cost for a unit of leaf area due to change in trait. Crucially,
-key elements in eq. \ref{eq:dg3} are modified by height.
+\section{Appendix C - Linking traits to whole-plant light compensation point} \label{app:traits-RGR}
 
-We can also solve for trait values $x$ maximising growth rate, i.e. the value
-giving $\partial G /\partial x = 0$:
+\footnotesize
 
-\begin{equation}\label{eq:max}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-=
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
+\bibliography{references}
 
-\subsection{Leaf construction cost}
-
-If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand
-eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm
-{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
-
-\begin{equation}\label{eq:dbdt2}
-\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
-\end{equation}
-
-where
-$c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$
-is independent of $\phi$, and thus
-
-\begin{equation}\label{eq:dbdt3}
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
-(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
-\end{equation}
-
-Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent
-of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
-{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
-
-Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
-with traits if
-
-\begin{equation} \label{eq:G6}
-\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
-< \frac{1}{\phi
- + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
- {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
-\end{equation}
-
-\subsection{Wood construction cost}
-
-Do similar to previous section for wood construction cost.
-
-\end{appendices}
 \end{document}

### 2016-07-29 sensitivity.R

@@ -1,89 +1,0 @@
-# Fitting TREE to observed data
-
-source_dir <- function(path, ...) {
-  for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
-    source(file.path(path, nm), ...)
-  }
-}
-source_dir("R")
-
-library(plant)
-suppressPackageStartupMessages({
-  library(plyr)
-  library(dplyr)
-})
-
-# First, look at plot: challenge is to move peak to left and pull down elevation of curve.
-data <- remake::make("BCI_species_data") %>%
-  mutate(dbh=at) %>%
-  filter(!is.na(lma*rho*hmat*dbh*diameter_stem_dt), dbh %in% c(0.01), hmat > 10) %>%
-  select(dbh, lma, rho, diameter_stem_dt)
-
-d <- trait_effects_data("lma", "diameter")
-d <- d[d$size_class == 2L,]
-
-plot(data$lma, data$diameter_stem_dt, las=1, log="x",
-     xlab=name_pretty("lma"), ylab=name_pretty("diameter_stem_dt"),
-     xlim=range(data$lma, d$lma),
-     ylim=range(data$diameter_stem_dt, d$diameter_stem_dt))
-lines(d$lma, d$diameter_stem_dt, col="red")
-
-# given a dbh x, traits and parameters, grows plant to right size and estimates growth rate
-growth <- function(x, trait, dbh, E=1,
-                   strategy=default_strategy()) {
-  strategy[trait] <- x
-  pp <- grow_plant_to_size(Plant(strategy), dbh, "diameter",
-                           fixed_environment(E))
-  sapply(pp$plant, function(p) p$vars_growth[["diameter_stem_dt"]])
-}
-
-# finds trait value in range that maximises growth rate at given size and light
-maximise_growth_rate_by_trait <- function(strategy,
-                                          trait="lma",
-                                          range=c(0.001, 1),
-                                          dbh=0.01, E=1) {
-  f <- function(x) {
-    growth(x, trait, dbh, E, strategy)
-  }
-  ret <- optimise(f, range, maximum=TRUE, tol=1e-4)
-  list(optimal_trait=ret$maximum,
-       growth_rate=ret$objective)
-}
-
-# finds trait value in range that maximises growth rate at given size
-# and light
[…]

### 2016-08-04 ms/figure_conceptual.tex
Writing introduction and model

@@
 \documentclass{beamer}
-\usepackage[size=custom, width=21, height=27, scale=1.0, orientation=portrait]{beamerposter}
+\usepackage[size=custom, width=21, height=28, scale=1.0, orientation=portrait]{beamerposter}
 \setbeamertemplate{navigation symbols}{}
@@
 
-\newcommand{\rowspace}{3.2}
+\newcommand{\rowspace}{3}
 \newcommand{\colspace}{3.5}
@@
   \node[anchor=center, rectangle]
-    at (#1) {\includegraphics[height=1.65cm]{#2}};
+    at (#1) {\includegraphics[height=1.85cm]{#2}};
 }
@@
 \newcommand{\defineRow}[2]{
-  \node at ($(top) + (-9.5cm, -#1*\rowspace cm) + (0cm, -1.5cm)$) (#2){};
+  \node at ($(top) + (-9.5cm, -#1*\rowspace cm)$) (#2){};
 }
@@
 \newcommand{\rowtitle}[1]{
-  \node at ($(thisrow) + (0.25cm, 1.5cm)$) [anchor=west, text width=8cm, text height=0.5cm]{
-  {\footnotesize #1}};
+  \node at ($(thisrow) + (0.25cm, 1.5cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
+  {\bf \footnotesize #1}};
 }
@@
   \node at ($(thisrow) + (-0.75cm, -0.3cm)$)[anchor=west, text width=2cm] {
-  \footnotesize {\color{orange!50}\begin{equation}\label{#1}\end{equation}}};
+ {\color{orange!50}\begin{equation}\label{#1}\end{equation}}};
 }
@@
 
-\node at ($(current page.south)+ (0cm, 20cm)$) (top){};
+\node at ($(current page.north)+ (0cm, -7.5cm)$) (top){};
 
-\node[rectangle, anchor =south] at ($(top)+ (-5cm, 2cm)$){
-  \includegraphics[height= 2cm]{downloads/seedling.png}};
-  \textbox{$(top)+ (-5cm, 6.2cm)$}{Seedling}
+\node[rectangle, anchor = center] at ($(top)+ (0cm, 4cm)$){
+  \includegraphics[width=18cm]{output/main/mass_fraction.pdf}};
 
-\node[rectangle, anchor =south] at ($(top)+ (5cm, 0.5cm)$){
-  \includegraphics[height= 5cm]{downloads/tree.png}};
-  \textbox{$(top)+ (5cm, 6.25cm)$}{Adult}
+\node at ($(top) + (-9.25cm, 6.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
+  {\bf \footnotesize Standing mass}};
+
+\node at ($(top) + (-10.5cm, 6.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
+  {\bf \footnotesize a)}};
+
+\node at ($(top) + (-10.5cm, 1.5cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
+  {\bf \footnotesize b)}};
+
+% \node[rectangle, anchor =center] at ($(top)+ (6cm, 4cm)$){
+%   \includegraphics[height= 3cm]{downloads/tree.png}};
+%   \textbox{$(top)+ (8.5cm, 4cm)$}{Adult}
+% \node[rectangle, anchor =center] at ($(top)+ (-6cm, 4cm)$){
+%   \includegraphics[height= 1cm]{downloads/seedling.png}};
+%   \textbox{$(top)+ (-8cm, 4cm)$}{Seedling}
 
 \defineRow{0}{thisrow}
-\rowtitle{Biomass production rate}
+\rowtitle{Biomass production}
 \firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/net_mass_production_dt_net_mass_production_dt}{}
-\othercol{2}{$\alpha_y$}{output/con/net_mass_production_dt_yield}{$\times$}
-\othercol{3}{$\Bigg({\rm P}$}{output/con/net_mass_production_dt_assimilation}{$-$}
-\othercol{4}{${\rm R}$}{output/con/net_mass_production_dt_respiration}{$\Bigg)$}
-\othercol{5}{$-{\rm T}$}{output/con/net_mass_production_dt_turnover}{}
+\othercol{2}{\scriptsize $\quad y$}{output/con/net_mass_production_dt_yield}{}
+\othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/con/net_mass_production_dt_assimilation}{}
+\othercol{4}{\hspace{-1em}-$\Sigma$\scriptsize $M_{\rm i} r_{\rm i}$}{output/con/net_mass_production_dt_respiration}{\hspace{-0.5em}$\Bigg)$}
+\othercol{5}{\hspace{-0.75em}-$\Sigma$\scriptsize $M_{\rm i} k_{\rm i}$}{output/con/net_mass_production_dt_turnover}{}
 \rowlabel{eq:dbdt}
@@
 \defineRow{1}{thisrow}
-\rowtitle{Leaf deployment per mass}
-\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm t}}}$}{output/con/darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
+\rowtitle{Mass growth}
+\firstcol{1}{$\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$}{output/con/mass_total_dt_mass_total_dt}{}
+\othercol{2}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/con/mass_total_dt_fraction_allocation_growth}{$\times$}
+\othercol{3}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/mass_total_dt_net_mass_production_dt}{$+$}
+\othercol{4}{$\frac{{\rm d}M_{{\rm h}}}{{\rm d}t}$}{output/con/mass_total_dt_mass_heartwood_dt}{}
+\rowlabel{eq:dMtdt}
+
+\defineRow{2}{thisrow}
+\rowtitle{Leaf deployment per mass invested}
+\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/con/darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
 \othercol{2}{$\frac{{\rm d}M_{{\rm l}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
-\othercol{3}{$\frac{{\rm d}M_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{+}
+\othercol{3}{$\frac{{\rm d}M_{{\rm r}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_root_darea_leaf}{+}
 \othercol{4}{$\frac{{\rm d}M_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
-\othercol{5}{$\frac{{\rm d}M_{{\rm r}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_root_darea_leaf}{$\Bigg)^{-1}$}
+\othercol{5}{$\frac{{\rm d}M_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{$\Bigg)^{-1}$}
  \rowlabel{eq:daldmt}
-\defineRow{2}{thisrow}
-\rowtitle{Leaf-area growth rate}
+
+\defineRow{3}{thisrow}
+\rowtitle{Leaf-area growth}
 \firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
-\othercol{2}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm t}}}$}{output/con/area_leaf_dt_darea_leaf_dmass_live}{$\times$}
-\othercol{3}{$\frac{{\rm d}M_{{\rm t}}}{{\rm d}B}$}{output/con/area_leaf_dt_fraction_allocation_growth}{$\times$}
+\othercol{2}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/con/area_leaf_dt_darea_leaf_dmass_live}{$\times$}
+\othercol{3}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/con/area_leaf_dt_fraction_allocation_growth}{$\times$}
 \othercol{4}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/area_leaf_dt_net_mass_production_dt}{}
@@
 
-\defineRow{3}{thisrow}
-\rowtitle{Height growth rate}
+\defineRow{4}{thisrow}
+\rowtitle{Height growth}
 \firstcol{1}{$\frac{{\rm d}H}{{\rm d}t}$}{output/con/height_dt_height_dt}{}
@@
 
-\defineRow{4}{thisrow}
-\rowtitle{Stem-area growth rate}
+\defineRow{5}{thisrow}
+\rowtitle{Stem-area growth}
 \firstcol{1}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/con/area_stem_dt_area_stem_dt}{$\Bigg($}
-\othercol{2}{$\frac{{\rm d}A_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_sapwood_darea_leaf}{$+$}
-\othercol{3}{$\frac{{\rm d}A_{\textrm{sb}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_bark_darea_leaf}{\hspace{-0.3em}$\Bigg)$}
+\othercol{2}{$\frac{{\rm d}A_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_bark_darea_leaf}{$+$}
+\othercol{3}{$\frac{{\rm d}A_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_sapwood_darea_leaf}{\hspace{-0.3em}$\Bigg)$}
 \othercol{4}{\hspace{-0.4em}$\times\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_stem_dt_area_leaf_dt}{$+$}
@@
 
-\defineRow{5}{thisrow}
-\rowtitle{Stem-diameter growth rate}
+\defineRow{6}{thisrow}
+\rowtitle{Stem-diameter growth}
 \firstcol{1}{$\frac{{\rm d}D}{{\rm d}t}$}{output/con/diameter_stem_dt_diameter_stem_dt}{}
@@
 
-\textbox{$(current page.south)+ (0cm, 0.5cm)$}{Plant height (m)}
+\textbox{$(top)+ (0cm, -20cm)$}{{\bf Plant height (m)}}
 
@@
 
-\defineRow{0}{rowOne}
-\defineCol{rowOne}{\colNum}{A}
+\defineRow{0}{rowZero}
+\defineCol{rowZero}{\colNum}{A}
 
-\defineRow{1}{rowTwo}
+\defineRow{1}{rowOne}
+\defineCol{rowOne}{\colNum}{B1}
+
+\defineRow{2}{rowTwo}
 \defineCol{rowTwo}{\colNum}{B}
 
-\defineRow{2}{rowThree}
+\defineRow{3}{rowThree}
 \defineCol{rowThree}{\colNum}{C}
 
-\defineRow{3}{rowFour}
+\defineRow{4}{rowFour}
 \defineCol{rowFour}{\colNum}{D}
 
-\defineRow{4}{rowFive}
+\defineRow{5}{rowFive}
 \defineCol{rowFive}{\colNum}{E}
 
-\defineRow{5}{rowSix}
+\defineRow{6}{rowSix}
 \defineCol{rowSix}{\colNum}{F}
@@
 
-\draw [s1] (A) to [bend right, looseness=1] (C);
+\draw [s1] (A) to [bend right, looseness=0.75] (C);
+\draw [s1] (A) to [bend right] (B1);
 \draw [s1] (B) to [bend right] (C);
-\draw [s1] (C) to [bend right, looseness=1] (D);
+\draw [s1] (C) to [bend right, looseness=0.75] (D);
 \draw [s1] (C) to [bend right, looseness=1] (E);

### 2016-08-04 ms/ms.tex
Writing introduction and model

@@
 \floatpagestyle{empty}
+
+% Enable new column type -- centered with fixed width
+% Following http://texblog.org/2008/05/07/fwd-equal-cell-width-right-and-centre-aligned-content/
+\usepackage{array}
+\newcolumntype{x}[1]{%
+>{\raggedleft}p{#1}}%
 
@@
 
-% Enable cross referencing to Evidence document
+% Enable cross referencing to external documents
 \usepackage{xr}
@@
 \externaldocument{figure_conceptual}
+% Needed to set equation counter, following inclusion of figure 1
+\setcounter{equation}{7}
 
+% Input common definitions from external file
 \input{ms-common-defs}
@@
 
-\keywords{growth strategy, life-history traits, maximum height, ontogenetic stage, photosynthetic rate, plant development, seed mass, specific leaf area, wood density}
-
-%appendix
-\usepackage[title,titletoc,toc]{appendix}
-\usepackage{setspace}  % for double spacing
+\keywords{growth strategy, life-history,  ontogeny, size, leaf mass per area, wood density, shade-tolerance, growth rate}
 
@@
 % \linenumbers
-\section{Summary}
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item We ...
-\end{enumerate}
 
-% Consider Ecol Letters I deas and Perspectives: maximum of 7500 words (main text), and 10 figures, tables or boxes.
+\tableofcontents
+% Consider Ecol Letters "Ideas and Perspectives": maximum of 7500 words (main text), and 10 figures, tables or boxes.
 
-\section{To do}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item New refs: \citep{Ishihara-2016, Duursma-2016, Gibert-2016, Falster-2016, Kunstler-2015,Falster-2015b, Falster-2015a,Wenk-2015,Li-2014}
-\item Table near intro summarising empirical patterns to be explained by integrated model?
-\end{enumerate}
+\section{Abstract}
+We present a new theoretical framework for understanding the effect of diverse functional traits on plant growth.
 
@@
 
-Functional traits capture core differences in the strategies
-plants use to generate and invest surplus energy
- \citep{Wright-2004, Chave-2009, Westoby-2002}.
-Although most plants have the same basic physiological function
-and key resource requirements (carbon, nitrogen and water), species differ considerably in
-rates at which resources are acquired and invested into different tissues.
-Much of this functional variation relates to underlying
-differences in some prominent traits, including leaf
-mass per area (LMA), wood density (WD), leaf nitrogen per area, vessel
-diameter,  maximum  height (HMAX) and seed size  \citep{Wright-2004, Chave-2009}.
-Data for these traits now exists for up to 10\% of the world's 250000 plant species
-\citep{Cornwell-2014}. Increasingly, researchers have been looking to traits
-as potential indicators of the growth strategy and ecology of a species.
+Functional traits capture core differences in the strategies plants use to generate and invest surplus energy \citep{Wright-2004, Chave-2009, Westoby-2002}.
+Although most plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in rates at which resources are acquired and invested into different tissues.
+During the last decade, trade-offs related to prominent traits -- such as leaf mass per leaf area, wood density, leaf nitrogen per area, vessel diameter, maximum  height, and seed size -- have been quantified and trait values exist for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
+Increasingly, researchers have been using traits as potential indicators of the growth strategy and ecology of a species.
 
-While the influence of plant traits on elements of plant physiological
-function is increasingly understood, attempts at using traits to understand
-demographic rates have met with mixed success
- \citep{Wright-2010, Poorter-2008}. In seedlings, the trait LMA
--- the central element of the leaf economics spectrum
- \citep{Wright-2004} -- has been tightly related to relative growth
-rate in plant mass \citep{Lambers-1992, Wright-2000}. LMA and
-its correlate leaf lifespan have also been linked to height growth rate for
-small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These
-early successes prompted researchers to search for similar relationships in
-large plants. However, the results showed that in trees LMA was not correlated
-with diameter growth rate  \citep{Wright-2010, Poorter-2008,
-Herault-2011}. Meanwhile, other traits such as wood density showed
-strong relationships to growth in large plants \citep{Wright-2010},
-but only weak relationships in seedlings  \citep{Castro-1998}. Thus
-far, most trait growth relationships have been quantified for one or few
-traits and at a single life- stage. Yet the picture is emerging is that the
-link between traits and growth may be modified by plant size
- \citep{Ruger-2012}. This finding challenges a wide-spread
-assumption regarding traits: that particular ends of the trait spectrum translate directly into
-faster or slower growth (refs XX).
+While the influence of plant traits on elements of plant physiological function is increasingly understood, attempts at using traits to understand demographic rates however have only met with mixed success \citep{Wright-2010, Poorter-2008}.
+In seedlings, the trait leaf mass per leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- has been tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}.
+Leaf mass per leaf area and its correlate leaf lifespan have also been linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
+These early successes prompted researchers to search for similar relationships in large plants.
+However, the results showed that in trees leaf mass per leaf area was not correlated with diameter growth rate \citep{Poorter-2008, Wright-2010, Herault-2011}.
+Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010}, but only weak relationships in seedlings \citep{Castro-1998}.
+Yet, until recently, most trait growth relationships had been quantified for one or few traits and at a single life- stage.
 
-The challenge of interpreting diverse empirical results seeking to link
-traits to growth rate is made all the harder because we currently lack clear
-expectations on the expected signal. Generating such expectations is one of the
-primary role for theory \citep{Kokko-2007}. Current empirical
-results suggest the effect of traits on growth changes with size and possibly
-also light environment, yet current theory says little about how such
-relationships might come about. A widely-used model shows that for seedlings,
-mass-based relative growth rate is linearly and negatively related to LMA
- \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
- An extension of the model suggests a similar
-relationship may hold at larger sizes \citep{Enquist-2007}. But as
-described above, this prediction is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth
-are largely absent.
+The picture is now emerging that the effect of traits on plant growth may be  modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
+In a recent meta-analysis of more than  500 correlations recorded from 103 studies, \citet{Gibert-2016} showed that the sign of the correlation between some traits (such as leaf mass per leaf area and maximum height) and growth rate was modified by size, while for other traits (such as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
+Importantly, the direction of the correlations and their shifts with size, supported predictions made based on understanding of trait-based trade-offs and their interaction with plant function \citep{Gibert-2016}.
 
-A further problem with existing theory is that the effects of traits are all
-realised via influences on mass production \citep{Enquist-2007},
-whereas the effect of traits such as LMA, WD, and Hmax is via allocated
-to different tissues \citep{Falster-2011}. There are in fact two problems with theory
-focussed solely on mass production. The first is that measuring mass
-production is only practical for small plants that can be easily harvested.
-For larger plants, growth is more often measured as increment in either diameter
-or height. For example, the CTFS monitors diameter growth in 49 large permanent
-plots across forests worldwide \citep{Anderson-2015};
- forestry have likewise recorded growth in diameter across
-large tracts of the globe \citep{Purves-2008}. The second problem is
-that, as we outline further below, because mass production is only one
-component of plant growth; theory focussing solely on mass production will
- always be limited.
+The challenge of interpreting diverse empirical results seeking to link traits to growth rate has been all the harder because we currently lack clear expectations on the expected signal.
+Generating such expectations is one of the primary role for theory \citep{Kokko-2007}.
+Current empirical results suggest the effect of traits on growth changes with size and possibly also light environment, yet current theory says little about how such relationships might come about.
+A widely-used model shows that for seedlings, mass-based relative growth rate is linearly and negatively related to leaf mass per leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+An extension of the model suggests a similar relationship may hold at larger sizes \citep{Enquist-2007}.
+But as described above, this prediction is not supported by empirical results.
+Meanwhile, theoretical predictions on how other traits should influence growth are largely absent..
 
-Effect of traits on shade tolerance also porrly udnerstood. Empirical -- shade
+A further problem with existing theory is that the effects of traits are all realised via influences on mass production \citep{Enquist-2007}, whereas the effect of traits such as leaf mass per leaf area and wood density is via allocation to different tissues \citep{Falster-2011, Gibert-2016}.
+A second issue with theory focussed solely on mass production is that measuring mass production is only practical for small plants that can be easily harvested.
+For larger plants, growth is more often measured as increment in either diameter or height.
+For example, the Centre for Tropical Science monitors diameter growth in 49 large permanent plots across forests worldwide \citep{Anderson-2015}; forestry agencies have likewise recorded growth in diameter in 1000's of plots across the globe \citep{Purves-2008, Kunstler-2015}.
+But as we outline further below, mass production is only one component of plant growth; meaning theory focussing solely on mass production will always be limited.
 
-(reich 1992, Niinements 2010, Gower-1993 - both show higher LAI in high LMA / long LL stands)
-
-Ultimately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny.
+Ultimately, trait based understanding must b integrated within a more general model of plant growth, also able to capture patterns of growth through ontogeny.
 Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
@@
 All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of mainatining stems \citep{Givnish-1988}
+Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of maintaining stems \citep{Givnish-1988}, and also vary with traits \citep{Lusk-2006,Poorter-2006}.
+These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are in need of a comprehensive and integrated explanation.
 
-In this paper we provide a mechanistic framework for describing the potential
-effect of traits on growth throughout a plant's life-cycle, linking to
-commonly used metrics such as height and diameter growth. To help motivate our
-study we re-analyse demographic and trait data from an earlier study
- \citep{Wright-2010}, to show 1) that the link between traits and
-growth is stronger than previously thought, 2) that the influence of three
-prominent traits on growth rate is moderated by plant size, and 3) that the
-impact of size on the nature of relationships varies among traits
- \citep[see also]{Ruger-2012}. We then show how these results align with new
-set of theoretical predictions. Analysis extends ...
-Moreover, we show how traits moderate plant
-responses to light environment and also determine shade tolerance, supporting
-diverse empirical results \citep{Ruger-2012, Poorter-2006}. By
-disentangling the effects of plant size, light environment and traits on
-growth rates, we hope to provide a solid theoretical foundation for trait
-ecology and thus provide a platform for understanding growth across diverse
-species around the world.
+In this paper we show how a recently published mechanistic framework \plant{} \citep{Falster-2016} describing the potential effect of traits on growth throughout a plant's life-cycle can explain diverse phenomena (Table \ref{tab:phenomena}).
+The model employs the same framework for partitioning growth used by \citet{Gibert-2016} in their meta-analysis, building on and extending several related approaches \citep{Givnish-1988, Yokozawa-1995, Makela-1997, Moorcroft-2001, Falster-2011, King-2011} in which metrics such as height, diameter and mass growth are calculated as an emergent outcomes of several, independent, physiological processes.
+This approach is quite different to models derived from metabolic scaling theory \citep{Enquist-1999, Enquist-2007}, which attempt to organise everything around a single master ``scaling'' equation for mass growth, or statistically fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014}.
+By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.
 
-\section{Conceptual framework for understanding height, basal area and diameter growth}
+\section{Framework for understanding the effects of traits on growth}
 
-To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in
-seedlings to LMA \citep{Lambers-1992, Wright-2000}, to explicitly include influences of
-size, light environment, and other prominent traits. This model
+Our approach builds on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within the plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
+\ref{fig:conceptual}a).
+Here we consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, with the subscript $i$ indicating: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
+The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
+A full list of variables, units and definitions is given in Table \ref{tab:params}.
 
-Builds on 2011 paper - highlight what's new from there.
+We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant.
+Adopting a standard model, the amount of biomass available for growth, ${\rm d}B / {\rm d}t$ is given by the difference between income (total photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000, Falster-2011} (Fig.
+\ref{fig:conceptual}a, eqn \ref{eq:dbdt}).
+Photosynthesis is proportional to the average photosynthetic rate per unit leaf area $\bar{p}$ and total leaf area, $A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per leaf area.
+The constant $y$ in  eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass.
+Although many models include a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, net biomass production eventually comes down to simple subtraction of income and loss.
 
-First derive ... generally true.
+The only potentially controversial element of in eqn \ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocated to either new growth (i.e.
+growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$ or reproduction \citep{Thornley-2000}.
+This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction often entails run-down in the vegetative part of the plant (refs XX).
 
-Equations \ref{eq:dbdt} -\ref{eq:dDdt} shown in Fig. \ref{fig:conceptual}.
-\setcounter{equation}{6}
+\subsection{Decomposing of growth rate into components}
 
-Requires definitions from Table \ref{tab:definitions}.
+To model growth in either height, leaf area, basal area, standing mass or stem diameter requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
+Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig.
+\ref{fig:conceptual}b, eqns \ref{eq:dbdt}-\ref{eq:dDdt}).
+Importantly, most of these terms vary intrinsically with size.
+The insets in Fig.
+\ref{fig:conceptual}b show size-related patterns for a typical plant, from a specific allocation model derived below.
+However, it is important to recognise that eqns \ref{eq:dbdt}-\ref{eq:dDdt} are mathematically true, derived using the chain rule of differential calculus, and thus must hold true for any class of models where carbon allocation is important.
 
-(check this material covered in table: Here, $M, r$, and $k$ refer to the mass, respiration rate, and
-turnover rate of different tissues, denoted by subscripts $l$=leaves,
-$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
-rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
-assimilated carbon fixed in biomass (the remaining fraction being lost
-as growth respiration) (see Table \ref{tab:params} for units and
-definitions). Photosynthesis is proportional to leaf area,
-$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
-The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}.$
-)
+The first point to note is that the growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$.
+We assume the remaining fraction, $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$, is allocated to reproduction.
+The term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, presumably at 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants.
+It is important to note here that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
+So even a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ does not increase any further.
 
-We begin with a standard model for the amount of biomass available for
-growth, ${\rm d}B / {\rm d}t$, given by the difference between income
-(total photosynthesis) and losses (respiration and turnover) within the
-plant \citep{Makela-1997, Thornley-2000, Falster-2011}:
+The growth rate in total standing mass of the plant, $\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$ is then sum of sapwood turnover (which is replaced by an equivalent amount) and growth in live mass (eqn \ref{eq:dMtdt}).
 
-\begin{equation}\label{eq:dbdt_full}
-\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
-  = \underbrace{\strut y}_\textrm{yield}
-    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
-     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
-    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
-\end{equation}
+The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and support tissues of bark, sapwood and roots (eqn \ref{eq:daldmt}).
+The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of tissue-level construction costs per unit of leaf area, with one of these being $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$: the trait leaf mass per area.
 
-The only potentially controversial element of in eq.
-\ref{eq:dbdt} \ref{eq:dbdt_full} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants, because the .....
-However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegetative part of the plant (refs XX) \citep{ Thornley-2000}.
+The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
+This variable accounts for the architectural strategy of the plant.
+Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
+Meanwhile, basal area ($A_{\rm st}$) increment can be expressed as the sum of increments in sapwood, bark \& heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
+These in turn are related to ratios of sapwood or bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
+Finally, diameter growth is then given by the geometric relationship between stem diameter ($D$) and $A_{\rm st}$ (eqn \ref{eq:dDdt}).
 
-To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
-Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Figure \ref{fig:conceptual})
+\subsection{A functional-balance model for plant construction}
 
-The first term on the right of eq \ref{eq:dhdt},
-${\rm d}H / {\rm d}A_{\rm l}$, is the growth in plant height
-per unit growth in total leaf area -- accounting for the architectural
-strategy of the plant. Some species tend to leaf out more than grow
-tall, while other species emphasise vertical
-extension \citep{Poorter-2006}.
+As noted above, most of eqns in Fig.
+\ref{fig:conceptual} arise as mathematical truisms.
+To make predictions from eqns \ref{eq:dbdt}-\ref{eq:dDdt}, however, requires a specific model of plant construction.
+The model described below can be considered a first-order functional-balance model, in that it seeks to maintain a balance between various elements of plant construction through ontogeny.
+Table \ref{tab:allometry} provides key equations of the model, while Table \ref{tab:params} provides estimates on key variable and fitted parameters.
 
-The second term, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$,
-accounts for the marginal cost of deploying an additional unit of leaf
-area, including construction of the leaf itself and various support
-structures. As such, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$
-can itself be expressed as a sum of construction costs per unit leaf
-area:
+Our model is founded on four key assumptions:
 
-The third term in eq \ref{eq:dhdt},
-${\rm d}M_{\rm a} / {\rm d}P$, gives the fraction of net
-biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
-rather than reproduction or storage.
+\noindent{\bf Assumption  1} -- There is an allometric (log-log) scaling relationship between the accumulated leaf area of a plant and its height.
+Relationship represented by eqn \ref{eq:ha}.
+Empirically motivated.
+Describes architectural strategy of plant.
 
-In a similar way, basal area ($A_{\rm st}$) increment can be
-expressed as the sum of increments in sapwood, bark \& heartwood areas
-($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively):
-$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
-We now have an equation for basal area growth that contains many of the same
-elements as eq. \ref{eq:dhdt}, see eq. \ref{eq:dast}.
+\noindent{\bf Assumptions 2-4} -- An isometric relationship between the total leaf area and each of sapwood cross-sectional area (2), bark cross-sectional area (3), and total root mass (4).
+As per the pipe model, \citep{Shinozaki-1964} we assume that each unit of sapwood area supports a fixed area of leaf, so that the total canopy area of a plant relates to basal sapwood area eqn \ref{eq:as}.
+This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (eqn \ref{eq:ms2}).
+Bark and phloem tissue are modelled using an analogue of the pipe model (eqn \ref{eq:as}), leading to a similar equation as that for sapwood mass (eqn \ref{eq:msb}).
+Also consistent with pipe-model assumption, we assume a fixed ratio of root mass per unit leaf area (eqn \ref{eq:mr}).
 
-Diameter growth is then given by the geometric relationship between stem
-diameter ($D$) and $A_{\rm st}$, see eq. \ref{eq:dDdt}
+\noindent{\bf Assumption 5} -- Vertical distribution of leaf within the canopy of a plant remains relatively constant through ontogeny.
+We follow the model of \citep{Yokozawa-1995} describing the vertical distribution of leaf area within the crowns of individual plants.
+This model can account for a variety of canopy profiles through a single parameter $\eta$.
+Setting $\eta=1$ results in a conical canopy, as seen in many conifers, while higher values, e.g.
+$\eta=12$ , give a top-weighted canopy profile similar to those seen among angiosperms.
 
-\subsubsection{A functional-balance model for allocation}
+Assuming lead immediately to an allometric model for plant size (Table \ref{tab:allometry}, see Supplementary material for derivations).
 
-To address the first of these challenges, we use a functional balance
-model \citep{Yokozawa-1995, Falster-2011} assuming
-1) an allometric relation between total leaf area and plant height, and
-2) a constant ratio between total leaf area and each of sapwood
-cross-sectional area, bark cross-sectional area, and total root mass.
-These assumptions, which are well-supported by data (Fig.
-\ref{fig:assumptions}), lead immediately to an allometric model for plant
-size (Table \ref{tab:allometry}, see Supplementary material for
-derivations). Substituting from Table \ref{tab:allometry} into eqs.
-\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
-for plant growth in relation to size.
+To first approximation, the assumptions, which are well-supported by data (Fig. \ref{fig:assumptions}).
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  four key assumptions.
-\item
-  simple linear functions non-critical for results of this paper,
-\item
-  explore using an example model, but any model with basic size effects
-  will display same behaviour
+Simple linear functions non-critical for results of this paper, explore using an example model, but any model with basic size effects will display same behaviour.
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    details about our model ()
-  \item
-    show argument hold intuitively
-  \end{itemize}
-\end{itemize}
+Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model for plant growth in relation to size.
 
-\section{Derivation of a simple allometric model of plant function}
+\subsection{Trait-based trade-offs}
 
-Here we describe an allometric model linking the various size dimensions
-of a plant required by most ecologically realistic vegetation models
-(i.e. =mass of leaves, mass of sapwood, mass of bark, mass of fine
-roots) to a plant height. Table \ref{tab:allometry} provides a summary
-of the derived model, while Table \ref{tab:params} provdies estimates on
-key parameters.
+\subsection{Parameters}
 
-Based on empirically observed allometries (see main text), we assume an
-allometric log-log scaling relationship between the accumulated leaf
-area of a plant and its height: Eq. \ref{eq:ha}.
+\citep{Falster-2015b}: BAAD
+\citep{Falster-2016}: \plant
 
-Note, scaling relationship reversed from \citep{Falster-2011}.
-
-We follow the model of \citep{Yokozawa-1995} describing the
-vertical distribution of leaf area within the crowns of individual
-plants. This model can account for a variety of canopy profiles through
-a single parameter $\eta$. Setting $\eta=1$ results in a conical
-canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
-, give a top-weighted canopy profile similar to those seen among
-angiosperms. We also assume that each unit of sapwood area supports a fixed area of
-leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
-canopy area of a plant relates to basal sapwood area. This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (Eq. \ref{eq:ms2}).
-
-Bark and phloem tissue are modelled using an analogue of the pipe model,
-leading to a similar equation as that for sapwood mass (eq.
-\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
-assumed to be a constant fraction $b$ of sapwood area per unit leaf
-area such that (Eq. \ref{eq:mb}).
-
-Also consistent with pipe-model assumption, we assume a fixed ratio of
-root mass per unit leaf area (Eq. \ref{eq:mr}).
-
-Even though nitrogen and water uptake are not modelled explicitly,
-imposing a fixed ratio of root mass to leaf area ensures that
-approximate costs of root production are included in calculations of
-carbon budget.
-
-\subsubsection{Embedding trait-related trade-offs}
-
-use table with arrows to
-
-By definition, the main effect of increasing $\phi$ is to decrease the
-marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
-growth rate. However, to fully account for the influence of $\phi$ on
-growth we must also account for the decrease in leaf turnover that
-results from superior leaf construction \citep{Wright-2004}.
-This is achieved by linking the turnover rate of leaf in eq.
-\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
-the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
-shown to describe patterns across diverse
-ecosystems \citep{Wright-2004} (Fig. \ref{fS-leaf}).
-
-As for leaf, cheaper stem construction (lower $\rho$) decreases the
-cost of deploying a unit of leaf area, and may thereby increase growth
-rates. However, the cost of cheap stem construction must also be
-accounted for. However, the direct physiological trade-offs of lower
-$\rho$ are less understood than for $\phi$. One possibility is that
-lower $\rho$ decreases mechanical strength resulting in higher
-mortality \citep{Chave-2009, Wright-2010}. Under
-this scenario, decreasing $\rho$ will always provide a growth
-advantage, because the costs of low $\rho$ are not realised within the
-terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
-\ref{tab:trade-offs}). In addition, lower $\rho$ may increase turnover
-of sapwood. Mirroring our assumptions for leaves, we let sapwood
-turnover decrease with $\rho$,
-i.e.$k_{\rm s}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
-
-Unlike leaf and stem traits, height at maturation ($h_m$) moderates
-growth by adjusting the amount of energy invested in growth, i.e.~the
-term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqs.
-\ref{eq:dhdt} and \ref{eq:dast}.
-
-\subsection{Analysis}
-
-\newpage
-\section{Results and Discussion}
-
-Section 3.3 of King (2005) shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
-
-King 1999 also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
-Basically, the much greater path length from roots to leaves in trees vs.
-seedlings results in a much greater increment in support cost per increment in crown area for trees.
-This explanation is treated by the growth model of King (1999), which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
+\section{Analysis}
 
@@
 
-This new growth model captures the intrinsically size-dependent nature
-of plant growth, recovering well-known
-empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern
-with size, first increasing then decreasing. This pattern results from
-systematic changes in the four components of eq. \ref{eq:dhdt} with size
-(Fig. \ref{fig:conceptual}), including a strong decline in the fraction of plant
-that is leaf declines with increasing size (Fig. \ref{fig:mass_fraction}),
-and increasing reproductive allocation. In contrast, basal-area growth
-continues to increase with size \citep{Sillett-2010, Stephenson-2014},
-due to an increasing influence of stem
-turnover. Diameter growth shows a weakly hump shaped
-curve \citep{Herault-2011}, tapering off slightly at larger
-sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
-mainly because of increased reproductive allocation in older trees
-(Fig?). All growth measures decrease sharply with size when expressed as
-relative growth rates \citep{Iida-2014}.
+This new growth model captures the intrinsically size-dependent nature of plant growth, recovering well-known empirical patterns \citep{Sillett-2010, King-2011}.
+Height growth shows a hump-shaped pattern with size, first increasing then decreasing.
+This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size (Fig.
+\ref{fig:conceptual}), including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig.
+\ref{fig:conceptual}), and increasing reproductive allocation.
+In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover.
+Diameter growth shows a weakly hump shaped curve \citep{Herault-2011}, tapering off slightly at larger sizes, in part because of allometric effects (eqn \ref{eq:dDdt}, but mainly because of increased reproductive allocation in older trees (Fig?).
+All growth measures decrease sharply with size when expressed as relative growth rates \citep{Iida-2014}.
 
@@
 
+Distinguish 3 types of trait.
+
+Growth traits
+
 For growth traits can derive rule:
 \begin{equation}\label{eq:dg4}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-<
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
+\frac{\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}} {\frac{ {\rm d}B} {{\rm d}t}} < \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
+\end{equation}
+In words, there will be a negative relationship between trait x and growth if the relative change in production due to change in traits is less than the relative change in the total construction cost for a unit of leaf area due to change in trait.
+Crucially, key elements in eqn \ref{eq:dg3} are modified by height.
+
+We can also solve for trait values $x$ maximising growth rate, i.e. the value giving $\partial G /\partial x = 0$:
+\begin{equation}\label{eq:max}
+\frac{\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} } {\frac{ {\rm d}B} {{\rm d}t}} = \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
 \end{equation}
 
-In words, there will be a negative relationship between trait x and growth if the
-relative change in production due to change in traits is less than the relative change
-in the total construction cost for a unit of leaf area due to change in trait. Crucially,
-key elements in eq. \ref{eq:dg3} are modified by height.
+\subsubsection{Seed size, $M_{\rm t,0}$}
 
-We can also solve for trait values $x$ maximising growth rate, i.e. the value
-giving $\partial G /\partial x = 0$:
-\begin{equation}\label{eq:max}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-=
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
+Seed size: increases starting size
 
-\subsubsection{LMA}
+\subsubsection{Height at maturation, $H_{\rm m}$}
 
-Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
-\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
-changes with plant size ( \ref{fig:lmA_growth_size}).
-Decreasing $\phi$ has two impacts on growth rate. First, lower
-$\phi$ increases marginal leaf deployment
-(${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on
-construction costs. Second, lower $\phi$ decreases net production
-(${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether
-lower $\phi$ increases growth thus depends on the relative magnitude
-of these two effects. When plants are small the effect on leaf
-deployment rate is bigger and so decreasing $\phi$ increases growth
-rate. When plants are large, the influence of $\phi$ on leaf
-deployment rate is diminished, because the costs of building other
-supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
-\ref{fig:mass_fraction}). The net result is that at larger sizes, low
-$\phi$ is no longer advantageous for growth (Fig.
+Height at maturation: ($h_m$) moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
+
+Greater height at maturation ($h_m$) can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ .
+At smaller sizes, there is no differentiation among species, because most species are focusing on growth.
+At larger sizes, individuals of some species are allocating a majority of their surplus energy to growth, leaving a strong signal in potential growth rate.
+
+\citet{Wenk-2015}
+
+\subsubsection{Leaf mass per leaf area, $\phi$}
+
+By definition, the main effect of increasing $\phi$ is to decrease the marginal leaf deployment per mass (eqn \ref{eq:daldmt}) and thus reduce growth rate.
+However, to fully account for the influence of $\phi$ on growth we must also account for the decrease in leaf turnover that results from superior leaf construction \citep{Wright-2004}.
+This is achieved by linking the turnover rate of leaf in eqn \ref{eq:dbdt} to $\phi$.
+We use an allometric scaling relationship of the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been shown to describe patterns across diverse ecosystems \citep{Wright-2004} (Fig.
+\ref{fS-leaf}).
+
+Unlike eqn \ref{eq:RGR}, eqs.
+\ref{eq:dhdt}, \ref{eq:dast} and \ref{eq:dDdt} predict a relationship between $\phi$ and growth that changes with plant size ( \ref{fig:lmA_growth_size}).
+Decreasing $\phi$ has two impacts on growth rate.
+First, lower $\phi$ increases marginal leaf deployment (${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on construction costs.
+Second, lower $\phi$ decreases net production (${\rm d}P / {\rm d}t$), due to increased leaf turnover.
+Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects.
+When plants are small the effect on leaf deployment rate is bigger and so decreasing $\phi$ increases growth rate.
+When plants are large, the influence of $\phi$ on leaf deployment rate is diminished, because the costs of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) are larger (Fig.
+\ref{fig:conceptual}).
+The net result is that at larger sizes, low $\phi$ is no longer advantageous for growth (Fig.
 \ref{fig:lmA_growth_size}d).
 
-If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand
-eq. \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm
-{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eq. \ref{eq:dbdt} we obtain,
-
+If the trait of interest is leaf-construction cost, i.e.
+$x=\phi$, we can expand eqn \ref{eq:dg4} as follows.
+Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$.
+Also, from eqn \ref{eq:dbdt} we obtain,
 \begin{equation}\label{eq:dbdt2}
@@
 \end{equation}
-
-where
-$c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$
-is independent of $\phi$, and thus
-
+where $c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$is independent of $\phi$, and thus
 \begin{equation}\label{eq:dbdt3}
@@
 \end{equation}
+eqn \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent of $\phi$.
+In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
 
-Eq. \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent
-of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm
-{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
-
-Substituting into eq. \ref{eq:dg4}, we have the result that growth rate correlates
-with traits if
-
+Substituting into eqn \ref{eq:dg4}, we have the result that growth rate correlates negatively with traits if
 \begin{equation} \label{eq:G6}
@@
 
-\subsubsection{Wood density}
+\subsubsection{Leaf nitrogen content per leaf area, $\nu$}
 
-Under this scenario, decreasing $\rho$ provides an advantage in height
-growth only when the benefits of low $\rho$ outweigh the costs. This
-effect is strongest at intermediate sizes. The decline in height growth
-at low $\rho$ with size is more marked than for either stem basal-area
-and diameter growth, because high sapwood turnover also contributes to
-this growth measure. Thus diameter growth continues, even as net
-production approaches zero.
+\subsubsection{Wood density, $\rho$}
 
-\subsubsection{hmat}
+As for leaf, cheaper stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates.
+However, the cost of cheap stem construction must also be accounted for.
+However, the direct physiological trade-offs of lower $\rho$ are less understood than for $\phi$.
+One possibility is that lower $\rho$ decreases mechanical strength resulting in higher mortality \citep{Chave-2009, Wright-2010}.
+Under this scenario, decreasing $\rho$ will always provide a growth advantage, because the costs of low $\rho$ are not realised within the terms of eqs.
+\ref{eq:dhdt} and \ref{eq:dast} (Table \ref{tab:trade-offs}).
+In addition, lower $\rho$ may increase turnover of sapwood.
+Mirroring our assumptions for leaves, we let sapwood turnover decrease with $\rho$, i.e.$k_{\rm s}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
 
- Greater height at maturation ($h_m$)
-can thus lead to a growth advantage by increasing
-$\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ . At smaller sizes, there is no differentiation
-among species, because most species are focusing on growth. At larger sizes, individuals of some species are
-allocating a majority of their surplus energy to growth, leaving a
-strong signal in potential growth rate.
+Under this scenario, decreasing $\rho$ provides an advantage in height growth only when the benefits of low $\rho$ outweigh the costs.
+This effect is strongest at intermediate sizes.
+The decline in height growth at low $\rho$ with size is more marked than for either stem basal-area and diameter growth, because high sapwood turnover also contributes to this growth measure.
+Thus diameter growth continues, even as net production approaches zero.
 
-\subsubsection{Responsiveness to light}
+\subsubsection{Sapwood area per leaf area, $\theta$}
 
-Two other important predictions arise directly from linking traits to
-growth rate. First, traits moderate the
-responsiveness of growth to changes in light environment. This response
-arises simply because species with higher potential growth rate have
-greater potential plasticity. Recent analyses have shown that species
-with low $rho$ increase growth more substantially with incraeses in
-light \citep{Ruger-2012}. This is exactly the behaviour
-exhibited in our model (Fig. \ref{fig:growth_light_dia}), at least for small
-to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
-of low $\phi$ are lost (Fig. \ref{fig:growth_light_dia}). This results also
-agrees with empirical results showing that growth for species with low
-$rho$ actually decreases as size
-increases \citep{Ruger-2012}. In contrast, growth rate tends
-to increase in size for the species with highest
-$\rho$ \citep{Ruger-2012}.
+\subsubsection{Vessel size, $\kappa$}
 
-Variation in $\phi$ also moderates the response of growth to changes
-in light, with species having the lowest $rho$ being most responsive.
-However, unlike for $\rho$, the effect appears only for the smallest
-size classes. Accordingly, empirical analyses have detected such a link
-in small plants but not for plants \textgreater{} 0.01m
-dbh \citep{Ruger-2012}.
+\subsubsection{Root mass per leaf area, $\alpha_{\rm r1}$}
 
-\subsection{Growth is more than just photosynthesis}
+\subsection{Responsiveness of growth rate to light}
 
-Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
-\citep{Lambers-1992, Cornelissen-1996, Wright-2000, Enquist-2007},
-${\rm d}P / {\rm d}t / M_{\rm a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
-(see Appendix \ref{app:traits-RGR}).
+Two other important predictions arise directly from linking traits to growth rate.
+First, traits moderate the responsiveness of growth to changes in light environment.
+This response arises simply because species with higher potential growth rate have greater potential plasticity.
+Recent analyses have shown that species with low $\rho$ increase growth more substantially with increases in light \citep{Ruger-2012}.
+This is exactly the behaviour exhibited in our model (Fig.
+\ref{fig:growth_light_dia}), at least for small to intermediate sizes.
+At larger sizes (D=0.2m) however, the advantages of low $\phi$ are lost (Fig.
+\ref{fig:growth_light_dia}).
+This results also agrees with empirical results showing that growth for species with low $\rho$ actually decreases as size increases \citep{Ruger-2012}.
+In contrast, growth rate tends to increase in size for the species with highest $\rho$ \citep{Ruger-2012}.
 
-Recently realisation that allocation and turnover major areas of
-uncertainty
+Variation in $\phi$ also moderates the response of growth to changes in light, with species having the lowest $\rho$ being most responsive.
+However, unlike for $\rho$, the effect appears only for the smallest size classes.
+Accordingly, empirical analyses have detected such a link in small plants but not for plants \textgreater{} 0.01m dbh \citep{Ruger-2012}.
 
-Model reconciles idea that traits can have negative impact on carbon
-budget, but still have positive influence on growth rates
+\subsection{Shade tolerance}
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Enquist 2007: sla and wd increase mass-production, where as in our
-  theory effects on mass production are negative. (also contradicts
-  earlier model from Enquist 1999, where argued wood density had no
-  effect on mass production)
-\end{itemize}
+Define as \textsc{wplcp}
 
-Assumes low lma and wd used to reduce costs of building leaf and stem,
-ie. total leaf area and stem cross section remains same. In this way
-differs to assumptions of recent work by \citep{Anten-2010}, \citep{Larjavaara-2010}.
+Second, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
+\ref{fig:wplcp}).
+At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset these costs.
+As previously suggested \citep{Givnish-1988}, shade-tolerance also decreases with height because as size increase, the total amount of energy needed to offset respiratory and turnover costs in the stem also increases.
+These finding matches well known empirical patterns, where both $\phi$ and $\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,Osunkoya-1996}.
+
+Results from \citet{Poorter-2006}: survival strongest for leaf lifespan.
+
+Also \citep{Baltzer-2007}.
+
+\subsection{Changes in trait values through ontogeny}
+
+\section{Discussion}
 
@@
 
-Plasticity in trait values predicted from model:
+Key patterns explained
 
-\citep{Osazuwa-2014} - increases in WD through ontogeny.
+Towards a global model of growth:
 
-\subsection{Shade tolerance}
+- mechanistic model of height and dbh growth which accounts for effects
+  of size, light environment and traits
+- need to account for size long recognized, but not straightforward
+- look at growth in wide range of measures: mass, height, dbh
+- Compared to previous work, shows effects of traits more about
+  allocation than mass production
 
-Second, species with low $\phi$ and low $\rho$ are predicted to be
-less shade tolerant (Fig. \ref{fig:wplcp}). At low $\phi$ ($\rho$),
-leaf (sapwood) turnover is higher and thus a greater light income is
-needed to offset these costs. As previously suggested
- \citep{Givnish-1988}, shade-tolerance also decreases with
-height because as size increase, the total amount of energy needed to
-offset respiratory and turnover costs in the stem also increases. These
-finding matches well known empirical patterns, where both $\phi$ and
-$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,Osunkoya-1996}.
+Plasticity in trait values predicted from model
 
-Results fo poort \& binger for surival strongest for LL.
+\subsection{Comparison with existing frameworks}
 
-Also baltzer.
+Growth is more than just photosynthesis:
 
-\subsection{Implications for trait-based approaches}
+Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000, Enquist-2007}, ${\rm d}P / {\rm d}t / M_{\rm a}$.
+This relationship can be derived from eqn \ref{eq:dbdt}  by ...
+(see Appendix \ref{app:traits-RGR}).
+To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in seedlings to leaf mass per leaf area \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
+This model.....Builds on 2011 paper - highlight what's new from there.
 
-How do comparative people interpret sla now - beyond fast slow.
+Recently realisation that allocation and turnover major areas of uncertainty
+
+Model reconciles idea that traits can have negative impact on carbon budget, but still have positive influence on growth rates
+
+Assumes low lma and wd used to reduce costs of building leaf and stem, ie.
+total leaf area and stem cross section remains same.
+In this way differs to assumptions of recent work by \citep{Anten-2010}, \citep{Larjavaara-2010}.
+
+\begin{itemize}
+  \item \citet{Ishihara-2016}
+  \item  \citet{Enquist-2007}: sla and wd increase mass-production, where as in our theory effects on mass production are negative.
+(also contradicts earlier model from \citet{Enquist-1999}, where argued wood density had no effect on mass production)
+  \item \citet{Ryan-2006}: reviewed evidence for hydraulic limitation of tree height.
+Designed to explain decrease in height growth rate for trees.
+Found mixed support: some species experience decreases in physiological function with height, especially very large species \citep{Koch-2004}, but certainly not not all species \citep{Ryan-2006}.
+Important to emphasise that demographic patterns observed by \citep{Ryan-2006} (decreases in height growth rate, no decrease in biomass growth) predicted from theory here, without invoking hydraulic limitation.
+So hydraulic limitation may amplify size related effects but not root cause.
+  \item \citet{King-2005} Section 3.3 of King  shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
+\citet{King-1999} also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
+Basically, the much greater path length from roots to leaves in trees vs.
+seedlings results in a much greater increment in support cost per increment in crown area for trees.
+This explanation is treated by the growth model of \citet{King-1999}, which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
+  \item \citet{King-1999}:
+  \item \citet{Li-2014}:
+  \item \citep{Givnish-1988,Yokozawa-1995,Makela-1997,King-2011,Moorcroft-2001}
+  \item \citep{Coomes-2011}:
+\end{itemize}
+
+\subsection{Plasticity}
+
+other plasticity \citep{Thomas-2010}
+
+\subsection{Implications and challenges for trait-based approaches}
+
+How do comparative people interpret $\phi$ now - beyond fast slow.
 
@@
 
-Selection on traits in different parts of lifecycle
+Selection on traits in different parts of life cycle
 
@@
 
-Framework for generalising about effect of traits
-
-Towards a global model of growth:
+Framework for generalising about effect of traits. New traits must have effects via pathways outlined in Fig. 1. Assuming plant works in way described by Fig 1.
 
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  mechanistic model of height and dbh growth which accounts for effects
-  of size, light environment and traits
-
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    need to account for size long recognized, but not straightforward
-  \item
-    look at growth in wide range of measures: mass, height, dbh
-  \end{itemize}
-\item
-  Compared to previous work, shows effects of traits more about
-  allocation than mass production
-\end{itemize}
-
-Key patterns explained
-
-Data challenges
-
-Reproductive allocation
-
-Detecting trait signals in data.
-
-\begin{itemize}
+\item Reproductive allocation
+\item Detecting trait signals in data.
 \item Need more than simply look for correlation with mean growth rate
-\item
-  understand influences of size on trait of interest
-\item
-  traits define potential growth rate --\textgreater{} how to extract
+\item Understand influences of size on trait of interest
+\item Traits define potential growth rate --\textgreater{} how to extract
   this
-\item
-  ideally start to use mechanistic models
+\item Ideally start to use mechanistic models
 \end{itemize}
@@
 
-Figure 2 uses image by Tracey Saxby, IAN Image Library http://ian.umces.edu/imagelibrary.
+Fig. 2 uses image by Tracey Saxby, IAN Image Library \smurl{http://ian.umces.edu/imagelibrary}.
 
@@
 \begin{table}[ht]
- \caption{Key phenomena to be explained}
-\centering
-
-{\footnotesize  % smaller font, double space
- \begin{doublespace}
-  \begin{tabular}{p{7cm}p{4cm}}
+\caption{Key phenomena}
+{\footnotesize
+{\centering
+  \begin{tabular}{p{9cm}p{6cm}}
   \hline
@@
   \hline
-  hump shape growth & \citep{King-2005} \\
+
+  \\\multicolumn{2}{l}{\bf{Patterns of growth with increasing size}}\\
+  Hump-shaped pattern of net biomass production & \citet{Givnish-1988, Koch-2004} \\
+  Increasing rate of plant mass and stem area growth & \citet{Stephenson-2014} \\
+  Hump-shaped pattern of $H$, $D$, $A_{\rm l}$ \& $A_{\rm st}$ growth& \citet{King-2005, Sillett-2010, King-2011, Herault-2011Ryan-2006} \\
+  Decreasing relative growth rate (for all variables) & \citet{Rees-2010, Iida-2014}\\
+  \\
+  \\ \multicolumn{2}{l}{\bf{Effect of traits on growth rate}}\\
+  Low $\phi$ species$^1$  have a growth advantage when small, that dissipates with increasing size & Meta-analysis by \citet{Gibert-2016}\\
+  Low $\rho$ species have a growth advantage at all sizes & Meta-analysis by \citet{Gibert-2016}\\
+  High $H_{\rm m}$ species have a growth advantage when small & Meta-analysis by \citet{Gibert-2016}\\
+  High $H_{\rm t,0}$ species have an advantage in absolute growth rate, while low $H_{\rm t,0}$ species have faster RGR & Meta-analysis by \citet{Gibert-2016}\\
+
+  \\ \multicolumn{2}{l}{\bf{Shade tolerance}}\\
+  \textsc{wplcp} decreases with height & \citet{Givnish-1988}\\
+  Low $\phi$ species have higher \textsc{wplcp} & \citet{Poorter-2006, Lusk-2008}\\
+  \quad $\rightarrow$ Stands of low $\phi$ species$^1$  have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
+  Low $\rho$ species have higher \textsc{wplcp} &  \citet{Osunkoya-1996}\\
+
+  \\ \multicolumn{2}{l}{\bf{Changes in traits through ontogeny}}\\
+  $\phi$ $^1$ tends to increase through ontogeny & \\
+  $\rho$ tends to increase through ontogeny & \citet{Osazuwa-2014}\\
+
   \hline
+  \\
   \end{tabular}
-\end{doublespace}}
+  }
+$^1$ Or leaf lifespan
+}
 \label{tab:phenomena}
@@
 {\footnotesize  % smaller font, double space
- \begin{doublespace}
   \begin{tabular}{p{2cm}p{2cm}p{7cm}}
@@
   \end{tabular}
-\end{doublespace}}
+}
 \label{tab:definitions}
@@
   \begin{tabular}{p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \hline
-  Variable & Function & Marginal cost & Growth rate & Eq.\\ \hline
+  Variable & Function & Marginal cost & Growth rate & Eqn\\ \hline
   \multicolumn{5}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
@@
     $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:a}\\
+    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
   bark area &
@@
     $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:a}\\  \\
+    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\  \\
   \multicolumn{5}{l}{\textbf{b) Derived equation for mass of tissue }} \\
@@
   sapwood mass &
-    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
+    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( H + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
     $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
   bark mass &
-    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
+    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( H + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
     $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
@@
 \end{tabular}
-NB -  theta is sapwood area per leaf area. In tree 2 has been opposite, but that makes notation more clumsy.
 \end{doublespace}}
@@
 \newcommand{\sepp}{{\color{grey}/}}
+\newcommand{\upup}{$\uparrow\,\uparrow$}
+\newcommand{\updo}{$\uparrow\,\downarrow$}
+\newcommand{\dodo}{$\downarrow\,\downarrow$}
+\newcommand{\upfl}{$\uparrow$\,--}
+\newcommand{\flup}{--$\,\uparrow$}
+\newcommand{\dofl}{$\downarrow\,$--}
+\newcommand{\doup}{$\downarrow\,\uparrow$}
+\newcommand{\fldo}{--$\,\downarrow$}
 
@@
 Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
-Traits are: specific leaf area (= SLA), wood density (= WD), Maximum height (= $H_{max}$), Seed mass, and Maximum photosynthesis per unit leaf area (= Area).
+Traits are: Seed mass ($M_{\rm t,0}$), Height at maturation ($H_{\rm m}$), Nitrogen content per leaf area ($\nu$), Leaf mass per unit leaf area ($\phi$),  wood density ($\rho$), Sapwood area per leaf area ($\theta$), Vessel size ($\kappa$), and root mass per leaf area ($\alpha_{\rm r1}$).
 For further details, see main text.
-Adapted and expnaded from \citet{Gibert-2016}.
-}
+Adapted and expanded from \citet{Gibert-2016}.}
 {\footnotesize
 \vspace{1cm}
-  \begin{tabular}{lllcccccc}
+  \begin{tabular}{lllcx{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}p{0.1cm}}
   \hline
-  & & & {\bf Symbol} & {\bf SLA} & {\bf WD} & {$\bf H_{max}$} & {\bf Seed mass} & {$\bf A_{area}$} \\ \hline
+  & & &  &  \multicolumn{2}{c}{\bf Ontogeny} & \multicolumn{2}{c}{\bf Leaf} & \multicolumn{3}{c}{\bf Stem}& {\bf Root} & \\
+  & & &  {\bf Symbol} & \boldmath$M_{\rm t,0}$ & \boldmath$H_{\rm m}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \boldmath$\theta$ & \boldmath$\kappa$ & \boldmath$\alpha_{\rm r1}$ & \\ \hline
+  \multicolumn{13}{l} {\textbf{Elements of equations \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
+  & \multicolumn{2}{l} {Net biomass production} & $\ud B / \ud t$ & & & & & & & & & \\
+  & & \tabitem photosynthetic rate & & - & - & $\uparrow$  & - & - & & & & \\
+  & & \tabitem respiration rate  & & - & - & $\uparrow$  & - & - & & & & \\
+  & & \tabitem turnover rate & & - & - & - & $\uparrow$ & - & & & & \\
   \\
-  \multicolumn{9}{ l } {\textbf{Elements of equations \ref{eq:eq0} -- \ref{eq:eq4}}}  \\
-  & \multicolumn{2}{ l } {Leaf deployment per mass}  & $\ud A_{\rm l} / \ud M_{\rm a} $ & \\
-    & &     \tabitem leaf  &  & $\uparrow$ & - & - & - & - \\
-    & &     \tabitem sapwood & & - & $\downarrow$ & - & - & - \\
-    & &     \tabitem root & & - & - & - & - & - \\
-  & \multicolumn{2}{ c } {Allocation to vegetative growth} & $\ud M_{\rm a} / \ud B$ & - & - & $\uparrow$ & - & - \\
-  & & & & & & \\
-  & \multicolumn{2}{ l } {Net live biomass production} & $\ud B / \ud t$ & & & & & \\
-  & &     \tabitem photosynthetic rate & & - & - & - & - & $\uparrow$  \\
-  & &     \tabitem respiration rate  & & - & - & - & - & $\uparrow$  \\
-  & &     \tabitem leaf turnover rate & & $\downarrow$ & - & - & - & - \\
-  & \multicolumn{2}{ l } {Architecture} & $\ud H / \ud A_{\rm l}$ & - & - & - & - & - \\
-  & \multicolumn{2}{ l } {Sapwood area per leaf} & $ \ud A_{\rm s} / \ud A_{\rm l}$ & - & - & - & - & - \\
-  & \multicolumn{2}{ l } {Sapwood turnover} & $k_{\rm s} \, A_{\rm s}$ & - & - & - & - & - \\ \hline
+  & \multicolumn{2}{l} {Allocation to growth} & $\ud M_{\rm a} / \ud B$ & - & $\uparrow$ & - & - & - & & & & \\
   \\
-  \multicolumn{9}{ l } {\textbf{Predicted effect on plant growth rate} (small plant {\sepp} large plant)} \\
-  & \multicolumn{2}{ l } {in mass} \\
-  & & \tabitem{Absolute} & $\ud M_{\rm a} / \ud t$ & $\uparrow$\sepp$\downarrow$ & $\downarrow${\sepp}$\downarrow$ &  $-${\sepp}$\uparrow$ & $\uparrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
-  & & \tabitem{Relative} & $\ud M_{\rm a} / (\ud t .
-M_{\rm a} )$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ &  $-${\sepp}$\uparrow$ & $\downarrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
-  & \multicolumn{2}{ l } {in height} \\
-  & & \tabitem{Absolute} & $\ud H / \ud t$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\uparrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
-  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\downarrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
-  & \multicolumn{2}{ l } {in diameter} \\
-  & & \tabitem{Absolute} & $\ud D / \ud t$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\uparrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
-  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & $\uparrow${\sepp}$\downarrow$ & $\downarrow${\sepp}$\downarrow$ & $-${\sepp}$\uparrow$ & $\downarrow${\sepp}$-$ & $\uparrow${\sepp}$\uparrow$ \\
+  & \multicolumn{2}{l} {Leaf deployment per mass}  & $\ud A_{\rm l} / \ud M_{\rm a} $ & & \\
+    & & \tabitem leaf  &  & - & - & - & $\downarrow$ & - & & & & \\
+    & & \tabitem sapwood & & - & - & - & - & $\downarrow$ & & & & \\
+    & & \tabitem root & & - & - & - & - & - & & & & \\
+  \\
+  & \multicolumn{2}{l} {Architecture} & $\ud H / \ud A_{\rm l}$ & - & - & - & - & - & & & & \\
+  \\
+  \multicolumn{13}{l} {\textbf{Predicted effect on plant growth rate} (small plant {\sepp} large plant)} \\
+  & \multicolumn{12}{l} {in mass} \\
+  & & \tabitem{Absolute} & $\ud M_{\rm a} / \ud t$ & \upfl &  \flup & \upup & \doup & \dodo & & & & \\
+  & & \tabitem{Relative} & $\ud M_{\rm a} / (\ud t . M_{\rm a} )$ & \dofl &  \flup & \upup & \doup & \dodo & & & & \\
+  \\
+  & \multicolumn{12}{l} {in height} \\
+  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & & & & \\
+  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \doup & \dodo & & & & \\
+  \\& \multicolumn{12}{l} {in stem area {\bf (Check)}} \\
+  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & & & & \\
+  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \doup & \dodo & & & & \\
+  \\& \multicolumn{12}{l} {in stem diameter} \\
+  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & & & & \\
+  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \doup & \dodo & & & & \\
 \hline
@@
 assumption (see Supplementary text for details). \label{fig:assumptions}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/mass_fraction.pdf}
-\caption{\textbf{Change in allocation with size.}
-\label{fig:mass_fraction}}
 \end{figure}
@@
 because of costs of high turnover.} Panels show effect of traits on
-maximum amount of shading that can be endured before net production (eq.
+maximum amount of shading that can be endured before net production (eqn
 \ref{eq:dbdt}) reaches zero. Lines indicate relationship for plants of a
@@
 
-\section{Appendix A - Relationship between traits and relative growth rate} \label{app:traits-RGR}
+\section{Appendices}
+\subsection{A - Relationship between traits and relative growth rate} \label{app:traits-RGR}
 
-We start with eq. \ref{eq:dbdt} for mass-based growth. For seedlings, which are
+We start with eqn \ref{eq:dbdt} for mass-based growth. For seedlings, which are
 young and mostly leaf, it is reasonable to ignore all turnover terms as
@@
 
-Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
+Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in
 relation to $\phi$\citep{Wright-2000}, this
@@
 
-\section{Appendix B - Effect of traits on height growth} \label{app:traits-RGR}
+\subsection{B - Effect of traits on height growth} \label{app:traits-RGR}
 
@@
 by looking at the effect of
-traits on the components of growth rate (eq. \ref{eq:dhdt}. To make
+traits on the components of growth rate (eqn \ref{eq:dhdt}. To make
 the analysis more tractable, we will assume we are dealing with a plant of
@@
 \end{equation}
-where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eq.
+where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eqn
 \ref{eq:G2} is a product of the form $Y(x) = W(x) \times Z(x)$. For,
@@
 
-We can further simplify by expanding the derivative of eq. \ref{eq:daldmt}:
+We can further simplify by expanding the derivative of eqn \ref{eq:daldmt}:
 \begin{equation} \label{eq:dA_dx}
@@
 
-\section{Appendix C - Linking traits to whole-plant light compensation point} \label{app:traits-RGR}
+\subsection{C - Linking traits to whole-plant light compensation point} \label{app:traits-RGR}
+
+\section{Supplementary materials}
+
+\subsection{A -- Full description of the \texttt{FF16} growth model}\label{sec:ff16}
+
+\subsection{B -- Supplementary Figures}\label{sec:SM_figs}
+
+\subsection{C -- Supplementary Tables}\label{sec:SM_tabs}
 

### 2016-08-04 R/figures.R

@@ -85,14 +85,29 @@ figure_rate_vs_size_panels <- function(data, type, path) {
   data[["darea_sapwood_darea_leaf"]] <- data[["area_sapwood"]] / data[["area_leaf"]]
   data[["darea_bark_darea_leaf"]] <- data[["area_bark"]] / data[["area_leaf"]]
 
+  get_col <- function(yvar){
+
+    cols <- RColorBrewer::brewer.pal(9, "Set1")
+    switch(yvar,
+      net_mass_production_dt=cols[1],
+      fraction_allocation_growth=cols[2],
+      darea_leaf_dmass_live =cols[3],
+      area_stem_dt =cols[4],
+      area_leaf_dt =cols[5],
+      "black")
+  }
+
   for (v in yvars) {
     filename <- file.path(path, sprintf("%s_%s.pdf", type, v))
-    pdf(filename,width=5, height=5)
-    par(oma=c(0,0,0,0), mar=rep(0.1,4))
+    pdf(filename,width=5, height=6)
+    par(oma=c(0,0,0,0), mar=c(6, 0.1,0.1,0.1))
     plot(data[["height"]], data[[v]], type="l", ann=FALSE, axes=FALSE,
        xlim=c(0, 25),
-  #    ylim=c(0, max(1,data[[type]], na.rm=TRUE)),
-      col="green", lwd=3)
+#       xaxs="i", yaxs = "i",
+      col=get_col(v), lwd=8)
+    if(type == "diameter_stem_dt" || (type == "area_stem_dt"  && v %in% c( "area_leaf_dt", "area_heartwood_dt"))){
+      axis(1, at=seq(0,30, by=10), tck=-0.05, cex.axis=4, mgp=c(3, 4, 0))
+    }
     box()
     dev.off()
   }
@@ -108,6 +123,8 @@ figure_rate_vs_size_data <- function(canopy_openness=1,
 figure_rate_vs_size_cols <- function(type) {
   if (type == "net_mass_production_dt") {
     c("net_mass_production_dt", "yield", "assimilation", "respiration", "turnover")
+  } else if (type == "mass_total_dt") {
+    c("mass_total_dt", "fraction_allocation_growth", "net_mass_production_dt", "mass_heartwood_dt")
   } else if (type == "darea_leaf_dmass_live") {
[…]

### 2016-08-04 R/figures_data.R

@@ -77,9 +77,13 @@ plot_with_sma_by_var_within_group <- function(data, xvar,yvar,group="vegetation"
 
   col.table <- make.color.table(unique(data[[group]]))
 
+  # dlply <- function(data, group, f, ...) {
+  #   ds <- split(data, data[[group]])
+  #   lapply(ds, f, ...)
+  # }
+
   fits <- dlply(data, group, add.sma.by.group, xvar=xvar, yvar=yvar, col.table=col.table,
     smagroup="species",colorBygroup=group)
-
 
   get.significant.groups <- function(fit, p=0.1){
     if(is.null(fit)) return(FALSE)

### 2016-08-05 ms/ms-supp.Rnw
Outline of main structure, embed eqn references in table and figure
Allocation model and figures

@@
 \includegraphics{SI/lma_tradeoff.pdf}
-\caption{\textbf{Leaf turnover decreases with leaf-construction cost.}
+\caption{\textbf{Leaf turnover decreases with leaf-mass per unit leaf area.}
 Data from \citep{Wright-2004} for 678 species from 51 sites, each
@@
 \centering
-\includegraphics{SI/mass_fraction.pdf}
-\caption{\textbf{Change in allocation with size.}
-\label{f-mass_fraction}}
+\includegraphics{SI/growth_light_dia.pdf}
+\caption{\textbf{Traits moderate the responsiveness of growth to changes
+in light environment (diameter growth).} Panels show predicted relationship between
+specific trait and diameter growth rate, for a plant of specified
+diameter and under a range of shading environments.
+\label{fig:growth_light_dia}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\centering
+\includegraphics{SI/growth_light_mass.pdf}
+\caption{\textbf{Traits moderate the responsiveness of growth to changes
+in light environment (mass growth).} Panels show predicted relationship between
+specific trait and mass growth rate, for a plant of specified
+mass and under a range of shading environments.
+\label{fig:growth_light_mass}}
 \end{figure}
@@
 \includegraphics{SI/lma_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
+\caption{\textbf{The expected correlation between leaf-mass per unit leaf area
 and growth rate changes with plant size.} \label{f-lma_growth_size}}
@@
 \includegraphics{SI/lma_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between leaf-construction cost
+\caption{\textbf{The expected correlation between leaf-mass per unit leaf area
 and relative growth rate changes with plant size.}
@@
 \includegraphics{SI/rho_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
+\caption{\textbf{The expected correlation between wood density
 and growth rate changes with plant size.} \label{f-rho_growth_size}}
@@
 \includegraphics{SI/rho_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between stem-construction cost
+\caption{\textbf{The expected correlation between wood density
 and relative growth rate changes with plant size.}

### 2016-08-05 ms/ms-supp.tex
Separate doc for supp mat, use ms style from plant paper
Outline of main structure, embed eqn references in table and figure
Allocation model and figures

--- 2016-07-29 ms/ms-supp.Rnw
@@
-\documentclass[10pt,twoside]{article}
+\documentclass[10pt,twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \date{}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}

### 2016-08-05 ms/ms.tex
Allocation model and figures

@@
 Although most plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in rates at which resources are acquired and invested into different tissues.
-During the last decade, trade-offs related to prominent traits -- such as leaf mass per leaf area, wood density, leaf nitrogen per area, vessel diameter, maximum  height, and seed size -- have been quantified and trait values exist for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
-Increasingly, researchers have been using traits as potential indicators of the growth strategy and ecology of a species.
+During the last decade, trade-offs related to prominent traits have been quantified and values for some traits -- such as leaf mass per leaf area, wood density, and seed size -- now exist for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
+Concomitantly, researchers have been using traits as potential indicators of the growth strategy and ecology of a species.
 
-While the influence of plant traits on elements of plant physiological function is increasingly understood, attempts at using traits to understand demographic rates however have only met with mixed success \citep{Wright-2010, Poorter-2008}.
-In seedlings, the trait leaf mass per leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- has been tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}.
-Leaf mass per leaf area and its correlate leaf lifespan have also been linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
+While the influence of plant traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have only met with mixed success \citep{Wright-2010, Poorter-2008}.
+In seedlings, the trait leaf mass per leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}.
+Leaf mass per leaf area and its correlate leaf lifespan were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
 These early successes prompted researchers to search for similar relationships in large plants.
 However, the results showed that in trees leaf mass per leaf area was not correlated with diameter growth rate \citep{Poorter-2008, Wright-2010, Herault-2011}.
-Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010}, but only weak relationships in seedlings \citep{Castro-1998}.
-Yet, until recently, most trait growth relationships had been quantified for one or few traits and at a single life- stage.
+Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010}, but less so in small plants \citep{Castro-1998}.
+In summary, traits did not correlate simply and consistently to growth rate.
 
-The picture is now emerging that the effect of traits on plant growth may be  modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-In a recent meta-analysis of more than  500 correlations recorded from 103 studies, \citet{Gibert-2016} showed that the sign of the correlation between some traits (such as leaf mass per leaf area and maximum height) and growth rate was modified by size, while for other traits (such as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Importantly, the direction of the correlations and their shifts with size, supported predictions made based on understanding of trait-based trade-offs and their interaction with plant function \citep{Gibert-2016}.
-
-The challenge of interpreting diverse empirical results seeking to link traits to growth rate has been all the harder because we currently lack clear expectations on the expected signal.
-Generating such expectations is one of the primary role for theory \citep{Kokko-2007}.
-Current empirical results suggest the effect of traits on growth changes with size and possibly also light environment, yet current theory says little about how such relationships might come about.
-A widely-used model shows that for seedlings, mass-based relative growth rate is linearly and negatively related to leaf mass per leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+The challenge of interpreting diverse empirical results seeking to link traits to growth rate is all the harder because while we lack clear theoretical expectations.
+Generating such expectations is one of the primary roles for theory \citep{Kokko-2007}.
+A widely-used model shows that for seedlings, mass-based relative growth rate may be linearly and negatively related to leaf mass per leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
 An extension of the model suggests a similar relationship may hold at larger sizes \citep{Enquist-2007}.
 But as described above, this prediction is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth are largely absent..
+Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
+Without further guidance, many researchers have simply assumed traits to linearly map onto growth rates, i.e.
+there is a ``fast'' and ``slow'' growth end of the trait spectrum.
 
-A further problem with existing theory is that the effects of traits are all realised via influences on mass production \citep{Enquist-2007}, whereas the effect of traits such as leaf mass per leaf area and wood density is via allocation to different tissues \citep{Falster-2011, Gibert-2016}.
+A further problem with existing theory is that the effects of traits are all realised via influences on mass production \citep{Enquist-2007}, whereas the influence of some traits such as leaf mass per leaf area and wood density is via allocation to different tissues rather than mass production \citep{Falster-2011, Gibert-2016}.
 A second issue with theory focussed solely on mass production is that measuring mass production is only practical for small plants that can be easily harvested.
 For larger plants, growth is more often measured as increment in either diameter or height.
-For example, the Centre for Tropical Science monitors diameter growth in 49 large permanent plots across forests worldwide \citep{Anderson-2015}; forestry agencies have likewise recorded growth in diameter in 1000's of plots across the globe \citep{Purves-2008, Kunstler-2015}.
-But as we outline further below, mass production is only one component of plant growth; meaning theory focussing solely on mass production will always be limited.
+For example, natural research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
+As we outline further below, mass production is only one component of plant growth; meaning theory focussing solely on mass production will always be limited.
 
-Ultimately, trait based understanding must b integrated within a more general model of plant growth, also able to capture patterns of growth through ontogeny.
+Ultimately, trait based understanding must be integrated within a more general model of plant growth, also able to capture patterns of growth through ontogeny.
 Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
 In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover.
-All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of maintaining stems \citep{Givnish-1988}, and also vary with traits \citep{Lusk-2006,Poorter-2006}.
-These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are in need of a comprehensive and integrated explanation.
+And all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
+Meanwhile shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of maintaining stems \citep{Givnish-1988}, and also vary with traits \citep{Lusk-2006,Poorter-2006}.
+These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.
 
-In this paper we show how a recently published mechanistic framework \plant{} \citep{Falster-2016} describing the potential effect of traits on growth throughout a plant's life-cycle can explain diverse phenomena (Table \ref{tab:phenomena}).
-The model employs the same framework for partitioning growth used by \citet{Gibert-2016} in their meta-analysis, building on and extending several related approaches \citep{Givnish-1988, Yokozawa-1995, Makela-1997, Moorcroft-2001, Falster-2011, King-2011} in which metrics such as height, diameter and mass growth are calculated as an emergent outcomes of several, independent, physiological processes.
-This approach is quite different to models derived from metabolic scaling theory \citep{Enquist-1999, Enquist-2007}, which attempt to organise everything around a single master ``scaling'' equation for mass growth, or statistically fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014}.
-By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.
+In this paper we show how a new mechanistic framework -- called {\plant} \citep{Falster-2016} -- offers fresh insight into the way traits influence plant growth throughout across the life-cycle and explain diverse phenomena (Table \ref{tab:phenomena}).
+Our primary aim is to explain a pattern that has been gradually emerging and recently solidified: that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
+A recent meta-analysis of 103 studies reporting \textgreater 500 correlations by \citet{Gibert-2016} provides the most compelling evidence.
+\citet{Gibert-2016} showed that the sign of the correlation between some traits (such as leaf mass per leaf area and maximum height) and growth rate was modified by size, while for other traits (such as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
+Importantly, the direction of the correlations and their shifts with size supported predictions generated via a framework describing how traits influence growth.
+Here we extend the results of taken by \citet{Gibert-2016} to provide a complete, functioning model and explain, physiologically, why it makes sense that the effect of traits changes with size.
+By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation linking trait-based trade-offs to demography and thus provide a platform for understanding growth across diverse species around the world.
 
@@
 
-Our approach builds on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within the plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
+We build on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within the plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
 \ref{fig:conceptual}a).
@@
 The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A full list of variables, units and definitions is given in Table \ref{tab:params}.
+A full list of variables, units and definitions is given in Table \ref{tab:definitions} and Table \ref{tab:traits}, with further detail on the parameter values applied given in Table \ref{tab:params}.
 
 We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant.
-Adopting a standard model, the amount of biomass available for growth, ${\rm d}B / {\rm d}t$ is given by the difference between income (total photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000, Falster-2011} (Fig.
+According to a standard model, the amount of biomass available for growth, ${\rm d}B / {\rm d}t$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig.
 \ref{fig:conceptual}a, eqn \ref{eq:dbdt}).
@@
 The constant $y$ in eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass.
-Although many models include a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, net biomass production eventually comes down to simple subtraction of income and loss.
+We assume that  that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ by trait (see below).
+While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
+as functions of temperature, such detail is not needed here.
+Even with detailed models, ${\rm d}B / {\rm d}t$ always comes down to simple subtraction of income and losses.
 
-The only potentially controversial element of in eqn \ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocated to either new growth (i.e.
-growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$ or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction often entails run-down in the vegetative part of the plant (refs XX).
-
-\subsection{Decomposing of growth rate into components}
+\subsection{Decomposing of growth rates into components}
 
 To model growth in either height, leaf area, basal area, standing mass or stem diameter requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
-Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig.
-\ref{fig:conceptual}b, eqns \ref{eq:dbdt}-\ref{eq:dDdt}).
+Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}.
+Relevant equations (eqns \ref{eq:dbdt}-\ref{eq:dDdt}) are given Fig.
+\ref{fig:conceptual}b.
 Importantly, most of these terms vary intrinsically with size.
 The insets in Fig.
-\ref{fig:conceptual}b show size-related patterns for a typical plant, from a specific allocation model derived below.
-However, it is important to recognise that eqns \ref{eq:dbdt}-\ref{eq:dDdt} are mathematically true, derived using the chain rule of differential calculus, and thus must hold true for any class of models where carbon allocation is important.
+\ref{fig:conceptual}b show size-related patterns in different variables for a typical plant obtained from applying the allocation model derived below.
+However, it is important to recognise that eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are mathematically true, derived using the chain rule of differential calculus, and thus must hold true for any class of models where carbon allocation is important.
 
-The first point to note is that the growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$.
-We assume the remaining fraction, $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$, is allocated to reproduction.
-The term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, presumably at 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants.
-It is important to note here that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
-So even a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ does not increase any further.
+Note that the growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
+The remaining fraction,$1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$, is allocated to reproduction.
+$\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, presumably at 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants.
+It is important to realise that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
+So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even while the net amount of live mass $M_{\rm a}$ does not increase any further.
 
@@
 To make predictions from eqns \ref{eq:dbdt}-\ref{eq:dDdt}, however, requires a specific model of plant construction.
-The model described below can be considered a first-order functional-balance model, in that it seeks to maintain a balance between various elements of plant construction through ontogeny.
-Table \ref{tab:allometry} provides key equations of the model, while Table \ref{tab:params} provides estimates on key variable and fitted parameters.
+The model described below can be considered a first-order functional-balance model.
+We might also call it ``isometric'', because the assumptions ensure the proportional relationships among the relevant variables are preserved as the plant grows.
+Table \ref{tab:allometry} provides key equations of the model (see Supplementary material \ref{sec:ff16} or \citet{Falster-2016} for full derivation).
+In particular, we assume:
+\begin{description}
+\item[Assumption 1:] A log-log scaling relationship with slope of 0.5 between the leaf area of a plant and its height (eqn \ref{eq:ha}).
+\item[Assumptions 2:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of sapwood in a plant (eqn \ref{eq:as}).
+\item[Assumptions 3:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of bark and phloem in a plant (eqn \ref{eq:ab}).
+\item[Assumptions 4:] A log-log scaling relationship with slope of 1.0 between the leaf area and mass of fine roots in a plant (eqn \ref{eq:mr}).
+\item[Assumption 5:] The vertical distribution of leaf area within the canopy of an individual plant remains relatively constant through ontogeny.
+\end{description}
 
-Our model is founded on four key assumptions:
+Assumption 1 accounts for the architectural layout of the plant.
+Assumptions 2-4 are implementations of the pipe model \citep{Shinozaki-1964}, whereby the area of conducting tissues are proportional to leaf area.
+To describe the vertical distribution of leaf area within the canopy of an individual plant, we follow the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1.
+Low values result in a conical canopy, as seen in many conifers, while higher values give a top-weighted canopy profile similar to those seen among angiosperms.
 
-\noindent{\bf Assumption  1} -- There is an allometric (log-log) scaling relationship between the accumulated leaf area of a plant and its height.
-Relationship represented by eqn \ref{eq:ha}.
-Empirically motivated.
-Describes architectural strategy of plant.
+The functional-balance assumptions are well-supported by data, at least to first approximation.
+Fig.
+\ref{fig:assumptions} shows data sourced from the Biomass and Allometry Database (\textsc{baad}) \citep{Falster-2015b}, which includes records for 21084 individuals on 656 species for various size metrics of individual plants.
+Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a.
+As is evident, the lines fitted within each species roughly align with predicted functional-balance assumption.
 
-\noindent{\bf Assumptions 2-4} -- An isometric relationship between the total leaf area and each of sapwood cross-sectional area (2), bark cross-sectional area (3), and total root mass (4).
-As per the pipe model, \citep{Shinozaki-1964} we assume that each unit of sapwood area supports a fixed area of leaf, so that the total canopy area of a plant relates to basal sapwood area eqn \ref{eq:as}.
-This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (eqn \ref{eq:ms2}).
-Bark and phloem tissue are modelled using an analogue of the pipe model (eqn \ref{eq:as}), leading to a similar equation as that for sapwood mass (eqn \ref{eq:msb}).
-Also consistent with pipe-model assumption, we assume a fixed ratio of root mass per unit leaf area (eqn \ref{eq:mr}).
+Combined, these assumptions in Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
+Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives a the necessary components in Fig.
+\ref{fig:conceptual} and thereby an explicit model for plant growth in relation to size.
+Note also that no assumptions are made about the diameter and total mass of the stem in relation to other components, e.g.
+height: these relationships arise as emergent properties, via integrating stem turnover (eqns \ref{eq:ha}- \ref{eq:hm}).
 
-\noindent{\bf Assumption 5} -- Vertical distribution of leaf within the canopy of a plant remains relatively constant through ontogeny.
-We follow the model of \citep{Yokozawa-1995} describing the vertical distribution of leaf area within the crowns of individual plants.
-This model can account for a variety of canopy profiles through a single parameter $\eta$.
-Setting $\eta=1$ results in a conical canopy, as seen in many conifers, while higher values, e.g.
-$\eta=12$ , give a top-weighted canopy profile similar to those seen among angiosperms.
-
-Assuming lead immediately to an allometric model for plant size (Table \ref{tab:allometry}, see Supplementary material for derivations).
-
-To first approximation, the assumptions, which are well-supported by data (Fig. \ref{fig:assumptions}).
-
-Simple linear functions non-critical for results of this paper, explore using an example model, but any model with basic size effects will display same behaviour.
-
-Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model for plant growth in relation to size.
+The most important prediction arising from the eqns in Table \ref{tab:allometry} is that the amount of live tissue supporting each unit of leaf area increases with plant height.
+We can derive an explicit prediction in relation to the assumed allocation model:
+\begin{equation}\label{eq:msml}
+\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
+\end{equation}
+Note that $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are traits, i.e.
+properties that are assumed approximately constant through ontogeny, so that their product is a constant.
+Again, the predicted relationship is supported by available data from the \textsc{baad} (Fig.
+\ref{fig:support_costs}).
 
@@
 
-\subsection{Parameters}
-
-\citep{Falster-2015b}: BAAD
-\citep{Falster-2016}: \plant
+Now consider how trait-based trade-offs can be embedded into the above growth model.
+It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect never-ending directional trait values towards more beneficial values.
+For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}.
+In postulating potential benefits and costs, we only consider direct biophysical consequences of varying a trait, i.e.
+we do not try to account for every single variable that may sometimes correlate with a given trait, only those for which there is thought to be an inevitable trade-off.
+The trade-offs described in Table \ref{tab:traits} are then implemented as follows.
+\begin{description}
+\item[Seed mass] Effects of the trait seed mass, denoted by $\omega$, are naturally embedded in the equations determining fecundity and size of seeds.
+We assume larger seeds result in larger seedlings, thus the energetic tradeoff arising from seed size translates into a demographic tradeoff between the number of seedlings produced by a parent and the initial size of seedlings.
+\item[Height at maturation] ........
+\item[Nitrogen per unit leaf area] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area $\nu$, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
+Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
+\item[Leaf mass per unit area] The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$.
+In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
+\item[Wood density] The trait wood density, denoted by $\rho$, directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$.
+The cost of cheaper wood is increased mortality rate, which is not included in the growth equations but itself an important demographic outcome.
+The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as $r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},$ where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood.
+Similarly, the rate of bark respiration per unit mass varies as $r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},$ with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.
+ \end{description}
 
@@
 
+The 
+Parameters
+
+\citep{Falster-2016}: \plant
+
 \subsection{Changes in growth rate with size}
 
-This new growth model captures the intrinsically size-dependent nature of plant growth, recovering well-known empirical patterns \citep{Sillett-2010, King-2011}.
+This new growth model captures the intrinsically size-dependent nature of plant growth, recovering well-known empirical patterns (Table \ref{tab:phenomena}).
 Height growth shows a hump-shaped pattern with size, first increasing then decreasing.
@@
 Decreasing $\phi$ has two impacts on growth rate.
-First, lower $\phi$ increases marginal leaf deployment (${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on construction costs.
+First, lower $\phi$ increases marginal leaf deployment ($\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ by economising on construction costs.
 Second, lower $\phi$ decreases net production (${\rm d}P / {\rm d}t$), due to increased leaf turnover.
@@
 
-If the trait of interest is leaf-construction cost, i.e.
+If the trait of interest is leaf-mass per unit leaf area, i.e.
 $x=\phi$, we can expand eqn \ref{eq:dg4} as follows.
-Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$.
+Note that $\frac{\partial \left(\frac{{\rm d}M_{\rm t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$.
 Also, from eqn \ref{eq:dbdt} we obtain,
@@
 \end{equation}
-where $c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$is independent of $\phi$, and thus
+where $c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_{i={\rm b,s,r}}{M_{i} \, r_{i}}) - (\sum_{i={\rm b,s,r}}{M_{i} \, ???})$is independent of $\phi$, and thus
 \begin{equation}\label{eq:dbdt3}
@@
 eqn \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent of $\phi$.
-In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
+In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ {\rm d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
 
@@
 < \frac{1}{\phi
- + \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
- {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
+ + \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
 \end{equation}
@@
 Thus diameter growth continues, even as net production approaches zero.
-
-\subsubsection{Sapwood area per leaf area, $\theta$}
-
-\subsubsection{Vessel size, $\kappa$}
-
-\subsubsection{Root mass per leaf area, $\alpha_{\rm r1}$}
 
@@
 Model is realisation of several ideas that have been around for some
-time.
+time.Employs the same framework for partitioning growth used by \citet{Gibert-2016} in their meta-analysis, which builds on and extends several related approaches \citep{Givnish-1988, Yokozawa-1995, Makela-1997, Moorcroft-2001, Falster-2011, King-2011} in which metrics such as height, diameter and mass growth are calculated as an emergent outcomes of several, independent, physiological processes.
+This approach is quite different to models derived from metabolic scaling theory \citep{Enquist-1999, Enquist-2007}, which attempt to organise everything around a single master ``scaling'' equation for mass growth; and also different from statistically fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it posits rather than tests for traits based effects.
+Our goal is to help
 
@@
 
+The only potentially controversial element of in eqn \ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocated to either new growth (i.e.
+growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$ or reproduction \citep{Thornley-2000}.
+This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction often entails run-down in the vegetative part of the plant (refs XX).
+
+In what follows, the assumption of exact isometry is non-critical, what matters is that the amount of live biomass (i.e.
+excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
+\ref{fig:support_costs}).
+Note that similar increases in construction and maintenance costs with plant size are embedded within various models of plant function \citep[e.g.][]{Givnish-1988, Makela-1997}.
+
+The list of functional traits that are known to differ among plant species is already long and ever-increasing. However, far fewer have well established tradeoffs.
+
 \subsection{Plasticity}
@@
   \begin{tabular}{p{9cm}p{6cm}}
+  \\
   \hline
@@
   Increasing rate of plant mass and stem area growth & \citet{Stephenson-2014} \\
-  Hump-shaped pattern of $H$, $D$, $A_{\rm l}$ \& $A_{\rm st}$ growth& \citet{King-2005, Sillett-2010, King-2011, Herault-2011Ryan-2006} \\
+  Hump-shaped pattern of $H$, $D$, $A_{\rm l}$ \& $A_{\rm st}$ growth& \citet{King-2005, Ryan-2006, Sillett-2010, Herault-2011, King-2011} \\
   Decreasing relative growth rate (for all variables) & \citet{Rees-2010, Iida-2014}\\
@@
   $\rho$ tends to increase through ontogeny & \citet{Osazuwa-2014}\\
-
-  \hline
+  \\ \hline
   \\
@@
 \centering
-
-{\footnotesize  % smaller font, double space
+{\footnotesize 
   \begin{tabular}{p{2cm}p{2cm}p{7cm}}
+  \\
   \hline
@@
   \hline
-
+\\
   $B$   & kg  & biomass originating from parent plant\\
@@
   $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
-  \hline
+  \\\hline
   \end{tabular}
@@
 \begin{table}[ht]
-\caption{Equations for an allometric growth model, based on functional balance assumption in {\textbf a}. }
+ \caption{Key traits and trade-offs}
+\centering
+{\footnotesize
+\raggedright
+  \begin{tabular}{p{3cm}p{1cm}p{1.5cm}p{5cm}p{3cm}}
+  \\
+  \hline
+  Name & Symbol & Unit & Trade-off (benefit {\bf vs.} cost) & References \\
+  \hline
+  \\
+  Seed mass & $M_{\rm t, 0}$ & kg & $\uparrow$ initial size seedlings {\bf vs.} $\downarrow$  maternal fecundity & \citet{Moles-2006}\\
+  \\
+  Height at maturation & $H_m$ & m  & $\uparrow$ growth {\bf vs.} $\downarrow$ fecundity & \citet{Thomas-2011, Wenk-2015}\\
+\\
+  Leaf nitrogen per unit leaf area & $\nu$ & kg m$^{-2}$ & $\uparrow$ photosynthesis in high light {\bf vs.} $\uparrow$ respiration rate & \citet{Wright-2004}\\
+  \\
+  Leaf mass per unit leaf area & $\phi$ & kg m$^{-2}$ & $\downarrow$ cost of building leaf {\bf vs.} $\uparrow$ leaf turnover & \citet{Wright-2004}\\
+  \\
+  Wood density & $\rho$ &  kg m$^{-3}$ & $\downarrow$ cost of building stem {\bf vs.} $\uparrow$ mortality rate & \citet{Chave-2009}\\
+  \\
+  \hline
+  \end{tabular}
+}
+\label{tab:traits}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
+\caption{Equations for a functional-balance allocation model.
+The first column of part {\textbf a} provides cores assumptions between various size metrics and leaf area.
+Eqns in the middle and right columns of {\textbf b}, and in {\textbf b}can then be derived from the assumptions in {\textbf a}.}
 \centering
 
-{\footnotesize  % smaller font, double space
- \begin{doublespace}
-  \begin{tabular}{p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \hline
+{\footnotesize
+  \begin{tabular}{p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \\ \hline
   Variable & Function & Marginal cost & Growth rate & Eqn\\ \hline
-  \multicolumn{5}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
-  height &
-    $H = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
-    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
+  \\
+  \multicolumn{5}{l}{\textbf{a) Assumptions}} \\
+  Height &
+    $H = \alpha_{\rm l1} \, A_{\rm l}^{0.5}$ &
+    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} = 0.5 \alpha_{\rm l1} A_{\rm l}^{-0.5}$ &
     $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
-  sapwood area &
+  Sapwood area &
     $A_{\rm s} = \theta \, A_{\rm l}$ &
@@
     $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
-  bark area &
+  Bark area &
     $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
     $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\  \\
-  \multicolumn{5}{l}{\textbf{b) Derived equation for mass of tissue }} \\
-  leaf mass &
+    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
+  Root mass &
+    $M_{\rm r} = \alpha_{\rm r1} \, A_{\rm l}$ &
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_{\rm r1}$  &
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
+  Heartwood area & & &
+    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\ \\
+
+  \multicolumn{5}{l}{\textbf{b) Derived quantities}} \\
+  Leaf mass &
     $M_{\rm l} = \phi \, A_{\rm l} $ &
@@
     $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
-  sapwood mass &
+  Sapwood mass &
     $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
@@
     $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
-  bark mass &
+  Bark mass &
     $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
@@
     $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
-  root mass &
-    $M_{\rm r} = \alpha_3 \, A_{\rm l}$ &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\ \\
-
-  \multicolumn{5}{l}{\textbf{c) Integrated quantities}} \\
-  heartwood area &
+  Heartwood area &
     $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
      &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:a} \\
-  heartwood mass &
+    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:ha} \\
+  Heartwood mass &
     $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
      &
-    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:a} \\
-  \hline \\
-\end{tabular}
-\end{doublespace}}
+    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:hm} \\
+    \\
+  \hline
+\end{tabular}}
 \label{tab:allometry}
@@
 \centering
-\caption{Predicted effects of traits on key elements of plant function determining growth rate.
+\caption{{\bf Predicted effects of traits on key elements of plant function determining growth rate.} Adapted and expanded from \citet{Gibert-2016}.
 Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
-Traits are: Seed mass ($M_{\rm t,0}$), Height at maturation ($H_{\rm m}$), Nitrogen content per leaf area ($\nu$), Leaf mass per unit leaf area ($\phi$),  wood density ($\rho$), Sapwood area per leaf area ($\theta$), Vessel size ($\kappa$), and root mass per leaf area ($\alpha_{\rm r1}$).
-For further details, see main text.
-Adapted and expanded from \citet{Gibert-2016}.}
+Traits are: Seed mass ($M_{\rm t,0}$), Height at maturation ($H_{\rm m}$), Nitrogen content per leaf area ($\nu$), Leaf mass per unit leaf area ($\phi$),  Wood density ($\rho$), Sapwood area per leaf area ($\theta$), Vessel size ($\kappa$), and root mass per leaf area ($\alpha_{\rm r1}$).}
 {\footnotesize
 \vspace{1cm}
-  \begin{tabular}{lllcx{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}x{0.5cm}p{0.1cm}}
+  \begin{tabular}{lllcx{0.8cm}x{0.8cm}x{0.8cm}x{0.8cm}x{0.8cm}p{0.1cm}}
   \hline
-  & & &  &  \multicolumn{2}{c}{\bf Ontogeny} & \multicolumn{2}{c}{\bf Leaf} & \multicolumn{3}{c}{\bf Stem}& {\bf Root} & \\
-  & & &  {\bf Symbol} & \boldmath$M_{\rm t,0}$ & \boldmath$H_{\rm m}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \boldmath$\theta$ & \boldmath$\kappa$ & \boldmath$\alpha_{\rm r1}$ & \\ \hline
-  \multicolumn{13}{l} {\textbf{Elements of equations \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
-  & \multicolumn{2}{l} {Net biomass production} & $\ud B / \ud t$ & & & & & & & & & \\
-  & & \tabitem photosynthetic rate & & - & - & $\uparrow$  & - & - & & & & \\
-  & & \tabitem respiration rate  & & - & - & $\uparrow$  & - & - & & & & \\
-  & & \tabitem turnover rate & & - & - & - & $\uparrow$ & - & & & & \\
+  & & &  &  \multicolumn{2}{c}{\bf Ontogeny} & \multicolumn{2}{c}{\bf Leaf} & \multicolumn{1}{c}{\bf Stem}& \\
+  & & &  {\bf Symbol} & \boldmath$M_{\rm t,0}$ & \boldmath$H_{\rm m}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\ \hline
   \\
-  & \multicolumn{2}{l} {Allocation to growth} & $\ud M_{\rm a} / \ud B$ & - & $\uparrow$ & - & - & - & & & & \\
+  \multicolumn{10}{l} {\textbf{a) Effect on elements of equations \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
+  & \multicolumn{2}{l} {Net biomass production} & $\ud B / \ud t$ & & & & & & \\
+  & & \tabitem Photosynthesis & & - & - & $\uparrow$  & - & - & \\
+  & & \tabitem Respiration  & & - & - & $\uparrow$  & - & - & \\
+  & & \tabitem Turnover & & - & - & - & $\uparrow$ & - & \\
+  \\
+  & \multicolumn{2}{l} {Allocation to growth} & $\ud M_{\rm a} / \ud B$ & - & $\uparrow$ & - & - & - & \\
   \\
   & \multicolumn{2}{l} {Leaf deployment per mass}  & $\ud A_{\rm l} / \ud M_{\rm a} $ & & \\
-    & & \tabitem leaf  &  & - & - & - & $\downarrow$ & - & & & & \\
-    & & \tabitem sapwood & & - & - & - & - & $\downarrow$ & & & & \\
-    & & \tabitem root & & - & - & - & - & - & & & & \\
+    & & \tabitem Leaf  &  & - & - & - & $\downarrow$ & - & \\
+    & & \tabitem Sapwood & & - & - & - & - & $\downarrow$ & \\
+    & & \tabitem Root & & - & - & - & - & - & \\
   \\
-  & \multicolumn{2}{l} {Architecture} & $\ud H / \ud A_{\rm l}$ & - & - & - & - & - & & & & \\
+  & \multicolumn{2}{l} {Architecture} & $\ud H / \ud A_{\rm l}$ & - & - & - & - & - & \\
   \\
-  \multicolumn{13}{l} {\textbf{Predicted effect on plant growth rate} (small plant {\sepp} large plant)} \\
-  & \multicolumn{12}{l} {in mass} \\
-  & & \tabitem{Absolute} & $\ud M_{\rm a} / \ud t$ & \upfl &  \flup & \upup & \doup & \dodo & & & & \\
-  & & \tabitem{Relative} & $\ud M_{\rm a} / (\ud t . M_{\rm a} )$ & \dofl &  \flup & \upup & \doup & \dodo & & & & \\
+  \multicolumn{10}{l} {\textbf{b) Predicted effect on plant growth rate for a small and large plant}} \\
+  & \multicolumn{9}{l} {in mass {\color{red} (check, live or total?)}} \\
+  & & \tabitem{Absolute} & $\ud M_{\rm a} / \ud t$ & \upfl &  \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Relative} & $\ud M_{\rm a} / (\ud t . M_{\rm a} )$ & \dofl &  \flup & \upup & \doup & \dodo & \\
   \\
-  & \multicolumn{12}{l} {in height} \\
-  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & & & & \\
-  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \doup & \dodo & & & & \\
-  \\& \multicolumn{12}{l} {in stem area {\bf (Check)}} \\
-  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & & & & \\
-  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \doup & \dodo & & & & \\
-  \\& \multicolumn{12}{l} {in stem diameter} \\
-  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & & & & \\
-  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \doup & \dodo & & & & \\
+  & \multicolumn{9}{l} {in height} \\
+  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \doup & \dodo & \\
+  \\& \multicolumn{9}{l} {in stem area {\color{red} (check)}} \\
+  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \doup & \dodo & \\
+  \\& \multicolumn{9}{l} {in stem diameter} \\
+  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \doup & \dodo & \\
+  \\
 \hline
@@
 \centering
-\includegraphics{main/allometry.pdf}
-\caption{\textbf{Key assumptions of a functional balance and trait
-trade-offs model, evaluated using global dataset.} We used the biomass and
-allometry database to evaluate model assumptions about \textbf{a,}
-scaling of leaf area with plant height, \textbf{b} Scaling of sapwood
-area with leaf area, and \textbf{c} scaling of root mass with leaf area.
-Each dot is a single plant. Lines show standardised major axis lines
-fitted to data from each site, with intensity of shading adjusted
-according to strength of the relationship. Colours indicate vegetation
-type. Dashed black lines show values expected under functional-balance
-assumption (see Supplementary text for details). \label{fig:assumptions}}
+\includegraphics{main/assumptions.pdf}
+\caption{\textbf{Key assumptions of a functional balance model for plant construction}.
+Data are from the \textsc{baad} \citep{Falster-2015b}.
+Each dot is a single plant.
+Lines show standardised major axis lines fitted by species, with intensity of shading adjusted according to strength of the relationship.
+Dashed lines indicate slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}.
+\label{fig:assumptions}}
 \end{figure}
 
+\begin{figure}[ht]
+\centering
+\includegraphics{main/support_costs.pdf}
+\caption{\textbf{Amount of live stem tissue per unit leaf area increases with plant height}.
+Data are from the \textsc{baad} \citep{Falster-2015b}.
+Each dot is a single plant.
+Plants \textless{}  1m tall were assumed to have no heartwood.
+Lines show standardised major axis lines fitted by species, with intensity of shading adjusted according to strength of the relationship.
+Dashed lines indicate slope of 1, being the prediction of the model in Table \ref{tab:allometry}.
+The dark line shows the relationship assumed in this paper.
+\label{fig:support_costs}}
+\end{figure}
 \newpage
@@
 \centering
-\includegraphics{main/growth_light_dia.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (diameter growth).} Panels show predicted relationship between
-specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{fig:growth_light_dia}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/growth_light_mass.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (mass growth).} Panels show predicted relationship between
-specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{fig:growth_light_mass}}
+\includegraphics{main/growth_light_height.pdf}
+\caption{\textbf{Effects of some traits on height growth changes with plant size.} Panels show predicted relationship between specific trait and height growth rate, for a plant of specified height and under a range of shading environments.
+\label{fig:growth_light_height}}
 \end{figure}
@@
 rate a linear function of $\phi$:
-
 \begin{equation}\label{eq:RGR}
-\underbrace{\strut\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}. \end{equation}
-
+\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}. \end{equation}
 Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to $\phi$\citep{Wright-2000}, this
+relation to $\phi$ \citep{Wright-2000}, this
 approximation does not directly link to other traits, or to the
 variables that are routinely collected for large trees: namely plant
-height ($h$) and stem cross-sectional area ($A_{\rm st}$) or
+height ($H$) and stem cross-sectional area ($A_{\rm st}$) or
 diameter $D$.
@@
 by looking at the effect of
-traits on the components of growth rate (eqn \ref{eq:dhdt}. To make
+traits on the components of growth rate (eqn \ref{eq:dhdt}). To make
 the analysis more tractable, we will assume we are dealing with a plant of
@@
 where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eqn
-\ref{eq:G2} is a product of the form $Y(x) = W(x) \times Z(x)$. For,
+\ref{eq:G2} is a product of the form $Y(x) = W(x) \, Z(x)$. For,
 equations of this type, the derivative $\partial{Y}/\partial{x}$ is

### 2016-08-05 R/figures.R

@@ -9,6 +9,10 @@ default_strategy <- function() {
       )
 }
 
+color_pallete <- function(n) {
+  RColorBrewer::brewer.pal(n, "Blues")
+}
+
 run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
                                time_max=300, filter=FALSE) {
   pl <- FF16_PlantPlus(strategy)
@@ -150,7 +154,7 @@ figure_dY_dt <- function(dat) {
 
   ymax <- max(sapply(dat[traits], function(x) max(x[[vars[2]]], na.rm=TRUE)))
   ylim <- c(0, ymax * 1.1)
-  cols <- rev(RColorBrewer::brewer.pal(length(lai) + 3, "Blues")[-(1:3)])
+  cols <- rev(color_pallete(length(lai) + 3)[-(1:3)])
   par(mfcol=c(length(sizes), length(traits)),
       oma=c(4, 5, 0, 1.5), mar=c(1, 1, 1, 1))
   for (v in traits) {
@@ -173,6 +177,14 @@ figure_dY_dt <- function(dat) {
   legend("topright", paste(rev(lai), "m2"), lty=1, col=cols, bty="n")
 }
 
+
+figure_height_dt_data <- function() {
+  ret <- figure_dY_dt_data(sizes =  c(0.5, 2, 10, 20),
+        vars = c("height", "height_dt")
+        )
+  ret[["label"]] <- function(x) sprintf("H=%sm",x)
+  ret
+}
 
 figure_diameter_stem_dt_data <- function() {
   ret <- figure_dY_dt_data(sizes =  c(0.005, 0.01, 0.1, 0.2),
@@ -268,7 +280,7 @@ figure_lcp_whole_plant <- function() {
     heights <- c(0.1, 2, 10, 20)
     x <- seq_log_range(xlim, 20)
     s <- default_strategy()
-    cols <- RColorBrewer::brewer.pal(length(heights) + 3, "Blues")[-(1:3)]
+    cols <- color_pallete(length(heights) + 3)[-(1:3)]
 
     lcp <- sapply(heights, function(h)
                   lcp_whole_plant_with_trait(trait_matrix(x, trait), h, s))

### 2016-08-05 R/figures_data.R

@@ -44,7 +44,13 @@ figure_lma_tradeoff <- function(data) {
          pch=16, col=col_table, cex=1, title=title)
 }
 
-addIsoclines <- function(A= seq(-10,10),b=1,...){
+addIsoclines <- function(A=NULL,b=NULL,...){
+  if(is.null(A)) {
+    A <- seq(-10,10)
+  }
+  if(is.null(b)) {
+    b <- 1
+  }
   for(a in A) abline(a,b,lty="dotted",...)
 }
 
@@ -64,23 +70,17 @@ make.color.table <- function(names){
 }
 
 # make a plot, fits lines to each within each group, applyiong same colour within that group
-plot_with_sma_by_var_within_group <- function(data, xvar,yvar,group="vegetation",location=NULL, isoclines=NULL,...){
-
+plot_with_sma_by_var_within_group <- function(data, xvar,yvar,group="vegetation",location=NULL, isoclines=NULL, A = NULL,...){
   plot(data[[xvar]], data[[yvar]],  type= 'n', log="xy", las=1, yaxt="n", xaxt="n",...)
   axis.log10(1)
   axis.log10(2)
 
   if(!is.null(isoclines))
-    addIsoclines(b=isoclines)
+    addIsoclines(b=isoclines, A=A)
 
   points(data[[xvar]], data[[yvar]],  type= 'p', col = make.transparent("grey", 0.4), pch=16)
 
   col.table <- make.color.table(unique(data[[group]]))
-
-  # dlply <- function(data, group, f, ...) {
-  #   ds <- split(data, data[[group]])
-  #   lapply(ds, f, ...)
-  # }
 
   fits <- dlply(data, group, add.sma.by.group, xvar=xvar, yvar=yvar, col.table=col.table,
     smagroup="species",colorBygroup=group)
@@ -98,6 +98,34 @@ plot_with_sma_by_var_within_group <- function(data, xvar,yvar,group="vegetation"
   fits
 }
 
+# make a plot, fits lines to each spp
+plot_with_sma_by_spp <- function(data, xvar, yvar, location=NULL, isoclines=NULL, A = NULL,
+  xlab=NULL, ylab=NULL, ...){
[…]

### 2016-09-06 ms/ms.tex
Refinements of text and figures

@@
 The constant $y$ in eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass.
-We assume that  that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ by trait (see below).
+We assume that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ by trait (see below).
 While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
@@
 
+\subsection{Shade tolerance}
+A useful measure of a plant's shade tolerance is the point at which photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988,Baltzer-2007,Lusk-2013}, sometimes referred to as the the ``whole-plant light compensation point'' (\wplcp{}).
+In general, the average assimilation rate per leaf area $\bar{p}$ is expected to increase with light level or canopy openness, $E$.
+The \wplcp can be estimated by solving the value of $E=E^*$ where $\frac{{\rm d}B} {{\rm d}t}(E)$ from eqn \ref{eq:dbdt} is zero, which occurs when:
+\begin{equation}\label{eq:wplcp}
+\bar{p}(E^*) =\frac{\sum \, M_{\rm i} \left(k_{\rm i} y + r_{\rm i}\right)}{A_{\rm l}}.
+\end{equation}
+Here $\bar{p}(E)$ is the rate of photosynthesis at canopy openness $E$. Below we express
+
 \subsection{A functional-balance model for plant construction}
@@
 Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a.
-As is evident, the lines fitted within each species roughly align with predicted functional-balance assumption.
+As is evident, the lines fitted within each species roughly align with predicted functional-balance assumption Fig.
+\ref{fig:assumptions}a-c.
 
@@
 Again, the predicted relationship is supported by available data from the \textsc{baad} (Fig.
-\ref{fig:support_costs}).
+\ref{fig:assumptions}d).
 
@@
 
-Define as \textsc{wplcp}
-
 Second, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
@@
 Designed to explain decrease in height growth rate for trees.
-Found mixed support: some species experience decreases in physiological function with height, especially very large species \citep{Koch-2004}, but certainly not not all species \citep{Ryan-2006}.
+Found mixed support: some species experience decreases in physiological function with height, especially very large species \citep{Koch-2004}, but certainly not all species \citep{Ryan-2006}.
 Important to emphasise that demographic patterns observed by \citep{Ryan-2006} (decreases in height growth rate, no decrease in biomass growth) predicted from theory here, without invoking hydraulic limitation.
@@
 excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
-\ref{fig:support_costs}).
+\ref{fig:assumptions}d).
 Note that similar increases in construction and maintenance costs with plant size are embedded within various models of plant function \citep[e.g.][]{Givnish-1988, Makela-1997}.
@@
 Dashed lines indicate slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}.
+The dark line shows the relationship assumed in this paper.
 \label{fig:assumptions}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/support_costs.pdf}
-\caption{\textbf{Amount of live stem tissue per unit leaf area increases with plant height}.
-Data are from the \textsc{baad} \citep{Falster-2015b}.
-Each dot is a single plant.
-Plants \textless{}  1m tall were assumed to have no heartwood.
-Lines show standardised major axis lines fitted by species, with intensity of shading adjusted according to strength of the relationship.
-Dashed lines indicate slope of 1, being the prediction of the model in Table \ref{tab:allometry}.
-The dark line shows the relationship assumed in this paper.
-\label{fig:support_costs}}
-\end{figure}
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/growth_light_height.pdf}
-\caption{\textbf{Effects of some traits on height growth changes with plant size.} Panels show predicted relationship between specific trait and height growth rate, for a plant of specified height and under a range of shading environments.
-\label{fig:growth_light_height}}
 \end{figure}
@@
 \centering
-\includegraphics{main/max_leaf_above.pdf}
-\caption{\textbf{The effect of traits on growth changes with size and
-light environment.} Traits moderate the responsiveness of growth to changes in light
-environmentThis figure needs to show more of key result with respect to to size.
-Possibly separate figure. \label{fig:shifts}}
+\includegraphics{main/growth_light_height.pdf}
+\caption{\textbf{Three different responses of growth rate to changes in size, with traits.} Panels show predicted relationship between specific trait and height growth rate, for a plant of specified height and under a range of shading environments.
+\label{fig:growth_light_height}}
 \end{figure}
+
+% \newpage
+
+% \begin{figure}[ht]
+% \centering
+% \includegraphics{main/max_leaf_above.pdf}
+% \caption{\textbf{The effect of traits on growth changes with size and
+% light environment.} Traits moderate the responsiveness of growth to changes in light
+% environmentThis figure needs to show more of key result with respect to size.
+% Possibly separate figure. \label{fig:shifts}}
+% \end{figure}
 
@@
 \includegraphics{main/max_leaf_above.pdf}
-\caption{\textbf{Low construction cost leads to shade intolerance,
-because of costs of high turnover.} Panels show effect of traits on
-maximum amount of shading that can be endured before net production (eqn
-\ref{eq:dbdt}) reaches zero. Lines indicate relationship for plants of a
-given height. \label{fig:wplcp}}
+\caption{\textbf{Effect of traits on shade tolerance.} Panels show effect of traits on maximum amount of shading that can be endured before net production (eqn \ref{eq:dbdt}) reaches zero.
+Lines indicate relationship for plants of a given height.
+\label{fig:wplcp}}
 \end{figure}

### 2016-09-06 R/axis.R

@@ -5,7 +5,7 @@ name_pretty <- function(trait) {
          narea=expression(paste("Nitrogen per leaf area (kg ", m^-2, " )")),
          hmat="Height at maturation (m)",
          diameter_stem_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
-         shading=expression(paste("Maximum leaf area above (", m^2, " )")),
+         shading="WPLCP (0-1)",
          leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, " )")),
 
          height_dt=expression(paste("Height growth (m ", yr^-1, " )")),

### 2016-09-06 R/figures.R

@@ -202,7 +202,7 @@ default_parameters <- function() {
     a_f1 = 0.8,
     a_f2 = 20,
     a_l1 = 2.17,
-    a_l2 = 0.546,
+    a_l2 = 0.5,
     k_l  = 0.4565855 / 3)
 
   FF16_Parameters(
@@ -301,9 +301,9 @@ figure_rate_vs_size_panels <- function(data, type, path) {
     switch(yvar,
       net_mass_production_dt=cols[1],
       fraction_allocation_growth=cols[2],
-      darea_leaf_dmass_live =cols[3],
+      darea_leaf_dmass_live =cols[8],
       area_stem_dt =cols[4],
-      area_leaf_dt =cols[5],
+      area_leaf_dt ="lightgreen",
       "black")
   }
 
@@ -483,28 +483,35 @@ lcp_whole_plant_with_trait <- function(x, height) {
 }
 
 figure_lcp_whole_plant <- function() {
-  op <- par(oma=c(0, 4, 1, 1), mar=c(4, 1, 1, 1), mfrow=c(3, 1))
+
+  mylabel <- function(txt) label(txt, -0.2, 1.2, xpd=NA, cex=1.5)
+
+  op <- par(oma=c(0, 2, 3, 0), mar=c(5, 3, 1, 1), mfrow=c(1,3))
+  lab = list(narea="a", lma= "b", rho = "c")
   for (trait in c("narea", "lma", "rho")) {
     xlim <- trait_range(trait)
-    ylim <- c(0.2, 5)
+    ylim <- c(0, 1)
 
-    heights <- c(0.5, 2, 10, 15, 20)
-    x <- seq_log_range(xlim, 60)
+    heights <- c(0.5, 5, 10, 20)
+    x <- seq_log_range(xlim, 100)
     cols <- color_pallete(length(heights) + 3)[-(1:3)]
 
     lcp <- sapply(heights, function(h)
                   lcp_whole_plant_with_trait(trait_matrix(x, trait), h))
-    lai <- log(lcp) / (-default_parameters()$k_I)
-
-    matplot(x, lai, type="l", lty=1, col=cols, log="xy", ylim=ylim,
-            xlab=name_pretty(trait), ylab=name_pretty("shading"),
-            yaxt="n")
+    matplot(x, lcp, type="l", lty=1, col=cols, log="x", ylim=ylim,
+            xlab="", ylab="", yaxt="n")
     axis(2, labels=TRUE, las=1)
-    if (trait == "lma") {
[…]

### 2016-09-06 R/figures_data.R

@@ -120,13 +120,14 @@ plot_with_sma_by_spp <- function(data, xvar, yvar, location=NULL, isoclines=NULL
 
   fit <- sma(data[[yvar]]~data[[xvar]]*data[["species"]], log="xy")
 
-  plot(fit, type='l', col=cols[9], p.lines.transparent=0.1, add=TRUE,..., lwd=1)
+  plot(fit, type='l', col=cols[8], p.lines.transparent=0.1, add=TRUE,..., lwd=1)
 
   invisible(fit)
 }
 
 
 figure_assumptions <- function(baad){
+
 
   data <- baad[["data"]]
 
@@ -138,34 +139,37 @@ figure_assumptions <- function(baad){
   i <- is.na(data[["sap"]]) & !is.na(data[["a.stba"]]) & !is.na(data[["h.t"]]) & data[["h.t"]] < 1
   data[["sap"]][i] <- data[["a.stba"]][i]/(1.2)
 
-  xlim <- c(5E-6, 1E4)
+  xlim <- c(1E-5, 1E4)
 
-  par(mfrow = c(3,1), mar=c(5, 5, 0.5, 0.5), oma=c(0,0,3,0))
+  mylabel <- function(txt) label(txt, -0.25, 1.1, xpd=NA, cex=1.5)
+  s <- default_strategy()
+  x <- seq_log_range(c(1E-6, 2E6), 100)
+
+  par(mfrow = c(2,2), mar=c(5, 7, 3, 0.5), oma=c(0,0,0,0))
   plot_with_sma_by_spp(data,"a.lf", "h.t",
     ylim=c(1.5E-3, 1.5E2), xlim=xlim,
-    ylab="Height (m)", xlab=NULL,location="topleft",
+    ylab="Height (m)", xlab=expression(paste("Leaf area (",m^2,")")), location="topleft",
     isoclines=0.5, A= seq(-10,10))
-  mtext("a) Architectural layout", 3, line=1)
+
+  lines(x, s$a_l1*x^s$a_l2, lwd=2)
+  mylabel("a) Architectural layout")
   legend("bottomright", legend = "Slope = 0.5", bty = "n", cex=1)
 
   plot_with_sma_by_spp(data,"a.lf", "sap",
     xlim=xlim, ylim=c(1E-9, 1E1),
-    ylab=expression(paste("Sapwood area (",m^2,")")), xlab=NULL,location="topleft",
[…]

### 2016-09-06 R/utils-plot.R

@@ -15,7 +15,86 @@ nice_colors <- function(n = 80) {
   cols[1:n]
 }
 
+# Color settings
+my_cols <- function() {
+  c("#01ABE9","#1B346C","#F34B1A")  # Zissou palette, from http://wesandersonpalettes.tumblr.com/
+}
+
 make_transparent <- function(col, opacity = 0.5) {
   tmp <- col2rgb(col)/255
   rgb(tmp[1, ], tmp[2, ], tmp[3, ], alpha = opacity)
+}
+
+
+axis.log10 <- function(side=1, horiz=FALSE, labels=TRUE, baseAxis = TRUE, wholenumbers=T, labelEnds=T,las=1, at=NULL) {
+
+  fg <- par("fg")
+
+  if(is.null(at)){
+
+    #get range on axis
+    if(side ==1 | side ==3) {
+      r <- par("usr")[1:2]   #upper and lower limits of x-axis
+    } else {
+      r <- par("usr")[3:4] #upper and lower limits of y-axis
+    }
+
+    #make pertty intervals
+    at <- pretty(r)
+    #drop ends if desirbale
+    if(!labelEnds)
+      at <- at[at > r[1] & at < r[2]]
+  }
+  #restrict to whole numbers if desriable
+  if(wholenumbers)
+    at<-at[is.wholenumber(at)]
+
+  lab <- do.call(expression, lapply(at, function(i) bquote(10^.(i))))
+
+  #convert at if
+  if(baseAxis)
+    at<-10^at
+
+  #make labels
+  if ( labels )
+    axis(side, at=at, lab, col=if(horiz) fg else NA,
+         col.ticks=fg, las=las)
+  else
+    axis(side, at=at, FALSE, col=if(horiz) fg else NA,
+         col.ticks=fg, las=las)
+}
+
+is.wholenumber <-  function(x, tol = .Machine$double.eps^0.5) {
+  abs(x - round(x)) < tol
+}
+
+colour.by.category <- function(x, table) {
+ unname(table[x])
+}
+
+
+label <- function(text, px=0.03, py=NULL, ..., adj=c(0, 1)) {
+  if (is.null(py)) {
+    fin <- par("fin")
+    r <- fin[[1]] / fin[[2]]
+    if (r > 1) { # x is longer.
+      py <- 1 - px
+      px <- (1 - py) / r
+    } else {
[…]

### 2016-09-16 ms/ms-supp.Rnw
Towards a full draft

@@
 
-\titleprefix{Supporting information for:}
+\titleprefix{Supporting Materials for:}
 
@@
 \caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (diameter growth).} Panels show predicted relationship between
-specific trait and diameter growth rate, for a plant of specified
+in light environment (stem-diameter growth).} Panels show predicted relationship between specific trait and diameter growth rate, for a plant of specified
 diameter and under a range of shading environments.
 \label{fig:growth_light_dia}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\centering
+\includegraphics{SI/growth_light_area.pdf}
+\caption{\textbf{Traits moderate the responsiveness of growth to changes
+in light environment (stem-area growth).} Panels show predicted relationship between specific trait and diameter growth rate, for a plant of specified
+diameter and under a range of shading environments.
+\label{fig:growth_light_area}}
 \end{figure}
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lma_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between leaf-mass per unit leaf area
-and growth rate changes with plant size.} \label{f-lma_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lma_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between leaf-mass per unit leaf area
-and relative growth rate changes with plant size.}
-\label{f-lma_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/rho_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between wood density
-and growth rate changes with plant size.} \label{f-rho_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/rho_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between wood density
-and relative growth rate changes with plant size.}
-\label{f-rho_growth_size_relative}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/hmat_effects_at_diameters.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-growth rate changes with plant size.} \label{f-hmat_growth_size}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/hmat_effects_at_diameters_relative.pdf}
-\caption{\textbf{The expected correlation between maximum height and
-relative growth rate changes with plant size.}
-\label{f-hmat_growth_size_relative}}
-\end{figure}
-
 \clearpage

### 2016-09-16 ms/ms.tex
Towards a full draft

@@
 
-\tableofcontents
+% \tableofcontents
 % Consider Ecol Letters "Ideas and Perspectives": maximum of 7500 words (main text), and 10 figures, tables or boxes.
@@
 
+By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation linking trait-based trade-offs to demography and thus provide a platform for understanding growth across diverse species around the world.
+
 \section{Introduction}
 
-Functional traits capture core differences in the strategies plants use to generate and invest surplus energy \citep{Wright-2004, Chave-2009, Westoby-2002}.
-Although most plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in rates at which resources are acquired and invested into different tissues.
-During the last decade, trade-offs related to prominent traits have been quantified and values for some traits -- such as leaf mass per leaf area, wood density, and seed size -- now exist for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
-Concomitantly, researchers have been using traits as potential indicators of the growth strategy and ecology of a species.
+In this article we outline how the growth of individual plants is influenced by various functional traits, also the plant's size and light environment.
+Functional traits capture core differences in the strategies plants use to generate and invest surplus energy \citep{Westoby-2002, Wright-2004, Chave-2009}.
+Although most woody plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in the rates at which resources are acquired and invested into different tissues.
+During the last two decades, trade-offs related to some prominent traits have been quantified, with values for traits such as leaf mass per unit leaf area, wood density, and seed size now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
+As data has accumulated, researchers have looked to traits as potential indicators of plant growth and demography \citep{Poorter-2008, Wright-2010, Adler-2014}.\footnote{Add references from applied fields, such as invasive species, conservation.}
 
-While the influence of plant traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have only met with mixed success \citep{Wright-2010, Poorter-2008}.
-In seedlings, the trait leaf mass per leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}.
-Leaf mass per leaf area and its correlate leaf lifespan were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
+While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
+In seedlings, the trait leaf mass per unit leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}.
+leaf mass per unit leaf area and its correlate leaf lifespan were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
 These early successes prompted researchers to search for similar relationships in large plants.
-However, the results showed that in trees leaf mass per leaf area was not correlated with diameter growth rate \citep{Poorter-2008, Wright-2010, Herault-2011}.
-Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010}, but less so in small plants \citep{Castro-1998}.
-In summary, traits did not correlate simply and consistently to growth rate.
+However, the results showed that in saplings and trees, leaf mass per unit leaf area was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
+Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
+So traits do not correlate simply and consistently to growth rate.
 
-The challenge of interpreting diverse empirical results seeking to link traits to growth rate is all the harder because while we lack clear theoretical expectations.
-Generating such expectations is one of the primary roles for theory \citep{Kokko-2007}.
-A widely-used model shows that for seedlings, mass-based relative growth rate may be linearly and negatively related to leaf mass per leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-An extension of the model suggests a similar relationship may hold at larger sizes \citep{Enquist-2007}.
-But as described above, this prediction is not supported by empirical results.
+It is now becoming increasingly clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
+A recent meta-analysis of 103 studies reporting \textgreater 500 correlations by \citet{Gibert-2016} provides the most compelling evidence.
+\citet{Gibert-2016} showed that the strength of the correlation between some traits (including leaf mass per unit leaf area and maximum height) and growth rate was modified by size, while for other traits (including as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
+Importantly, the direction of the correlations and their shifts with size supported predictions generated via a framework describing how traits influence growth.
+
+The challenge of interpreting diverse empirical results seeking to link traits to growth rate was all the harder until the paper by \citet{Gibert-2016}, because we lacked any clear theoretical expectations on why the effect of traits may be moderated by size.
+Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
+A widely-used model for seedlings suggests a seedling's relative growth rate in mass is linearly and negatively related to the trait leaf mass per unit leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+An extension of the model suggests a similar relationship should hold at larger sizes \citep{Enquist-2007}.
+But as described above, the prediction for large plants is not supported by empirical results.
 Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
-Without further guidance, many researchers have simply assumed traits to linearly map onto growth rates, i.e.
-there is a ``fast'' and ``slow'' growth end of the trait spectrum.
+Without further guidance, many researchers expected traits to linearly map onto growth rates, i.e.
+there is a ``fast'' and ``slow'' growth end for each trait \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
 
-A further problem with existing theory is that the effects of traits are all realised via influences on mass production \citep{Enquist-2007}, whereas the influence of some traits such as leaf mass per leaf area and wood density is via allocation to different tissues rather than mass production \citep{Falster-2011, Gibert-2016}.
-A second issue with theory focussed solely on mass production is that measuring mass production is only practical for small plants that can be easily harvested.
-For larger plants, growth is more often measured as increment in either diameter or height.
-For example, natural research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-As we outline further below, mass production is only one component of plant growth; meaning theory focussing solely on mass production will always be limited.
+One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on mass production \citep{Enquist-2007}.
+By contrast, the physiology of traits such as leaf mass per unit leaf area and wood density suggests they have influence by altering the allocation of energy among different tissues, rather than changing mass production \citep{Falster-2011, Gibert-2016}.
+A second concern for theory focussed on mass production is that altered mass production is only possible outcome.
+Measuring mass production is really only practical for small plants that can be easily harvested.
+For larger plants, growth is more-often measured as increment in either diameter or height.
+For example, research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
+Theory focussing on mass production is therefore of limited potential use.
 
-Ultimately, trait based understanding must be integrated within a more general model of plant growth, also able to capture patterns of growth through ontogeny.
+In this paper we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth (Table \ref{tab:phenomena}); and thereby offer fresh insight into the way traits influence plant demography throughout across the life-cycle.
+This model can be viewed as an integration of several related approaches \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
+Our primary focus in this article is to explain the pattern that has been gradually emerging: that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
+As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth.
+Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of energy uptake and allocation, why it makes sense that the effect of some traits changes with size.
+We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
+We submit that, trait-based approaches -- which focus on differences among species -- should seek to be integrated within a general model of plant growth, also able to capture patterns of growth through ontogeny.
 Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
-In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover.
+In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
 And all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Meanwhile shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of maintaining stems \citep{Givnish-1988}, and also vary with traits \citep{Lusk-2006,Poorter-2006}.
+Shade tolerance also varies among species, correlates with traits \citep{Lusk-2008,Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Lusk-2008}.
 These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.
-
-In this paper we show how a new mechanistic framework -- called {\plant} \citep{Falster-2016} -- offers fresh insight into the way traits influence plant growth throughout across the life-cycle and explain diverse phenomena (Table \ref{tab:phenomena}).
-Our primary aim is to explain a pattern that has been gradually emerging and recently solidified: that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-A recent meta-analysis of 103 studies reporting \textgreater 500 correlations by \citet{Gibert-2016} provides the most compelling evidence.
-\citet{Gibert-2016} showed that the sign of the correlation between some traits (such as leaf mass per leaf area and maximum height) and growth rate was modified by size, while for other traits (such as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Importantly, the direction of the correlations and their shifts with size supported predictions generated via a framework describing how traits influence growth.
-Here we extend the results of taken by \citet{Gibert-2016} to provide a complete, functioning model and explain, physiologically, why it makes sense that the effect of traits changes with size.
-By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation linking trait-based trade-offs to demography and thus provide a platform for understanding growth across diverse species around the world.
 
@@
 
-We build on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within the plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
+The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
 \ref{fig:conceptual}a).
-Here we consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, with the subscript $i$ indicating: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
+We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
 The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A full list of variables, units and definitions is given in Table \ref{tab:definitions} and Table \ref{tab:traits}, with further detail on the parameter values applied given in Table \ref{tab:params}.
+A summary of all variables, units and definitions is given in Table \ref{tab:definitions}-\ref{tab:traits}, with further details on the parameter values applied given in Table \ref{tab:params}.
 
 We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant.
-According to a standard model, the amount of biomass available for growth, ${\rm d}B / {\rm d}t$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig.
+Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}}t$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig.
 \ref{fig:conceptual}a, eqn \ref{eq:dbdt}).
-Photosynthesis is proportional to the average photosynthetic rate per unit leaf area $\bar{p}$ and total leaf area, $A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per leaf area.
+Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}$) and total leaf area ($A_{\rm l}$).
 The constant $y$ in eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass.
-We assume that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ by trait (see below).
+We assume that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ with traits (see below).
 While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
-as functions of temperature, such detail is not needed here.
-Even with detailed models, ${\rm d}B / {\rm d}t$ always comes down to simple subtraction of income and losses.
+as functions of temperature, such detail is not needed here, as even with detailed models, $\frac{{\rm d}B}{{\rm d}}t$ always comes down to simple subtraction of income and losses.
 
@@
 
-To model growth in either height, leaf area, basal area, standing mass or stem diameter requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
+To model growth in either height, leaf area, basal area, standing mass or stem diameter then requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
 Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}.
-Relevant equations (eqns \ref{eq:dbdt}-\ref{eq:dDdt}) are given Fig.
-\ref{fig:conceptual}b.
-Importantly, most of these terms vary intrinsically with size.
+Relevant equations are given in Fig.
+\ref{fig:conceptual}b, (eqns \ref{eq:dbdt}-\ref{eq:dDdt}).
+As is evident in eqn  Fig.
+\ref{fig:conceptual}b, the growth rate in plant weight (eqn \ref{eq:dMtdt}), leaf area (eqn \ref{eq:daldt}), height (eqn \ref{eq:dhdt}), stem basal area (eqn \ref{eq:dast}), and stem diameter (eqn \ref{eq:dDdt}) all share some terms in common.
+Many of the terms in Fig.
+\ref{fig:conceptual} also vary intrinsically with size.
 The insets in Fig.
-\ref{fig:conceptual}b show size-related patterns in different variables for a typical plant obtained from applying the allocation model derived below.
-However, it is important to recognise that eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are mathematically true, derived using the chain rule of differential calculus, and thus must hold true for any class of models where carbon allocation is important.
+\ref{fig:conceptual}b show the size-related patterns in different variables for a typical plant, obtained from applying the specific allocation model in the next section.
 
-Note that the growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
-The remaining fraction,$1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$, is allocated to reproduction.
-$\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, presumably at 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants.
-It is important to realise that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
-So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even while the net amount of live mass $M_{\rm a}$ does not increase any further.
+The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
+The remaining  $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
+Presumably, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, close to 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants.
+Note also that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
+So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ is not increasing.
 
-The growth rate in total standing mass of the plant, $\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$ is then sum of sapwood turnover (which is replaced by an equivalent amount) and growth in live mass (eqn \ref{eq:dMtdt}).
+The growth rate in the total standing mass of the plant, $\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$ is then the sum of sapwood turnover (which is replaced by an equivalent amount) and increment in live mass (eqn \ref{eq:dMtdt}).
 
-The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and support tissues of bark, sapwood and roots (eqn \ref{eq:daldmt}).
+The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood and roots (eqn \ref{eq:daldmt}).
 The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of tissue-level construction costs per unit of leaf area, with one of these being $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$: the trait leaf mass per area.
@@
 Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
-Meanwhile, basal area ($A_{\rm st}$) increment can be expressed as the sum of increments in sapwood, bark \& heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
-These in turn are related to ratios of sapwood or bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
-Finally, diameter growth is then given by the geometric relationship between stem diameter ($D$) and $A_{\rm st}$ (eqn \ref{eq:dDdt}).
+
+The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
+These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
+
+Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
+Note that we make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integrating of stem turnover (eqns \ref{eq:ha}- \ref{eq:htm}).
 
 \subsection{Shade tolerance}
-A useful measure of a plant's shade tolerance is the point at which photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988,Baltzer-2007,Lusk-2013}, sometimes referred to as the the ``whole-plant light compensation point'' (\wplcp{}).
-In general, the average assimilation rate per leaf area $\bar{p}$ is expected to increase with light level or canopy openness, $E$.
-The \wplcp can be estimated by solving the value of $E=E^*$ where $\frac{{\rm d}B} {{\rm d}t}(E)$ from eqn \ref{eq:dbdt} is zero, which occurs when:
+
+Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance, sometimes referred to as the the ``whole-plant-light-compensation point'' ({\wplcp}): the  the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013}.
+In general, average assimilation rate per leaf area $\bar{p}$ is expected to increase with light level or canopy openness, $E$.
+The {\wplcp} can be estimated by solving the value $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$.
+From eqn \ref{eq:dbdt}, this occurs when
 \begin{equation}\label{eq:wplcp}
-\bar{p}(E^*) =\frac{\sum \, M_{\rm i} \left(k_{\rm i} y + r_{\rm i}\right)}{A_{\rm l}}.
+\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}},
 \end{equation}
-Here $\bar{p}(E)$ is the rate of photosynthesis at canopy openness $E$. Below we express
+where $\bar{p}(E)$ is the rate of photosynthesis at canopy openness $E$.
+The {\wplcp} occurs at the points where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig.
+\ref{fig:wplcp_idea}).
+Traits will influence the {\wplcp} if they effect either income or costs.
+Also, since the amount of stem support naturally increases with height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig.
+\ref{fig:wplcp_idea}).
 
@@
 
-As noted above, most of eqns in Fig.
-\ref{fig:conceptual} arise as mathematical truisms.
-To make predictions from eqns \ref{eq:dbdt}-\ref{eq:dDdt}, however, requires a specific model of plant construction.
-The model described below can be considered a first-order functional-balance model.
-We might also call it ``isometric'', because the assumptions ensure the proportional relationships among the relevant variables are preserved as the plant grows.
-Table \ref{tab:allometry} provides key equations of the model (see Supplementary material \ref{sec:ff16} or \citet{Falster-2016} for full derivation).
+It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are derived derived using the chain rule of differential calculus, they are mathematically true, and must therefore hold for a wide range of potential growth models.
+To make explicit predictions via this framework, however, requires a explicit model of plant construction and function; i.e.
+we must put equations behind all there terms in Fig.
+\ref{fig:conceptual}b.
+
+Here we adopt a model of plant construction and function that can be considered a first-order functional-balance model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
+We could also call it ``isometric'', because the assumptions preserve the proportions among the relevant variables as the plant grows.
+Table \ref{tab:allometry} provides key equations of the model (see Supplementary Material \ref{sec:ff16} or \citet{Falster-2016} for full derivation).
 In particular, we assume:
@@
 \item[Assumptions 4:] A log-log scaling relationship with slope of 1.0 between the leaf area and mass of fine roots in a plant (eqn \ref{eq:mr}).
-\item[Assumption 5:] The vertical distribution of leaf area within the canopy of an individual plant remains relatively constant through ontogeny.
+\item[Assumption 5:] The vertical distribution of leaf within the canopy of an individual plant remains constant through ontogeny.
 \end{description}
+Assumption 1 accounts for the architectural layout of the plant.
+Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
+To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).
 
-Assumption 1 accounts for the architectural layout of the plant.
-Assumptions 2-4 are implementations of the pipe model \citep{Shinozaki-1964}, whereby the area of conducting tissues are proportional to leaf area.
-To describe the vertical distribution of leaf area within the canopy of an individual plant, we follow the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1.
-Low values result in a conical canopy, as seen in many conifers, while higher values give a top-weighted canopy profile similar to those seen among angiosperms.
+Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height, described in Table \ref{tab:allometry}b.
+Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
+\ref{fig:conceptual}.
 
-The functional-balance assumptions are well-supported by data, at least to first approximation.
-Fig.
-\ref{fig:assumptions} shows data sourced from the Biomass and Allometry Database (\textsc{baad}) \citep{Falster-2015b}, which includes records for 21084 individuals on 656 species for various size metrics of individual plants.
-Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a.
-As is evident, the lines fitted within each species roughly align with predicted functional-balance assumption Fig.
-\ref{fig:assumptions}a-c.
-
-Combined, these assumptions in Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
-Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives a the necessary components in Fig.
-\ref{fig:conceptual} and thereby an explicit model for plant growth in relation to size.
-Note also that no assumptions are made about the diameter and total mass of the stem in relation to other components, e.g.
-height: these relationships arise as emergent properties, via integrating stem turnover (eqns \ref{eq:ha}- \ref{eq:hm}).
-
-The most important prediction arising from the eqns in Table \ref{tab:allometry} is that the amount of live tissue supporting each unit of leaf area increases with plant height.
-We can derive an explicit prediction in relation to the assumed allocation model:
+An important prediction arising from the eqns in Table \ref{tab:allometry} is that the amount of live stem tissue supporting each unit of leaf area increases linearly with  height:
 \begin{equation}\label{eq:msml}
-\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
+\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H,
 \end{equation}
-Note that $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are traits, i.e.
-properties that are assumed approximately constant through ontogeny, so that their product is a constant.
-Again, the predicted relationship is supported by available data from the \textsc{baad} (Fig.
-\ref{fig:assumptions}d).
+where $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits, i.e. properties that are assumed approximately constant through ontogeny.
 
@@
 
-Now consider how trait-based trade-offs can be embedded into the above growth model.
+We now consider how trait-based trade-offs enter into the above growth model.
 It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect never-ending directional trait values towards more beneficial values.
 For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}.
-In postulating potential benefits and costs, we only consider direct biophysical consequences of varying a trait, i.e.
-we do not try to account for every single variable that may sometimes correlate with a given trait, only those for which there is thought to be an inevitable trade-off.
-The trade-offs described in Table \ref{tab:traits} are then implemented as follows.
+In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait.
+The trade-offs are then implemented as follows.
 \begin{description}
-\item[Seed mass] Effects of the trait seed mass, denoted by $\omega$, are naturally embedded in the equations determining fecundity and size of seeds.
-We assume larger seeds result in larger seedlings, thus the energetic tradeoff arising from seed size translates into a demographic tradeoff between the number of seedlings produced by a parent and the initial size of seedlings.
-\item[Height at maturation] ........
-\item[Nitrogen per unit leaf area] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area $\nu$, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
-Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
-\item[Leaf mass per unit area] The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$.
-In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-\item[Wood density] The trait wood density, denoted by $\rho$, directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$.
-The cost of cheaper wood is increased mortality rate, which is not included in the growth equations but itself an important demographic outcome.
-The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as $r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},$ where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood.
-Similarly, the rate of bark respiration per unit mass varies as $r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},$ with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.
+\item[Seed mass, $\omega$:] There is a direct energetic trade-off between the a plant's fecundity and the size of it's seeds.
+Moreover, we assume larger seeds result in larger seedlings; the seed-size number tradeoff thereby translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
+
+\item[Height at maturation, $H_{\rm m}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction.
+We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{\rm m}$ via a function $r(H, H_{\rm m}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{\rm m}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
+% By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{\rm m}$ (Fig. \ref{fig:S3}).
+
+\item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
+Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.\footnote{Add plot of this relationship, from data in \citet{Wright-2004}.}
+
+\item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
+The rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
+
+\item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
+The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly  whenever $\rho$ is varied.
  \end{description}
 
-\section{Analysis}
+\section{Methods}
 
-The 
-Parameters
+The growth model described above has been implemented as the \texttt{FF16} physiological model within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
+The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
+To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits.
+Parameters used in the current analysis were sourced from \citet{Falster-2016}, with the exception of XXXX (see Table \ref{tab:params} for values).
 
-\citep{Falster-2016}: \plant
+The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database (\textsc{baad}) \citep{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
+We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
+We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed to come from normal distributions.
+The means and variances of these distributions are then fit as part of the model-fitting procedure.
+
+The analyses presented employ best practises in scientific computing, as defined by \citep{Wilson-2014}, and are fully reproducible, via code available at
+\href{https://github.com/traitecoevo/growth\_trajectories}{github.com/traitecoevo/growth\_trajectories}.
+
+\section{Results}
+
+\subsection{Model assumptions}
+
+The functional-balance assumptions outlined in Table \ref{tab:allometry}a are well-supported by the available data, at least to first degree.
+Fig.
+\ref{fig:assumptions} shows data sourced from the \textsc{baad} \citep{Falster-2015b}, with solid lines indicating {\sma} lines fit to each species in the data.
+Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a.
+As expected, species differed in elevation, but less so in the slope of the fitted lines, with slopes aligning with those predicted by the functional-balance assumptions.
 
@@
 
-This new growth model captures the intrinsically size-dependent nature of plant growth, recovering well-known empirical patterns (Table \ref{tab:phenomena}).
+Our growth model suggests an intrinsically size-dependent pattern of growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
+The panels in Fig.
+\ref{fig:conceptual} show the expected patterns for a typical woody plant.
 Height growth shows a hump-shaped pattern with size, first increasing then decreasing.
@@
 \ref{fig:conceptual}), and increasing reproductive allocation.
-In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover.
-Diameter growth shows a weakly hump shaped curve \citep{Herault-2011}, tapering off slightly at larger sizes, in part because of allometric effects (eqn \ref{eq:dDdt}, but mainly because of increased reproductive allocation in older trees (Fig?).
-All growth measures decrease sharply with size when expressed as relative growth rates \citep{Iida-2014}.
+In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
+Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}, and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
+All growth measures decrease sharply with size when expressed as relative growth rates.
 
-\subsection{How traits influence growth rate}
+% Importantly, we can remove these size dependent effects by altering XXX of the assumptions in Table \ref{tab:allometry} (??).
+% \footnote{What are key assumptions that induce all size-related effects?
+My guess is  1) changing sapwood area so that volume scales isometrically with height will do it, and 2) removing reproductive allocation.
+Add figure in SI?}
 
-Distinguish 3 types of trait.
+\subsection{Changes in height growth rate with trait}
 
-Growth traits
+We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}).
+We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
+For the five different traits, we observed four relatively distinct types of response.
+These responses are summarised in Table \ref{tab:responses} and described further below.
+The first two traits considered modify behaviour primarily at the start and end of an individuals growth trajectory, and are therefore termed ``ontogenetic traits''.
+The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individuals growth trajectory.
 
-For growth traits can derive rule:
-\begin{equation}\label{eq:dg4}
-\frac{\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}} {\frac{ {\rm d}B} {{\rm d}t}} < \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
-In words, there will be a negative relationship between trait x and growth if the relative change in production due to change in traits is less than the relative change in the total construction cost for a unit of leaf area due to change in trait.
-Crucially, key elements in eqn \ref{eq:dg3} are modified by height.
+\subsubsection{Ontogenetic traits}
 
-We can also solve for trait values $x$ maximising growth rate, i.e. the value giving $\partial G /\partial x = 0$:
-\begin{equation}\label{eq:max}
-\frac{\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} } {\frac{ {\rm d}B} {{\rm d}t}} = \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
+\noindent{\bf Seed size, $\omega$:}
+Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease.
+As such, the only effect of seed size on growth comes from changing the plant's initial  size.
+The plots in Fig.
+\ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
+Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in diameter, height, and mass), because of their greater total leaf area.
+At the same time, relative growth rate is predicted to decrease with size, because the ratio of leaf area to support mass decreases with plant size.
+As plants grow, differences in initial mass will decrease in importance, relative to other factors influencing growth through the life-cycle.
+We therefore expect the correlations between seed size and growth rate observed for seedlings to disappear among larger plants.
 
-\subsubsection{Seed size, $M_{\rm t,0}$}
+\noindent{\bf Height at maturation, $H_{\rm m}$:}
+$H_{\rm m}$ moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
+Greater height at maturation ($H_{\rm m}$) can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
+\ref{fig:growth_light_height}).
+At smaller sizes, there is no differentiation among species, because most individuals are focusing on growth.
+At larger sizes, individuals of some species are allocating a majority of their surplus energy to reproduction, which reduces their growth rate relative to those species with greater $H_{\rm m}$(Fig.
+\ref{fig:growth_light_height}).
 
-Seed size: increases starting size
+\subsubsection{Development traits}
 
-\subsubsection{Height at maturation, $H_{\rm m}$}
+The remaining three traits ($\nu$, $\phi$, \& $\rho$) moderate growth rates at a point along an individuals growth trajectory (Fig. \ref{fig:growth_light_height}).
 
-Height at maturation: ($h_m$) moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
+% To understand the response of growth to these traits the following derivation is helpful.
 
-Greater height at maturation ($h_m$) can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ .
-At smaller sizes, there is no differentiation among species, because most species are focusing on growth.
-At larger sizes, individuals of some species are allocating a majority of their surplus energy to growth, leaving a strong signal in potential growth rate.
+% Consider the effect of traits on the components of height growth rate (eqn \ref{eq:dhdt}).
+To make analysis more tractable, assume we are dealing with a plant investing all its energy in growth, i.e.
+$\frac{{\rm d}M_{\rm a}}{{\rm d}B}=1$.
+Note also that none of  the traits considered here influence $\frac{{\rm d}H} {{\rm d}A_{\rm l}}$.
+Therefore,
+% \begin{equation} \label{eq:G2}
+% \frac{{\rm d}H}{{\rm d}t} \propto \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}  \times \frac{ {\rm d}B} {{\rm d}t}.
+% \end{equation}
+% The correlation of traits with growth at a given height will be given by the derivative of eqn \ref{eq:G2} with respect to some trait $x$.
+The derivative of $\frac{{\rm d}H}{{\rm d}t}$ with respect to some trait $x$ is then \footnote{Eqn \ref{eq:G2} is a product of the form $Y(x) = W(x) \, Z(x)$, whose derivative is $\frac{\partial{Y}}{\partial{x}} = \frac{\partial{W}(x)}{\partial{x}} \, Z(x) + W(x) \, \frac{\partial{Z}(x)}{\partial{x}}$.}
+% \begin{equation} \label{eq:dg1}
+% \frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x} \propto
+% \frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}{\partial x}
+%  \, \frac{ {\rm d}B} {{\rm d}t}
+% + \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
+% \, \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}.
+% \end{equation}
+% If this derivative is positive (negative), there will be a positive (negative) relationship between traits and growth at the height where it is evaluated.
+We can simplify eqn \ref{eq:dg1} further by substituting $\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}
+% {\partial x} = -{\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}}^2
+% \, \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
+% }{\partial x}$ (from eqn \ref{eq:daldmt}). Noting also that $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}>0$ and $\frac{{\rm d}B} {{\rm d}t} > 0$, we find that
+% \begin{equation}\label{eq:dg2}
+% \frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x}  \propto \frac{
+% \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
+% {\frac{ {\rm d}B} {{\rm d}t}} - \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
+% }{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
+% \end{equation}
 
-\citet{Wenk-2015}
+% Eqn \ref{eq:dg2} implies that there will be a negative relationship between trait $x$ and height growth if the relative change in net mass production, due to a change in trait, is less than the relative change in the total construction cost per unit leaf area, due to a change in trait.
+We can also solve for trait values $x$ maximising growth rate by solving for the value of $x$ that gives $\frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x}= 0$.
+Crucially, key elements in eqn \ref{eq:dg2} will be modified by height.
+With these general results in hand, let us now consider the influence of three traits on height growth rate.
 
-\subsubsection{Leaf mass per leaf area, $\phi$}
+\noindent{\bf Leaf nitrogen content per leaf area, $\nu$}: The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given for any given light environment $E$ and does not vary with height (Fig.
+\ref{fig:growth_light_height}).
+As $E$ increases from low to high, the optimal $\nu$ also increases.
+The invariance of the growth-trait relationship with respect to size can be understood as follows.
+The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rates.
+Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$
+%  $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}$.
+From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial nu} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial nu}  - \frac{ \partial r_{\rm l}}{\partial nu} )$.
+Since both and $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of plant height.
 
-By definition, the main effect of increasing $\phi$ is to decrease the marginal leaf deployment per mass (eqn \ref{eq:daldmt}) and thus reduce growth rate.
-However, to fully account for the influence of $\phi$ on growth we must also account for the decrease in leaf turnover that results from superior leaf construction \citep{Wright-2004}.
-This is achieved by linking the turnover rate of leaf in eqn \ref{eq:dbdt} to $\phi$.
-We use an allometric scaling relationship of the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been shown to describe patterns across diverse ecosystems \citep{Wright-2004} (Fig.
-\ref{fS-leaf}).
-
-Unlike eqn \ref{eq:RGR}, eqs.
-\ref{eq:dhdt}, \ref{eq:dast} and \ref{eq:dDdt} predict a relationship between $\phi$ and growth that changes with plant size ( \ref{fig:lmA_growth_size}).
-Decreasing $\phi$ has two impacts on growth rate.
-First, lower $\phi$ increases marginal leaf deployment ($\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ by economising on construction costs.
-Second, lower $\phi$ decreases net production (${\rm d}P / {\rm d}t$), due to increased leaf turnover.
+\noindent{\bf leaf mass per unit leaf area, $\phi$}:
+Unlike $\nu$, the response of growth rate to changes in $\phi$ varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
+\ref{fig:growth_light_height}).
+As a result, the value of $\phi$ that optimises plant growth increases with plant size, and the direction of correlation between height growth rate and $\phi$ shifts from negative to positive, as plant's get taller.
+Decreasing $\phi$ has two impacts on height growth rate.
+First, lower $\phi$ increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$ by economising on construction costs.
+Second, lower $\phi$ decreases net production $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
 Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects.
@@
 The net result is that at larger sizes, low $\phi$ is no longer advantageous for growth (Fig.
-\ref{fig:lmA_growth_size}d).
+\ref{fig:growth_light_height}).
 
-If the trait of interest is leaf-mass per unit leaf area, i.e.
-$x=\phi$, we can expand eqn \ref{eq:dg4} as follows.
-Note that $\frac{\partial \left(\frac{{\rm d}M_{\rm t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$.
-Also, from eqn \ref{eq:dbdt} we obtain,
-\begin{equation}\label{eq:dbdt2}
-\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
-\end{equation}
-where $c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_{i={\rm b,s,r}}{M_{i} \, r_{i}}) - (\sum_{i={\rm b,s,r}}{M_{i} \, ???})$is independent of $\phi$, and thus
-\begin{equation}\label{eq:dbdt3}
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
-(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
-\end{equation}
-eqn \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent of $\phi$.
-In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ {\rm d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.
+% Substituting $x=\phi$ into eqn \ref{eq:dg2}, we can come up with a more precise expressions for some of the terms.
+Note that we are evaluating these derivatives at a given value of $H$ and therefore $A_{\rm l}$ (via the allometry in eqn \ref{eq:ha}).
+Varying $\phi$ therefore changes $M_{\rm l}$, not $A_{\rm l}$.
+We can then reorganise eqn \ref{eq:dbdt} as,
+% \begin{equation}\label{eq:dbdt2}
+% \frac{{\rm d}B}{{\rm d}t} = c_1 - c_2  \, \phi^{1 -\beta_{{\rm kl2}}} \end{equation}
+% where
+% $c_1 = y \, A_{\rm l} \, (\bar{p}(E) - \nu \beta_{{\rm lf4}}) - \sum_{i={\rm b,s,r}} M_{i} \left(y\, r_{i} + k{i}\right)$
+% and
+% $c_2 =  A_{\rm l} \, \beta_{{\rm kl1}} \, \phi_0^{\beta_{{\rm kl2}}}$ are independent of $\phi$.
+% % The derivative of mass production with respect to $\phi$ is therefore
+% % \begin{equation}\label{eq:dbdt3}
+% % \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
+% % (\beta_{{\rm kl2}}-1) \phi ^{-\beta_{{\rm kl2}}}.
+% % \end{equation}
+% Eqn \ref{eq:dbdt2} indicates that if $\beta_{{\rm kl2}}=1$, mass production rate $\frac{ {\rm d}B} {{\rm d}t}$ will be independent of $\phi$.
+In that case, lower $\phi$ will always give a growth advantage, irrespective of plant size.
+However,
+% evidence strongly suggests $\beta_{{\rm kl2}}>1$ \citep[see Fig.
+\ref{fS-leaf}, after][]{Wright-2004}.
+In that case, $\frac{ {\rm d}B} {{\rm d}t}$ will increase with $\phi$, because of lower losses via turnover.
+As $\phi$ increases, the relative increase in $\frac{ {\rm d}B} {{\rm d}t}$  eventually outweighs the relative change in $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$, causing the direction of correlation between growth rate and trait to shift from negative to positive.
+Moreover, the value of $\phi$ at which this shift occurs, which is also the value of $\phi$ optimising growth rate, is higher at greater sizes.
 
-Substituting into eqn \ref{eq:dg4}, we have the result that growth rate correlates negatively with traits if
-\begin{equation} \label{eq:G6}
-\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
-< \frac{1}{\phi
- + \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
-\end{equation}
+\noindent{\bf Wood density, $\rho$}:
+As for leaf, cheaper stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates (Fig.
+\ref{fig:growth_light_height}).
+In contrast to $\phi$, the benefits of cheaper stem construction become more pronounced at larger sizes, as more of the plant becomes wood (Fig.
+\ref{fig:conceptual}a).
+Thus, $\rho$ has only a weak effect on growth rates for small plants, but a strong effect for intermediate to large plants.
+As the plant becomes very tall (H \textgreater 15m), the benefits start to low $\rho$ finally begin to diminish.
 
-\subsubsection{Leaf nitrogen content per leaf area, $\nu$}
+% Substituting in $x=\rho$, we can come up with a more precise expressions for some of the terms in eqn \ref{eq:dg2}.
+Note that we are evaluating these derivatives at a given value of $H$ and therefore $A_{\rm l}$ and thus the volume of bark and sapwood (via the allometry in eqns \ref{eq:ha}--\ref{eq:ab}).
+Varying $\rho$ therefore changes $M_{\rm s}$, not $A_{\rm st}$.
+We can then reorganise eqn \ref{eq:dbdt} as,
+% \begin{equation}\label{eq:dbdt3}
+% \frac{{\rm d}B}{{\rm d}t} = c_3 - c_4 \, \rho^{1-\beta_{{\rm ks2}}}
+% \end{equation}
+% where
+% $c_3 = y \, A_{\rm l} \, (\bar{p}(E) - \sum_{i={\rm l,b,s,r}} M_{i}\, r_{i})  - sum_{i={\rm l,r}} M_{i} k{i}$
+% and
+% $c_4 =  (1+b) \theta \,  \eta_c \, A_{\rm l} \, H \,  \beta_{{\rm ks1}} \, \rho_0^{\beta_{{\rm ks2}}}$ are independent of $\rho$.
+% Eqn \ref{eq:dbdt3} indicates that if $\beta_{{\rm ks2}}=1$, mass production rate $\frac{ {\rm d}B} {{\rm d}t}$ will be independent of $\rho$.
+In that case, lower $\rho$ will always give a growth advantage, irrespective of plant size.
+In this analysis we assume $\beta_{{\rm ks2}}=1.25$, so that mass production decreases slightly as $\rho$ decreases.
 
-\subsubsection{Wood density, $\rho$}
+\subsection{Changes in other growth rates with trait}
 
-As for leaf, cheaper stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates.
-However, the cost of cheap stem construction must also be accounted for.
-However, the direct physiological trade-offs of lower $\rho$ are less understood than for $\phi$.
-One possibility is that lower $\rho$ decreases mechanical strength resulting in higher mortality \citep{Chave-2009, Wright-2010}.
-Under this scenario, decreasing $\rho$ will always provide a growth advantage, because the costs of low $\rho$ are not realised within the terms of eqs.
-\ref{eq:dhdt} and \ref{eq:dast} (Table \ref{tab:trade-offs}).
-In addition, lower $\rho$ may increase turnover of sapwood.
-Mirroring our assumptions for leaves, we let sapwood turnover decrease with $\rho$, i.e.$k_{\rm s}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.
+The results reported above and shown in Fig.
+\ref{fig:growth_light_height} focus on height growth rate (eqn \ref{eq:dhdt}).
+Corresponding results for growth rates in stem-diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
+\ref{fig:growth_light_dia}--Fig.
+\ref{fig:growth_light_mass} of the Supplementary Materials.
+In each figure, plants were grown to a suitable diameter, area, or mass.
+As such, changes in relative growth rates with traits show a similar patterns as absolute growth rates.
 
-Under this scenario, decreasing $\rho$ provides an advantage in height growth only when the benefits of low $\rho$ outweigh the costs.
-This effect is strongest at intermediate sizes.
-The decline in height growth at low $\rho$ with size is more marked than for either stem basal-area and diameter growth, because high sapwood turnover also contributes to this growth measure.
-Thus diameter growth continues, even as net production approaches zero.
+We find that for seed mass, leaf nitrogen, leaf mass per area and height at maturation, the patterns of growth rate in stem-diameter, stem-area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}).
+The only trait where a slightly different response was observed was for wood density.
+Whereas the effect of wood density on height growth tended to diminish slightly at larger sizes (Fig.
+\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem-diameter, stem-area or above-ground mass.
+Recall that sapwood lost via turnover is turns into heartwood.
+Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.
 
@@
 
-Two other important predictions arise directly from linking traits to growth rate.
-First, traits moderate the responsiveness of growth to changes in light environment.
-This response arises simply because species with higher potential growth rate have greater potential plasticity.
-Recent analyses have shown that species with low $\rho$ increase growth more substantially with increases in light \citep{Ruger-2012}.
-This is exactly the behaviour exhibited in our model (Fig.
-\ref{fig:growth_light_dia}), at least for small to intermediate sizes.
-At larger sizes (D=0.2m) however, the advantages of low $\phi$ are lost (Fig.
-\ref{fig:growth_light_dia}).
-This results also agrees with empirical results showing that growth for species with low $\rho$ actually decreases as size increases \citep{Ruger-2012}.
-In contrast, growth rate tends to increase in size for the species with highest $\rho$ \citep{Ruger-2012}.
-
-Variation in $\phi$ also moderates the response of growth to changes in light, with species having the lowest $\rho$ being most responsive.
+The predictions in Figs.
+\ref{fig:growth_light_height} and \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} illustrate how traits impact on growth rate under different light environments and at different sizes.
+An additional outcome that arises directly from these analyses is that traits moderate the responsiveness of growth to changes in light environment.
+This response arises because individuals with higher potential growth rate naturally have greater potential plasticity in growth.
+Our results therefore support findings that species with low $\rho$ increase growth more substantially with increases in light (Table \ref{tab:phenomena}).
+Variation in $\phi$ also moderates the response of growth to changes in light, with species having the lowest $\phi$ being most responsive.
 However, unlike for $\rho$, the effect appears only for the smallest size classes.
-Accordingly, empirical analyses have detected such a link in small plants but not for plants \textgreater{} 0.01m dbh \citep{Ruger-2012}.
 
 \subsection{Shade tolerance}
-
-Second, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
+Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
+\begin{equation}\label{eq:wplcp_2}
+\bar{p}(E^*) =
+      \phi \left(\frac{k_{\rm l}}{y} + r_{\rm l}\right) +
+      \left(\theta \, \rho \, \eta_c \, H\right)
+        \left(b\left(\frac{k_{\rm b}}{y} + r_{\rm b}\right)
+            + \left(\frac{k_{\rm s}}{y} + r_{\rm s}\right)\right) +
+      \alpha_{\rm r1} \left(\frac{k_{\rm r}}{y} + r_{\rm r}\right).
+\end{equation}
+Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with traits $\nu$, $\phi$, and $\rho$.
+With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$.
+Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$.
+The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm kl2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
+Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
 \ref{fig:wplcp}).
 At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset these costs.
-As previously suggested \citep{Givnish-1988}, shade-tolerance also decreases with height because as size increase, the total amount of energy needed to offset respiratory and turnover costs in the stem also increases.
-These finding matches well known empirical patterns, where both $\phi$ and $\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,Osunkoya-1996}.
-
-Results from \citet{Poorter-2006}: survival strongest for leaf lifespan.
-
-Also \citep{Baltzer-2007}.
-
-\subsection{Changes in trait values through ontogeny}
+{\wplcp} also decreases with height because as size increase, the total amount of energy needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
+In addition, {\wplcp} varies with $\nu$.
+At small sizes, {\wplcp} increases with $\nu$ across the band of values typically observed in real plants, i.e.
+high leaf nitrogen makes seedlings shade intolerant.
+At larger sizes, as net production declines to zero, {\wplcp} begins to increase again for very low values of $\nu$.
+All of these patterns match empirically observed patterns (Table \ref{tab:phenomena}).
 
@@
 
-Model is realisation of several ideas that have been around for some
-time.Employs the same framework for partitioning growth used by \citet{Gibert-2016} in their meta-analysis, which builds on and extends several related approaches \citep{Givnish-1988, Yokozawa-1995, Makela-1997, Moorcroft-2001, Falster-2011, King-2011} in which metrics such as height, diameter and mass growth are calculated as an emergent outcomes of several, independent, physiological processes.
-This approach is quite different to models derived from metabolic scaling theory \citep{Enquist-1999, Enquist-2007}, which attempt to organise everything around a single master ``scaling'' equation for mass growth; and also different from statistically fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it posits rather than tests for traits based effects.
-Our goal is to help
+Using a model relating plant physiological function and energy allocation to five prominent traits, we have shown how traits impact on plant growth across the life cycle.
+This approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait leaf mass per unit leaf area \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
+During the last two decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}.
+However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}.
+Matching a growing amount of empirical evidence (Table \ref{tab:phenomena}), this study outlines when and why the direction or strength of correlation between traits and growth rate shifts with plant size.
+Moreover, we show that different traits and trade-offs generate different types of response.
+Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life (from slow to fast growth) \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
+Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by light environment.
+The {\plant} model, used here, builds on and extends several related approaches \citet[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits.
+This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which attempts to derive everything from a single master ``scaling'' equation for mass growth.
+Thus far, the \textsc{mst} has as been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}.
+Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects.
+In this sense, our model is designed to both explained observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
 
-Key patterns explained
+\subsection{Generalising to other traits and trade-offs}
+
+The list of functional traits that are known to differ among plant species is already long and ever-increasing \citep{Perez-2013}.
+While we have focussed on understanding the effects of five specific traits on growth rate, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
+The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass or carbon, and can therefore be embedded within the eqns in Fig.
+\ref{fig:conceptual}b.
+A well-developed trait-based -- or some might say evolutionary -- trade-off must include two opposing forces, that operate at some point in the organisms life cycle.
+It is not necessary that benefit and cost both enter into Fig.
+\ref{fig:conceptual}; for example the trait wood density is sometimes thought of a trade-off between the costs of tissue construction and the rate of stem mortality.
+In that case, lower wood density would always increase growth rate, while higher wood density would always increase survival.
+However, while the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
+
+Our framework also highlights what is needed for traits to impact on growth rates and shade-tolerance.
+While traits can influence many aspects of plant function, these  influences must operate via the pathways outlined in Fig.
+\ref{fig:conceptual} if the trait is going to impact on growth.
+For example, many studies have focused on traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
+These traits will inevitably influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}) by altering conductivity of water to the leaf.
+The potential costs of larger vessel size might be higher rates of stem turnover, which would appear in the term $k_{\rm s}$ in eqn \ref{eq:dbdt}.
+The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in our framework via the parameter $\theta$ (Table \ref{tab:allometry}).
+The effect of both these traits on growth rate should be expected to vary with plant size.
+
+% The only circumstances whereby traits impact on growth but not via the pathways in Fig. \ref{fig:conceptual}, would be where growth rate was entirely disconnected from the plant's carbon budget.
+
+\subsection{Relationship to a classic model for seedling growth}
+
+The model presented here extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
+Importantly, the seedling model can be derived as a special case of our more-extended approach.
+For seedlings, which are young and mostly leaf, we can ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
+Net production (from eqn \ref{eq:dbdt}) then becomes a linear function of leaf area, making relative growth rate in mass a linear function of $\phi$:
+\begin{equation}\label{eq:RGR}
+\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
+where $A_\textrm{net} = y(\bar{p}(E) - r_{\rm l})$.
+Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees: namely plant height ($H$), stem-basal area ($A_{\rm st}$) or stem diameter $D$.
+The refined model makes these links clear.
+
+Unlike the seeding model, the new model also predicts a relationship between $\phi$ and growth rate that changes with size.
+In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig.
+\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem-diameter, stem-area or or total mass.
+This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size.
+As seedlings, leaves comprise a large part of the plant (Fig.
+\ref{fig:conceptual}a).
+Decreasing $\phi$ has an overwhelmingly positive effect on height and diameter growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
+As plants increase in size, however, the amount of supporting tissue increases, decreasing the benefit of cheaper leaf construction in eqn \ref{eq:daldmt}.
+Consequently, the effect of $\phi$ on leaf turnover comes to dominate at larger sizes, and as such, the effect of $\phi$ on height, diameter and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
+
+\subsection{Implications for trait-based approaches}
+
+There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.
+
+First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
+Much of current ecosystem research focusses on  photosynthesis, respiration, and resultant fluxes of carbon, with little attention devoted to allocation and turnover \citep[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}.
+Yet, for four of the five traits considered here, trait values do not influence photosynthesis, respiration, or net primary production (=photosynthesis - respiration).
+In fact, the analysis with $\phi$ shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
+Low $\phi$ results in high leaf turnover, such that individuals with a $\phi$ have lower mass production.
+It is this property that makes them shade intolerant.
+And yet they can still achieve a growth advantage (when small), because the benefits of cheap leaf construction outweigh the costs of high leaf turnover.
+
+% - \citet{King-2005} shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
+
+Second, our results demand a shift in the way plant species are sometimes described as, being of fast or slow growth.
+To the extent the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum from slow to fast growth.
+Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{CSR} triangle, the r-K spectrum, and coexistence models base on the Lotka- Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.
+Researchers using functional traits have also tended to describe species as fast or slow growing \citep[e.g.][]{Adler-2014,Diaz-2016}.
+Our results suggest a more-nuanced approach.
+Plants that are fast growing as seedlings may not be fast growing as saplings or adults, or under low light.
+Plants that are fast growing as adults may not be fast growing as seedlings.
+This more-nuanced perspective tends to mirror observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
+% So we should instead be thinking of a species traits as defining a trajectory of potential growth for an individual, and one that is moderated by plant size and light environment.
+
+Third, our results suggest that even if traits define a potential growth trajectory, researchers seeking to link traits to growth rate must probe deeper into the data than simply looking for a linear relationship.
+None of the predicted relationships between traits and growth is linear across the range of sizes and light environments sampled.
+As such,  we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits.
+Controlling for size, site and light environment will therefore be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}.
+
+\subsection{Comparison with other frameworks}
+
+It is also worth comparing our approach to other frameworks.
+As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citet{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}, and in particular those by \citet{Makela-1997} and \citet{Moorcroft-2001}.
+These models have several properties in common, including that they all have growth being energetically driven by the gross amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
+Many models also make functional-balance assumptions, for example linking the cross-section of sapwood to leaf area \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-2005, Moorcroft-2001}.
+We note that an assumption of exact functional balance is non-critical for our results, what matters is that the amount of live biomass (i.e.
+excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
+\ref{fig:assumptions}d).
+Some models also differ from in that they directly link a plant's stem diameter to it's height, e.g.
+\citet{Yokozawa-1995, King-1999, Moorcroft-2001, Li-2014}, whereas we let this scaling arise as an emergent outcome of growth and sapwood turnover.
+Again, we suspect this difference will not effect the core of results presented.
+
+A feature distinguishing model from most of the mentioned above is the explicit linking to trait-based trade-offs.
+While such a linkage was also made by \citet{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have mainly focussed on ecosystem-level outcomes rather than the growth of individual plants.
+Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
+\citet{King-1999} also connected his model of growth to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
+\ref{fig:growth_light_height}).
+This flattening out arose for a similar reason: the value of cheap leaf construction decreased with size, as more of the plant became stem.
+In our study however, there was additional cost of increased leaf turnover, that further penalised low $\phi$ strategies when large.
+
+Perhaps the two most controversial elements of our approach concerns our assumptions about tissue replacement and reproductive allocation.
+Many vegetation models determine allocation based on net primary production(photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing surplus energy.
+This is because  we assume tissues lost via turnover are replaced before energy is allocated to either new growth (i.e.
+growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
+This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant (??refs XX??).
+The second assumption we make is that when mature, plant's allocate a substantial fraction of their surplus energy to reproduction.
+While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
+Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}.
+Reproductive allocation is rarely given much attention in models of ecosystem flux.
+For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
+In the case of growth, differences in reproductive allocation offer a clear pathway for explaining patterns linking a plants maximum size to the growth rate of large individuals \citep[e.g.][]{Wright-2010}.
+
+Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}.
+In this framework, mass-based growth is expressed as a 3/4 scaling power of total plant mass, and diameter growth is expressed as a 1/3 power of stem diameter.
+An extension of the model, connecting it to eqn \ref{eq:RGR} for relative growth rate, makes the connection to traits more explicit \citep{Enquist-2007}.
+Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
+First, the \textsc{mst}-derived model suggests diameter growth continues to increase as plants grow, whereas empirical data suggests growth rate declines for larger plants \citep{Herault-2011}.
+Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size.
+Predicted effects are for a linear increase in growth with decreases in either $\phi$ and lower $\rho$, that apply irrespective of size.
+However, at least for $\phi$ such effects in large trees have not been observed.
+Finally, the \textsc{mst} models are difficult to connect with growth rates in other terms such as height.
+Other predictions from the two frameworks  agree, but for fundamentally different reasons.
+Specifically, both frameworks predict that the growth rate in total plant mass will continue to increase with size, which matches empirical evidence \citep{Stephenson-2014}.
+In the \textsc{mst} framework, this occurs because all mass (including heartwood) contributes to mass production.
+In our framework, this occurs because the turnover of sapwood also contributes substantially to whole-plant mass growth.
+This latter pattern matches data for large trees, where it was found that while height growth declined, mass growth continued to increase, mostly because of increased mass production \citep{Sillett-2010}.
+
+\subsection{A prediction and closing remarks}
+While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
+So let us conclude with a new, testable prediction arising from our results: which is that the trait $\phi$ should increase through ontogeny.
+This prediction arises from the fact that as plants grow larger, the value of cheap leaf construction diminishes, compared to the costs of high leaf turnover.
+It therefore pays for all plants, but especially those with low $\phi$, to increase their $\phi$ as they grow larger.
+While trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
+This hypothesis attempts to give meaning to some of that variation.
+
+So we should instead be thinking of a species traits as defining a trajectory of potential growth for an individual, and one that is moderated by plant size and light environment.
 
@@
 
-- mechanistic model of height and dbh growth which accounts for effects
-  of size, light environment and traits
+- mechanistic model of height and dbh growth which accounts for effects of size, light environment and traits
 - need to account for size long recognized, but not straightforward
@@
 
-Plasticity in trait values predicted from model
-
-\subsection{Comparison with existing frameworks}
-
-Growth is more than just photosynthesis:
-
-Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000, Enquist-2007}, ${\rm d}P / {\rm d}t / M_{\rm a}$.
-This relationship can be derived from eqn \ref{eq:dbdt}  by ...
-(see Appendix \ref{app:traits-RGR}).
-To make sense of empirical patterns, we extend a widely-used theoretical model that links growth rate in seedlings to leaf mass per leaf area \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
-This model.....Builds on 2011 paper - highlight what's new from there.
-
-Recently realisation that allocation and turnover major areas of uncertainty
-
-Model reconciles idea that traits can have negative impact on carbon budget, but still have positive influence on growth rates
-
-Assumes low lma and wd used to reduce costs of building leaf and stem, ie.
-total leaf area and stem cross section remains same.
-In this way differs to assumptions of recent work by \citep{Anten-2010}, \citep{Larjavaara-2010}.
-
-\begin{itemize}
-  \item \citet{Ishihara-2016}
-  \item  \citet{Enquist-2007}: sla and wd increase mass-production, where as in our theory effects on mass production are negative.
-(also contradicts earlier model from \citet{Enquist-1999}, where argued wood density had no effect on mass production)
-  \item \citet{Ryan-2006}: reviewed evidence for hydraulic limitation of tree height.
-Designed to explain decrease in height growth rate for trees.
-Found mixed support: some species experience decreases in physiological function with height, especially very large species \citep{Koch-2004}, but certainly not all species \citep{Ryan-2006}.
-Important to emphasise that demographic patterns observed by \citep{Ryan-2006} (decreases in height growth rate, no decrease in biomass growth) predicted from theory here, without invoking hydraulic limitation.
-So hydraulic limitation may amplify size related effects but not root cause.
-  \item \citet{King-2005} Section 3.3 of King  shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
-\citet{King-1999} also predicts a change in the SLA optimising growth rate, via a cross correlation with self shading.
-Basically, the much greater path length from roots to leaves in trees vs.
-seedlings results in a much greater increment in support cost per increment in crown area for trees.
-This explanation is treated by the growth model of \citet{King-1999}, which predicts that the SLA maximizing growth rate declines as seedlings grow larger, even before the onset of leaf turnover.
-  \item \citet{King-1999}:
-  \item \citet{Li-2014}:
-  \item \citep{Givnish-1988,Yokozawa-1995,Makela-1997,King-2011,Moorcroft-2001}
-  \item \citep{Coomes-2011}:
-\end{itemize}
-
-The only potentially controversial element of in eqn \ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocated to either new growth (i.e.
-growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$ or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction often entails run-down in the vegetative part of the plant (refs XX).
-
-In what follows, the assumption of exact isometry is non-critical, what matters is that the amount of live biomass (i.e.
-excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
-\ref{fig:assumptions}d).
-Note that similar increases in construction and maintenance costs with plant size are embedded within various models of plant function \citep[e.g.][]{Givnish-1988, Makela-1997}.
-
-The list of functional traits that are known to differ among plant species is already long and ever-increasing. However, far fewer have well established tradeoffs.
-
-\subsection{Plasticity}
-
-other plasticity \citep{Thomas-2010}
-
-\subsection{Implications and challenges for trait-based approaches}
-
-How do comparative people interpret $\phi$ now - beyond fast slow.
-
-Think about traits as defining potential trajectory, rather than making
-species fast or slow growing per se.
-
-Selection on traits in different parts of life cycle
-
-Trait plasticity - can explain, no clear advantage of low lma at large
-sizes
-
-Framework for generalising about effect of traits. New traits must have effects via pathways outlined in Fig. 1. Assuming plant works in way described by Fig 1.
-
-\begin{itemize}
-\item Reproductive allocation
-\item Detecting trait signals in data.
-\item Need more than simply look for correlation with mean growth rate
-\item Understand influences of size on trait of interest
-\item Traits define potential growth rate --\textgreater{} how to extract
-  this
-\item Ideally start to use mechanistic models
-\end{itemize}
-
 \section{Acknowledgements}
 
-Fig. 2 uses image by Tracey Saxby, IAN Image Library \smurl{http://ian.umces.edu/imagelibrary}.
+We thank J Camac, R Duursma, A Gibert, G Kunstler, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
+Falster was supported by an Australian Research Council discovery grant (DP110102086).
+FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174).
+The authors have no conflicts of interest to declare.
+
+\section{Supplementary Materials}
+
+\subsection{A -- Full description of the \texttt{FF16} growth model}\label{sec:ff16}
+
+\subsection{B -- Supplementary Figures}\label{sec:SM_figs}
+
+\subsection{C -- Supplementary Tables}\label{sec:SM_tabs}
 
@@
 
-\section{Tables}
+\section{Tables and Figures}
 
@@
 
-  \\\multicolumn{2}{l}{\bf{Patterns of growth with increasing size}}\\
+  \\ \multicolumn{2}{l}{\bf{Patterns of growth with increasing size}}\\
   Hump-shaped pattern of net biomass production & \citet{Givnish-1988, Koch-2004} \\
-  Increasing rate of plant mass and stem area growth & \citet{Stephenson-2014} \\
+  Increasing rate of plant mass and stem area growth & \citet{Sillett-2010, Stephenson-2014} \\
   Hump-shaped pattern of $H$, $D$, $A_{\rm l}$ \& $A_{\rm st}$ growth& \citet{King-2005, Ryan-2006, Sillett-2010, Herault-2011, King-2011} \\
@@
   \\ \multicolumn{2}{l}{\bf{Effect of traits on growth rate}}\\
+  Low $\omega$ species produce smaller seedlings, resulting in lower absolute growth rates but higher relative growth rates. & \citet{Turnbull-2012};  Meta-analysis by \citet{Gibert-2016};\\
   Low $\phi$ species$^1$ have a growth advantage when small, that dissipates with increasing size & Meta-analysis by \citet{Gibert-2016}\\
   Low $\rho$ species have a growth advantage at all sizes & Meta-analysis by \citet{Gibert-2016}\\
-  High $H_{\rm m}$ species have a growth advantage when small & Meta-analysis by \citet{Gibert-2016}\\
-  High $H_{\rm t,0}$ species have an advantage in absolute growth rate, while low $H_{\rm t,0}$ species have faster RGR & Meta-analysis by \citet{Gibert-2016}\\
+  High $H_{\rm m}$ species have a growth advantage when large & Meta-analysis by \citet{Gibert-2016}\\
+
+  \\ \multicolumn{2}{l}{\bf{Responsiveness of growth rate to light}}\\
+  Low $\rho$ species respond more to changes in light & \citep{Ruger-2012}\\
 
@@
   \textsc{wplcp} decreases with height & \citet{Givnish-1988}\\
-  Low $\phi$ species have higher \textsc{wplcp} & \citet{Poorter-2006, Lusk-2008}\\
+  Seedlings of higher $\nu$ species have higher \textsc{wplcp} &  \citet{Baltzer-2007}\\
+  Low $\phi$ species have higher \textsc{wplcp} & \citet{Poorter-2006, Baltzer-2007, Lusk-2008}\\
   \quad $\rightarrow$ Stands of low $\phi$ species$^1$  have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
   Low $\rho$ species have higher \textsc{wplcp} &  \citet{Osunkoya-1996}\\
-
-  \\ \multicolumn{2}{l}{\bf{Changes in traits through ontogeny}}\\
-  $\phi$ $^1$ tends to increase through ontogeny & \\
-  $\rho$ tends to increase through ontogeny & \citet{Osazuwa-2014}\\
   \\ \hline
@@
 \raggedright
-  \begin{tabular}{p{3cm}p{1cm}p{1.5cm}p{5cm}p{3cm}}
+  \begin{tabular}{p{3cm}p{1.5cm}p{6cm}p{3cm}}
   \\
   \hline
-  Name & Symbol & Unit & Trade-off (benefit {\bf vs.} cost) & References \\
+  Name, Symbol & Unit & Trade-off & References \\
   \hline
   \\
-  Seed mass & $M_{\rm t, 0}$ & kg & $\uparrow$ initial size seedlings {\bf vs.} $\downarrow$  maternal fecundity & \citet{Moles-2006}\\
+  Seed mass, $\omega$ & kg & $\uparrow$ initial size seedlings {\bf vs.} $\downarrow$  maternal fecundity & \citet{Moles-2006}\\
   \\
-  Height at maturation & $H_m$ & m  & $\uparrow$ growth {\bf vs.} $\downarrow$ fecundity & \citet{Thomas-2011, Wenk-2015}\\
+  Height at maturation, $H_{\rm m}$ & m  & $\uparrow$ growth {\bf vs.} $\downarrow$ fecundity & \citet{Thomas-2011, Wenk-2015}\\
 \\
-  Leaf nitrogen per unit leaf area & $\nu$ & kg m$^{-2}$ & $\uparrow$ photosynthesis in high light {\bf vs.} $\uparrow$ respiration rate & \citet{Wright-2004}\\
+  Leaf nitrogen per unit leaf area, $\nu$ & kg m$^{-2}$ & $\uparrow$ photosynthesis in high light {\bf vs.} $\uparrow$ respiration rate & \citet{Wright-2004}\\
   \\
-  Leaf mass per unit leaf area & $\phi$ & kg m$^{-2}$ & $\downarrow$ cost of building leaf {\bf vs.} $\uparrow$ leaf turnover & \citet{Wright-2004}\\
+  Leaf mass per unit leaf area, $\phi$ & kg m$^{-2}$ & $\downarrow$ cost of building leaf {\bf vs.} $\uparrow$ leaf turnover & \citet{Wright-2004}\\
   \\
-  Wood density & $\rho$ &  kg m$^{-3}$ & $\downarrow$ cost of building stem {\bf vs.} $\uparrow$ mortality rate & \citet{Chave-2009}\\
+  Wood density, $\rho$ &  kg m$^{-3}$ & $\downarrow$ cost of building stem {\bf vs.} $\uparrow$ mortality rate & \citet{Chave-2009}\\
   \\
@@
      &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:ha} \\
+    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
   Heartwood mass &
@@
      &
-    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:hm} \\
+    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:htm} \\
     \\
@@
 Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
-Traits are: Seed mass ($M_{\rm t,0}$), Height at maturation ($H_{\rm m}$), Nitrogen content per leaf area ($\nu$), Leaf mass per unit leaf area ($\phi$),  Wood density ($\rho$), Sapwood area per leaf area ($\theta$), Vessel size ($\kappa$), and root mass per leaf area ($\alpha_{\rm r1}$).}
+Traits are: seed mass ($\omega$), height at maturation ($H_{\rm m}$), nitrogen content per leaf area ($\nu$), leaf mass per unit leaf area ($\phi$), and wood density ($\rho$) (see Table \ref{tab:traits} for more details).}
 {\footnotesize
@@
   \hline
-  & & &  &  \multicolumn{2}{c}{\bf Ontogeny} & \multicolumn{2}{c}{\bf Leaf} & \multicolumn{1}{c}{\bf Stem}& \\
-  & & &  {\bf Symbol} & \boldmath$M_{\rm t,0}$ & \boldmath$H_{\rm m}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\ \hline
+  & & &  &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
+  & & &  {\bf Symbol} & \boldmath$\omega$ & \boldmath$H_{\rm m}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\ \hline
   \\
-  \multicolumn{10}{l} {\textbf{a) Effect on elements of equations \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
+  \multicolumn{10}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
   & \multicolumn{2}{l} {Net biomass production} & $\ud B / \ud t$ & & & & & & \\
@@
   & & \tabitem Respiration  & & - & - & $\uparrow$  & - & - & \\
-  & & \tabitem Turnover & & - & - & - & $\uparrow$ & - & \\
-  \\
+  & & \tabitem Turnover & & - & - & - & $\uparrow$ & $\uparrow$ & \\
   & \multicolumn{2}{l} {Allocation to growth} & $\ud M_{\rm a} / \ud B$ & - & $\uparrow$ & - & - & - & \\
-  \\
   & \multicolumn{2}{l} {Leaf deployment per mass}  & $\ud A_{\rm l} / \ud M_{\rm a} $ & & \\
@@
     & & \tabitem Root & & - & - & - & - & - & \\
-  \\
   & \multicolumn{2}{l} {Architecture} & $\ud H / \ud A_{\rm l}$ & - & - & - & - & - & \\
   \\
-  \multicolumn{10}{l} {\textbf{b) Predicted effect on plant growth rate for a small and large plant}} \\
-  & \multicolumn{9}{l} {in mass {\color{red} (check, live or total?)}} \\
+  \multicolumn{10}{l} {\textbf{b) Predicted effect of trait on growths rate for a small and large plant}} \\
+  & \multicolumn{9}{l} {Height} \\
+  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \doup & \dodo & \\
+  & \multicolumn{9}{l} {Stem area {\color{red} (check)}} \\
+  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \doup & \dodo & \\
+  & \multicolumn{9}{l} {Stem diameter} \\
+  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \doup & \dodo & \\
+  & \multicolumn{9}{l} {Mass {\color{red} (check, live or total?)}} \\
   & & \tabitem{Absolute} & $\ud M_{\rm a} / \ud t$ & \upfl &  \flup & \upup & \doup & \dodo & \\
   & & \tabitem{Relative} & $\ud M_{\rm a} / (\ud t . M_{\rm a} )$ & \dofl &  \flup & \upup & \doup & \dodo & \\
-  \\
-  & \multicolumn{9}{l} {in height} \\
-  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
-  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \doup & \dodo & \\
-  \\& \multicolumn{9}{l} {in stem area {\color{red} (check)}} \\
-  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
-  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \doup & \dodo & \\
-  \\& \multicolumn{9}{l} {in stem diameter} \\
-  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
-  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \doup & \dodo & \\
   \\
@@
   }
-\label{tab:trade-offs}
+\label{tab:responses}
 \end{table}
 
-\newpage
-\section{Figures}\label{figures}
+\clearpage
 
@@
 \includegraphics{main/figure_conceptual.pdf}
-\caption{\textbf{Conceptual framework for understanding size-dependent
+\caption{\textbf{Conceptual framework for understanding size- and trait-dependent
 changes in growth rate.} Hump-shaped relationship between height growth rate and
 size. \label{fig:conceptual}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+\centering
+\includegraphics{main/lcp_schematic.pdf}
+\caption{\textbf{Conceptual framework for understanding size- and trait-dependent
+changes in shade tolerance.} Do we need a figure? \label{fig:wplcp_idea}}
 \end{figure}
@@
 
-% \newpage
-
-% \begin{figure}[ht]
-% \centering
-% \includegraphics{main/max_leaf_above.pdf}
-% \caption{\textbf{The effect of traits on growth changes with size and
-% light environment.} Traits moderate the responsiveness of growth to changes in light
-% environmentThis figure needs to show more of key result with respect to size.
-% Possibly separate figure. \label{fig:shifts}}
-% \end{figure}
-
 \newpage
@@
 \centering
-\includegraphics{main/max_leaf_above.pdf}
-\caption{\textbf{Effect of traits on shade tolerance.} Panels show effect of traits on maximum amount of shading that can be endured before net production (eqn \ref{eq:dbdt}) reaches zero.
-Lines indicate relationship for plants of a given height.
+\includegraphics{main/lcp_trait.pdf}
+\caption{\textbf{Effect of traits on shade tolerance.} Panels show effect of traits on level of canopy openness that causes net production (from eqn \ref{eq:dbdt}, Fig.
+\ref{fig:conceptual}) to be zero.
+Different lines indicate relationship for plants with specified height.
 \label{fig:wplcp}}
@@
 \clearpage
-
-\section{Appendices}
-\subsection{A - Relationship between traits and relative growth rate} \label{app:traits-RGR}
-
-We start with eqn \ref{eq:dbdt} for mass-based growth. For seedlings, which are
-young and mostly leaf, it is reasonable to ignore all turnover terms as
-well as the respiration terms for non-leaf tissues. Net production then
-becomes a linear function of leaf area and net photosynthesis per leaf
-area ($A_\textrm{net} = y(A - r_{\rm l})$), making relative growth
-rate a linear function of $\phi$:
-\begin{equation}\label{eq:RGR}
-\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}. \end{equation}
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in
-relation to $\phi$ \citep{Wright-2000}, this
-approximation does not directly link to other traits, or to the
-variables that are routinely collected for large trees: namely plant
-height ($H$) and stem cross-sectional area ($A_{\rm st}$) or
-diameter $D$.
-
-\subsection{B - Effect of traits on height growth} \label{app:traits-RGR}
-
-We want to show how two traits $\phi$ and $\rho$ influence growth rate,
-by looking at the effect of
-traits on the components of growth rate (eqn \ref{eq:dhdt}). To make
-the analysis more tractable, we will assume we are dealing with a plant of
-given height where 100\% of available energy is allocated to growth,
-i.e. $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=1$. Note also
-that by the assumptions of our model and above, traits do not influence
-$\frac{{\rm d}H} {{\rm d}A_{\rm l}}$. Thus we have
-\begin{equation} \label{eq:G2}
-G = \frac{{\rm d}H}{{\rm d}t} = c_1   \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}  \frac{ {\rm d}B} {{\rm d}t} \right),
-\end{equation}
-where $c_1 = \frac{{\rm d}H}{{\rm d}A_{\rm l}}$. Eqn
-\ref{eq:G2} is a product of the form $Y(x) = W(x) \, Z(x)$. For,
-equations of this type, the derivative $\partial{Y}/\partial{x}$ is
-given by $\partial{W}/\partial{x} \, Z + W \, \partial{Z}/\partial{x}$. Thus
-the derivative of $G$ with respect to some trait $x$  is given by
-\begin{equation} \label{eq:dG}
-\frac{\partial G} {\partial x} =
-\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}{\partial x}
- \, \frac{ {\rm d}B} {{\rm d}t}
-+ \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\, \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}.
-\end{equation}
-If this derivative is positive (negative), there will be a positive (negative) relationship
-between traits and growth at the height where it is evaluated.
-
-We can further simplify by expanding the derivative of eqn \ref{eq:daldmt}:
-\begin{equation} \label{eq:dA_dx}
-\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}
-{\partial x} = -\left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\ \right)^2
-\, \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}.
-\end{equation}
-Combing eqs. \ref{eq:dG}-\ref{eq:dA_dx}, we have
-\begin{equation} \label{eq:dG2}
-\frac{\partial G} {\partial x} =
-\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\left(
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
-- \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}
- \, \frac{ {\rm d}B} {{\rm d}t}
-\right).
-\end{equation}
-
-Since is $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}>0$ always, the sign
-of $\frac{\partial G} {\partial x}$ depends on the terms inside the brackets on the
-RHS. Rearranging we find that $\frac
-{\partial G} {\partial x} < 0$ if
-\begin{equation}\label{eq:dg3}
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}
-< \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-\,  \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x} \, \frac{ {\rm d}B} {{\rm d}t},
-\end{equation}
-or equivalently
-\begin{equation}\label{eq:dg4_orig}
-\frac{
-\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-{\frac{ {\rm d}B} {{\rm d}t}}
-<
-\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-\end{equation}
-
-\subsection{C - Linking traits to whole-plant light compensation point} \label{app:traits-RGR}
-
-\section{Supplementary materials}
-
-\subsection{A -- Full description of the \texttt{FF16} growth model}\label{sec:ff16}
-
-\subsection{B -- Supplementary Figures}\label{sec:SM_figs}
-
-\subsection{C -- Supplementary Tables}\label{sec:SM_tabs}
 

### 2016-09-16 R/axis.R

@@ -1,14 +1,14 @@
 name_pretty <- function(trait) {
   switch(trait,
-         lma=expression(paste("Leaf-mass per unit leaf area (kg ", m^-2, " )")),
+         lma=expression(paste("Leaf-mass per area (kg ", m^-2, " )")),
          rho=expression(paste("Wood density (kg ", m^-3, " )")),
-         narea=expression(paste("Nitrogen per leaf area (kg ", m^-2, " )")),
-         hmat="Height at maturation (m)",
+         narea=expression(paste("Nitrogen per area (kg ", m^-2, " )")),
+         hmat=expression(paste("Height at maturation (m)")),
          diameter_stem_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
-         shading="WPLCP (0-1)",
+         shading=expression(paste("WPLCP, ",E^"*"," (0-1)")),
          leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, " )")),
 
-         height_dt=expression(paste("Height growth (m ", yr^-1, " )")),
+         height_dt=expression(paste("Height growth rate (m ", yr^-1, " )")),
          diameter_stem_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
          area_stem_dt=expression(paste("Basal area growth (", m^2, " ", yr^-1, ")")),
          mass_above_ground_dt=expression(paste("Mass growth (kg ", yr^-1, ")")),
@@ -23,9 +23,33 @@ name_pretty <- function(trait) {
 
 trait_range <- function(trait) {
   switch(trait,
-         lma=c(0.02, 1.0),
-         rho=c(200,  1200),
-         narea =c(5E-4, 2E-2),
-         hmat=c(3, 60),
+         lma=c(5E-3, 3),
+         rho=c(90, 3000),
+         narea =c(1E-4, 5E-2),
+         hmat=c(1, 100),
          stop("No standard range for trait ", trait))
+}
+
+trait_range_obs <- function(trait) {
+  switch(trait,
+         lma=c(0.05, 0.3),
+         rho=c(300,  970),
+         narea =c(5E-4, 1E-2),
+         hmat=c(4, 30),
+         stop("No standard range for trait ", trait))
+}
+
+get_by <- function(ymax) {
+    by <- diff(pretty(c(0, ymax)))[1]
+    if(ymax > 0.6)
+        by <- 0.2
+    if(ymax > 1.25)
[…]

### 2016-09-16 R/figures.R

@@ -38,7 +38,7 @@ make_hyperpar2 <- function(
                                 narea=1.87e-3,
                                 narea_0=1.87e-3,
                                 B_lf1=5120.738 * 1.87e-3 * 24 * 3600 / 1e+06,
-                                B_lf2=0.5,
+                                B_lf2=0.75,
                                 B_lf3=0.04,
                                 B_lf4=21000,
                                 B_lf5=1,
@@ -198,15 +198,15 @@ make_hyperpar2 <- function(
 default_parameters <- function() {
 
   s <- FF16_Strategy(
-    hmat = 30.0,
+    hmat = 15.0,
     a_f1 = 0.8,
-    a_f2 = 20,
+    a_f2 = 10,
     a_l1 = 2.17,
     a_l2 = 0.5,
     k_l  = 0.4565855 / 3)
 
   FF16_Parameters(
-    hyperpar=make_hyperpar2(B_ks2=1.25),
+    hyperpar=make_hyperpar2(B_ks2=1.25, B_lf5=0.5, B_lf4=21000*0.75),
     strategy_default = s
     )
 }
@@ -217,8 +217,19 @@ default_strategy <- function() {
   p$strategy_default
 }
 
-color_pallete <- function(n=9) {
-  RColorBrewer::brewer.pal(n, "Blues")
+color_pallete1 <- function(n) {
+  RColorBrewer::brewer.pal(n+3, "Blues")[-(1:3)]
+}
+
+color_pallete2 <- function() {
+  RColorBrewer::brewer.pal(n=9, "Blues")
+}
+
+color_pallete3 <- function() {
+
+  cols <- RColorBrewer::brewer.pal(9, "YlOrBr")
+  c(mass_leaf="forestgreen", mass_root=cols[4], mass_bark=cols[6],
+            mass_sapwood=cols[7], mass_heartwood=cols[9])
 }
 
 run_plant_to_sizes <- function(sizes, size_variable, strategy, env,
@@ -243,36 +254,6 @@ extract_plant_info <- function(plant, env) {
   x[sprintf("%s_dt_relative", v)] <- x[sprintf("%s_dt", v)] / x[v]
 
   data.frame(t(x))
-}
-
-## Code for trait derivative figure
-## TODO: Can grader help here?
-figure_trait_derivative <- function(type, trait_name="lma", canopy_openness=1) {
-  strategy <- default_strategy()
-
[…]

### 2016-09-16 R/figures_data.R

@@ -88,7 +88,7 @@ figure_assumptions <- function(fit_a.lf_h.t,  fit_a.lf_sap, fit_a.lf_m.rf, fit_h
   myaxislabel <- function(txt, side) mtext(txt, side, line =3.5)
   mylegend <- function(txt) legend("bottomright", legend = txt, bty = "n", cex=1)
 
-  cols <- color_pallete()
+  cols <- color_pallete2()
   s <- default_strategy()
   x <- seq_log_range(c(1E-6, 2E6), 100)
 

### 2016-09-16 R/wright_2004.R

@@ -56,10 +56,11 @@ process_wright_2004 <- function(filename, sitevars) {
   data[["lma"]] <- data[["lma"]]/1000
   ## Convert to Kg from g
   data[["n.area"]] <- data[["n.area"]]/1000
-  ## Convert to mol/kg/yr from micro-mol/g/s
-  data[["a.area"]] <- (data[["a.area"]] * 31557.6) * 10^-6
-  ## Convert to mol/kg/yr from micro-mol/g/s
-  data[["rd.area"]] <- (data[["rd.area"]] * 31557.6) * 10^-6
+  ## Convert to mol/m2/d from micro-mol/m2/s
+  data[["a.area"]] <- (data[["a.area"]] * 24 * 3600) * 10^-6
+
+  ## Convert to mol/m2/d from micro-mol/m2/s
+  data[["rd.area"]] <- (data[["rd.area"]] * 24 * 3600) * 10^-6
 
   ## Convert to years from month
   data[["leaflifespan"]] <- data[["leaflifespan"]]/12

### 2016-09-20 ms/figure_conceptual.tex
Refinements of text and figures
Full draft
- Add abstract, conclusions, minor edits throughout
- Improve figure and table captions
- Add Remko as coauthor
- Add tables in SuppMat, details about hyper-parameterisation
- Better table of contents in SuppMat

@@
   \node at ($(thisrow) + (-0.75cm, -0.3cm)$)[anchor=west, text width=2cm] {
- {\color{orange!50}\begin{equation}\label{#1}\end{equation}}};
+ {\color{black!30}\begin{equation}\label{#1}\end{equation}}};
 }
@@
 \firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/net_mass_production_dt_net_mass_production_dt}{}
-\othercol{2}{\scriptsize $\quad y$}{output/con/net_mass_production_dt_yield}{}
+\othercol{2}{\scriptsize $\quad c$}{output/con/net_mass_production_dt_yield}{}
 \othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/con/net_mass_production_dt_assimilation}{}
@@
 
-\tikzstyle{s1}=[->,line width=2pt, draw=orange!50]
+\tikzstyle{s1}=[->,line width=2pt, draw=black!20]
 

### 2016-09-20 ms/ms-common-defs.tex
Writing introduction and model
Allocation model and figures
Refinements of text and figures
Towards a full draft
Full draft
- Add abstract, conclusions, minor edits throughout
- Improve figure and table captions
- Add Remko as coauthor
- Add tables in SuppMat, details about hyper-parameterisation
- Better table of contents in SuppMat

@@
 \newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\sma}{\textsc{sma}}
 
-\title{Untangling the link between traits, size and growth rate in plants}
-\runninghead{I\lowercase{nfluence of traits on growth}}
+\title{Growth trajectories: how functional traits influence plant growth}
+\runninghead{H\lowercase{ow traits influence growth}}
 
-\author{Daniel S. Falster\textasteriskcentered \& Richard G. FitzJohn}
-
-\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia
-\\
-\textasteriskcentered Correspondence author. E-mail: \texttt{adaptive.plant@gmail.com}
+\author{Daniel S. Falster\textsuperscript{1,\textasteriskcentered}, Remko Duursma\textsuperscript{2}, \& Richard G. FitzJohn\textsuperscript{1,3}}
+\affiliation{\textsuperscript{1} Biological Sciences, Macquarie University NSW 2109, Australia \\
+\textsuperscript{2} Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia \\
+\textsuperscript{3} Imperial College, London, United Kingdom \\
+\textsuperscript{\textasteriskcentered} Correspondence author. E-mail: \email{adaptive.plant@gmail.com}
 }
@@
 
+\pdfminorversion=4 % tell pdflatex to generate PDF in version 1.4
+% Needed for problem-free conversion on manuscript central (arrgh)
+% http://timotheepoisot.fr/2013/06/24/manuscript-central-pdf-fix/
+
 \usepackage{graphicx}
@@
 
-% We will generate all images so they have a width \maxwidth. This means
+% We will generate all images so they have a max width and height. This means
 % that they will get their normal width if they fit onto the page, but
 % are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
-
 \let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=15cm,height=21cm,keepaspectratio]{#1}}

### 2016-09-20 ms/ms-supp.Rnw
Full draft
- Add abstract, conclusions, minor edits throughout
- Improve figure and table captions
- Add Remko as coauthor
- Add tables in SuppMat, details about hyper-parameterisation
- Better table of contents in SuppMat

@@
 \usepackage{listings}
-\usepackage[T1]{fontenc} % Better underscores, shockingly.
+\usepackage[T1]{fontenc} % Better underscores
 
@@
 
+% Figure command -- also adds figure to table of contents
+\newcommand{\myfigure}[4]{
+    \begin{figure}
+            \centering
+            \includegraphics{#1}
+            \caption{\textbf{#2} #3}
+            \label{#4}
+    \end{figure}
+  \addcontentsline{toc}{subsection}{Fig \ref{#4} -- #2}
+}
+
 \date{}
+
+<<preload, echo=FALSE>>=
+library(knitr)
+library(xtable)
+opts_chunk$set(echo=FALSE)
+@
 
@@
 
-\section{Supporting Tables}\label{app:supp_info_table}
+\section{A -- Additional details on the \texttt{FF16} growth model in the {\plant} package}
 
-\begin{table}[ht]
-\caption{Model parameters}
-\centering
-{\footnotesize  % smaller font, double space
-\begin{doublespace}
-% \include{table-pars}
+In this paper we use the \texttt{FF16} growth model from the package {\plant} \citep{Falster-2016}, which is accessed via \textsc{r} \citep{R-2015}.
+A full derivation of  Eqns 1-21 from the main text are described is provided in Appendix of \citet{Falster-2016}, see also \citep{Falster-2011}.
+Below we repeat some additional details from \citet{Falster-2016}, so that readers have a complete picture of the model.
 
-\end{doublespace}
-}
-\label{tab:params}
-\end{table}
+The \texttt{FF16} physiological model includes default values for all needed parameters (Table \ref{tab:params_core}).
+Species are known to vary considerably in many of these parameters, such as $\phi$, $\rho$, $\nu$, and $\omega$; so by varying parameters one can account for species differences.
+When altering a parameter in the model, however, one must also consider whether there are trade-offs linking parameters.
+This is achieved via a hyper-parameterisation function.
+
+\subsection{Hyper-parameterisation of the \texttt{FF16} model via traits}
+
+{\plant} allows for the hyper-parameterisation of the \texttt{FF16} physiological model via plant functional traits: this enables simultaneous variation in multiple parameters in accordance with an assumed trade-off.
+The hyper-parameterisation function used in the current analysis differs slightly from that in \citet{Falster-2016}.
+The functions implemented are as follows.
+
+\subsubsection{Leaf mass per unit area}
+
+The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$.
+In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citet{Wright-2004},
+$$k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.$$
+This relationship is normalised around $\phi_0$, the global mean of $\phi$.
+This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.
+
+We also vary the mass-based leaf respiration rate so that it stays constant per unit leaf area and varies with $\phi$ and nitrogen per unit leaf area $\nu$, as empirically observed by \citet{Wright-2004},
+$$r_{\rm l} = \frac{\beta_{{\rm lf4}}\, \nu}{\phi}.$$
+
+\subsubsection{Wood density}
+
+The trait wood density, denoted by $\rho$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$.
+In addition, we allow for $\rho$ to influence the rate of growth-independent mortality,
+$$d_{\rm I} = \beta_{{\rm dI1}} \, \left(\frac{\rho}{\rho_0}\right) ^ {-\beta_{{\rm dI2}}},$$
+and also the rate of sapwood turnover,
+$$k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^ {-\beta_{{\rm ks2}}}.$$
+As for $\phi$, these relationships are normalized around $\rho_0$, the global mean of $\rho$.
+By default, $\beta_{{\rm kI2}}$ and $\beta_{{\rm ks2}}$ are set to zero, so these linkages only become present when these parameters are set to something other than their default values.
+
+The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as
+$$r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},$$
+where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood. Similarly, the rate of bark respiration per unit mass varies as
+$$r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},$$
+with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.
+
+\subsubsection{Seed mass}
+
+Effects of the trait seed mass, denoted by $\omega$, are naturally embedded in the equation determining fecundity (Eq.
+\ref{eq:fecundity}) and the initial height of seedlings.
+In addition, we let the accessory cost per seed be a multiple of seed size,
+$$\alpha_{{\rm f3}} = \beta_{{\rm f1}} \omega,$$
+as empirically observed \citep{Henery-2001}.
+
+\subsubsection{Nitrogen per unit leaf area}
+
+Photosynthesis per unit leaf area and respiration rates per unit leaf mass (or area) are assumed to vary with leaf nitrogen per unit area, $\nu$.
+The calculation of respiration rates is already described above.
+To calculate the average annual photosynthesis for a leaf, we integrate the instantaneous rate per unit leaf area over the annual solar trajectory, using a rectangular-hyperbolic photosynthesis light response curve,
+
+$$p(\nu, E) = \frac1{365 \textrm{d}}\int_0^{365 \textrm{d}}\frac{Y(t) + A_{\rm max} - \sqrt{(Y(t) + A_{\rm max})^2 - 4 \beta_{{\rm lf2}} Y(t) A_{\rm max} }}{2 \beta_{{\rm lf2}}} {\rm d}t,$$
+where
+\begin{itemize}
+\item $A_{\rm max}$ is the maximum photosynthetic capacity of the leaf,
+\item $\beta_{{\rm lf2}}$ is the curvature of the light response curve,
+\item $Y(t) = \beta_{{\rm lf3}} I(t)$ is the initial yield of the light response curve, with $\beta_{{\rm lf3}}$ being the quantum yield parameter,
+\item $I(t) = k_{\rm I} \, I_0(t)\, E$ is the intensity of light on the leaf surface, and
+\item $I_0(t)$ is light incident on a surface perpendicular to the sun's rays directly above the canopy at time $t$.
+\end{itemize}
+The profile of $I_0(t)$ is given by a solar model adapted from \citet{TerSteege-1997}.
+
+We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as
+\begin{equation} \label{eq:amax}
+A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}},
+\end{equation}
+where $\beta_{{\rm lf1}}$ and $\beta_{{\rm lf5}}$ are constants. The relationships is normalized around $\nu_0$, the global mean of leaf nitrogen per unit area.
+
+Values of $p(\nu, E)$ are calculated across a range of values of $E$, and then an expression of the form in Eq. \ref{eq:photosynthesis} is fitted to extract the parameters
+$\alpha_{{\rm p1}}$ and $\alpha_{{\rm p2}}$, such that these become functions of $\nu$.
+
+\subsubsection{Implementation of hyper-parameterisation function}
+
+The hyper-parameterisation function used in this analysis is implemented as follows:
+
+<<echo=FALSE>>=
+make_hyperpar2
+@
+
+\clearpage
+
+\section{B -- Supporting Tables}\label{app:supp_info_table}
+
+\addcontentsline{toc}{subsection}{Table \ref{tab:params_core} -- Core parameters of the \texttt{FF16} physiological model}
+
+<<echo=FALSE,results='asis'>>=
+tab <- table_plant_parameters("../data/growth_model_pars_core.csv")
+print(xtable(tab,
+            hline.after=c(1),
+            align='lp{7cm}llll',
+            label="tab:params_core",
+            caption="Core parameter of the \texttt{FF16} physiological model."),
+      sanitize.text.function=I,
+      include.rownames=FALSE, caption.placement = "top")
+@
 
@@
 
-\section{Supporting Figures}\label{app:supp_info_figures}
+\addcontentsline{toc}{subsection}{Table \ref{tab:params_hyper} -- Parameters for hyper-parameterisation of the \texttt{FF16} physiological model}
 
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/lma_tradeoff.pdf}
-\caption{\textbf{Leaf turnover decreases with leaf-mass per unit leaf area.}
-Data from \citep{Wright-2004} for 678 species from 51 sites, each
+<<echo=FALSE,results='asis'>>=
+tab <- table_plant_parameters("../data/growth_model_pars_hyper.csv")
+print(xtable(tab,
+            hline.after=c(1),
+            align='lp{7cm}llll',
+            label="tab:params_hyper",
+            caption="Parameters for hyper-parameterisation of the \\texttt{FF16} physiological model."),
+      sanitize.text.function=I,
+      include.rownames=FALSE, caption.placement = "top")
+@
+
+\clearpage
+
+\section{C -- Supporting Figures}\label{app:supp_info_figures}
+
+\myfigure{SI/lma_tradeoff.pdf}{Leaf turnover decreases with leaf-mass per unit leaf area.}{Data from \citep{Wright-2004} for 678 species from 51 sites, each
 point giving a species-average. Lines show standardised major axis lines
 fitted to datsa from each site, with intensity of shading adjusted
-according to strength of the relationship.\label{fS-leaf}}
-\end{figure}
+according to strength of the relationship.}{fS-leaf}
 
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/growth_light_dia.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (stem-diameter growth).} Panels show predicted relationship between specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{fig:growth_light_dia}}
-\end{figure}
+\myfigure{SI/growth_light_dia.pdf}{Effect of four traits on stem-diameter growth rate for different-sized plants.}{
+Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
+Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
+For any given value of trait and $E$, plants were grown to the desired stem-diameter and their growth rate estimated.
+Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_dia}
 
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/growth_light_area.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (stem-area growth).} Panels show predicted relationship between specific trait and diameter growth rate, for a plant of specified
-diameter and under a range of shading environments.
-\label{fig:growth_light_area}}
-\end{figure}
+\myfigure{SI/growth_light_area.pdf}{Effect of four traits on stem-area growth rate for different-sized plants.}{
+Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
+Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
+For any given value of trait and $E$, plants were grown to the desired stem-area and their growth rate estimated.
+Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_area}
 
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics{SI/growth_light_mass.pdf}
-\caption{\textbf{Traits moderate the responsiveness of growth to changes
-in light environment (mass growth).} Panels show predicted relationship between
-specific trait and mass growth rate, for a plant of specified
-mass and under a range of shading environments.
-\label{fig:growth_light_mass}}
-\end{figure}
-
-\newpage
+\myfigure{SI/growth_light_mass.pdf}{Effect of four traits on plant mass growth rate for different-sized plants.}{
+Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
+Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
+For any given value of trait and $E$, plants were grown to the desired mass and their growth rate estimated.
+Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_mass}
 
@@
 
+\addcontentsline{toc}{section}{References}
 \bibliography{references}

### 2016-09-20 ms/ms.tex
Full draft
- Add abstract, conclusions, minor edits throughout
- Improve figure and table captions
- Add Remko as coauthor
- Add tables in SuppMat, details about hyper-parameterisation
- Better table of contents in SuppMat

@@
 
-\keywords{growth strategy, life-history,  ontogeny, size, leaf mass per area, wood density, shade-tolerance, growth rate}
+\keywords{growth rate, life-history, ontogeny, size, leaf mass per area, wood density, shade-tolerance, seed mass, height at maturation, leaf nitrogen}
 
@@
 \section{Abstract}
-We present a new theoretical framework for understanding the effect of diverse functional traits on plant growth.
 
-By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation linking trait-based trade-offs to demography and thus provide a platform for understanding growth across diverse species around the world.
+Plant species differ in many functional traits, including aspects of tissue construction and energy allocation, that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover.
+Yet, it remains unclear how -- and even if -- such traits influence whole-plant growth, with the simple linear relationships predicted by existing theory often lacking empirical support.
+Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth, extending a widely-used theoretical model that links growth rate in seedlings with a single leaf trait to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
+Using a model based on energy uptake and allocation, we explain why the influence of prominent traits on growth rate often varies with plant size and why the impact of size varies among traits.
+Considering a plant's growth rate in height, we find the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) is present across sizes but is strongest at intermediate sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
+Our theoretical model shows how these changes in trait-growth relationships with size arise via changes in biomass partitioning and reproductive allocation.
+Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results.
+By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.
 
@@
 
-One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on mass production \citep{Enquist-2007}.
+One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Enquist-2007}.
 By contrast, the physiology of traits such as leaf mass per unit leaf area and wood density suggests they have influence by altering the allocation of energy among different tissues, rather than changing mass production \citep{Falster-2011, Gibert-2016}.
@@
 For example, research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-Theory focussing on mass production is therefore of limited potential use.
+Theory considering only primary production is therefore of limited potential use.
 
@@
 This model can be viewed as an integration of several related approaches \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
-Our primary focus in this article is to explain the pattern that has been gradually emerging: that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
+Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
 As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth.
@@
 We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
-We submit that, trait-based approaches -- which focus on differences among species -- should seek to be integrated within a general model of plant growth, also able to capture patterns of growth through ontogeny.
+Our view is that trait-based approaches -- which focus on differences among species -- should be integrated within a general model of plant growth, also able to capture patterns of growth through ontogeny.
 Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
@@
 We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant.
-Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}}t$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig.
+Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig.
 \ref{fig:conceptual}a, eqn \ref{eq:dbdt}).
@@
 While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
-as functions of temperature, such detail is not needed here, as even with detailed models, $\frac{{\rm d}B}{{\rm d}}t$ always comes down to simple subtraction of income and losses.
+as functions of temperature, such detail is not needed here, as even with detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.
 
@@
 
-Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance, sometimes referred to as the the ``whole-plant-light-compensation point'' ({\wplcp}): the  the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013}.
-In general, average assimilation rate per leaf area $\bar{p}$ is expected to increase with light level or canopy openness, $E$.
-The {\wplcp} can be estimated by solving the value $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$.
+Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance, sometimes referred to as the ``whole-plant-light-compensation point'' ({\wplcp}): the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013}.
+In general, average assimilation rate per leaf area $\bar{p}$ increases with light level or canopy openness, $E$.
+The {\wplcp} can be estimated by solving for the the value of canopy openness $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$ (Fig.
+\ref{fig:wplcp_idea}).
 From eqn \ref{eq:dbdt}, this occurs when
@@
 
-It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are derived derived using the chain rule of differential calculus, they are mathematically true, and must therefore hold for a wide range of potential growth models.
-To make explicit predictions via this framework, however, requires a explicit model of plant construction and function; i.e.
+It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are derived using the chain rule of differential calculus, they are mathematically true and must therefore hold for any potential growth model.
+To make explicit predictions via this framework, then requires an explicit model of plant construction and function; i.e.
 we must put equations behind all there terms in Fig.
@@
 We could also call it ``isometric'', because the assumptions preserve the proportions among the relevant variables as the plant grows.
-Table \ref{tab:allometry} provides key equations of the model (see Supplementary Material \ref{sec:ff16} or \citet{Falster-2016} for full derivation).
+Table \ref{tab:allometry} provides key equations of the model (see \citet{Falster-2016} for full derivation).
 In particular, we assume:
@@
 
-Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height, described in Table \ref{tab:allometry}b.
+Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
 Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
@@
 \end{equation}
-where $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits, i.e. properties that are assumed approximately constant through ontogeny.
+where $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits (Table \ref{tab:traits}), i.e. properties that are assumed approximately constant through ontogeny.
 
@@
 Moreover, we assume larger seeds result in larger seedlings; the seed-size number tradeoff thereby translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
-
-\item[Height at maturation, $H_{\rm m}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction.
-We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{\rm m}$ via a function $r(H, H_{\rm m}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{\rm m}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-% By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{\rm m}$ (Fig. \ref{fig:S3}).
-
+\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction.
+We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{{\rm mat}}$ via a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
 \item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
 Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.\footnote{Add plot of this relationship, from data in \citet{Wright-2004}.}
-
 \item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
@@
 The rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
-
 \item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
@@
 The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits.
-Parameters used in the current analysis were sourced from \citet{Falster-2016}, with the exception of XXXX (see Table \ref{tab:params} for values).
+To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for full details see Supplementary Materials -- A).
+
+For the most part, parameters used in the current analysis were sourced from \citet{Falster-2016} (see Table \ref{tab:params} for values).
+The only exceptions are i))parameters effecting the relationships outlined in Table \ref{tab:allometry}a, which were fit using data described below, and ii) parameters describing the function for reproductive allocation.
+By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
+\ref{fig:conceptual}).
 
@@
 
-The functional-balance assumptions outlined in Table \ref{tab:allometry}a are well-supported by the available data, at least to first degree.
+The functional-balance assumptions outlined in Table \ref{tab:allometry}a are well-supported by the available data.
 Fig.
@@
 The panels in Fig.
-\ref{fig:conceptual} show the expected patterns for a typical woody plant.
+\ref{fig:conceptual} show the expected patterns for a typical woody plant, obtained by applying the functional-balance model of function (Table \ref{tab:allometry}).
 Height growth shows a hump-shaped pattern with size, first increasing then decreasing.
-This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size (Fig.
-\ref{fig:conceptual}), including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig.
-\ref{fig:conceptual}), and increasing reproductive allocation.
+This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size, including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig.
+\ref{fig:conceptual}), and increasing reproductive allocation (Fig.
+\ref{fig:conceptual}).
 In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
 Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}, and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
-All growth measures decrease sharply with size when expressed as relative growth rates.
+All growth measures decrease sharply with size when expressed as relative growth rates (results not shown).
 
@@
 We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
-For the five different traits, we observed four relatively distinct types of response.
-These responses are summarised in Table \ref{tab:responses} and described further below.
 The first two traits considered modify behaviour primarily at the start and end of an individuals growth trajectory, and are therefore termed ``ontogenetic traits''.
 The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individuals growth trajectory.
+Across the five different traits, we observed four relatively distinct types of response.
+These responses are summarised in Table \ref{tab:responses} and described further below.
 
@@
 
-\noindent{\bf Seed size, $\omega$:}
+\noindent{\bf Seed size, $\omega$.}
 Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease.
@@
 \ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
-Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in diameter, height, and mass), because of their greater total leaf area.
+Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates in all metrics  because of their greater total leaf area.
 At the same time, relative growth rate is predicted to decrease with size, because the ratio of leaf area to support mass decreases with plant size.
@@
 
-\noindent{\bf Height at maturation, $H_{\rm m}$:}
-$H_{\rm m}$ moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
-Greater height at maturation ($H_{\rm m}$) can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
+\noindent{\bf Height at maturation, $H_{{\rm mat}}$.}
+$H_{{\rm mat}}$ moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
+Greater $H_{{\rm mat}}$ can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
 \ref{fig:growth_light_height}).
 At smaller sizes, there is no differentiation among species, because most individuals are focusing on growth.
-At larger sizes, individuals of some species are allocating a majority of their surplus energy to reproduction, which reduces their growth rate relative to those species with greater $H_{\rm m}$(Fig.
+At larger sizes, individuals of some species are allocating a majority of their surplus energy to reproduction, which reduces their growth rate relative to those species with greater $H_{{\rm mat}}$ (Fig.
 \ref{fig:growth_light_height}).
@@
 
-\noindent{\bf Leaf nitrogen content per leaf area, $\nu$}: The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given for any given light environment $E$ and does not vary with height (Fig.
+\noindent{\bf Leaf nitrogen content per leaf area, $\nu$.} The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given for any given light environment $E$ and does not vary with height (Fig.
 \ref{fig:growth_light_height}).
 As $E$ increases from low to high, the optimal $\nu$ also increases.
-The invariance of the growth-trait relationship with respect to size can be understood as follows.
+The invariance of the growth-trait relationship with respect to size arises as follows.
 The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rates.
-Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$
+Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
 %  $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}$.
@@
 
-\noindent{\bf leaf mass per unit leaf area, $\phi$}:
+\noindent{\bf Leaf mass per unit leaf area, $\phi$.}
 Unlike $\nu$, the response of growth rate to changes in $\phi$ varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
@@
 
+% Possible add some bits to suppmat?
 % Substituting $x=\phi$ into eqn \ref{eq:dg2}, we can come up with a more precise expressions for some of the terms.
@@
 
-\noindent{\bf Wood density, $\rho$}:
+\noindent{\bf Wood density, $\rho$.}
 As for leaf, cheaper stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates (Fig.
@@
 With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$.
-Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$.
+Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$\footnote{Add derivation to suppmat}.
 The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm kl2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
@@
 Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by light environment.
-The {\plant} model, used here, builds on and extends several related approaches \citet[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits.
-This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which attempts to derive everything from a single master ``scaling'' equation for mass growth.
-Thus far, the \textsc{mst} has as been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}.
+The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
+This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}.
 Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects.
@@
 \ref{fig:conceptual}b.
-A well-developed trait-based -- or some might say evolutionary -- trade-off must include two opposing forces, that operate at some point in the organisms life cycle.
+While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
+A well-developed  trade-off must include two opposing forces, that operate at some point in the organisms life cycle.
 It is not necessary that benefit and cost both enter into Fig.
@@
 In that case, lower wood density would always increase growth rate, while higher wood density would always increase survival.
-However, while the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
 
-Our framework also highlights what is needed for traits to impact on growth rates and shade-tolerance.
+Our framework also highlights what is needed for traits to impact on growth rate and shade-tolerance.
 While traits can influence many aspects of plant function, these  influences must operate via the pathways outlined in Fig.
@@
 In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig.
-\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem-diameter, stem-area or or total mass.
+\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem-diameter, stem-area or total mass.
 This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size.
@@
 \ref{fig:conceptual}a).
-Decreasing $\phi$ has an overwhelmingly positive effect on height and diameter growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
-As plants increase in size, however, the amount of supporting tissue increases, decreasing the benefit of cheaper leaf construction in eqn \ref{eq:daldmt}.
+Decreasing $\phi$ then has an overwhelmingly positive effect on height and diameter growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
+As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheaper leaf construction in eqn \ref{eq:daldmt}.
 Consequently, the effect of $\phi$ on leaf turnover comes to dominate at larger sizes, and as such, the effect of $\phi$ on height, diameter and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
@@
 First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
-Much of current ecosystem research focusses on  photosynthesis, respiration, and resultant fluxes of carbon, with little attention devoted to allocation and turnover \citep[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}.
-Yet, for four of the five traits considered here, trait values do not influence photosynthesis, respiration, or net primary production (=photosynthesis - respiration).
+Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with little attention devoted to allocation and turnover \citep[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}.
+Yet, for four of the five traits considered here, trait values do not influence net primary production.
 In fact, the analysis with $\phi$ shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
@@
 To the extent the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum from slow to fast growth.
-Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{CSR} triangle, the r-K spectrum, and coexistence models base on the Lotka- Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.
-Researchers using functional traits have also tended to describe species as fast or slow growing \citep[e.g.][]{Adler-2014,Diaz-2016}.
+Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.
+Researchers using functional traits have also tended to describe species as fast or slow growing \citep[e.g.][]{Adler-2014, Diaz-2016}.
 Our results suggest a more-nuanced approach.
@@
 
-Third, our results suggest that even if traits define a potential growth trajectory, researchers seeking to link traits to growth rate must probe deeper into the data than simply looking for a linear relationship.
+Third, our results suggest that even if traits define a potential growth trajectory, researchers seeking to link traits to growth rate must probe deeper into the data than simply looking for a linear relationship between traits and average growth, to recover the expected relationships.
 None of the predicted relationships between traits and growth is linear across the range of sizes and light environments sampled.
-As such,  we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits.
-Controlling for size, site and light environment will therefore be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}.
+As such, we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits \citep[e.g.][]{Poorter-2008,Paine-2015}.
+Controlling for size, site and light environment will be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}, as will having a clear expectation for the hypothesised relationship.
+
+While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
+To that end, we can make a further prediction arising from our results, which is that the trait $\phi$ should increase through ontogeny for all individuals, across all species.
+Such shifts have been observed across a variety of species \citet{King-1999,Thomas-1999,Koch-2004}.
+The prediction arises from the fact that as plants grow larger, the value of cheap leaf construction diminishes, compared to the costs of high leaf turnover.
+It therefore pays for plants, but especially those with low $\phi$, to increase their $\phi$ as they grow larger.
+\citet{King-1999} made a similar prediction for a single species of \emph{Eucalyptus}, but here we can extend the idea across species.
+While trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
+This hypothesis attempts to give meaning to some of that variation, and shows how variation across and within species might be understood within a single framework.
 
@@
 
-It is also worth comparing our approach to other frameworks.
 As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citet{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}, and in particular those by \citet{Makela-1997} and \citet{Moorcroft-2001}.
@@
 Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
-\citet{King-1999} also connected his model of growth to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
+\citet{King-1999} also connected his model of growth for a single species to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
 \ref{fig:growth_light_height}).
@@
 growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant (??refs XX??).
+This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant
+% TODO: ??refs XX??)
+.
 The second assumption we make is that when mature, plant's allocate a substantial fraction of their surplus energy to reproduction.
@@
 Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}.
-Reproductive allocation is rarely given much attention in models of ecosystem flux.
+Reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
 For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
@@
 Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}.
-In this framework, mass-based growth is expressed as a 3/4 scaling power of total plant mass, and diameter growth is expressed as a 1/3 power of stem diameter.
-An extension of the model, connecting it to eqn \ref{eq:RGR} for relative growth rate, makes the connection to traits more explicit \citep{Enquist-2007}.
+In this framework, mass-based growth is expressed as a $\frac{3}{4}$ scaling power of total plant mass, and diameter growth is expressed as a $\frac{1}{3}$ power of stem diameter.
+An extension of the model makes the connection to traits more explicit, and also links the model to the seedling model for relative growth rate in eqn \ref{eq:RGR} \citep{Enquist-2007}.
 Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
@@
 
-\subsection{A prediction and closing remarks}
-While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
-So let us conclude with a new, testable prediction arising from our results: which is that the trait $\phi$ should increase through ontogeny.
-This prediction arises from the fact that as plants grow larger, the value of cheap leaf construction diminishes, compared to the costs of high leaf turnover.
-It therefore pays for all plants, but especially those with low $\phi$, to increase their $\phi$ as they grow larger.
-While trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
-This hypothesis attempts to give meaning to some of that variation.
+\subsection{Closing remarks}
 
-So we should instead be thinking of a species traits as defining a trajectory of potential growth for an individual, and one that is moderated by plant size and light environment.
-
-Towards a global model of growth:
-
-- mechanistic model of height and dbh growth which accounts for effects of size, light environment and traits
-- need to account for size long recognized, but not straightforward
-- look at growth in wide range of measures: mass, height, dbh
-- Compared to previous work, shows effects of traits more about
-  allocation than mass production
+Overall we have shown how diverse phenomena related to plant growth can be understood via a model accounting for processes generating photosynthetic income and allocating this among different tissues.
+The need to consider effects of size, alongside trait-based differences among species, has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001,, Westoby-2002, Enquist-2007}.
+Here we have provided a framework for achieving this.
+By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.
 
@@
 
-We thank J Camac, R Duursma, A Gibert, G Kunstler, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
+We thank J Camac, A Gibert, G Kunstler, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
 Falster was supported by an Australian Research Council discovery grant (DP110102086).
@@
 
-\subsection{A -- Full description of the \texttt{FF16} growth model}\label{sec:ff16}
+\noindent A -- Additional details on the \texttt{FF16} growth model
 
-\subsection{B -- Supplementary Figures}\label{sec:SM_figs}
+\noindent B -- Supplementary Tables
 
-\subsection{C -- Supplementary Tables}\label{sec:SM_tabs}
+\noindent C -- Supplementary Figures
+
+\vspace{10em}
+
+\section{Tables and Figures}
 
@@
 
-\section{Tables and Figures}
-
 \begin{table}[ht]
-\caption{Key phenomena}
+\caption{\textbf{Key empirical phenomena explained via the framework presented in this paper.}
+Select references are provided for each phenomena, with preference for meta-analyses, where available.
+The traits considered are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$), leaf mass per area ($\phi$)$^1$, and wood density ($\rho$).
+See Tables \ref{tab:definitions} and \ref{tab:traits} for further details.
+Abbreviations: \textsc{rgr}=relative growth ratel \textsc{lai} = leaf area index.
+}
 {\footnotesize
 {\centering
-  \begin{tabular}{p{9cm}p{6cm}}
+  \begin{tabular}{p{0.1cm}p{9cm}p{6cm}}
   \\
   \hline
-  Phenomena & References \\
+  \multicolumn{2}{l}{ Phenomena} & References \\
   \hline
+  \\
+  \multicolumn{3}{l}{\bf{Change in growth rate with increasing size} (Fig. \ref{fig:conceptual})}\\
+  & Net biomass production: Hump-shaped & \citet{Givnish-1988, Koch-2004} \\
+  & Plant mass: Increasing & \citet{Sillett-2010, Stephenson-2014} \\
+  & Height: Hump-shaped &  \citet{Ryan-2006, Sillett-2010, King-2011} \\
+  & Stem-diameter: Hump-shaped &  \citet{Herault-2011} \\
+  & Stem-area: Increasing & \citet{Sillett-2010} \\
+  & \textsc{rgr} (all variables): decreasing & \citet{Rees-2010, Iida-2014}\\
+  \\
+  \multicolumn{3}{l}{\bf{Effect of traits on growth rate} (Fig. \ref{fig:growth_light_height})}\\
+  & $\omega$: low values produce smaller seedlings, resulting in lower absolute growth but higher \textsc{rgr}. & Meta-analysis by \citet{Gibert-2016} \\
+  & $H_{\rm m}$: low values slow growth, but only at larger sizes & Meta-analysis by \citet{Gibert-2016}\\
+  & $\nu$: low values increase growth irrespective of size, but only in high light & Meta-analysis by \citet{Gibert-2016}\\
+  & $\phi$: low values increase growth when small, not at mid-large sizes & Meta-analysis by \citet{Gibert-2016}\\
+  & $\rho$: low values increase growth, except at largest sizes & Meta-analysis by \citet{Gibert-2016}\\
 
-  \\ \multicolumn{2}{l}{\bf{Patterns of growth with increasing size}}\\
-  Hump-shaped pattern of net biomass production & \citet{Givnish-1988, Koch-2004} \\
-  Increasing rate of plant mass and stem area growth & \citet{Sillett-2010, Stephenson-2014} \\
-  Hump-shaped pattern of $H$, $D$, $A_{\rm l}$ \& $A_{\rm st}$ growth& \citet{King-2005, Ryan-2006, Sillett-2010, Herault-2011, King-2011} \\
-  Decreasing relative growth rate (for all variables) & \citet{Rees-2010, Iida-2014}\\
-  \\
-  \\ \multicolumn{2}{l}{\bf{Effect of traits on growth rate}}\\
-  Low $\omega$ species produce smaller seedlings, resulting in lower absolute growth rates but higher relative growth rates. & \citet{Turnbull-2012};  Meta-analysis by \citet{Gibert-2016};\\
-  Low $\phi$ species$^1$ have a growth advantage when small, that dissipates with increasing size & Meta-analysis by \citet{Gibert-2016}\\
-  Low $\rho$ species have a growth advantage at all sizes & Meta-analysis by \citet{Gibert-2016}\\
-  High $H_{\rm m}$ species have a growth advantage when large & Meta-analysis by \citet{Gibert-2016}\\
+  \\ \multicolumn{3}{l}{\bf{Responsiveness of growth rate to changes in light, $E$} (Fig. \ref{fig:growth_light_height})}\\
+  & $\nu$: high values respond more to $E$, changes ranking of growth rate across species & \\
+  & $\phi$: low values respond more to $E$, does not change ranking of growth rate across species & \\
+  & $\rho$: low values respond more to $E$, does not change ranking of growth rate across species & \citet{Ruger-2012}\\
 
-  \\ \multicolumn{2}{l}{\bf{Responsiveness of growth rate to light}}\\
-  Low $\rho$ species respond more to changes in light & \citep{Ruger-2012}\\
-
-  \\ \multicolumn{2}{l}{\bf{Shade tolerance}}\\
-  \textsc{wplcp} decreases with height & \citet{Givnish-1988}\\
-  Seedlings of higher $\nu$ species have higher \textsc{wplcp} &  \citet{Baltzer-2007}\\
-  Low $\phi$ species have higher \textsc{wplcp} & \citet{Poorter-2006, Baltzer-2007, Lusk-2008}\\
-  \quad $\rightarrow$ Stands of low $\phi$ species$^1$  have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
-  Low $\rho$ species have higher \textsc{wplcp} &  \citet{Osunkoya-1996}\\
+  \\ \multicolumn{3}{l}{\bf{Shade tolerance, \textsc{wplcp}}(Fig. \ref{fig:wplcp})}\\
+  & Decreasing with size & \citet{Givnish-1988, Lusk-2008}\\
+  & $\nu$: low values more shade tolerant &  \citet{Baltzer-2007}\\
+  & $\phi$: low values less shade tolerant & \citet{Poorter-2006, Baltzer-2007, Lusk-2008}\\
+  & \quad $\rightarrow$ stands have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
+  & $\rho$: low values less shade tolerant  &  \citet{Osunkoya-1996}\\
   \\ \hline
@@
 \begin{table}[ht]
- \caption{Key variables}
+ \caption{\textbf{Key variables of the \texttt{FF16} physiological model from {\plant}, and used in this analysis.}. For mass ($M$), respiration ($r$), and turnover ($k$)
+variables, subscripts refer to any of the following tissues: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,
+\({\rm r}\) = roots, \({\rm a}\) = all living tissue.
+For area \(A\) variables, subscripts refer to any of the following: \({\rm l}\) = leaves, \({\rm st}\) = total stem cross-section, \({\rm s}\) = sapwood cross-section, \({\rm b}\) = bark cross-section, \({\rm h}\) = heartwood cross-section.}
 \centering
 {\footnotesize 
-  \begin{tabular}{p{2cm}p{2cm}p{7cm}}
+\begin{tabular}{p{0.1cm}p{2cm}p{2.5cm}p{8cm}}
 \\
 \hline
-  Symbol & Unit & Description \\
+& Symbol & Unit & Description \\
 \hline
-\\
-  $B$   & kg  & biomass originating from parent plant\\
-  $M_i$ & kg  & mass retained on plant in tissue $i$\\
-  $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
-  $H$   & m  & height of plant\\
-  $D$   & m  & diameter of plant stem\\
-  $\alpha_y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
-  $P$ & kg yr$^{-1}$ & total plant photosynthesis \\
-  $R$ & kg yr$^{-1}$ & total plant respiration \\
-  $T$ & kg yr$^{-1}$ & total plant turnover \\
-  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthetic rate per unit area \\
-  $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
-  $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
-  \\\hline
+\multicolumn{4}{l}{\textbf{Environmental variables}} \\
+& $E$ & & Canopy openness\\
+\multicolumn{4}{l}{\textbf{State variables (may vary through ontogeny)}} \\
+& $H$ & m & Height of a plant\\
+& $B$ & kg & Biomass originating from parent plant\\
+& $D$ & m & Stem diameter\\
+& $M_i$ & kg & Mass of tissue type $i$ retained on plant \\
+& $A_i$ & m$^2$ & Surface area or area of cross-section of tissue type $i$\\
+& $p,\bar{p}$ & mol yr$^{-1}$ m$^{-2}$ & Photosynthetic rate per unit area \\
+& $r_i$ & mol yr$^{-1}$ kg$^{-1}$  & Respiration rate per unit mass of tissue type $i$ \\
+& $k_i$ & yr$^{-1}$ & Turnover rate for tissue type $i$ \\
+
+\multicolumn{4}{l}{\textbf{Traits (constant through ontogeny)}} \\
+& $\omega$ & kg & Seed mass  \\
+& $H_{{\rm mat}}$ & m & Height at maturation\\
+& $\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf nitrogen per unit leaf area  \\
+& $\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf mass per area \\
+& $\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & Wood density \\
+\multicolumn{4}{l}{\textbf{Other parameters (constant through ontogeny)}} \\
+& $c$ & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Adjusts for growth respiration and carbon content\\
+& $\eta_c$ & & Crown-shape parameter\\
+& $\theta$ &  & Sapwood area per unit leaf area\\
+& $\alpha_{\rm l1}$ & m & Height of plant with leaf area of 1m$^2$ \\
+& $\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Root mass per unit leaf area \\
+& $\alpha_{\rm b1}$ &  & Ratio of bark area to sapwood area\\
+\hline
 \end{tabular}
@@
 \begin{table}[ht]
- \caption{Key traits and trade-offs}
+ \caption{\textbf{Key trade-offs for the five traits considered here.}
+For each trait we list the benefit and cost of increased trait values, as encoded into the {\plant} model.
+ }
 \centering
@@
 \raggedright
-  \begin{tabular}{p{3cm}p{1.5cm}p{6cm}p{3cm}}
+  \begin{tabular}{p{3cm}p{3cm}p{3cm}p{4cm}}
   \\
   \hline
-  Name, Symbol & Unit & Trade-off & References \\
+  Name &  Benefit & Cost & References \\
   \hline
   \\
-  Seed mass, $\omega$ & kg & $\uparrow$ initial size seedlings {\bf vs.} $\downarrow$  maternal fecundity & \citet{Moles-2006}\\
+  Seed mass, $\omega$ & $\uparrow$ size seedlings & $\downarrow$  maternal fecundity & \citet{Moles-2006}\\
   \\
-  Height at maturation, $H_{\rm m}$ & m  & $\uparrow$ growth {\bf vs.} $\downarrow$ fecundity & \citet{Thomas-2011, Wenk-2015}\\
+  Height at maturation, $H_{{\rm mat}}$  & $\uparrow$ growth & $\downarrow$ fecundity & \citet{Thomas-2011, Wenk-2015}\\
 \\
-  Leaf nitrogen per unit leaf area, $\nu$ & kg m$^{-2}$ & $\uparrow$ photosynthesis in high light {\bf vs.} $\uparrow$ respiration rate & \citet{Wright-2004}\\
+  Leaf nitrogen per unit leaf area, $\nu$ & $\uparrow$ photosynthesis in high light & $\uparrow$ respiration rate & \citet{Wright-2004}\\
   \\
-  Leaf mass per unit leaf area, $\phi$ & kg m$^{-2}$ & $\downarrow$ cost of building leaf {\bf vs.} $\uparrow$ leaf turnover & \citet{Wright-2004}\\
+  Leaf mass per unit leaf area, $\phi$ & $\downarrow$ leaf turnover & $\uparrow$ cost building leaf & \citet{Wright-2004}\\
   \\
-  Wood density, $\rho$ &  kg m$^{-3}$ & $\downarrow$ cost of building stem {\bf vs.} $\uparrow$ mortality rate & \citet{Chave-2009}\\
+  Wood density, $\rho$ & $\downarrow$ sapwood turnover & $\uparrow$ cost building stem & \citet{Chave-2009}\\
   \\
@@
 \begin{table}[ht]
-\caption{Equations for a functional-balance allocation model.
-The first column of part {\textbf a} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of {\textbf b}, and in {\textbf b}can then be derived from the assumptions in {\textbf a}.}
+\caption{ \textbf{Equations for a ``functional-balance'' model of plant construction.}
+The first column of part \textbf{a} provides cores assumptions between various size metrics and leaf area.
+Eqns in the middle and right columns of \textbf{b}, and in \textbf{b} can then be derived from the assumptions in the left column of \textbf{a}.
+The column ``Eqn'' indicates equation numbers referred to in the main text.
+See Table \ref{tab:definitions} for a list of variable names and definitions.
+}
 \centering
@@
 {\footnotesize
-  \begin{tabular}{p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \\ \hline
-  Variable & Function & Marginal cost & Growth rate & Eqn\\ \hline
-  \\
-  \multicolumn{5}{l}{\textbf{a) Assumptions}} \\
-  Height &
+  \begin{tabular}{p{0.1cm}p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \\ \hline
+  & Variable & Function & Marginal cost & Growth rate & Eqn\\ \hline
+  \multicolumn{6}{l}{\textbf{a) Functional-balance assumptions}} \\
+  & Height &
     $H = \alpha_{\rm l1} \, A_{\rm l}^{0.5}$ &
@@
     $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
-  Sapwood area &
+  & Sapwood area &
     $A_{\rm s} = \theta \, A_{\rm l}$ &
@@
     $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
-  Bark area &
+  & Bark area &
     $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
@@
     $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
-  Root mass &
+  & Root mass &
     $M_{\rm r} = \alpha_{\rm r1} \, A_{\rm l}$ &
@@
     $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
-  Heartwood area & & &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\ \\
-
-  \multicolumn{5}{l}{\textbf{b) Derived quantities}} \\
-  Leaf mass &
+  & Heartwood area & & &
+    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\
+  \multicolumn{6}{l}{\textbf{b) Derived quantities}} \\
+  & Leaf mass &
     $M_{\rm l} = \phi \, A_{\rm l} $ &
@@
     $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
-  Sapwood mass &
+  & Sapwood mass &
     $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
@@
     $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
-  Bark mass &
+  & Bark mass &
     $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
@@
     $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
-  Heartwood area &
+  & Heartwood area &
     $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
@@
     $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
-  Heartwood mass &
+  & Heartwood mass &
     $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
@@
 \centering
-\caption{{\bf Predicted effects of traits on key elements of plant function determining growth rate.} Adapted and expanded from \citet{Gibert-2016}.
-Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
-Traits are: seed mass ($\omega$), height at maturation ($H_{\rm m}$), nitrogen content per leaf area ($\nu$), leaf mass per unit leaf area ($\phi$), and wood density ($\rho$) (see Table \ref{tab:traits} for more details).}
+\caption{\textbf{Predicted effects of traits on key elements of plant function determining growth rate.} Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
+Traits are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), nitrogen content per leaf area ($\nu$), leaf mass per unit leaf area ($\phi$), and wood density ($\rho$) (see Table \ref{tab:traits} for more details).
+Adapted and expanded from \citet{Gibert-2016}.}
 {\footnotesize
@@
   & & &  &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
-  & & &  {\bf Symbol} & \boldmath$\omega$ & \boldmath$H_{\rm m}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\ \hline
+  & & &  {\bf Symbol} & \boldmath$\omega$ & \boldmath$H_{{\rm mat}}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\ \hline
   \\
@@
   & \multicolumn{9}{l} {Height} \\
-  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
-  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \doup & \dodo & \\
-  & \multicolumn{9}{l} {Stem area {\color{red} (check)}} \\
-  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
-  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
+  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
+  & \multicolumn{9}{l} {Stem area} \\
+  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
+  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \dofl & \dodo & \\
   & \multicolumn{9}{l} {Stem diameter} \\
-  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \doup & \dodo & \\
-  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \doup & \dodo & \\
-  & \multicolumn{9}{l} {Mass {\color{red} (check, live or total?)}} \\
-  & & \tabitem{Absolute} & $\ud M_{\rm a} / \ud t$ & \upfl &  \flup & \upup & \doup & \dodo & \\
-  & & \tabitem{Relative} & $\ud M_{\rm a} / (\ud t . M_{\rm a} )$ & \dofl &  \flup & \upup & \doup & \dodo & \\
+  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
+  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
+  & \multicolumn{9}{l} {Mass} \\
+  & & \tabitem{Absolute} & $\ud M_{\rm t} / \ud t$ & \upfl &  \flup & \upup & \dofl & \dodo & \\
+  & & \tabitem{Relative} & $\ud M_{\rm t} / (\ud t . M_{\rm t} )$ & \dofl &  \flup & \upup & \dofl & \dodo & \\
   \\
@@
 \includegraphics{main/figure_conceptual.pdf}
-\caption{\textbf{Conceptual framework for understanding size- and trait-dependent
-changes in growth rate.} Hump-shaped relationship between height growth rate and
-size. \label{fig:conceptual}}
+\caption{\textbf{Conceptual framework linking growth rate to plant size and traits.}
+\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
+\textbf{b)} Equations describing the rates of biomass production and growth in various dimensions of the plant.
+In the first line the symbol $\sum$ means ``sum'', across tissues, where $i = {\rm l,b,sr}$.
+The grey numbers indicate equation numbers referred to in the main text.
+The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}.
+Colours highlight where the same metric appears repeatedly in different equations.
+For a full list of variable names see Table \ref{tab:definitions}.
+\label{fig:conceptual}}
 \end{figure}
@@
 \includegraphics{main/lcp_schematic.pdf}
-\caption{\textbf{Conceptual framework for understanding size- and trait-dependent
-changes in shade tolerance.} Do we need a figure? \label{fig:wplcp_idea}}
+\caption{\textbf{Conceptual framework linking shade-tolerance to plant size and traits.}
+Shade tolerance is the level of canopy openness  $E$ where photosynthetic income (dashed line) intersects with the sum of respiration and turnover costs over tissues (solid black lines).
+The black circles indicate the point of intersection for plants with different heights.
+All income and cost is expressed in units of dry mass produced  per unit leaf area per year.
+Note the costs of sapwood and bark increases with height.
+Traits can impact on shade tolerance by, in the case of leaf nitrogen content, shifting the income line up or down, or in the case of leaf mass per unit leaf area or wood density, causing the cost components to increase or decrease.
+Adapted and expanded from \citet{Givnish-1988}.
+\label{fig:wplcp_idea}}
 \end{figure}
@@
 \includegraphics{main/assumptions.pdf}
-\caption{\textbf{Key assumptions of a functional balance model for plant construction}.
-Data are from the \textsc{baad} \citep{Falster-2015b}.
-Each dot is a single plant.
-Lines show standardised major axis lines fitted by species, with intensity of shading adjusted according to strength of the relationship.
-Dashed lines indicate slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}.
-The dark line shows the relationship assumed in this paper.
+\caption{\textbf{Key assumptions of a functional balance model for plant construction.}
+Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
+Blue lines indicate the show standardised major axis of the bivariate cloud for each species.
+Dotted lines indicate slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}.
+The dark line shows the relationship assumed in the {\plant} model and applied throughout the paper.
 \label{fig:assumptions}}
@@
 \includegraphics{main/growth_light_height.pdf}
-\caption{\textbf{Three different responses of growth rate to changes in size, with traits.} Panels show predicted relationship between specific trait and height growth rate, for a plant of specified height and under a range of shading environments.
+\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
+Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}.
+Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
+For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated.
+Figs.
+\ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass.
+Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
 \label{fig:growth_light_height}}
@@
 \includegraphics{main/lcp_trait.pdf}
-\caption{\textbf{Effect of traits on shade tolerance.} Panels show effect of traits on level of canopy openness that causes net production (from eqn \ref{eq:dbdt}, Fig.
-\ref{fig:conceptual}) to be zero.
-Different lines indicate relationship for plants with specified height.
+\caption{\textbf{Effect of three ontogenetic traits on shade tolerance.}
+Panels show effect of traits on level of canopy openness that causes net production (eqn \ref{eq:dbdt}) to be zero.
+Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
 \label{fig:wplcp}}

### 2016-09-20 R/figures.R

@@ -31,7 +31,7 @@ make_hyperpar2 <- function(
                                 B_dI1=0.01,
                                 B_dI2=0.0,
                                 B_ks1=0.2,
-                                B_ks2=0.0,
+                                B_ks2=1.25, #0.0,
                                 B_rs1=4012.0,
                                 B_rb1=2.0*4012.0,
                                 B_f1 =3.0,
@@ -40,8 +40,8 @@ make_hyperpar2 <- function(
                                 B_lf1=5120.738 * 1.87e-3 * 24 * 3600 / 1e+06,
                                 B_lf2=0.75,
                                 B_lf3=0.04,
-                                B_lf4=21000,
-                                B_lf5=1,
+                                B_lf4=21000*0.75,
+                                B_lf5=0.5,
                                 k_I=0.5,
                                 latitude=0) {
   assert_scalar <- function(x, name=deparse(substitute(x))) {
@@ -206,7 +206,7 @@ default_parameters <- function() {
     k_l  = 0.4565855 / 3)
 
   FF16_Parameters(
-    hyperpar=make_hyperpar2(B_ks2=1.25, B_lf5=0.5, B_lf4=21000*0.75),
+    hyperpar=make_hyperpar2(),
     strategy_default = s
     )
 }

### 2016-09-20 R/tables.R

@@ -1,9 +1,44 @@
-table_plant_parameters <- function(plant_parameters, dest){
-  x <- xtable(plant_parameters[-c(1,2,4,7,10,21,22,25:31),1:4],
-              hline.after=c(1),
-              align='lp{5cm}lll')
-  y <- print(x, sanitize.text.function=as.character,
-             include.rownames=FALSE, floating=FALSE,
-             print.results=FALSE)
-  cat(y, file=dest)
+table_plant_parameters <- function(filename){
+  tab <- read.csv(filename, stringsAsFactors=FALSE,
+                  check.names=FALSE, comment.char="#")
+
+  to_bold <- tab$Symbol == ""
+  tab$Description[to_bold] <- sprintf("\\textbf{%s}", tab$Description[to_bold])
+
+  to_math <- tab$Symbol != ""
+  tab$Symbol[to_math] <- sprintf("$%s$", tab$Symbol[to_math])
+
+  to_math <- grepl("[_^]", tab$Unit)
+  tab$Unit[to_math] <-
+    sprintf("$%s$",
+            gsub("(mol|kg|m|yr|d)", "\\\\mathrm{\\1}", tab$Unit[to_math]))
+  tab$Unit[to_math] <-
+            gsub(" ", "\\,", tab$Unit[to_math], fixed =TRUE)
+
+  s <- c(default_strategy(),
+         lapply(as.list(args(make_hyperpar2)), eval))
+
+  tab$Value <- NA_real_
+
+  # Now extract actual values from plant
+  to_code <- tab[["Code"]] != "" & tab[["Code"]] %in% names(s)
+  code <- tab[["Code"]][to_code]
+
+  val <- s[code]
+  err <- sapply(val, is.null)
+  if (any(err)) {
+    message("Missing parameters: ", paste(code[err], collapse=", "))
+  }
+
+  oo <- options(scipen=999)
+  on.exit(options(oo))
+  tab$Value[to_code] <- unlist(lapply(val, prettyNum))
+
+  # Display names in latex
+  to_code <- tab[["Code"]] != ""
+
+  tab[["Code"]][to_code] <- sprintf("\\texttt{%s}",
+                                      gsub("_", "\\\\_", tab[["Code"]][to_code]))
+
+  tab
 }

### 2016-09-22 R/build.R
Import call_system function from callr to avoid package dependency
Another package has recently appeared using name callr, so installation becomes fraught.
For now I thought it safer to remove the dependency.

@@ -64,11 +64,11 @@ run_latex <- function(filename, interaction="nonstopmode", engine="pdflatex") {
   args <- c(paste0("-interaction=", interaction),
             "-halt-on-error",
             filename)
-  callr::call_system(Sys_which(engine), args)
+  call_system(Sys_which(engine), args)
 }
 
 run_bibtex <- function(filename) {
-  callr::call_system(Sys_which("bibtex"), sub(".tex$", "", filename))
+  call_system(Sys_which("bibtex"), sub(".tex$", "", filename))
 }
 
 Sys_which <- function(x) {
@@ -79,3 +79,85 @@ Sys_which <- function(x) {
   ret
 }
 
+
+##' Function imported from callr package;  makes it easy to call a
+##' system command from R and have it behave.
+##'
+##' This function uses \code{system2} to call a system command fairly
+##' portably.  What it adds is a particular way of dealing with
+##' errors.  \code{call_system} runs the command \code{command} with
+##' arguments \code{args} (and with optionally set environment
+##' variables \code{env}) and hides \emph{all} produced output to
+##' stdout and stderr.  If the command fails (currently any nonzero
+##' exit code is counted as a failure) then \code{call_system} will
+##' throw an R error giving
+##' \itemize{
+##' \item the full string of the command run
+##' \item the exit code of the command
+##' \item any \code{errmsg} attribute that might have been returned
+##' \item all output that the program produced to either stdout and
+##' stderr
+##' }
+##'
+##' This means that a successful invocation of a program produces no
+##' output while the unsuccessful invocation throws an error and
+##' prints all information to the screen (though this is delayed until
+##' failure happens).
+##'
+##'
+##' \code{call_system} also returns the contents of both stderr and
+##' stdout \emph{invisibly} so that it can be inspected if needed.
+##'
+##' The function \code{run_system} does the same thing and will be
+##' removed as soon as code that depends on it is out of use.
+##'
+##' @title Run a system command, stopping on error
+##' @param command The system command to be invoked, as a character
+##' string.  \code{\link{Sys.which}} is useful here.
+##' @param args A character vector of arguments to \code{command}
+##' @param env A character vector of name=value pairs to be set as
+##' environment variables (see \code{\link{system2}}).
[…]

### 2016-10-13 ms/ms.tex
Incorporate suggestions from R Duursma throughout article

@@
 \externaldocument{figure_conceptual}
-% Needed to set equation counter, following inclusion of figure 1
-\setcounter{equation}{7}
 
@@
 
-Plant species differ in many functional traits, including aspects of tissue construction and energy allocation, that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover.
+Plant species differ in many functional traits that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover.
 Yet, it remains unclear how -- and even if -- such traits influence whole-plant growth, with the simple linear relationships predicted by existing theory often lacking empirical support.
-Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth, extending a widely-used theoretical model that links growth rate in seedlings with a single leaf trait to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
-Using a model based on energy uptake and allocation, we explain why the influence of prominent traits on growth rate often varies with plant size and why the impact of size varies among traits.
-Considering a plant's growth rate in height, we find the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) is present across sizes but is strongest at intermediate sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
-Our theoretical model shows how these changes in trait-growth relationships with size arise via changes in biomass partitioning and reproductive allocation.
+Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth and shade-tolerance, extending a widely-used theoretical model that links growth rate in seedlings with a single leaf trait to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
+Based on biomass production and allocation, this framework explains why the influence of prominent traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
+Considering  growth rate in height, we find the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes but is strongest at intermediate sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
 Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results.
@@
 
+% XX - The main comment I have is whether the strength of the 'correlation' between traits and growth rates can be summarized in some metric, so that it can be directly shown.
+To me Fig 4 is not very clear on showing how size affects the trait-growth correlations.
+For example, for wood density, I see a relation for all size classes.
+For the largest trees, it may look like a weaker correlation but isn't that because growth rates are lower?
+The same with Narea, to me they all look the same (except again, larger trees grow slower but it’s hard to tell there for me whether the trait-growth correlation is any different)
+% XX -  The shade tolerance (WPLCP) analysis seems off-topic (it also not mentioned in the abstract).
+Can this be integrated more closely with the main objective, i.e.
+correlation between traits and growth rate and its dependence on size?
+
 In this article we outline how the growth of individual plants is influenced by various functional traits, also the plant's size and light environment.
@@
 During the last two decades, trade-offs related to some prominent traits have been quantified, with values for traits such as leaf mass per unit leaf area, wood density, and seed size now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
-As data has accumulated, researchers have looked to traits as potential indicators of plant growth and demography \citep{Poorter-2008, Wright-2010, Adler-2014}.\footnote{Add references from applied fields, such as invasive species, conservation.}
+As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
 
 While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the trait leaf mass per unit leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}.
-leaf mass per unit leaf area and its correlate leaf lifespan were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
+In seedlings, the trait leaf mass per unit leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly related to relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a mathematical model of growth rate (see below).
+Leaf mass per unit leaf area and its correlate leaf lifespan were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
 These early successes prompted researchers to search for similar relationships in large plants.
@@
 
-It is now becoming increasingly clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
+Recently, it has become clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
 A recent meta-analysis of 103 studies reporting \textgreater 500 correlations by \citet{Gibert-2016} provides the most compelling evidence.
@@
 
-The challenge of interpreting diverse empirical results seeking to link traits to growth rate was all the harder until the paper by \citet{Gibert-2016}, because we lacked any clear theoretical expectations on why the effect of traits may be moderated by size.
+The challenge of interpreting diverse empirical results seeking to link traits to growth rate was all the harder because, until the paper by \citet{Gibert-2016}, we lacked any clear theoretical expectations on why the effect of traits may be moderated by size.
 Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
-A widely-used model for seedlings suggests a seedling's relative growth rate in mass is linearly and negatively related to the trait leaf mass per unit leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to the trait leaf mass per unit leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
 An extension of the model suggests a similar relationship should hold at larger sizes \citep{Enquist-2007}.
-But as described above, the prediction for large plants is not supported by empirical results.
+But as noted above, the prediction for large plants is not supported by empirical results.
 Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
@@
 
-One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Enquist-2007}.
-By contrast, the physiology of traits such as leaf mass per unit leaf area and wood density suggests they have influence by altering the allocation of energy among different tissues, rather than changing mass production \citep{Falster-2011, Gibert-2016}.
-A second concern for theory focussed on mass production is that altered mass production is only possible outcome.
-Measuring mass production is really only practical for small plants that can be easily harvested.
+One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Wright-2000,Enquist-2007}.
+By contrast, the physiology of traits such as leaf mass per unit leaf area and wood density suggests they have influence by altering the allocation of energy among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
+A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants that can be easily harvested.
 For larger plants, growth is more-often measured as increment in either diameter or height.
 For example, research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-Theory considering only primary production is therefore of limited potential use.
+Theory considering only primary production therefore offers limited insight.
 
-In this paper we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth (Table \ref{tab:phenomena}); and thereby offer fresh insight into the way traits influence plant demography throughout across the life-cycle.
-This model can be viewed as an integration of several related approaches \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
+In this paper we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and shade tolerance (Table \ref{tab:phenomena}); and thereby offer fresh insight into the way traits influence plant demography across the life-cycle.
+Broadly, the {\plant} model builds off several related approaches, focussing on production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
 Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
@@
 Our view is that trait-based approaches -- which focus on differences among species -- should be integrated within a general model of plant growth, also able to capture patterns of growth through ontogeny.
-Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.
-In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
+Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}.
+In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
 And all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
@@
 
-The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
+The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
 \ref{fig:conceptual}a).
@@
 The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A summary of all variables, units and definitions is given in Table \ref{tab:definitions}-\ref{tab:traits}, with further details on the parameter values applied given in Table \ref{tab:params}.
+A summary of all variables, units and definitions is given in Table \ref{tab:definitions}-\ref{tab:traits}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
 
-We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant.
-Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig.
-\ref{fig:conceptual}a, eqn \ref{eq:dbdt}).
-Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}$) and total leaf area ($A_{\rm l}$).
-The constant $y$ in eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass.
-We assume that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ with traits (see below).
-While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
-as functions of temperature, such detail is not needed here, as even with detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.
+We assume growth is fundamentally driven by biomass production and its subsequent distribution throughout the plant.
+Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000}:
+\begin{equation}\label{eq:dbdt}
+\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
+= \underbrace{\strut \alpha_{\rm bio} }_\textrm{mass per C} \,
+\underbrace{\strut \alpha_{\rm y} }_\textrm{yield}
+\big( \underbrace{\strut A_{\rm l} \, \bar{p}(E)}_\textrm{photosynthesis} -
+\underbrace{\strut \, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}}_\textrm{respiration}\big)
+- \underbrace{\strut \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}}_\textrm{turnover}.
+\end{equation}
+Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}(E)$) and total leaf area ($A_{\rm l}$).
+We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (Fig.
+\ref{fig:wplcp_idea}; for details, see Supplementary Materials), and respiration and turnover rates of different tissues are constants that might differ with traits (see below).
+The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for growth respiration and the conversion of CO$_2$ into units of biomass, respectively.
+While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment, so that we can better understand the intrinsic trait- and size-related effects.
+Also, many vegetation models use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
+as functions of temperature.
+Such detail is not needed here because it will not change general model behaviour, even it would alter  he absolute values of predicted growth rates.
+Even with detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.
+
+\subsection{Original model for mass-based relative growth rate}
+
+Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+In fact, that model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows.
+For seedlings, which are young and mostly leaf, we first ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
+Net production (from eqn \ref{eq:dbdt}) then becomes a linear function of leaf area, making relative growth rate in mass a linear function of $\phi$:
+\begin{equation}\label{eq:RGR}
+\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx P_{\rm net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
+where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$.
+Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees: namely plant height ($H$), stem-basal area ($A_{\rm st}$) or stem diameter $D$.
+The derivations below makes these links clear.
 
@@
 
-To model growth in either height, leaf area, basal area, standing mass or stem diameter then requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
+% Update equation counter, following inclusion of figure 1
+\setcounter{equation}{9}
+
+To model growth in either plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) then requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
 Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}.
 Relevant equations are given in Fig.
-\ref{fig:conceptual}b, (eqns \ref{eq:dbdt}-\ref{eq:dDdt}).
+\ref{fig:conceptual}b, (eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
 As is evident in eqn  Fig.
-\ref{fig:conceptual}b, the growth rate in plant weight (eqn \ref{eq:dMtdt}), leaf area (eqn \ref{eq:daldt}), height (eqn \ref{eq:dhdt}), stem basal area (eqn \ref{eq:dast}), and stem diameter (eqn \ref{eq:dDdt}) all share some terms in common.
+\ref{fig:conceptual}b, the growth rate in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) all share some terms in common.
 Many of the terms in Fig.
@@
 The remaining  $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
-Presumably, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, close to 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants.
+In plants, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants \citep{Wenk-2015}.
 Note also that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
@@
 
-The growth rate in the total standing mass of the plant, $\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$ is then the sum of sapwood turnover (which is replaced by an equivalent amount) and increment in live mass (eqn \ref{eq:dMtdt}).
+The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
 
@@
 \begin{equation}\label{eq:wplcp}
-\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}},
+\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
 \end{equation}
-where $\bar{p}(E)$ is the rate of photosynthesis at canopy openness $E$.
 The {\wplcp} occurs at the points where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig.
@@
 
-It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are derived using the chain rule of differential calculus, they are mathematically true and must therefore hold for any potential growth model.
+It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication and differentiation, and must therefore hold for any potential growth model where biomass allocation is important.
 To make explicit predictions via this framework, then requires an explicit model of plant construction and function; i.e.
@@
 
-Here we adopt a model of plant construction and function that can be considered a first-order functional-balance model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
-We could also call it ``isometric'', because the assumptions preserve the proportions among the relevant variables as the plant grows.
-Table \ref{tab:allometry} provides key equations of the model (see \citet{Falster-2016} for full derivation).
-In particular, we assume:
+The {\plant} adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
+We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics \citep{Huxley-1932}.
+Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation).
+In particular, we assume that as a plant grows:
 \begin{description}
-\item[Assumption 1:] A log-log scaling relationship with slope of 0.5 between the leaf area of a plant and its height (eqn \ref{eq:ha}).
-\item[Assumptions 2:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of sapwood in a plant (eqn \ref{eq:as}).
-\item[Assumptions 3:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of bark and phloem in a plant (eqn \ref{eq:ab}).
-\item[Assumptions 4:] A log-log scaling relationship with slope of 1.0 between the leaf area and mass of fine roots in a plant (eqn \ref{eq:mr}).
-\item[Assumption 5:] The vertical distribution of leaf within the canopy of an individual plant remains constant through ontogeny.
+\item[Assumption 1:] Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha}).
+\item[Assumptions 2:] The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as}).
+\item[Assumptions 3:] The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab}).
+\item[Assumptions 4:] The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}).
+\item[Assumption 5:] The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
 \end{description}
@@
 
-An important prediction arising from the eqns in Table \ref{tab:allometry} is that the amount of live stem tissue supporting each unit of leaf area increases linearly with  height:
-\begin{equation}\label{eq:msml}
-\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H,
-\end{equation}
-where $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits (Table \ref{tab:traits}), i.e. properties that are assumed approximately constant through ontogeny.
-
 \subsection{Trait-based trade-offs}
@@
 \begin{description}
-\item[Seed mass, $\omega$:] There is a direct energetic trade-off between the a plant's fecundity and the size of it's seeds.
-Moreover, we assume larger seeds result in larger seedlings; the seed-size number tradeoff thereby translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
+\item[Seed mass, $\omega$:] There is a direct energetic trade-off between a plant's fecundity and the size of it's seeds.
+Moreover, we assume larger seeds result in larger seedlings; so this the seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
 \item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction.
@@
 \item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
-Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.\footnote{Add plot of this relationship, from data in \citet{Wright-2004}.}
+Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
 \item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
 In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-The rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
+Following \citet{Wright-2004}, the rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
 \item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
@@
 The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly whenever $\rho$ is varied.
+There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is more speculative than the equivalent assumption for leaves, which is well supported empirically.
  \end{description}
@@
 
-For the most part, parameters used in the current analysis were sourced from \citet{Falster-2016} (see Table \ref{tab:params} for values).
-The only exceptions are i))parameters effecting the relationships outlined in Table \ref{tab:allometry}a, which were fit using data described below, and ii) parameters describing the function for reproductive allocation.
+For the most part, parameters used in the current analysis were sourced from \citet{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
+The only exceptions are: i) parameters effecting the relationships outlined in Table \ref{tab:allometry}a, which were fit using data described below, and ii) parameters describing the function for reproductive allocation.
 By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
@@
 
-The analyses presented employ best practises in scientific computing, as defined by \citep{Wilson-2014}, and are fully reproducible, via code available at
+The analyses presented employ best practises in scientific computing, as defined by \citet{Wilson-2014}, and are fully reproducible, via code available at
 \href{https://github.com/traitecoevo/growth\_trajectories}{github.com/traitecoevo/growth\_trajectories}.
@@
 
-The functional-balance assumptions outlined in Table \ref{tab:allometry}a are well-supported by the available data.
+To verify model assumptions we compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the \textsc{baad} \citep{Falster-2015b}.
+Additionally, we evaluated an important prediction arising from the eqns in Table \ref{tab:allometry}a, that the amount of live stem tissue supporting each unit of leaf area increases linearly with height, as per the equation
+\begin{equation}\label{eq:msml}
+\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
+\end{equation}
+Note that $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits, i.e.
+properties that are assumed approximately constant through ontogeny for any given species.
 Fig.
-\ref{fig:assumptions} shows data sourced from the \textsc{baad} \citep{Falster-2015b}, with solid lines indicating {\sma} lines fit to each species in the data.
-Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a.
-As expected, species differed in elevation, but less so in the slope of the fitted lines, with slopes aligning with those predicted by the functional-balance assumptions.
+\ref{fig:assumptions} shows that the three functional-balance assumptions outlined in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
+The solid lines indicate {\sma} lines fit to each species in the dataset.
+Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a and eqn \ref{eq:msml}.
+As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions.
 
@@
 
-Our growth model suggests an intrinsically size-dependent pattern of growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
+Our growth model suggests an intrinsically size-dependent pattern of biomass-production and growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
 The panels in Fig.
-\ref{fig:conceptual} show the expected patterns for a typical woody plant, obtained by applying the functional-balance model of function (Table \ref{tab:allometry}).
-Height growth shows a hump-shaped pattern with size, first increasing then decreasing.
+\ref{fig:conceptual} show the expected patterns for a typical woody plant, obtained by applying the functional-balance model encoded in {\plant}.
+Biomass production shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
+Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
 This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size, including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig.
-\ref{fig:conceptual}), and increasing reproductive allocation (Fig.
-\ref{fig:conceptual}).
+\ref{fig:conceptual}), increasing reproductive allocation (Fig.
+\ref{fig:conceptual}), and declining mass production.
 In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
@@
 
-% Importantly, we can remove these size dependent effects by altering XXX of the assumptions in Table \ref{tab:allometry} (??).
+% Importantly, we can remove these size dependent effects by altering ??? of the assumptions in Table \ref{tab:allometry} (??).
 % \footnote{What are key assumptions that induce all size-related effects?
@@
 With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$.
-Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$\footnote{Add derivation to suppmat}.
+Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$.
 The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm kl2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
@@
 
-The list of functional traits that are known to differ among plant species is already long and ever-increasing \citep{Perez-2013}.
-While we have focussed on understanding the effects of five specific traits on growth rate, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
-The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass or carbon, and can therefore be embedded within the eqns in Fig.
-\ref{fig:conceptual}b.
-While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
-A well-developed  trade-off must include two opposing forces, that operate at some point in the organisms life cycle.
-It is not necessary that benefit and cost both enter into Fig.
-\ref{fig:conceptual}; for example the trait wood density is sometimes thought of a trade-off between the costs of tissue construction and the rate of stem mortality.
-In that case, lower wood density would always increase growth rate, while higher wood density would always increase survival.
-
-Our framework also highlights what is needed for traits to impact on growth rate and shade-tolerance.
-While traits can influence many aspects of plant function, these  influences must operate via the pathways outlined in Fig.
-\ref{fig:conceptual} if the trait is going to impact on growth.
-For example, many studies have focused on traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
-These traits will inevitably influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}) by altering conductivity of water to the leaf.
-The potential costs of larger vessel size might be higher rates of stem turnover, which would appear in the term $k_{\rm s}$ in eqn \ref{eq:dbdt}.
-The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in our framework via the parameter $\theta$ (Table \ref{tab:allometry}).
-The effect of both these traits on growth rate should be expected to vary with plant size.
-
-% The only circumstances whereby traits impact on growth but not via the pathways in Fig. \ref{fig:conceptual}, would be where growth rate was entirely disconnected from the plant's carbon budget.
-
-\subsection{Relationship to a classic model for seedling growth}
-
 The model presented here extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
 Importantly, the seedling model can be derived as a special case of our more-extended approach.
-For seedlings, which are young and mostly leaf, we can ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
-Net production (from eqn \ref{eq:dbdt}) then becomes a linear function of leaf area, making relative growth rate in mass a linear function of $\phi$:
-\begin{equation}\label{eq:RGR}
-\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
-where $A_\textrm{net} = y(\bar{p}(E) - r_{\rm l})$.
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees: namely plant height ($H$), stem-basal area ($A_{\rm st}$) or stem diameter $D$.
-The refined model makes these links clear.
-
-Unlike the seeding model, the new model also predicts a relationship between $\phi$ and growth rate that changes with size.
+Unlike the original model for seedlings (eqn \ref{eq:RGR}), the new model also predicts a relationship between $\phi$ and growth rate that changes with size.
 In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig.
@@
 
+The list of functional traits that are known to differ among plant species is already long and ever-increasing \citep{Perez-2013}.
+While we have focussed on understanding the effects of five specific traits on growth rate, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
+The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass or carbon, and can therefore be embedded within the eqns in Fig.
+\ref{fig:conceptual}b.
+While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
+A well-developed  trade-off must include two opposing forces, that operate at some point in the organisms life cycle.
+It is not necessary that benefit and cost both enter into Fig.
+\ref{fig:conceptual}; for example the trait wood density is sometimes thought of a trade-off between the costs of tissue construction and the rate of stem mortality.
+In that case, the costs of lower wood density would not appear within the eqns in Fig.
+\ref{fig:conceptual}b, so lower wood density would always increase growth rate.
+
+Our framework also highlights what is needed for traits to impact on growth rate and shade-tolerance.
+While traits can influence many aspects of plant function, these  influences must operate via the pathways outlined in Fig.
+\ref{fig:conceptual} if the trait is going to impact on growth.
+For example, many studies have focused on traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
+These traits will inevitably influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}) by altering conductance of water to the leaf.
+The potential costs of larger vessel size might be higher rates of stem turnover, which would appear in the term $k_{\rm s}$ in eqn \ref{eq:dbdt}.
+The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in our framework via the parameter $\theta$ (Table \ref{tab:allometry}).
+The effect of both these traits on growth rate should be expected to vary with plant size.
+
+% The only circumstances whereby traits impact on growth but not via the pathways in Fig. \ref{fig:conceptual}, would be where growth rate was entirely disconnected from the plant's carbon budget.
+
 \subsection{Implications for trait-based approaches}
@@
 First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
-Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with little attention devoted to allocation and turnover \citep[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}.
+% Remko thinks the following sentence is over the top but Martin de K thought it was fine.
+Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
 Yet, for four of the five traits considered here, trait values do not influence net primary production.
@@
 Such shifts have been observed across a variety of species \citet{King-1999,Thomas-1999,Koch-2004}.
-The prediction arises from the fact that as plants grow larger, the value of cheap leaf construction diminishes, compared to the costs of high leaf turnover.
-It therefore pays for plants, but especially those with low $\phi$, to increase their $\phi$ as they grow larger.
+Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low $\phi$ -- to increase their $\phi$ as they grow larger.
 \citet{King-1999} made a similar prediction for a single species of \emph{Eucalyptus}, but here we can extend the idea across species.
@@
 \ref{fig:assumptions}d).
-Some models also differ from in that they directly link a plant's stem diameter to it's height, e.g.
+Some models also differ from in that they directly link a plant's stem diameter to its height, e.g.
 \citet{Yokozawa-1995, King-1999, Moorcroft-2001, Li-2014}, whereas we let this scaling arise as an emergent outcome of growth and sapwood turnover.
@@
 \citet{King-1999} also connected his model of growth for a single species to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
-\ref{fig:growth_light_height}).
-This flattening out arose for a similar reason: the value of cheap leaf construction decreased with size, as more of the plant became stem.
+\ref{fig:growth_light_height}), because the value of cheap leaf construction decreased with size.
 In our study however, there was additional cost of increased leaf turnover, that further penalised low $\phi$ strategies when large.
@@
 Perhaps the two most controversial elements of our approach concerns our assumptions about tissue replacement and reproductive allocation.
-Many vegetation models determine allocation based on net primary production(photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing surplus energy.
+Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing surplus biomass.
 This is because we assume tissues lost via turnover are replaced before energy is allocated to either new growth (i.e.
 growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant
-% TODO: ??refs XX??)
-.
-The second assumption we make is that when mature, plant's allocate a substantial fraction of their surplus energy to reproduction.
+This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant.
+The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus energy to reproduction.
 While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
@@
 Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}.
-In this framework, mass-based growth is expressed as a $\frac{3}{4}$ scaling power of total plant mass, and diameter growth is expressed as a $\frac{1}{3}$ power of stem diameter.
-An extension of the model makes the connection to traits more explicit, and also links the model to the seedling model for relative growth rate in eqn \ref{eq:RGR} \citep{Enquist-2007}.
 Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
@@
 However, at least for $\phi$ such effects in large trees have not been observed.
-Finally, the \textsc{mst} models are difficult to connect with growth rates in other terms such as height.
-Other predictions from the two frameworks  agree, but for fundamentally different reasons.
-Specifically, both frameworks predict that the growth rate in total plant mass will continue to increase with size, which matches empirical evidence \citep{Stephenson-2014}.
-In the \textsc{mst} framework, this occurs because all mass (including heartwood) contributes to mass production.
-In our framework, this occurs because the turnover of sapwood also contributes substantially to whole-plant mass growth.
-This latter pattern matches data for large trees, where it was found that while height growth declined, mass growth continued to increase, mostly because of increased mass production \citep{Sillett-2010}.
 
@@
 Overall we have shown how diverse phenomena related to plant growth can be understood via a model accounting for processes generating photosynthetic income and allocating this among different tissues.
-The need to consider effects of size, alongside trait-based differences among species, has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001,, Westoby-2002, Enquist-2007}.
+The need to consider effects of size, alongside trait-based differences among species, has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
 Here we have provided a framework for achieving this.
@@
   & Stem-diameter: Hump-shaped &  \citet{Herault-2011} \\
-  & Stem-area: Increasing & \citet{Sillett-2010} \\
+  & Stem-area: unknown & \\
   & \textsc{rgr} (all variables): decreasing & \citet{Rees-2010, Iida-2014}\\
@@
 
-  \\ \multicolumn{3}{l}{\bf{Shade tolerance, \textsc{wplcp}}(Fig. \ref{fig:wplcp})}\\
+  \\ \multicolumn{3}{l}{\bf{Shade tolerance, \textsc{wplcp}} (Fig. \ref{fig:wplcp})}\\
   & Decreasing with size & \citet{Givnish-1988, Lusk-2008}\\
@@
 \multicolumn{4}{l}{\textbf{Other parameters (constant through ontogeny)}} \\
-& $c$ & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Adjusts for growth respiration and carbon content\\
+& $\alpha_{\rm y}$ &  & Yield -- fraction of carbon fixed converted into mass\\
+& $\alpha_{\rm bio}$  & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Biomass per mol carbon \\
 & $\eta_c$ & & Crown-shape parameter\\
@@
 \includegraphics{main/lcp_schematic.pdf}
-\caption{\textbf{Conceptual framework linking shade-tolerance to plant size and traits.}
+\caption{\textbf{Conceptual framework linking shade-tolerance to plant size and traits,}
+adapted from \citet{Givnish-1988}.
 Shade tolerance is the level of canopy openness $E$ where photosynthetic income (dashed line) intersects with the sum of respiration and turnover costs over tissues (solid black lines).
@@
 Traits can impact on shade tolerance by, in the case of leaf nitrogen content, shifting the income line up or down, or in the case of leaf mass per unit leaf area or wood density, causing the cost components to increase or decrease.
-Adapted and expanded from \citet{Givnish-1988}.
 \label{fig:wplcp_idea}}
@@
 For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated.
+The white regions indicate trait ranges that are typically observed in real systems.
 Figs.
@@
 Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
+The white regions indicate trait ranges that are typically observed in real systems.
 \label{fig:wplcp}}

### 2016-10-13 R/figures.R
Experimenting with new figure

@@ -221,8 +221,11 @@ color_pallete1 <- function(n) {
   RColorBrewer::brewer.pal(n+3, "Blues")[-(1:3)]
 }
 
-color_pallete2 <- function() {
-  RColorBrewer::brewer.pal(n=9, "Blues")
+color_pallete2 <- function(n=9) {
+  if(n <= 9)
+    RColorBrewer::brewer.pal(n=n, "Blues")
+  else
+    colorRampPalette(brewer.pal(9,"Blues")[-(1)])(n)
 }
 
 color_pallete3 <- function() {
@@ -383,12 +386,102 @@ figure_dY_dt <- function(dat) {
   legend("topright", legend = E, lty=1, col=rev(cols), bty="n", cex=0.5)
 }
 
-
 figure_height_dt_data <- function() {
   ret <- figure_dY_dt_data(sizes =  c(0.5, 2, 10, 15, 20),
         vars = c("height", "height_dt")
         )
   ret[["label"]] <- function(x) sprintf("H=%sm",x)
+  ret
+}
+
+
+figure_dY_dt_grid <- function(dat) {
+
+  vars <- dat[["vars"]]
+  sizes <- dat[["sizes"]]
+  E <- dat[["E"]]
+  traits <- dat[["traits"]]
+
+  par(oma=c(4, 5, 0, 3), mar=c(1, 1, 1, 0.9))
+
+  layout(matrix(1:5, 1, 5, byrow = TRUE), c(1, 1, 1, 1, 0.3), c(1))
+
+  nz <- 100
+  cols <- rev(colorRampPalette(brewer.pal(9,"RdYlBu"))(nz))
+  zmax <- 3
+  levels <- seq_range(c(0, zmax), nz)
+
+  for (v in traits) {
+
+    out <- collapse_grid(dat$data[[v]], "class", v, "height_dt")
+    x <- log10(out[[v]])
+    y <- dat[["sizes"]]
+    z <- out[["height_dt"]]
+
+    # cap z data
+    z[z > zmax] <- zmax
+
+    # make plot boundaries
+    xlim <- range(x)
+    ylim <- range(y)
+    plot.new()
+    plot.window(xlim = xlim, ylim = ylim, "", xaxs = "i", yaxs = "i", las = 1)
+    box()
+    rect(xlim[1], ylim[1], xlim[2], ylim[2], col="grey")
+
+    # add contour
+    .filled.contour(x,y,z, levels, cols)
+
+    # Calculate maximum of x for each value of y
+    # index_max <- function(y) {
[…]

### 2016-10-13 R/utils.R

@@ -38,3 +38,23 @@ linear_rescale <- function(x, range, scale = range(x)) {
   p <- (x - scale[[1]])/(scale[[2]] - scale[[1]])
   range[[1]] + p * (range[[2]] - range[[1]])
 }
+
+# Does reverse of expand.grid function
+# Takes a dataframe with up to 2 or 3 cols and converts into x,y,z vectors
+# such that data[,c("x","y)] = expand.grid(x,y)
+collapse_grid <- function(data, yvar, xvar, zvar){
+  n <- length(data[[yvar]])
+
+  #Find dimensions of X and Y vectors
+  ncol <- match(FALSE, data[[yvar]][seq_len(n-1) +1] > data[[yvar]][seq_len(n-1)])
+  nrow <- length(data[[xvar]])/ncol
+
+  #Extract x and y vectors
+  out <- list()
+  out[[yvar]] <- data[[yvar]][1:ncol]
+  out[[xvar]] <- data[[xvar]][ncol*seq(0,(nrow-1))+1]
+
+  #Reshape Z data into matrix
+  out[[zvar]] <- matrix(data[[zvar]], ncol =ncol, nrow=nrow, byrow=TRUE)
+  out
+}

### 2016-10-17 ms/figure_conceptual.tex
Incorporate suggestions from R Duursma throughout article
Cannot have 'con' directory on windows

@@
 \begin{document}
+
+% Needed to set equation counter, following inclusion of figure 1
+\setcounter{equation}{2}
 
@@
 \rowtitle{Biomass production}
-\firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/net_mass_production_dt_net_mass_production_dt}{}
-\othercol{2}{\scriptsize $\quad c$}{output/con/net_mass_production_dt_yield}{}
-\othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/con/net_mass_production_dt_assimilation}{}
-\othercol{4}{\hspace{-1em}-$\Sigma$\scriptsize $M_{\rm i} r_{\rm i}$}{output/con/net_mass_production_dt_respiration}{\hspace{-0.5em}$\Bigg)$}
-\othercol{5}{\hspace{-0.75em}-$\Sigma$\scriptsize $M_{\rm i} k_{\rm i}$}{output/con/net_mass_production_dt_turnover}{}
-\rowlabel{eq:dbdt}
+\firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/net_mass_production_dt_net_mass_production_dt}{}
+\othercol{2}{\hspace{-1em} \scriptsize $\alpha_{\rm \tiny bio} \alpha_{\rm \tiny y}$}{output/conn/net_mass_production_dt_yield}{}
+\othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/conn/net_mass_production_dt_assimilation}{}
+\othercol{4}{\hspace{-1em}-$\Sigma$\scriptsize $M_{\rm i} r_{\rm i}$}{output/conn/net_mass_production_dt_respiration}{\hspace{-0.5em}$\Bigg)$}
+\othercol{5}{\hspace{-0.75em}-$\Sigma$\scriptsize $M_{\rm i} k_{\rm i}$}{output/conn/net_mass_production_dt_turnover}{}
+\rowlabel{eq:dbdt_2}
 
@@
 \rowtitle{Mass growth}
-\firstcol{1}{$\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$}{output/con/mass_total_dt_mass_total_dt}{}
-\othercol{2}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/con/mass_total_dt_fraction_allocation_growth}{$\times$}
-\othercol{3}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/mass_total_dt_net_mass_production_dt}{$+$}
-\othercol{4}{$\frac{{\rm d}M_{{\rm h}}}{{\rm d}t}$}{output/con/mass_total_dt_mass_heartwood_dt}{}
+\firstcol{1}{$\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$}{output/conn/mass_total_dt_mass_total_dt}{}
+\othercol{2}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/conn/mass_total_dt_fraction_allocation_growth}{$\times$}
+\othercol{3}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/mass_total_dt_net_mass_production_dt}{$+$}
+\othercol{4}{$\frac{{\rm d}M_{{\rm h}}}{{\rm d}t}$}{output/conn/mass_total_dt_mass_heartwood_dt}{}
 \rowlabel{eq:dMtdt}
@@
 \rowtitle{Leaf deployment per mass invested}
-\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/con/darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
-\othercol{2}{$\frac{{\rm d}M_{{\rm l}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
-\othercol{3}{$\frac{{\rm d}M_{{\rm r}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_root_darea_leaf}{+}
-\othercol{4}{$\frac{{\rm d}M_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
-\othercol{5}{$\frac{{\rm d}M_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{$\Bigg)^{-1}$}
+\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/conn/darea_leaf_dmass_live_darea_leaf_dmass_live}{$\Bigg($}
+\othercol{2}{$\frac{{\rm d}M_{{\rm l}}}{{\rm d}A_{{\rm l}}}$}{output/conn/darea_leaf_dmass_live_dmass_leaf_darea_leaf}{+}
+\othercol{3}{$\frac{{\rm d}M_{{\rm r}}}{{\rm d}A_{{\rm l}}}$}{output/conn/darea_leaf_dmass_live_dmass_root_darea_leaf}{+}
+\othercol{4}{$\frac{{\rm d}M_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/conn/darea_leaf_dmass_live_dmass_bark_darea_leaf}{+}
+\othercol{5}{$\frac{{\rm d}M_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/conn/darea_leaf_dmass_live_dmass_sapwood_darea_leaf}{$\Bigg)^{-1}$}
  \rowlabel{eq:daldmt}
@@
 \rowtitle{Leaf-area growth}
-\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
-\othercol{2}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/con/area_leaf_dt_darea_leaf_dmass_live}{$\times$}
-\othercol{3}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/con/area_leaf_dt_fraction_allocation_growth}{$\times$}
-\othercol{4}{$\frac{{\rm d}B}{{\rm d}t}$}{output/con/area_leaf_dt_net_mass_production_dt}{}
+\firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/conn/area_leaf_dt_area_leaf_dt}{}
+\othercol{2}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/conn/area_leaf_dt_darea_leaf_dmass_live}{$\times$}
+\othercol{3}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/conn/area_leaf_dt_fraction_allocation_growth}{$\times$}
+\othercol{4}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/area_leaf_dt_net_mass_production_dt}{}
 \rowlabel{eq:daldt}
@@
 \rowtitle{Height growth}
-\firstcol{1}{$\frac{{\rm d}H}{{\rm d}t}$}{output/con/height_dt_height_dt}{}
-\othercol{2}{$\frac{{\rm d}H}{{\rm d}A_{{\rm l}}}$}{output/con/height_dt_dheight_darea_leaf}{$\times$}
-\othercol{3}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_leaf_dt_area_leaf_dt}{}
+\firstcol{1}{$\frac{{\rm d}H}{{\rm d}t}$}{output/conn/height_dt_height_dt}{}
+\othercol{2}{$\frac{{\rm d}H}{{\rm d}A_{{\rm l}}}$}{output/conn/height_dt_dheight_darea_leaf}{$\times$}
+\othercol{3}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/conn/area_leaf_dt_area_leaf_dt}{}
 \rowlabel{eq:dhdt}
@@
 \rowtitle{Stem-area growth}
-\firstcol{1}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/con/area_stem_dt_area_stem_dt}{$\Bigg($}
-\othercol{2}{$\frac{{\rm d}A_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_bark_darea_leaf}{$+$}
-\othercol{3}{$\frac{{\rm d}A_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/con/area_stem_dt_darea_sapwood_darea_leaf}{\hspace{-0.3em}$\Bigg)$}
-\othercol{4}{\hspace{-0.4em}$\times\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/con/area_stem_dt_area_leaf_dt}{$+$}
-\othercol{5}{$\frac{{\rm d}A_{{\rm h}}}{{\rm d}t}$}{output/con/area_stem_dt_area_heartwood_dt}{}
+\firstcol{1}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/conn/area_stem_dt_area_stem_dt}{$\Bigg($}
+\othercol{2}{$\frac{{\rm d}A_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/conn/area_stem_dt_darea_bark_darea_leaf}{$+$}
+\othercol{3}{$\frac{{\rm d}A_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/conn/area_stem_dt_darea_sapwood_darea_leaf}{\hspace{-0.3em}$\Bigg)$}
+\othercol{4}{\hspace{-0.4em}$\times\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/conn/area_stem_dt_area_leaf_dt}{$+$}
+\othercol{5}{$\frac{{\rm d}A_{{\rm h}}}{{\rm d}t}$}{output/conn/area_stem_dt_area_heartwood_dt}{}
 \rowlabel{eq:dast}
@@
 \rowtitle{Stem-diameter growth}
-\firstcol{1}{$\frac{{\rm d}D}{{\rm d}t}$}{output/con/diameter_stem_dt_diameter_stem_dt}{}
-\othercol{2}{$\frac{{\rm d}D}{{\rm d}A_{{\rm st}}}$}{output/con/diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
-\othercol{3}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/con/diameter_stem_dt_area_stem_dt}{}
+\firstcol{1}{$\frac{{\rm d}D}{{\rm d}t}$}{output/conn/diameter_stem_dt_diameter_stem_dt}{}
+\othercol{2}{$\frac{{\rm d}D}{{\rm d}A_{{\rm st}}}$}{output/conn/diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
+\othercol{3}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/conn/diameter_stem_dt_area_stem_dt}{}
 \rowlabel{eq:dDdt}

### 2016-10-17 ms/ms.tex
Edits to introduction

@@
 
-In this article we outline how the growth of individual plants is influenced by various functional traits, also the plant's size and light environment.
-Functional traits capture core differences in the strategies plants use to generate and invest surplus energy \citep{Westoby-2002, Wright-2004, Chave-2009}.
-Although most woody plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in the rates at which resources are acquired and invested into different tissues.
+Functional traits capture core differences in the strategies plants use to generate and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
+Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients and water), species differ considerably in the rates at which resources are acquired and invested into different tissues.
 During the last two decades, trade-offs related to some prominent traits have been quantified, with values for traits such as leaf mass per unit leaf area, wood density, and seed size now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
 As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
+In this article we outline how the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.
 
 While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the trait leaf mass per unit leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly related to relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a mathematical model of growth rate (see below).
-Leaf mass per unit leaf area and its correlate leaf lifespan were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
+In seedlings, the leaf mass per unit leaf area (LMA) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
+LMA and leaf lifespan, itself closely correlated with LMA, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
 These early successes prompted researchers to search for similar relationships in large plants.
-However, the results showed that in saplings and trees, leaf mass per unit leaf area was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
+However, the results showed that in saplings and trees, LMA was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
 Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
-So traits do not correlate simply and consistently to growth rate.
+Clearly, traits do not correlate simply and consistently to growth rate.
 
@@
 A recent meta-analysis of 103 studies reporting \textgreater 500 correlations by \citet{Gibert-2016} provides the most compelling evidence.
-\citet{Gibert-2016} showed that the strength of the correlation between some traits (including leaf mass per unit leaf area and maximum height) and growth rate was modified by size, while for other traits (including as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
+\citet{Gibert-2016} showed that the strength of the correlation between some traits (including LMA and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
 Importantly, the direction of the correlations and their shifts with size supported predictions generated via a framework describing how traits influence growth.
 
-The challenge of interpreting diverse empirical results seeking to link traits to growth rate was all the harder because, until the paper by \citet{Gibert-2016}, we lacked any clear theoretical expectations on why the effect of traits may be moderated by size.
+Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear theoretical expectations on why the effect of traits may be moderated by size.
 Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
-A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to the trait leaf mass per unit leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-An extension of the model suggests a similar relationship should hold at larger sizes \citep{Enquist-2007}.
+A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to LMA \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+An extension of the model suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
 But as noted above, the prediction for large plants is not supported by empirical results.
@@
 One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Wright-2000,Enquist-2007}.
-By contrast, the physiology of traits such as leaf mass per unit leaf area and wood density suggests they have influence by altering the allocation of energy among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
+By contrast, the physiology of traits such as LMA and wood density suggests they have influence by altering the allocation of biomass among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
 A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants that can be easily harvested.
-For larger plants, growth is more-often measured as increment in either diameter or height.
-For example, research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
+For larger plants, growth is more often measured as increment in either diameter or height.
+For example, stem diameter growth has been measured for many decades across many thousands of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
 Theory considering only primary production therefore offers limited insight.
 
-In this paper we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and shade tolerance (Table \ref{tab:phenomena}); and thereby offer fresh insight into the way traits influence plant demography across the life-cycle.
-Broadly, the {\plant} model builds off several related approaches, focussing on production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
-Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
+Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle.
+Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
+Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
 As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth.
-Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of energy uptake and allocation, why it makes sense that the effect of some traits changes with size.
+Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size.
 We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
-Our view is that trait-based approaches -- which focus on differences among species -- should be integrated within a general model of plant growth, also able to capture patterns of growth through ontogeny.
+Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
 Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}.
-In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
-And all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
+In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}, and all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
 Shade tolerance also varies among species, correlates with traits \citep{Lusk-2008,Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Lusk-2008}.

### 2016-10-17 ms/ms.tex
More edits to text, up to and including wood density result

@@
 
-\subsection{Original model for mass-based relative growth rate}
+\subsection{Classic model for mass-based relative growth rate}
+
+% Could cite Blackman 1919 "The Compound Interest Law and Plant Growth" AnnBot - this approach is very old.
 
@@
 In fact, that model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows.
-For seedlings, which are young and mostly leaf, we first ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
+For seedlings, which are young and mostly consist of leaf biomass, we can as a first approximation ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
 Net production (from eqn \ref{eq:dbdt}) then becomes a linear function of leaf area, making relative growth rate in mass a linear function of $\phi$:
@@
 
-\subsection{Decomposing of growth rates into components}
+\subsection{Decomposition of growth rates into components}
 
@@
 
-To model growth in either plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) then requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
+To model growth in either plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
 Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}.
@@
 
-The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
+The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
 The remaining  $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
@@
 Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
-Note that we make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integrating of stem turnover (eqns \ref{eq:ha}- \ref{eq:htm}).
+Note that we make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}- \ref{eq:htm}).
 
@@
 
-Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance, sometimes referred to as the ``whole-plant-light-compensation point'' ({\wplcp}): the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013}.
+Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance: the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013} (the ``whole-plant-light-compensation point'', {\wplcp}).
 In general, average assimilation rate per leaf area $\bar{p}$ increases with light level or canopy openness, $E$.
@@
 \ref{fig:wplcp_idea}).
-Traits will influence the {\wplcp} if they effect either income or costs.
-Also, since the amount of stem support naturally increases with height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig.
+Traits will influence the {\wplcp} if they effect either carbon uptake or costs (respiration, turnover).
+Also, since the amount of stem support increases with plant height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig.
 \ref{fig:wplcp_idea}).
@@
 
-It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication and differentiation, and must therefore hold for any potential growth model where biomass allocation is important.
+It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication and differentiation, they hold for any potential growth model where biomass allocation is important.
 To make explicit predictions via this framework, then requires an explicit model of plant construction and function; i.e.
-we must put equations behind all there terms in Fig.
+we must put quantify all the terms in Fig.
 \ref{fig:conceptual}b.
@@
 We now consider how trait-based trade-offs enter into the above growth model.
-It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect never-ending directional trait values towards more beneficial values.
+It is essential that any trait includes both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values.
 For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}.
@@
 \begin{description}
-\item[Seed mass, $\omega$:] There is a direct energetic trade-off between a plant's fecundity and the size of it's seeds.
-Moreover, we assume larger seeds result in larger seedlings; so this the seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
+\item[Seed mass, $\omega$:] There is a direct energetic trade-off between a plant's fecundity and the size of its seeds.
+Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
 \item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction.
 We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{{\rm mat}}$ via a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
+% Is this a direct biophysical consequence of varying this trait, as outlined as a necessary condition in the above?
 \item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
@@
 For the most part, parameters used in the current analysis were sourced from \citet{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters effecting the relationships outlined in Table \ref{tab:allometry}a, which were fit using data described below, and ii) parameters describing the function for reproductive allocation.
+The only exceptions are: i) parameters affecting the relationships outlined in Table \ref{tab:allometry}a, which were estimated from data described below, and ii) parameters describing the function for reproductive allocation.
 By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
@@
 As plants grow, differences in initial mass will decrease in importance, relative to other factors influencing growth through the life-cycle.
-We therefore expect the correlations between seed size and growth rate observed for seedlings to disappear among larger plants.
+As a result, the correlations between seed size and growth rate observed for seedlings disappear among larger plants.
 
@@
 At smaller sizes, there is no differentiation among species, because most individuals are focusing on growth.
-At larger sizes, individuals of some species are allocating a majority of their surplus energy to reproduction, which reduces their growth rate relative to those species with greater $H_{{\rm mat}}$ (Fig.
+At larger sizes, individuals of some species are allocating a larger fraction of biomass to reproduction, which reduces their growth rate relative to those species with greater $H_{{\rm mat}}$ (Fig.
 \ref{fig:growth_light_height}).
@@
 
-\noindent{\bf Leaf nitrogen content per leaf area, $\nu$.} The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given for any given light environment $E$ and does not vary with height (Fig.
+\noindent{\bf Leaf nitrogen content per leaf area, $\nu$.} The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
 \ref{fig:growth_light_height}).
@@
 The invariance of the growth-trait relationship with respect to size arises as follows.
-The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rates.
+The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
 Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
@@
 \ref{fig:growth_light_height}).
-As a result, the value of $\phi$ that optimises plant growth increases with plant size, and the direction of correlation between height growth rate and $\phi$ shifts from negative to positive, as plant's get taller.
+As a result, the value of $\phi$ that optimises plant growth increases with plant size, and the direction of correlation between height growth rate and $\phi$ shifts from negative to positive, as plants increase in height.
 Decreasing $\phi$ has two impacts on height growth rate.
@@
 Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects.
-When plants are small the effect on leaf deployment rate is bigger and so decreasing $\phi$ increases growth rate.
-When plants are large, the influence of $\phi$ on leaf deployment rate is diminished, because the costs of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) are larger (Fig.
+When plants are small the effect on leaf deployment rate is larger and so decreasing $\phi$ increases growth rate.
+When plants are large, the influence of $\phi$ on leaf deployment rate is diminished, because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger (Fig.
 \ref{fig:conceptual}).
@@
 \noindent{\bf Wood density, $\rho$.}
-As for leaf, cheaper stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates (Fig.
+As for leaves, lower cost of stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates (Fig.
 \ref{fig:growth_light_height}).
-In contrast to $\phi$, the benefits of cheaper stem construction become more pronounced at larger sizes, as more of the plant becomes wood (Fig.
+In contrast to $\phi$, the benefits of cheaper stem construction become more pronounced at larger sizes, as an increasingly large fraction of the plant is wood (Fig.
 \ref{fig:conceptual}a).
 Thus, $\rho$ has only a weak effect on growth rates for small plants, but a strong effect for intermediate to large plants.
-As the plant becomes very tall (H \textgreater 15m), the benefits start to low $\rho$ finally begin to diminish.
+As the plant becomes very tall (H \textgreater 15m), the benefits of low $\rho$ finally begin to diminish.
 

### 2016-10-18 ms/ms.tex
Incorporate suggestions from Georges Kunstler
also addresses some issues raised by Remko

@@
 correlation between traits and growth rate and its dependence on size?
+% XX - Georges suggests the graident of blues in Fig 4-5 not disntinct enough
 
@@
 
-Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear theoretical expectations on why the effect of traits may be moderated by size.
+Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
 Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
@@
 
-One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Wright-2000,Enquist-2007}.
+One possible reason that theoretical predictions have either been lacking or not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
 By contrast, the physiology of traits such as LMA and wood density suggests they have influence by altering the allocation of biomass among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
@@
 Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth.
-Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size.
+The paper by \citet{Gibert-2016} used the framework of the {\plant} model to make a conceptual argument why the effect of traits might change with size.
+Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to how, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size.
 We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
@@
 In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}, and all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Shade tolerance also varies among species, correlates with traits \citep{Lusk-2008,Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Lusk-2008}.
+Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
 These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.
@@
 \ref{fig:conceptual}a).
-We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
+We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
 The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
@@
 While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment, so that we can better understand the intrinsic trait- and size-related effects.
+We also do not yet consider any traits influencing levels of self-shading.
 Also, many vegetation models use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
@@
 
-% Could cite Blackman 1919 "The Compound Interest Law and Plant Growth" AnnBot - this approach is very old.
-
-Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-In fact, that model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows.
+Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
+To aid comparison with this literature, we will first show how that model is derived from eqn \ref{eq:dbdt}.
+In fact, the seedling model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows.
 For seedlings, which are young and mostly consist of leaf biomass, we can as a first approximation ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
@@
 The {\plant} adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
-We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics \citep{Huxley-1932}.
+We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
 Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation).
@@
 Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
-\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction.
-We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{{\rm mat}}$ via a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-% Is this a direct biophysical consequence of varying this trait, as outlined as a necessary condition in the above?
+\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates an inevitable energetic trade-off between growth and reproduction, that operates at all times through the lifestyle.
+Mass invested in growth cannot be invested in reproduction.
+To describe how the fraction of mass allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$, changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $H_{{\rm mat}}$ is height at maturation, $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
+The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction.
+The trait $H_{{\rm mat}}$ then describes the size at which this shift occurs, with direct biophysical consequences for growth.
 \item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
@@
 
-\subsection{Changes in height growth rate with trait}
+\subsection{Changes in height growth rate with traits}
 
@@
 We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
-The first two traits considered modify behaviour primarily at the start and end of an individuals growth trajectory, and are therefore termed ``ontogenetic traits''.
-The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individuals growth trajectory.
+The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory, and are therefore termed ``ontogenetic traits''.
+The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individual's growth trajectory.
 Across the five different traits, we observed four relatively distinct types of response.
-These responses are summarised in Table \ref{tab:responses} and described further below.
+These responses are summarised in Table \ref{tab:responses} and described in more detail below.
 
@@
 
-The remaining three traits ($\nu$, $\phi$, \& $\rho$) moderate growth rates at a point along an individuals growth trajectory (Fig. \ref{fig:growth_light_height}).
+The remaining three traits ($\nu$, $\phi$, \& $\rho$) moderate growth rates at a point along an individual's growth trajectory (Fig. \ref{fig:growth_light_height}).
 
@@
 \ref{fig:growth_light_height} focus on height growth rate (eqn \ref{eq:dhdt}).
-Corresponding results for growth rates in stem-diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
+Corresponding results for growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
 \ref{fig:growth_light_dia}--Fig.
@@
 
-We find that for seed mass, leaf nitrogen, leaf mass per area and height at maturation, the patterns of growth rate in stem-diameter, stem-area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}).
+We find that for seed mass, leaf nitrogen, leaf mass per area and height at maturation, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}).
 The only trait where a slightly different response was observed was for wood density.
 Whereas the effect of wood density on height growth tended to diminish slightly at larger sizes (Fig.
-\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem-diameter, stem-area or above-ground mass.
+\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area or above-ground mass.
 Recall that sapwood lost via turnover is turns into heartwood.
@@
 Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$.
-The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm kl2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
+The parameters $\beta_{{\rm ks2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
 Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
@@
 At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset these costs.
-{\wplcp} also decreases with height because as size increase, the total amount of energy needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
+{\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
 In addition, {\wplcp} varies with $\nu$.
@@
 
-Using a model relating plant physiological function and energy allocation to five prominent traits, we have shown how traits impact on plant growth across the life cycle.
+Using a model relating plant physiological function and carbon allocation to five prominent traits, we have shown how traits impact on plant growth across the life cycle.
 This approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait leaf mass per unit leaf area \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
@@
 Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life (from slow to fast growth) \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
-Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by light environment.
+Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment.
 The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
@@
 The model presented here extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
-Importantly, the seedling model can be derived as a special case of our more-extended approach.
+Importantly, the seedling model can be derived as a special case of our extended approach.
 Unlike the original model for seedlings (eqn \ref{eq:RGR}), the new model also predicts a relationship between $\phi$ and growth rate that changes with size.
 In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig.
-\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem-diameter, stem-area or total mass.
+\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem diameter, stem area or total mass.
 This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size.
@@
 
-The list of functional traits that are known to differ among plant species is already long and ever-increasing \citep{Perez-2013}.
+The list of functional traits that are known to differ among plant species is long and ever-increasing \citep{Perez-2013}.
 While we have focussed on understanding the effects of five specific traits on growth rate, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
@@
 It is not necessary that benefit and cost both enter into Fig.
-\ref{fig:conceptual}; for example the trait wood density is sometimes thought of a trade-off between the costs of tissue construction and the rate of stem mortality.
+\ref{fig:conceptual}; for example wood density is sometimes viewed as a trade-off between the costs of tissue construction and the rate of stem mortality.
 In that case, the costs of lower wood density would not appear within the eqns in Fig.
@@
 
-Our framework also highlights what is needed for traits to impact on growth rate and shade-tolerance.
+Our framework also highlights what is needed for traits to impact on growth rate and shade tolerance.
 While traits can influence many aspects of plant function, these  influences must operate via the pathways outlined in Fig.
@@
 % Remko thinks the following sentence is over the top but Martin de K thought it was fine.
+% :) 
 Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
@@
 Third, our results suggest that even if traits define a potential growth trajectory, researchers seeking to link traits to growth rate must probe deeper into the data than simply looking for a linear relationship between traits and average growth, to recover the expected relationships.
-None of the predicted relationships between traits and growth is linear across the range of sizes and light environments sampled.
+None of the predicted relationships between traits and growth is linear across the range of sizes and light environments tested.
 As such, we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits \citep[e.g.][]{Poorter-2008,Paine-2015}.
@@
 As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citet{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}, and in particular those by \citet{Makela-1997} and \citet{Moorcroft-2001}.
-These models have several properties in common, including that they all have growth being energetically driven by the gross amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
+These models have several properties in common, including that they all have growth being driven by the gross amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
 Many models also make functional-balance assumptions, for example linking the cross-section of sapwood to leaf area \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-2005, Moorcroft-2001}.
-We note that an assumption of exact functional balance is non-critical for our results, what matters is that the amount of live biomass (i.e.
+We note that an assumption of exact functional balance is not critical for our results, what matters is that the amount of live biomass (i.e.
 excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
@@
 \citet{Yokozawa-1995, King-1999, Moorcroft-2001, Li-2014}, whereas we let this scaling arise as an emergent outcome of growth and sapwood turnover.
-Again, we suspect this difference will not effect the core of results presented.
+Again, we expect this difference will not affect the main results.
 
-A feature distinguishing model from most of the mentioned above is the explicit linking to trait-based trade-offs.
+A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs.
 While such a linkage was also made by \citet{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have mainly focussed on ecosystem-level outcomes rather than the growth of individual plants.
@@
 \citet{King-1999} also connected his model of growth for a single species to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
-\ref{fig:growth_light_height}), because the value of cheap leaf construction decreased with size.
-In our study however, there was additional cost of increased leaf turnover, that further penalised low $\phi$ strategies when large.
+\ref{fig:growth_light_height}), because the influence of cheap leaf construction decreased with size.
+In our study however, there was additional cost of increased leaf turnover, that further penalised low $\phi$ strategies when plants were large.
 
-Perhaps the two most controversial elements of our approach concerns our assumptions about tissue replacement and reproductive allocation.
+Perhaps the two most controversial elements of our approach concerns the assumptions about tissue replacement and reproductive allocation.
 Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing surplus biomass.
-This is because we assume tissues lost via turnover are replaced before energy is allocated to either new growth (i.e.
+This is because we assume tissues lost via turnover are replaced before carbon is allocated to either new growth (i.e.
 growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
 This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant.
-The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus energy to reproduction.
+The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus carbon to reproduction.
 While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
@@
 
-Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}.
+Another class of model dealing explicitly with size-related effects includes those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}.
 Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
@@
 However, at least for $\phi$ such effects in large trees have not been observed.
+% yes, I like this shortened version!
 
@@
 
-Overall we have shown how diverse phenomena related to plant growth can be understood via a model accounting for processes generating photosynthetic income and allocating this among different tissues.
-The need to consider effects of size, alongside trait-based differences among species, has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
+Overall we have shown how diverse phenomena related to plant growth can be understood with a model accounting for processes generating photosynthetic income and allocating this among different tissues.
+The need to consider effects of plant size, alongside trait-based differences among species, has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
 Here we have provided a framework for achieving this.
@@
 Select references are provided for each phenomena, with preference for meta-analyses, where available.
-The traits considered are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$), leaf mass per area ($\phi$)$^1$, and wood density ($\rho$).
+The traits considered are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$)$^1$, leaf mass per area ($\phi$)$^2$, and wood density ($\rho$).
 See Tables \ref{tab:definitions} and \ref{tab:traits} for further details.
@@
   \\ \multicolumn{3}{l}{\bf{Shade tolerance, \textsc{wplcp}} (Fig. \ref{fig:wplcp})}\\
-  & Decreasing with size & \citet{Givnish-1988, Lusk-2008}\\
-  & $\nu$: low values more shade tolerant &  \citet{Baltzer-2007}\\
-  & $\phi$: low values less shade tolerant & \citet{Poorter-2006, Baltzer-2007, Lusk-2008}\\
+  & Decreasing with size & \citet{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
+  & $\nu$: low values more shade tolerant &  \citet{Messier-1999}\\
+  & $\phi$: low values less shade tolerant & \citet{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}\\
   & \quad $\rightarrow$ stands have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
@@
   }
-$^1$ Or leaf lifespan
+$^1$ Or maximum photosynthetic rate per leaf area
+$^2$ Or leaf lifespan
 }
@@
 \includegraphics{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three ontogenetic traits on shade tolerance.}
+\caption{\textbf{Effect of three development traits on shade tolerance.}
 Panels show effect of traits on level of canopy openness that causes net production (eqn \ref{eq:dbdt}) to be zero.

### 2016-10-18 ms/ms.tex
final minor editorial edits

@@
 correlation between traits and growth rate and its dependence on size?
-% XX - Georges suggests the graident of blues in Fig 4-5 not disntinct enough
 
@@
 
-Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
+Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear theoretical expectations on why the effect of traits may be moderated by size.
 Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
@@
 
-One possible reason that theoretical predictions have either been lacking or not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
+One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Wright-2000,Enquist-2007}.
 By contrast, the physiology of traits such as LMA and wood density suggests they have influence by altering the allocation of biomass among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
@@
 Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-The paper by \citet{Gibert-2016} used the framework of the {\plant} model to make a conceptual argument why the effect of traits might change with size.
-Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to how, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size.
+As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth.
+Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size.
 We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
@@
 In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}, and all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
+Shade tolerance also varies among species, correlates with traits \citep{Lusk-2008,Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Lusk-2008}.
 These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.
@@
 \ref{fig:conceptual}a).
-We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
+We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
 The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
@@
 While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment, so that we can better understand the intrinsic trait- and size-related effects.
-We also do not yet consider any traits influencing levels of self-shading.
 Also, many vegetation models use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
@@
 
-Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
-To aid comparison with this literature, we will first show how that model is derived from eqn \ref{eq:dbdt}.
-In fact, the seedling model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows.
+% Could cite Blackman 1919 "The Compound Interest Law and Plant Growth" AnnBot - this approach is very old.
+
+Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+In fact, that model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows.
 For seedlings, which are young and mostly consist of leaf biomass, we can as a first approximation ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
@@
 The {\plant} adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
-We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
+We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics \citep{Huxley-1932}.
 Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation).
@@
 Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
-\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates an inevitable energetic trade-off between growth and reproduction, that operates at all times through the lifestyle.
-Mass invested in growth cannot be invested in reproduction.
-To describe how the fraction of mass allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$, changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $H_{{\rm mat}}$ is height at maturation, $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction.
-The trait $H_{{\rm mat}}$ then describes the size at which this shift occurs, with direct biophysical consequences for growth.
+\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction.
+We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{{\rm mat}}$ via a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
+% Is this a direct biophysical consequence of varying this trait, as outlined as a necessary condition in the above?
 \item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
@@
 Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$.
-The parameters $\beta_{{\rm ks2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
+The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm kl2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
 Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
@@
 Select references are provided for each phenomena, with preference for meta-analyses, where available.
-The traits considered are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$)$^1$, leaf mass per area ($\phi$)$^2$, and wood density ($\rho$).
+The traits considered are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$), leaf mass per area ($\phi$)$^1$, and wood density ($\rho$).
 See Tables \ref{tab:definitions} and \ref{tab:traits} for further details.
@@
   \\ \multicolumn{3}{l}{\bf{Shade tolerance, \textsc{wplcp}} (Fig. \ref{fig:wplcp})}\\
-  & Decreasing with size & \citet{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
-  & $\nu$: low values more shade tolerant &  \citet{Messier-1999}\\
-  & $\phi$: low values less shade tolerant & \citet{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}\\
+  & Decreasing with size & \citet{Givnish-1988, Lusk-2008}\\
+  & $\nu$: low values more shade tolerant &  \citet{Baltzer-2007}\\
+  & $\phi$: low values less shade tolerant & \citet{Poorter-2006, Baltzer-2007, Lusk-2008}\\
   & \quad $\rightarrow$ stands have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
@@
   }
-$^1$ Or maximum photosynthetic rate per leaf area
-$^2$ Or leaf lifespan
+$^1$ Or leaf lifespan
 }
@@
 \includegraphics{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three development traits on shade tolerance.}
+\caption{\textbf{Effect of three ontogenetic traits on shade tolerance.}
 Panels show effect of traits on level of canopy openness that causes net production (eqn \ref{eq:dbdt}) to be zero.

### 2016-10-19 ms/ms.tex
Polishing
Closes #46

--- 2016-10-18 ms/ms.tex
@@
 
-% XX - The main comment I have is whether the strength of the 'correlation' between traits and growth rates can be summarized in some metric, so that it can be directly shown.
-To me Fig 4 is not very clear on showing how size affects the trait-growth correlations.
-For example, for wood density, I see a relation for all size classes.
-For the largest trees, it may look like a weaker correlation but isn't that because growth rates are lower?
-The same with Narea, to me they all look the same (except again, larger trees grow slower but it’s hard to tell there for me whether the trait-growth correlation is any different)
-% XX -  The shade tolerance (WPLCP) analysis seems off-topic (it also not mentioned in the abstract).
-Can this be integrated more closely with the main objective, i.e.
-correlation between traits and growth rate and its dependence on size?
-% XX - Georges suggests the graident of blues in Fig 4-5 not disntinct enough
-
 Functional traits capture core differences in the strategies plants use to generate and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
@@
 
-Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle.
+Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle.
 Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
 Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-The paper by \citet{Gibert-2016} used the framework of the {\plant} model to make a conceptual argument why the effect of traits might change with size.
-Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to how, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size.
+Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below \citep[from][]{Falster-2011}, \citet{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
+Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of primary production and allocation, how and why the effect of some traits on growth rates changes with size.
 We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
 Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
-Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}.
-In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}, and all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
+Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Canham-2004, Canham-2006, Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}.
+In contrast, the growth rate of standing plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
+All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
 Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
@@
 \ref{fig:wplcp_idea}).
-Traits will influence the {\wplcp} if they effect either carbon uptake or costs (respiration, turnover).
+Traits can influence the {\wplcp} if they effect either carbon uptake or costs (respiration, turnover).
 Also, since the amount of stem support increases with plant height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig.
@@
 
-The {\plant} adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
+The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
 We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
@@
 
-% Importantly, we can remove these size dependent effects by altering ??? of the assumptions in Table \ref{tab:allometry} (??).
-% \footnote{What are key assumptions that induce all size-related effects?
-My guess is  1) changing sapwood area so that volume scales isometrically with height will do it, and 2) removing reproductive allocation.
-Add figure in SI?}
-
 \subsection{Changes in height growth rate with traits}
@@
 
-% To understand the response of growth to these traits the following derivation is helpful.
-
-% Consider the effect of traits on the components of height growth rate (eqn \ref{eq:dhdt}).
-To make analysis more tractable, assume we are dealing with a plant investing all its energy in growth, i.e.
-$\frac{{\rm d}M_{\rm a}}{{\rm d}B}=1$.
-Note also that none of  the traits considered here influence $\frac{{\rm d}H} {{\rm d}A_{\rm l}}$.
-Therefore,
-% \begin{equation} \label{eq:G2}
-% \frac{{\rm d}H}{{\rm d}t} \propto \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}  \times \frac{ {\rm d}B} {{\rm d}t}.
-% \end{equation}
-% The correlation of traits with growth at a given height will be given by the derivative of eqn \ref{eq:G2} with respect to some trait $x$.
-The derivative of $\frac{{\rm d}H}{{\rm d}t}$ with respect to some trait $x$ is then \footnote{Eqn \ref{eq:G2} is a product of the form $Y(x) = W(x) \, Z(x)$, whose derivative is $\frac{\partial{Y}}{\partial{x}} = \frac{\partial{W}(x)}{\partial{x}} \, Z(x) + W(x) \, \frac{\partial{Z}(x)}{\partial{x}}$.}
-% \begin{equation} \label{eq:dg1}
-% \frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x} \propto
-% \frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}{\partial x}
-%  \, \frac{ {\rm d}B} {{\rm d}t}
-% + \frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}
-% \, \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}.
-% \end{equation}
-% If this derivative is positive (negative), there will be a positive (negative) relationship between traits and growth at the height where it is evaluated.
-We can simplify eqn \ref{eq:dg1} further by substituting $\frac{\partial \left(\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}\right)}
-% {\partial x} = -{\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}}^2
-% \, \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-% }{\partial x}$ (from eqn \ref{eq:daldmt}). Noting also that $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}>0$ and $\frac{{\rm d}B} {{\rm d}t} > 0$, we find that
-% \begin{equation}\label{eq:dg2}
-% \frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x}  \propto \frac{
-% \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
-% {\frac{ {\rm d}B} {{\rm d}t}} - \frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
-% }{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
-% \end{equation}
-
-% Eqn \ref{eq:dg2} implies that there will be a negative relationship between trait $x$ and height growth if the relative change in net mass production, due to a change in trait, is less than the relative change in the total construction cost per unit leaf area, due to a change in trait.
-We can also solve for trait values $x$ maximising growth rate by solving for the value of $x$ that gives $\frac{\partial \left(\frac{{\rm d}H}{{\rm d}t} \right)} {\partial x}= 0$.
-Crucially, key elements in eqn \ref{eq:dg2} will be modified by height.
-With these general results in hand, let us now consider the influence of three traits on height growth rate.
-
 \noindent{\bf Leaf nitrogen content per leaf area, $\nu$.} The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
@@
 Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
-%  $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x}$.
 From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial nu} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial nu}  - \frac{ \partial r_{\rm l}}{\partial nu} )$.
@@
 
-% Possible add some bits to suppmat?
-% Substituting $x=\phi$ into eqn \ref{eq:dg2}, we can come up with a more precise expressions for some of the terms.
-Note that we are evaluating these derivatives at a given value of $H$ and therefore $A_{\rm l}$ (via the allometry in eqn \ref{eq:ha}).
-Varying $\phi$ therefore changes $M_{\rm l}$, not $A_{\rm l}$.
-We can then reorganise eqn \ref{eq:dbdt} as,
-% \begin{equation}\label{eq:dbdt2}
-% \frac{{\rm d}B}{{\rm d}t} = c_1 - c_2  \, \phi^{1 -\beta_{{\rm kl2}}} \end{equation}
-% where
-% $c_1 = y \, A_{\rm l} \, (\bar{p}(E) - \nu \beta_{{\rm lf4}}) - \sum_{i={\rm b,s,r}} M_{i} \left(y\, r_{i} + k{i}\right)$
-% and
-% $c_2 =  A_{\rm l} \, \beta_{{\rm kl1}} \, \phi_0^{\beta_{{\rm kl2}}}$ are independent of $\phi$.
-% % The derivative of mass production with respect to $\phi$ is therefore
-% % \begin{equation}\label{eq:dbdt3}
-% % \frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
-% % (\beta_{{\rm kl2}}-1) \phi ^{-\beta_{{\rm kl2}}}.
-% % \end{equation}
-% Eqn \ref{eq:dbdt2} indicates that if $\beta_{{\rm kl2}}=1$, mass production rate $\frac{ {\rm d}B} {{\rm d}t}$ will be independent of $\phi$.
-In that case, lower $\phi$ will always give a growth advantage, irrespective of plant size.
-However,
-% evidence strongly suggests $\beta_{{\rm kl2}}>1$ \citep[see Fig.
-\ref{fS-leaf}, after][]{Wright-2004}.
-In that case, $\frac{ {\rm d}B} {{\rm d}t}$ will increase with $\phi$, because of lower losses via turnover.
-As $\phi$ increases, the relative increase in $\frac{ {\rm d}B} {{\rm d}t}$  eventually outweighs the relative change in $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$, causing the direction of correlation between growth rate and trait to shift from negative to positive.
-Moreover, the value of $\phi$ at which this shift occurs, which is also the value of $\phi$ optimising growth rate, is higher at greater sizes.
-
 \noindent{\bf Wood density, $\rho$.}
@@
 As the plant becomes very tall (H \textgreater 15m), the benefits of low $\rho$ finally begin to diminish.
-
-% Substituting in $x=\rho$, we can come up with a more precise expressions for some of the terms in eqn \ref{eq:dg2}.
-Note that we are evaluating these derivatives at a given value of $H$ and therefore $A_{\rm l}$ and thus the volume of bark and sapwood (via the allometry in eqns \ref{eq:ha}--\ref{eq:ab}).
-Varying $\rho$ therefore changes $M_{\rm s}$, not $A_{\rm st}$.
-We can then reorganise eqn \ref{eq:dbdt} as,
-% \begin{equation}\label{eq:dbdt3}
-% \frac{{\rm d}B}{{\rm d}t} = c_3 - c_4 \, \rho^{1-\beta_{{\rm ks2}}}
-% \end{equation}
-% where
-% $c_3 = y \, A_{\rm l} \, (\bar{p}(E) - \sum_{i={\rm l,b,s,r}} M_{i}\, r_{i})  - sum_{i={\rm l,r}} M_{i} k{i}$
-% and
-% $c_4 =  (1+b) \theta \,  \eta_c \, A_{\rm l} \, H \,  \beta_{{\rm ks1}} \, \rho_0^{\beta_{{\rm ks2}}}$ are independent of $\rho$.
-% Eqn \ref{eq:dbdt3} indicates that if $\beta_{{\rm ks2}}=1$, mass production rate $\frac{ {\rm d}B} {{\rm d}t}$ will be independent of $\rho$.
-In that case, lower $\rho$ will always give a growth advantage, irrespective of plant size.
-In this analysis we assume $\beta_{{\rm ks2}}=1.25$, so that mass production decreases slightly as $\rho$ decreases.
 
@@
 With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$.
-Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm kl2}} >1$.
-The parameters $\beta_{{\rm ks2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
-Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
+Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm ks2}} >1$.
+The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
+Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant because of disproportionate increases in turnover costs (Fig.
 \ref{fig:wplcp}).
@@
 Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life (from slow to fast growth) \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
-Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment.
+Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e.
+lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}.
 The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
@@
 
-% The only circumstances whereby traits impact on growth but not via the pathways in Fig. \ref{fig:conceptual}, would be where growth rate was entirely disconnected from the plant's carbon budget.
-
 \subsection{Implications for trait-based approaches}
@@
 First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
-% Remko thinks the following sentence is over the top but Martin de K thought it was fine.
-% :) 
 Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
@@
 This more-nuanced perspective tends to mirror observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
-% So we should instead be thinking of a species traits as defining a trajectory of potential growth for an individual, and one that is moderated by plant size and light environment.
 
@@
 Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
-First, the \textsc{mst}-derived model suggests diameter growth continues to increase as plants grow, whereas empirical data suggests growth rate declines for larger plants \citep{Herault-2011}.
+First, the \textsc{mst}-derived model suggests diameter growth continues to increase as plants grow, whereas empirical data suggests growth rate declines for larger plants \citep{Canham-2004, Canham-2006, Herault-2011}.
 Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size.
@@
 However, at least for $\phi$ such effects in large trees have not been observed.
-% yes, I like this shortened version!
 
@@
 
+\newpage
 \section{Tables and Figures}
-
-\newpage
 
@@
 {\footnotesize
-{\centering
-  \begin{tabular}{p{0.1cm}p{9cm}p{6cm}}
+{
+  \begin{tabular}{p{0.01cm}p{7.5cm}p{7cm}}
   \\
@@
   \hline
-  \\
   \multicolumn{3}{l}{\bf{Change in growth rate with increasing size} (Fig. \ref{fig:conceptual})}\\
@@
   & Height: Hump-shaped &  \citet{Ryan-2006, Sillett-2010, King-2011} \\
-  & Stem-diameter: Hump-shaped &  \citet{Herault-2011} \\
+  & Stem-diameter: Hump-shaped &  \citet{Canham-2004, Canham-2006, Herault-2011} \\
   & Stem-area: unknown & \\
@@
   & Decreasing with size & \citet{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
-  & $\nu$: low values more shade tolerant &  \citet{Messier-1999}\\
-  & $\phi$: low values less shade tolerant & \citet{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}\\
+  & $\nu$: low values more shade tolerant &  \citet{Messier-1999, Craine-2005,Baltzer-2007}\\
+  & $\phi$: low values less shade tolerant & \citet{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$\\
   & \quad $\rightarrow$ stands have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
   & $\rho$: low values less shade tolerant  &  \citet{Osunkoya-1996}\\
-  \\ \hline
-  \\
+  \hline
   \end{tabular}
   }
-$^1$ Or maximum photosynthetic rate per leaf area
-$^2$ Or leaf lifespan
+$^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area, or dark respiration rate per leaf area. Here these terms are directly related to $\nu$.\\
+$^2$ Similar responses are predicted for leaf lifespan. In our analysis this terms is directly related to $\phi$.\\
+$^3$ Note also that the article by \citet{Baltzer-2007} reports a very strong relationship between \textsc{wplcp} and respiration rate measured on either area or mass basis.\\
+
 }

### 2016-10-19 R/figures.R

@@ -218,7 +218,7 @@ default_strategy <- function() {
 }
 
 color_pallete1 <- function(n) {
-  RColorBrewer::brewer.pal(n+3, "Blues")[-(1:3)]
+  RColorBrewer::brewer.pal(n+1, "Blues")[-(1)]
 }
 
 color_pallete2 <- function(n=9) {
@@ -343,7 +343,7 @@ figure_dY_dt <- function(dat) {
   E <- dat[["E"]]
   traits <- dat[["traits"]]
 
-  cols <- rev(color_pallete1(length(E)))
+  cols <- color_pallete1(length(E))
   par(mfrow=c(length(sizes), length(traits)),
       oma=c(4, 5, 0, 1.5), mar=c(1, 1, 1, 0.9))
 
@@ -383,7 +383,7 @@ figure_dY_dt <- function(dat) {
   }
   mtext(name_pretty(vars[2]), line=3, side=2, cex=0.9 , outer=TRUE)
 
-  legend("topright", legend = E, lty=1, col=rev(cols), bty="n", cex=0.5)
+  legend("topright", legend = E, lty=1, col=rev(cols), bty="n", cex=0.9)
 }
 
 figure_height_dt_data <- function() {
@@ -603,7 +603,7 @@ figure_lcp_trait <- function(data) {
 
   mylabel <- function(txt) label(txt, -0.2, 1.2, xpd=NA, cex=1.5)
 
-  cols <- color_pallete1(length(data[["heights"]]))
+  cols <- rev(color_pallete1(length(data[["heights"]])))
 
   par(oma=c(0, 2, 3, 0), mar=c(5, 3, 1, 1), mfrow=c(1,3))
   lab = list(narea="a", lma= "b", rho = "c")

### 2016-11-14 ms/ms.tex
Incorporate feedback from Gibert

@@
 Based on biomass production and allocation, this framework explains why the influence of prominent traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
-Considering  growth rate in height, we find the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes but is strongest at intermediate sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
+Considering  growth rate in height, we demonstrate why the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes but is strongest at intermediate sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
 Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results.
-By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.
+By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for understanding growth across diverse species around the world.
 
@@
 Functional traits capture core differences in the strategies plants use to generate and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
-Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients and water), species differ considerably in the rates at which resources are acquired and invested into different tissues.
+Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover.
 During the last two decades, trade-offs related to some prominent traits have been quantified, with values for traits such as leaf mass per unit leaf area, wood density, and seed size now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
 As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
-In this article we outline how the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.
+In this article we outline the mechanisms by which the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.
 
@@
 \citet{Gibert-2016} showed that the strength of the correlation between some traits (including LMA and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Importantly, the direction of the correlations and their shifts with size supported predictions generated via a framework describing how traits influence growth.
+Importantly, \citet{Gibert-2016} proposed putative mechanism to explain their results.
+In this paper, we test these mechanisms by providing a fully functioning model of plant growth.
 
@@
 For example, stem diameter growth has been measured for many decades across many thousands of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-Theory considering only primary production therefore offers limited insight.
+As we outline below, mass production is only one of the factors influencing diameter growth; theory considering only primary production therefore offers limited insight.
 
 Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle.
-Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}.
+Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
 Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
@@
 For the most part, parameters used in the current analysis were sourced from \citet{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters affecting the relationships outlined in Table \ref{tab:allometry}a, which were estimated from data described below, and ii) parameters describing the function for reproductive allocation.
+The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from data described below, and ii) parameters describing the function for reproductive allocation.
 By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
@@
 
-The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database (\textsc{baad}) \citep{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
+The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
 We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
@@
 In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
-Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}, and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
+Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
 All growth measures decrease sharply with size when expressed as relative growth rates (results not shown).
@@
 Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects.
-In this sense, our model is designed to both explained observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
+In this sense, our model is designed to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
 

### 2016-11-14 R/figures.R
Figures for talk

@@ -386,8 +386,8 @@ figure_dY_dt <- function(dat) {
   legend("topright", legend = E, lty=1, col=rev(cols), bty="n", cex=0.9)
 }
 
-figure_height_dt_data <- function() {
-  ret <- figure_dY_dt_data(sizes =  c(0.5, 2, 10, 15, 20),
+figure_height_dt_data <- function(sizes =  c(0.5, 2, 10, 15, 20)) {
+  ret <- figure_dY_dt_data(sizes =  sizes,
         vars = c("height", "height_dt")
         )
   ret[["label"]] <- function(x) sprintf("H=%sm",x)

### 2016-11-14 R/figures_data.R

@@ -132,3 +132,52 @@ figure_assumptions <- function(fit_a.lf_h.t,  fit_a.lf_sap, fit_a.lf_m.rf, fit_h
   lines(x, (1+s$a_b1)*s$theta*s$rho*eta_c*x, lwd=2)
 }
 
+
+
+figure_assumption_talk <- function(fit, show_data=TRUE){
+
+  xlim <- c(1E-5, 1E4)
+
+  mylegend <- function(txt) legend("bottomright", legend = txt, bty = "n", cex=1.5)
+  myaxislabel <- function(txt, side) mtext(txt, side, line =3.5, cex=1.5)
+
+  cols <- color_pallete2()
+  s <- default_strategy()
+  x <- seq_log_range(c(1E-6, 2E6), 100)
+
+  par(mar=c(5, 5.5, 0.5, 0.5), oma=c(0,0,0,0))
+
+  isoclines <- 1
+
+  if(fit$data$yvar == "h.t"){
+    isoclines <- 0.5
+    ylab <- "Height (m)"
+    y <- s$a_l1*x^s$a_l2
+  } else if(fit$data$yvar == "sap"){
+    ylab <-   expression(paste("Sapwood area (",m^2,")"))
+    y <- s$theta*x
+  } else if(fit$data$yvar == "m.rf"){
+    ylab <-  "Mass of fine roots (kg)"
+    y <- s$a_r1*x
+  }
+
+  if(show_data) {
+    figure_fit_by_group_stan(fit, isoclines=isoclines, A=NULL,
+      col.l = "darkgreen", col.p = "grey")
+  } else {
+    figure_fit_by_group_stan(fit, isoclines=isoclines, A=NULL,
+      col.l = NULL, col.p = NULL)
+  }
+
+
+  mylegend(sprintf("Slope =%0.1f\n\n",isoclines))
+
+  if(show_data)
+    mylegend(sprintf("%d individuals, %d species", fit$data$n_obs, fit$data$n_groups))
+
+  myaxislabel(ylab,2)
+  myaxislabel(expression(paste("Leaf area (",m^2,")")),1)
+
+  # Now plot expected relationship from model
+  lines(x, y, lwd=2)
+}

### 2016-11-14 R/stan.R

@@ -139,7 +139,8 @@ figure_fit_by_group <- function(data, isoclines=NULL, A=NULL,
   if(!is.null(isoclines))
     add_isoclines(A=A, b=isoclines)
 
-  points(df[,1], df[,2],  type= 'p', col=col, pch=16)
+  if(!is.null(col))
+    points(df[,1], df[,2],  type= 'p', col=col, pch=16)
 
   mtext(xlab, 1, line=3)
   mtext(ylab, 2, line=3)
@@ -189,10 +190,7 @@ figure_fit_by_group_stan <- function(fit_stan, isoclines=NULL,
 }
 
 add_isoclines <- function(A=NULL, b=NULL,...){
-  if(is.null(A)) {
-    A <- seq(-10,10)
-  }
-  if(is.null(b)) {
+  if(!is.null(A) & ! is.null(b)) {
     b <- 1
   }
   for(a in A) abline(a,b,lty="dotted",...)

### 2016-11-14 R/talk.R

@@ -1,0 +1,78 @@
+
+
+library(gridBase)
+library(grImport)
+
+get.pic <- function(file) {
+  base <- tools::file_path_sans_ext(file)
+  file.xml <- sprintf("%s.xml", base)
+  on.exit({
+    file.remove(file.xml)
+    file.remove(sprintf("capture%s", basename(file)))
+  })
+  PostScriptTrace(file, file.xml)
+  readPicture(file.xml)
+}
+
+# establishes a new viewport to plot onto at RHS of plotting window
+# bottom, right, top in inches
+setup_viewport <- function(bottom, right, top){
+  #gridBase  - works out coordinates of different plots
+  vps <- baseViewports()
+  pushViewport(vps$inner)  #put current plot onto global viewport stack
+  #make new viewport
+  vp <- viewport(x=unit(1, "npc") - unit(right, "inches"),
+                 y=unit(bottom, "inches"),
+                 width=unit(right, "inches"),
+                 height=unit(1, "npc") - unit(bottom, "inches") - unit(top, "inches"),
+                 just=c("left", "bottom"))
+  #push onto stack
+  pushViewport(vp)
+}
+
+figure_dY_dt_talk <- function(v, i, dat, tree) {
+
+  vars <- dat[["vars"]]
+  E <- dat[["E"]]
+  cols <- "forestgreen"
+  H <- dat[["sizes"]][[i]]
+  Hmax <- max(dat[["sizes"]])
+  n <- length(E)
+
+  extract_f <- function(v, i, dat) {
+      dat_v <- unname(split(dat$data[[v]], dat$data[[v]]$class))
+      long_to_wide(dat_v[[i]], "canopy_openness",
+                           c(v, vars[2]))
+  }
+
+  dsub <- extract_f(v, i, dat)
+
+  ymax <-  3
+  par(mai=c(1.2, 1.0, 0.6, 2))
+  plot(NA, type="n", log="x",
+          xaxt="n", yaxt="n", xlab="", ylab="",
+          ylim = c(0, ymax*1.05),
+          xlim = range(dsub[[1]]))
+
+  usr <- par("usr")
+  rect(1E-5, -1, 1E4, 2*ymax, col = make_transparent("grey", 0.3), border=NA)
+  obs <- trait_range_obs(v)
[…]

### 2016-11-15 ms/ms.tex
Format for PNAS

@@
-\documentclass[a4paper,11pt]{article}
-\usepackage[osf]{mathpazo}
-\usepackage{ms}
-\usepackage[]{natbib}
-\usepackage{floatpag}
-\floatpagestyle{empty}
+\documentclass[9pt,twocolumn,twoside]{pnas-new}
 
-% Enable new column type -- centered with fixed width
+% Enable new column type -- centered with fixed
 % Following http://texblog.org/2008/05/07/fwd-equal-cell-width-right-and-centre-aligned-content/
@@
 
-%\raggedright
+\usepackage{graphicx}
+\graphicspath{{../output/}}
 
-\usepackage{graphicx}
+\newcommand{\smurl}[1]{\url{#1}}
+\newcommand{\ud}{\ensuremath{\rm{d}}}
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\sma}{\textsc{sma}}
 
-% Enable cross referencing to external documents
+\templatetype{pnasresearcharticle} % Choose template
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
+
+% Enable cross referencing to Evidence document
 \usepackage{xr}
-\externaldocument{ms-supp}
+\externaldocument{supp}
 \externaldocument{figure_conceptual}
 
-% Input common definitions from external file
-\input{ms-common-defs}
-\date{}
+\title{Trajectories: how functional traits influence plant growth and shade tolerance across the life-cycle}
 
-\mstype{Research Article}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Daniel S. Falster}
+\author[b]{Remko A. Duursma}
+\author[a,c]{Richard G. FitzJohn}
 
-\keywords{growth rate, life-history, ontogeny, size, leaf mass per area, wood density, shade-tolerance, seed mass, height at maturation, leaf nitrogen}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\affil[a]{Department of Biological Sciences, Macquarie University NSW 2109, Australia}
+\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
+\affil[c]{Imperial College, London, United Kingdom}
 
-\begin{document}
-\mstitlepage
-\noindent
-\parindent=1.5em
-\addtolength{\parskip}{.3em}
-% \doublespacing
-% \linenumbers
+% Please give the surname of the lead author for the running footer
+\leadauthor{Falster}
 
-% \tableofcontents
-% Consider Ecol Letters "Ideas and Perspectives": maximum of 7500 words (main text), and 10 figures, tables or boxes.
+% Please add here a significance statement to explain the relevance of your work
+% 120 words
+\significancestatement{ Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
+The Significance Statement appears in the paper itself and is required for all research papers.
+Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
+The Significance Statement appears in the paper itself.
+}
 
-\section{Abstract}
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{D.F. designed the study; D.F and R.F performed the analysis. All authors contributed to the writing of this manuscript.}
 
+\authordeclaration{The authors declare that they have no competing financial interests.}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: adaptive.plant@gmail.com}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{growth rate $|$ traits $|$ trees $|$ wood density}
+
+% 250 words
+\begin{abstract}
 Plant species differ in many functional traits that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover.
@@
 By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for understanding growth across diverse species around the world.
+\end{abstract}
 
-\section{Introduction}
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
-Functional traits capture core differences in the strategies plants use to generate and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\begin{document}
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalized the article contents.
+\verticaladjustment{-2pt}
+
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
+
+\dropcap{F}unctional traits capture core differences in the strategies plants use to generate and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
 Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover.
@@
 Recently, it has become clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
-A recent meta-analysis of 103 studies reporting \textgreater 500 correlations by \citet{Gibert-2016} provides the most compelling evidence.
-\citet{Gibert-2016} showed that the strength of the correlation between some traits (including LMA and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Importantly, \citet{Gibert-2016} proposed putative mechanism to explain their results.
+A recent meta-analysis of 103 studies reporting \textgreater 500 correlations provides the most compelling evidence\citep{Gibert-2016}.
+That study showed that the strength of the correlation between some traits (including LMA and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
+Importantly, they proposed putative mechanisms to explain their results \citep{Gibert-2016}.
 In this paper, we test these mechanisms by providing a fully functioning model of plant growth.
 
-Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
+\begin{table}[!ht]
+\caption{Empirical phenomena explained via the framework presented in this paper.}
+\begin{tabular}{p{8cm}}
+\toprule
+\textbf{Change in growth rate with increasing size (Fig. \ref{fig:conceptual})}\\
+ Net biomass production: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
+ Plant mass: Increasing \citep{Sillett-2010, Stephenson-2014} \\
+ Height: Hump-shaped \citep{Ryan-2006, Sillett-2010, King-2011} \\
+ Stem-diameter: Hump-shaped \citep{Canham-2004, Canham-2006, Herault-2011} \\
+ \textsc{rgr} (all variables): decreasing \citep{Rees-2010, Iida-2014}\\
+\textbf{Effect of traits on growth rate (Fig. \ref{fig:growth_light_height})}\\
+ $\omega$: low values $\rightarrow$ smaller seedlings, lower absolute, higher relative \citep{Gibert-2016} \\
+ $H_{\rm m}$: low values $\rightarrow$ slower growth at at larger sizes \citep{Gibert-2016}\\
+ $\nu$: low values $\uparrow$ growth irrespective of size, but only in high light \citep{Gibert-2016}\\
+ $\phi$: low values $\uparrow$ growth when small, not at mid-large sizes \citep{Gibert-2016}\\
+ $\rho$: low values $\uparrow$ growth, except at largest sizes \citep{Gibert-2016}\\
+\textbf{Responsiveness of growth rate to changes in light, $E$ (Fig. \ref{fig:growth_light_height})}\\
+ $\nu$: high values respond more \\
+ $\phi$: low values respond more \\
+ $\rho$: low values respond more\citep{Ruger-2012}\\
+\textbf{Shade tolerance, \textsc{wplcp} (Fig. \ref{fig:wplcp})}\\
+ Decreasing with size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
+ $\nu$: low values $\uparrow$ shade tolerant \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
+ $\phi$: low values $\downarrow$ shade tolerant \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$, higher \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
+ $\rho$: low values less shade tolerant \citep{Osunkoya-1996}\\
+  \bottomrule
+  \end{tabular}
+\addtabletext{The traits are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$)$^1$, leaf mass per area ($\phi$)$^2$, and wood density ($\rho$).
+See Tables \ref{tab:definitions} and \ref{tab:traits} for further details.
+textsc{lai} = leaf area index.
+Footnotes: $^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area \& dark respiration rate per leaf area.
+Here these terms are directly related to $\nu$.
+$^2$ Similar responses are predicted for leaf lifespan.
+Here this terms is directly related to $\phi$.
+$^3$ \citep{Baltzer-2007} reports a strongs relationship between \textsc{wplcp} and respiration rate measured per unit leaf area and mass.}
+\label{tab:phenomena}
+\end{table}
+
+\begin{figure}[!hb]
+\centering
+\includegraphics[width=\linewidth]{main/figure_conceptual.pdf}
+\caption{Conceptual framework linking growth rate to plant size and traits.
+\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
+\textbf{b)} Equations describing the rates of biomass production and growth in various dimensions of the plant.
+In the first line the symbol $\sum$ means ``sum'', across tissues, where $i = {\rm l,b,sr}$.
+The grey numbers indicate equation numbers referred to in the main text.
+The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}.
+Colours highlight where the same metric appears repeatedly in different equations.
+For a full list of variable names see Table \ref{tab:definitions}.}
+\label{fig:conceptual}
+\end{figure}
+
+Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
 Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
@@
 
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
+\caption{\textbf{Conceptual framework linking shade-tolerance to plant size and traits,}
+adapted from \citep{Givnish-1988}.
+Shade tolerance is the level of canopy openness $E$ where photosynthetic income (dashed line) intersects with the sum of respiration and turnover costs over tissues (solid black lines).
+The black circles indicate the point of intersection for plants with different heights.
+All income and cost is expressed in units of dry mass produced  per unit leaf area per year.
+Note the costs of sapwood and bark increases with height.
+Traits can impact on shade tolerance by, in the case of leaf nitrogen content, shifting the income line up or down, or in the case of leaf mass per unit leaf area or wood density, causing the cost components to increase or decrease.
+\label{fig:wplcp_idea}}
+\end{figure}
+
+\begin{table}[!hb]
+ \caption{Variables of the \texttt{FF16} physiological model from {\plant} used in this analysis.}
+\centering
+\begin{tabular}{p{0.5cm}p{2cm}p{5cm}}
+\toprule
+Symbol & Unit & Description \\
+\midrule
+\multicolumn{3}{l}{\textbf{Traits (constant through ontogeny)}}\\
+$\omega$ & kg & Seed mass  \\
+$H_{{\rm mat}}$ & m & Height at maturation\\
+$\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf nitrogen per unit leaf area  \\
+$\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf mass per area \\
+$\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & Wood density \\
+\multicolumn{3}{l}{\textbf{State variables (may vary through ontogeny)}} \\
+$H$ & m & Height of a plant\\
+$B$ & kg & Biomass originating from parent plant\\
+$D$ & m & Stem diameter\\
+$M_i$ & kg & Mass of tissue type $i$ retained on plant \\
+$A_i$ & m$^2$ & Surface area or area of cross-section of tissue type $i$\\
+$p,\bar{p}$ & mol yr$^{-1}$ m$^{-2}$ & Photosynthetic rate per unit area \\
+$r_i$ & mol yr$^{-1}$ kg$^{-1}$  & Respiration rate per unit mass of tissue type $i$ \\
+$k_i$ & yr$^{-1}$ & Turnover rate for tissue type $i$ \\
+\multicolumn{3}{l}{\textbf{Environmental variables (fixed)}} \\
+$E$ & & Canopy openness\\
+\multicolumn{3}{l}{\textbf{Other parameters (constant throughout)}} \\
+$\alpha_{\rm y}$ &  & Yield -- fraction of carbon fixed converted into mass\\
+$\alpha_{\rm bio}$  & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Biomass per mol carbon \\
+$\eta_c$ & & Crown-shape parameter\\
+$\theta$ &  & Sapwood area per unit leaf area\\
+$\alpha_{\rm l1}$ & m & Height of plant with leaf area of 1m$^2$ \\
+$\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Root mass per unit leaf area \\
+$\alpha_{\rm b1}$ &  & Ratio of bark area to sapwood area\\
+\bottomrule
+\end{tabular}
+ \addtabletext{For mass ($M$), respiration ($r$), and turnover ($k$)
+variables, subscripts refer to any of the following tissues: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,
+\({\rm r}\) = roots, \({\rm a}\) = all living tissue.
+For area \(A\) variables, subscripts refer to any of the following: \({\rm l}\) = leaves, \({\rm st}\) = total stem cross-section, \({\rm s}\) = sapwood cross-section, \({\rm b}\) = bark cross-section, \({\rm h}\) = heartwood cross-section.}
+\label{tab:definitions}
+\end{table}
+
 Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle.
@@
 Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below \citep[from][]{Falster-2011}, \citet{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
-Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of primary production and allocation, how and why the effect of some traits on growth rates changes with size.
+Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below \citep[from][]{Falster-2011}, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
+Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of primary production and allocation, how and why the effect of some traits on growth rates changes with size.
 We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
@@
 
-\section{Framework for understanding the effects of traits on growth}
+\section*{Framework for understanding the effects of traits on growth}
 
@@
 \begin{equation}\label{eq:dbdt}
-\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
-= \underbrace{\strut \alpha_{\rm bio} }_\textrm{mass per C} \,
-\underbrace{\strut \alpha_{\rm y} }_\textrm{yield}
-\big( \underbrace{\strut A_{\rm l} \, \bar{p}(E)}_\textrm{photosynthesis} -
-\underbrace{\strut \, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}}_\textrm{respiration}\big)
-- \underbrace{\strut \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}}_\textrm{turnover}.
+\frac{{\rm d}B}{{\rm d}t}
+= \alpha_{\rm bio} \,
+\alpha_{\rm y}
+\big( A_{\rm l} \, \bar{p}(E) -
+\, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}\big)
+- \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}.
 \end{equation}
@@
 The derivations below makes these links clear.
+
+\begin{table}[t]
+ \caption{Key trade-offs (benefit and cost) for the five traits considered, as encoded into the {\plant} model.
+ }
+\centering
+  \begin{tabular}{p{0.5cm}p{3.0cm}p{2.5cm}p{1cm}}
+  \toprule
+  Trait &  Benefit & Cost & Ref. \\
+  \midrule
+  $\omega$ & $\uparrow$ size seedlings & $\downarrow$ fecundity & \citep{Moles-2006}\\
+  $H_{{\rm mat}}$  & $\uparrow$ growth & $\downarrow$ fecundity & \citep{Thomas-2011, Wenk-2015}\\
+  $\nu$ & $\uparrow$ photosynthesis (high light) & $\uparrow$ respiration rate & \citep{Wright-2004}\\
+  $\phi$ & $\downarrow$ leaf turnover & $\uparrow$ cost building leaf & \citep{Wright-2004}\\
+  $\rho$ & $\downarrow$ sapwood turnover & $\uparrow$ cost building stem & \citep{Chave-2009}\\
+  \bottomrule
+  \end{tabular}
+\label{tab:traits}
+\end{table}
+
+\begin{table*}
+\caption{Equations for a ``functional-balance'' model of plant construction. See Table \ref{tab:definitions} for a list of variable names and definitions.}
+\centering
+  \begin{tabular}{p{0.1cm}p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}
+  \toprule
+  & Variable & Function & Marginal cost & Growth rate & Eqn\\
+  \midrule
+  \multicolumn{6}{l}{\textbf{a) Functional-balance assumptions}} \\
+  & Height &
+    $H = \alpha_{\rm l1} \, A_{\rm l}^{0.5}$ &
+    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} = 0.5 \alpha_{\rm l1} A_{\rm l}^{-0.5}$ &
+    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
+  & Sapwood area &
+    $A_{\rm s} = \theta \, A_{\rm l}$ &
+    $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
+    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
+  & Bark area &
+    $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
+    $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
+    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
+  & Root mass &
+    $M_{\rm r} = \alpha_{\rm r1} \, A_{\rm l}$ &
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_{\rm r1}$  &
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
+  & Heartwood area & & &
+    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\
+  \multicolumn{6}{l}{\textbf{b) Derived quantities}} \\
+  & Leaf mass &
+    $M_{\rm l} = \phi \, A_{\rm l} $ &
+    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
+    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
+  & Sapwood mass &
+    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
+    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( H + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
+  & Bark mass &
+    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
+    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( H + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
+    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
+  & Heartwood area &
+    $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
+     &
+    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
+  & Heartwood mass &
+    $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
+     &
+    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:htm} \\
+  \bottomrule
+\end{tabular}
+\addtabletext{\\The first column of part \textbf{a} provides cores assumptions between various size metrics and leaf area.
+Eqns in the middle and right columns of \textbf{b}, and in \textbf{b} can then be derived from the assumptions in the left column of \textbf{a}.
+The column ``Eqn'' indicates equation numbers referred to in the main text.
+}
+\label{tab:allometry}
+\end{table*}
 
@@
 
-The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}.
+The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citep{Makela-1997} and \citep{Moorcroft-2001}.
 We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
@@
 In particular, we assume that as a plant grows:
-\begin{description}
-\item[Assumption 1:] Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha}).
-\item[Assumptions 2:] The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as}).
-\item[Assumptions 3:] The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab}).
-\item[Assumptions 4:] The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}).
-\item[Assumption 5:] The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
-\end{description}
+\textbf{1)} Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
+\textbf{2)} The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
+\textbf{3)} The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
+\textbf{4)} The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and \textbf{5)} The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
 Assumption 1 accounts for the architectural layout of the plant.
 Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
-To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).
+To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).
 
@@
 The trade-offs are then implemented as follows.
-\begin{description}
-\item[Seed mass, $\omega$:] There is a direct energetic trade-off between a plant's fecundity and the size of its seeds.
-Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
-\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates an inevitable energetic trade-off between growth and reproduction, that operates at all times through the lifestyle.
-Mass invested in growth cannot be invested in reproduction.
-To describe how the fraction of mass allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$, changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $H_{{\rm mat}}$ is height at maturation, $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction.
-The trait $H_{{\rm mat}}$ then describes the size at which this shift occurs, with direct biophysical consequences for growth.
-\item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
-Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
-\item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-Following \citet{Wright-2004}, the rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
-\item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
-The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly whenever $\rho$ is varied.
-There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is more speculative than the equivalent assumption for leaves, which is well supported empirically.
- \end{description}
 
-\section{Methods}
+ \section*{Results}
 
-The growth model described above has been implemented as the \texttt{FF16} physiological model within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
-The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for full details see Supplementary Materials -- A).
-
-For the most part, parameters used in the current analysis were sourced from \citet{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from data described below, and ii) parameters describing the function for reproductive allocation.
-By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
-\ref{fig:conceptual}).
-
-The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
-We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
-We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed to come from normal distributions.
-The means and variances of these distributions are then fit as part of the model-fitting procedure.
-
-The analyses presented employ best practises in scientific computing, as defined by \citet{Wilson-2014}, and are fully reproducible, via code available at
-\href{https://github.com/traitecoevo/growth\_trajectories}{github.com/traitecoevo/growth\_trajectories}.
-
-\section{Results}
+\begin{figure}[!hb]
+\centering
+\includegraphics[width=.95\linewidth]{main/assumptions.pdf}
+\caption{Assumptions of a functional balance model for plant construction.
+Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
+Blue lines indicate the show standardised major axis of the bivariate cloud for each species.
+Dotted lines indicate slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}.
+The dark line shows the relationship assumed in the {\plant} model and applied throughout the paper.
+}
+\label{fig:assumptions}
+\end{figure}
 
@@
 
-\noindent{\bf Seed size, $\omega$.}
+Seed size.
 Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease.
@@
 
-\noindent{\bf Height at maturation, $H_{{\rm mat}}$.}
+Height at maturation.
 $H_{{\rm mat}}$ moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
@@
 
-\noindent{\bf Leaf nitrogen content per leaf area, $\nu$.} The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
+Leaf nitrogen content per leaf area.
+The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
 \ref{fig:growth_light_height}).
@@
 
-\noindent{\bf Leaf mass per unit leaf area, $\phi$.}
+Leaf mass per unit leaf area.
 Unlike $\nu$, the response of growth rate to changes in $\phi$ varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
@@
 
-\noindent{\bf Wood density, $\rho$.}
+Wood density.
 As for leaves, lower cost of stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates (Fig.
@@
 
+\begin{SCfigure*}[\sidecaptionrelwidth][t]
+\centering
+\includegraphics[width=11.4cm,height=11.4cm, keepaspectratio]{main/growth_light_height.pdf}
+\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
+Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}.
+Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
+For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated.
+The white regions indicate trait ranges that are typically observed in real systems.
+Figs.
+\ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass.
+Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
+\label{fig:growth_light_height}}
+\end{SCfigure*}
 \subsection{Changes in other growth rates with trait}
@@
 
+\begin{SCfigure*}[\sidecaptionrelwidth][t]
+\centering
+\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
+\caption{\textbf{Effect of three development traits on shade tolerance.}
+Panels show effect of traits on level of canopy openness that causes net production (eqn \ref{eq:dbdt}) to be zero.
+Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
+The white regions indicate trait ranges that are typically observed in real systems.
+\label{fig:wplcp}}
+\end{SCfigure*}
+
 \subsection{Responsiveness of growth rate to light}
@@
 \bar{p}(E^*) =
-      \phi \left(\frac{k_{\rm l}}{y} + r_{\rm l}\right) +
+      \phi c_{rm l} +
       \left(\theta \, \rho \, \eta_c \, H\right)
-        \left(b\left(\frac{k_{\rm b}}{y} + r_{\rm b}\right)
-            + \left(\frac{k_{\rm s}}{y} + r_{\rm s}\right)\right) +
-      \alpha_{\rm r1} \left(\frac{k_{\rm r}}{y} + r_{\rm r}\right).
+        \left(b c_{\rm b}
+            + c_{\rm s}\right) +
+      \alpha_{\rm r1} c_{\rm r},
 \end{equation}
+where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
 Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with traits $\nu$, $\phi$, and $\rho$.
@@
 
-\section{Discussion}
+\newcommand{\sepp}{{\color{grey}/}}
+\newcommand{\upup}{$\uparrow\,\uparrow$}
+\newcommand{\updo}{$\uparrow\,\downarrow$}
+\newcommand{\dodo}{$\downarrow\,\downarrow$}
+\newcommand{\upfl}{$\uparrow$\,--}
+\newcommand{\flup}{--$\,\uparrow$}
+\newcommand{\dofl}{$\downarrow\,$--}
+\newcommand{\doup}{$\downarrow\,\uparrow$}
+\newcommand{\fldo}{--$\,\downarrow$}
+
+\begin{table}[t!]
+\centering
+\caption{Predicted effects of traits on key elements of plant function determining growth rate.}
+  \begin{tabular}{llx{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}p{0.01cm}}
+  \toprule
+  & &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
+  & & \boldmath$\omega$ & \boldmath$H_{{\rm mat}}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\
+  \midrule
+  \multicolumn{8}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
+  \multicolumn{2}{l} {Net biomass production, $\ud B / \ud t$} & & & & & & \\
+  & \tabitem Photosynthesis & - & - & $\uparrow$  & - & - & \\
+  & \tabitem Respiration & - & - & $\uparrow$  & - & - & \\
+  & \tabitem Turnover & - & - & - & $\uparrow$ & $\uparrow$ & \\
+  \multicolumn{2}{l} {Allocation to growth, $\ud M_{\rm a} / \ud B$} & - & $\uparrow$ & - & - & - & \\
+  \multicolumn{2}{l} {Leaf deployment, $\ud A_{\rm l} / \ud M_{\rm a}$} & & \\
+  & \tabitem Leaf  & - & - & - & $\downarrow$ & - & \\
+  & \tabitem Sapwood & - & - & - & - & $\downarrow$ & \\
+  & \tabitem Root & - & - & - & - & - & \\
+  \multicolumn{2}{l} {Architecture, $\ud H / \ud A_{\rm l}$} & - & - & - & - & - & \\
+\midrule
+ \multicolumn{8}{l} {\textbf{b) Predicted effect of trait on growths rate for a small and large plant}} \\
+  \multicolumn{8}{l} {Height} \\
+  & \tabitem{Absolute}, $\ud H / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
+  & \tabitem{Relative}, $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
+  \multicolumn{8}{l} {Stem area} \\
+  & \tabitem{Absolute}, $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
+  & \tabitem{Relative}, $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \dofl & \dodo & \\
+  \multicolumn{8}{l} {Stem diameter} \\
+  & \tabitem{Absolute}, $\ud D / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
+  & \tabitem{Relative}, $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
+  \multicolumn{8}{l} {Mass} \\
+  & \tabitem{Absolute}, $\ud M_{\rm t} / \ud t$ & \upfl &  \flup & \upup & \dofl & \dodo & \\
+  & \tabitem{Relative}, $\ud M_{\rm t} / (\ud t . M_{\rm t} )$ & \dofl &  \flup & \upup & \dofl & \dodo & \\
+  \bottomrule
+  \end{tabular}
+\addtabletext{Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
+See Table \ref{tab:traits} for trait definitions.
+Adapted and expanded from \citep{Gibert-2016}.}
+\label{tab:responses}
+\end{table}
+
+\section*{Discussion}
 
@@
 
-% - \citet{King-2005} shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
+% - \citep{King-2005} shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
 
@@
 To that end, we can make a further prediction arising from our results, which is that the trait $\phi$ should increase through ontogeny for all individuals, across all species.
-Such shifts have been observed across a variety of species \citet{King-1999,Thomas-1999,Koch-2004}.
+Such shifts have been observed across a variety of species \citep{King-1999,Thomas-1999,Koch-2004}.
 Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low $\phi$ -- to increase their $\phi$ as they grow larger.
-\citet{King-1999} made a similar prediction for a single species of \emph{Eucalyptus}, but here we can extend the idea across species.
+\citep{King-1999} made a similar prediction for a single species of \emph{Eucalyptus}, but here we can extend the idea across species.
 While trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
@@
 
-As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citet{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}, and in particular those by \citet{Makela-1997} and \citet{Moorcroft-2001}.
+As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}, and in particular those by \citep{Makela-1997} and \citep{Moorcroft-2001}.
 These models have several properties in common, including that they all have growth being driven by the gross amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
@@
 Some models also differ from in that they directly link a plant's stem diameter to its height, e.g.
-\citet{Yokozawa-1995, King-1999, Moorcroft-2001, Li-2014}, whereas we let this scaling arise as an emergent outcome of growth and sapwood turnover.
+\citep{Yokozawa-1995, King-1999, Moorcroft-2001, Li-2014}, whereas we let this scaling arise as an emergent outcome of growth and sapwood turnover.
 Again, we expect this difference will not affect the main results.
@@
 A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs.
-While such a linkage was also made by \citet{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have mainly focussed on ecosystem-level outcomes rather than the growth of individual plants.
+While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have mainly focussed on ecosystem-level outcomes rather than the growth of individual plants.
 Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
-\citet{King-1999} also connected his model of growth for a single species to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
+\citep{King-1999} also connected his model of growth for a single species to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
 \ref{fig:growth_light_height}), because the influence of cheap leaf construction decreased with size.
@@
 
-\section{Acknowledgements}
+\matmethods{
+The growth model described above has been implemented as the \texttt{FF16} physiological model within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
+The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
+To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for full details see Supplementary Materials -- A).
+
+The trade-offs are then implemented as follows.
+
+\textbf{Seed mass, $\omega$:} There is a direct energetic trade-off between a plant's fecundity and the size of its seeds.
+Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
+
+\textbf{Height at maturation, $H_{{\rm mat}}$:} This trait moderates an inevitable energetic trade-off between growth and reproduction, that operates at all times through the lifestyle.
+Mass invested in growth cannot be invested in reproduction.
+To describe how the fraction of mass allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$, changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $H_{{\rm mat}}$ is height at maturation, $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
+The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction.
+The trait $H_{{\rm mat}}$ then describes the size at which this shift occurs, with direct biophysical consequences for growth.
+
+\textbf{Nitrogen per unit leaf area,  $\nu$:} We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
+Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
+
+\textbf{Leaf mass per unit area, $\phi$:} This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
+Following \citep{Wright-2004}, the rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
+
+\textbf{Wood density, $\rho$:} This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
+The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly whenever $\rho$ is varied.
+There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is more speculative than the equivalent assumption for leaves, which is well supported empirically.
+
+For the most part, parameters used in the current analysis were sourced from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
+The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from data described below, and ii) parameters describing the function for reproductive allocation.
+By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
+\ref{fig:conceptual}).
+
+The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
+We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
+We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed to come from normal distributions.
+The means and variances of these distributions are then fit as part of the model-fitting procedure.
+
+The analyses presented employ best practises in scientific computing, as defined by \citep{Wilson-2014}, and are fully reproducible, via code available at
+\href{https://github.com/traitecoevo/growth\_trajectories}{github.com/traitecoevo/growth\_trajectories}.
+}
+
+\showmatmethods % Display the Materials and Methods section
+
+\acknow{
 
@@
 The authors have no conflicts of interest to declare.
+}
 
-\section{Supplementary Materials}
-
-\noindent A -- Additional details on the \texttt{FF16} growth model
-
-\noindent B -- Supplementary Tables
-
-\noindent C -- Supplementary Figures
-
-\vspace{10em}
-
-\newpage
-\section{Tables and Figures}
-
-\begin{table}[ht]
-\caption{\textbf{Key empirical phenomena explained via the framework presented in this paper.}
-Select references are provided for each phenomena, with preference for meta-analyses, where available.
-The traits considered are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$)$^1$, leaf mass per area ($\phi$)$^2$, and wood density ($\rho$).
-See Tables \ref{tab:definitions} and \ref{tab:traits} for further details.
-Abbreviations: \textsc{rgr}=relative growth ratel \textsc{lai} = leaf area index.
-}
-{\footnotesize
-{
-  \begin{tabular}{p{0.01cm}p{7.5cm}p{7cm}}
-  \\
-  \hline
-  \multicolumn{2}{l}{ Phenomena} & References \\
-  \hline
-  \multicolumn{3}{l}{\bf{Change in growth rate with increasing size} (Fig. \ref{fig:conceptual})}\\
-  & Net biomass production: Hump-shaped & \citet{Givnish-1988, Koch-2004} \\
-  & Plant mass: Increasing & \citet{Sillett-2010, Stephenson-2014} \\
-  & Height: Hump-shaped &  \citet{Ryan-2006, Sillett-2010, King-2011} \\
-  & Stem-diameter: Hump-shaped &  \citet{Canham-2004, Canham-2006, Herault-2011} \\
-  & Stem-area: unknown & \\
-  & \textsc{rgr} (all variables): decreasing & \citet{Rees-2010, Iida-2014}\\
-  \\
-  \multicolumn{3}{l}{\bf{Effect of traits on growth rate} (Fig. \ref{fig:growth_light_height})}\\
-  & $\omega$: low values produce smaller seedlings, resulting in lower absolute growth but higher \textsc{rgr}. & Meta-analysis by \citet{Gibert-2016} \\
-  & $H_{\rm m}$: low values slow growth, but only at larger sizes & Meta-analysis by \citet{Gibert-2016}\\
-  & $\nu$: low values increase growth irrespective of size, but only in high light & Meta-analysis by \citet{Gibert-2016}\\
-  & $\phi$: low values increase growth when small, not at mid-large sizes & Meta-analysis by \citet{Gibert-2016}\\
-  & $\rho$: low values increase growth, except at largest sizes & Meta-analysis by \citet{Gibert-2016}\\
-
-  \\ \multicolumn{3}{l}{\bf{Responsiveness of growth rate to changes in light, $E$} (Fig. \ref{fig:growth_light_height})}\\
-  & $\nu$: high values respond more to $E$, changes ranking of growth rate across species & \\
-  & $\phi$: low values respond more to $E$, does not change ranking of growth rate across species & \\
-  & $\rho$: low values respond more to $E$, does not change ranking of growth rate across species & \citet{Ruger-2012}\\
-
-  \\ \multicolumn{3}{l}{\bf{Shade tolerance, \textsc{wplcp}} (Fig. \ref{fig:wplcp})}\\
-  & Decreasing with size & \citet{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
-  & $\nu$: low values more shade tolerant &  \citet{Messier-1999, Craine-2005,Baltzer-2007}\\
-  & $\phi$: low values less shade tolerant & \citet{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$\\
-  & \quad $\rightarrow$ stands have higher \textsc{lai} &  \citet{Reich-1992, Gower-1993, Niinemets-2010} \\
-  & $\rho$: low values less shade tolerant  &  \citet{Osunkoya-1996}\\
-  \hline
-  \end{tabular}
-  }
-$^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area, or dark respiration rate per leaf area. Here these terms are directly related to $\nu$.\\
-$^2$ Similar responses are predicted for leaf lifespan. In our analysis this terms is directly related to $\phi$.\\
-$^3$ Note also that the article by \citet{Baltzer-2007} reports a very strong relationship between \textsc{wplcp} and respiration rate measured on either area or mass basis.\\
-
-}
-\label{tab:phenomena}
-\end{table}
-
-\newpage
-
-\begin{table}[ht]
- \caption{\textbf{Key variables of the \texttt{FF16} physiological model from {\plant}, and used in this analysis.}. For mass ($M$), respiration ($r$), and turnover ($k$)
-variables, subscripts refer to any of the following tissues: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,
-\({\rm r}\) = roots, \({\rm a}\) = all living tissue.
-For area \(A\) variables, subscripts refer to any of the following: \({\rm l}\) = leaves, \({\rm st}\) = total stem cross-section, \({\rm s}\) = sapwood cross-section, \({\rm b}\) = bark cross-section, \({\rm h}\) = heartwood cross-section.}
-\centering
-{\footnotesize 
-\begin{tabular}{p{0.1cm}p{2cm}p{2.5cm}p{8cm}}
-\\
-\hline
-& Symbol & Unit & Description \\
-\hline
-\multicolumn{4}{l}{\textbf{Environmental variables}} \\
-& $E$ & & Canopy openness\\
-\multicolumn{4}{l}{\textbf{State variables (may vary through ontogeny)}} \\
-& $H$ & m & Height of a plant\\
-& $B$ & kg & Biomass originating from parent plant\\
-& $D$ & m & Stem diameter\\
-& $M_i$ & kg & Mass of tissue type $i$ retained on plant \\
-& $A_i$ & m$^2$ & Surface area or area of cross-section of tissue type $i$\\
-& $p,\bar{p}$ & mol yr$^{-1}$ m$^{-2}$ & Photosynthetic rate per unit area \\
-& $r_i$ & mol yr$^{-1}$ kg$^{-1}$  & Respiration rate per unit mass of tissue type $i$ \\
-& $k_i$ & yr$^{-1}$ & Turnover rate for tissue type $i$ \\
-
-\multicolumn{4}{l}{\textbf{Traits (constant through ontogeny)}} \\
-& $\omega$ & kg & Seed mass  \\
-& $H_{{\rm mat}}$ & m & Height at maturation\\
-& $\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf nitrogen per unit leaf area  \\
-& $\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf mass per area \\
-& $\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & Wood density \\
-\multicolumn{4}{l}{\textbf{Other parameters (constant through ontogeny)}} \\
-& $\alpha_{\rm y}$ &  & Yield -- fraction of carbon fixed converted into mass\\
-& $\alpha_{\rm bio}$  & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Biomass per mol carbon \\
-& $\eta_c$ & & Crown-shape parameter\\
-& $\theta$ &  & Sapwood area per unit leaf area\\
-& $\alpha_{\rm l1}$ & m & Height of plant with leaf area of 1m$^2$ \\
-& $\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Root mass per unit leaf area \\
-& $\alpha_{\rm b1}$ &  & Ratio of bark area to sapwood area\\
-\hline
-\end{tabular}
-}
-\label{tab:definitions}
-\end{table}
-
-\newpage
-
-\begin{table}[ht]
- \caption{\textbf{Key trade-offs for the five traits considered here.}
-For each trait we list the benefit and cost of increased trait values, as encoded into the {\plant} model.
- }
-\centering
-{\footnotesize
-\raggedright
-  \begin{tabular}{p{3cm}p{3cm}p{3cm}p{4cm}}
-  \\
-  \hline
-  Name &  Benefit & Cost & References \\
-  \hline
-  \\
-  Seed mass, $\omega$ & $\uparrow$ size seedlings & $\downarrow$  maternal fecundity & \citet{Moles-2006}\\
-  \\
-  Height at maturation, $H_{{\rm mat}}$  & $\uparrow$ growth & $\downarrow$ fecundity & \citet{Thomas-2011, Wenk-2015}\\
-\\
-  Leaf nitrogen per unit leaf area, $\nu$ & $\uparrow$ photosynthesis in high light & $\uparrow$ respiration rate & \citet{Wright-2004}\\
-  \\
-  Leaf mass per unit leaf area, $\phi$ & $\downarrow$ leaf turnover & $\uparrow$ cost building leaf & \citet{Wright-2004}\\
-  \\
-  Wood density, $\rho$ & $\downarrow$ sapwood turnover & $\uparrow$ cost building stem & \citet{Chave-2009}\\
-  \\
-  \hline
-  \end{tabular}
-}
-\label{tab:traits}
-\end{table}
-
-\newpage
-
-\begin{table}[ht]
-\caption{ \textbf{Equations for a ``functional-balance'' model of plant construction.}
-The first column of part \textbf{a} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of \textbf{b}, and in \textbf{b} can then be derived from the assumptions in the left column of \textbf{a}.
-The column ``Eqn'' indicates equation numbers referred to in the main text.
-See Table \ref{tab:definitions} for a list of variable names and definitions.
-}
-\centering
-
-{\footnotesize
-  \begin{tabular}{p{0.1cm}p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \\ \hline
-  & Variable & Function & Marginal cost & Growth rate & Eqn\\ \hline
-  \multicolumn{6}{l}{\textbf{a) Functional-balance assumptions}} \\
-  & Height &
-    $H = \alpha_{\rm l1} \, A_{\rm l}^{0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} = 0.5 \alpha_{\rm l1} A_{\rm l}^{-0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
-  & Sapwood area &
-    $A_{\rm s} = \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
-  & Bark area &
-    $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
-  & Root mass &
-    $M_{\rm r} = \alpha_{\rm r1} \, A_{\rm l}$ &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_{\rm r1}$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
-  & Heartwood area & & &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\
-  \multicolumn{6}{l}{\textbf{b) Derived quantities}} \\
-  & Leaf mass &
-    $M_{\rm l} = \phi \, A_{\rm l} $ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
-  & Sapwood mass &
-    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( H + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
-  & Bark mass &
-    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( H + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
-  & Heartwood area &
-    $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
-  & Heartwood mass &
-    $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:htm} \\
-    \\
-  \hline
-\end{tabular}}
-\label{tab:allometry}
-\end{table}
-
-\newpage
-
-\newcommand{\sepp}{{\color{grey}/}}
-\newcommand{\upup}{$\uparrow\,\uparrow$}
-\newcommand{\updo}{$\uparrow\,\downarrow$}
-\newcommand{\dodo}{$\downarrow\,\downarrow$}
-\newcommand{\upfl}{$\uparrow$\,--}
-\newcommand{\flup}{--$\,\uparrow$}
-\newcommand{\dofl}{$\downarrow\,$--}
-\newcommand{\doup}{$\downarrow\,\uparrow$}
-\newcommand{\fldo}{--$\,\downarrow$}
-
-\begin{table}[h!]
-\centering
-\caption{\textbf{Predicted effects of traits on key elements of plant function determining growth rate.} Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
-Traits are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), nitrogen content per leaf area ($\nu$), leaf mass per unit leaf area ($\phi$), and wood density ($\rho$) (see Table \ref{tab:traits} for more details).
-Adapted and expanded from \citet{Gibert-2016}.}
-{\footnotesize
-\vspace{1cm}
-  \begin{tabular}{lllcx{0.8cm}x{0.8cm}x{0.8cm}x{0.8cm}x{0.8cm}p{0.1cm}}
-  \hline
-  & & &  &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
-  & & &  {\bf Symbol} & \boldmath$\omega$ & \boldmath$H_{{\rm mat}}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\ \hline
-  \\
-  \multicolumn{10}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
-  & \multicolumn{2}{l} {Net biomass production} & $\ud B / \ud t$ & & & & & & \\
-  & & \tabitem Photosynthesis & & - & - & $\uparrow$  & - & - & \\
-  & & \tabitem Respiration  & & - & - & $\uparrow$  & - & - & \\
-  & & \tabitem Turnover & & - & - & - & $\uparrow$ & $\uparrow$ & \\
-  & \multicolumn{2}{l} {Allocation to growth} & $\ud M_{\rm a} / \ud B$ & - & $\uparrow$ & - & - & - & \\
-  & \multicolumn{2}{l} {Leaf deployment per mass}  & $\ud A_{\rm l} / \ud M_{\rm a} $ & & \\
-    & & \tabitem Leaf  &  & - & - & - & $\downarrow$ & - & \\
-    & & \tabitem Sapwood & & - & - & - & - & $\downarrow$ & \\
-    & & \tabitem Root & & - & - & - & - & - & \\
-  & \multicolumn{2}{l} {Architecture} & $\ud H / \ud A_{\rm l}$ & - & - & - & - & - & \\
-  \\
-  \multicolumn{10}{l} {\textbf{b) Predicted effect of trait on growths rate for a small and large plant}} \\
-  & \multicolumn{9}{l} {Height} \\
-  & & \tabitem{Absolute} & $\ud H / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & & \tabitem{Relative} & $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  & \multicolumn{9}{l} {Stem area} \\
-  & & \tabitem{Absolute} & $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & & \tabitem{Relative} & $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  & \multicolumn{9}{l} {Stem diameter} \\
-  & & \tabitem{Absolute} & $\ud D / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & & \tabitem{Relative} & $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  & \multicolumn{9}{l} {Mass} \\
-  & & \tabitem{Absolute} & $\ud M_{\rm t} / \ud t$ & \upfl &  \flup & \upup & \dofl & \dodo & \\
-  & & \tabitem{Relative} & $\ud M_{\rm t} / (\ud t . M_{\rm t} )$ & \dofl &  \flup & \upup & \dofl & \dodo & \\
-  \\
-\hline
-  \end{tabular}
-  }
-\label{tab:responses}
-\end{table}
-
-\clearpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/figure_conceptual.pdf}
-\caption{\textbf{Conceptual framework linking growth rate to plant size and traits.}
-\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
-\textbf{b)} Equations describing the rates of biomass production and growth in various dimensions of the plant.
-In the first line the symbol $\sum$ means ``sum'', across tissues, where $i = {\rm l,b,sr}$.
-The grey numbers indicate equation numbers referred to in the main text.
-The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}.
-Colours highlight where the same metric appears repeatedly in different equations.
-For a full list of variable names see Table \ref{tab:definitions}.
-\label{fig:conceptual}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/lcp_schematic.pdf}
-\caption{\textbf{Conceptual framework linking shade-tolerance to plant size and traits,}
-adapted from \citet{Givnish-1988}.
-Shade tolerance is the level of canopy openness $E$ where photosynthetic income (dashed line) intersects with the sum of respiration and turnover costs over tissues (solid black lines).
-The black circles indicate the point of intersection for plants with different heights.
-All income and cost is expressed in units of dry mass produced  per unit leaf area per year.
-Note the costs of sapwood and bark increases with height.
-Traits can impact on shade tolerance by, in the case of leaf nitrogen content, shifting the income line up or down, or in the case of leaf mass per unit leaf area or wood density, causing the cost components to increase or decrease.
-\label{fig:wplcp_idea}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/assumptions.pdf}
-\caption{\textbf{Key assumptions of a functional balance model for plant construction.}
-Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
-Blue lines indicate the show standardised major axis of the bivariate cloud for each species.
-Dotted lines indicate slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}.
-The dark line shows the relationship assumed in the {\plant} model and applied throughout the paper.
-\label{fig:assumptions}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/growth_light_height.pdf}
-\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
-Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}.
-Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
-For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated.
-The white regions indicate trait ranges that are typically observed in real systems.
-Figs.
-\ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass.
-Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
-\label{fig:growth_light_height}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three development traits on shade tolerance.}
-Panels show effect of traits on level of canopy openness that causes net production (eqn \ref{eq:dbdt}) to be zero.
-Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
-The white regions indicate trait ranges that are typically observed in real systems.
-\label{fig:wplcp}}
-\end{figure}
-
-\clearpage
-
-\footnotesize
-
+% Bibliography
 \bibliography{references}
-
+%\pnasbreak
 \end{document}

### 2016-11-15 ms/ms.tex
Tightening down to page limit

@@
 \mbox{}\refstepcounter{equation}%
-$(\theequation)$%
+$[\theequation]$%
 \label{#1}
@@
 \usepackage{xr}
-\externaldocument{supp}
+\externaldocument{ms-supp}
 \externaldocument{figure_conceptual}
@@
 % 120 words
-\significancestatement{ Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+% Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
 The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
@@
 The Significance Statement appears in the paper itself.
+\significancestatement{
+Plant species differ in many functional traits -- features of specific tissues and allocation of energy among them.
+While traits have been used in many correlative approaches to describe communities and demography, it has remained unclear how and why traits should influence whole-plant growth.
+Here we present a new theoretical framework for understanding the effect of traits on plant growth and shade-tolerance.
+This framework captures diverse patterns of growth in relation to size, and explains why the effect of traits on growth changes through ontogeny.
+By disentangling the effects of plant size, light environment and traits on growth rates, this study provides a theoretical foundation for understanding growth across diverse tree species around the world.
 }
@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{growth rate $|$ traits $|$ trees $|$ wood density}
+\keywords{growth rate $|$ traits $|$ trees $|$ allometry $|$ model}
 
@@
 Based on biomass production and allocation, this framework explains why the influence of prominent traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
-Considering  growth rate in height, we demonstrate why the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes but is strongest at intermediate sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
+Specifically, we demonstrate why for height growth the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
 Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results.
-By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for understanding growth across diverse species around the world.
 \end{abstract}
@@
 While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the leaf mass per unit leaf area (LMA) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
-LMA and leaf lifespan, itself closely correlated with LMA, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
+In seedlings, the leaf mass per unit leaf area ($\phi$) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
+$\phi$ and leaf lifespan, itself closely correlated with $\phi$, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
 These early successes prompted researchers to search for similar relationships in large plants.
-However, the results showed that in saplings and trees, LMA was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
+However, the results showed that in saplings and trees, $\phi$ was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
 Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
@@
 Recently, it has become clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
-A recent meta-analysis of 103 studies reporting \textgreater 500 correlations provides the most compelling evidence\citep{Gibert-2016}.
-That study showed that the strength of the correlation between some traits (including LMA and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
+A recent meta-analysis of 103 studies reporting \textgreater 500 correlations provides the most compelling evidence \citep{Gibert-2016}.
+That study showed that the strength of the correlation between some traits (including $\phi$ and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
 Importantly, they proposed putative mechanisms to explain their results \citep{Gibert-2016}.
@@
 \begin{table}[!ht]
-\caption{Empirical phenomena explained via the framework presented in this paper.}
+\caption{Empirical phenomena explained in this paper.}
 \begin{tabular}{p{8cm}}
@@
 \textbf{Change in growth rate with increasing size (Fig. \ref{fig:conceptual})}\\
- Net biomass production: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
- Plant mass: Increasing \citep{Sillett-2010, Stephenson-2014} \\
- Height: Hump-shaped \citep{Ryan-2006, Sillett-2010, King-2011} \\
- Stem-diameter: Hump-shaped \citep{Canham-2004, Canham-2006, Herault-2011} \\
- \textsc{rgr} (all variables): decreasing \citep{Rees-2010, Iida-2014}\\
+ $\,$ Net biomass production: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
+ $\,$ Plant mass: Increasing \citep{Sillett-2010, Stephenson-2014} \\
+ $\,$ Height: Hump-shaped \citep{Ryan-2006, Sillett-2010, King-2011} \\
+ $\,$ Stem-diameter: Hump-shaped \citep{Canham-2004, Canham-2006, Herault-2011} \\
+ $\,$ Relative growth rate (all variables): decreasing \citep{Rees-2010, Iida-2014}\\
 \textbf{Effect of traits on growth rate (Fig. \ref{fig:growth_light_height})}\\
- $\omega$: low values $\rightarrow$ smaller seedlings, lower absolute, higher relative \citep{Gibert-2016} \\
- $H_{\rm m}$: low values $\rightarrow$ slower growth at at larger sizes \citep{Gibert-2016}\\
- $\nu$: low values $\uparrow$ growth irrespective of size, but only in high light \citep{Gibert-2016}\\
- $\phi$: low values $\uparrow$ growth when small, not at mid-large sizes \citep{Gibert-2016}\\
- $\rho$: low values $\uparrow$ growth, except at largest sizes \citep{Gibert-2016}\\
+ $\,$ $\omega$: low values $\rightarrow$ smaller seedlings, lower absolute \& higher relative growth \citep{Gibert-2016} \\
+ $\,$ $H_{\rm m}$: low values $\rightarrow$ slower growth at at larger sizes \citep{Gibert-2016}\\
+ $\,$  $\nu$: low values $\uparrow$ growth irrespective of size, but only in high light \citep{Gibert-2016}\\
+ $\,$ $\phi$: low values $\uparrow$ growth when small, not at mid-large sizes \citep{Gibert-2016}\\
+ $\,$ $\rho$: low values $\uparrow$ growth, except at largest sizes \citep{Gibert-2016}\\
 \textbf{Responsiveness of growth rate to changes in light, $E$ (Fig. \ref{fig:growth_light_height})}\\
- $\nu$: high values respond more \\
- $\phi$: low values respond more \\
- $\rho$: low values respond more\citep{Ruger-2012}\\
+ $\,$ $\nu$: high values respond more \\
+ $\,$ $\phi$: low values respond more \\
+ $\,$ $\rho$: low values respond more \citep{Ruger-2012}\\
 \textbf{Shade tolerance, \textsc{wplcp} (Fig. \ref{fig:wplcp})}\\
- Decreasing with size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
- $\nu$: low values $\uparrow$ shade tolerant \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
- $\phi$: low values $\downarrow$ shade tolerant \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$, higher \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
- $\rho$: low values less shade tolerant \citep{Osunkoya-1996}\\
+ $\,$ Decreasing with size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
+ $\,$ $\nu$: low values $\uparrow$ shade tolerant \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
+ $\,$ $\phi$: low values $\downarrow$ shade tolerant \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$, higher \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
+ $\,$ $\rho$: low values less shade tolerant \citep{Osunkoya-1996}\\
   \bottomrule
 \end{tabular}
-\addtabletext{The traits are: seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$)$^1$, leaf mass per area ($\phi$)$^2$, and wood density ($\rho$).
-See Tables \ref{tab:definitions} and \ref{tab:traits} for further details.
-textsc{lai} = leaf area index.
-Footnotes: $^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area \& dark respiration rate per leaf area.
-Here these terms are directly related to $\nu$.
-$^2$ Similar responses are predicted for leaf lifespan.
-Here this terms is directly related to $\phi$.
-$^3$ \citep{Baltzer-2007} reports a strongs relationship between \textsc{wplcp} and respiration rate measured per unit leaf area and mass.}
+\addtabletext{Traits are seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$), leaf mass per unit leaf area ($\phi$), and wood density ($\rho$); for details see Tables \ref{tab:definitions}-\ref{tab:traits}.
+\textsc{lai} = leaf area index.
+Footnotes: $^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area \& dark respiration rate per leaf area; here both directly related to $\nu$.
+$^2$ Similar responses are predicted for leaf lifespan; here directly related to $\phi$.
+$^3$ \citep{Baltzer-2007} reports a strong relationship between \textsc{wplcp} and leaf respiration rate per unit leaf area or mass.}
 \label{tab:phenomena}
@@
 \textbf{b)} Equations describing the rates of biomass production and growth in various dimensions of the plant.
-In the first line the symbol $\sum$ means ``sum'', across tissues, where $i = {\rm l,b,sr}$.
-The grey numbers indicate equation numbers referred to in the main text.
+In the first line the symbol $\Sigma$ means ``sum'' across tissues, where $i = {\rm l,b,s,r}$.
+Grey numbers indicate equation numbers referred to in the main text.
 The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}.
 Colours highlight where the same metric appears repeatedly in different equations.
-For a full list of variable names see Table \ref{tab:definitions}.}
+For a list of variable names see Table \ref{tab:definitions}.}
 \label{fig:conceptual}
@@
 
-Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
+Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until recently \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
 Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
-A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to LMA \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to $\phi$ \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
 An extension of the model suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
@@
 
-One possible reason that theoretical predictions have either been lacking or not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
-By contrast, the physiology of traits such as LMA and wood density suggests they have influence by altering the allocation of biomass among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
-A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants that can be easily harvested.
-For larger plants, growth is more often measured as increment in either diameter or height.
-For example, stem diameter growth has been measured for many decades across many thousands of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-As we outline below, mass production is only one of the factors influencing diameter growth; theory considering only primary production therefore offers limited insight.
+One reason theoretical predictions have been lacking or not been supported is that in existing theory the effects of traits is realised mainly via influences on primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
+By contrast, the physiology of traits such as $\phi$ and wood density suggests they have influence by altering the allocation of biomass among different tissues rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
+A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants.
+On larger plants, growth is measured mainly as increment in either stem diameter or height \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
+As we outline below, mass production is only one of the factors influencing diameter growth; making theory considering only primary production of limited utility.
 
@@
 \includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
-\caption{\textbf{Conceptual framework linking shade-tolerance to plant size and traits,}
-adapted from \citep{Givnish-1988}.
+\caption{Conceptual framework linking shade-tolerance to plant size and traits, adapted from \citep{Givnish-1988}.
 Shade tolerance is the level of canopy openness $E$ where photosynthetic income (dashed line) intersects with the sum of respiration and turnover costs over tissues (solid black lines).
@@
 \begin{table}[!hb]
- \caption{Variables of the \texttt{FF16} physiological model from {\plant} used in this analysis.}
+ \caption{Variable definitions.}
 \centering
-\begin{tabular}{p{0.5cm}p{2cm}p{5cm}}
+\begin{tabular}{p{0.5cm}p{1.75cm}p{5cm}}
 \toprule
@@
 $M_i$ & kg & Mass of tissue type $i$ retained on plant \\
-$A_i$ & m$^2$ & Surface area or area of cross-section of tissue type $i$\\
+$A_i$ & m$^2$ & Surface or cross-section are of tissue type $i$\\
 $p,\bar{p}$ & mol yr$^{-1}$ m$^{-2}$ & Photosynthetic rate per unit area \\
@@
 \multicolumn{3}{l}{\textbf{Other parameters (constant throughout)}} \\
-$\alpha_{\rm y}$ &  & Yield -- fraction of carbon fixed converted into mass\\
+$\alpha_{\rm y}$ &  & Yield, fraction of carbon fixed into biomass\\
 $\alpha_{\rm bio}$  & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Biomass per mol carbon \\
@@
 \end{tabular}
- \addtabletext{For mass ($M$), respiration ($r$), and turnover ($k$)
-variables, subscripts refer to any of the following tissues: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,
-\({\rm r}\) = roots, \({\rm a}\) = all living tissue.
-For area \(A\) variables, subscripts refer to any of the following: \({\rm l}\) = leaves, \({\rm st}\) = total stem cross-section, \({\rm s}\) = sapwood cross-section, \({\rm b}\) = bark cross-section, \({\rm h}\) = heartwood cross-section.}
+\addtabletext{For $M_i$, $r_i$, $k_i$ and \(A_i\), subscripts refer to: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,  \({\rm h}\) = heartwood, \({\rm r}\) = roots, \({\rm st}\) = total stem, \({\rm a}\) =  alive tissue.}
 \label{tab:definitions}
@@
 Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle.
-Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
-Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
+The {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
+Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
 Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below \citep[from][]{Falster-2011}, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
@@
 Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
-These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.
+The diverse phenomena in Table \ref{tab:phenomena} deserve a comprehensive and integrated explanation.
 
@@
 
-We assume growth is fundamentally driven by biomass production and its subsequent distribution throughout the plant.
-Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000}:
+We assume growth is driven by biomass production and its subsequent distribution throughout the plant.
+Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover)\citep{Makela-1997, Thornley-2000}:
 \begin{equation}\label{eq:dbdt}
@@
 \end{equation}
-Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}(E)$) and total leaf area ($A_{\rm l}$).
+Photosynthesis is the product of the average photosynthetic rate per unit leaf area, $\bar{p}(E)$, and total leaf area, $A_{\rm l}$.
 We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (Fig.
-\ref{fig:wplcp_idea}; for details, see Supplementary Materials), and respiration and turnover rates of different tissues are constants that might differ with traits (see below).
+\ref{fig:wplcp_idea}; for details, see SI), and respiration and turnover rates of different tissues are constants that might differ with traits (see below).
 The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for growth respiration and the conversion of CO$_2$ into units of biomass, respectively.
-While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment, so that we can better understand the intrinsic trait- and size-related effects.
-We also do not yet consider any traits influencing levels of self-shading.
-Also, many vegetation models use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
+While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment so that we can better understand the intrinsic trait- and size-related effects.
+Many vegetation models also use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
 as functions of temperature.
-Such detail is not needed here because it will not change general model behaviour, even it would alter  he absolute values of predicted growth rates.
-Even with detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.
+Such detail is not needed here because it will not change general model behaviour, even it would alter the absolute values of predicted growth rates.
+Even in detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.
 
@@
 
-Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
-To aid comparison with this literature, we will first show how that model is derived from eqn \ref{eq:dbdt}.
-In fact, the seedling model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows.
-For seedlings, which are young and mostly consist of leaf biomass, we can as a first approximation ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}.
-Net production (from eqn \ref{eq:dbdt}) then becomes a linear function of leaf area, making relative growth rate in mass a linear function of $\phi$:
+Earlier studies focussing on seedlings related relative growth rate in mass to the trait $\phi$ \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
+This model can be derived from eqn \ref{eq:dbdt}, as a special case of the more-extended model described in the following sections, as follows.
+Ignoring all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}, net production becomes a linear function of leaf area, making relative growth rate in mass:
 \begin{equation}\label{eq:RGR}
@@
 where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$.
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees: namely plant height ($H$), stem-basal area ($A_{\rm st}$) or stem diameter $D$.
+Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
 The derivations below makes these links clear.
@@
 \centering
-  \begin{tabular}{p{0.5cm}p{3.0cm}p{2.5cm}p{1cm}}
+  \begin{tabular}{p{0.5cm}p{3.25cm}p{2.25cm}p{1cm}}
   \toprule
@@
 \end{table}
+
+\subsection{Decomposition of growth rates into components}
+
+% Update equation counter, following inclusion of figure 1
+\setcounter{equation}{9}
+
+To model growth in plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for net mass production, but also for the costs of building new tissues and allocation to reproduction.
+Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}.
+Equations are given in Fig.
+\ref{fig:conceptual}b, (eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
+As is evident in eqn  Fig.
+\ref{fig:conceptual}b, the growth rate in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) all share some terms in common.
+Many of the terms in Fig.
+\ref{fig:conceptual} also vary intrinsically with size.
+The insets in Fig.
+\ref{fig:conceptual}b show the size-related patterns in different variables for a typical plant, obtained from applying the specific allocation model in the next section.
+
+The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
+The remaining  $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
+In plants, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants \citep{Wenk-2015}.
+Note also that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
+So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ is not increasing.
+
+The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
+
+The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood and roots (eqn \ref{eq:daldmt}).
+The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of tissue-level construction costs per unit of leaf area, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$.
+
+The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
+This variable accounts for the architectural strategy of the plant.
+Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
+
+The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
+These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
+
+Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
+We make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}-\ref{eq:htm}).
 
@@
 \addtabletext{\\The first column of part \textbf{a} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of \textbf{b}, and in \textbf{b} can then be derived from the assumptions in the left column of \textbf{a}.
+Eqns in the middle and right columns of \textbf{a} and in \textbf{b} can then be derived from the assumptions in the left column of \textbf{a}.
 The column ``Eqn'' indicates equation numbers referred to in the main text.
@@
 
-\subsection{Decomposition of growth rates into components}
-
-% Update equation counter, following inclusion of figure 1
-\setcounter{equation}{9}
-
-To model growth in either plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout.
-Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}.
-Relevant equations are given in Fig.
-\ref{fig:conceptual}b, (eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
-As is evident in eqn  Fig.
-\ref{fig:conceptual}b, the growth rate in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) all share some terms in common.
-Many of the terms in Fig.
-\ref{fig:conceptual} also vary intrinsically with size.
-The insets in Fig.
-\ref{fig:conceptual}b show the size-related patterns in different variables for a typical plant, obtained from applying the specific allocation model in the next section.
-
-The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
-The remaining  $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
-In plants, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants \citep{Wenk-2015}.
-Note also that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
-So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ is not increasing.
-
-The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
-
-The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood and roots (eqn \ref{eq:daldmt}).
-The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of tissue-level construction costs per unit of leaf area, with one of these being $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$: the trait leaf mass per area.
-
-The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
-This variable accounts for the architectural strategy of the plant.
-Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
-
-The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
-These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
-
-Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
-Note that we make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}- \ref{eq:htm}).
-
 \subsection{Shade tolerance}
 
-Eqn \ref{eq:dbdt} can also be used to estimate a measure of shade tolerance: the light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration \citep{Givnish-1988, Baltzer-2007, Lusk-2013} (the ``whole-plant-light-compensation point'', {\wplcp}).
-In general, average assimilation rate per leaf area $\bar{p}$ increases with light level or canopy openness, $E$.
-The {\wplcp} can be estimated by solving for the the value of canopy openness $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$ (Fig.
+Eqn \ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'', \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
+In general, assimilation rate per leaf area $\bar{p}$ increases with light level, $E$.
+The {\wplcp} is light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration and can be estimated by solving for the the value $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$ (Fig.
 \ref{fig:wplcp_idea}).
@@
 We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
-Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation).
+Table \ref{tab:allometry} provides key equations (see \citep{Falster-2016} for full derivation).
 In particular, we assume that as a plant grows:
-\textbf{1)} Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
-\textbf{2)} The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
-\textbf{3)} The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
-\textbf{4)} The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and \textbf{5)} The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
+1) Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
+2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
+3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
+4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and
+5) The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
 Assumption 1 accounts for the architectural layout of the plant.
 Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
-To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).
+To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see SI).
 
@@
 
-We now consider how trait-based trade-offs enter into the above growth model.
+For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}.
 It is essential that any trait includes both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values.
-For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}.
 In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait.
-The trade-offs are then implemented as follows.
 
@@
 
+\subsection{Model assumptions}
+
+We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the \textsc{baad} \citep{Falster-2015b}.
+We also evaluated an additional prediction arising from the eqns in Table \ref{tab:allometry}a.
+The amount of live stem tissue supporting each unit of leaf area is predicted to increase linearly with height as
+\begin{equation}\label{eq:msml}
+\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
+\end{equation}
+Note that $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits and so assumed approximately constant for any given species.
+
 \begin{figure}[!hb]
 \centering
-\includegraphics[width=.95\linewidth]{main/assumptions.pdf}
+\includegraphics[width=1\linewidth]{main/assumptions.pdf}
 \caption{Assumptions of a functional balance model for plant construction.
 Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
-Blue lines indicate the show standardised major axis of the bivariate cloud for each species.
-Dotted lines indicate slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}.
-The dark line shows the relationship assumed in the {\plant} model and applied throughout the paper.
+Blue lines show standardised major axis lines fit to different species.
+The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table \ref{tab:allometry}.
 }
@@
 
-\subsection{Model assumptions}
-
-To verify model assumptions we compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the \textsc{baad} \citep{Falster-2015b}.
-Additionally, we evaluated an important prediction arising from the eqns in Table \ref{tab:allometry}a, that the amount of live stem tissue supporting each unit of leaf area increases linearly with height, as per the equation
-\begin{equation}\label{eq:msml}
-\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
-\end{equation}
-Note that $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits, i.e.
-properties that are assumed approximately constant through ontogeny for any given species.
 Fig.
 \ref{fig:assumptions} shows that the three functional-balance assumptions outlined in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
-The solid lines indicate {\sma} lines fit to each species in the dataset.
-Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a and eqn \ref{eq:msml}.
-As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions.
+The solid blue lines indicate {\sma} lines fit to each species in the dataset.
+As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions (Table \ref{tab:allometry}a and eqn \ref{eq:msml}).
 
-\subsection{Changes in growth rate with size}
-
-Our growth model suggests an intrinsically size-dependent pattern of biomass-production and growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
-The panels in Fig.
-\ref{fig:conceptual} show the expected patterns for a typical woody plant, obtained by applying the functional-balance model encoded in {\plant}.
-Biomass production shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
-Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
-This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size, including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig.
-\ref{fig:conceptual}), increasing reproductive allocation (Fig.
-\ref{fig:conceptual}), and declining mass production.
-In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
-Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
-All growth measures decrease sharply with size when expressed as relative growth rates (results not shown).
-
-\subsection{Changes in height growth rate with traits}
-
-We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}).
-We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
-The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory, and are therefore termed ``ontogenetic traits''.
-The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individual's growth trajectory.
-Across the five different traits, we observed four relatively distinct types of response.
-These responses are summarised in Table \ref{tab:responses} and described in more detail below.
-
-\subsubsection{Ontogenetic traits}
-
-Seed size.
-Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease.
-As such, the only effect of seed size on growth comes from changing the plant's initial  size.
-The plots in Fig.
-\ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
-Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates in all metrics  because of their greater total leaf area.
-At the same time, relative growth rate is predicted to decrease with size, because the ratio of leaf area to support mass decreases with plant size.
-As plants grow, differences in initial mass will decrease in importance, relative to other factors influencing growth through the life-cycle.
-As a result, the correlations between seed size and growth rate observed for seedlings disappear among larger plants.
-
-Height at maturation.
-$H_{{\rm mat}}$ moderates growth by adjusting the amount of energy invested in growth, i.e.~the term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
-Greater $H_{{\rm mat}}$ can thus lead to a growth advantage by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
-\ref{fig:growth_light_height}).
-At smaller sizes, there is no differentiation among species, because most individuals are focusing on growth.
-At larger sizes, individuals of some species are allocating a larger fraction of biomass to reproduction, which reduces their growth rate relative to those species with greater $H_{{\rm mat}}$ (Fig.
-\ref{fig:growth_light_height}).
-
-\subsubsection{Development traits}
-
-The remaining three traits ($\nu$, $\phi$, \& $\rho$) moderate growth rates at a point along an individual's growth trajectory (Fig. \ref{fig:growth_light_height}).
-
-Leaf nitrogen content per leaf area.
-The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
-\ref{fig:growth_light_height}).
-As $E$ increases from low to high, the optimal $\nu$ also increases.
-The invariance of the growth-trait relationship with respect to size arises as follows.
-The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
-Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
-From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial nu} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial nu}  - \frac{ \partial r_{\rm l}}{\partial nu} )$.
-Since both and $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of plant height.
-
-Leaf mass per unit leaf area.
-Unlike $\nu$, the response of growth rate to changes in $\phi$ varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
-\ref{fig:growth_light_height}).
-As a result, the value of $\phi$ that optimises plant growth increases with plant size, and the direction of correlation between height growth rate and $\phi$ shifts from negative to positive, as plants increase in height.
-Decreasing $\phi$ has two impacts on height growth rate.
-First, lower $\phi$ increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$ by economising on construction costs.
-Second, lower $\phi$ decreases net production $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
-Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects.
-When plants are small the effect on leaf deployment rate is larger and so decreasing $\phi$ increases growth rate.
-When plants are large, the influence of $\phi$ on leaf deployment rate is diminished, because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger (Fig.
-\ref{fig:conceptual}).
-The net result is that at larger sizes, low $\phi$ is no longer advantageous for growth (Fig.
-\ref{fig:growth_light_height}).
-
-Wood density.
-As for leaves, lower cost of stem construction (lower $\rho$) decreases the cost of deploying a unit of leaf area, and may thereby increase growth rates (Fig.
-\ref{fig:growth_light_height}).
-In contrast to $\phi$, the benefits of cheaper stem construction become more pronounced at larger sizes, as an increasingly large fraction of the plant is wood (Fig.
-\ref{fig:conceptual}a).
-Thus, $\rho$ has only a weak effect on growth rates for small plants, but a strong effect for intermediate to large plants.
-As the plant becomes very tall (H \textgreater 15m), the benefits of low $\rho$ finally begin to diminish.
-
-\begin{SCfigure*}[\sidecaptionrelwidth][t]
+\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
 \centering
@@
 \end{SCfigure*}
+
+\subsection{Changes in growth rate with size}
+
+Our model suggests an intrinsically size-dependent pattern of biomass-production and growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
+The panels in Fig.
+\ref{fig:conceptual} show the predicted patterns for a typical woody plant, obtained by applying the functional-balance model Table \ref{tab:allometry}a in {\plant}.
+Biomass production shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
+Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
+This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with increasing size, including a strong decline in the fraction of plant that is leaf (Fig.
+\ref{fig:conceptual}), increasing reproductive allocation (Fig.
+\ref{fig:conceptual}), and declining mass production.
+In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
+Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
+All growth measures decrease sharply with size when expressed as relative growth rates.
+
+\subsection{Changes in height growth rate with traits}
+
+We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}).
+We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
+The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
+The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individual's growth trajectory.
+Across the five different traits, we observed four relatively distinct types of response.
+These responses are summarised in Table \ref{tab:responses} and described in more detail below.
+
+\subsubsection{Ontogenetic traits}  Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease.
+As such, the only effect of $\omega$ on growth comes from changing the plant's initial size.
+The plots in Fig.
+\ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
+Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates in all metrics because of their greater total leaf area.
+At the same time, relative growth rate is predicted to decrease with size, because the ratio of leaf area to support mass decreases with plant size.
+As plants grow, differences in initial mass will decrease in importance, relative to other factors influencing growth through the life-cycle.
+As a result, the correlations between $\omega$ and growth rate observed for seedlings disappear among larger plants.
+
+Height at maturation, $H_{{\rm mat}}$, moderates growth by adjusting the amount of energy invested in growth, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
+Greater $H_{{\rm mat}}$ causes a growth advantage at larger sizes by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
+\ref{fig:growth_light_height}).
+At smaller sizes, there is no differentiation among species, as all plants focus on growth.
+
+\subsubsection{Development traits}  The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
+\ref{fig:growth_light_height}).
+As $E$ increases, the optimal $\nu$ also increases.
+The invariance of the growth-trait relationship with respect to size arises as follows.
+The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
+Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
+From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial nu} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial nu}  - \frac{ \partial r_{\rm l}}{\partial nu} )$.
+As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.
+
+Unlike $\nu$, the response of growth rate to changes in $\phi$ varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
+\ref{fig:growth_light_height}).
+As a result, the value of $\phi$ that optimises plant growth increases with height, and the direction of correlation between height growth rate and $\phi$ shifts from negative to positive, as plants increase in height.
+Decreasing $\phi$ has two impacts on height growth rate.
+First, lower $\phi$ increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$.
+Second, lower $\phi$ decreases net production $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
+Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects.
+When plants are small, the effect on leaf deployment rate is larger and so decreasing $\phi$ increases growth rate.
+When plants are large, the influence of $\phi$ on leaf deployment per mass is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
+At larger sizes, low $\phi$ is thus no longer advantageous for growth (Fig.
+\ref{fig:growth_light_height}).
+
+Reducing the cost of stem construction via lower $\rho$ decreases the cost of deploying a unit of leaf area and thereby increasing growth rate (Fig.
+\ref{fig:growth_light_height}).
+In contrast to $\phi$, the benefits of cheaper stem construction become more pronounced at intermediate sizes, as an increasingly large fraction of the plant is wood (Fig.
+\ref{fig:conceptual}a).
+
 \subsection{Changes in other growth rates with trait}
 
-The results reported above and shown in Fig.
-\ref{fig:growth_light_height} focus on height growth rate (eqn \ref{eq:dhdt}).
-Corresponding results for growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
-\ref{fig:growth_light_dia}--Fig.
-\ref{fig:growth_light_mass} of the Supplementary Materials.
-In each figure, plants were grown to a suitable diameter, area, or mass.
-As such, changes in relative growth rates with traits show a similar patterns as absolute growth rates.
+The results reported above and in Fig.
+\ref{fig:growth_light_height} focus on height growth rate.
+Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
+\ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}.
+For each, plants were grown to a suitable diameter, area, or mass.
+Changes in relative growth rates with traits therefore show a similar pattern.
 
-We find that for seed mass, leaf nitrogen, leaf mass per area and height at maturation, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}).
+We find that for $\omega$, $\nu$, $\phi$, and $H_{\rm mat}$, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}).
 The only trait where a slightly different response was observed was for wood density.
@@
 \ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area or above-ground mass.
-Recall that sapwood lost via turnover is turns into heartwood.
+Sapwood lost via turnover is turns into heartwood.
 Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.
 
-\begin{SCfigure*}[\sidecaptionrelwidth][t]
+\begin{SCfigure*}[\sidecaptionrelwidth][b]
 \centering
@@
 \bar{p}(E^*) =
-      \phi c_{rm l} +
+      \phi c_{\rm l} +
       \left(\theta \, \rho \, \eta_c \, H\right)
@@
 where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
-Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with traits $\nu$, $\phi$, and $\rho$.
+Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with $\nu$, $\phi$, and $\rho$.
 With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$.
 Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm ks2}} >1$.
-The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rate to $\phi$ and $\rho$, respectively.
-Since in this analysis, we have assumed these criteria to hold, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant because of disproportionate increases in turnover costs (Fig.
+The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to $\phi$ and $\rho$, respectively.
+Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
 \ref{fig:wplcp}).
-At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset these costs.
+At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover.
 {\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
@@
 At larger sizes, as net production declines to zero, {\wplcp} begins to increase again for very low values of $\nu$.
-All of these patterns match empirically observed patterns (Table \ref{tab:phenomena}).
+All these results  match empirical patterns (Table \ref{tab:phenomena}).
 
@@
 \centering
-\caption{Predicted effects of traits on key elements of plant function determining growth rate.}
+\caption{Predicted effects of traits on components of plant function determining growth rate. Adapted and expanded from \citep{Gibert-2016}.}
   \begin{tabular}{llx{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}p{0.01cm}}
@@
   & \tabitem Sapwood & - & - & - & - & $\downarrow$ & \\
-  & \tabitem Root & - & - & - & - & - & \\
-  \multicolumn{2}{l} {Architecture, $\ud H / \ud A_{\rm l}$} & - & - & - & - & - & \\
 \midrule
@@
   \end{tabular}
-\addtabletext{Arrows indicate the effect an increase in trait value would have on each element of the equations, with dashes indicating no effect.
-See Table \ref{tab:traits} for trait definitions.
-Adapted and expanded from \citep{Gibert-2016}.}
+\addtabletext{Arrows indicate the effect of increased trait value on each component; dashes indicating no effect. See Table \ref{tab:traits} for trait definitions.}
 \label{tab:responses}
@@
 The model presented here extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
-Importantly, the seedling model can be derived as a special case of our extended approach.
-Unlike the original model for seedlings (eqn \ref{eq:RGR}), the new model also predicts a relationship between $\phi$ and growth rate that changes with size.
+Importantly, the seedling model can be derived as a special case of the new-extended model (eqn \ref{eq:RGR}).
+Unlike the original, the extended model also predicts a relationship between $\phi$ and growth rate that changes with size.
 In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig.
@@
 \ref{fig:conceptual}a).
-Decreasing $\phi$ then has an overwhelmingly positive effect on height and diameter growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
-As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheaper leaf construction in eqn \ref{eq:daldmt}.
+Decreasing $\phi$ then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
+As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheap leaf construction.
 Consequently, the effect of $\phi$ on leaf turnover comes to dominate at larger sizes, and as such, the effect of $\phi$ on height, diameter and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
 
-The list of functional traits that are known to differ among plant species is long and ever-increasing \citep{Perez-2013}.
-While we have focussed on understanding the effects of five specific traits on growth rate, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
-The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass or carbon, and can therefore be embedded within the eqns in Fig.
-\ref{fig:conceptual}b.
+The list of functional traits known to differ among plant species is ever-increasing \citep{Perez-2013}.
+While we have focussed on understanding the effects of five traits on growth rate, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
+The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be embedded within the eqns \ref{eq:dbdt}--\ref{eq:dDdt}.
 While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
-A well-developed  trade-off must include two opposing forces, that operate at some point in the organisms life cycle.
-It is not necessary that benefit and cost both enter into Fig.
-\ref{fig:conceptual}; for example wood density is sometimes viewed as a trade-off between the costs of tissue construction and the rate of stem mortality.
-In that case, the costs of lower wood density would not appear within the eqns in Fig.
-\ref{fig:conceptual}b, so lower wood density would always increase growth rate.
+A well-developed trade-off must include two opposing forces that operate at some point in the life cycle.
 
@@
 The potential costs of larger vessel size might be higher rates of stem turnover, which would appear in the term $k_{\rm s}$ in eqn \ref{eq:dbdt}.
-The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in our framework via the parameter $\theta$ (Table \ref{tab:allometry}).
+The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in via the parameter $\theta$ (Table \ref{tab:allometry}).
 The effect of both these traits on growth rate should be expected to vary with plant size.
@@
 While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
-To that end, we can make a further prediction arising from our results, which is that the trait $\phi$ should increase through ontogeny for all individuals, across all species.
+To that end, we make a further prediction arising from our results, which is that the trait $\phi$ should increase through ontogeny for all individuals.
 Such shifts have been observed across a variety of species \citep{King-1999,Thomas-1999,Koch-2004}.
 Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low $\phi$ -- to increase their $\phi$ as they grow larger.
-\citep{King-1999} made a similar prediction for a single species of \emph{Eucalyptus}, but here we can extend the idea across species.
-While trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
+While a similar prediction was made for a species of \emph{Eucalyptus} \citep{King-1999}, we extend the idea across species.
+Although trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
 This hypothesis attempts to give meaning to some of that variation, and shows how variation across and within species might be understood within a single framework.
@@
 
-As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}, and in particular those by \citep{Makela-1997} and \citep{Moorcroft-2001}.
+As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}.
 These models have several properties in common, including that they all have growth being driven by the gross amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
@@
 \ref{fig:assumptions}d).
-Some models also differ from in that they directly link a plant's stem diameter to its height, e.g.
-\citep{Yokozawa-1995, King-1999, Moorcroft-2001, Li-2014}, whereas we let this scaling arise as an emergent outcome of growth and sapwood turnover.
-Again, we expect this difference will not affect the main results.
 
@@
 Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
-\citep{King-1999} also connected his model of growth for a single species to the trait $\phi$, and like the current study, predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
-\ref{fig:growth_light_height}), because the influence of cheap leaf construction decreased with size.
-In our study however, there was additional cost of increased leaf turnover, that further penalised low $\phi$ strategies when plants were large.
+Another study \citep{King-1999} connected a model of growth for a single species to the trait $\phi$, and likewise predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
+\ref{fig:growth_light_height}).
+Here we added an additional cost of increased leaf turnover, that further penalised low $\phi$ strategies when large.
 
@@
 For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
-In the case of growth, differences in reproductive allocation offer a clear pathway for explaining patterns linking a plants maximum size to the growth rate of large individuals \citep[e.g.][]{Wright-2010}.
+Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a plant's maximum size and growth rate in adult plants \citep[e.g.][]{Wright-2010}.
 
@@
 
-Overall we have shown how diverse phenomena related to plant growth can be understood with a model accounting for processes generating photosynthetic income and allocating this among different tissues.
-The need to consider effects of plant size, alongside trait-based differences among species, has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
+We have shown how diverse phenomena related to plant growth can be understood with a model accounting for processes generating photosynthetic income and allocating this among different tissues.
+The need to consider effects of plant size, alongside trait-based differences among species has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
 Here we have provided a framework for achieving this.
-By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.
+By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid  foundation for understanding and modelling growth across diverse species around the world.
 
 \matmethods{
-The growth model described above has been implemented as the \texttt{FF16} physiological model within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
-The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for full details see Supplementary Materials -- A).
+\subsection{Trait-based trade-offs}
+The trade-offs described in Table \ref{tab:traits} were embedded within the growth model as follows.
 
-The trade-offs are then implemented as follows.
+We assume a direct trade-off between a plant's fecundity and $\omega$, with larger seeds resulting in larger seedlings.
 
-\textbf{Seed mass, $\omega$:} There is a direct energetic trade-off between a plant's fecundity and the size of its seeds.
-Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
+$H_{{\rm mat}}$ moderates an inevitable energetic trade-off between growth and reproduction that operates at all times through the lifestyle.
+To describe how the fraction of mass allocated to reproduction changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
+The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction.
 
-\textbf{Height at maturation, $H_{{\rm mat}}$:} This trait moderates an inevitable energetic trade-off between growth and reproduction, that operates at all times through the lifestyle.
-Mass invested in growth cannot be invested in reproduction.
-To describe how the fraction of mass allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$, changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $H_{{\rm mat}}$ is height at maturation, $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction.
-The trait $H_{{\rm mat}}$ then describes the size at which this shift occurs, with direct biophysical consequences for growth.
-
-\textbf{Nitrogen per unit leaf area,  $\nu$:} We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
+We allow for the maximum photosynthetic capacity of the leaf to vary with $\nu$ as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\nu / \nu_0 \right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
 Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
 
-\textbf{Leaf mass per unit area, $\phi$:} This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-Following \citep{Wright-2004}, the rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
+$\phi$ influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
+The rate of leaf respiration per unit area is assumed independent of $\phi$\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
 
-\textbf{Wood density, $\rho$:} This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
+$\rho$ directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\rho / \rho_0\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
 The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly whenever $\rho$ is varied.
-There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is more speculative than the equivalent assumption for leaves, which is well supported empirically.
+There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is somewhat speculative.
 
-For the most part, parameters used in the current analysis were sourced from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from data described below, and ii) parameters describing the function for reproductive allocation.
-By default, we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H=H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
+\subsection{Parameters}
+Parameters were sourced mainly from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
+The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
 \ref{fig:conceptual}).
 
+\subsection{Assumptions}
 The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
 We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
-We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed to come from normal distributions.
-The means and variances of these distributions are then fit as part of the model-fitting procedure.
+We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed Gaussian.
 
-The analyses presented employ best practises in scientific computing, as defined by \citep{Wilson-2014}, and are fully reproducible, via code available at
+\subsection{Software}
+The growth model used has been implemented as the \texttt{FF16} physiological module within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
+The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
+To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
+The analyses presented employ best practises in scientific computing \citep{Wilson-2014} and are fully reproducible via code available at
 \href{https://github.com/traitecoevo/growth\_trajectories}{github.com/traitecoevo/growth\_trajectories}.
@@
 \acknow{
-
-We thank J Camac, A Gibert, G Kunstler, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
-Falster was supported by an Australian Research Council discovery grant (DP110102086).
-FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174).
+We thank J Camac, A Gibert, G Kunstler, H Muller-Landau, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
+DF was supported by an Australian Research Council discovery grant (DP110102086).
+RF was supported by the Science and Industry Endowment Fund (RP04-174).
 The authors have no conflicts of interest to declare.

### 2016-11-15 R/figures_data.R

@@ -94,7 +94,7 @@ figure_assumptions <- function(fit_a.lf_h.t,  fit_a.lf_sap, fit_a.lf_m.rf, fit_h
 
   par(mfrow = c(2,2), mar=c(5, 7, 3, 0.5), oma=c(0,0,0,0))
 
-  figure_fit_by_group_stan(fit_a.lf_h.t, isoclines=0.5, col.l = cols[8])
+  figure_fit_by_group_stan(fit_a.lf_h.t, col.l = cols[8])
   myaxislabel("Height (m)",2)
   myaxislabel(expression(paste("Leaf area (",m^2,")")),1)
   mylegend("Slope = 0.5")
@@ -102,8 +102,7 @@ figure_assumptions <- function(fit_a.lf_h.t,  fit_a.lf_sap, fit_a.lf_m.rf, fit_h
   # Now plot expected relationship from model
   lines(x, s$a_l1*x^s$a_l2, lwd=2)
 
-  figure_fit_by_group_stan(fit_a.lf_sap, isoclines=1,
-    A= seq(-15,10, by=2), col.l = cols[8])
+  figure_fit_by_group_stan(fit_a.lf_sap, col.l = cols[8])
   myaxislabel(expression(paste("Sapwood area (",m^2,")")),2)
   myaxislabel(expression(paste("Leaf area (",m^2,")")),1)
   mylegend("Slope = 1")
@@ -111,8 +110,7 @@ figure_assumptions <- function(fit_a.lf_h.t,  fit_a.lf_sap, fit_a.lf_m.rf, fit_h
   # Now plot expected relationship from model
   lines(x, s$theta*x, lwd=2)
 
-  figure_fit_by_group_stan(fit_a.lf_m.rf, isoclines=1,
-    A=seq(-10,10,by=2), col.l = cols[8])
+  figure_fit_by_group_stan(fit_a.lf_m.rf, col.l = cols[8])
   myaxislabel("Mass of fine roots (kg)",2)
   myaxislabel(expression(paste("Leaf area (",m^2,")")),1)
   mylegend("Slope = 1")
@@ -120,8 +118,7 @@ figure_assumptions <- function(fit_a.lf_h.t,  fit_a.lf_sap, fit_a.lf_m.rf, fit_h
   # Now plot expected relationship from model
   lines(x, s$a_r1*x, lwd=2)
 
-  figure_fit_by_group_stan(fit_h.t_msal, isoclines=1,
[…]

### 2016-11-15 ms/ms.tex
Minor edits text and fixed a reference

@@
 A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to $\phi$ \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-An extension of the model suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
+An extension of this equation suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
 But as noted above, the prediction for large plants is not supported by empirical results.
@@
 One reason theoretical predictions have been lacking or not been supported is that in existing theory the effects of traits is realised mainly via influences on primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
-By contrast, the physiology of traits such as $\phi$ and wood density suggests they have influence by altering the allocation of biomass among different tissues rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
+By contrast, the physiology of traits such as $\phi$ and wood density suggests they influence allocation of biomass among different tissues rather than mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
 A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants.
@@
 Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of primary production and allocation, how and why the effect of some traits on growth rates changes with size.
-We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
+We also show how our approach can account for related phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
 Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
@@
 as functions of temperature.
-Such detail is not needed here because it will not change general model behaviour, even it would alter the absolute values of predicted growth rates.
-Even in detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.
+Such detail is not needed here because it will not change general model behaviour, though it alters the absolute values of predicted growth rates.
 
@@
 
-Earlier studies focussing on seedlings related relative growth rate in mass to the trait $\phi$ \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
+Earlier studies on seedlings related relative growth rate in mass to $\phi$ \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
 This model can be derived from eqn \ref{eq:dbdt}, as a special case of the more-extended model described in the following sections, as follows.
@@
 Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
-The derivations below makes these links clear.
+The derivations below clarify these links.
 
@@
 As is evident in eqn  Fig.
-\ref{fig:conceptual}b, the growth rate in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) all share some terms in common.
+\ref{fig:conceptual}b, the growth rate in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) all share some terms.
 Many of the terms in Fig.
@@
 \ref{fig:wplcp_idea}).
-Traits can influence the {\wplcp} if they effect either carbon uptake or costs (respiration, turnover).
+Traits can influence the {\wplcp} if they affect either carbon uptake or costs (respiration, turnover).
 Also, since the amount of stem support increases with plant height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig.
@@
 
-It worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication and differentiation, they hold for any potential growth model where biomass allocation is important.
-To make explicit predictions via this framework, then requires an explicit model of plant construction and function; i.e.
-we must put quantify all the terms in Fig.
+It is worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication and differentiation, they hold for any potential growth model where biomass allocation is important.
+To make explicit predictions via this framework, an explicit model of plant construction and function is needed; i.e.
+we must quantify all the terms in Fig.
 \ref{fig:conceptual}b.
@@
 
-We find that for $\omega$, $\nu$, $\phi$, and $H_{\rm mat}$, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}).
+We find that for $\omega$, $\nu$, $\phi$, and $H_{\rm mat}$, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to height growth (Table \ref{tab:responses}).
 The only trait where a slightly different response was observed was for wood density.
@@
 \subsection{Implications for trait-based approaches}
-
 There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.
 
-First, our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
-Much of current ecosystem research focusses on factors effecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
+Our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
+Much of current ecosystem research focusses on factors affecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
 Yet, for four of the five traits considered here, trait values do not influence net primary production.
-In fact, the analysis with $\phi$ shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
-Low $\phi$ results in high leaf turnover, such that individuals with a $\phi$ have lower mass production.
+In fact, our analysis  shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
+Low $\phi$ results in high leaf turnover, such that individuals with a low $\phi$ have lower mass production.
 It is this property that makes them shade intolerant.
@@
 
-Second, our results demand a shift in the way plant species are sometimes described as, being of fast or slow growth.
-To the extent the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum from slow to fast growth.
+Second, our results demand a shift in the simple division of slow or fast growing plant species.
+To the extent that the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum from slow to fast growth.
 Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.

### 2017-08-09 ms/ms-common-defs.tex
Incorporate suggestions from R Duursma throughout article
minor edits, address change

@@
 
-\title{Growth trajectories: how functional traits influence plant growth}
+\title{Trajectories: how functional traits influence plant growth and shade tolerance across the life-cycle}
 \runninghead{H\lowercase{ow traits influence growth}}
 
-\author{Daniel S. Falster\textsuperscript{1,\textasteriskcentered}, Remko Duursma\textsuperscript{2}, \& Richard G. FitzJohn\textsuperscript{1,3}}
-\affiliation{\textsuperscript{1} Biological Sciences, Macquarie University NSW 2109, Australia \\
+\author{Daniel S. Falster\textsuperscript{1,\textasteriskcentered}, Remko A. Duursma\textsuperscript{2}, \& Richard G. FitzJohn\textsuperscript{3}}
+\affiliation{\textsuperscript{1} Evolution \& Ecology Research Centre, School of Biological, Earth and
+Environmental Sciences, University of New South Wales, Sydney, NSW 2052, Australia \\
 \textsuperscript{2} Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia \\

### 2017-08-09 ms/ms.tex
minor edits, address change

@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\affil[a]{Department of Biological Sciences, Macquarie University NSW 2109, Australia}
+\affil[a]{Evolution \& Ecology Research Centre, School of Biological, Earth and
+Environmental Sciences, University of New South Wales, Sydney, NSW 2052, Australia}
 \affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
@@
 
-\authordeclaration{The authors declare that they have no competing financial interests.}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: adaptive.plant@gmail.com}
+\authordeclaration{The authors declare that they have no competing financial interests.\\
+
+Code reproducing the results is available at \href{https://github.com/traitecoevo/growth_trajectories}{github.com/traitecoevo/growth\_trajectories}.
+}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: daniel.falster@unsw.edu.au}
 
@@
 Yet, it remains unclear how -- and even if -- such traits influence whole-plant growth, with the simple linear relationships predicted by existing theory often lacking empirical support.
-Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth and shade-tolerance, extending a widely-used theoretical model that links growth rate in seedlings with a single leaf trait to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
-Based on biomass production and allocation, this framework explains why the influence of prominent traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
+Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth and shade-tolerance by extending a widely-used model, linking growth rate in seedlings with a single leaf trait, to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
+Based on biomass production and allocation, this framework explains why the influence of traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
 Specifically, we demonstrate why for height growth the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
@@
 While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the leaf mass per unit leaf area ($\phi$) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
+In seedlings, the trait leaf mass per unit leaf area ($\phi$) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
 $\phi$ and leaf lifespan, itself closely correlated with $\phi$, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
@@
 That study showed that the strength of the correlation between some traits (including $\phi$ and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Importantly, they proposed putative mechanisms to explain their results \citep{Gibert-2016}.
-In this paper, we test these mechanisms by providing a fully functioning model of plant growth.
+Here, we extend previous work \citep{Gibert-2016} to provide a mechanistic explanation of these empirical patterns via a model of plant growth and allocation.
 
@@
 $^2$ Similar responses are predicted for leaf lifespan; here directly related to $\phi$.
-$^3$ \citep{Baltzer-2007} reports a strong relationship between \textsc{wplcp} and leaf respiration rate per unit leaf area or mass.}
+$^3$ \citep{Baltzer-2007} reports a strong relationship between \textsc{wplcp} and leaf respiration rate per unit leaf area or mass.
+
+}
 \label{tab:phenomena}
@@
 Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until recently \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
-Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}.
 A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to $\phi$ \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
@@
 Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
-Without further guidance, many researchers expected traits to linearly map onto growth rates, i.e.
-there is a ``fast'' and ``slow'' growth end for each trait \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
+Without further guidance, many researchers have expected traits to map directy to growth rates, i.e.
+there is a ``fast'' and ``slow'' growth traits \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
 
@@
 On larger plants, growth is measured mainly as increment in either stem diameter or height \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-As we outline below, mass production is only one of the factors influencing diameter growth; making theory considering only primary production of limited utility.
+As we outline below, height and diameter growth depend not only on mass production but also allocation of biomass and tissue construction costs, meaning theory considering only primary production of limited utility.
 
@@
 \end{tabular}
-\addtabletext{For $M_i$, $r_i$, $k_i$ and \(A_i\), subscripts refer to: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,  \({\rm h}\) = heartwood, \({\rm r}\) = roots, \({\rm st}\) = total stem, \({\rm a}\) =  alive tissue.}
+\addtabletext{For $M_i$, $r_i$, $k_i$ and \(A_i\), subscripts refer to: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,  \({\rm h}\) = heartwood, \({\rm r}\) = roots, \({\rm st}\) = total stem, \({\rm a}\) =  alive tissue.
+
+}
 \label{tab:definitions}
@@
 Earlier studies on seedlings related relative growth rate in mass to $\phi$ \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
-This model can be derived from eqn \ref{eq:dbdt}, as a special case of the more-extended model described in the following sections, as follows.
+This model can be derived from eqn \ref{eq:dbdt} as follows and is a special case of the more-extended model described below.
 Ignoring all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}, net production becomes a linear function of leaf area, making relative growth rate in mass:
@@
 
-We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the \textsc{baad} \citep{Falster-2015b}.
+We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}.
 We also evaluated an additional prediction arising from the eqns in Table \ref{tab:allometry}a.
@@
 \centering
-\includegraphics[width=11.4cm,height=11.4cm, keepaspectratio]{main/growth_light_height.pdf}
+\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{main/growth_light_height.pdf}
 \caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
@@
 The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
-The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individual's growth trajectory.
+The remaining three traits are termed ``developmental traits'' because they moderate the rate of movement along an individual's growth trajectory.
 Across the five different traits, we observed four relatively distinct types of response.
@@
 
-\subsubsection{Development traits}  The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
+\subsubsection{Developmental traits}  The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
 \ref{fig:growth_light_height}).
@@
 \includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three development traits on shade tolerance.}
+\caption{\textbf{Effect of three developmental traits on shade tolerance.}
 Panels show effect of traits on level of canopy openness that causes net production (eqn \ref{eq:dbdt}) to be zero.
@@
 
-The model presented here extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
+The model presented here extends a widely-used model for seedlings, linking mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
 Importantly, the seedling model can be derived as a special case of the new-extended model (eqn \ref{eq:RGR}).
@@
 It is this property that makes them shade intolerant.
-And yet they can still achieve a growth advantage (when small), because the benefits of cheap leaf construction outweigh the costs of high leaf turnover.
-
-% - \citep{King-2005} shows that the cost of turning over leaves, branches and fine roots may halve the height growth rate that could otherwise be attained.
+And yet they can still achieve a growth advantage (when small).
 
@@
 As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}.
-These models have several properties in common, including that they all have growth being driven by the gross amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
+These models have several properties in common, including that they all have growth being driven by the net amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
 Many models also make functional-balance assumptions, for example linking the cross-section of sapwood to leaf area \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-2005, Moorcroft-2001}.
@@
 \matmethods{
+\vspace{-1em}
 \subsection{Trait-based trade-offs}
@@
 
-$\phi$ influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+The trait $\phi$ influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
 In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
@@
 
-$\rho$ directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+The trait $\rho$ directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
 In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\rho / \rho_0\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
@@
 To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
-The analyses presented employ best practises in scientific computing \citep{Wilson-2014} and are fully reproducible via code available at
-\href{https://github.com/traitecoevo/growth\_trajectories}{github.com/traitecoevo/growth\_trajectories}.
+
 }

### 2017-08-09 R/figures.R

@@ -132,7 +132,9 @@ make_hyperpar2 <- function(
         ret <- c(last(AA), 0)
         names(ret) <- c("p1","p2")
       } else {
-        fit <- nls(AA ~ p1 * E/(p2 + E), data.frame(E = E, AA = AA), start = list(p1 = 100, p2 = 0.2))
+        fit <- nls(AA ~ p1 * E/(p2 + E),
+                  data.frame(E = E, AA = AA),
+                  start = list(p1 = 100, p2 = 0.2))
         ret <- coef(fit)
       }
       ret

### 2018-04-05 ms/figure_conceptual.tex
Tightening down to page limit
Revise paper after reviewers comments

@@
 \documentclass{beamer}
-\usepackage[size=custom, width=21, height=28, scale=1.0, orientation=portrait]{beamerposter}
+\usepackage[size=custom, width=21, height=29, scale=1.0, orientation=portrait]{beamerposter}
 \setbeamertemplate{navigation symbols}{}
@@
 \graphicspath{{../}}
+
+% Use square brackets for equations
+\usepackage{mathtools}
+\newtagform{brackets}{[}{]}
+\usetagform{brackets}
 
@@
   \node at (#1)[anchor=center] {
-  {\footnotesize #2}};
+  {\small #2}};
 }
@@
 
-\node at ($(current page.north)+ (0cm, -7.5cm)$) (top){};
+\node at ($(current page.north)+ (0cm, -8.5cm)$) (top){};
 
-\node[rectangle, anchor = center] at ($(top)+ (0cm, 4cm)$){
+\node[rectangle, anchor = center] at ($(top)+ (0.75cm, 5.25cm)$){
   \includegraphics[width=18cm]{output/main/mass_fraction.pdf}};
 
-\node at ($(top) + (-9.25cm, 6.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
+\node[rectangle, anchor = center, rotate=90, text width =6cm] at ($(top)+ (-8.75cm, 7cm)$){
+  \small Fraction\\[-1ex] of mass
+  };
+
+\node at ($(top) + (-9.25cm, 7.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
   {\bf \footnotesize Standing mass}};
 
-\node at ($(top) + (-10.5cm, 6.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
-  {\bf \footnotesize a)}};
+\node at ($(top) + (-10.5cm, 7.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
+  {\bf \small a)}};
 
 \node at ($(top) + (-10.5cm, 1.5cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
-  {\bf \footnotesize b)}};
+  {\bf \small b)}};
 
@@
 \defineRow{0}{thisrow}
-\rowtitle{Biomass production}
+\rowtitle{Biomass growth}
 \firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/net_mass_production_dt_net_mass_production_dt}{}
-\othercol{2}{\hspace{-1em} \scriptsize $\alpha_{\rm \tiny bio} \alpha_{\rm \tiny y}$}{output/conn/net_mass_production_dt_yield}{}
+\othercol{2}{\hspace{-1em} \scriptsize $\,\alpha_{\rm \tiny bio} \alpha_{\rm \tiny y}$}{output/conn/net_mass_production_dt_yield}{}
 \othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/conn/net_mass_production_dt_assimilation}{}
@@
 
-\textbox{$(top)+ (0cm, -20cm)$}{{\bf Plant height (m)}}
+\textbox{$(top)+ (0cm, -20cm)$}{Plant height (m)}
+\textbox{$(top)+ (0cm, 2.5cm)$}{Plant height (m)}
 

### 2018-04-05 ms/ms-common-defs.tex
Revise paper after reviewers comments

--- 2016-07-29 ms/ms-common-defs.tex
@@
-\usepackage[rgb,dvipsnames]{xcolor}
-\definecolor{grey}{rgb}{0.5, 0.5, 0.5}
-
+% abbreviations
 \newcommand{\smurl}[1]{\url{#1}}
@@
 \newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
-
-\title{Untangling the link between traits, size and growth rate in plants}
-\runninghead{I\lowercase{nfluence of traits on growth}}
-
-\author{Daniel S. Falster\textasteriskcentered \& Richard G. FitzJohn}
-
-\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia
-\\
-\textasteriskcentered Correspondence author. E-mail: \texttt{adaptive.plant@gmail.com}
-}
-
-\usepackage{natbib}
-\bibliographystyle{ecol_let}
-\setcitestyle{authoryear,open={(},close={)}}
-% allows for numbered referencing, as required by ecol letters
-\usepackage{etoolbox}
-\newbool{MyRefNumbers}
-\booltrue{MyRefNumbers} % comment to remove numbers in reference list
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\lma}{\textsc{lma}}
+\newcommand{\wood}{\textsc{wd}}
+\newcommand{\seed}{\textsc{sm}}
+\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
+\newcommand{\nitrogen}{\textsc{n}}
+\newcommand{\sma}{\textsc{sma}}
 
@@
 
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Daniel S. Falster}
+\author[b]{Remko A. Duursma}
+\author[a,c]{Richard G. FitzJohn}
 
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
+Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
+\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
+\affil[c]{Imperial College, London, United Kingdom}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Falster}

### 2018-04-05 ms/ms-supp.Rnw
Add longtable package
minor edits, address change
Revise paper after reviewers comments

@@
-\documentclass[10pt,twoside]{article}
+\documentclass[9pt,onecolumn,twoside]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
 
-\usepackage{suppmat}
-\usepackage{listings}
-\usepackage[T1]{fontenc} % Better underscores
+\templatetype{pnasresearcharticle}
+
+\title{
+{\normalfont \fontsize{12pt}{16pt}
+{\color{black50} Supporting information for:}\\} \vskip6pt
+\fontsize{16pt}{20pt} How functional traits influence plant growth and shade tolerance across the life-cycle
+}
 
@@
 
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\dates{This supplement was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
-\titleprefix{Supporting Materials for:}
-
-% Figure command -- also adds figure to table of contents
-\newcommand{\myfigure}[4]{
-    \begin{figure}
-            \centering
-            \includegraphics{#1}
-            \caption{\textbf{#2} #3}
-            \label{#4}
-    \end{figure}
-  \addcontentsline{toc}{subsection}{Fig \ref{#4} -- #2}
+\hypersetup{
+     linkcolor=pnasbluetext
 }
-
-\date{}
-
-<<preload, echo=FALSE>>=
-library(knitr)
-library(xtable)
-opts_chunk$set(echo=FALSE)
-@
 
@@
 
-\tableofcontents
-
 \renewcommand{\thefigure}{S\arabic{figure}}
 \renewcommand{\thetable}{S\arabic{table}}
-\setcounter{secnumdepth}{0}
+\renewcommand{\theequation}{S\arabic{equation}}
+\renewcommand{\thesection}{S\arabic{section}}
 
-\clearpage
+% Figure command -- also adds figure to table of contents
+\newcommand{\myfigure}[4]{
+    \begin{figure*}[!hb]
+            \centering
+             \includegraphics[width=0.9\textwidth]{#1}
+            \caption{\textbf{#2} #3}
+            \label{#4}
+    \end{figure*}
+}
 
-\section{A -- Additional details on the \texttt{FF16} growth model in the {\plant} package}
+\section{Additional details on the \texttt{FF16} growth model in the {\plant} package}
 
 In this paper we use the \texttt{FF16} growth model from the package {\plant} \citep{Falster-2016}, which is accessed via \textsc{r} \citep{R-2015}.
-A full derivation of  Eqns 1-21 from the main text are described is provided in Appendix of \citet{Falster-2016}, see also \citep{Falster-2011}.
-Below we repeat some additional details from \citet{Falster-2016}, so that readers have a complete picture of the model.
+A full derivation of  Eqns 1-21 from the main text are described is provided in Appendix of \citep{Falster-2016}, see also \citep{Falster-2011}.
+Below we repeat some additional details from \citep{Falster-2016}, so that readers have a complete picture of the model.
 
@@
 {\plant} allows for the hyper-parameterisation of the \texttt{FF16} physiological model via plant functional traits: this enables simultaneous variation in multiple parameters in accordance with an assumed trade-off.
-The hyper-parameterisation function used in the current analysis differs slightly from that in \citet{Falster-2016}.
+The hyper-parameterisation function used in the current analysis differs slightly from that in \citep{Falster-2016}.
 The functions implemented are as follows.
-
-\subsubsection{Leaf mass per unit area}
-
-The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$.
-In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citet{Wright-2004},
-$$k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.$$
-This relationship is normalised around $\phi_0$, the global mean of $\phi$.
-This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.
-
-We also vary the mass-based leaf respiration rate so that it stays constant per unit leaf area and varies with $\phi$ and nitrogen per unit leaf area $\nu$, as empirically observed by \citet{Wright-2004},
-$$r_{\rm l} = \frac{\beta_{{\rm lf4}}\, \nu}{\phi}.$$
-
-\subsubsection{Wood density}
-
-The trait wood density, denoted by $\rho$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$.
-In addition, we allow for $\rho$ to influence the rate of growth-independent mortality,
-$$d_{\rm I} = \beta_{{\rm dI1}} \, \left(\frac{\rho}{\rho_0}\right) ^ {-\beta_{{\rm dI2}}},$$
-and also the rate of sapwood turnover,
-$$k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^ {-\beta_{{\rm ks2}}}.$$
-As for $\phi$, these relationships are normalized around $\rho_0$, the global mean of $\rho$.
-By default, $\beta_{{\rm kI2}}$ and $\beta_{{\rm ks2}}$ are set to zero, so these linkages only become present when these parameters are set to something other than their default values.
-
-The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as
-$$r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},$$
-where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood. Similarly, the rate of bark respiration per unit mass varies as
-$$r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},$$
-with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.
 
@@
 
-Effects of the trait seed mass, denoted by $\omega$, are naturally embedded in the equation determining fecundity (Eq.
-\ref{eq:fecundity}) and the initial height of seedlings.
+Effects of the trait seed mass, are naturally embedded in the equation determining fecundity and the initial height of seedlings.
 In addition, we let the accessory cost per seed be a multiple of seed size,
-$$\alpha_{{\rm f3}} = \beta_{{\rm f1}} \omega,$$
+$\alpha_{{\rm f3}} = \beta_{{\rm f1}} \omega,$
 as empirically observed \citep{Henery-2001}.
@@
 To calculate the average annual photosynthesis for a leaf, we integrate the instantaneous rate per unit leaf area over the annual solar trajectory, using a rectangular-hyperbolic photosynthesis light response curve,
-
-$$p(\nu, E) = \frac1{365 \textrm{d}}\int_0^{365 \textrm{d}}\frac{Y(t) + A_{\rm max} - \sqrt{(Y(t) + A_{\rm max})^2 - 4 \beta_{{\rm lf2}} Y(t) A_{\rm max} }}{2 \beta_{{\rm lf2}}} {\rm d}t,$$
-where
-\begin{itemize}
-\item $A_{\rm max}$ is the maximum photosynthetic capacity of the leaf,
-\item $\beta_{{\rm lf2}}$ is the curvature of the light response curve,
-\item $Y(t) = \beta_{{\rm lf3}} I(t)$ is the initial yield of the light response curve, with $\beta_{{\rm lf3}}$ being the quantum yield parameter,
-\item $I(t) = k_{\rm I} \, I_0(t)\, E$ is the intensity of light on the leaf surface, and
-\item $I_0(t)$ is light incident on a surface perpendicular to the sun's rays directly above the canopy at time $t$.
-\end{itemize}
-The profile of $I_0(t)$ is given by a solar model adapted from \citet{TerSteege-1997}.
+\begin{equation} \label{eq:pts}
+p(\nu, E) = \frac1{365 \textrm{d}}\int_0^{365 \textrm{d}}\frac{Y(t) + A_{\rm max} - \sqrt{(Y(t) + A_{\rm max})^2 - 4 \beta_{{\rm lf2}} Y(t) A_{\rm max} }}{2 \beta_{{\rm lf2}}} {\rm d}t,
+\end{equation}
+where $A_{\rm max}$ is the maximum photosynthetic capacity of the leaf, 
+$\beta_{{\rm lf2}}$ is the curvature of the light response curve,
+$Y(t) = \beta_{{\rm lf3}} I(t)$ is the initial yield of the light response curve, with $\beta_{{\rm lf3}}$ being the quantum yield parameter,
+$I(t) = k_{\rm I} \, I_0(t)\, E$ is the intensity of light on the leaf surface, and
+$I_0(t)$ is light incident on a surface perpendicular to the sun's rays directly above the canopy at time $t$.
+The profile of $I_0(t)$ is given by a solar model adapted from \citep{TerSteege-1997}.
 
@@
 
+\subsubsection{Leaf mass per unit area}
+
+The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$.
+In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citep{Wright-2004},
+\begin{equation}
+k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.
+\end{equation}
+This relationship is normalised around $\phi_0$, the global mean of $\phi$.
+This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.
+
+We also vary the mass-based leaf respiration rate so that it stays constant per unit leaf area and varies with $\phi$ and nitrogen per unit leaf area $\nu$, as empirically observed by \citep{Wright-2004},
+\begin{equation}
+r_{\rm l} = \frac{\beta_{{\rm lf4}}\, \nu}{\phi}.
+\end{equation}
+
+\subsubsection{Wood density}
+
+The trait wood density, denoted by $\rho$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$.
+As for $\phi$, these relationships are normalized around $\rho_0$, the global mean of $\rho$.
+By default, $\beta_{{\rm kI2}}$ and $\beta_{{\rm ks2}}$ are set to zero, so these linkages only become present when these parameters are set to something other than their default values.
+
+The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as
+\begin{equation}
+r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},
+\end{equation}
+where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood. Similarly, the rate of bark respiration per unit mass varies as
+\begin{equation}
+r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},
+\end{equation}
+with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.
+
 \subsubsection{Implementation of hyper-parameterisation function}
@@
 
-<<echo=FALSE>>=
+<<echo=FALSE, highlight=TRUE, comment = ''>>=
 make_hyperpar2
+#txt <- capture.output(make_hyperpar2)
+#writeLines(strwrap(txt, width = 150, exdent = 10))
 @
 
-\clearpage
+\section{References}
+\bibliography{references}
 
-\section{B -- Supporting Tables}\label{app:supp_info_table}
-
-\addcontentsline{toc}{subsection}{Table \ref{tab:params_core} -- Core parameters of the \texttt{FF16} physiological model}
+\section{Supporting Tables and Figures}\label{app:supp_info_table}
 
@@
             caption="Core parameter of the \texttt{FF16} physiological model."),
+      floating.environment="table*",
       sanitize.text.function=I,
-      include.rownames=FALSE, caption.placement = "top")
+      include.rownames=FALSE, 
+      caption.placement = "top")
 @
-
-\newpage
-
-\addcontentsline{toc}{subsection}{Table \ref{tab:params_hyper} -- Parameters for hyper-parameterisation of the \texttt{FF16} physiological model}
 
@@
             hline.after=c(1),
-            align='lp{7cm}llll',
+            align='lp{9cm}llll',
             label="tab:params_hyper",
@@
       sanitize.text.function=I,
-      include.rownames=FALSE, caption.placement = "top")
+      floating.environment="table*",
+      include.rownames=FALSE, 
+      caption.placement = "top")
 @
 
-\clearpage
-
-\section{C -- Supporting Figures}\label{app:supp_info_figures}
+\newpage
 
@@
 according to strength of the relationship.}{fS-leaf}
+
+\newpage
+
+\myfigure{main/assumptions.pdf}{Assumptions of a functional balance model for plant construction.}{Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
+Blue lines show standardised major axis lines fit to different species.
+The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table \ref{tab:allometry}.}{fig:assumptions}
 
@@
 
-\clearpage
-
-\addcontentsline{toc}{section}{References}
-\bibliography{references}
-
 \end{document}

### 2018-04-05 ms/ms.tex
Enable line numbers, in grey
Revise paper after reviewers comments

@@
-\documentclass[9pt,twocolumn,twoside]{pnas-new}
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
 
@@
 
-\usepackage{graphicx}
-\graphicspath{{../output/}}
-
-\newcommand{\smurl}[1]{\url{#1}}
-\newcommand{\ud}{\ensuremath{\rm{d}}}
-\newcommand{\tabitem}{~~\llap{\textbullet}~~}
-\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
-\newcommand{\plant}{\texttt{plant}}
-\newcommand{\wplcp}{\textsc{wplcp}}
-\newcommand{\sma}{\textsc{sma}}
+\input{ms-common-defs}
 
@@
 
-\title{Trajectories: how functional traits influence plant growth and shade tolerance across the life-cycle}
-
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Daniel S. Falster}
-\author[b]{Remko A. Duursma}
-\author[a,c]{Richard G. FitzJohn}
-
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\affil[a]{Evolution \& Ecology Research Centre, School of Biological, Earth and
-Environmental Sciences, University of New South Wales, Sydney, NSW 2052, Australia}
-\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
-\affil[c]{Imperial College, London, United Kingdom}
-
-% Please give the surname of the lead author for the running footer
-\leadauthor{Falster}
+\title{How functional traits influence plant growth and shade tolerance across the life-cycle}
 
@@
 This framework captures diverse patterns of growth in relation to size, and explains why the effect of traits on growth changes through ontogeny.
-By disentangling the effects of plant size, light environment and traits on growth rates, this study provides a theoretical foundation for understanding growth across diverse tree species around the world.
+By disentangling the effects of plant size, light environment, and traits on growth rates, this study provides a theoretical foundation for understanding growth across diverse tree species around the world.
 }
@@
 Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth and shade-tolerance by extending a widely-used model, linking growth rate in seedlings with a single leaf trait, to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
-Based on biomass production and allocation, this framework explains why the influence of traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
+Based on biomass growth and allocation, this framework explains why the influence of traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
 Specifically, we demonstrate why for height growth the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
@@
 
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+
 \begin{document}
@@
 
-\dropcap{F}unctional traits capture core differences in the strategies plants use to generate and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
-Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover.
-During the last two decades, trade-offs related to some prominent traits have been quantified, with values for traits such as leaf mass per unit leaf area, wood density, and seed size now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
+\dropcap{F}unctional traits capture core differences in the strategies plants use to acquire and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
+Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients, and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover.
+During the last two decades, trade-offs related to some prominent traits have been quantified, with values for some leaf, wood, or seed traits now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
 As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
@@
 While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the trait leaf mass per unit leaf area ($\phi$) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
-$\phi$ and leaf lifespan, itself closely correlated with $\phi$, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
+In seedlings, the trait leaf mass per unit leaf area ({\lma}) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
+{\lma} and a close correlate leaf lifespan, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
 These early successes prompted researchers to search for similar relationships in large plants.
-However, the results showed that in saplings and trees, $\phi$ was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
-Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
+However, in saplings and trees showed, {\lma} was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
+Meanwhile, other traits such as wood density ({\wood}) showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
 Clearly, traits do not correlate simply and consistently to growth rate.
@@
 A recent meta-analysis of 103 studies reporting \textgreater 500 correlations provides the most compelling evidence \citep{Gibert-2016}.
-That study showed that the strength of the correlation between some traits (including $\phi$ and maximum height) and growth rate was modified by size, while for other traits (including wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
+That study showed that the strength of the correlation between some traits (including {\lma} and maximum height) and growth rate was modified by size, while for other traits (including {\wood} and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
 Here, we extend previous work \citep{Gibert-2016} to provide a mechanistic explanation of these empirical patterns via a model of plant growth and allocation.
@@
 \textbf{Change in growth rate with increasing size (Fig. \ref{fig:conceptual})}\\
- $\,$ Net biomass production: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
+ $\,$ Biomass growth: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
  $\,$ Plant mass: Increasing \citep{Sillett-2010, Stephenson-2014} \\
@@
 \textbf{Effect of traits on growth rate (Fig. \ref{fig:growth_light_height})}\\
- $\,$ $\omega$: low values $\rightarrow$ smaller seedlings, lower absolute \& higher relative growth \citep{Gibert-2016} \\
- $\,$ $H_{\rm m}$: low values $\rightarrow$ slower growth at at larger sizes \citep{Gibert-2016}\\
- $\,$  $\nu$: low values $\uparrow$ growth irrespective of size, but only in high light \citep{Gibert-2016}\\
- $\,$ $\phi$: low values $\uparrow$ growth when small, not at mid-large sizes \citep{Gibert-2016}\\
- $\,$ $\rho$: low values $\uparrow$ growth, except at largest sizes \citep{Gibert-2016}\\
+ $\,$ {\seed}: $\downarrow$ values $\downarrow$ seedling size, \& thus $\downarrow$ absolute \& $\uparrow$ relative growth rate \citep{Gibert-2016} \\
+ $\,$ $H_{\rm mat}$: $\downarrow$ values $\downarrow$ growth rate at at larger sizes \citep{Gibert-2016}\\
+ $\,$  {\nitrogen}: $\downarrow$ values $\uparrow$ growth rate irrespective of size, but only in high light \citep{Gibert-2016}\\
+ $\,$ {\lma}: $\downarrow$ values $\uparrow$ growth rate when small, not at mid-large sizes \citep{Gibert-2016}\\
+ $\,$ {\wood}: $\downarrow$ values $\uparrow$ growth rate, except at largest sizes \citep{Gibert-2016}\\
 \textbf{Responsiveness of growth rate to changes in light, $E$ (Fig. \ref{fig:growth_light_height})}\\
- $\,$ $\nu$: high values respond more \\
- $\,$ $\phi$: low values respond more \\
- $\,$ $\rho$: low values respond more \citep{Ruger-2012}\\
+ $\,$ {\nitrogen}: $\uparrow$ values respond more \\
+ $\,$ {\lma}: $\downarrow$ values respond more \\
+ $\,$ {\wood}: $\downarrow$ values respond more \citep{Ruger-2012}\\
 \textbf{Shade tolerance, \textsc{wplcp} (Fig. \ref{fig:wplcp})}\\
- $\,$ Decreasing with size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
- $\,$ $\nu$: low values $\uparrow$ shade tolerant \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
- $\,$ $\phi$: low values $\downarrow$ shade tolerant \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$, higher \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
- $\,$ $\rho$: low values less shade tolerant \citep{Osunkoya-1996}\\
+ $\,$ Size: decreasing \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
+ $\,$ {\nitrogen}: $\downarrow$ values $\uparrow$ shade tolerance \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
+ $\,$ {\lma}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$ \& $\uparrow$ \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
+ $\,$ {\wood}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Osunkoya-1996}\\
   \bottomrule
 \end{tabular}
-\addtabletext{Traits are seed mass ($\omega$), height at maturation ($H_{{\rm mat}}$), leaf nitrogen content per unit leaf area ($\nu$), leaf mass per unit leaf area ($\phi$), and wood density ($\rho$); for details see Tables \ref{tab:definitions}-\ref{tab:traits}.
-\textsc{lai} = leaf area index.
-Footnotes: $^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area \& dark respiration rate per leaf area; here both directly related to $\nu$.
-$^2$ Similar responses are predicted for leaf lifespan; here directly related to $\phi$.
-$^3$ \citep{Baltzer-2007} reports a strong relationship between \textsc{wplcp} and leaf respiration rate per unit leaf area or mass.
+\addtabletext{{\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density, \textsc{lai} = leaf area index.
+Footnotes: $^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area \& dark respiration rate per leaf area; here both related to {\nitrogen}.
+$^2$ Similar responses are predicted for leaf lifespan; here directly related to {\lma}.
+$^3$ \citep{Baltzer-2007} finds a relationship between \textsc{wplcp} \& leaf respiration rate expressed per unit leaf area or per unit leaf mass.
 
@@
 
-\begin{figure}[!hb]
+\begin{table}[!hb]
+ \caption{Key trade-offs (benefit and cost) for the five traits considered, as encoded into the {\plant} model.
+ }
+\centering
+  \begin{tabular}{p{0.5cm}p{3.6cm}p{2.5cm}p{0.4cm}}
+  \toprule
+  Trait &  Benefit & Cost & Ref. \\
+  \midrule
+  {\seed} & $\uparrow$ size of seedlings & $\downarrow$ seed production rate & \citep{Moles-2006}\\
+  {\hmat}  & $\uparrow$ growth rate & $\downarrow$ seed production rate & \citep{Thomas-2011}\\
+  {\nitrogen} & $\uparrow$ photosynthetic rate in high light & $\uparrow$ respiration rate & \citep{Wright-2004}\\
+  {\lma} & $\downarrow$ leaf turnover rate & $\uparrow$ cost of building leaf & \citep{Wright-2004}\\
+  {\wood} & $\downarrow$ sapwood turnover rate & $\uparrow$ cost of building stem & \citep{Chave-2009}\\
+  \bottomrule
+  \end{tabular}
+\addtabletext{Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density.
+
+}
+\label{tab:traits}
+\end{table}
+
+Interpreting diverse empirical results seeking to link traits to growth rate has been challenging because, until recently \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
+A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to {\lma} \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
+An extension of this equation suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
+But as noted above, the prediction for large plants is not supported by empirical results.
+Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
+Without further guidance, many researchers expected traits to map directly onto growth rates \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
+
+One reason theoretical predictions have been lacking or not been supported is that in existing theory the effects of traits is realised mainly via influences on net primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
+By contrast, the physiology of traits such as {\lma} and {\wood} suggests they influence allocation of biomass among different tissues, and their turnover rates, rather than net primary production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
+A second concern for theory focussed on net primary production is that measuring production is really only practical for small plants.
+On larger plants, growth is measured mainly as increment in either stem diameter or height \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
+As outlined below, growth in height and diameter also depend on the allocation of biomass and tissue construction costs.
+
+Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence key elements of plant demography across the life-cycle.
+The {\plant} model builds on past approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
+Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
+Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
+Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size.
+We consider five prominent traits representing specific physiological trade-offs: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).
+
+We also show how our approach can account for related phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment (Table \ref{tab:phenomena}).
+Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
+In absolute terms, growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011}, diameter \citep{Canham-2004, Canham-2006, Herault-2011} or biomass growth \citep{Givnish-1988, Koch-2004}.
+In contrast, the growth rate of standing plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
+Expressed as relative growth rates, all growth measures decrease sharply with size \citep{Rees-2010, Iida-2014}.
+Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
+These diverse phenomena deserve an integrated explanation.
+
+\begin{figure}[!ht]
 \centering
@@
 \textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
-\textbf{b)} Equations describing the rates of biomass production and growth in various dimensions of the plant.
+\textbf{b)} Equations describing the rates of biomass growth, and growth in various dimensions of the plant.
 In the first line the symbol $\Sigma$ means ``sum'' across tissues, where $i = {\rm l,b,s,r}$.
@@
 
-Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until recently \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
-A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to $\phi$ \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-An extension of this equation suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
-But as noted above, the prediction for large plants is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
-Without further guidance, many researchers have expected traits to map directy to growth rates, i.e.
-there is a ``fast'' and ``slow'' growth traits \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
+\section*{Modeling the effects of traits on demography}
 
-One reason theoretical predictions have been lacking or not been supported is that in existing theory the effects of traits is realised mainly via influences on primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
-By contrast, the physiology of traits such as $\phi$ and wood density suggests they influence allocation of biomass among different tissues rather than mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
-A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants.
-On larger plants, growth is measured mainly as increment in either stem diameter or height \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-As we outline below, height and diameter growth depend not only on mass production but also allocation of biomass and tissue construction costs, meaning theory considering only primary production of limited utility.
+The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
+\ref{fig:conceptual}a).
+We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive, t=total, and st=stem total (s+b+h).
+The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
+A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
 
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
-\caption{Conceptual framework linking shade-tolerance to plant size and traits, adapted from \citep{Givnish-1988}.
-Shade tolerance is the level of canopy openness $E$ where photosynthetic income (dashed line) intersects with the sum of respiration and turnover costs over tissues (solid black lines).
-The black circles indicate the point of intersection for plants with different heights.
-All income and cost is expressed in units of dry mass produced  per unit leaf area per year.
-Note the costs of sapwood and bark increases with height.
-Traits can impact on shade tolerance by, in the case of leaf nitrogen content, shifting the income line up or down, or in the case of leaf mass per unit leaf area or wood density, causing the cost components to increase or decrease.
-\label{fig:wplcp_idea}}
-\end{figure}
+Applying a standard approach, the amount of biomass growth on the plant, $\frac{{\rm d}B}{{\rm d}t}$, is given by the difference between income (photosynthesis) and losses (respiration and turnover)\citep{Makela-1997, Thornley-2000}:
+\begin{equation}\label{eq:dbdt}
+\frac{{\rm d}B}{{\rm d}t}
+= \alpha_{\rm bio} \,
+\alpha_{\rm y}
+\bigg( A_{\rm l} \, \bar{p}(E) -
+\, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}\bigg)
+- \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}.
+\end{equation}
+Photosynthesis is the product of the average photosynthetic rate per unit leaf area, $\bar{p}(E)$, and total leaf area, $A_{\rm l}$.
+We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (see SI for details), and respiration and turnover rates of different tissues are constants that potentially differ with traits.
+The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for the yield (fraction of C fixed in biomass) and the conversion of CO$_2$ into units of biomass, respectively.
+While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment so that we can better understand trait- and size-related effects.
+Many vegetation models also use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$ -- e.g.
+as functions of temperature -- though such detail will not qualitatively alter model behaviour.
 
@@
 \multicolumn{3}{l}{\textbf{Traits (constant through ontogeny)}}\\
-$\omega$ & kg & Seed mass  \\
-$H_{{\rm mat}}$ & m & Height at maturation\\
-$\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf nitrogen per unit leaf area  \\
-$\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf mass per area \\
-$\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & Wood density \\
-\multicolumn{3}{l}{\textbf{State variables (may vary through ontogeny)}} \\
+$\omega$ & kg & Seed mass ({\seed})  \\
+$H_{{\rm mat}}$ & m & Height at maturation ({\hmat})\\
+$\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf nitrogen per unit leaf area ({\nitrogen})  \\
+$\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf mass per area ({\lma}) \\
+$\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & Wood density ({\wood})\\
+\multicolumn{3}{l}{\textbf{State or rate variables (may vary through ontogeny)}} \\
 $H$ & m & Height of a plant\\
-$B$ & kg & Biomass originating from parent plant\\
 $D$ & m & Stem diameter\\
@@
 $A_i$ & m$^2$ & Surface or cross-section are of tissue type $i$\\
+$\frac{{\rm d}B}{{\rm d}t}$ & kg yr$^{-1}$& Biomass growth rate\\
 $p,\bar{p}$ & mol yr$^{-1}$ m$^{-2}$ & Photosynthetic rate per unit area \\
@@
 
-Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle.
-The {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
-Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below \citep[from][]{Falster-2011}, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
-Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of primary production and allocation, how and why the effect of some traits on growth rates changes with size.
-We also show how our approach can account for related phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment.
-Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
-Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Canham-2004, Canham-2006, Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}.
-In contrast, the growth rate of standing plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
-All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}.
-Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
-The diverse phenomena in Table \ref{tab:phenomena} deserve a comprehensive and integrated explanation.
-
-\section*{Framework for understanding the effects of traits on growth}
-
-The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
-\ref{fig:conceptual}a).
-We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h).
-The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A summary of all variables, units and definitions is given in Table \ref{tab:definitions}-\ref{tab:traits}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
-
-We assume growth is driven by biomass production and its subsequent distribution throughout the plant.
-Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover)\citep{Makela-1997, Thornley-2000}:
-\begin{equation}\label{eq:dbdt}
-\frac{{\rm d}B}{{\rm d}t}
-= \alpha_{\rm bio} \,
-\alpha_{\rm y}
-\big( A_{\rm l} \, \bar{p}(E) -
-\, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}\big)
-- \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}.
-\end{equation}
-Photosynthesis is the product of the average photosynthetic rate per unit leaf area, $\bar{p}(E)$, and total leaf area, $A_{\rm l}$.
-We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (Fig.
-\ref{fig:wplcp_idea}; for details, see SI), and respiration and turnover rates of different tissues are constants that might differ with traits (see below).
-The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for growth respiration and the conversion of CO$_2$ into units of biomass, respectively.
-While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment so that we can better understand the intrinsic trait- and size-related effects.
-Many vegetation models also use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g.
-as functions of temperature.
-Such detail is not needed here because it will not change general model behaviour, though it alters the absolute values of predicted growth rates.
-
 \subsection{Classic model for mass-based relative growth rate}
 
-Earlier studies on seedlings related relative growth rate in mass to $\phi$ \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}.
-This model can be derived from eqn \ref{eq:dbdt} as follows and is a special case of the more-extended model described below.
-Ignoring all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}, net production becomes a linear function of leaf area, making relative growth rate in mass:
+Earlier studies on seedlings related relative growth rate in mass to {\lma} \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}, using a model that can be derived from eqn \ref{eq:dbdt} as a special case of the more-extended model described below.
+Ignoring all turnover terms, as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt} and reproduction, biomass growth becomes a linear function of leaf area, making relative growth rate in mass:
 \begin{equation}\label{eq:RGR}
 \frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx P_{\rm net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
-where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$.
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
+where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$, and $\phi$ is {\lma}.
+Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to {\lma} \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
 The derivations below clarify these links.
-
-\begin{table}[t]
- \caption{Key trade-offs (benefit and cost) for the five traits considered, as encoded into the {\plant} model.
- }
-\centering
-  \begin{tabular}{p{0.5cm}p{3.25cm}p{2.25cm}p{1cm}}
-  \toprule
-  Trait &  Benefit & Cost & Ref. \\
-  \midrule
-  $\omega$ & $\uparrow$ size seedlings & $\downarrow$ fecundity & \citep{Moles-2006}\\
-  $H_{{\rm mat}}$  & $\uparrow$ growth & $\downarrow$ fecundity & \citep{Thomas-2011, Wenk-2015}\\
-  $\nu$ & $\uparrow$ photosynthesis (high light) & $\uparrow$ respiration rate & \citep{Wright-2004}\\
-  $\phi$ & $\downarrow$ leaf turnover & $\uparrow$ cost building leaf & \citep{Wright-2004}\\
-  $\rho$ & $\downarrow$ sapwood turnover & $\uparrow$ cost building stem & \citep{Chave-2009}\\
-  \bottomrule
-  \end{tabular}
-\label{tab:traits}
-\end{table}
 
@@
 
-To model growth in plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for net mass production, but also for the costs of building new tissues and allocation to reproduction.
-Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}.
-Equations are given in Fig.
-\ref{fig:conceptual}b, (eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
-As is evident in eqn  Fig.
-\ref{fig:conceptual}b, the growth rate in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) all share some terms.
-Many of the terms in Fig.
-\ref{fig:conceptual} also vary intrinsically with size.
-The insets in Fig.
-\ref{fig:conceptual}b show the size-related patterns in different variables for a typical plant, obtained from applying the specific allocation model in the next section.
+To model growth in plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for biomass growth, but also for the costs of building new tissues and allocation to reproduction.
+Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig.
+\ref{fig:conceptual}b, eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
+While other decompositions are possible, the approach used here arises from the idea that multiple aspects of function and allocation decisions are made in relation to the amount of leaf area on a plant  \citep{Falster-2011}.
+Growth rates in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), total leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) therefore share several terms.
+Many of these terms also vary intrinsically with size, as shown in the insets in Fig.
+\ref{fig:conceptual}b.
 
-The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
+The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass growth $\frac{{\rm d}B}{{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
 The remaining $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
@@
 
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
+
+\caption{Conceptual framework showing how shade-tolerance (as measured by \textsc{wplcp}) is affected by plant size and traits, adapted from \citep{Givnish-1988}.
+From eqn \ref{eq:wplcp}, \textsc{wplcp} is the canopy openness $E$ where the photosynthetic income per unit leaf area (dashed line) intersects the total respiration and turnover costs per unit leaf area (solid black lines).
+Colors indicate the contributions from different tissues to total costs.
+Because the costs of stem tissues increase with height, \textsc{wplcp} also increases.
+Traits {\wood}, {\nitrogen}, and {\lma} can alter \textsc{wplcp} by shifting either income or cost of different tissues (Table \ref{tab:responses}).
+\label{fig:wplcp_idea}}
+\end{figure}
+
 The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
 
-The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood and roots (eqn \ref{eq:daldmt}).
-The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of tissue-level construction costs per unit of leaf area, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$.
+The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood, and roots (eqn \ref{eq:daldmt}).
+The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of construction cost per unit leaf area for different tissues, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$ (eqn \ref{eq:daldmt}).
 
 The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
-This variable accounts for the architectural strategy of the plant.
+This variable accounts for the architectural strategy of the plant (\citep{Falster-2011}).
 Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
 
-The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
+The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark, and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
 These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
@@
 We make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}-\ref{eq:htm}).
+
+\subsection{Shade tolerance}
+
+Eqn \ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'' \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
+In general, photosynthesis rate per leaf area $\bar{p}$ increases with canopy openness, $E$.
+The {\wplcp} is then the value $E=E^*$ where photosynthetic gains balance the costs of tissue turnover and respiration, i.e.
+$\frac{{\rm d}B}{{\rm d}t}(E^*)=0$.
+From eqn \ref{eq:dbdt}, this occurs when
+\begin{equation}\label{eq:wplcp}
+\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
+\end{equation}
+Graphically, the {\wplcp} occurs at the point where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig.
+\ref{fig:wplcp_idea}).
+Traits influence {\wplcp} when they affect either photosynthesis, respiration, or turnover.
+Also, as the amount of stem support increases with plant height, {\wplcp} also increases with height \citep{Givnish-1988} (Fig.
+\ref{fig:wplcp_idea}).
+
+\subsection{A functional-balance model for plant construction}
+
+Since eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication, and differentiation, they may apply to a variety of growth model where biomass allocation is important.
+However, to make explicit predictions requires that we quantify all the terms in Fig.
+\ref{fig:conceptual}b via an explicit model of plant construction and function.
+
+The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citep{Makela-1997} and \citep{Moorcroft-2001}.
+We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
+Table \ref{tab:allometry} provides key equations (see \citep{Falster-2016} for full derivation).
+In particular, we assume that as a plant grows:
+1) Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
+2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
+3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
+4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and
+5) The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
+Assumption 1 accounts for the architectural layout of the plant.
+Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
+To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see SI).
+
+Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
+These eqns aso predict that the amount of live stem tissue supporting each unit of leaf area should increase linearly with height,
+\begin{equation}\label{eq:msml}
+\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
+\end{equation}
+Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
+\ref{fig:conceptual}.
 
@@
 \end{tabular}
-\addtabletext{\\The first column of part \textbf{a} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of \textbf{a} and in \textbf{b} can then be derived from the assumptions in the left column of \textbf{a}.
+\addtabletext{\\The first column of \textbf{a)} provides cores assumptions between various size metrics and leaf area.
+Eqns in the middle and right columns of \textbf{a)} and in \textbf{b)} can be derived from these assumptions.
 The column ``Eqn'' indicates equation numbers referred to in the main text.
+
 }
@@
 
-\subsection{Shade tolerance}
-
-Eqn \ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'', \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
-In general, assimilation rate per leaf area $\bar{p}$ increases with light level, $E$.
-The {\wplcp} is light level at which a plant's photosynthetic gains just balance the costs of tissue turnover and respiration and can be estimated by solving for the the value $E=E^*$ giving $\frac{{\rm d}B} {{\rm d}t}(E)=0$ (Fig.
-\ref{fig:wplcp_idea}).
-From eqn \ref{eq:dbdt}, this occurs when
-\begin{equation}\label{eq:wplcp}
-\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
-\end{equation}
-The {\wplcp} occurs at the points where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig.
-\ref{fig:wplcp_idea}).
-Traits can influence the {\wplcp} if they affect either carbon uptake or costs (respiration, turnover).
-Also, since the amount of stem support increases with plant height, the {\wplcp} also naturally increases with height \citep{Givnish-1988} (Fig.
-\ref{fig:wplcp_idea}).
-
-\subsection{A functional-balance model for plant construction}
-
-It is worth noting that because eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication and differentiation, they hold for any potential growth model where biomass allocation is important.
-To make explicit predictions via this framework, an explicit model of plant construction and function is needed; i.e.
-we must quantify all the terms in Fig.
-\ref{fig:conceptual}b.
-
-The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citep{Makela-1997} and \citep{Moorcroft-2001}.
-We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
-Table \ref{tab:allometry} provides key equations (see \citep{Falster-2016} for full derivation).
-In particular, we assume that as a plant grows:
-1) Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
-2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
-3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
-4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and
-5) The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
-Assumption 1 accounts for the architectural layout of the plant.
-Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
-To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see SI).
-
-Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
-Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
-\ref{fig:conceptual}.
-
 \subsection{Trait-based trade-offs}
 
-For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}.
-It is essential that any trait includes both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values.
+To capture the effects of functional traits on growth rates and shade tolerance requires that trait-based trade-offs be embedded within the model described above.
+Here we consider five traits, for which we can posit specific physiological costs and benefits -- described in detail below and summarised in Table \ref{tab:traits}.
+It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values.
 In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait.
+
+\subsubsection{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
+
+Seed mass: {\seed} moderates a trade-off between a plant's rate of seed production per unit mass invested in reproduction and the mass of each offspring  \citep{Moles-2006}.
+
+Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
+We assume the amount allocated to vegetative growth starts high amd then decreases once $H > H_{{\rm mat}}$  (Fig.
+\ref{fig:conceptual}b).
+The exact nature of the function is non-critical, what is important is that plants shift from a period of investing mainly in vegetative growth to investing mainly in reproduction.
+Here we let the fraction of mass allocated to reproduction vary with $H$ via the function $r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
+
+\subsubsection{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.
+
+Leaf nitrogen per unit leaf area: We assume that both the maximum photosynthetic capacity of a leaf and leaf respiration rate both vary with {\nitrogen} (=$\nu$), as $\beta_{{\rm lf1}} \, \left(\nu / \nu_0 \right)^{\beta_{{\rm lf5}}}$ and $\beta_{{\rm lf4}}\, \nu$ respectively, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf4}}$, and $\beta_{{\rm lf5}}$ are constants.
+
+Leaf mass per unit leaf area: {\lma} (=$\phi$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
+The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever {\lma} is varied.
+
+Wood density: {\wood} (=$\rho$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
+In addition, we link {\wood} to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\rho / \rho_0\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
+The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of {\wood}, as such the mass-based rate is adjusted accordingly whenever {\wood} is varied.
+There currently little data relating sapwood turnover or respiration to {\wood}.
 
@@
 We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}.
-We also evaluated an additional prediction arising from the eqns in Table \ref{tab:allometry}a.
-The amount of live stem tissue supporting each unit of leaf area is predicted to increase linearly with height as
-\begin{equation}\label{eq:msml}
-\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
-\end{equation}
-Note that $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are all traits and so assumed approximately constant for any given species.
-
-\begin{figure}[!hb]
-\centering
-\includegraphics[width=1\linewidth]{main/assumptions.pdf}
-\caption{Assumptions of a functional balance model for plant construction.
-Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
-Blue lines show standardised major axis lines fit to different species.
-The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table \ref{tab:allometry}.
-}
-\label{fig:assumptions}
-\end{figure}
-
 Fig.
-\ref{fig:assumptions} shows that the three functional-balance assumptions outlined in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
-The solid blue lines indicate {\sma} lines fit to each species in the dataset.
+\ref{fig:assumptions} shows that the three functional-balance assumptions in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
 As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions (Table \ref{tab:allometry}a and eqn \ref{eq:msml}).
@@
 The panels in Fig.
-\ref{fig:conceptual} show the predicted patterns for a typical woody plant, obtained by applying the functional-balance model Table \ref{tab:allometry}a in {\plant}.
-Biomass production shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
+\ref{fig:conceptual} show the predicted patterns for a typical woody plant, obtained by applying the functional-balance model from Table \ref{tab:allometry}a.
+Biomass growth shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
 Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
@@
 \ref{fig:conceptual}), increasing reproductive allocation (Fig.
-\ref{fig:conceptual}), and declining mass production.
+\ref{fig:conceptual}), and declining biomass growth.
 In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
@@
 We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
-The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
-The remaining three traits are termed ``developmental traits'' because they moderate the rate of movement along an individual's growth trajectory.
 Across the five different traits, we observed four relatively distinct types of response.
@@
 
-\subsubsection{Ontogenetic traits}  Increasing $\omega$ in our model causes seedlings to be larger and fecundity to decrease.
-As such, the only effect of $\omega$ on growth comes from changing the plant's initial size.
+\subsubsection{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease.
+The only effect of {\seed} on growth comes from changing the plant's initial size.
 The plots in Fig.
 \ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
-Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates in all metrics because of their greater total leaf area.
-At the same time, relative growth rate is predicted to decrease with size, because the ratio of leaf area to support mass decreases with plant size.
-As plants grow, differences in initial mass will decrease in importance, relative to other factors influencing growth through the life-cycle.
-As a result, the correlations between $\omega$ and growth rate observed for seedlings disappear among larger plants.
+Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in all metrics) because of their greater total leaf area.
+At the same time, relative growth rate is predicted to decrease with size, as the ratio of leaf area to support mass decreases.
+As plants grow, differences in initial mass decrease in importance relative to other factors influencing growth through the life-cycle.
+As a result, the correlations between {\seed} and growth rate disappear among larger plants.
 
-Height at maturation, $H_{{\rm mat}}$, moderates growth by adjusting the amount of energy invested in growth, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
-Greater $H_{{\rm mat}}$ causes a growth advantage at larger sizes by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
+Greater {\hmat} causes a growth advantage at larger sizes by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
 \ref{fig:growth_light_height}).
@@
 
-\subsubsection{Developmental traits}  The response of growth rate to changes in $\nu$ is relatively straightforward: there is an optimum value of $\nu$ that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
+\begin{SCfigure*}[\sidecaptionrelwidth][b]
+\centering
+\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
+\caption{\textbf{Effect of three developmental traits on shade tolerance.}
+Panels show effect of traits on level of canopy openness that causes biomass growth (eqn \ref{eq:dbdt}) to be zero.
+Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
+The white regions indicate trait ranges that are typically observed in real systems.
+\label{fig:wplcp}}
+\end{SCfigure*}
+
+\subsubsection{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
 \ref{fig:growth_light_height}).
-As $E$ increases, the optimal $\nu$ also increases.
+As $E$ increases, the optimal {\nitrogen} also increases.
 The invariance of the growth-trait relationship with respect to size arises as follows.
-The direct physiological effect of $\nu$ is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
-Both the cost and benefits of $\nu$ appear within $\frac{ {\rm d}B} {{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in mass production per $\nu$.
-From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial nu} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial nu}  - \frac{ \partial r_{\rm l}}{\partial nu} )$.
-As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.
+The direct physiological effect of {\nitrogen} is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
+Both the cost and benefits of {\nitrogen} appear within the term $\frac{ {\rm d}B}{{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in biomass growth per {\nitrogen}.
+From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
+u} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial
+u}  - \frac{ \partial r_{\rm l}}{\partial
+u} )$. As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.
 
-Unlike $\nu$, the response of growth rate to changes in $\phi$ varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
+Unlike {\nitrogen}, the response of growth rate to changes in {\lma} varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
 \ref{fig:growth_light_height}).
-As a result, the value of $\phi$ that optimises plant growth increases with height, and the direction of correlation between height growth rate and $\phi$ shifts from negative to positive, as plants increase in height.
-Decreasing $\phi$ has two impacts on height growth rate.
-First, lower $\phi$ increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$.
-Second, lower $\phi$ decreases net production $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
-Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects.
-When plants are small, the effect on leaf deployment rate is larger and so decreasing $\phi$ increases growth rate.
-When plants are large, the influence of $\phi$ on leaf deployment per mass is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
-At larger sizes, low $\phi$ is thus no longer advantageous for growth (Fig.
+As a result, the value of {\lma} that optimises plant growth increases with height, while the direction of correlation between height growth rate and {\lma} shifts from negative to positive, as plants increase in height.
+Decreasing {\lma} has two impacts on height growth rate.
+First, lower {\lma} increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$.
+Second, lower {\lma} decreases biomass growth $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
+Whether lower {\lma} increases height growth depends on the relative magnitude of these two effects.
+When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate.
+When plants are large, the influence of {\lma} on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
+Thus, at larger sizes low {\lma} no longer increases growth rate (Fig.
 \ref{fig:growth_light_height}).
 
-Reducing the cost of stem construction via lower $\rho$ decreases the cost of deploying a unit of leaf area and thereby increasing growth rate (Fig.
+Reducing the cost of stem construction, via lower {\wood}, decreases the cost of deploying a unit of leaf area, increasing growth rate (Fig.
 \ref{fig:growth_light_height}).
-In contrast to $\phi$, the benefits of cheaper stem construction become more pronounced at intermediate sizes, as an increasingly large fraction of the plant is wood (Fig.
+Unlike {\lma}, the benefits of cheaper stem construction become more pronounced at intermediate sizes, as a larger fraction of the plant is stem (Fig.
 \ref{fig:conceptual}a).
@@
 \ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}.
-For each, plants were grown to a suitable diameter, area, or mass.
-Changes in relative growth rates with traits therefore show a similar pattern.
+For each, plants were grown to a standard diameter, area, or mass.
+Changes in absolute and relative growth rates with traits therefore show similar patterns.
 
-We find that for $\omega$, $\nu$, $\phi$, and $H_{\rm mat}$, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to height growth (Table \ref{tab:responses}).
-The only trait where a slightly different response was observed was for wood density.
-Whereas the effect of wood density on height growth tended to diminish slightly at larger sizes (Fig.
-\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area or above-ground mass.
-Sapwood lost via turnover is turns into heartwood.
+We find that for {\seed}, {\nitrogen}, {\lma}, and {\hmat}, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to height growth (Table \ref{tab:responses}).
+The only trait where a slightly different response was observed was for {\wood}.
+Whereas the effect of {\wood} on height growth tended to diminish slightly at larger sizes (Fig.
+\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area, or above-ground mass.
+Sapwood lost via turnover becomes heartwood.
 Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.
-
-\begin{SCfigure*}[\sidecaptionrelwidth][b]
-\centering
-\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three developmental traits on shade tolerance.}
-Panels show effect of traits on level of canopy openness that causes net production (eqn \ref{eq:dbdt}) to be zero.
-Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
-The white regions indicate trait ranges that are typically observed in real systems.
-\label{fig:wplcp}}
-\end{SCfigure*}
 
@@
 This response arises because individuals with higher potential growth rate naturally have greater potential plasticity in growth.
-Our results therefore support findings that species with low $\rho$ increase growth more substantially with increases in light (Table \ref{tab:phenomena}).
-Variation in $\phi$ also moderates the response of growth to changes in light, with species having the lowest $\phi$ being most responsive.
-However, unlike for $\rho$, the effect appears only for the smallest size classes.
+Our results therefore support findings that species with low {\wood} increase growth more substantially with increases in light (Table \ref{tab:phenomena}).
+Variation in {\lma} also moderates the response of growth to changes in light, with species having the lowest {\lma} being most responsive.
+However, unlike for {\wood}, the effect appears only for the smallest size classes.
 
@@
 where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
-Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with $\nu$, $\phi$, and $\rho$.
-With some further manipulations, it is possible to show that {\wplcp} will decrease with $\phi$ if $\beta_{{\rm kl2}} >1$.
-Likewise {\wplcp} will decrease with $\rho$ if $\beta_{{\rm ks2}} >1$.
-The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to $\phi$ and $\rho$, respectively.
-Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low $\phi$ and low $\rho$ are predicted to be less shade tolerant (Fig.
+Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}.
+With some further manipulations, it is possible to show that {\wplcp} will decrease with {\lma} if $\beta_{{\rm kl2}} >1$.
+Likewise {\wplcp} will decrease with {\wood} if $\beta_{{\rm ks2}} >1$.
+The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to {\lma} and {\wood}, respectively.
+Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low {\lma} and low {\wood} are predicted to be less shade tolerant (Fig.
 \ref{fig:wplcp}).
-At low $\phi$ ($\rho$), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover.
+At low {\lma} ({\wood}), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover.
 {\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
-In addition, {\wplcp} varies with $\nu$.
-At small sizes, {\wplcp} increases with $\nu$ across the band of values typically observed in real plants, i.e.
-high leaf nitrogen makes seedlings shade intolerant.
-At larger sizes, as net production declines to zero, {\wplcp} begins to increase again for very low values of $\nu$.
+In addition, {\wplcp} varies with {\nitrogen}.
+At small sizes, {\wplcp} increases with {\nitrogen} across the band of values typically observed in real plants, i.e.
+high {\nitrogen} makes seedlings shade intolerant.
+At larger sizes, as biomass growth declines to zero, {\wplcp} begins to increase again for very low values of {\nitrogen}.
 All these results  match empirical patterns (Table \ref{tab:phenomena}).
@@
 
-\begin{table}[t!]
+\begin{table}[hb!]
 \centering
@@
   & &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
-  & & \boldmath$\omega$ & \boldmath$H_{{\rm mat}}$ & \boldmath$\nu$ & \boldmath$\phi$ & \boldmath$\rho$ & \\
+  & & {\seed} & $H_{{\rm mat}}$ & {\nitrogen} & {\lma} & {\wood} & \\
   \midrule
   \multicolumn{8}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
-  \multicolumn{2}{l} {Net biomass production, $\ud B / \ud t$} & & & & & & \\
+  \multicolumn{2}{l} {Biomass growth, $\ud B / \ud t$} & & & & & & \\
   & \tabitem Photosynthesis & - & - & $\uparrow$  & - & - & \\
@@
   \end{tabular}
-\addtabletext{Arrows indicate the effect of increased trait value on each component; dashes indicating no effect. See Table \ref{tab:traits} for trait definitions.}
+\addtabletext{Arrows indicate the effect of increased trait value on each component; dashes indicating no effect.
+Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density (see Tables \ref{tab:traits}-\ref{tab:definitions}).
+
+}
 \label{tab:responses}
@@
 
-Using a model relating plant physiological function and carbon allocation to five prominent traits, we have shown how traits impact on plant growth across the life cycle.
-This approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait leaf mass per unit leaf area \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
-During the last two decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}.
+Using a model relating plant physiological function and mass allocation to prominent traits, we have clarified how traits could impact on plant growth across the life cycle.
+Our approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
+During recent decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}.
 However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}.
-Matching a growing amount of empirical evidence (Table \ref{tab:phenomena}), this study outlines when and why the direction or strength of correlation between traits and growth rate shifts with plant size.
+Matching growing empirical evidence (Table \ref{tab:phenomena}), this study shows when and why the direction or strength of a correlation between traits and growth rate shifts with plant size.
 Moreover, we show that different traits and trade-offs generate different types of response.
-Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life (from slow to fast growth) \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
+Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
 Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e.
 lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}.
-The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
+The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter, and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
 This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}.
 Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects.
-In this sense, our model is designed to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
+Our aim is to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
 
-\subsection{Generalising to other traits and trade-offs}
+\subsection{LMA doesn't predict growth rates of large plants}
 
-The model presented here extends a widely-used model for seedlings, linking mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
-Importantly, the seedling model can be derived as a special case of the new-extended model (eqn \ref{eq:RGR}).
-Unlike the original, the extended model also predicts a relationship between $\phi$ and growth rate that changes with size.
-In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig.
+The model presented here extends a widely-used model for seedlings, linking mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
+Importantly, the seedling model (eqn \ref{eq:RGR}) can be derived as a special case of the new, extended model.
+Unlike eqn \ref{eq:RGR}, the extended model also predicts a relationship between {\lma} and growth rate that changes with size.
+In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even weakly positive in larger plants (Fig.
 \ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem diameter, stem area or total mass.
@@
 \ref{fig:conceptual}a).
-Decreasing $\phi$ then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
+Decreasing {\lma} then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
 As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheap leaf construction.
-Consequently, the effect of $\phi$ on leaf turnover comes to dominate at larger sizes, and as such, the effect of $\phi$ on height, diameter and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
+Consequently, the effect of {\lma} on leaf turnover comes to dominate at larger sizes, and as such, the effect of {\lma} on height, diameter, and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
+
+\subsection{Generalising to other traits and trade-offs}
 
 The list of functional traits known to differ among plant species is ever-increasing \citep{Perez-2013}.
-While we have focussed on understanding the effects of five traits on growth rate, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
-The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be embedded within the eqns \ref{eq:dbdt}--\ref{eq:dDdt}.
+While we have focussed on understanding the effects of five traits on elements of demography, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
+The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be connected with variables in eqns \ref{eq:dbdt}--\ref{eq:dDdt}.
 While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
@@
 
-Our framework also highlights what is needed for traits to impact on growth rate and shade tolerance.
+Our framework also highlights what is needed for traits to influence demography.
 While traits can influence many aspects of plant function, these influences must operate via the pathways outlined in Fig.
-\ref{fig:conceptual} if the trait is going to impact on growth.
-For example, many studies have focused on traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
-These traits will inevitably influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}) by altering conductance of water to the leaf.
-The potential costs of larger vessel size might be higher rates of stem turnover, which would appear in the term $k_{\rm s}$ in eqn \ref{eq:dbdt}.
+\ref{fig:conceptual} if the trait is going to impact on growth rate or shade tolerance.
+For example, many studies consider traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
+By altering conductance of water to the leaf, both traits will influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}), while vessel size might also affect rates of stem turnover ($k_{\rm s}$ in eqn \ref{eq:dbdt}).
 The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in via the parameter $\theta$ (Table \ref{tab:allometry}).
@@
 
-Our results highlight the importance of allocation decisions and turnover costs in determining growth dynamics.
+First, our results highlight the importance of allocation decisions and turnover costs in determining demography.
 Much of current ecosystem research focusses on factors affecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
-Yet, for four of the five traits considered here, trait values do not influence net primary production.
+Yet, four of the five traits considered here do not influence net primary production.
 In fact, our analysis shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
-Low $\phi$ results in high leaf turnover, such that individuals with a low $\phi$ have lower mass production.
+Low {\lma} results in high leaf turnover, such that individuals with a low {\lma} have lower biomass growth.
 It is this property that makes them shade intolerant.
@@
 
-Second, our results demand a shift in the simple division of slow or fast growing plant species.
-To the extent that the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum from slow to fast growth.
+Second, our results demand a shift away from viewing species as lying on a spectrum  from slow to fast growth.
+When the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum.
 Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.
-Researchers using functional traits have also tended to describe species as fast or slow growing \citep[e.g.][]{Adler-2014, Diaz-2016}.
+Researchers using functional traits have also tended to describe species as fast or slow growing \citep{Adler-2014, Diaz-2016}.
 Our results suggest a more-nuanced approach.
@@
 Plants that are fast growing as adults may not be fast growing as seedlings.
-This more-nuanced perspective tends to mirror observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
+This more-nuanced perspective reflects observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
 
@@
 As such, we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits \citep[e.g.][]{Poorter-2008,Paine-2015}.
-Controlling for size, site and light environment will be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}, as will having a clear expectation for the hypothesised relationship.
+Controlling for size, site, and light environment will be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}, as will having a clear expectation for the hypothesised relationship.
+
+\subsection{Plasticity of traits through ontogeny}
 
 While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
-To that end, we make a further prediction arising from our results, which is that the trait $\phi$ should increase through ontogeny for all individuals.
+To that end, we make a further prediction arising from our results, which is that the trait {\lma} should increase through ontogeny for all individuals.
 Such shifts have been observed across a variety of species \citep{King-1999,Thomas-1999,Koch-2004}.
-Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low $\phi$ -- to increase their $\phi$ as they grow larger.
+Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low {\lma} -- to increase their {\lma} as they grow larger.
 While a similar prediction was made for a species of \emph{Eucalyptus} \citep{King-1999}, we extend the idea across species.
@@
 A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs.
-While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have mainly focussed on ecosystem-level outcomes rather than the growth of individual plants.
+While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have focussed on ecosystem-level outcomes rather than the growth of individual plants.
 Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
-Another study \citep{King-1999} connected a model of growth for a single species to the trait $\phi$, and likewise predicted a gradual flattening out of the relationship between $\phi$ and growth rate with size (as in Fig.
+Another study \citep{King-1999} connected a model of growth for a single species to the trait {\lma}, and likewise predicted a gradual flattening out of the relationship between {\lma} and growth rate with size (as in Fig.
 \ref{fig:growth_light_height}).
-Here we added an additional cost of increased leaf turnover, that further penalised low $\phi$ strategies when large.
+Here, we include an additional cost of increased leaf turnover that further penalised low {\lma} strategies when large.
 
 Perhaps the two most controversial elements of our approach concerns the assumptions about tissue replacement and reproductive allocation.
-Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing surplus biomass.
-This is because we assume tissues lost via turnover are replaced before carbon is allocated to either new growth (i.e.
+Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing any surplus.
+This is because we assume tissues lost via turnover are replaced before mass is allocated to either new growth (i.e.
 growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
 This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant.
-The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus carbon to reproduction.
+The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus mass to reproduction.
 While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
 Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}.
-Reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
+Currently, reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
 For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
@@
 
-Another class of model dealing explicitly with size-related effects includes those derived from the metabolic scaling theory (\textsc{mst}) of ecology \citep{Enquist-1999, Enquist-2007}.
+Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory  of ecology (\textsc{mst}) \citep{Enquist-1999, Enquist-2007}.
 Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
-First, the \textsc{mst}-derived model suggests diameter growth continues to increase as plants grow, whereas empirical data suggests growth rate declines for larger plants \citep{Canham-2004, Canham-2006, Herault-2011}.
+First, the \textsc{mst} model suggests diameter growth continues to increase wit size, whereas empirical data suggests growth rate declines for larger plants \citep{Canham-2004, Canham-2006, Herault-2011}.
 Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size.
-Predicted effects are for a linear increase in growth with decreases in either $\phi$ and lower $\rho$, that apply irrespective of size.
-However, at least for $\phi$ such effects in large trees have not been observed.
+Predicted effects are for a linear increase in growth with decreases in either {\lma} and lower {\wood}, that apply irrespective of size.
+However, at least for {\lma} such effects in large trees have not been observed.
 
@@
 \vspace{-1em}
-\subsection{Trait-based trade-offs}
-The trade-offs described in Table \ref{tab:traits} were embedded within the growth model as follows.
-
-We assume a direct trade-off between a plant's fecundity and $\omega$, with larger seeds resulting in larger seedlings.
-
-$H_{{\rm mat}}$ moderates an inevitable energetic trade-off between growth and reproduction that operates at all times through the lifestyle.
-To describe how the fraction of mass allocated to reproduction changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction.
-
-We allow for the maximum photosynthetic capacity of the leaf to vary with $\nu$ as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\nu / \nu_0 \right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants.
-Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
-
-The trait $\phi$ influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-The rate of leaf respiration per unit area is assumed independent of $\phi$\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
-
-The trait $\rho$ directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\rho / \rho_0\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
-The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly whenever $\rho$ is varied.
-There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is somewhat speculative.
 
@@
 Parameters were sourced mainly from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at $H_{{\rm mat}}$ (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
+The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at {\hmat} (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
 \ref{fig:conceptual}).
@@
 The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity to in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
+To encode the trait-based trade-offs described above, we use the capacity in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
 
@@
 We thank J Camac, A Gibert, G Kunstler, H Muller-Landau, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
-DF was supported by an Australian Research Council discovery grant (DP110102086).
+DF was supported by fellowships from the  Australian Research Council discovery (DP110102086, FT160100113).
 RF was supported by the Science and Industry Endowment Fund (RP04-174).

### 2018-04-05 R/figures.R

@@ -297,15 +297,16 @@ figure_rate_vs_size_panels <- function(data, type, path) {
   for (v in yvars) {
     filename <- file.path(path, sprintf("%s_%s.pdf", type, v))
     pdf(filename,width=5, height=6)
-    par(oma=c(0,0,0,0), mar=c(6, 0.1,0.1,0.1))
+    par(oma=c(0,0,0,0), mar=c(7, 0.1,0.1,0.1))
     plot(data[["height"]], data[[v]], type="l", ann=FALSE, axes=FALSE,
-       xlim=c(0, 25),
+       xlim=c(-1, 25),
 #       xaxs="i", yaxs = "i",
-      col=get_col(v), lwd=8)
+      col=get_col(v), lwd=20)
     if(type == "diameter_stem_dt" || (type == "area_stem_dt"  && v %in% c( "area_leaf_dt", "area_heartwood_dt"))){
-      axis(1, at=seq(0,30, by=10), tck=-0.05, cex.axis=4, mgp=c(3, 4, 0))
+      f <- function(at) axis(1, at=at, tck=-0.05, col="grey30", col.ticks = "grey30", cex.axis=7, mgp=c(3, 6, 0))
+      f(0); f(10); f(20);
     }
-    box()
+    box(col="grey30")
     dev.off()
   }
 }
@@ -677,15 +678,18 @@ figure_lcp_schematic <- function() {
   plot(NA, xlab = "", ylab="", xaxs="i", yaxs="i",
       xlim=c(0,1), ylim=c(0, 2.5), las=1)
   mtext("Canopy openness, E (0-1)", 1, line=3)
-  mtext(expression(paste("Mass production or loss (kg ", m^-2~yr^-1, ")")), 2, line=3)
+  mtext(expression(paste("Income or loss (kg ", m^-2~yr^-1, ")")), 2, line=3)
 
   # colour background
   rect(0, 0, 1, 2.5, col = cols["bark"], border=NA)
 
   # leaf and root for smallest size
-  x <- data[[1]]$all
-  rect(0, 0, 1, x[,"leaf"], col = cols["leaf"], border=NA)
-  rect(0, x[,"leaf"], 1, x[,"root"]+x[,"leaf"], col = cols["root"], border=NA)
+
+  rect(0, 0, 1, data[[1]]$all$leaf, col = cols["leaf"], border=NA)
[…]

### 2018-04-05 R/talk.R

@@ -19,7 +19,7 @@ get.pic <- function(file) {
 setup_viewport <- function(bottom, right, top){
   #gridBase  - works out coordinates of different plots
   vps <- baseViewports()
-  pushViewport(vps$inner)  #put current plot onto global viewport stack
+  pushViewport(vps$inner)  #put current inner onto global viewport stack
   #make new viewport
   vp <- viewport(x=unit(1, "npc") - unit(right, "inches"),
                  y=unit(bottom, "inches"),

### 2018-04-05 build.R

@@ -1,0 +1,78 @@
+library("baad.data")
+library("rmarkdown")
+library("plant")
+library("xtable")
+library("RColorBrewer")
+library("dplyr")
+library("knitr")
+library("parallel")
+library("rstan")
+source("R/axis.R")
+source("R/build.R")
+source("R/figures_data.R")
+source("R/figures.R")
+source("R/stan.R")
+source("R/tables.R")
+source("R/talk.R")
+source("R/utils-plot.R")
+source("R/utils.R")
+source("R/wright_2004.R")
+dir.create("downloads", FALSE, TRUE)
+dir.create("output/main", FALSE, TRUE)
+dir.create("output/SI", FALSE, TRUE)
+dir.create("output/conn", FALSE, TRUE)
+dir.create("output/talk", FALSE, TRUE)
+download.file("http://www.nature.com/nature/journal/v428/n6985/extref/nature02403-s2.xls", "downloads/wright_2004.xls")
+fit_a.lf_h.t <- fit_SMA_baad("a.lf", "h.t")
+fit_a.lf_m.rf <- fit_SMA_baad("a.lf", "m.rf")
+fit_a.lf_sap <- fit_SMA_baad("a.lf", "sap")
+fit_h.t_msal <- fit_SMA_baad("h.t", "msal")
+library("remake")
+make_script(file = "build.R")
+height_dt_data <- figure_height_dt_data()
+diameter_stem_dt_data <- figure_diameter_stem_dt_data()
+area_stem_dt_data <- figure_area_stem_dt_data()
+mass_above_ground_dt_data <- figure_mass_above_ground_dt_data()
+pdf("output/main/lcp_schematic.pdf", width = 6L, height = 4L)
+figure_lcp_schematic()
+dev.off()
+lcp_trait_data <- figure_lcp_trait_data()
+rate_vs_size_data <- figure_rate_vs_size_data()
+knitr::knit("ms/ms-supp.Rnw", "ms/ms-supp.tex")
+library("readxl")
+glopnet <- process_wright_2004("downloads/wright_2004.xls")
+pdf("output/main/assumptions.pdf", width = 9L, height = 9L)
+figure_assumptions(fit_a.lf_h.t, fit_a.lf_sap, fit_a.lf_m.rf, fit_h.t_msal)
+dev.off()
+fit_lma_tradeoff <- fit_SMA_lma(glopnet)
+pdf("output/SI/lma_tradeoff.pdf", width = 6L, height = 6L)
[…]

### 2018-04-15 R/talk.R
Additional figure for talks

@@ -76,3 +76,55 @@ figure_dY_dt_talk <- function(v, i, dat, tree) {
   grid.text(dat[["label"]](H), x=unit(.5, "npc"), y=unit(1.0, "npc"), just=c("center", "bottom"))
   grid.picture(tree, x=unit(.5, "npc"), y=unit(0, "npc"), just=c("center", "bottom"), height=unit(0.2+0.8*H/Hmax, "npc"))
  }
+
+figure_dY_dt_talk2 <- function(dat) {
+
+  vars <- dat[["vars"]]
+  sizes <- dat[["sizes"]][1]
+  E <- dat[["E"]]
+  traits <- dat[["traits"]][2:4]
+
+  cols <- color_pallete1(length(E))
+  par(mfrow=c(1, length(traits)),
+      oma=c(4, 5, 0, 1.5), mar=c(1, 1, 1, 0.9))
+
+  extract_f <- function(v, i, dat) {
+      dat_v <- unname(split(dat$data[[v]], dat$data[[v]]$class))
+      dsub <- long_to_wide(dat_v[[i]], "canopy_openness",
+                           c(v, vars[2]))
+  }
+
+  for (i in seq_along(sizes)) {
+
+    dat2 <- lapply(traits, extract_f, i, dat)
+    names(dat2) <- traits
+    ymax <- max(sapply(dat2, function(x) max(x[[2]],  na.rm=TRUE)))
+    for (v in traits) {
+      dsub <- dat2[[v]]
+      matplot(dsub[[1]], dsub[[2]], type="n", log="x",
+              xaxt="n", yaxt="n", xlab="", ylab="",
+              ylim = c(0, ymax*1.05))
+      usr <- par("usr")
+#      rect(1E-5, -1, 1E4, 2*ymax, col = make_transparent("grey", 0.3), border=NA)
+      obs <- trait_range_obs(v)
+      rect(obs[1], -1, obs[2], 2*ymax, col = "white", border=NA)
+      box()
+      matplot(dsub[[1]], dsub[[2]], type="l", col=cols, lty=1, add=TRUE)
+
+      if(v== traits[1])
+        mtext(name_pretty(vars[2]), line=3, side=2, cex=0.9 , outer=FALSE)
+      # axes
+      axis.log10(1, labels = (i == 1), las=1)
+      axis(2, labels = (v == traits[1] ), las=1, at = seq(0, 20, by=get_by(ymax)))
[…]

### 2018-04-15 ms/ms-supp.Rnw
Minor modifications for submission

@@
 
-\input{ms-common-defs}
+% abbreviations
+\newcommand{\smurl}[1]{\url{#1}}
+\newcommand{\ud}{\ensuremath{\rm{d}}}
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\lma}{\textsc{lma}}
+\newcommand{\wood}{\textsc{wd}}
+\newcommand{\seed}{\textsc{sm}}
+\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
+\newcommand{\nitrogen}{\textsc{n}}
+\newcommand{\sma}{\textsc{sma}}
+
+\usepackage{graphicx}
+\graphicspath{{../output/SI/}}
+
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Daniel S. Falster}
+\author[b]{Remko A. Duursma}
+\author[c]{Richard G. FitzJohn}
+
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
+Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
+\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
+\affil[c]{Imperial College, London, United Kingdom}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Falster}
 
@@
 }
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \renewcommand{\thesection}{S\arabic{section}}
-
-% Figure command -- also adds figure to table of contents
-\newcommand{\myfigure}[4]{
-    \begin{figure*}[!hb]
-            \centering
-             \includegraphics[width=0.9\textwidth]{#1}
-            \caption{\textbf{#2} #3}
-            \label{#4}
-    \end{figure*}
-}
 
@@
 
-Values of $p(\nu, E)$ are calculated across a range of values of $E$, and then an expression of the form in Eq. \ref{eq:photosynthesis} is fitted to extract the parameters
-$\alpha_{{\rm p1}}$ and $\alpha_{{\rm p2}}$, such that these become functions of $\nu$.
-
 \subsubsection{Leaf mass per unit area}
@@
 The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$.
-In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citep{Wright-2004},
+In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citep{Wright-2004} (Fig.
+\ref{fig:leafturn}),
 \begin{equation}
@@
 
-<<echo=FALSE,results='asis'>>=
-tab <- table_plant_parameters("../data/growth_model_pars_core.csv")
-print(xtable(tab,
-            hline.after=c(1),
-            align='lp{7cm}llll',
-            label="tab:params_core",
-            caption="Core parameter of the \texttt{FF16} physiological model."),
-      floating.environment="table*",
-      sanitize.text.function=I,
-      include.rownames=FALSE, 
-      caption.placement = "top")
-@
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Thu Apr  5 13:08:59 2018
+\begin{table*}[ht]
+\centering
+\caption{Core parameter of the 	exttt{FF16} physiological model.} 
+\label{tab:params_core}
+\begin{tabular}{p{7cm}llll}
+  \hline
+Description & Symbol & Unit & Code & Value \\ 
+  \hline
+\textbf{Plant construction} &  &  &  &  \\ 
+  Crown-shape parameter & $\eta$ &  & \texttt{eta} & 12 \\ 
+  Leaf mass per area & $\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{lma} & 0.1978791 \\ 
+  Wood density & $\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & \texttt{rho} & 608 \\ 
+  Sapwood area per unit leaf area & $\theta$ &  & \texttt{theta} & 0.0002141786 \\ 
+  Height of plant with leaf area of 1m$^2$ & $\alpha_{\rm l1}$ & m & \texttt{a\_l1} & 2.17 \\ 
+  Exponent of relationship between height and leaf area & $\alpha_{\rm l2}$ &  & \texttt{a\_l2} & 0.5 \\ 
+  Root mass per unit leaf area & $\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{a\_r1} & 0.07 \\ 
+  Ratio of bark area to sapwood area & $\alpha_{\rm b1}$ &  & \texttt{a\_b1} & 0.17 \\ 
+  \textbf{Production} &  &  &  &  \\ 
+  Leaf photosynthesis per area & $\alpha_{\rm p1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-2}$ & \texttt{a\_p1} & 151.1778 \\ 
+  Saturation of leaf photosynthesis per area & $\alpha_{\rm p2}$ &  & \texttt{a\_p2} & 0.2047162 \\ 
+  Yield = fraction of carbon fixed converted into mass & $\alpha_{\rm y}$ &  & \texttt{a\_y} & 0.7 \\ 
+  Biomass per mol carbon & $\alpha_{\rm bio}$ & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & \texttt{a\_bio} & 0.0245 \\ 
+  Leaf respiration per mass & $r_{\rm l}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_l} & 198.4545 \\ 
+  Fine-root respiration per mass & $r_{\rm r}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_r} & 217 \\ 
+  Sapwood respiration per mass & $r_{\rm s}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_s} & 6.598684 \\ 
+  Bark respiration per mass & $r_{\rm b}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_b} & 13.19737 \\ 
+  Turnover rate for leaves & $k_{\rm l}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_l} & 0.1521952 \\ 
+  Turnover rate for sapwood & $k_{\rm s}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_s} & 0.2 \\ 
+  Turnover rate for bark & $k_{\rm b}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_b} & 0.2 \\ 
+  Turnover rate for fine roots & $k_{\rm r}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_r} & 1 \\ 
+  \textbf{Fecundity} &  &  &  &  \\ 
+  Seed mass & $\omega$ & kg & \texttt{omega} & 0.000038 \\ 
+  Height at maturation & $H_{{\rm mat}}$ & m & \texttt{hmat} & 15 \\ 
+  Maximum allocation to reproduction & $\alpha_{\rm f1}$ &  & \texttt{a\_f1} & 0.8 \\ 
+  Parameter determining rate of change in $r(x, m_\textrm{l})$ around $H_{\rm mat}$ & $\alpha_{\rm f2}$ &  & \texttt{a\_f2} & 10 \\ 
+   \hline
+\end{tabular}
+\end{table*}
 
-<<echo=FALSE,results='asis'>>=
-tab <- table_plant_parameters("../data/growth_model_pars_hyper.csv")
-print(xtable(tab,
-            hline.after=c(1),
-            align='lp{9cm}llll',
-            label="tab:params_hyper",
-            caption="Parameters for hyper-parameterisation of the \\texttt{FF16} physiological model."),
-      sanitize.text.function=I,
-      floating.environment="table*",
-      include.rownames=FALSE, 
-      caption.placement = "top")
-@
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Thu Apr  5 13:08:59 2018
+\begin{table*}[ht]
+\centering
+\caption{Parameters for hyper-parameterisation of the \texttt{FF16} physiological model.} 
+\label{tab:params_hyper}
+\begin{tabular}{p{9cm}llll}
+  \hline
+Description & Symbol & Unit & Code & Value \\ 
+  \hline
+\textbf{Leaf turnover} &  &  &  &  \\ 
+  Global average leaf mass per area & $\phi_0$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{lma\_0} & 0.1978791 \\ 
+  Rate of leaf turnover at average leaf mass per unit leaf area, $\phi_0$ & $\beta_{\rm kl1}$ & $\mathrm{yr}^{-1}$ & \texttt{B\_kl1} & 0.4565855 \\ 
+  Scaling exponent for $\phi$ in leaf turnover & $\beta_{\rm kl2}$ &  & \texttt{B\_kl2} & 1.71 \\ 
+  \textbf{Sapwood turnover} &  &  &  &  \\ 
+  Global average wood density & $\rho_0$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & \texttt{rho\_0} & 608 \\ 
+  Rate of sapwood turnover at average wood density, $\rho_0$ & $\beta_{\rm ks1}$ & $\mathrm{yr}^{-1}$ & \texttt{B\_ks1} & 0.2 \\ 
+  Scaling exponent for $\rho$ in sapwood turnover & $\beta_{\rm ks2}$ &  & \texttt{B\_ks2} & 1.25 \\ 
+  \textbf{Photosynthesis} &  &  &  &  \\ 
+  Leaf nitrogen per unit leaf area & $\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{narea} & 0.00187 \\ 
+  Global average nitrogen per unit leaf area & $\nu_0$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{narea\_0} & 0.00187 \\ 
+  Potential CO$_2$ photosynthesis at average leaf nitrogen, $\nu_0$ & $\beta_{\rm lf1}$ & $\mathrm{mol}\,\mathrm{d}^{-1}\,\mathrm{m}^{-2}$ & \texttt{B\_lf1} & 0.8273474 \\ 
+  Curvature of light response curve & $\beta_{\rm lf2}$ &  & \texttt{B\_lf2} & 0.75 \\ 
+  Quantum yield of leaf photosynthesis (CO$_2$ per unit photosynthetically active radiation) & $\beta_{\rm lf3}$ &  & \texttt{B\_lf3} & 0.04 \\ 
+  Scaling exponent for leaf nitrogen in maximum leaf photosynthesis & $\beta_{\rm lf5}$ &  & \texttt{B\_lf5} & 0.5 \\ 
+  \textbf{Respiration} &  &  &  &  \\ 
+  CO$_2$ respiration per unit leaf nitrogen & $\beta_{\rm lf4}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{B\_lf4} & 15750 \\ 
+  CO$_2$ respiration per unit sapwood volume & $\beta_{\rm rs1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-3}$ & \texttt{B\_rs1} & 4012 \\ 
+  CO$_2$respiration per unit bark volume & $\beta_{\rm rb1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-3}$ & \texttt{B\_rb1} & 8024 \\ 
+   \hline
+\end{tabular}
+\end{table*}
+
+\newpage
+\begin{figure*}[!hb]
+\centering
+\includegraphics[width=0.9\textwidth]{lma_tradeoff.pdf}
+\caption{\textbf{Leaf turnover decreases with leaf-mass per unit leaf area.}
+Data from \citep{Wright-2004} for 678 species from 51 sites, each
+point giving a species-average. Lines show standardised major axis lines
+fitted to data from each site, with intensity of shading adjusted
+according to strength of the relationship.}
+\label{fig:leafturn}
+\end{figure*}
+\newpage
+
+\begin{figure*}[!hb]
+\centering
+\includegraphics[width=0.9\textwidth]{assumptions.pdf}
+\caption{\textbf{Assumptions of a functional balance model for plant construction.}
+Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
+Blue lines show standardised major axis lines fit to different species.
+The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table 4 from main text.}
+\label{fig:assumptions}
+\end{figure*}
 
@@
 
-\myfigure{SI/lma_tradeoff.pdf}{Leaf turnover decreases with leaf-mass per unit leaf area.}{Data from \citep{Wright-2004} for 678 species from 51 sites, each
-point giving a species-average. Lines show standardised major axis lines
-fitted to datsa from each site, with intensity of shading adjusted
-according to strength of the relationship.}{fS-leaf}
+\begin{figure*}[!hb]
+\centering
+\includegraphics[width=0.9\textwidth]{growth_light_dia.pdf}
+\caption{\textbf{Effect of four traits on stem-diameter growth rate for different-sized plants.}
+Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
+Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
+For any given value of trait and $E$, plants were grown to the desired stem-diameter and their growth rate estimated.
+Changes in trait-growth relationships are summarised in Table 5.}
+\label{fig:growth_light_dia}
+\end{figure*}
 
@@
 
-\myfigure{main/assumptions.pdf}{Assumptions of a functional balance model for plant construction.}{Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
-Blue lines show standardised major axis lines fit to different species.
-The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table \ref{tab:allometry}.}{fig:assumptions}
+\begin{figure*}[!hb]
+\centering
+\includegraphics[width=0.9\textwidth]{growth_light_area.pdf}
+\caption{\textbf{Effect of four traits on stem-area growth rate for different-sized plants.}
+Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
+Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
+For any given value of trait and $E$, plants were grown to the desired stem-area and their growth rate estimated.
+Changes in trait-growth relationships are summarised in Table 5.}
+\label{fig:growth_light_area}
+\end{figure*}
 
@@
 
-\myfigure{SI/growth_light_dia.pdf}{Effect of four traits on stem-diameter growth rate for different-sized plants.}{
-Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
-Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
-For any given value of trait and $E$, plants were grown to the desired stem-diameter and their growth rate estimated.
-Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_dia}
-
-\newpage
-
-\myfigure{SI/growth_light_area.pdf}{Effect of four traits on stem-area growth rate for different-sized plants.}{
-Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
-Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
-For any given value of trait and $E$, plants were grown to the desired stem-area and their growth rate estimated.
-Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_area}
-
-\newpage
-
-\myfigure{SI/growth_light_mass.pdf}{Effect of four traits on plant mass growth rate for different-sized plants.}{
+\begin{figure*}[!hb]
+\centering
+\includegraphics[width=0.9\textwidth]{growth_light_mass.pdf}
+\caption{\textbf{Effect of four traits on plant mass growth rate for different-sized plants.}
 Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3.
@@
 For any given value of trait and $E$, plants were grown to the desired mass and their growth rate estimated.
-Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_mass}
+Changes in trait-growth relationships are summarised in Table 5.}
+\label{fig:growth_light_mass}
+\end{figure*}
 

### 2018-04-15 ms/ms.tex
Minor modifications for submission

@@
 
-\input{ms-common-defs}
+\usepackage{graphicx}
+\graphicspath{{../output/main/}}
+
+% abbreviations
+\newcommand{\smurl}[1]{\url{#1}}
+\newcommand{\ud}{\ensuremath{\rm{d}}}
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\lma}{\textsc{lma}}
+\newcommand{\wood}{\textsc{wd}}
+\newcommand{\seed}{\textsc{sm}}
+\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
+\newcommand{\nitrogen}{\textsc{n}}
+\newcommand{\sma}{\textsc{sma}}
+
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Daniel S. Falster}
+\author[b]{Remko A. Duursma}
+\author[c]{Richard G. FitzJohn}
+
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
+Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
+\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
+\affil[c]{Imperial College, London, United Kingdom}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Falster}
 
@@
 % Enable cross referencing to Evidence document
-\usepackage{xr}
-\externaldocument{ms-supp}
-\externaldocument{figure_conceptual}
+% \usepackage{xr}
+% \externaldocument{ms-supp}
+% \externaldocument{figure_conceptual}
 
@@
 \centering
-\includegraphics[width=\linewidth]{main/figure_conceptual.pdf}
+\includegraphics[width=\linewidth]{figure_conceptual.pdf}
 \caption{Conceptual framework linking growth rate to plant size and traits.
@@
 
-\section*{Modeling the effects of traits on demography}
+\section*{A framework for understanding the effects of traits on growth rate and shade tolerance}
 
@@
 
+\refstepcounter{equation}\label{eq:dbdt_2}
+\refstepcounter{equation}\label{eq:dMtdt}
+\refstepcounter{equation}\label{eq:daldmt}
+\refstepcounter{equation}\label{eq:daldt}
+\refstepcounter{equation}\label{eq:dhdt}
+\refstepcounter{equation}\label{eq:dast}
+\refstepcounter{equation}\label{eq:dDdt}
+
 % Update equation counter, following inclusion of figure 1
-\setcounter{equation}{9}
+% \setcounter{equation}{9}
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
+\includegraphics[width=.9\linewidth]{lcp_schematic.pdf}
 
@@
 Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
-We assume the amount allocated to vegetative growth starts high amd then decreases once $H > H_{{\rm mat}}$  (Fig.
+We assume the amount allocated to vegetative growth starts high and then decreases once $H > H_{{\rm mat}}$  (Fig.
 \ref{fig:conceptual}b).
@@
 Leaf mass per unit leaf area: {\lma} (=$\phi$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever {\lma} is varied.
+In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004} (Fig.
+\ref{fig:leafturn}): $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
+The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted whenever {\lma} is varied.
 
@@
 \centering
-\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{main/growth_light_height.pdf}
+\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{growth_light_height.pdf}
 \caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
@@
 \centering
-\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
+\includegraphics[width=11.4cm,keepaspectratio]{lcp_trait.pdf}
 \caption{\textbf{Effect of three developmental traits on shade tolerance.}
@@
 
+% Defines references from supplemntary section
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{table}{0}
+\setcounter{figure}{0}
+
+\refstepcounter{table}\label{tab:params_core}
+\refstepcounter{table}\label{tab:params_hyper}
+\refstepcounter{figure}\label{fig:leafturn}
+\refstepcounter{figure}\label{fig:assumptions}
+\refstepcounter{figure}\label{fig:growth_light_dia}
+\refstepcounter{figure}\label{fig:growth_light_area}
+\refstepcounter{figure}\label{fig:growth_light_mass}
+
 % Bibliography

### 2018-05-01 ms/figure_conceptual.tex
Tweaks for technical check from journal

@@
 
+% \makeatletter
+% \newcommand\thefontsize[1]{{#1 The current font size is: \f@size pt\par}}
+% \makeatother
+
 %----------------------------------------------------------------------------------------
 \begin{document}
+
+% script = 14.4, footnote = 17.28, small = 20.74
+
+% \sffamily
+% \thefontsize\scriptsize
 
@@
   \node at (#1)[anchor=west, text width=2cm, minimum height=1cm] {
-  {\footnotesize #2}};
+  \fontsize{19.5pt}{15pt}\selectfont #2};
 }
@@
   \plot{  $(thiscol)$                 }{#3}
-  \ylabel{$(thiscol) + (0.7cm,   0cm)$}{=}
+  \ylabel{$(thiscol) + (0.7cm,   0cm)$}{\footnotesize=}
   \ylabel{$(thiscol) + (1.1cm,   0cm)$}{#4}
@@
 \firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/net_mass_production_dt_net_mass_production_dt}{}
-\othercol{2}{\hspace{-1em} \scriptsize $\,\alpha_{\rm \tiny bio} \alpha_{\rm \tiny y}$}{output/conn/net_mass_production_dt_yield}{}
-\othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/conn/net_mass_production_dt_assimilation}{}
-\othercol{4}{\hspace{-1em}-$\Sigma$\scriptsize $M_{\rm i} r_{\rm i}$}{output/conn/net_mass_production_dt_respiration}{\hspace{-0.5em}$\Bigg)$}
-\othercol{5}{\hspace{-0.75em}-$\Sigma$\scriptsize $M_{\rm i} k_{\rm i}$}{output/conn/net_mass_production_dt_turnover}{}
+\othercol{2}{\hspace{-0.8em} \fontsize{14}{10}\selectfont $\alpha_{\rm bio} \alpha_{\rm y}$}{output/conn/net_mass_production_dt_yield}{}
+\othercol{3}{\hspace{-1em}{\footnotesize$\times$}$\hspace{-0.1em}\Bigg($\hspace{-0.2em}\fontsize{16}{13}\selectfont $A_{\rm l} \bar{p}$}{output/conn/net_mass_production_dt_assimilation}{}
+\othercol{4}{\hspace{-1em}\fontsize{16}{13}\selectfont -$\Sigma M_{\rm i} r_{\rm i}$}{output/conn/net_mass_production_dt_respiration}{\hspace{-0.65em}$\Bigg)$}
+\othercol{5}{\hspace{-0.8em}\fontsize{16}{13}\selectfont -$\Sigma M_{\rm i} k_{\rm i}$}{output/conn/net_mass_production_dt_turnover}{}
 \rowlabel{eq:dbdt_2}
@@
 \firstcol{1}{$\frac{{\rm d}M_{{\rm t}}}{{\rm d}t}$}{output/conn/mass_total_dt_mass_total_dt}{}
-\othercol{2}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/conn/mass_total_dt_fraction_allocation_growth}{$\times$}
-\othercol{3}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/mass_total_dt_net_mass_production_dt}{$+$}
+\othercol{2}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/conn/mass_total_dt_fraction_allocation_growth}{\footnotesize$\times$}
+\othercol{3}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/mass_total_dt_net_mass_production_dt}{\footnotesize$+$}
 \othercol{4}{$\frac{{\rm d}M_{{\rm h}}}{{\rm d}t}$}{output/conn/mass_total_dt_mass_heartwood_dt}{}
@@
 \firstcol{1}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/conn/area_leaf_dt_area_leaf_dt}{}
-\othercol{2}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/conn/area_leaf_dt_darea_leaf_dmass_live}{$\times$}
-\othercol{3}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/conn/area_leaf_dt_fraction_allocation_growth}{$\times$}
+\othercol{2}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}M_{{\rm a}}}$}{output/conn/area_leaf_dt_darea_leaf_dmass_live}{\footnotesize$\times$}
+\othercol{3}{$\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$}{output/conn/area_leaf_dt_fraction_allocation_growth}{\footnotesize$\times$}
 \othercol{4}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/area_leaf_dt_net_mass_production_dt}{}
@@
 \firstcol{1}{$\frac{{\rm d}H}{{\rm d}t}$}{output/conn/height_dt_height_dt}{}
-\othercol{2}{$\frac{{\rm d}H}{{\rm d}A_{{\rm l}}}$}{output/conn/height_dt_dheight_darea_leaf}{$\times$}
+\othercol{2}{$\frac{{\rm d}H}{{\rm d}A_{{\rm l}}}$}{output/conn/height_dt_dheight_darea_leaf}{\footnotesize$\times$}
 \othercol{3}{$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/conn/area_leaf_dt_area_leaf_dt}{}
@@
 \firstcol{1}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/conn/area_stem_dt_area_stem_dt}{$\Bigg($}
-\othercol{2}{$\frac{{\rm d}A_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/conn/area_stem_dt_darea_bark_darea_leaf}{$+$}
+\othercol{2}{$\frac{{\rm d}A_{{\rm b}}}{{\rm d}A_{{\rm l}}}$}{output/conn/area_stem_dt_darea_bark_darea_leaf}{\footnotesize$+$}
 \othercol{3}{$\frac{{\rm d}A_{{\rm s}}}{{\rm d}A_{{\rm l}}}$}{output/conn/area_stem_dt_darea_sapwood_darea_leaf}{\hspace{-0.3em}$\Bigg)$}
-\othercol{4}{\hspace{-0.4em}$\times\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/conn/area_stem_dt_area_leaf_dt}{$+$}
+\othercol{4}{\hspace{-0.4em}{\footnotesize$\times$}$\frac{{\rm d}A_{{\rm l}}}{{\rm d}t}$}{output/conn/area_stem_dt_area_leaf_dt}{\footnotesize$+$}
 \othercol{5}{$\frac{{\rm d}A_{{\rm h}}}{{\rm d}t}$}{output/conn/area_stem_dt_area_heartwood_dt}{}
@@
 \firstcol{1}{$\frac{{\rm d}D}{{\rm d}t}$}{output/conn/diameter_stem_dt_diameter_stem_dt}{}
-\othercol{2}{$\frac{{\rm d}D}{{\rm d}A_{{\rm st}}}$}{output/conn/diameter_stem_dt_ddiameter_stem_darea_stem}{$\times$}
+\othercol{2}{$\frac{{\rm d}D}{{\rm d}A_{{\rm st}}}$}{output/conn/diameter_stem_dt_ddiameter_stem_darea_stem}{\footnotesize$\times$}
 \othercol{3}{$\frac{{\rm d}A_{{\rm st}}}{{\rm d}t}$}{output/conn/diameter_stem_dt_area_stem_dt}{}

### 2018-05-01 ms/ms-supp.Rnw
Tweaks for technical check from journal

@@
-\documentclass[9pt,onecolumn,twoside]{pnas-new}
+\documentclass[9pt,twoside]{pnas-new}
 % Use the lineno option to display guide line numbers if required.
-% Note that the use of elements such as single-column equations
-% may affect the guide line number alignment.
 
-\templatetype{pnasresearcharticle}
+\templatetype{pnassupportinginfo}
 
-\title{
-{\normalfont \fontsize{12pt}{16pt}
-{\color{black50} Supporting information for:}\\} \vskip6pt
-\fontsize{16pt}{20pt} How functional traits influence plant growth and shade tolerance across the life-cycle
-}
+% Use the lineno option to display guide line numbers if required.
+
+\title{How functional traits influence plant growth and shade tolerance across the life-cycle}
+\author{Daniel S. Falster Remko A. Duursma, Richard G. FitzJohn}
+\correspondingauthor{Corresponding Daniel Falster.\\E-mail: daniel.falster@unsw.edu.au}
+
+\usepackage{graphicx}
+\graphicspath{{../output/SI/}}
 
@@
 
-\usepackage{graphicx}
-\graphicspath{{../output/SI/}}
+\begin{document}
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Daniel S. Falster}
-\author[b]{Remko A. Duursma}
-\author[c]{Richard G. FitzJohn}
-
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
-Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
-\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
-\affil[c]{Imperial College, London, United Kingdom}
-
-% Please give the surname of the lead author for the running footer
-\leadauthor{Falster}
-
-\dates{This supplement was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\hypersetup{
-     linkcolor=pnasbluetext
-}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\begin{document}
+%% Comment/remove this line before generating final copy for submission
+%%\instructionspage
 
@@
 
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\theequation}{S\arabic{equation}}
-\renewcommand{\thesection}{S\arabic{section}}
+%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
+\SItext
 
-\section{Additional details on the \texttt{FF16} growth model in the {\plant} package}
+\section*{Additional details on the \texttt{FF16} growth model in the {\plant} package}
 
@@
 
-\subsection{Hyper-parameterisation of the \texttt{FF16} model via traits}
+\subsection*{Hyper-parameterisation of the \texttt{FF16} model via traits}
 
@@
 
-\subsubsection{Seed mass}
+\subsubsection*{Seed mass}
 
@@
 
-\subsubsection{Nitrogen per unit leaf area}
+\subsubsection*{Nitrogen per unit leaf area}
 
@@
 
-\subsubsection{Leaf mass per unit area}
+\subsubsection*{Leaf mass per unit area}
 
@@
 
-\subsubsection{Wood density}
+\subsubsection*{Wood density}
 
@@
 
-\subsubsection{Implementation of hyper-parameterisation function}
+\subsubsection*{Implementation of hyper-parameterisation function}
 
@@
 
-\section{References}
-\bibliography{references}
-
-\section{Supporting Tables and Figures}\label{app:supp_info_table}
-
 % latex table generated in R 3.4.0 by xtable 1.8-2 package
 % Thu Apr  5 13:08:59 2018
-\begin{table*}[ht]
+\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table*}
+\end{table}
 
@@
 % Thu Apr  5 13:08:59 2018
-\begin{table*}[ht]
+\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\end{table*}
+\end{table}
 
-\newpage
-\begin{figure*}[!hb]
+\clearpage
+\begin{figure}[!hb]
 \centering
@@
 \label{fig:leafturn}
-\end{figure*}
+\end{figure}
 \newpage
 
-\begin{figure*}[!hb]
+\begin{figure}[!hb]
 \centering
@@
 \label{fig:assumptions}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[!hb]
+\begin{figure}[!hb]
 \centering
@@
 \label{fig:growth_light_dia}
-\end{figure*}
+\end{figure}
 
@@
 \label{fig:growth_light_area}
-\end{figure*}
+\end{figure}
 
@@
 \label{fig:growth_light_mass}
-\end{figure*}
+\end{figure}
+
+\newpage
+\bibliography{references}
 

### 2018-05-01 ms/ms.tex
Tweaks for technical check from journal

@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Daniel S. Falster}
-\author[b]{Remko A. Duursma}
-\author[c]{Richard G. FitzJohn}
+\author[a,b,1]{Daniel S. Falster}
+\author[c]{Remko A. Duursma}
+\author[d,b]{Richard G. FitzJohn}
 
@@
 Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
-\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
-\affil[c]{Imperial College, London, United Kingdom}
+\affil[b]{Previous address: Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia}
+\affil[c]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
+\affil[d]{Imperial College, London, United Kingdom}
 
@@
 As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
-In this article we outline the mechanisms by which the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.
+In this article we outline the mechanisms by which the growth of individual plants may be influenced by various functional traits, as well as plant size and the light environment.
 
@@
 Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
-Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size.
-We consider five prominent traits representing specific physiological trade-offs: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).
+Based on the same decomposition of growth rates as is implemented in the {\plant} model \citep[from][]{Falster-2011, Falster-2017}, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
+Here we extend the results of \citep{Gibert-2016} to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size.
+We consider five prominent traits, each capturing a specific physiological trade-off: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).
 
@@
 The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
+A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in SI Appendix, Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
 
@@
 
-\subsection{Classic model for mass-based relative growth rate}
+\subsection*{Classic model for mass-based relative growth rate}
 
@@
 
-\subsection{Decomposition of growth rates into components}
+\subsection*{Decomposition of growth rates into components}
 
@@
 
-\subsection{Shade tolerance}
+\subsection*{Shade tolerance}
 
@@
 
-\subsection{A functional-balance model for plant construction}
+\subsection*{A functional-balance model for plant construction}
 
@@
 
-\subsection{Trait-based trade-offs}
+\subsection*{Trait-based trade-offs}
 
@@
 
-\subsubsection{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
+\subsubsection*{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
 
@@
 
-\subsubsection{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.
+\subsubsection*{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.
 
@@
 
-\subsection{Model assumptions}
+\subsection*{Model assumptions}
 
@@
 The white regions indicate trait ranges that are typically observed in real systems.
-Figs.
+In the SI Appendix, Figs.
 \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass.
@@
 
-\subsection{Changes in growth rate with size}
+\subsection*{Changes in growth rate with size}
 
@@
 
-\subsection{Changes in height growth rate with traits}
+\subsection*{Changes in height growth rate with traits}
 
@@
 
-\subsubsection{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease.
+\subsubsection*{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease.
 The only effect of {\seed} on growth comes from changing the plant's initial size.
@@
 
-\subsubsection{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
+\subsubsection*{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
 \ref{fig:growth_light_height}).
@@
 When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate.
-When plants are large, the influence of {\lma} on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
+When plants are large, the effect on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
 Thus, at larger sizes low {\lma} no longer increases growth rate (Fig.
@@
 
-\subsection{Changes in other growth rates with trait}
+\subsection*{Changes in other growth rates with trait}
 
@@
 \ref{fig:growth_light_height} focus on height growth rate.
-Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
+Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in the SI Appendix, Figs.
 \ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}.
@@
 
-\subsection{Responsiveness of growth rate to light}
+\subsection*{Responsiveness of growth rate to light}
 
-The predictions in Figs.
-\ref{fig:growth_light_height} and \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} illustrate how traits impact on growth rate under different light environments and at different sizes.
+The predictions in Fig.
+\ref{fig:growth_light_height} and Figs.
+\ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} in the SI Appendix illustrate how traits impact on growth rate under different light environments and at different sizes.
 An additional outcome that arises directly from these analyses is that traits moderate the responsiveness of growth to changes in light environment.
@@
 
-\subsection{Shade tolerance}
+\subsection*{Shade tolerance}
 Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
@@
 Moreover, we show that different traits and trade-offs generate different types of response.
-Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
+Combined with the available empirical evidence, these results demand a shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
 Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e.
@@
 
-\subsection{LMA doesn't predict growth rates of large plants}
+\subsection*{LMA doesn't predict growth rates of large plants}
 
@@
 
-\subsection{Generalising to other traits and trade-offs}
+\subsection*{Generalising to other traits and trade-offs}
 
@@
 
-\subsection{Implications for trait-based approaches}
+\subsection*{Implications for trait-based approaches}
 There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.
@@
 
-\subsection{Plasticity of traits through ontogeny}
+\subsection*{Plasticity of traits through ontogeny}
 
@@
 
-\subsection{Comparison with other frameworks}
+\subsection*{Comparison with other frameworks}
 
@@
 
-A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs.
+A feature distinguishing our approach \citep[see also][]{Falster-2011, Falster-2016, Falster-2017} from most of those mentioned above is the explicit linking to trait-based trade-offs.
+This link enables demographic outcomes and trade-offs to emerge from underlying phsyiologcal function.
 While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have focussed on ecosystem-level outcomes rather than the growth of individual plants.
@@
 While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
-Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}.
-Currently, reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
+Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by][]{Wenk-2015}.
+Currently, reproductive allocation receives little attention in models focussing on production \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
 For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
-Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a plant's maximum size and growth rate in adult plants \citep[e.g.][]{Wright-2010}.
+Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a maximum size and growth rate \citep[e.g.][]{Wright-2010}.
 
@@
 
-\subsection{Closing remarks}
+\subsection*{Closing remarks}
 
@@
 \matmethods{
-\vspace{-1em}
+\vspace{-2em}
 
-\subsection{Parameters}
-Parameters were sourced mainly from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
+\subsection*{Parameters}
+Parameters were sourced mainly from \citep{Falster-2016} (see SI Appendix, Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
 The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at {\hmat} (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
@@
 
-\subsection{Assumptions}
+\subsection*{Assumptions}
 The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
@@
 
-\subsection{Software}
-The growth model used has been implemented as the \texttt{FF16} physiological module within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
-The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
+\subsection*{Software}
+The growth model applied here is implemented as the \texttt{FF16} physiological module within v1.1 of the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
+The {\plant} package makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
+To encode the trait-based trade-offs described above, we {\plant}'s ability to provide a ``hyper-parameterisation'' function, which causes various parameters to covary with traits (for details see SI).
 
@@
 \acknow{
-We thank J Camac, A Gibert, G Kunstler, H Muller-Landau, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
+We thank two anonymous reviewers and many colleagues for helpful discussions and comments on earlier versions of this paper.
 DF was supported by fellowships from the Australian Research Council discovery (DP110102086, FT160100113).

### 2018-05-01 R/axis.R

@@ -1,14 +1,14 @@
 name_pretty <- function(trait) {
   switch(trait,
-         lma=expression(paste("Leaf-mass per area (kg ", m^-2, " )")),
-         rho=expression(paste("Wood density (kg ", m^-3, " )")),
-         narea=expression(paste("Nitrogen per area (kg ", m^-2, " )")),
+         lma=expression(paste("Leaf-mass per area (kg ", m^-2, ")")),
+         rho=expression(paste("Wood density (kg ", m^-3, ")")),
+         narea=expression(paste("Nitrogen per area (kg ", m^-2, ")")),
          hmat=expression(paste("Height at maturation (m)")),
          diameter_stem_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
          shading=expression(paste("WPLCP, ",E^"*"," (0-1)")),
-         leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, " )")),
+         leaf_turnover=expression(paste("Leaf turnover rate (", yr^-1, ")")),
 
-         height_dt=expression(paste("Height growth rate (m ", yr^-1, " )")),
+         height_dt=expression(paste("Height growth rate (m ", yr^-1, ")")),
          diameter_stem_dt=expression(paste("Diameter growth (m ", yr^-1, ")")),
          area_stem_dt=expression(paste("Basal area growth (", m^2, " ", yr^-1, ")")),
          mass_above_ground_dt=expression(paste("Mass growth (kg ", yr^-1, ")")),
@@ -42,7 +42,7 @@ trait_range_obs <- function(trait) {
 get_by <- function(ymax) {
     by <- diff(pretty(c(0, ymax)))[1]
     if(ymax > 0.6)
-        by <- 0.2
+        by <- 0.5
     if(ymax > 1.25)
         by <- 0.5
     if(ymax > 2.5)

### 2018-05-01 R/figures.R

@@ -303,8 +303,8 @@ figure_rate_vs_size_panels <- function(data, type, path) {
 #       xaxs="i", yaxs = "i",
       col=get_col(v), lwd=20)
     if(type == "diameter_stem_dt" || (type == "area_stem_dt"  && v %in% c( "area_leaf_dt", "area_heartwood_dt"))){
-      f <- function(at) axis(1, at=at, tck=-0.05, col="grey30", col.ticks = "grey30", cex.axis=7, mgp=c(3, 6, 0))
-      f(0); f(10); f(20);
+      f <- function(at) axis(1, at=at, tck=-0.05, col="grey30", col.ticks = "grey30", cex.axis=9, mgp=c(3, 6, 0))
+      f(0); f(20);
     }
     box(col="grey30")
     dev.off()
@@ -347,8 +347,10 @@ figure_dY_dt <- function(dat) {
   traits <- dat[["traits"]]
 
   cols <- color_pallete1(length(E))
+  is_main <- function(){vars[2]== "height_dt"}
+
   par(mfrow=c(length(sizes), length(traits)),
-      oma=c(4, 5, 0, 1.5), mar=c(1, 1, 1, 0.9))
+      oma=c(3, ifelse(is_main(), 4, 5), 0, 1.5), mar=c(0.8, 0.8, 0.8, 0.8))
 
   extract_f <- function(v, i, dat) {
       dat_v <- unname(split(dat$data[[v]], dat$data[[v]]$class))
@@ -376,15 +378,16 @@ figure_dY_dt <- function(dat) {
       axis.log10(1, labels = (i == length(sizes)), las=1)
       axis(2, labels = (v == traits[1] ), las=1, at = seq(0, 20, by=get_by(ymax)))
       if (v == last(traits)) {
-        mtext(dat[["label"]](sizes[[i]]), 4, cex=0.9, line=1)
+        
+        mtext(dat[["label"]](sizes[[i]]), 4, cex=0.8, line=1)
       }
       if (i == length(sizes)) {
-        mtext(name_pretty(v), 1, cex=0.9, line=3)
+        mtext(name_pretty(v), 1, cex=0.78, line=2.5)
       }
 
     }
   }
-  mtext(name_pretty(vars[2]), line=3, side=2, cex=0.9 , outer=TRUE)
[…]

### 2018-05-02 ms/Table1.tex
Separate files for Tables
Requested by PNAS office

--- 2018-04-05 ms/ms.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\documentclass[9pt,onecolumn]{pnas-new}
 
-% Enable new column type -- centered with fixed
-% Following http://texblog.org/2008/05/07/fwd-equal-cell-width-right-and-centre-aligned-content/
-\usepackage{array}
-\newcolumntype{x}[1]{%
->{\raggedleft}p{#1}}%
-
-% Enable equation numbers in Table
-% Following http://tex.stackexchange.com/questions/62974/numbering-equations-in-tabular-environment
-\newcommand{\eqncounter}[1]{%
-\mbox{}\refstepcounter{equation}%
-$[\theequation]$%
-\label{#1}
-}
-
-\input{ms-common-defs}
-
-\templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
-
-% Enable cross referencing to Evidence document
-\usepackage{xr}
-\externaldocument{ms-supp}
-\externaldocument{figure_conceptual}
-
-\title{How functional traits influence plant growth and shade tolerance across the life-cycle}
-
-% Please add here a significance statement to explain the relevance of your work
-% 120 words
-% Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself and is required for all research papers.
-Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself.
-\significancestatement{
-Plant species differ in many functional traits -- features of specific tissues and allocation of energy among them.
-While traits have been used in many correlative approaches to describe communities and demography, it has remained unclear how and why traits should influence whole-plant growth.
-Here we present a new theoretical framework for understanding the effect of traits on plant growth and shade-tolerance.
-This framework captures diverse patterns of growth in relation to size, and explains why the effect of traits on growth changes through ontogeny.
-By disentangling the effects of plant size, light environment, and traits on growth rates, this study provides a theoretical foundation for understanding growth across diverse tree species around the world.
-}
-
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{D.F. designed the study; D.F and R.F performed the analysis. All authors contributed to the writing of this manuscript.}
-
-\authordeclaration{The authors declare that they have no competing financial interests.\\
-
-Code reproducing the results is available at \href{https://github.com/traitecoevo/growth_trajectories}{github.com/traitecoevo/growth\_trajectories}.
-}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: daniel.falster@unsw.edu.au}
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{growth rate $|$ traits $|$ trees $|$ allometry $|$ model}
-
-% 250 words
-\begin{abstract}
-Plant species differ in many functional traits that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover.
-Yet, it remains unclear how -- and even if -- such traits influence whole-plant growth, with the simple linear relationships predicted by existing theory often lacking empirical support.
-Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth and shade-tolerance by extending a widely-used model, linking growth rate in seedlings with a single leaf trait, to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
-Based on biomass growth and allocation, this framework explains why the influence of traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
-Specifically, we demonstrate why for height growth the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
-Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results.
-\end{abstract}
-
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\lma}{\textsc{lma}}
+\newcommand{\wood}{\textsc{wd}}
+\newcommand{\seed}{\textsc{sm}}
+\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
+\newcommand{\nitrogen}{\textsc{n}}
 
 \begin{document}
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalized the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-% If your first paragraph (i.e.
-with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
-If this is the case, add \parshape=0 to the end of the list environment.
-
-\dropcap{F}unctional traits capture core differences in the strategies plants use to acquire and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
-Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients, and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover.
-During the last two decades, trade-offs related to some prominent traits have been quantified, with values for some leaf, wood, or seed traits now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
-As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
-In this article we outline the mechanisms by which the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.
-
-While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the trait leaf mass per unit leaf area ({\lma}) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
-{\lma} and a close correlate leaf lifespan, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
-These early successes prompted researchers to search for similar relationships in large plants.
-However, in saplings and trees showed, {\lma} was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
-Meanwhile, other traits such as wood density ({\wood}) showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
-Clearly, traits do not correlate simply and consistently to growth rate.
-
-Recently, it has become clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
-A recent meta-analysis of 103 studies reporting \textgreater 500 correlations provides the most compelling evidence \citep{Gibert-2016}.
-That study showed that the strength of the correlation between some traits (including {\lma} and maximum height) and growth rate was modified by size, while for other traits (including {\wood} and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Here, we extend previous work \citep{Gibert-2016} to provide a mechanistic explanation of these empirical patterns via a model of plant growth and allocation.
 
@@
 \toprule
-\textbf{Change in growth rate with increasing size (Fig. \ref{fig:conceptual})}\\
- $\,$ Biomass growth: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
- $\,$ Plant mass: Increasing \citep{Sillett-2010, Stephenson-2014} \\
- $\,$ Height: Hump-shaped \citep{Ryan-2006, Sillett-2010, King-2011} \\
- $\,$ Stem-diameter: Hump-shaped \citep{Canham-2004, Canham-2006, Herault-2011} \\
- $\,$ Relative growth rate (all variables): decreasing \citep{Rees-2010, Iida-2014}\\
-\textbf{Effect of traits on growth rate (Fig. \ref{fig:growth_light_height})}\\
- $\,$ {\seed}: $\downarrow$ values $\downarrow$ seedling size, \& thus $\downarrow$ absolute \& $\uparrow$ relative growth rate \citep{Gibert-2016} \\
- $\,$ $H_{\rm mat}$: $\downarrow$ values $\downarrow$ growth rate at at larger sizes \citep{Gibert-2016}\\
- $\,$  {\nitrogen}: $\downarrow$ values $\uparrow$ growth rate irrespective of size, but only in high light \citep{Gibert-2016}\\
- $\,$ {\lma}: $\downarrow$ values $\uparrow$ growth rate when small, not at mid-large sizes \citep{Gibert-2016}\\
- $\,$ {\wood}: $\downarrow$ values $\uparrow$ growth rate, except at largest sizes \citep{Gibert-2016}\\
-\textbf{Responsiveness of growth rate to changes in light, $E$ (Fig. \ref{fig:growth_light_height})}\\
+\textbf{Change in growth rate with increasing size (Fig. 1)}\\
+ $\,$ Biomass growth: Hump-shaped (22, 23) \\
+ $\,$ Plant mass: Increasing (24, 25) \\
+ $\,$ Height: Hump-shaped (24, 26, 27) \\
+ $\,$ Stem-diameter: Hump-shaped (10, 28, 29) \\
+ $\,$ Relative growth rate (all variables): decreasing (19, 30)\\
+\textbf{Effect of traits on growth rate (Fig. 3)}\\
+ $\,$ {\seed}: $\downarrow$ values $\downarrow$ seedling size, \& thus $\downarrow$ absolute \& $\uparrow$ relative growth rate (21) \\
+ $\,$ $H_{\rm mat}$: $\downarrow$ values $\downarrow$ growth rate at at larger sizes (21)\\
+ $\,$  {\nitrogen}: $\downarrow$ values $\uparrow$ growth rate irrespective of size, but only in high light (21)\\
+ $\,$ {\lma}: $\downarrow$ values $\uparrow$ growth rate when small, not at mid-large sizes (21)\\
+ $\,$ {\wood}: $\downarrow$ values $\uparrow$ growth rate, except at largest sizes (21)\\
+\textbf{Responsiveness of growth rate to changes in light, $E$ (Fig. 3)}\\
  $\,$ {\nitrogen}: $\uparrow$ values respond more \\
  $\,$ {\lma}: $\downarrow$ values respond more \\
- $\,$ {\wood}: $\downarrow$ values respond more \citep{Ruger-2012}\\
-\textbf{Shade tolerance, \textsc{wplcp} (Fig. \ref{fig:wplcp})}\\
- $\,$ Size: decreasing \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
- $\,$ {\nitrogen}: $\downarrow$ values $\uparrow$ shade tolerance \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
- $\,$ {\lma}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$ \& $\uparrow$ \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
- $\,$ {\wood}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Osunkoya-1996}\\
+ $\,$ {\wood}: $\downarrow$ values respond more (18)\\
+\textbf{Shade tolerance, \textsc{wplcp} (Fig. 4)}\\
+ $\,$ Size: decreasing  (22, 31, 32)\\
+ $\,$ {\nitrogen}: $\downarrow$ values $\uparrow$ shade tolerance (33–35)\\
+ $\,$ {\lma}: $\downarrow$ values $\downarrow$ shade tolerance (9, 32, 33, 35)$^3$ \& $\uparrow$ \textsc{lai} (15, 36, 37) \\
+ $\,$ {\wood}: $\downarrow$ values $\downarrow$ shade tolerance (38)\\
   \bottomrule
 \end{tabular}
+
 \addtabletext{{\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density, \textsc{lai} = leaf area index.
@@
 $^2$ Similar responses are predicted for leaf lifespan; here directly related to {\lma}.
-$^3$ \citep{Baltzer-2007} finds a relationship between \textsc{wplcp} \& leaf respiration rate expressed per unit leaf area or per unit leaf mass.
+$^3$ (35) finds a relationship between \textsc{wplcp} \& leaf respiration rate expressed per unit leaf area or per unit leaf mass.
 
@@
 
-\begin{table}[!hb]
- \caption{Key trade-offs (benefit and cost) for the five traits considered, as encoded into the {\plant} model.
- }
-\centering
-  \begin{tabular}{p{0.5cm}p{3.6cm}p{2.5cm}p{0.4cm}}
-  \toprule
-  Trait &  Benefit & Cost & Ref. \\
-  \midrule
-  {\seed} & $\uparrow$ size of seedlings & $\downarrow$ seed production rate & \citep{Moles-2006}\\
-  {\hmat}  & $\uparrow$ growth rate & $\downarrow$ seed production rate & \citep{Thomas-2011}\\
-  {\nitrogen} & $\uparrow$ photosynthetic rate in high light & $\uparrow$ respiration rate & \citep{Wright-2004}\\
-  {\lma} & $\downarrow$ leaf turnover rate & $\uparrow$ cost of building leaf & \citep{Wright-2004}\\
-  {\wood} & $\downarrow$ sapwood turnover rate & $\uparrow$ cost of building stem & \citep{Chave-2009}\\
-  \bottomrule
-  \end{tabular}
-\addtabletext{Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density.
-
-}
-\label{tab:traits}
-\end{table}
-
-Interpreting diverse empirical results seeking to link traits to growth rate has been challenging because, until recently \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
-A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to {\lma} \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-An extension of this equation suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
-But as noted above, the prediction for large plants is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
-Without further guidance, many researchers expected traits to map directly onto growth rates \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
-
-One reason theoretical predictions have been lacking or not been supported is that in existing theory the effects of traits is realised mainly via influences on net primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
-By contrast, the physiology of traits such as {\lma} and {\wood} suggests they influence allocation of biomass among different tissues, and their turnover rates, rather than net primary production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
-A second concern for theory focussed on net primary production is that measuring production is really only practical for small plants.
-On larger plants, growth is measured mainly as increment in either stem diameter or height \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-As outlined below, growth in height and diameter also depend on the allocation of biomass and tissue construction costs.
-
-Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence key elements of plant demography across the life-cycle.
-The {\plant} model builds on past approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
-Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
-Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size.
-We consider five prominent traits representing specific physiological trade-offs: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).
-
-We also show how our approach can account for related phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment (Table \ref{tab:phenomena}).
-Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
-In absolute terms, growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011}, diameter \citep{Canham-2004, Canham-2006, Herault-2011} or biomass growth \citep{Givnish-1988, Koch-2004}.
-In contrast, the growth rate of standing plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
-Expressed as relative growth rates, all growth measures decrease sharply with size \citep{Rees-2010, Iida-2014}.
-Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
-These diverse phenomena deserve an integrated explanation.
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=\linewidth]{main/figure_conceptual.pdf}
-\caption{Conceptual framework linking growth rate to plant size and traits.
-\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
-\textbf{b)} Equations describing the rates of biomass growth, and growth in various dimensions of the plant.
-In the first line the symbol $\Sigma$ means ``sum'' across tissues, where $i = {\rm l,b,s,r}$.
-Grey numbers indicate equation numbers referred to in the main text.
-The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}.
-Colours highlight where the same metric appears repeatedly in different equations.
-For a list of variable names see Table \ref{tab:definitions}.}
-\label{fig:conceptual}
-\end{figure}
-
-\section*{Modeling the effects of traits on demography}
-
-The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
-\ref{fig:conceptual}a).
-We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive, t=total, and st=stem total (s+b+h).
-The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
-
-Applying a standard approach, the amount of biomass growth on the plant, $\frac{{\rm d}B}{{\rm d}t}$, is given by the difference between income (photosynthesis) and losses (respiration and turnover)\citep{Makela-1997, Thornley-2000}:
-\begin{equation}\label{eq:dbdt}
-\frac{{\rm d}B}{{\rm d}t}
-= \alpha_{\rm bio} \,
-\alpha_{\rm y}
-\bigg( A_{\rm l} \, \bar{p}(E) -
-\, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}\bigg)
-- \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}.
-\end{equation}
-Photosynthesis is the product of the average photosynthetic rate per unit leaf area, $\bar{p}(E)$, and total leaf area, $A_{\rm l}$.
-We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (see SI for details), and respiration and turnover rates of different tissues are constants that potentially differ with traits.
-The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for the yield (fraction of C fixed in biomass) and the conversion of CO$_2$ into units of biomass, respectively.
-While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment so that we can better understand trait- and size-related effects.
-Many vegetation models also use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$ -- e.g.
-as functions of temperature -- though such detail will not qualitatively alter model behaviour.
-
-\begin{table}[!hb]
- \caption{Variable definitions.}
-\centering
-\begin{tabular}{p{0.5cm}p{1.75cm}p{5cm}}
-\toprule
-Symbol & Unit & Description \\
-\midrule
-\multicolumn{3}{l}{\textbf{Traits (constant through ontogeny)}}\\
-$\omega$ & kg & Seed mass ({\seed})  \\
-$H_{{\rm mat}}$ & m & Height at maturation ({\hmat})\\
-$\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf nitrogen per unit leaf area ({\nitrogen})  \\
-$\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf mass per area ({\lma}) \\
-$\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & Wood density ({\wood})\\
-\multicolumn{3}{l}{\textbf{State or rate variables (may vary through ontogeny)}} \\
-$H$ & m & Height of a plant\\
-$D$ & m & Stem diameter\\
-$M_i$ & kg & Mass of tissue type $i$ retained on plant \\
-$A_i$ & m$^2$ & Surface or cross-section are of tissue type $i$\\
-$\frac{{\rm d}B}{{\rm d}t}$ & kg yr$^{-1}$& Biomass growth rate\\
-$p,\bar{p}$ & mol yr$^{-1}$ m$^{-2}$ & Photosynthetic rate per unit area \\
-$r_i$ & mol yr$^{-1}$ kg$^{-1}$  & Respiration rate per unit mass of tissue type $i$ \\
-$k_i$ & yr$^{-1}$ & Turnover rate for tissue type $i$ \\
-\multicolumn{3}{l}{\textbf{Environmental variables (fixed)}} \\
-$E$ & & Canopy openness\\
-\multicolumn{3}{l}{\textbf{Other parameters (constant throughout)}} \\
-$\alpha_{\rm y}$ &  & Yield, fraction of carbon fixed into biomass\\
-$\alpha_{\rm bio}$  & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Biomass per mol carbon \\
-$\eta_c$ & & Crown-shape parameter\\
-$\theta$ &  & Sapwood area per unit leaf area\\
-$\alpha_{\rm l1}$ & m & Height of plant with leaf area of 1m$^2$ \\
-$\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Root mass per unit leaf area \\
-$\alpha_{\rm b1}$ &  & Ratio of bark area to sapwood area\\
-\bottomrule
-\end{tabular}
-\addtabletext{For $M_i$, $r_i$, $k_i$ and \(A_i\), subscripts refer to: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,  \({\rm h}\) = heartwood, \({\rm r}\) = roots, \({\rm st}\) = total stem, \({\rm a}\) =  alive tissue.
-
-}
-\label{tab:definitions}
-\end{table}
-
-\subsection{Classic model for mass-based relative growth rate}
-
-Earlier studies on seedlings related relative growth rate in mass to {\lma} \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}, using a model that can be derived from eqn \ref{eq:dbdt} as a special case of the more-extended model described below.
-Ignoring all turnover terms, as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt} and reproduction, biomass growth becomes a linear function of leaf area, making relative growth rate in mass:
-\begin{equation}\label{eq:RGR}
-\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx P_{\rm net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
-where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$, and $\phi$ is {\lma}.
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to {\lma} \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
-The derivations below clarify these links.
-
-\subsection{Decomposition of growth rates into components}
-
-% Update equation counter, following inclusion of figure 1
-\setcounter{equation}{9}
-
-To model growth in plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for biomass growth, but also for the costs of building new tissues and allocation to reproduction.
-Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig.
-\ref{fig:conceptual}b, eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
-While other decompositions are possible, the approach used here arises from the idea that multiple aspects of function and allocation decisions are made in relation to the amount of leaf area on a plant  \citep{Falster-2011}.
-Growth rates in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), total leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) therefore share several terms.
-Many of these terms also vary intrinsically with size, as shown in the insets in Fig.
-\ref{fig:conceptual}b.
-
-The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass growth $\frac{{\rm d}B}{{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
-The remaining $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
-In plants, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants \citep{Wenk-2015}.
-Note also that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
-So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ is not increasing.
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
-
-\caption{Conceptual framework showing how shade-tolerance (as measured by \textsc{wplcp}) is affected by plant size and traits, adapted from \citep{Givnish-1988}.
-From eqn \ref{eq:wplcp}, \textsc{wplcp} is the canopy openness $E$ where the photosynthetic income per unit leaf area (dashed line) intersects the total respiration and turnover costs per unit leaf area (solid black lines).
-Colors indicate the contributions from different tissues to total costs.
-Because the costs of stem tissues increase with height, \textsc{wplcp} also increases.
-Traits {\wood}, {\nitrogen}, and {\lma} can alter \textsc{wplcp} by shifting either income or cost of different tissues (Table \ref{tab:responses}).
-\label{fig:wplcp_idea}}
-\end{figure}
-
-The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
-
-The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood, and roots (eqn \ref{eq:daldmt}).
-The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of construction cost per unit leaf area for different tissues, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$ (eqn \ref{eq:daldmt}).
-
-The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
-This variable accounts for the architectural strategy of the plant (\citep{Falster-2011}).
-Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
-
-The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark, and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
-These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
-
-Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
-We make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}-\ref{eq:htm}).
-
-\subsection{Shade tolerance}
-
-Eqn \ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'' \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
-In general, photosynthesis rate per leaf area $\bar{p}$ increases with canopy openness, $E$.
-The {\wplcp} is then the value $E=E^*$ where photosynthetic gains balance the costs of tissue turnover and respiration, i.e.
-$\frac{{\rm d}B}{{\rm d}t}(E^*)=0$.
-From eqn \ref{eq:dbdt}, this occurs when
-\begin{equation}\label{eq:wplcp}
-\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
-\end{equation}
-Graphically, the {\wplcp} occurs at the point where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig.
-\ref{fig:wplcp_idea}).
-Traits influence {\wplcp} when they affect either photosynthesis, respiration, or turnover.
-Also, as the amount of stem support increases with plant height, {\wplcp} also increases with height \citep{Givnish-1988} (Fig.
-\ref{fig:wplcp_idea}).
-
-\subsection{A functional-balance model for plant construction}
-
-Since eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication, and differentiation, they may apply to a variety of growth model where biomass allocation is important.
-However, to make explicit predictions requires that we quantify all the terms in Fig.
-\ref{fig:conceptual}b via an explicit model of plant construction and function.
-
-The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citep{Makela-1997} and \citep{Moorcroft-2001}.
-We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
-Table \ref{tab:allometry} provides key equations (see \citep{Falster-2016} for full derivation).
-In particular, we assume that as a plant grows:
-1) Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
-2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
-3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
-4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and
-5) The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
-Assumption 1 accounts for the architectural layout of the plant.
-Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
-To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see SI).
-
-Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
-These eqns aso predict that the amount of live stem tissue supporting each unit of leaf area should increase linearly with height,
-\begin{equation}\label{eq:msml}
-\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
-\end{equation}
-Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
-\ref{fig:conceptual}.
-
-\begin{table*}
-\caption{Equations for a ``functional-balance'' model of plant construction. See Table \ref{tab:definitions} for a list of variable names and definitions.}
-\centering
-  \begin{tabular}{p{0.1cm}p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}
-  \toprule
-  & Variable & Function & Marginal cost & Growth rate & Eqn\\
-  \midrule
-  \multicolumn{6}{l}{\textbf{a) Functional-balance assumptions}} \\
-  & Height &
-    $H = \alpha_{\rm l1} \, A_{\rm l}^{0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} = 0.5 \alpha_{\rm l1} A_{\rm l}^{-0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
-  & Sapwood area &
-    $A_{\rm s} = \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
-  & Bark area &
-    $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
-  & Root mass &
-    $M_{\rm r} = \alpha_{\rm r1} \, A_{\rm l}$ &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_{\rm r1}$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
-  & Heartwood area & & &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\
-  \multicolumn{6}{l}{\textbf{b) Derived quantities}} \\
-  & Leaf mass &
-    $M_{\rm l} = \phi \, A_{\rm l} $ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
-  & Sapwood mass &
-    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( H + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
-  & Bark mass &
-    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( H + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
-  & Heartwood area &
-    $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
-  & Heartwood mass &
-    $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:htm} \\
-  \bottomrule
-\end{tabular}
-\addtabletext{\\The first column of \textbf{a)} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of \textbf{a)} and in \textbf{b)} can be derived from these assumptions.
-The column ``Eqn'' indicates equation numbers referred to in the main text.
-
-}
-\label{tab:allometry}
-\end{table*}
-
-\subsection{Trait-based trade-offs}
-
-To capture the effects of functional traits on growth rates and shade tolerance requires that trait-based trade-offs be embedded within the model described above.
-Here we consider five traits, for which we can posit specific physiological costs and benefits -- described in detail below and summarised in Table \ref{tab:traits}.
-It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values.
-In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait.
-
-\subsubsection{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
-
-Seed mass: {\seed} moderates a trade-off between a plant's rate of seed production per unit mass invested in reproduction and the mass of each offspring  \citep{Moles-2006}.
-
-Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
-We assume the amount allocated to vegetative growth starts high amd then decreases once $H > H_{{\rm mat}}$  (Fig.
-\ref{fig:conceptual}b).
-The exact nature of the function is non-critical, what is important is that plants shift from a period of investing mainly in vegetative growth to investing mainly in reproduction.
-Here we let the fraction of mass allocated to reproduction vary with $H$ via the function $r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-
-\subsubsection{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.
-
-Leaf nitrogen per unit leaf area: We assume that both the maximum photosynthetic capacity of a leaf and leaf respiration rate both vary with {\nitrogen} (=$\nu$), as $\beta_{{\rm lf1}} \, \left(\nu / \nu_0 \right)^{\beta_{{\rm lf5}}}$ and $\beta_{{\rm lf4}}\, \nu$ respectively, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf4}}$, and $\beta_{{\rm lf5}}$ are constants.
-
-Leaf mass per unit leaf area: {\lma} (=$\phi$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever {\lma} is varied.
-
-Wood density: {\wood} (=$\rho$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link {\wood} to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\rho / \rho_0\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
-The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of {\wood}, as such the mass-based rate is adjusted accordingly whenever {\wood} is varied.
-There currently little data relating sapwood turnover or respiration to {\wood}.
-
-\section*{Results}
-
-\subsection{Model assumptions}
-
-We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}.
-Fig.
-\ref{fig:assumptions} shows that the three functional-balance assumptions in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
-As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions (Table \ref{tab:allometry}a and eqn \ref{eq:msml}).
-
-\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
-\centering
-\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{main/growth_light_height.pdf}
-\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
-Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}.
-Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
-For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated.
-The white regions indicate trait ranges that are typically observed in real systems.
-Figs.
-\ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass.
-Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
-\label{fig:growth_light_height}}
-\end{SCfigure*}
-
-\subsection{Changes in growth rate with size}
-
-Our model suggests an intrinsically size-dependent pattern of biomass-production and growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
-The panels in Fig.
-\ref{fig:conceptual} show the predicted patterns for a typical woody plant, obtained by applying the functional-balance model from Table \ref{tab:allometry}a.
-Biomass growth shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
-Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
-This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with increasing size, including a strong decline in the fraction of plant that is leaf (Fig.
-\ref{fig:conceptual}), increasing reproductive allocation (Fig.
-\ref{fig:conceptual}), and declining biomass growth.
-In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
-Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
-All growth measures decrease sharply with size when expressed as relative growth rates.
-
-\subsection{Changes in height growth rate with traits}
-
-We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}).
-We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
-Across the five different traits, we observed four relatively distinct types of response.
-These responses are summarised in Table \ref{tab:responses} and described in more detail below.
-
-\subsubsection{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease.
-The only effect of {\seed} on growth comes from changing the plant's initial size.
-The plots in Fig.
-\ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
-Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in all metrics) because of their greater total leaf area.
-At the same time, relative growth rate is predicted to decrease with size, as the ratio of leaf area to support mass decreases.
-As plants grow, differences in initial mass decrease in importance relative to other factors influencing growth through the life-cycle.
-As a result, the correlations between {\seed} and growth rate disappear among larger plants.
-
-Greater {\hmat} causes a growth advantage at larger sizes by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
-\ref{fig:growth_light_height}).
-At smaller sizes, there is no differentiation among species, as all plants focus on growth.
-
-\begin{SCfigure*}[\sidecaptionrelwidth][b]
-\centering
-\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three developmental traits on shade tolerance.}
-Panels show effect of traits on level of canopy openness that causes biomass growth (eqn \ref{eq:dbdt}) to be zero.
-Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
-The white regions indicate trait ranges that are typically observed in real systems.
-\label{fig:wplcp}}
-\end{SCfigure*}
-
-\subsubsection{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
-\ref{fig:growth_light_height}).
-As $E$ increases, the optimal {\nitrogen} also increases.
-The invariance of the growth-trait relationship with respect to size arises as follows.
-The direct physiological effect of {\nitrogen} is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
-Both the cost and benefits of {\nitrogen} appear within the term $\frac{ {\rm d}B}{{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in biomass growth per {\nitrogen}.
-From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
-u} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial
-u}  - \frac{ \partial r_{\rm l}}{\partial
-u} )$. As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.
-
-Unlike {\nitrogen}, the response of growth rate to changes in {\lma} varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
-\ref{fig:growth_light_height}).
-As a result, the value of {\lma} that optimises plant growth increases with height, while the direction of correlation between height growth rate and {\lma} shifts from negative to positive, as plants increase in height.
-Decreasing {\lma} has two impacts on height growth rate.
-First, lower {\lma} increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$.
-Second, lower {\lma} decreases biomass growth $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
-Whether lower {\lma} increases height growth depends on the relative magnitude of these two effects.
-When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate.
-When plants are large, the influence of {\lma} on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
-Thus, at larger sizes low {\lma} no longer increases growth rate (Fig.
-\ref{fig:growth_light_height}).
-
-Reducing the cost of stem construction, via lower {\wood}, decreases the cost of deploying a unit of leaf area, increasing growth rate (Fig.
-\ref{fig:growth_light_height}).
-Unlike {\lma}, the benefits of cheaper stem construction become more pronounced at intermediate sizes, as a larger fraction of the plant is stem (Fig.
-\ref{fig:conceptual}a).
-
-\subsection{Changes in other growth rates with trait}
-
-The results reported above and in Fig.
-\ref{fig:growth_light_height} focus on height growth rate.
-Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
-\ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}.
-For each, plants were grown to a standard diameter, area, or mass.
-Changes in absolute and relative growth rates with traits therefore show similar patterns.
-
-We find that for {\seed}, {\nitrogen}, {\lma}, and {\hmat}, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to height growth (Table \ref{tab:responses}).
-The only trait where a slightly different response was observed was for {\wood}.
-Whereas the effect of {\wood} on height growth tended to diminish slightly at larger sizes (Fig.
-\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area, or above-ground mass.
-Sapwood lost via turnover becomes heartwood.
-Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.
-
-\subsection{Responsiveness of growth rate to light}
-
-The predictions in Figs.
-\ref{fig:growth_light_height} and \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} illustrate how traits impact on growth rate under different light environments and at different sizes.
-An additional outcome that arises directly from these analyses is that traits moderate the responsiveness of growth to changes in light environment.
-This response arises because individuals with higher potential growth rate naturally have greater potential plasticity in growth.
-Our results therefore support findings that species with low {\wood} increase growth more substantially with increases in light (Table \ref{tab:phenomena}).
-Variation in {\lma} also moderates the response of growth to changes in light, with species having the lowest {\lma} being most responsive.
-However, unlike for {\wood}, the effect appears only for the smallest size classes.
-
-\subsection{Shade tolerance}
-Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
-\begin{equation}\label{eq:wplcp_2}
-\bar{p}(E^*) =
-      \phi c_{\rm l} +
-      \left(\theta \, \rho \, \eta_c \, H\right)
-        \left(b c_{\rm b}
-            + c_{\rm s}\right) +
-      \alpha_{\rm r1} c_{\rm r},
-\end{equation}
-where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
-Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}.
-With some further manipulations, it is possible to show that {\wplcp} will decrease with {\lma} if $\beta_{{\rm kl2}} >1$.
-Likewise {\wplcp} will decrease with {\wood} if $\beta_{{\rm ks2}} >1$.
-The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to {\lma} and {\wood}, respectively.
-Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low {\lma} and low {\wood} are predicted to be less shade tolerant (Fig.
-\ref{fig:wplcp}).
-At low {\lma} ({\wood}), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover.
-{\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
-In addition, {\wplcp} varies with {\nitrogen}.
-At small sizes, {\wplcp} increases with {\nitrogen} across the band of values typically observed in real plants, i.e.
-high {\nitrogen} makes seedlings shade intolerant.
-At larger sizes, as biomass growth declines to zero, {\wplcp} begins to increase again for very low values of {\nitrogen}.
-All these results  match empirical patterns (Table \ref{tab:phenomena}).
-
-\newcommand{\sepp}{{\color{grey}/}}
-\newcommand{\upup}{$\uparrow\,\uparrow$}
-\newcommand{\updo}{$\uparrow\,\downarrow$}
-\newcommand{\dodo}{$\downarrow\,\downarrow$}
-\newcommand{\upfl}{$\uparrow$\,--}
-\newcommand{\flup}{--$\,\uparrow$}
-\newcommand{\dofl}{$\downarrow\,$--}
-\newcommand{\doup}{$\downarrow\,\uparrow$}
-\newcommand{\fldo}{--$\,\downarrow$}
-
-\begin{table}[hb!]
-\centering
-\caption{Predicted effects of traits on components of plant function determining growth rate. Adapted and expanded from \citep{Gibert-2016}.}
-  \begin{tabular}{llx{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}p{0.01cm}}
-  \toprule
-  & &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
-  & & {\seed} & $H_{{\rm mat}}$ & {\nitrogen} & {\lma} & {\wood} & \\
-  \midrule
-  \multicolumn{8}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
-  \multicolumn{2}{l} {Biomass growth, $\ud B / \ud t$} & & & & & & \\
-  & \tabitem Photosynthesis & - & - & $\uparrow$  & - & - & \\
-  & \tabitem Respiration & - & - & $\uparrow$  & - & - & \\
-  & \tabitem Turnover & - & - & - & $\uparrow$ & $\uparrow$ & \\
-  \multicolumn{2}{l} {Allocation to growth, $\ud M_{\rm a} / \ud B$} & - & $\uparrow$ & - & - & - & \\
-  \multicolumn{2}{l} {Leaf deployment, $\ud A_{\rm l} / \ud M_{\rm a}$} & & \\
-  & \tabitem Leaf  & - & - & - & $\downarrow$ & - & \\
-  & \tabitem Sapwood & - & - & - & - & $\downarrow$ & \\
-\midrule
- \multicolumn{8}{l} {\textbf{b) Predicted effect of trait on growths rate for a small and large plant}} \\
-  \multicolumn{8}{l} {Height} \\
-  & \tabitem{Absolute}, $\ud H / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Stem area} \\
-  & \tabitem{Absolute}, $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Stem diameter} \\
-  & \tabitem{Absolute}, $\ud D / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Mass} \\
-  & \tabitem{Absolute}, $\ud M_{\rm t} / \ud t$ & \upfl &  \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud M_{\rm t} / (\ud t . M_{\rm t} )$ & \dofl &  \flup & \upup & \dofl & \dodo & \\
-  \bottomrule
-  \end{tabular}
-\addtabletext{Arrows indicate the effect of increased trait value on each component; dashes indicating no effect.
-Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density (see Tables \ref{tab:traits}-\ref{tab:definitions}).
-
-}
-\label{tab:responses}
-\end{table}
-
-\section*{Discussion}
-
-Using a model relating plant physiological function and mass allocation to prominent traits, we have clarified how traits could impact on plant growth across the life cycle.
-Our approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
-During recent decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}.
-However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}.
-Matching growing empirical evidence (Table \ref{tab:phenomena}), this study shows when and why the direction or strength of a correlation between traits and growth rate shifts with plant size.
-Moreover, we show that different traits and trade-offs generate different types of response.
-Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
-Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e.
-lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}.
-The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter, and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
-This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}.
-Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects.
-Our aim is to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
-
-\subsection{LMA doesn't predict growth rates of large plants}
-
-The model presented here extends a widely-used model for seedlings, linking mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
-Importantly, the seedling model (eqn \ref{eq:RGR}) can be derived as a special case of the new, extended model.
-Unlike eqn \ref{eq:RGR}, the extended model also predicts a relationship between {\lma} and growth rate that changes with size.
-In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even weakly positive in larger plants (Fig.
-\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem diameter, stem area or total mass.
-This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size.
-As seedlings, leaves comprise a large part of the plant (Fig.
-\ref{fig:conceptual}a).
-Decreasing {\lma} then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
-As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheap leaf construction.
-Consequently, the effect of {\lma} on leaf turnover comes to dominate at larger sizes, and as such, the effect of {\lma} on height, diameter, and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
-
-\subsection{Generalising to other traits and trade-offs}
-
-The list of functional traits known to differ among plant species is ever-increasing \citep{Perez-2013}.
-While we have focussed on understanding the effects of five traits on elements of demography, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
-The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be connected with variables in eqns \ref{eq:dbdt}--\ref{eq:dDdt}.
-While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
-A well-developed trade-off must include two opposing forces that operate at some point in the life cycle.
-
-Our framework also highlights what is needed for traits to influence demography.
-While traits can influence many aspects of plant function, these influences must operate via the pathways outlined in Fig.
-\ref{fig:conceptual} if the trait is going to impact on growth rate or shade tolerance.
-For example, many studies consider traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
-By altering conductance of water to the leaf, both traits will influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}), while vessel size might also affect rates of stem turnover ($k_{\rm s}$ in eqn \ref{eq:dbdt}).
-The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in via the parameter $\theta$ (Table \ref{tab:allometry}).
-The effect of both these traits on growth rate should be expected to vary with plant size.
-
-\subsection{Implications for trait-based approaches}
-There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.
-
-First, our results highlight the importance of allocation decisions and turnover costs in determining demography.
-Much of current ecosystem research focusses on factors affecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
-Yet, four of the five traits considered here do not influence net primary production.
-In fact, our analysis shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
-Low {\lma} results in high leaf turnover, such that individuals with a low {\lma} have lower biomass growth.
-It is this property that makes them shade intolerant.
-And yet they can still achieve a growth advantage (when small).
-
-Second, our results demand a shift away from viewing species as lying on a spectrum  from slow to fast growth.
-When the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum.
-Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.
-Researchers using functional traits have also tended to describe species as fast or slow growing \citep{Adler-2014, Diaz-2016}.
-Our results suggest a more-nuanced approach.
-Plants that are fast growing as seedlings may not be fast growing as saplings or adults, or under low light.
-Plants that are fast growing as adults may not be fast growing as seedlings.
-This more-nuanced perspective reflects observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
-
-Third, our results suggest that even if traits define a potential growth trajectory, researchers seeking to link traits to growth rate must probe deeper into the data than simply looking for a linear relationship between traits and average growth, to recover the expected relationships.
-None of the predicted relationships between traits and growth is linear across the range of sizes and light environments tested.
-As such, we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits \citep[e.g.][]{Poorter-2008,Paine-2015}.
-Controlling for size, site, and light environment will be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}, as will having a clear expectation for the hypothesised relationship.
-
-\subsection{Plasticity of traits through ontogeny}
-
-While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
-To that end, we make a further prediction arising from our results, which is that the trait {\lma} should increase through ontogeny for all individuals.
-Such shifts have been observed across a variety of species \citep{King-1999,Thomas-1999,Koch-2004}.
-Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low {\lma} -- to increase their {\lma} as they grow larger.
-While a similar prediction was made for a species of \emph{Eucalyptus} \citep{King-1999}, we extend the idea across species.
-Although trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
-This hypothesis attempts to give meaning to some of that variation, and shows how variation across and within species might be understood within a single framework.
-
-\subsection{Comparison with other frameworks}
-
-As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}.
-These models have several properties in common, including that they all have growth being driven by the net amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
-Many models also make functional-balance assumptions, for example linking the cross-section of sapwood to leaf area \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-2005, Moorcroft-2001}.
-We note that an assumption of exact functional balance is not critical for our results, what matters is that the amount of live biomass (i.e.
-excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
-\ref{fig:assumptions}d).
-
-A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs.
-While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have focussed on ecosystem-level outcomes rather than the growth of individual plants.
-Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
-Another study \citep{King-1999} connected a model of growth for a single species to the trait {\lma}, and likewise predicted a gradual flattening out of the relationship between {\lma} and growth rate with size (as in Fig.
-\ref{fig:growth_light_height}).
-Here, we include an additional cost of increased leaf turnover that further penalised low {\lma} strategies when large.
-
-Perhaps the two most controversial elements of our approach concerns the assumptions about tissue replacement and reproductive allocation.
-Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing any surplus.
-This is because we assume tissues lost via turnover are replaced before mass is allocated to either new growth (i.e.
-growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant.
-The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus mass to reproduction.
-While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
-Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}.
-Currently, reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
-For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
-Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a plant's maximum size and growth rate in adult plants \citep[e.g.][]{Wright-2010}.
-
-Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory  of ecology (\textsc{mst}) \citep{Enquist-1999, Enquist-2007}.
-Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
-First, the \textsc{mst} model suggests diameter growth continues to increase wit size, whereas empirical data suggests growth rate declines for larger plants \citep{Canham-2004, Canham-2006, Herault-2011}.
-Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size.
-Predicted effects are for a linear increase in growth with decreases in either {\lma} and lower {\wood}, that apply irrespective of size.
-However, at least for {\lma} such effects in large trees have not been observed.
-
-\subsection{Closing remarks}
-
-We have shown how diverse phenomena related to plant growth can be understood with a model accounting for processes generating photosynthetic income and allocating this among different tissues.
-The need to consider effects of plant size, alongside trait-based differences among species has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
-Here we have provided a framework for achieving this.
-By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid  foundation for understanding and modelling growth across diverse species around the world.
-
-\matmethods{
-\vspace{-1em}
-
-\subsection{Parameters}
-Parameters were sourced mainly from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at {\hmat} (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
-\ref{fig:conceptual}).
-
-\subsection{Assumptions}
-The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
-We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
-We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed Gaussian.
-
-\subsection{Software}
-The growth model used has been implemented as the \texttt{FF16} physiological module within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
-The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
-
-}
-
-\showmatmethods % Display the Materials and Methods section
-
-\acknow{
-We thank J Camac, A Gibert, G Kunstler, H Muller-Landau, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
-DF was supported by fellowships from the  Australian Research Council discovery (DP110102086, FT160100113).
-RF was supported by the Science and Industry Endowment Fund (RP04-174).
-The authors have no conflicts of interest to declare.
-}
-
-% Bibliography
-\bibliography{references}
-%\pnasbreak
 \end{document}

### 2018-05-02 ms/Table2.tex
Separate files for Tables
Requested by PNAS office

--- 2018-04-05 ms/ms.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\documentclass[9pt,onecolumn]{pnas-new}
 
-% Enable new column type -- centered with fixed
-% Following http://texblog.org/2008/05/07/fwd-equal-cell-width-right-and-centre-aligned-content/
-\usepackage{array}
-\newcolumntype{x}[1]{%
->{\raggedleft}p{#1}}%
-
-% Enable equation numbers in Table
-% Following http://tex.stackexchange.com/questions/62974/numbering-equations-in-tabular-environment
-\newcommand{\eqncounter}[1]{%
-\mbox{}\refstepcounter{equation}%
-$[\theequation]$%
-\label{#1}
-}
-
-\input{ms-common-defs}
-
-\templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
-
-% Enable cross referencing to Evidence document
-\usepackage{xr}
-\externaldocument{ms-supp}
-\externaldocument{figure_conceptual}
-
-\title{How functional traits influence plant growth and shade tolerance across the life-cycle}
-
-% Please add here a significance statement to explain the relevance of your work
-% 120 words
-% Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself and is required for all research papers.
-Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself.
-\significancestatement{
-Plant species differ in many functional traits -- features of specific tissues and allocation of energy among them.
-While traits have been used in many correlative approaches to describe communities and demography, it has remained unclear how and why traits should influence whole-plant growth.
-Here we present a new theoretical framework for understanding the effect of traits on plant growth and shade-tolerance.
-This framework captures diverse patterns of growth in relation to size, and explains why the effect of traits on growth changes through ontogeny.
-By disentangling the effects of plant size, light environment, and traits on growth rates, this study provides a theoretical foundation for understanding growth across diverse tree species around the world.
-}
-
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{D.F. designed the study; D.F and R.F performed the analysis. All authors contributed to the writing of this manuscript.}
-
-\authordeclaration{The authors declare that they have no competing financial interests.\\
-
-Code reproducing the results is available at \href{https://github.com/traitecoevo/growth_trajectories}{github.com/traitecoevo/growth\_trajectories}.
-}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: daniel.falster@unsw.edu.au}
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{growth rate $|$ traits $|$ trees $|$ allometry $|$ model}
-
-% 250 words
-\begin{abstract}
-Plant species differ in many functional traits that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover.
-Yet, it remains unclear how -- and even if -- such traits influence whole-plant growth, with the simple linear relationships predicted by existing theory often lacking empirical support.
-Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth and shade-tolerance by extending a widely-used model, linking growth rate in seedlings with a single leaf trait, to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
-Based on biomass growth and allocation, this framework explains why the influence of traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
-Specifically, we demonstrate why for height growth the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
-Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results.
-\end{abstract}
-
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\lma}{\textsc{lma}}
+\newcommand{\wood}{\textsc{wd}}
+\newcommand{\seed}{\textsc{sm}}
+\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
+\newcommand{\nitrogen}{\textsc{n}}
 
 \begin{document}
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalized the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-% If your first paragraph (i.e.
-with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
-If this is the case, add \parshape=0 to the end of the list environment.
-
-\dropcap{F}unctional traits capture core differences in the strategies plants use to acquire and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
-Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients, and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover.
-During the last two decades, trade-offs related to some prominent traits have been quantified, with values for some leaf, wood, or seed traits now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
-As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
-In this article we outline the mechanisms by which the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.
-
-While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the trait leaf mass per unit leaf area ({\lma}) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
-{\lma} and a close correlate leaf lifespan, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
-These early successes prompted researchers to search for similar relationships in large plants.
-However, in saplings and trees showed, {\lma} was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
-Meanwhile, other traits such as wood density ({\wood}) showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
-Clearly, traits do not correlate simply and consistently to growth rate.
-
-Recently, it has become clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
-A recent meta-analysis of 103 studies reporting \textgreater 500 correlations provides the most compelling evidence \citep{Gibert-2016}.
-That study showed that the strength of the correlation between some traits (including {\lma} and maximum height) and growth rate was modified by size, while for other traits (including {\wood} and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Here, we extend previous work \citep{Gibert-2016} to provide a mechanistic explanation of these empirical patterns via a model of plant growth and allocation.
 
 \begin{table}[!ht]
-\caption{Empirical phenomena explained in this paper.}
-\begin{tabular}{p{8cm}}
-\toprule
-\textbf{Change in growth rate with increasing size (Fig. \ref{fig:conceptual})}\\
- $\,$ Biomass growth: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
- $\,$ Plant mass: Increasing \citep{Sillett-2010, Stephenson-2014} \\
- $\,$ Height: Hump-shaped \citep{Ryan-2006, Sillett-2010, King-2011} \\
- $\,$ Stem-diameter: Hump-shaped \citep{Canham-2004, Canham-2006, Herault-2011} \\
- $\,$ Relative growth rate (all variables): decreasing \citep{Rees-2010, Iida-2014}\\
-\textbf{Effect of traits on growth rate (Fig. \ref{fig:growth_light_height})}\\
- $\,$ {\seed}: $\downarrow$ values $\downarrow$ seedling size, \& thus $\downarrow$ absolute \& $\uparrow$ relative growth rate \citep{Gibert-2016} \\
- $\,$ $H_{\rm mat}$: $\downarrow$ values $\downarrow$ growth rate at at larger sizes \citep{Gibert-2016}\\
- $\,$  {\nitrogen}: $\downarrow$ values $\uparrow$ growth rate irrespective of size, but only in high light \citep{Gibert-2016}\\
- $\,$ {\lma}: $\downarrow$ values $\uparrow$ growth rate when small, not at mid-large sizes \citep{Gibert-2016}\\
- $\,$ {\wood}: $\downarrow$ values $\uparrow$ growth rate, except at largest sizes \citep{Gibert-2016}\\
-\textbf{Responsiveness of growth rate to changes in light, $E$ (Fig. \ref{fig:growth_light_height})}\\
- $\,$ {\nitrogen}: $\uparrow$ values respond more \\
- $\,$ {\lma}: $\downarrow$ values respond more \\
- $\,$ {\wood}: $\downarrow$ values respond more \citep{Ruger-2012}\\
-\textbf{Shade tolerance, \textsc{wplcp} (Fig. \ref{fig:wplcp})}\\
- $\,$ Size: decreasing \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
- $\,$ {\nitrogen}: $\downarrow$ values $\uparrow$ shade tolerance \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
- $\,$ {\lma}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$ \& $\uparrow$ \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
- $\,$ {\wood}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Osunkoya-1996}\\
-  \bottomrule
-\end{tabular}
-\addtabletext{{\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density, \textsc{lai} = leaf area index.
-Footnotes: $^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area \& dark respiration rate per leaf area; here both related to {\nitrogen}.
-$^2$ Similar responses are predicted for leaf lifespan; here directly related to {\lma}.
-$^3$ \citep{Baltzer-2007} finds a relationship between \textsc{wplcp} \& leaf respiration rate expressed per unit leaf area or per unit leaf mass.
-
-}
-\label{tab:phenomena}
-\end{table}
-
-\begin{table}[!hb]
  \caption{Key trade-offs (benefit and cost) for the five traits considered, as encoded into the {\plant} model.
@@
   \midrule
-  {\seed} & $\uparrow$ size of seedlings & $\downarrow$ seed production rate & \citep{Moles-2006}\\
-  {\hmat}  & $\uparrow$ growth rate & $\downarrow$ seed production rate & \citep{Thomas-2011}\\
-  {\nitrogen} & $\uparrow$ photosynthetic rate in high light & $\uparrow$ respiration rate & \citep{Wright-2004}\\
-  {\lma} & $\downarrow$ leaf turnover rate & $\uparrow$ cost of building leaf & \citep{Wright-2004}\\
-  {\wood} & $\downarrow$ sapwood turnover rate & $\uparrow$ cost of building stem & \citep{Chave-2009}\\
+  {\seed} & $\uparrow$ size of seedlings & $\downarrow$ seed production rate & (39)\\
+  {\hmat}  & $\uparrow$ growth rate & $\downarrow$ seed production rate & (40)\\
+  {\nitrogen} & $\uparrow$ photosynthetic rate in high light & $\uparrow$ respiration rate & (2)\\
+  {\lma} & $\downarrow$ leaf turnover rate & $\uparrow$ cost of building leaf & (2)\\
+  {\wood} & $\downarrow$ sapwood turnover rate & $\uparrow$ cost of building stem & (2)\\
   \bottomrule
   \end{tabular}
+
 \addtabletext{Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density.
@@
 
-Interpreting diverse empirical results seeking to link traits to growth rate has been challenging because, until recently \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
-A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to {\lma} \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-An extension of this equation suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
-But as noted above, the prediction for large plants is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
-Without further guidance, many researchers expected traits to map directly onto growth rates \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
-
-One reason theoretical predictions have been lacking or not been supported is that in existing theory the effects of traits is realised mainly via influences on net primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
-By contrast, the physiology of traits such as {\lma} and {\wood} suggests they influence allocation of biomass among different tissues, and their turnover rates, rather than net primary production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
-A second concern for theory focussed on net primary production is that measuring production is really only practical for small plants.
-On larger plants, growth is measured mainly as increment in either stem diameter or height \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-As outlined below, growth in height and diameter also depend on the allocation of biomass and tissue construction costs.
-
-Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence key elements of plant demography across the life-cycle.
-The {\plant} model builds on past approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
-Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
-Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size.
-We consider five prominent traits representing specific physiological trade-offs: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).
-
-We also show how our approach can account for related phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment (Table \ref{tab:phenomena}).
-Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
-In absolute terms, growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011}, diameter \citep{Canham-2004, Canham-2006, Herault-2011} or biomass growth \citep{Givnish-1988, Koch-2004}.
-In contrast, the growth rate of standing plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
-Expressed as relative growth rates, all growth measures decrease sharply with size \citep{Rees-2010, Iida-2014}.
-Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
-These diverse phenomena deserve an integrated explanation.
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=\linewidth]{main/figure_conceptual.pdf}
-\caption{Conceptual framework linking growth rate to plant size and traits.
-\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
-\textbf{b)} Equations describing the rates of biomass growth, and growth in various dimensions of the plant.
-In the first line the symbol $\Sigma$ means ``sum'' across tissues, where $i = {\rm l,b,s,r}$.
-Grey numbers indicate equation numbers referred to in the main text.
-The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}.
-Colours highlight where the same metric appears repeatedly in different equations.
-For a list of variable names see Table \ref{tab:definitions}.}
-\label{fig:conceptual}
-\end{figure}
-
-\section*{Modeling the effects of traits on demography}
-
-The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
-\ref{fig:conceptual}a).
-We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive, t=total, and st=stem total (s+b+h).
-The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
-
-Applying a standard approach, the amount of biomass growth on the plant, $\frac{{\rm d}B}{{\rm d}t}$, is given by the difference between income (photosynthesis) and losses (respiration and turnover)\citep{Makela-1997, Thornley-2000}:
-\begin{equation}\label{eq:dbdt}
-\frac{{\rm d}B}{{\rm d}t}
-= \alpha_{\rm bio} \,
-\alpha_{\rm y}
-\bigg( A_{\rm l} \, \bar{p}(E) -
-\, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}\bigg)
-- \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}.
-\end{equation}
-Photosynthesis is the product of the average photosynthetic rate per unit leaf area, $\bar{p}(E)$, and total leaf area, $A_{\rm l}$.
-We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (see SI for details), and respiration and turnover rates of different tissues are constants that potentially differ with traits.
-The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for the yield (fraction of C fixed in biomass) and the conversion of CO$_2$ into units of biomass, respectively.
-While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment so that we can better understand trait- and size-related effects.
-Many vegetation models also use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$ -- e.g.
-as functions of temperature -- though such detail will not qualitatively alter model behaviour.
-
-\begin{table}[!hb]
- \caption{Variable definitions.}
-\centering
-\begin{tabular}{p{0.5cm}p{1.75cm}p{5cm}}
-\toprule
-Symbol & Unit & Description \\
-\midrule
-\multicolumn{3}{l}{\textbf{Traits (constant through ontogeny)}}\\
-$\omega$ & kg & Seed mass ({\seed})  \\
-$H_{{\rm mat}}$ & m & Height at maturation ({\hmat})\\
-$\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf nitrogen per unit leaf area ({\nitrogen})  \\
-$\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Leaf mass per area ({\lma}) \\
-$\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & Wood density ({\wood})\\
-\multicolumn{3}{l}{\textbf{State or rate variables (may vary through ontogeny)}} \\
-$H$ & m & Height of a plant\\
-$D$ & m & Stem diameter\\
-$M_i$ & kg & Mass of tissue type $i$ retained on plant \\
-$A_i$ & m$^2$ & Surface or cross-section are of tissue type $i$\\
-$\frac{{\rm d}B}{{\rm d}t}$ & kg yr$^{-1}$& Biomass growth rate\\
-$p,\bar{p}$ & mol yr$^{-1}$ m$^{-2}$ & Photosynthetic rate per unit area \\
-$r_i$ & mol yr$^{-1}$ kg$^{-1}$  & Respiration rate per unit mass of tissue type $i$ \\
-$k_i$ & yr$^{-1}$ & Turnover rate for tissue type $i$ \\
-\multicolumn{3}{l}{\textbf{Environmental variables (fixed)}} \\
-$E$ & & Canopy openness\\
-\multicolumn{3}{l}{\textbf{Other parameters (constant throughout)}} \\
-$\alpha_{\rm y}$ &  & Yield, fraction of carbon fixed into biomass\\
-$\alpha_{\rm bio}$  & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & Biomass per mol carbon \\
-$\eta_c$ & & Crown-shape parameter\\
-$\theta$ &  & Sapwood area per unit leaf area\\
-$\alpha_{\rm l1}$ & m & Height of plant with leaf area of 1m$^2$ \\
-$\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & Root mass per unit leaf area \\
-$\alpha_{\rm b1}$ &  & Ratio of bark area to sapwood area\\
-\bottomrule
-\end{tabular}
-\addtabletext{For $M_i$, $r_i$, $k_i$ and \(A_i\), subscripts refer to: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,  \({\rm h}\) = heartwood, \({\rm r}\) = roots, \({\rm st}\) = total stem, \({\rm a}\) =  alive tissue.
-
-}
-\label{tab:definitions}
-\end{table}
-
-\subsection{Classic model for mass-based relative growth rate}
-
-Earlier studies on seedlings related relative growth rate in mass to {\lma} \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}, using a model that can be derived from eqn \ref{eq:dbdt} as a special case of the more-extended model described below.
-Ignoring all turnover terms, as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt} and reproduction, biomass growth becomes a linear function of leaf area, making relative growth rate in mass:
-\begin{equation}\label{eq:RGR}
-\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx P_{\rm net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
-where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$, and $\phi$ is {\lma}.
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to {\lma} \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
-The derivations below clarify these links.
-
-\subsection{Decomposition of growth rates into components}
-
-% Update equation counter, following inclusion of figure 1
-\setcounter{equation}{9}
-
-To model growth in plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for biomass growth, but also for the costs of building new tissues and allocation to reproduction.
-Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig.
-\ref{fig:conceptual}b, eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
-While other decompositions are possible, the approach used here arises from the idea that multiple aspects of function and allocation decisions are made in relation to the amount of leaf area on a plant  \citep{Falster-2011}.
-Growth rates in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), total leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) therefore share several terms.
-Many of these terms also vary intrinsically with size, as shown in the insets in Fig.
-\ref{fig:conceptual}b.
-
-The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass growth $\frac{{\rm d}B}{{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
-The remaining $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
-In plants, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants \citep{Wenk-2015}.
-Note also that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
-So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ is not increasing.
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
-
-\caption{Conceptual framework showing how shade-tolerance (as measured by \textsc{wplcp}) is affected by plant size and traits, adapted from \citep{Givnish-1988}.
-From eqn \ref{eq:wplcp}, \textsc{wplcp} is the canopy openness $E$ where the photosynthetic income per unit leaf area (dashed line) intersects the total respiration and turnover costs per unit leaf area (solid black lines).
-Colors indicate the contributions from different tissues to total costs.
-Because the costs of stem tissues increase with height, \textsc{wplcp} also increases.
-Traits {\wood}, {\nitrogen}, and {\lma} can alter \textsc{wplcp} by shifting either income or cost of different tissues (Table \ref{tab:responses}).
-\label{fig:wplcp_idea}}
-\end{figure}
-
-The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
-
-The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood, and roots (eqn \ref{eq:daldmt}).
-The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of construction cost per unit leaf area for different tissues, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$ (eqn \ref{eq:daldmt}).
-
-The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
-This variable accounts for the architectural strategy of the plant (\citep{Falster-2011}).
-Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
-
-The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark, and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
-These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
-
-Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
-We make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}-\ref{eq:htm}).
-
-\subsection{Shade tolerance}
-
-Eqn \ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'' \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
-In general, photosynthesis rate per leaf area $\bar{p}$ increases with canopy openness, $E$.
-The {\wplcp} is then the value $E=E^*$ where photosynthetic gains balance the costs of tissue turnover and respiration, i.e.
-$\frac{{\rm d}B}{{\rm d}t}(E^*)=0$.
-From eqn \ref{eq:dbdt}, this occurs when
-\begin{equation}\label{eq:wplcp}
-\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
-\end{equation}
-Graphically, the {\wplcp} occurs at the point where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig.
-\ref{fig:wplcp_idea}).
-Traits influence {\wplcp} when they affect either photosynthesis, respiration, or turnover.
-Also, as the amount of stem support increases with plant height, {\wplcp} also increases with height \citep{Givnish-1988} (Fig.
-\ref{fig:wplcp_idea}).
-
-\subsection{A functional-balance model for plant construction}
-
-Since eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication, and differentiation, they may apply to a variety of growth model where biomass allocation is important.
-However, to make explicit predictions requires that we quantify all the terms in Fig.
-\ref{fig:conceptual}b via an explicit model of plant construction and function.
-
-The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citep{Makela-1997} and \citep{Moorcroft-2001}.
-We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
-Table \ref{tab:allometry} provides key equations (see \citep{Falster-2016} for full derivation).
-In particular, we assume that as a plant grows:
-1) Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
-2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
-3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
-4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and
-5) The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
-Assumption 1 accounts for the architectural layout of the plant.
-Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
-To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see SI).
-
-Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
-These eqns aso predict that the amount of live stem tissue supporting each unit of leaf area should increase linearly with height,
-\begin{equation}\label{eq:msml}
-\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
-\end{equation}
-Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
-\ref{fig:conceptual}.
-
-\begin{table*}
-\caption{Equations for a ``functional-balance'' model of plant construction. See Table \ref{tab:definitions} for a list of variable names and definitions.}
-\centering
-  \begin{tabular}{p{0.1cm}p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}
-  \toprule
-  & Variable & Function & Marginal cost & Growth rate & Eqn\\
-  \midrule
-  \multicolumn{6}{l}{\textbf{a) Functional-balance assumptions}} \\
-  & Height &
-    $H = \alpha_{\rm l1} \, A_{\rm l}^{0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} = 0.5 \alpha_{\rm l1} A_{\rm l}^{-0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
-  & Sapwood area &
-    $A_{\rm s} = \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
-  & Bark area &
-    $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
-  & Root mass &
-    $M_{\rm r} = \alpha_{\rm r1} \, A_{\rm l}$ &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_{\rm r1}$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
-  & Heartwood area & & &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\
-  \multicolumn{6}{l}{\textbf{b) Derived quantities}} \\
-  & Leaf mass &
-    $M_{\rm l} = \phi \, A_{\rm l} $ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
-  & Sapwood mass &
-    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( H + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
-  & Bark mass &
-    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( H + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
-  & Heartwood area &
-    $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
-  & Heartwood mass &
-    $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:htm} \\
-  \bottomrule
-\end{tabular}
-\addtabletext{\\The first column of \textbf{a)} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of \textbf{a)} and in \textbf{b)} can be derived from these assumptions.
-The column ``Eqn'' indicates equation numbers referred to in the main text.
-
-}
-\label{tab:allometry}
-\end{table*}
-
-\subsection{Trait-based trade-offs}
-
-To capture the effects of functional traits on growth rates and shade tolerance requires that trait-based trade-offs be embedded within the model described above.
-Here we consider five traits, for which we can posit specific physiological costs and benefits -- described in detail below and summarised in Table \ref{tab:traits}.
-It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values.
-In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait.
-
-\subsubsection{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
-
-Seed mass: {\seed} moderates a trade-off between a plant's rate of seed production per unit mass invested in reproduction and the mass of each offspring  \citep{Moles-2006}.
-
-Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
-We assume the amount allocated to vegetative growth starts high amd then decreases once $H > H_{{\rm mat}}$  (Fig.
-\ref{fig:conceptual}b).
-The exact nature of the function is non-critical, what is important is that plants shift from a period of investing mainly in vegetative growth to investing mainly in reproduction.
-Here we let the fraction of mass allocated to reproduction vary with $H$ via the function $r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-
-\subsubsection{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.
-
-Leaf nitrogen per unit leaf area: We assume that both the maximum photosynthetic capacity of a leaf and leaf respiration rate both vary with {\nitrogen} (=$\nu$), as $\beta_{{\rm lf1}} \, \left(\nu / \nu_0 \right)^{\beta_{{\rm lf5}}}$ and $\beta_{{\rm lf4}}\, \nu$ respectively, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf4}}$, and $\beta_{{\rm lf5}}$ are constants.
-
-Leaf mass per unit leaf area: {\lma} (=$\phi$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever {\lma} is varied.
-
-Wood density: {\wood} (=$\rho$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link {\wood} to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\rho / \rho_0\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
-The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of {\wood}, as such the mass-based rate is adjusted accordingly whenever {\wood} is varied.
-There currently little data relating sapwood turnover or respiration to {\wood}.
-
-\section*{Results}
-
-\subsection{Model assumptions}
-
-We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}.
-Fig.
-\ref{fig:assumptions} shows that the three functional-balance assumptions in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
-As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions (Table \ref{tab:allometry}a and eqn \ref{eq:msml}).
-
-\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
-\centering
-\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{main/growth_light_height.pdf}
-\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
-Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}.
-Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
-For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated.
-The white regions indicate trait ranges that are typically observed in real systems.
-Figs.
-\ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass.
-Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
-\label{fig:growth_light_height}}
-\end{SCfigure*}
-
-\subsection{Changes in growth rate with size}
-
-Our model suggests an intrinsically size-dependent pattern of biomass-production and growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
-The panels in Fig.
-\ref{fig:conceptual} show the predicted patterns for a typical woody plant, obtained by applying the functional-balance model from Table \ref{tab:allometry}a.
-Biomass growth shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
-Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
-This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with increasing size, including a strong decline in the fraction of plant that is leaf (Fig.
-\ref{fig:conceptual}), increasing reproductive allocation (Fig.
-\ref{fig:conceptual}), and declining biomass growth.
-In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
-Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
-All growth measures decrease sharply with size when expressed as relative growth rates.
-
-\subsection{Changes in height growth rate with traits}
-
-We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}).
-We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
-Across the five different traits, we observed four relatively distinct types of response.
-These responses are summarised in Table \ref{tab:responses} and described in more detail below.
-
-\subsubsection{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease.
-The only effect of {\seed} on growth comes from changing the plant's initial size.
-The plots in Fig.
-\ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
-Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in all metrics) because of their greater total leaf area.
-At the same time, relative growth rate is predicted to decrease with size, as the ratio of leaf area to support mass decreases.
-As plants grow, differences in initial mass decrease in importance relative to other factors influencing growth through the life-cycle.
-As a result, the correlations between {\seed} and growth rate disappear among larger plants.
-
-Greater {\hmat} causes a growth advantage at larger sizes by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
-\ref{fig:growth_light_height}).
-At smaller sizes, there is no differentiation among species, as all plants focus on growth.
-
-\begin{SCfigure*}[\sidecaptionrelwidth][b]
-\centering
-\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three developmental traits on shade tolerance.}
-Panels show effect of traits on level of canopy openness that causes biomass growth (eqn \ref{eq:dbdt}) to be zero.
-Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
-The white regions indicate trait ranges that are typically observed in real systems.
-\label{fig:wplcp}}
-\end{SCfigure*}
-
-\subsubsection{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
-\ref{fig:growth_light_height}).
-As $E$ increases, the optimal {\nitrogen} also increases.
-The invariance of the growth-trait relationship with respect to size arises as follows.
-The direct physiological effect of {\nitrogen} is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
-Both the cost and benefits of {\nitrogen} appear within the term $\frac{ {\rm d}B}{{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in biomass growth per {\nitrogen}.
-From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
-u} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial
-u}  - \frac{ \partial r_{\rm l}}{\partial
-u} )$. As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.
-
-Unlike {\nitrogen}, the response of growth rate to changes in {\lma} varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
-\ref{fig:growth_light_height}).
-As a result, the value of {\lma} that optimises plant growth increases with height, while the direction of correlation between height growth rate and {\lma} shifts from negative to positive, as plants increase in height.
-Decreasing {\lma} has two impacts on height growth rate.
-First, lower {\lma} increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$.
-Second, lower {\lma} decreases biomass growth $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
-Whether lower {\lma} increases height growth depends on the relative magnitude of these two effects.
-When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate.
-When plants are large, the influence of {\lma} on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
-Thus, at larger sizes low {\lma} no longer increases growth rate (Fig.
-\ref{fig:growth_light_height}).
-
-Reducing the cost of stem construction, via lower {\wood}, decreases the cost of deploying a unit of leaf area, increasing growth rate (Fig.
-\ref{fig:growth_light_height}).
-Unlike {\lma}, the benefits of cheaper stem construction become more pronounced at intermediate sizes, as a larger fraction of the plant is stem (Fig.
-\ref{fig:conceptual}a).
-
-\subsection{Changes in other growth rates with trait}
-
-The results reported above and in Fig.
-\ref{fig:growth_light_height} focus on height growth rate.
-Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
-\ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}.
-For each, plants were grown to a standard diameter, area, or mass.
-Changes in absolute and relative growth rates with traits therefore show similar patterns.
-
-We find that for {\seed}, {\nitrogen}, {\lma}, and {\hmat}, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to height growth (Table \ref{tab:responses}).
-The only trait where a slightly different response was observed was for {\wood}.
-Whereas the effect of {\wood} on height growth tended to diminish slightly at larger sizes (Fig.
-\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area, or above-ground mass.
-Sapwood lost via turnover becomes heartwood.
-Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.
-
-\subsection{Responsiveness of growth rate to light}
-
-The predictions in Figs.
-\ref{fig:growth_light_height} and \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} illustrate how traits impact on growth rate under different light environments and at different sizes.
-An additional outcome that arises directly from these analyses is that traits moderate the responsiveness of growth to changes in light environment.
-This response arises because individuals with higher potential growth rate naturally have greater potential plasticity in growth.
-Our results therefore support findings that species with low {\wood} increase growth more substantially with increases in light (Table \ref{tab:phenomena}).
-Variation in {\lma} also moderates the response of growth to changes in light, with species having the lowest {\lma} being most responsive.
-However, unlike for {\wood}, the effect appears only for the smallest size classes.
-
-\subsection{Shade tolerance}
-Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
-\begin{equation}\label{eq:wplcp_2}
-\bar{p}(E^*) =
-      \phi c_{\rm l} +
-      \left(\theta \, \rho \, \eta_c \, H\right)
-        \left(b c_{\rm b}
-            + c_{\rm s}\right) +
-      \alpha_{\rm r1} c_{\rm r},
-\end{equation}
-where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
-Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}.
-With some further manipulations, it is possible to show that {\wplcp} will decrease with {\lma} if $\beta_{{\rm kl2}} >1$.
-Likewise {\wplcp} will decrease with {\wood} if $\beta_{{\rm ks2}} >1$.
-The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to {\lma} and {\wood}, respectively.
-Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low {\lma} and low {\wood} are predicted to be less shade tolerant (Fig.
-\ref{fig:wplcp}).
-At low {\lma} ({\wood}), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover.
-{\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
-In addition, {\wplcp} varies with {\nitrogen}.
-At small sizes, {\wplcp} increases with {\nitrogen} across the band of values typically observed in real plants, i.e.
-high {\nitrogen} makes seedlings shade intolerant.
-At larger sizes, as biomass growth declines to zero, {\wplcp} begins to increase again for very low values of {\nitrogen}.
-All these results  match empirical patterns (Table \ref{tab:phenomena}).
-
-\newcommand{\sepp}{{\color{grey}/}}
-\newcommand{\upup}{$\uparrow\,\uparrow$}
-\newcommand{\updo}{$\uparrow\,\downarrow$}
-\newcommand{\dodo}{$\downarrow\,\downarrow$}
-\newcommand{\upfl}{$\uparrow$\,--}
-\newcommand{\flup}{--$\,\uparrow$}
-\newcommand{\dofl}{$\downarrow\,$--}
-\newcommand{\doup}{$\downarrow\,\uparrow$}
-\newcommand{\fldo}{--$\,\downarrow$}
-
-\begin{table}[hb!]
-\centering
-\caption{Predicted effects of traits on components of plant function determining growth rate. Adapted and expanded from \citep{Gibert-2016}.}
-  \begin{tabular}{llx{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}p{0.01cm}}
-  \toprule
-  & &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
-  & & {\seed} & $H_{{\rm mat}}$ & {\nitrogen} & {\lma} & {\wood} & \\
-  \midrule
-  \multicolumn{8}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
-  \multicolumn{2}{l} {Biomass growth, $\ud B / \ud t$} & & & & & & \\
-  & \tabitem Photosynthesis & - & - & $\uparrow$  & - & - & \\
-  & \tabitem Respiration & - & - & $\uparrow$  & - & - & \\
-  & \tabitem Turnover & - & - & - & $\uparrow$ & $\uparrow$ & \\
-  \multicolumn{2}{l} {Allocation to growth, $\ud M_{\rm a} / \ud B$} & - & $\uparrow$ & - & - & - & \\
-  \multicolumn{2}{l} {Leaf deployment, $\ud A_{\rm l} / \ud M_{\rm a}$} & & \\
-  & \tabitem Leaf  & - & - & - & $\downarrow$ & - & \\
-  & \tabitem Sapwood & - & - & - & - & $\downarrow$ & \\
-\midrule
- \multicolumn{8}{l} {\textbf{b) Predicted effect of trait on growths rate for a small and large plant}} \\
-  \multicolumn{8}{l} {Height} \\
-  & \tabitem{Absolute}, $\ud H / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Stem area} \\
-  & \tabitem{Absolute}, $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Stem diameter} \\
-  & \tabitem{Absolute}, $\ud D / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Mass} \\
-  & \tabitem{Absolute}, $\ud M_{\rm t} / \ud t$ & \upfl &  \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud M_{\rm t} / (\ud t . M_{\rm t} )$ & \dofl &  \flup & \upup & \dofl & \dodo & \\
-  \bottomrule
-  \end{tabular}
-\addtabletext{Arrows indicate the effect of increased trait value on each component; dashes indicating no effect.
-Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density (see Tables \ref{tab:traits}-\ref{tab:definitions}).
-
-}
-\label{tab:responses}
-\end{table}
-
-\section*{Discussion}
-
-Using a model relating plant physiological function and mass allocation to prominent traits, we have clarified how traits could impact on plant growth across the life cycle.
-Our approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
-During recent decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}.
-However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}.
-Matching growing empirical evidence (Table \ref{tab:phenomena}), this study shows when and why the direction or strength of a correlation between traits and growth rate shifts with plant size.
-Moreover, we show that different traits and trade-offs generate different types of response.
-Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
-Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e.
-lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}.
-The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter, and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
-This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}.
-Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects.
-Our aim is to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
-
-\subsection{LMA doesn't predict growth rates of large plants}
-
-The model presented here extends a widely-used model for seedlings, linking mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
-Importantly, the seedling model (eqn \ref{eq:RGR}) can be derived as a special case of the new, extended model.
-Unlike eqn \ref{eq:RGR}, the extended model also predicts a relationship between {\lma} and growth rate that changes with size.
-In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even weakly positive in larger plants (Fig.
-\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem diameter, stem area or total mass.
-This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size.
-As seedlings, leaves comprise a large part of the plant (Fig.
-\ref{fig:conceptual}a).
-Decreasing {\lma} then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
-As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheap leaf construction.
-Consequently, the effect of {\lma} on leaf turnover comes to dominate at larger sizes, and as such, the effect of {\lma} on height, diameter, and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
-
-\subsection{Generalising to other traits and trade-offs}
-
-The list of functional traits known to differ among plant species is ever-increasing \citep{Perez-2013}.
-While we have focussed on understanding the effects of five traits on elements of demography, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
-The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be connected with variables in eqns \ref{eq:dbdt}--\ref{eq:dDdt}.
-While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
-A well-developed trade-off must include two opposing forces that operate at some point in the life cycle.
-
-Our framework also highlights what is needed for traits to influence demography.
-While traits can influence many aspects of plant function, these influences must operate via the pathways outlined in Fig.
-\ref{fig:conceptual} if the trait is going to impact on growth rate or shade tolerance.
-For example, many studies consider traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
-By altering conductance of water to the leaf, both traits will influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}), while vessel size might also affect rates of stem turnover ($k_{\rm s}$ in eqn \ref{eq:dbdt}).
-The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in via the parameter $\theta$ (Table \ref{tab:allometry}).
-The effect of both these traits on growth rate should be expected to vary with plant size.
-
-\subsection{Implications for trait-based approaches}
-There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.
-
-First, our results highlight the importance of allocation decisions and turnover costs in determining demography.
-Much of current ecosystem research focusses on factors affecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
-Yet, four of the five traits considered here do not influence net primary production.
-In fact, our analysis shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
-Low {\lma} results in high leaf turnover, such that individuals with a low {\lma} have lower biomass growth.
-It is this property that makes them shade intolerant.
-And yet they can still achieve a growth advantage (when small).
-
-Second, our results demand a shift away from viewing species as lying on a spectrum  from slow to fast growth.
-When the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum.
-Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.
-Researchers using functional traits have also tended to describe species as fast or slow growing \citep{Adler-2014, Diaz-2016}.
-Our results suggest a more-nuanced approach.
-Plants that are fast growing as seedlings may not be fast growing as saplings or adults, or under low light.
-Plants that are fast growing as adults may not be fast growing as seedlings.
-This more-nuanced perspective reflects observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
-
-Third, our results suggest that even if traits define a potential growth trajectory, researchers seeking to link traits to growth rate must probe deeper into the data than simply looking for a linear relationship between traits and average growth, to recover the expected relationships.
-None of the predicted relationships between traits and growth is linear across the range of sizes and light environments tested.
-As such, we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits \citep[e.g.][]{Poorter-2008,Paine-2015}.
-Controlling for size, site, and light environment will be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}, as will having a clear expectation for the hypothesised relationship.
-
-\subsection{Plasticity of traits through ontogeny}
-
-While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
-To that end, we make a further prediction arising from our results, which is that the trait {\lma} should increase through ontogeny for all individuals.
-Such shifts have been observed across a variety of species \citep{King-1999,Thomas-1999,Koch-2004}.
-Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low {\lma} -- to increase their {\lma} as they grow larger.
-While a similar prediction was made for a species of \emph{Eucalyptus} \citep{King-1999}, we extend the idea across species.
-Although trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
-This hypothesis attempts to give meaning to some of that variation, and shows how variation across and within species might be understood within a single framework.
-
-\subsection{Comparison with other frameworks}
-
-As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}.
-These models have several properties in common, including that they all have growth being driven by the net amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
-Many models also make functional-balance assumptions, for example linking the cross-section of sapwood to leaf area \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-2005, Moorcroft-2001}.
-We note that an assumption of exact functional balance is not critical for our results, what matters is that the amount of live biomass (i.e.
-excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
-\ref{fig:assumptions}d).
-
-A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs.
-While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have focussed on ecosystem-level outcomes rather than the growth of individual plants.
-Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
-Another study \citep{King-1999} connected a model of growth for a single species to the trait {\lma}, and likewise predicted a gradual flattening out of the relationship between {\lma} and growth rate with size (as in Fig.
-\ref{fig:growth_light_height}).
-Here, we include an additional cost of increased leaf turnover that further penalised low {\lma} strategies when large.
-
-Perhaps the two most controversial elements of our approach concerns the assumptions about tissue replacement and reproductive allocation.
-Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing any surplus.
-This is because we assume tissues lost via turnover are replaced before mass is allocated to either new growth (i.e.
-growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant.
-The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus mass to reproduction.
-While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
-Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}.
-Currently, reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
-For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
-Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a plant's maximum size and growth rate in adult plants \citep[e.g.][]{Wright-2010}.
-
-Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory  of ecology (\textsc{mst}) \citep{Enquist-1999, Enquist-2007}.
-Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
-First, the \textsc{mst} model suggests diameter growth continues to increase wit size, whereas empirical data suggests growth rate declines for larger plants \citep{Canham-2004, Canham-2006, Herault-2011}.
-Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size.
-Predicted effects are for a linear increase in growth with decreases in either {\lma} and lower {\wood}, that apply irrespective of size.
-However, at least for {\lma} such effects in large trees have not been observed.
-
-\subsection{Closing remarks}
-
-We have shown how diverse phenomena related to plant growth can be understood with a model accounting for processes generating photosynthetic income and allocating this among different tissues.
-The need to consider effects of plant size, alongside trait-based differences among species has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
-Here we have provided a framework for achieving this.
-By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid  foundation for understanding and modelling growth across diverse species around the world.
-
-\matmethods{
-\vspace{-1em}
-
-\subsection{Parameters}
-Parameters were sourced mainly from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at {\hmat} (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
-\ref{fig:conceptual}).
-
-\subsection{Assumptions}
-The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
-We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
-We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed Gaussian.
-
-\subsection{Software}
-The growth model used has been implemented as the \texttt{FF16} physiological module within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
-The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
-
-}
-
-\showmatmethods % Display the Materials and Methods section
-
-\acknow{
-We thank J Camac, A Gibert, G Kunstler, H Muller-Landau, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
-DF was supported by fellowships from the  Australian Research Council discovery (DP110102086, FT160100113).
-RF was supported by the Science and Industry Endowment Fund (RP04-174).
-The authors have no conflicts of interest to declare.
-}
-
-% Bibliography
-\bibliography{references}
-%\pnasbreak
 \end{document}

### 2018-05-02 ms/Table3.tex
Separate files for Tables
Requested by PNAS office

--- 2018-04-05 ms/ms.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\documentclass[9pt,twcolumn]{pnas-new}
 
-% Enable new column type -- centered with fixed
-% Following http://texblog.org/2008/05/07/fwd-equal-cell-width-right-and-centre-aligned-content/
-\usepackage{array}
-\newcolumntype{x}[1]{%
->{\raggedleft}p{#1}}%
-
-% Enable equation numbers in Table
-% Following http://tex.stackexchange.com/questions/62974/numbering-equations-in-tabular-environment
-\newcommand{\eqncounter}[1]{%
-\mbox{}\refstepcounter{equation}%
-$[\theequation]$%
-\label{#1}
-}
-
-\input{ms-common-defs}
-
-\templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
-
-% Enable cross referencing to Evidence document
-\usepackage{xr}
-\externaldocument{ms-supp}
-\externaldocument{figure_conceptual}
-
-\title{How functional traits influence plant growth and shade tolerance across the life-cycle}
-
-% Please add here a significance statement to explain the relevance of your work
-% 120 words
-% Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself and is required for all research papers.
-Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself.
-\significancestatement{
-Plant species differ in many functional traits -- features of specific tissues and allocation of energy among them.
-While traits have been used in many correlative approaches to describe communities and demography, it has remained unclear how and why traits should influence whole-plant growth.
-Here we present a new theoretical framework for understanding the effect of traits on plant growth and shade-tolerance.
-This framework captures diverse patterns of growth in relation to size, and explains why the effect of traits on growth changes through ontogeny.
-By disentangling the effects of plant size, light environment, and traits on growth rates, this study provides a theoretical foundation for understanding growth across diverse tree species around the world.
-}
-
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{D.F. designed the study; D.F and R.F performed the analysis. All authors contributed to the writing of this manuscript.}
-
-\authordeclaration{The authors declare that they have no competing financial interests.\\
-
-Code reproducing the results is available at \href{https://github.com/traitecoevo/growth_trajectories}{github.com/traitecoevo/growth\_trajectories}.
-}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: daniel.falster@unsw.edu.au}
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{growth rate $|$ traits $|$ trees $|$ allometry $|$ model}
-
-% 250 words
-\begin{abstract}
-Plant species differ in many functional traits that drive differences in rates of photosynthesis, biomass allocation, and tissue turnover.
-Yet, it remains unclear how -- and even if -- such traits influence whole-plant growth, with the simple linear relationships predicted by existing theory often lacking empirical support.
-Here we present a new theoretical framework for understanding the effect of diverse functional traits on plant growth and shade-tolerance by extending a widely-used model, linking growth rate in seedlings with a single leaf trait, to explicitly include influences of size, light environment, and five other prominent traits: seed mass, height at maturation, leaf mass per unit leaf area, leaf nitrogen per unit leaf area, and wood density.
-Based on biomass growth and allocation, this framework explains why the influence of traits on growth rate and shade tolerance often varies with plant size and why the impact of size on growth varies among traits.
-Specifically, we demonstrate why for height growth the influence of: i) leaf mass per unit leaf area is strong in small plants but weakens with size, ii) leaf nitrogen per unit leaf area does not change with size, iii) wood density is present across sizes, iv) height at maturation strengthens with size, and v) seed mass decreases with size.
-Moreover, we show how traits moderate plant responses to light environment and also determine shade tolerance, supporting diverse empirical results.
-\end{abstract}
-
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\newcommand{\plant}{\texttt{plant}}
+\newcommand{\wplcp}{\textsc{wplcp}}
+\newcommand{\lma}{\textsc{lma}}
+\newcommand{\wood}{\textsc{wd}}
+\newcommand{\seed}{\textsc{sm}}
+\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
+\newcommand{\nitrogen}{\textsc{n}}
 
 \begin{document}
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalized the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-% If your first paragraph (i.e.
-with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
-If this is the case, add \parshape=0 to the end of the list environment.
-
-\dropcap{F}unctional traits capture core differences in the strategies plants use to acquire and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}.
-Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients, and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover.
-During the last two decades, trade-offs related to some prominent traits have been quantified, with values for some leaf, wood, or seed traits now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}.
-As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}.
-In this article we outline the mechanisms by which the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.
-
-While the influence of traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have met with mixed success \citep{Poorter-2006, Poorter-2008,Wright-2010,Herault-2011,Paine-2015}.
-In seedlings, the trait leaf mass per unit leaf area ({\lma}) -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly correlated with relative growth rate in plant mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}, as predicted from a simple mathematical model of growth rate (see below).
-{\lma} and a close correlate leaf lifespan, were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}.
-These early successes prompted researchers to search for similar relationships in large plants.
-However, in saplings and trees showed, {\lma} was not correlated with growth rate \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015}.
-Meanwhile, other traits such as wood density ({\wood}) showed strong relationships to growth in large plants \citep{Wright-2010,Herault-2011}, but less so in small plants \citep{Castro-1998}.
-Clearly, traits do not correlate simply and consistently to growth rate.
-
-Recently, it has become clear that the effect of traits on plant growth can be modified by plant size \citep{Falster-2011, Ruger-2012, Iida-2014, Visser-2016, Gibert-2016}.
-A recent meta-analysis of 103 studies reporting \textgreater 500 correlations provides the most compelling evidence \citep{Gibert-2016}.
-That study showed that the strength of the correlation between some traits (including {\lma} and maximum height) and growth rate was modified by size, while for other traits (including {\wood} and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size.
-Here, we extend previous work \citep{Gibert-2016} to provide a mechanistic explanation of these empirical patterns via a model of plant growth and allocation.
 
 \begin{table}[!ht]
-\caption{Empirical phenomena explained in this paper.}
-\begin{tabular}{p{8cm}}
-\toprule
-\textbf{Change in growth rate with increasing size (Fig. \ref{fig:conceptual})}\\
- $\,$ Biomass growth: Hump-shaped \citep{Givnish-1988, Koch-2004} \\
- $\,$ Plant mass: Increasing \citep{Sillett-2010, Stephenson-2014} \\
- $\,$ Height: Hump-shaped \citep{Ryan-2006, Sillett-2010, King-2011} \\
- $\,$ Stem-diameter: Hump-shaped \citep{Canham-2004, Canham-2006, Herault-2011} \\
- $\,$ Relative growth rate (all variables): decreasing \citep{Rees-2010, Iida-2014}\\
-\textbf{Effect of traits on growth rate (Fig. \ref{fig:growth_light_height})}\\
- $\,$ {\seed}: $\downarrow$ values $\downarrow$ seedling size, \& thus $\downarrow$ absolute \& $\uparrow$ relative growth rate \citep{Gibert-2016} \\
- $\,$ $H_{\rm mat}$: $\downarrow$ values $\downarrow$ growth rate at at larger sizes \citep{Gibert-2016}\\
- $\,$  {\nitrogen}: $\downarrow$ values $\uparrow$ growth rate irrespective of size, but only in high light \citep{Gibert-2016}\\
- $\,$ {\lma}: $\downarrow$ values $\uparrow$ growth rate when small, not at mid-large sizes \citep{Gibert-2016}\\
- $\,$ {\wood}: $\downarrow$ values $\uparrow$ growth rate, except at largest sizes \citep{Gibert-2016}\\
-\textbf{Responsiveness of growth rate to changes in light, $E$ (Fig. \ref{fig:growth_light_height})}\\
- $\,$ {\nitrogen}: $\uparrow$ values respond more \\
- $\,$ {\lma}: $\downarrow$ values respond more \\
- $\,$ {\wood}: $\downarrow$ values respond more \citep{Ruger-2012}\\
-\textbf{Shade tolerance, \textsc{wplcp} (Fig. \ref{fig:wplcp})}\\
- $\,$ Size: decreasing \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}\\
- $\,$ {\nitrogen}: $\downarrow$ values $\uparrow$ shade tolerance \citep{Messier-1999, Craine-2005,Baltzer-2007}\\
- $\,$ {\lma}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Messier-1999, Poorter-2006, Baltzer-2007, Lusk-2008}$^3$ \& $\uparrow$ \textsc{lai} \citep{Reich-1992, Gower-1993, Niinemets-2010} \\
- $\,$ {\wood}: $\downarrow$ values $\downarrow$ shade tolerance \citep{Osunkoya-1996}\\
-  \bottomrule
-\end{tabular}
-\addtabletext{{\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density, \textsc{lai} = leaf area index.
-Footnotes: $^1$ Similar responses are predicted for maximum photosynthetic rate per leaf area \& dark respiration rate per leaf area; here both related to {\nitrogen}.
-$^2$ Similar responses are predicted for leaf lifespan; here directly related to {\lma}.
-$^3$ \citep{Baltzer-2007} finds a relationship between \textsc{wplcp} \& leaf respiration rate expressed per unit leaf area or per unit leaf mass.
-
-}
-\label{tab:phenomena}
-\end{table}
-
-\begin{table}[!hb]
- \caption{Key trade-offs (benefit and cost) for the five traits considered, as encoded into the {\plant} model.
- }
-\centering
-  \begin{tabular}{p{0.5cm}p{3.6cm}p{2.5cm}p{0.4cm}}
-  \toprule
-  Trait &  Benefit & Cost & Ref. \\
-  \midrule
-  {\seed} & $\uparrow$ size of seedlings & $\downarrow$ seed production rate & \citep{Moles-2006}\\
-  {\hmat}  & $\uparrow$ growth rate & $\downarrow$ seed production rate & \citep{Thomas-2011}\\
-  {\nitrogen} & $\uparrow$ photosynthetic rate in high light & $\uparrow$ respiration rate & \citep{Wright-2004}\\
-  {\lma} & $\downarrow$ leaf turnover rate & $\uparrow$ cost of building leaf & \citep{Wright-2004}\\
-  {\wood} & $\downarrow$ sapwood turnover rate & $\uparrow$ cost of building stem & \citep{Chave-2009}\\
-  \bottomrule
-  \end{tabular}
-\addtabletext{Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density.
-
-}
-\label{tab:traits}
-\end{table}
-
-Interpreting diverse empirical results seeking to link traits to growth rate has been challenging because, until recently \citep{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size.
-A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to {\lma} \citep{Lambers-1992, Cornelissen-1996, Wright-2000}.
-An extension of this equation suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}.
-But as noted above, the prediction for large plants is not supported by empirical results.
-Meanwhile, theoretical predictions on how other traits should influence growth are largely absent.
-Without further guidance, many researchers expected traits to map directly onto growth rates \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.
-
-One reason theoretical predictions have been lacking or not been supported is that in existing theory the effects of traits is realised mainly via influences on net primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}.
-By contrast, the physiology of traits such as {\lma} and {\wood} suggests they influence allocation of biomass among different tissues, and their turnover rates, rather than net primary production \citep{Falster-2011, Duursma-2016, Gibert-2016}.
-A second concern for theory focussed on net primary production is that measuring production is really only practical for small plants.
-On larger plants, growth is measured mainly as increment in either stem diameter or height \citep{Purves-2008, Anderson-2015, Kunstler-2016}.
-As outlined below, growth in height and diameter also depend on the allocation of biomass and tissue construction costs.
-
-Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence key elements of plant demography across the life-cycle.
-The {\plant} model builds on past approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}.
-Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}.
-Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size.
-Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size.
-We consider five prominent traits representing specific physiological trade-offs: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).
-
-We also show how our approach can account for related phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment (Table \ref{tab:phenomena}).
-Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny.
-In absolute terms, growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011}, diameter \citep{Canham-2004, Canham-2006, Herault-2011} or biomass growth \citep{Givnish-1988, Koch-2004}.
-In contrast, the growth rate of standing plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}.
-Expressed as relative growth rates, all growth measures decrease sharply with size \citep{Rees-2010, Iida-2014}.
-Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}.
-These diverse phenomena deserve an integrated explanation.
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=\linewidth]{main/figure_conceptual.pdf}
-\caption{Conceptual framework linking growth rate to plant size and traits.
-\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
-\textbf{b)} Equations describing the rates of biomass growth, and growth in various dimensions of the plant.
-In the first line the symbol $\Sigma$ means ``sum'' across tissues, where $i = {\rm l,b,s,r}$.
-Grey numbers indicate equation numbers referred to in the main text.
-The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}.
-Colours highlight where the same metric appears repeatedly in different equations.
-For a list of variable names see Table \ref{tab:definitions}.}
-\label{fig:conceptual}
-\end{figure}
-
-\section*{Modeling the effects of traits on demography}
-
-The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig.
-\ref{fig:conceptual}a).
-We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive, t=total, and st=stem total (s+b+h).
-The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$.
-A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.
-
-Applying a standard approach, the amount of biomass growth on the plant, $\frac{{\rm d}B}{{\rm d}t}$, is given by the difference between income (photosynthesis) and losses (respiration and turnover)\citep{Makela-1997, Thornley-2000}:
-\begin{equation}\label{eq:dbdt}
-\frac{{\rm d}B}{{\rm d}t}
-= \alpha_{\rm bio} \,
-\alpha_{\rm y}
-\bigg( A_{\rm l} \, \bar{p}(E) -
-\, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}\bigg)
-- \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}.
-\end{equation}
-Photosynthesis is the product of the average photosynthetic rate per unit leaf area, $\bar{p}(E)$, and total leaf area, $A_{\rm l}$.
-We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (see SI for details), and respiration and turnover rates of different tissues are constants that potentially differ with traits.
-The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for the yield (fraction of C fixed in biomass) and the conversion of CO$_2$ into units of biomass, respectively.
-While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment so that we can better understand trait- and size-related effects.
-Many vegetation models also use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$ -- e.g.
-as functions of temperature -- though such detail will not qualitatively alter model behaviour.
-
-\begin{table}[!hb]
  \caption{Variable definitions.}
@@
 \end{tabular}
+
 \addtabletext{For $M_i$, $r_i$, $k_i$ and \(A_i\), subscripts refer to: \({\rm l}\) = leaves, \({\rm b}\) = bark, \({\rm s}\) = sapwood,  \({\rm h}\) = heartwood, \({\rm r}\) = roots, \({\rm st}\) = total stem, \({\rm a}\) =  alive tissue.
@@
 
-\subsection{Classic model for mass-based relative growth rate}
-
-Earlier studies on seedlings related relative growth rate in mass to {\lma} \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}, using a model that can be derived from eqn \ref{eq:dbdt} as a special case of the more-extended model described below.
-Ignoring all turnover terms, as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt} and reproduction, biomass growth becomes a linear function of leaf area, making relative growth rate in mass:
-\begin{equation}\label{eq:RGR}
-\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx P_{\rm net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
-where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$, and $\phi$ is {\lma}.
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to {\lma} \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
-The derivations below clarify these links.
-
-\subsection{Decomposition of growth rates into components}
-
-% Update equation counter, following inclusion of figure 1
-\setcounter{equation}{9}
-
-To model growth in plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for biomass growth, but also for the costs of building new tissues and allocation to reproduction.
-Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig.
-\ref{fig:conceptual}b, eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
-While other decompositions are possible, the approach used here arises from the idea that multiple aspects of function and allocation decisions are made in relation to the amount of leaf area on a plant  \citep{Falster-2011}.
-Growth rates in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), total leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) therefore share several terms.
-Many of these terms also vary intrinsically with size, as shown in the insets in Fig.
-\ref{fig:conceptual}b.
-
-The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass growth $\frac{{\rm d}B}{{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
-The remaining $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
-In plants, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants \citep{Wenk-2015}.
-Note also that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover.
-So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ is not increasing.
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}
-
-\caption{Conceptual framework showing how shade-tolerance (as measured by \textsc{wplcp}) is affected by plant size and traits, adapted from \citep{Givnish-1988}.
-From eqn \ref{eq:wplcp}, \textsc{wplcp} is the canopy openness $E$ where the photosynthetic income per unit leaf area (dashed line) intersects the total respiration and turnover costs per unit leaf area (solid black lines).
-Colors indicate the contributions from different tissues to total costs.
-Because the costs of stem tissues increase with height, \textsc{wplcp} also increases.
-Traits {\wood}, {\nitrogen}, and {\lma} can alter \textsc{wplcp} by shifting either income or cost of different tissues (Table \ref{tab:responses}).
-\label{fig:wplcp_idea}}
-\end{figure}
-
-The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
-
-The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood, and roots (eqn \ref{eq:daldmt}).
-The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of construction cost per unit leaf area for different tissues, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$ (eqn \ref{eq:daldmt}).
-
-The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
-This variable accounts for the architectural strategy of the plant (\citep{Falster-2011}).
-Some species tend to leaf out more than grow tall, while other species emphasise vertical extension \citep{Poorter-2006}.
-
-The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark, and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
-These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
-
-Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
-We make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}-\ref{eq:htm}).
-
-\subsection{Shade tolerance}
-
-Eqn \ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'' \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
-In general, photosynthesis rate per leaf area $\bar{p}$ increases with canopy openness, $E$.
-The {\wplcp} is then the value $E=E^*$ where photosynthetic gains balance the costs of tissue turnover and respiration, i.e.
-$\frac{{\rm d}B}{{\rm d}t}(E^*)=0$.
-From eqn \ref{eq:dbdt}, this occurs when
-\begin{equation}\label{eq:wplcp}
-\bar{p}(E^*) =\frac{\sum_{i={\rm l,b,s,r}} \, M_{\rm i} \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)}{A_{\rm l}}.
-\end{equation}
-Graphically, the {\wplcp} occurs at the point where the photosynthetic production (per unit leaf area) line intersects with the sum of maintenance and respiration costs (per unit leaf area) for each tissue (Fig.
-\ref{fig:wplcp_idea}).
-Traits influence {\wplcp} when they affect either photosynthesis, respiration, or turnover.
-Also, as the amount of stem support increases with plant height, {\wplcp} also increases with height \citep{Givnish-1988} (Fig.
-\ref{fig:wplcp_idea}).
-
-\subsection{A functional-balance model for plant construction}
-
-Since eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication, and differentiation, they may apply to a variety of growth model where biomass allocation is important.
-However, to make explicit predictions requires that we quantify all the terms in Fig.
-\ref{fig:conceptual}b via an explicit model of plant construction and function.
-
-The {\plant} package adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citep{Makela-1997} and \citep{Moorcroft-2001}.
-We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}.
-Table \ref{tab:allometry} provides key equations (see \citep{Falster-2016} for full derivation).
-In particular, we assume that as a plant grows:
-1) Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
-2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
-3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
-4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and
-5) The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
-Assumption 1 accounts for the architectural layout of the plant.
-Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area.
-To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citep{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see SI).
-
-Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
-These eqns aso predict that the amount of live stem tissue supporting each unit of leaf area should increase linearly with height,
-\begin{equation}\label{eq:msml}
-\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
-\end{equation}
-Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
-\ref{fig:conceptual}.
-
-\begin{table*}
-\caption{Equations for a ``functional-balance'' model of plant construction. See Table \ref{tab:definitions} for a list of variable names and definitions.}
-\centering
-  \begin{tabular}{p{0.1cm}p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}
-  \toprule
-  & Variable & Function & Marginal cost & Growth rate & Eqn\\
-  \midrule
-  \multicolumn{6}{l}{\textbf{a) Functional-balance assumptions}} \\
-  & Height &
-    $H = \alpha_{\rm l1} \, A_{\rm l}^{0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} = 0.5 \alpha_{\rm l1} A_{\rm l}^{-0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
-  & Sapwood area &
-    $A_{\rm s} = \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
-  & Bark area &
-    $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
-  & Root mass &
-    $M_{\rm r} = \alpha_{\rm r1} \, A_{\rm l}$ &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_{\rm r1}$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
-  & Heartwood area & & &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\
-  \multicolumn{6}{l}{\textbf{b) Derived quantities}} \\
-  & Leaf mass &
-    $M_{\rm l} = \phi \, A_{\rm l} $ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
-  & Sapwood mass &
-    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( H + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
-  & Bark mass &
-    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, H $ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( H + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
-  & Heartwood area &
-    $A_{\rm h} = \int_0^t \frac{{\rm d}A_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
-  & Heartwood mass &
-    $M_{\rm h} = \int_0^t \frac{{\rm d}M_{\rm h}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
-     &
-    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:htm} \\
-  \bottomrule
-\end{tabular}
-\addtabletext{\\The first column of \textbf{a)} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of \textbf{a)} and in \textbf{b)} can be derived from these assumptions.
-The column ``Eqn'' indicates equation numbers referred to in the main text.
-
-}
-\label{tab:allometry}
-\end{table*}
-
-\subsection{Trait-based trade-offs}
-
-To capture the effects of functional traits on growth rates and shade tolerance requires that trait-based trade-offs be embedded within the model described above.
-Here we consider five traits, for which we can posit specific physiological costs and benefits -- described in detail below and summarised in Table \ref{tab:traits}.
-It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values.
-In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait.
-
-\subsubsection{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.
-
-Seed mass: {\seed} moderates a trade-off between a plant's rate of seed production per unit mass invested in reproduction and the mass of each offspring  \citep{Moles-2006}.
-
-Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
-We assume the amount allocated to vegetative growth starts high amd then decreases once $H > H_{{\rm mat}}$  (Fig.
-\ref{fig:conceptual}b).
-The exact nature of the function is non-critical, what is important is that plants shift from a period of investing mainly in vegetative growth to investing mainly in reproduction.
-Here we let the fraction of mass allocated to reproduction vary with $H$ via the function $r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
-
-\subsubsection{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.
-
-Leaf nitrogen per unit leaf area: We assume that both the maximum photosynthetic capacity of a leaf and leaf respiration rate both vary with {\nitrogen} (=$\nu$), as $\beta_{{\rm lf1}} \, \left(\nu / \nu_0 \right)^{\beta_{{\rm lf5}}}$ and $\beta_{{\rm lf4}}\, \nu$ respectively, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf4}}$, and $\beta_{{\rm lf5}}$ are constants.
-
-Leaf mass per unit leaf area: {\lma} (=$\phi$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
-The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever {\lma} is varied.
-
-Wood density: {\wood} (=$\rho$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}).
-In addition, we link {\wood} to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\rho / \rho_0\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants.
-The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of {\wood}, as such the mass-based rate is adjusted accordingly whenever {\wood} is varied.
-There currently little data relating sapwood turnover or respiration to {\wood}.
-
-\section*{Results}
-
-\subsection{Model assumptions}
-
-We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}.
-Fig.
-\ref{fig:assumptions} shows that the three functional-balance assumptions in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
-As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions (Table \ref{tab:allometry}a and eqn \ref{eq:msml}).
-
-\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
-\centering
-\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{main/growth_light_height.pdf}
-\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
-Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}.
-Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$).
-For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated.
-The white regions indicate trait ranges that are typically observed in real systems.
-Figs.
-\ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass.
-Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
-\label{fig:growth_light_height}}
-\end{SCfigure*}
-
-\subsection{Changes in growth rate with size}
-
-Our model suggests an intrinsically size-dependent pattern of biomass-production and growth, which aligns with well-known empirical patterns (Table \ref{tab:phenomena}).
-The panels in Fig.
-\ref{fig:conceptual} show the predicted patterns for a typical woody plant, obtained by applying the functional-balance model from Table \ref{tab:allometry}a.
-Biomass growth shows a hump-shaped pattern with size, decreasing at larger sizes as the turnover and respiration of sapwood and bark increase.
-Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
-This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with increasing size, including a strong decline in the fraction of plant that is leaf (Fig.
-\ref{fig:conceptual}), increasing reproductive allocation (Fig.
-\ref{fig:conceptual}), and declining biomass growth.
-In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
-Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
-All growth measures decrease sharply with size when expressed as relative growth rates.
-
-\subsection{Changes in height growth rate with traits}
-
-We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}).
-We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter.
-Across the five different traits, we observed four relatively distinct types of response.
-These responses are summarised in Table \ref{tab:responses} and described in more detail below.
-
-\subsubsection{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease.
-The only effect of {\seed} on growth comes from changing the plant's initial size.
-The plots in Fig.
-\ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size.
-Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in all metrics) because of their greater total leaf area.
-At the same time, relative growth rate is predicted to decrease with size, as the ratio of leaf area to support mass decreases.
-As plants grow, differences in initial mass decrease in importance relative to other factors influencing growth through the life-cycle.
-As a result, the correlations between {\seed} and growth rate disappear among larger plants.
-
-Greater {\hmat} causes a growth advantage at larger sizes by increasing $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ (Fig.
-\ref{fig:growth_light_height}).
-At smaller sizes, there is no differentiation among species, as all plants focus on growth.
-
-\begin{SCfigure*}[\sidecaptionrelwidth][b]
-\centering
-\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
-\caption{\textbf{Effect of three developmental traits on shade tolerance.}
-Panels show effect of traits on level of canopy openness that causes biomass growth (eqn \ref{eq:dbdt}) to be zero.
-Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
-The white regions indicate trait ranges that are typically observed in real systems.
-\label{fig:wplcp}}
-\end{SCfigure*}
-
-\subsubsection{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig.
-\ref{fig:growth_light_height}).
-As $E$ increases, the optimal {\nitrogen} also increases.
-The invariance of the growth-trait relationship with respect to size arises as follows.
-The direct physiological effect of {\nitrogen} is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate.
-Both the cost and benefits of {\nitrogen} appear within the term $\frac{ {\rm d}B}{{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in biomass growth per {\nitrogen}.
-From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
-u} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial
-u}  - \frac{ \partial r_{\rm l}}{\partial
-u} )$. As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.
-
-Unlike {\nitrogen}, the response of growth rate to changes in {\lma} varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig.
-\ref{fig:growth_light_height}).
-As a result, the value of {\lma} that optimises plant growth increases with height, while the direction of correlation between height growth rate and {\lma} shifts from negative to positive, as plants increase in height.
-Decreasing {\lma} has two impacts on height growth rate.
-First, lower {\lma} increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$.
-Second, lower {\lma} decreases biomass growth $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover.
-Whether lower {\lma} increases height growth depends on the relative magnitude of these two effects.
-When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate.
-When plants are large, the influence of {\lma} on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
-Thus, at larger sizes low {\lma} no longer increases growth rate (Fig.
-\ref{fig:growth_light_height}).
-
-Reducing the cost of stem construction, via lower {\wood}, decreases the cost of deploying a unit of leaf area, increasing growth rate (Fig.
-\ref{fig:growth_light_height}).
-Unlike {\lma}, the benefits of cheaper stem construction become more pronounced at intermediate sizes, as a larger fraction of the plant is stem (Fig.
-\ref{fig:conceptual}a).
-
-\subsection{Changes in other growth rates with trait}
-
-The results reported above and in Fig.
-\ref{fig:growth_light_height} focus on height growth rate.
-Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs.
-\ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}.
-For each, plants were grown to a standard diameter, area, or mass.
-Changes in absolute and relative growth rates with traits therefore show similar patterns.
-
-We find that for {\seed}, {\nitrogen}, {\lma}, and {\hmat}, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to height growth (Table \ref{tab:responses}).
-The only trait where a slightly different response was observed was for {\wood}.
-Whereas the effect of {\wood} on height growth tended to diminish slightly at larger sizes (Fig.
-\ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area, or above-ground mass.
-Sapwood lost via turnover becomes heartwood.
-Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.
-
-\subsection{Responsiveness of growth rate to light}
-
-The predictions in Figs.
-\ref{fig:growth_light_height} and \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} illustrate how traits impact on growth rate under different light environments and at different sizes.
-An additional outcome that arises directly from these analyses is that traits moderate the responsiveness of growth to changes in light environment.
-This response arises because individuals with higher potential growth rate naturally have greater potential plasticity in growth.
-Our results therefore support findings that species with low {\wood} increase growth more substantially with increases in light (Table \ref{tab:phenomena}).
-Variation in {\lma} also moderates the response of growth to changes in light, with species having the lowest {\lma} being most responsive.
-However, unlike for {\wood}, the effect appears only for the smallest size classes.
-
-\subsection{Shade tolerance}
-Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
-\begin{equation}\label{eq:wplcp_2}
-\bar{p}(E^*) =
-      \phi c_{\rm l} +
-      \left(\theta \, \rho \, \eta_c \, H\right)
-        \left(b c_{\rm b}
-            + c_{\rm s}\right) +
-      \alpha_{\rm r1} c_{\rm r},
-\end{equation}
-where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
-Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}.
-With some further manipulations, it is possible to show that {\wplcp} will decrease with {\lma} if $\beta_{{\rm kl2}} >1$.
-Likewise {\wplcp} will decrease with {\wood} if $\beta_{{\rm ks2}} >1$.
-The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to {\lma} and {\wood}, respectively.
-Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low {\lma} and low {\wood} are predicted to be less shade tolerant (Fig.
-\ref{fig:wplcp}).
-At low {\lma} ({\wood}), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover.
-{\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}.
-In addition, {\wplcp} varies with {\nitrogen}.
-At small sizes, {\wplcp} increases with {\nitrogen} across the band of values typically observed in real plants, i.e.
-high {\nitrogen} makes seedlings shade intolerant.
-At larger sizes, as biomass growth declines to zero, {\wplcp} begins to increase again for very low values of {\nitrogen}.
-All these results  match empirical patterns (Table \ref{tab:phenomena}).
-
-\newcommand{\sepp}{{\color{grey}/}}
-\newcommand{\upup}{$\uparrow\,\uparrow$}
-\newcommand{\updo}{$\uparrow\,\downarrow$}
-\newcommand{\dodo}{$\downarrow\,\downarrow$}
-\newcommand{\upfl}{$\uparrow$\,--}
-\newcommand{\flup}{--$\,\uparrow$}
-\newcommand{\dofl}{$\downarrow\,$--}
-\newcommand{\doup}{$\downarrow\,\uparrow$}
-\newcommand{\fldo}{--$\,\downarrow$}
-
-\begin{table}[hb!]
-\centering
-\caption{Predicted effects of traits on components of plant function determining growth rate. Adapted and expanded from \citep{Gibert-2016}.}
-  \begin{tabular}{llx{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}x{0.4cm}p{0.01cm}}
-  \toprule
-  & &  \multicolumn{2}{c}{\bf Ontogenetic} & \multicolumn{3}{c}{\bf Development}& \\
-  & & {\seed} & $H_{{\rm mat}}$ & {\nitrogen} & {\lma} & {\wood} & \\
-  \midrule
-  \multicolumn{8}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
-  \multicolumn{2}{l} {Biomass growth, $\ud B / \ud t$} & & & & & & \\
-  & \tabitem Photosynthesis & - & - & $\uparrow$  & - & - & \\
-  & \tabitem Respiration & - & - & $\uparrow$  & - & - & \\
-  & \tabitem Turnover & - & - & - & $\uparrow$ & $\uparrow$ & \\
-  \multicolumn{2}{l} {Allocation to growth, $\ud M_{\rm a} / \ud B$} & - & $\uparrow$ & - & - & - & \\
-  \multicolumn{2}{l} {Leaf deployment, $\ud A_{\rm l} / \ud M_{\rm a}$} & & \\
-  & \tabitem Leaf  & - & - & - & $\downarrow$ & - & \\
-  & \tabitem Sapwood & - & - & - & - & $\downarrow$ & \\
-\midrule
- \multicolumn{8}{l} {\textbf{b) Predicted effect of trait on growths rate for a small and large plant}} \\
-  \multicolumn{8}{l} {Height} \\
-  & \tabitem{Absolute}, $\ud H / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud H / (\ud t . H)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Stem area} \\
-  & \tabitem{Absolute}, $\ud A_{\rm st} / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud A_{\rm st} / (\ud t .A_{\rm st})$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Stem diameter} \\
-  & \tabitem{Absolute}, $\ud D / \ud t$ & \upfl & \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud D / (\ud t . D)$ & \dofl & \flup & \upup & \dofl & \dodo & \\
-  \multicolumn{8}{l} {Mass} \\
-  & \tabitem{Absolute}, $\ud M_{\rm t} / \ud t$ & \upfl &  \flup & \upup & \dofl & \dodo & \\
-  & \tabitem{Relative}, $\ud M_{\rm t} / (\ud t . M_{\rm t} )$ & \dofl &  \flup & \upup & \dofl & \dodo & \\
-  \bottomrule
-  \end{tabular}
-\addtabletext{Arrows indicate the effect of increased trait value on each component; dashes indicating no effect.
-Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density (see Tables \ref{tab:traits}-\ref{tab:definitions}).
-
-}
-\label{tab:responses}
-\end{table}
-
-\section*{Discussion}
-
-Using a model relating plant physiological function and mass allocation to prominent traits, we have clarified how traits could impact on plant growth across the life cycle.
-Our approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits.
-During recent decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}.
-However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}.
-Matching growing empirical evidence (Table \ref{tab:phenomena}), this study shows when and why the direction or strength of a correlation between traits and growth rate shifts with plant size.
-Moreover, we show that different traits and trade-offs generate different types of response.
-Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}.
-Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e.
-lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}.
-The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter, and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}.
-This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}.
-Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects.
-Our aim is to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.
-
-\subsection{LMA doesn't predict growth rates of large plants}
-
-The model presented here extends a widely-used model for seedlings, linking mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
-Importantly, the seedling model (eqn \ref{eq:RGR}) can be derived as a special case of the new, extended model.
-Unlike eqn \ref{eq:RGR}, the extended model also predicts a relationship between {\lma} and growth rate that changes with size.
-In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even weakly positive in larger plants (Fig.
-\ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem diameter, stem area or total mass.
-This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size.
-As seedlings, leaves comprise a large part of the plant (Fig.
-\ref{fig:conceptual}a).
-Decreasing {\lma} then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
-As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheap leaf construction.
-Consequently, the effect of {\lma} on leaf turnover comes to dominate at larger sizes, and as such, the effect of {\lma} on height, diameter, and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
-
-\subsection{Generalising to other traits and trade-offs}
-
-The list of functional traits known to differ among plant species is ever-increasing \citep{Perez-2013}.
-While we have focussed on understanding the effects of five traits on elements of demography, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
-The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be connected with variables in eqns \ref{eq:dbdt}--\ref{eq:dDdt}.
-While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
-A well-developed trade-off must include two opposing forces that operate at some point in the life cycle.
-
-Our framework also highlights what is needed for traits to influence demography.
-While traits can influence many aspects of plant function, these influences must operate via the pathways outlined in Fig.
-\ref{fig:conceptual} if the trait is going to impact on growth rate or shade tolerance.
-For example, many studies consider traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
-By altering conductance of water to the leaf, both traits will influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}), while vessel size might also affect rates of stem turnover ($k_{\rm s}$ in eqn \ref{eq:dbdt}).
-The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in via the parameter $\theta$ (Table \ref{tab:allometry}).
-The effect of both these traits on growth rate should be expected to vary with plant size.
-
-\subsection{Implications for trait-based approaches}
-There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.
-
-First, our results highlight the importance of allocation decisions and turnover costs in determining demography.
-Much of current ecosystem research focusses on factors affecting primary production -- photosynthesis, respiration, and resultant fluxes of carbon -- with less attention devoted to allocation and turnover (\citealp{Friend-2014}; \citealp[for comparisons of models see][]{Sitch-2008, DeKauwe-2014}).
-Yet, four of the five traits considered here do not influence net primary production.
-In fact, our analysis shows that increased growth rate can occur even at a distinct cost to the plant's carbon budget.
-Low {\lma} results in high leaf turnover, such that individuals with a low {\lma} have lower biomass growth.
-It is this property that makes them shade intolerant.
-And yet they can still achieve a growth advantage (when small).
-
-Second, our results demand a shift away from viewing species as lying on a spectrum  from slow to fast growth.
-When the ranking of growth rates among individuals differing in traits shifts with either plant size or light environment, it is not possible to describe a species via a single point along a spectrum.
-Such a spectrum is implied by many of theoretical models used in community ecology, including Grime's \textsc{csr} triangle, the r-K spectrum, and coexistence models base on the Lotka-Volterra system of equations \citep[e.g.][]{Grime-1977,Chesson-2000}.
-Researchers using functional traits have also tended to describe species as fast or slow growing \citep{Adler-2014, Diaz-2016}.
-Our results suggest a more-nuanced approach.
-Plants that are fast growing as seedlings may not be fast growing as saplings or adults, or under low light.
-Plants that are fast growing as adults may not be fast growing as seedlings.
-This more-nuanced perspective reflects observed demographic patterns, where juvenile and adult growth rates are sometimes only loosely correlated \citep{Rees-2001}.
-
-Third, our results suggest that even if traits define a potential growth trajectory, researchers seeking to link traits to growth rate must probe deeper into the data than simply looking for a linear relationship between traits and average growth, to recover the expected relationships.
-None of the predicted relationships between traits and growth is linear across the range of sizes and light environments tested.
-As such, we should not be surprised if the mean growth rate across individuals spanning a range of sizes or light environments is only weakly or not correlated with traits \citep[e.g.][]{Poorter-2008,Paine-2015}.
-Controlling for size, site, and light environment will be essential for detecting significant patterns \citep[e.g.][]{Gibert-2016}, as will having a clear expectation for the hypothesised relationship.
-
-\subsection{Plasticity of traits through ontogeny}
-
-While our theory has succeeded in explaining some observed phenomena (Table \ref{tab:phenomena}), the test for good theory is that it also makes new predictions that enable the theory to be further refined and tested.
-To that end, we make a further prediction arising from our results, which is that the trait {\lma} should increase through ontogeny for all individuals.
-Such shifts have been observed across a variety of species \citep{King-1999,Thomas-1999,Koch-2004}.
-Since the value of cheap leaf construction diminishes with size, it pays for plants -- and especially those with low {\lma} -- to increase their {\lma} as they grow larger.
-While a similar prediction was made for a species of \emph{Eucalyptus} \citep{King-1999}, we extend the idea across species.
-Although trait-based research largely focusses on differences among species, it has long been recognised that traits also vary among individuals within a species and within individuals \citep{Westoby-2002}.
-This hypothesis attempts to give meaning to some of that variation, and shows how variation across and within species might be understood within a single framework.
-
-\subsection{Comparison with other frameworks}
-
-As noted above, the {\plant} model is closely related to models used in several other studies, including those by \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-1999, King-2005, Moorcroft-2001, Li-2014}.
-These models have several properties in common, including that they all have growth being driven by the net amount of photosynthetic income; have photosynthesis increasing non-linearly with light and leaf nitrogen content; and that they consider the costs of respiration and turnover in different tissues.
-Many models also make functional-balance assumptions, for example linking the cross-section of sapwood to leaf area \citep{Givnish-1988, Yokozawa-1995, Makela-1997, King-2005, Moorcroft-2001}.
-We note that an assumption of exact functional balance is not critical for our results, what matters is that the amount of live biomass (i.e.
-excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig.
-\ref{fig:assumptions}d).
-
-A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs.
-While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have focussed on ecosystem-level outcomes rather than the growth of individual plants.
-Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model.
-Another study \citep{King-1999} connected a model of growth for a single species to the trait {\lma}, and likewise predicted a gradual flattening out of the relationship between {\lma} and growth rate with size (as in Fig.
-\ref{fig:growth_light_height}).
-Here, we include an additional cost of increased leaf turnover that further penalised low {\lma} strategies when large.
-
-Perhaps the two most controversial elements of our approach concerns the assumptions about tissue replacement and reproductive allocation.
-Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing any surplus.
-This is because we assume tissues lost via turnover are replaced before mass is allocated to either new growth (i.e.
-growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}.
-This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant.
-The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus mass to reproduction.
-While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}.
-Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}.
-Currently, reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}.
-For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size.
-Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a plant's maximum size and growth rate in adult plants \citep[e.g.][]{Wright-2010}.
-
-Another class of model dealing explicitly with size-related effects are those derived from the metabolic scaling theory  of ecology (\textsc{mst}) \citep{Enquist-1999, Enquist-2007}.
-Several points suggest our new framework provides a better explanation to the growth phenomena in Table \ref{tab:phenomena} than the \textsc{mst} framework.
-First, the \textsc{mst} model suggests diameter growth continues to increase wit size, whereas empirical data suggests growth rate declines for larger plants \citep{Canham-2004, Canham-2006, Herault-2011}.
-Second, the \textsc{mst} model does not allow for the effects of traits to vary with plant size.
-Predicted effects are for a linear increase in growth with decreases in either {\lma} and lower {\wood}, that apply irrespective of size.
-However, at least for {\lma} such effects in large trees have not been observed.
-
-\subsection{Closing remarks}
-
-We have shown how diverse phenomena related to plant growth can be understood with a model accounting for processes generating photosynthetic income and allocating this among different tissues.
-The need to consider effects of plant size, alongside trait-based differences among species has has long been recognised in trait-based research \cite[e.g.][]{Givnish-1988, Thomas-1999, Moorcroft-2001, Westoby-2002, Enquist-2007}.
-Here we have provided a framework for achieving this.
-By disentangling the effects of plant size, light environment and traits on growth rates, our results provide a solid  foundation for understanding and modelling growth across diverse species around the world.
-
-\matmethods{
-\vspace{-1em}
-
-\subsection{Parameters}
-Parameters were sourced mainly from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values).
-The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at {\hmat} (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig.
-\ref{fig:conceptual}).
-
-\subsection{Assumptions}
-The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
-We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
-We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed Gaussian.
-
-\subsection{Software}
-The growth model used has been implemented as the \texttt{FF16} physiological module within the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}.
-The {\plant} package also makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}.
-To encode the trait-based trade-offs described above, we use the capacity in {\plant} to provide a ``hyper-parameterisation'' function, which enables various parameters to covary with traits (for details see SI).
-
-}
-
-\showmatmethods % Display the Materials and Methods section
-
-\acknow{
-We thank J Camac, A Gibert, G Kunstler, H Muller-Landau, C Prentice, E Wenk, M Westoby, I Wright, and SJ Wright for helpful discussions; J Camac for introducing us to the \texttt{stan} framework, and D Warton for discussions about {\sma} line fitting.
-DF was supported by fellowships from the  Australian Research Council discovery (DP110102086, FT160100113).
-RF was supported by the Science and Industry Endowment Fund (RP04-174).
-The authors have no conflicts of interest to declare.
-}
-
-% Bibliography
-\bibliography{references}
-%\pnasbreak
 \end{document}

### 2018-05-17 ms/ms.tex
Change labelling of equations in Figures and Tables
journal requested using numbers 1-n for equations in Fig 1 and Table 4

In the main text I have hard-coded the numbers

@@
 >{\raggedleft}p{#1}}%
-
-% Enable equation numbers in Table
-% Following http://tex.stackexchange.com/questions/62974/numbering-equations-in-tabular-environment
-\newcommand{\eqncounter}[1]{%
-\mbox{}\refstepcounter{equation}%
-$[\theequation]$%
-\label{#1}
-}
 
@@
 
-Earlier studies on seedlings related relative growth rate in mass to {\lma} \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}, using a model that can be derived from eqn \ref{eq:dbdt} as a special case of the more-extended model described below.
-Ignoring all turnover terms, as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt} and reproduction, biomass growth becomes a linear function of leaf area, making relative growth rate in mass:
+Earlier studies on seedlings related relative growth rate in mass to {\lma} \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}, using a model that can be derived from Eq.
+\ref{eq:dbdt} as a special case of the more-extended model described below.
+Ignoring all turnover terms, as well as the respiration terms for non-leaf tissues in Eq.
+\ref{eq:dbdt} and reproduction, biomass growth becomes a linear function of leaf area, making relative growth rate in mass:
 \begin{equation}\label{eq:RGR}
@@
 where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$, and $\phi$ is {\lma}.
-Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to {\lma} \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
+Although Eq.
+\ref{eq:RGR} captures patterns of growth in seedlings in relation to {\lma} \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees such as height or stem diameter.
 The derivations below clarify these links.
@@
 
-\refstepcounter{equation}\label{eq:dbdt_2}
-\refstepcounter{equation}\label{eq:dMtdt}
-\refstepcounter{equation}\label{eq:daldmt}
-\refstepcounter{equation}\label{eq:daldt}
-\refstepcounter{equation}\label{eq:dhdt}
-\refstepcounter{equation}\label{eq:dast}
-\refstepcounter{equation}\label{eq:dDdt}
-
-% Update equation counter, following inclusion of figure 1
-% \setcounter{equation}{9}
-
 To model growth in plant height ($H$), leaf area ($A_{\rm l}$), basal stem area ($ A_{\rm st}$), standing mass ($M_{\rm t}$) or stem diameter ($D$) requires that we account not just for biomass growth, but also for the costs of building new tissues and allocation to reproduction.
 Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig.
-\ref{fig:conceptual}b, eqns \ref{eq:dMtdt}-\ref{eq:dDdt}).
+\ref{fig:conceptual}b, Eqs.
+{\color{pnasbluetext}2}-{\color{pnasbluetext}7}).
 While other decompositions are possible, the approach used here arises from the idea that multiple aspects of function and allocation decisions are made in relation to the amount of leaf area on a plant  \citep{Falster-2011}.
-Growth rates in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; eqn \ref{eq:dMtdt}), total leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; eqn \ref{eq:daldt}), height ($\frac{\ud H}{\ud t}$; eqn \ref{eq:dhdt}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; eqn \ref{eq:dast}), and stem diameter ($\frac{\ud D}{\ud t}$; eqn \ref{eq:dDdt}) therefore share several terms.
+Growth rates in plant weight ($\frac{\ud M_{\rm t} }{\ud t}$; Fig.
+\ref{fig:conceptual}b, Eqs.
+{\color{pnasbluetext}2}), total leaf area ($\frac{\ud A_{\rm l}}{\ud t}$; Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}4}), height ($\frac{\ud H}{\ud t}$; Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}5}), stem basal area ($\frac{\ud  A_{\rm st}}{\ud t}$; Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}6}), and stem diameter ($\frac{\ud D}{\ud t}$; Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}7}) therefore share several terms.
 Many of these terms also vary intrinsically with size, as shown in the insets in Fig.
@@
 
-The growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depends on the product of biomass growth $\frac{{\rm d}B}{{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
+The growth rates of all size metrics (Fig.
+\ref{fig:conceptual}b, Eqs.
+{\color{pnasbluetext}2}-{\color{pnasbluetext}7}) depend on the product of biomass growth $\frac{{\rm d}B}{{\rm d}t}$ (from Eq.
+\ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1.
 The remaining $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$ fraction of mass produced is allocated to reproduction.
@@
 \caption{Conceptual framework showing how shade-tolerance (as measured by \textsc{wplcp}) is affected by plant size and traits, adapted from \citep{Givnish-1988}.
-From eqn \ref{eq:wplcp}, \textsc{wplcp} is the canopy openness $E$ where the photosynthetic income per unit leaf area (dashed line) intersects the total respiration and turnover costs per unit leaf area (solid black lines).
+From Eq.
+\ref{eq:wplcp}, \textsc{wplcp} is the canopy openness $E$ where the photosynthetic income per unit leaf area (dashed line) intersects the total respiration and turnover costs per unit leaf area (solid black lines).
 Colors indicate the contributions from different tissues to total costs.
@@
 
-The growth rate in the total standing mass of the plant (eqn \ref{eq:dMtdt}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
+The growth rate in the total standing mass of the plant (Fig.
+\ref{fig:conceptual}b, Eqs.
+{\color{pnasbluetext}2}) is then the sum of heartwood formation (=sapwood turnover) and any increment in live mass.
 
-The remaining growth rates (eqns \ref{eq:daldt}-\ref{eq:dDdt}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood, and roots (eqn \ref{eq:daldmt}).
-The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of construction cost per unit leaf area for different tissues, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$ (eqn \ref{eq:daldmt}).
+The remaining growth rates (Fig.
+\ref{fig:conceptual}b, Eqs.
+{\color{pnasbluetext}4}-{\color{pnasbluetext}7}) all depend on another variable, $\frac{{\rm d}A_{\rm l}} {{\rm d}M_{\rm a}}$, that accounts for the marginal cost of deploying an additional unit of leaf area, including construction of the leaf itself and supporting  bark, sapwood, and roots (Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}3}).
+The inverse of this term, $\frac{{\rm d}M_{\rm a} }{{\rm d}A_{\rm l}}$, is the whole plant construction cost per unit leaf area, which can be further decomposed as a sum of construction cost per unit leaf area for different tissues, with one of these being the trait $\phi = \frac{{\rm d}M_{\rm l} }{{\rm d}A_{\rm l}}$ (Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}3}).
 
-The rate of height growth (eqn \ref{eq:dhdt}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
+The rate of height growth (Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}5}) depends on an additional term, $\frac{{\rm d}H }{{\rm d}A_{\rm l}}$: the growth in plant height per unit growth in leaf area.
 This variable accounts for the architectural strategy of the plant (\citep{Falster-2011}).
@@
 
-The rate of stem-basal-area growth (eqn \ref{eq:dast}) can be expressed as the sum of increments in sapwood, bark, and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
-These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see eqn \ref{eq:dast}).
+The rate of stem-basal-area growth (Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}6}) can be expressed as the sum of increments in sapwood, bark, and heartwood areas ($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively): $\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
+These in turn are related to ratios of sapwood and bark area per leaf area, and sapwood turnover (see Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}6}).
 
-Finally, the rate of stem-diameter growth (eqn \ref{eq:dDdt}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
-We make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (eqns \ref{eq:ha}-\ref{eq:htm}).
+Finally, the rate of stem-diameter growth (Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}7}) is given by a geometric relationship between stem diameter ($D$) and stem area ($A_{\rm st}$).
+We make no assumptions about the relationship of stem diameter to height or leaf area: these arise as emergent properties, via integration of stem turnover (Table \ref{tab:allometry}, Eqs.
+{\color{pnasbluetext}9}-{\color{pnasbluetext}10}).
 
@@
 
-Eqn \ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'' \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
+Eq.
+\ref{eq:dbdt} can also be rearranged to obtain a measure of shade tolerance: the ``whole-plant-light-compensation point'' \citep[{\wplcp}][]{Givnish-1988, Baltzer-2007, Lusk-2013} .
 In general, photosynthesis rate per leaf area $\bar{p}$ increases with canopy openness, $E$.
@@
 $\frac{{\rm d}B}{{\rm d}t}(E^*)=0$.
-From eqn \ref{eq:dbdt}, this occurs when
+From Eq.
+\ref{eq:dbdt}, this occurs when
 \begin{equation}\label{eq:wplcp}
@@
 
-Since eqns \ref{eq:dMtdt}-\ref{eq:wplcp} are derived using standard rules of addition, multiplication, and differentiation, they may apply to a variety of growth model where biomass allocation is important.
+Since Eqs.
+in Fig.
+\ref{fig:conceptual}b and Eq.
+\ref{eq:wplcp} are derived using standard rules of addition, multiplication, and differentiation, they may apply to a variety of growth model where biomass allocation is important.
 However, to make explicit predictions requires that we quantify all the terms in Fig.
@@
 In particular, we assume that as a plant grows:
-1) Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha});
-2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as});
-3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab});
-4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}); and
+1) Its height scales to the 0.5 power of its leaf area (Table \ref{tab:allometry}, Eq. {\color{pnasbluetext}1});
+2) The cross-sectional area of sapwood in the stem is proportional to its leaf area (Table \ref{tab:allometry}, Eq. {\color{pnasbluetext}2});
+3) The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (Table \ref{tab:allometry}, Eq. {\color{pnasbluetext}3});
+4) The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (Table \ref{tab:allometry}, Eq. {\color{pnasbluetext}4}); and
 5) The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
@@
 Combined, the functional-balance assumptions from Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b).
-These eqns aso predict that the amount of live stem tissue supporting each unit of leaf area should increase linearly with height,
+These equations also predict that the amount of live stem tissue supporting each unit of leaf area should increase linearly with height,
 \begin{equation}\label{eq:msml}
@@
 \end{equation}
-Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives all the necessary terms needed to implement the growth model described in Fig.
-\ref{fig:conceptual}.
+Substituting from Table \ref{tab:allometry} Eqs.
+{\color{pnasbluetext}1}-{\color{pnasbluetext}10} into Fig.
+\ref{fig:conceptual}b Eqs.
+{\color{pnasbluetext}5}-{\color{pnasbluetext}7} then gives all the necessary terms needed to implement the growth model described in Fig.
+\ref{fig:conceptual}b.
 
@@
   \toprule
-  & Variable & Function & Marginal cost & Growth rate & Eqn\\
+  & Variable & Function & Marginal cost & Growth rate & Eq.\\
   \midrule
@@
     $\frac{{\rm d}H}{{\rm d}A_{\rm l}} = 0.5 \alpha_{\rm l1} A_{\rm l}^{-0.5}$ &
-    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
+    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  [1]\\
   & Sapwood area &
@@
     $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
-    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:as}\\
+    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & [2]\\
   & Bark area &
@@
     $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
-    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ab}\\
+    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & [3]\\
   & Root mass &
@@
     $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_{\rm r1}$  &
-    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\
+    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & [4]\\
   & Heartwood area & & &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & \eqncounter{eq:dah}\\
+    $\frac{{\rm d}A_{\rm h}}{{\rm d} t}  = k_{s} A_{\rm h} $ & [5]\\
   \multicolumn{6}{l}{\textbf{b) Derived quantities}} \\
@@
     $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
-    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
+    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & [6]\\
   & Sapwood mass &
@@
     $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( H + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
+    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & [7]\\
   & Bark mass &
@@
     $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( H + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
-    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
+    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & [8]\\
   & Heartwood area &
@@
      &
-    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & \eqncounter{eq:hta} \\
+    $\frac{{\rm d}A_{\rm h}}{{\rm d}t} = k_{\rm s} \, A_{\rm s}$ & [9] \\
   & Heartwood mass &
@@
      &
-    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & \eqncounter{eq:htm} \\
+    $\frac{{\rm d}M_{\rm h}}{{\rm d}t} = k_{\rm s} \, M_{\rm s}$ & [10]\\
   \bottomrule
@@
 \addtabletext{\\The first column of \textbf{a)} provides cores assumptions between various size metrics and leaf area.
-Eqns in the middle and right columns of \textbf{a)} and in \textbf{b)} can be derived from these assumptions.
-The column ``Eqn'' indicates equation numbers referred to in the main text.
-
+Eqs.
+in the middle and right columns of \textbf{a)} and in \textbf{b)} can be derived from these assumptions.
 }
@@
 
-Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}.
+Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested in reproduction and in growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in Fig.
+\ref{fig:conceptual}b Eq.
+{\color{pnasbluetext}2}-{\color{pnasbluetext}7}.
 We assume the amount allocated to vegetative growth starts high and then decreases once $H > H_{{\rm mat}}$  (Fig.
@@
 
-We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}.
+We compared the assumptions outlined in Table \ref{tab:allometry}a to data sourced from the Biomass and Allometry Database \citep{Falster-2015b}.
 Fig.
-\ref{fig:assumptions} shows that the three functional-balance assumptions in Table \ref{tab:allometry}a and the relationship in eqn \ref{eq:msml} are all well-supported by the available data.
-As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions (Table \ref{tab:allometry}a and eqn \ref{eq:msml}).
+\ref{fig:assumptions} shows that the three functional-balance assumptions in Table \ref{tab:allometry}a and the relationship in Eq.
+\ref{eq:msml} are all well-supported by the available data.
+As expected, species differed in elevation, but less so in the slope of the fitted lines; with slopes aligning with those predicted by the functional-balance assumptions (Table \ref{tab:allometry}a and Eq.
+\ref{eq:msml}).
 
@@
 Height growth also shows a hump-shaped pattern with size, first increasing then decreasing.
-This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with increasing size, including a strong decline in the fraction of plant that is leaf (Fig.
-\ref{fig:conceptual}), increasing reproductive allocation (Fig.
-\ref{fig:conceptual}), and declining biomass growth.
+This pattern results from systematic changes in the four components of Fig.
+\ref{fig:conceptual}b Eq.
+{\color{pnasbluetext}5} with increasing size, including a strong decline in the fraction of plant that is leaf (Fig.
+\ref{fig:conceptual}a), increasing reproductive allocation, and declining biomass growth.
 In contrast, basal-area growth continues to increase with size, due to an increasing influence of stem turnover.
-Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (eqn \ref{eq:dDdt}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
+Diameter growth shows a weakly hump shaped curve, tapering off slightly at larger sizes, in part because of the allometric conversion from basal area to diameter (Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}7}), and in part because of increased reproductive allocation in older trees (Fig \ref{fig:conceptual}).
 All growth measures decrease sharply with size when expressed as relative growth rates.
@@
 \caption{\textbf{Effect of three developmental traits on shade tolerance.}
-Panels show effect of traits on level of canopy openness that causes biomass growth (eqn \ref{eq:dbdt}) to be zero.
+Panels show effect of traits on level of canopy openness that causes biomass growth (Eq.
+\ref{eq:dbdt}) to be zero.
 Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m).
@@
 Both the cost and benefits of {\nitrogen} appear within the term $\frac{ {\rm d}B}{{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in biomass growth per {\nitrogen}.
-From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
+From Eq. \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
 u} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial
@@
 When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate.
-When plants are large, the effect on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger.
+When plants are large, the effect on leaf deployment is diminished because the cost of building other supportive tissues (other terms in Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}3}) is larger.
 Thus, at larger sizes low {\lma} no longer increases growth rate (Fig.
@@
 \ref{fig:growth_light_height} focus on height growth rate.
-Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in the SI Appendix, Figs.
+Corresponding results for absolute growth rates in stem diameter (Eq.
+Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}7}), stem basal area (Eq.
+Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}6}), and above-ground mass (Eq.
+Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}2}) are provided in the SI Appendix, Figs.
 \ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}.
@@
 \subsection*{Shade tolerance}
-Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
+Combining Eq. \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
 \begin{equation}\label{eq:wplcp_2}
@@
 where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
-Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}.
+Eq.
+\ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}.
 With some further manipulations, it is possible to show that {\wplcp} will decrease with {\lma} if $\beta_{{\rm kl2}} >1$.
@@
   \midrule
-  \multicolumn{8}{l} {\textbf{a) Effect on elements of eqns \ref{eq:dbdt} -- \ref{eq:dDdt}}}  \\
+  \multicolumn{8}{l} {\textbf{a) Effect on elements of Eq. \ref{eq:dbdt} and Fig. \ref{fig:conceptual}b Eqs. {\color{pnasbluetext}1} {\color{pnasbluetext}7} }}  \\
   \multicolumn{2}{l} {Biomass growth, $\ud B / \ud t$} & & & & & & \\
@@
 The model presented here extends a widely-used model for seedlings, linking mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to larger plants and other traits.
-Importantly, the seedling model (eqn \ref{eq:RGR}) can be derived as a special case of the new, extended model.
-Unlike eqn \ref{eq:RGR}, the extended model also predicts a relationship between {\lma} and growth rate that changes with size.
+Importantly, the seedling model (Eq.
+\ref{eq:RGR}) can be derived as a special case of the new, extended model.
+Unlike Eq.
+\ref{eq:RGR}, the extended model also predicts a relationship between {\lma} and growth rate that changes with size.
 In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even weakly positive in larger plants (Fig.
@@
 \ref{fig:conceptual}a).
-Decreasing {\lma} then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}.
+Decreasing {\lma} then has an overwhelmingly positive effect on growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in Fig.
+\ref{fig:conceptual}b, Eq.
+{\color{pnasbluetext}3}.
 As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheap leaf construction.
@@
 While we have focussed on understanding the effects of five traits on elements of demography, the framework presented can be extended to generate hypotheses about other traits and trade-offs.
-The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be connected with variables in eqns \ref{eq:dbdt}--\ref{eq:dDdt}.
+The main criteria for including new traits is that a clear trade-off has been established, with benefits and/or costs that ultimately translate into biomass and can therefore be connected with variables in Eq.
+\ref{eq:dbdt} and Fig.
+\ref{fig:conceptual}b Eqs.
+{\color{pnasbluetext}1}-{\color{pnasbluetext}7}.
 While the list of plant traits that have been measured is extensive, clear trade-offs have been established for only a few of these.
@@
 For example, many studies consider traits related to plant hydraulics, such as vessel size and increased sapwood area per leaf area \citep{Zanne-2010}.
-By altering conductance of water to the leaf, both traits will influence the rate of photosynthesis per leaf area ($\bar{p}$ in eqn \ref{eq:dbdt}), while vessel size might also affect rates of stem turnover ($k_{\rm s}$ in eqn \ref{eq:dbdt}).
+By altering conductance of water to the leaf, both traits will influence the rate of photosynthesis per leaf area ($\bar{p}$ in Eq.
+\ref{eq:dbdt}), while vessel size might also affect rates of stem turnover ($k_{\rm s}$ in Eq.
+\ref{eq:dbdt}).
 The costs of increased sapwood area per leaf area is increased allocation to stem, a factor which is already included in via the parameter $\theta$ (Table \ref{tab:allometry}).
@@
 \subsection*{Assumptions}
-The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
+The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species.
 We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships.
